### 2014-11-16 appendices/appendix-code.tex
initial commit

\section{Coding}\label{code}
In this Appendix we reproduce all the quantities  that should be coded.

Eqs.~\eqref{calvimchiewn}, \eqref{calvimchie2wn}, \eqref{calvimchiwn}
 and \eqref{calvimchi2wn}
\input{appendices/chis}
\noindent$\bullet$ Coding:
$\calv^{\gs,\rma,\ell}_{nm}\to$ \verb=calVsig=,
$r^\rma_{nm}\to$ \verb=posMatElem=,
$\left(\calv^{\gs,\rma,\ell}_{nm}\right)_{;k^\rmb}\to$ \verb=gdcalVsig=,
\\ $(r^\rma_{nm})_{;k^\rmb}\to$ \verb=derMatElem= 
$\gD^\rma_{nm}\to$ \verb=Delta= and $\go^\gs_n\to$ \verb=band(n)=

$\bullet$ proof:\\
To evaluate above expressions we need the following ($m_e=1$):
\begin{align}\label{}
\bfv^\lda_{nm}(\bfk) 
=(1/m_e)\bfp_{nm}(\bfk)+\bfv^\nl_{nm}(\bfk)
=\bfp_{nm}(\bfk)+\bfv^\nl_{nm}(\bfk)
,
\end{align}
that
 includes the local and nonlocal parts of the pseudopotential. They
 correspond to the following files:\\
$\bullet$ $\bfp_{nm}(\bfk)\to$ \verb=me_pmn_*=\\
$\bullet$ $\bfv^\nl_{nm}(\bfk)\to$ \verb=me_vnlnm_*=\\
where the \verb=nm= or \verb=mn= order in the files is irrelevant, and
ought to be fixed just for the {\it biuty} of it.
Option \verb=-n= in \verb=all_responses.sh= does
\begin{enumerate}
\item 
 \verb=> cp me_pmn_* me_pmn_*.o= 
\item adds \verb=me_pmn_*= and \verb=me_vnlnm_*= into
  \verb=me_pmn_*= 
\item calculates the response
\item \verb=> mv me_pmn_*.o me_pmn_*=
\end{enumerate}
so   
$\bfv^\lda_{nm}(\bfk)$, stored in \verb=vldaMatElem=
is available for the calculation of the response, and with it we calculate
(Eqs.~\eqref{chon.9} and \eqref{chon.10}),
\begin{align}\label{c-chon.98}
\bfv^\gs_{nm}(\bfk)
&=
\left(1+\frac{\gS}{\go_c(\bfk)-\go_v(\bfk)}\right)\bfv^\lda_{nm}(\bfk)
\quad\quad n\notin D_m
\nonumber\\
\bfv^\gs_{nn}(\bfk)
&=
\bfv^\lda_{nn}(\bfk)
\nonumber\\
\bfr_{nm}(\bfk)&=\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
=\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
\quad\quad n\notin D_m
.
\end{align}   
If option \verb=-n= is not chosen, then the contribution of $\bfv^\nl_{nm}(\bfk)$
is neglected in the calculation of any response. Obviously, in this
case the code only uses \verb=me_pmn_*= without adding \verb=me_vnlnm_*=

We need Eq.~\eqref{a.1} and \eqref{a.2}
\begin{align}\label{c-a.1}
\calv^{\gs,\rma,\ell}_{nm}
&=
\calv^{\lda,\rma,\ell}_{nm}
+
\calv^{\cals,\rma,\ell}_{nm}
\nonumber\\
\left(
\calv^{\gs,\rma,\ell}_{nm}
\right)_{;k^\rmb}
&=
\left(
\calv^{\lda,\rma,\ell}_{nm}
\right)_{;k^\rmb}
+
\left(
\calv^{\cals,\rma,\ell}_{nm}
\right)_{;k^\rmb}
.
\end{align}
 The first LDA term is
\begin{align}\label{c-a.2}
\calv^{\lda,\rma,\ell}_{nm}
&=
\frac{1}{2}\sum_q\left(
v^{\lda,\rma}_{nq}\calc^\ell_{qm}+\calc^\ell_{nq} v^{\lda,\rma}_{qm}
\right)
.
\end{align} 
If option \verb=-n= is not chosen in \verb=all_responses.sh=
Eq.~\eqref{c-a.2}
 is
not calculated and\\
$\bullet$ $\calv^{\lda,\rma,\ell}_{nm}\to$ \verb=me_cpmn_*=\\  
If option \verb=-n= is chosen Eq.~\eqref{c-a.2}
 must be calculated as given in
\verb=set_input_ascii.f90=. We mention that
$\calv^{\lda,\rma,\ell}_{nm}$ can be computed directly,\cite{nicolaspc}
avoiding the sum over the full set of bands $q$, however we chose to
compute Eq.~\eqref{c-a.2}, which is done in
\verb=functions.f90= under the name \verb=calVlda=.
Then, we need 
Eq.~\eqref{eni.4}
\begin{align}\label{c-eni.4}
\calc^\ell_{nm}(\bfk)&=
\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
\gd_{\bfG_\parallel \bfG'_\parallel}
f_\ell(G_\perp-G'_\perp)
\nonumber\\
\calc^\ell_{mn}(\bfk)&=
\big(\calc^\ell_{nm}(\bfk)\big)^*
,
\end{align} 
which is coded in \verb=sub_pmn_ascii.f90= within the same subroutine of $\calbv^\ell_{nm}$
calculated with Eq.~\eqref{eni.2}. However, Sean out of the blue, call
it \verb=me_cfmn_*= in \verb=run_tiniba.sh=,
 and Darwin won (what else? ID??), 
thus I call it \verb=cfMatElem= in \verb=SRC_1setinput=. ID would call
it  \verb=ccMatElem=
but long live CD!

 The second LDA term is
\begin{align}\label{c-a.2n}
\left(\calv^{\lda,\rma,\ell}_{nm}\right)_{;k^\rmb}
&=
\frac{1}{2}\sum_q\left(
(v^{\lda,\rma}_{nq})_{;k^\rmb}\calc^\ell_{qm}
+  
v^{\lda,\rma}_{nq}(\calc^\ell_{qm})_{;k^\rmb}
+
(\calc^\ell_{nq})_{;k^\rmb} v^{\lda,\rma}_{qm}
+
\calc^\ell_{nq} (v^{\lda,\rma}_{qm})_{;k^\rmb}
\right)
,
\end{align}  
where\\
$\bullet$ for $n\ne m$\\
Eq.~\eqref{a.3}
\begin{align}\label{c-a.3}
(v^{\lda,\rma}_{nm})_{;k^\rmb}
&=  
im_e\left(\gD^b_{nm}r^\rma_{nm}
+ 
\go^\lda_{nm}(r^\rma_{nm})_{;k^\rmb}
\right)
\nonumber\\
(v^{\lda,\rma}_{mn})_{;k^\rmb}
&=
\left((v^{\lda,\rma}_{nm})_{;k^\rmb}\right)^*
\quad\mathrm{for}\quad n\ne m
,
\end{align} 
with
Eq.~\eqref{eli.13}
\begin{align}\label{c-eli.13}
\gD_{nm}^{\rma}
=
v_{nn}^{\lda,\rma}-v_{mm}^{\lda,\rma}
,
\end{align}
and \eqref{na_rgendevn}
\begin{align}\label{c-na_rgendevn}
(r^{\rmb}_{nm})_{;k^{\rma}}
&=
-i\calt^{\rma\rmb}_{nm}
+
\frac{
r^{\rma}_{nm}
\Delta^{\rmb}_{mn}
+r^{\rmb}_{nm}
\Delta^{\rma}_{mn}
}
{\go^\lda_{nm}}
+
\frac{i}{\go^\lda_{nm}}
\sum_{\ell}
\bigg(
\go^\lda_{\ell m}
r^{\rma}_{n\ell}
r^{\rmb}_{\ell m}
-
\go^\lda_{n\ell}
r^{\rmb}_{n\ell}
r^{\rma}_{\ell m}
\bigg)
\nonumber\\
&\approx
\frac{
r^{\rma}_{nm}
\Delta^{\rmb}_{mn}
+r^{\rmb}_{nm}
\Delta^{\rma}_{mn}
}
{\go^\lda_{nm}}
+
\frac{i}{\go^\lda_{nm}}
\sum_{\ell}
\bigg(
\go^\lda_{\ell m}
r^{\rma}_{n\ell}
r^{\rmb}_{\ell m}
-
\go^\lda_{n\ell}
r^{\rmb}_{n\ell}
r^{\rma}_{\ell m}
\bigg)
\nonumber\\
(r^{\rmb}_{mn})_{;k^{\rma}}
&=
\left((r^{\rmb}_{nm})_{;k^{\rma}}\right)^*
,
\end{align}
where $\calt^{\rma\rmb}_{nm}\approx 0$.\\
$\bullet$ for $n=m$\\
Since 
$\calt^{\rma\rmb}_{nn}\approx (\hbar/m_e)\gd_{\rma\rmb}$,
Eq.~\eqref{a.3c} gives
\begin{align}\label{c-a.3c}
(v^{\lda,\rma}_{nn})_{;k^\rmb}
&=
-i\calt^{\rma\rmb}_{nn}
-
\sum_{\ell\ne n}
\go^\lda_{\ell n}
\bigg( 
r^{\rma}_{n\ell} 
r^\rmb_{\ell n}
+ 
r^\rmb_{n\ell} 
r^{\rma}_{\ell n}
\bigg)
\nonumber\\
&\approx
\frac{\hbar}{m_e}\gd_{\rma\rmb}
-
\sum_{\ell\ne n}
\go^\lda_{\ell n}
\bigg( 
r^{\rma}_{n\ell} 
r^\rmb_{\ell n}
+ 
r^\rmb_{n\ell} 
r^{\rma}_{\ell n}
\bigg)
.
\end{align} 
For Eq.~\eqref{c-a.2n} we need
\eqref{a.7}
\begin{align}\label{c-a.7}
 (\calc^\ell_{nm})_{;k^\rma}
&= 
i\sum_{q\ne nm}
\left(
r_{nq}^\rma
\calc^\ell_{qm}
-
\calc^\ell_{nq}
r_{qm}^\rma
\right)
+ir_{nm}^\rma(\calc^\ell_{mm}-\calc^\ell_{nn})
\nonumber\\
 (\calc^\ell_{mn})_{;\bfk}
&=
\big( (\calc^\ell_{nm})_{;\bfk}\big)^*
.
\end{align}

For the scissor related term we have:
Eq.~\eqref{a.3b} , \eqref{choni.1} and \eqref{chon.2}
\begin{align}\label{c-a.3b}
\calv^{\cals,\rma,\ell}_{nm}
&=
\frac{1}{2}\sum_q\left(    
v^{\cals,\rma}_{nq}\calc^\ell_{qm}+\calc^\ell_{nq} v^{\cals,\rma}_{qm}
\right)
\nonumber\\
\left(\calv^{\cals,\rma,\ell}_{nm}\right)_{;k^\rmb}
&=
\frac{1}{2}\sum_q\left(
(v^{\cals,\rma}_{nq})_{;k^\rmb}\calc^\ell_{qm}
+     
v^{\cals,\rma}_{nq}(\calc^\ell_{qm})_{;k^\rmb}
+
(\calc^\ell_{nq})_{;k^\rmb} v^{\cals,\rma}_{qm}
+
\calc^\ell_{nq} (v^{\cals,\rma}_{qm})_{;k^\rmb}
\right)
,
\end{align}     
with Eqs.~\eqref{chon.2} and \eqref{choni.1}
\begin{align}\label{c-chon.2} 
v^{\cals,\rma}_{nm}=i\gS f_{mn}r^\rma_{nm}
,
\end{align}
\begin{align}\label{c-choni.1}
(v^{\cals,\rma}_{nm})_{;k^\rmb}= 
i\gS f_{mn}(r^\rma_{nm})_{;k^\rmb}
,
\end{align} 
where $\hbar\gS$ is the scissors correction.
Notice that
$v^{\cals,\rma}_{nn}=0$ and 
$(v^{\cals,\rma}_{nn})_{;k^\rmb}=0$.
Substuiting Eq.~\eqref{c-chon.2} into \eqref{c-a.3b}, we obtain
\begin{align}\label{c-a.3bb}
\calv^{\cals,\rma,\ell}_{nm}
&=
\frac{i\gS}{2}\sum_q\left(  
f_{qn}r^\rma_{nq}\calc^\ell_{qm}+f_{mq}\calc^\ell_{nq} r^\rma_{qm}
\right)
,
\end{align}   
$\bullet$ Coding: \verb=functions.f90= array \verb=calVscissors= 
where $f_n$ is coded in \verb=set_input_ascii.f90=\\
Notice that $q=n$ and $q=m$ give zero contribution from the $f_{nm}$
factors, but we set in the code $r^\rma_{nn}=0$ so the program would
not complain that such values of the array \verb=posMatElem= do not
exist, since actually, the diagonal elements do not exist.
Explicitly (although, we don't code them), 
\begin{align}\label{vs.vc}
      \mathcal{V}^{S,a,\ell}_{vc}
&= 
      -\frac{i\Sigma}{2}
      \left[\sum_{v^{\prime}}r^{a}_{vv^{\prime}}C^{\ell}_{v^{\prime}c} 
          +
          \sum_{c^{\prime}}C^{\ell}_{vc^{\prime}}r^{a}_{c^{\prime}c}\right],
\nonumber\\
      \mathcal{V}^{S,a,\ell}_{cv}
&= 
      \frac{i\Sigma}{2}
      \left[\sum_{v^{\prime}}r^{a}_{cv^{\prime}}C^{\ell}_{v^{\prime}v} 
          + \sum_{c^{\prime}}C^{\ell}_{cc^{\prime}}r^{a}_{c^{\prime}v}\right],
\nonumber\\
      \mathcal{V}^{S,a,\ell}_{cv}
&= 
      (\mathcal{V}^{S,a,\ell}_{vc})^*
    \end{align}
and    
  \begin{align}\label{vs.cc}
    \mathcal{V}^{S,a,\ell}_{cc} 
    &= -\Sigma\sum_{v}
    \text{Im}\left[r^{a}_{cv}C^{\ell}_{vc}\right],
  \end{align}
 \begin{align}\label{vs.vv}
    \mathcal{V}^{S,a,\ell}_{vv} 
    &= \Sigma\sum_{c}
    \text{Im}\left[r^{a}_{vc}C^{\ell}_{cv}\right],
  \end{align}
where the last two are real functions as they must, since they are
velocities.

Substuiting Eqs.~\eqref{c-chon.2} and \eqref{c-choni.1} into \eqref{c-a.3b}, we obtain
\begin{align}\label{c-a.3bnn}
\left(\calv^{\cals,\rma,\ell}_{nm}\right)_{;k^\rmb}
&=
\frac{i\gS}{2}\sum_q\left(
f_{qn}\Big[
(r^\rma_{nq})_{;k^\rmb}\calc^\ell_{qm}
+    
r^\rma_{nq}(\calc^\ell_{qm})_{;k^\rmb}
\Big]
+
f_{mq}\Big[
(\calc^\ell_{nq})_{;k^\rmb} r^\rma_{qm}
+
\calc^\ell_{nq} (r^\rma_{qm})_{;k^\rmb}
\Big]
\right)
\nonumber\\
\left(\calv^{\cals,\rma,\ell}_{mn}\right)_{;k^\rmb}
&=
\left(\left(\calv^{\cals,\rma,\ell}_{nm}\right)_{;k^\rmb}\right)^*
,
\end{align}    
$\bullet$ Coding:\\
$(r^\rma_{nm})_{;k^\rmb}\to$ \verb=derMatElem=
$\calc^\ell_{nm}\to$ \verb=cfMatElem=
$r^\rma_{nm}\to$ \verb=posMatElem=
$(\calc^\ell_{nm})_{;k^\rmb}\to$ \verb=gdf=, and\\
$\left(\calv^{\cals,\rma,\ell}_{nm}\right)_{;k^\rmb}\to$ 
\verb=gdcalVS=\\
Also
\begin{align}\label{dgvs.cv}
\left(\calv^{\cals,\rma\,\ell}_{cv}\right)_{;k^\rmb}
&=
\frac{i\gS}{2}
\left(
\sum_{v'}\left(
 (r^\rma_{cv'})_{;k^\rmb}\calc^\ell_{v'v}
+  
r^\rma_{cv'}(\calc^\ell_{v'v})_{;k^\rmb}
\right)
+
\sum_{c'}\left(
(\calc^\ell_{cc'})_{;k^\rmb} r^\rma_{c'v}
+
\calc^\ell_{cc'} (r^\rma_{c'v})_{;k^\rmb}
\right)
\right)
\nonumber\\
\left(\calv^{\cals,\rma,\ell}_{vc}\right)_{;k^\rmb}
&=
\left(\left(\calv^{\cals,\rma,\ell}_{cv}\right)_{;k^\rmb}\right)^*
,
\end{align}  
\begin{align}\label{dgvs.cc}
\left(\calv^{\cals,\rma,\ell}_{cc}\right)_{;k^\rmb}
&=
-\gS
\sum_{v}
 \mathrm{Im}\left[(r^\rma_{cv})_{;k^\rmb}\calc^\ell_{vc}
+
r^\rma_{cv}
(\calc^\ell_{vc}) _{;k^\rmb}\right]
,
\end{align}  
and
\begin{align}\label{dgvs.vv}
\left(\calv^{\cals,\rma,\ell}_{vv}\right)_{;k^\rmb}
&=
\gS
\sum_{c}
 \mathrm{Im}\left[(r^\rma_{vc})_{;k^\rmb}\calc^\ell_{cv}
+
r^\rma_{vc}(\calc^\ell_{cv}) _{;k^\rmb}\right]
.
\end{align}

\subsection{Coding for $\calv^{\gs,\rma,\ell}_{nm}(\bfk)$} 
Recall that 
$\calv^{\lda,\rma,\ell}_{mn}=(\calv^{\lda,\rma,\ell}_{nm})^*$
and 
$\calv^{\cals,\rma,\ell}_{mn}=(\calv^{\cals,\rma,\ell}_{nm})^*$ 
\begin{itemize}
%%%
\item If \verb=-n= option is chosen in \verb=all_responses.sh=
\begin{itemize}
\item $\calv^{\lda,\rma,\ell}_{nm}$, comes from
  Eq.~\eqref{c-a.2}, coded in \verb=functions.f90= as \verb=calVlda=
\end{itemize}
%%%
\item If \verb=-n= option is NOT chosen in \verb=all_responses.sh=
\begin{itemize}
\item $\calv^{\lda,\rma,\ell}_{nm}$ 
is used from \verb=me_cpmn_*=
  which is Eq.~\eqref{eni.2} and is coded in \verb=sub_pmn_ascii.f90=
\end{itemize}
%%%
\end{itemize}
For either case
\begin{itemize}
\item $\calv^{\cals,\rma,\ell}_{nm}$ 
is obtained from
  Eqs.~\eqref{vs.vc}, \eqref{vs.cc} or \eqref{vs.vv}, depending on
  $nm$. This is coded in \verb=functions.f90= and used in 
\verb=set_input_ascii.f90=  
\end{itemize}
Thus,\\
$\bullet$ 
$\calv^{\gs,\rma,\ell}_{nm}(\bfk)=\calv^{\lda,\rma,\ell}_{nm}(\bfk)+\calv^{\cals,\rma,\ell}_{nm}(\bfk)$\\ 
is stored in \verb=calMomMatElem=  
array, constructed in 
\verb=set_input_ascii.f90=, and used in \verb=SRC_2latm= for
integrating the response function. A brave young soul, should change   
\verb=calMomMatElem= to \verb=calVelMatElem= in order to have a more
appropriate name. But as good old DNA, we construct upon available
ATGC; using the old structure, adding functionality and  keeping all
the usles non-codifying crap, thus making Darwin
 proud of us!

\subsection{$\gD^{\gs,\rma,\ell}_{nm}(\bfk)$}
$\gD^{\gs,\rma,\ell}_{nm}(\bfk)$  is given by
\begin{align}\label{caldelta}
\gD^{\gs,\rma,\ell}_{nm}(\bfk)&=\calv^{\gs,\rma,\ell}_{nn}(\bfk)-\calv^{\gs,\rma,\ell}_{mm}(\bfk)
\end{align}
\begin{align}\label{delta}
\gD^{\gs,\rma}_{nm}(\bfk) &=v^{\gs,\rma,\ell}_{nn}(\bfk)-v^{\gs,\rma,\ell}_{mm}(\bfk)
\nonumber\\
&=v^{\lda,\rma,\ell}_{nn}(\bfk)-v^{\lda,\rma,\ell}_{mm}(\bfk)
,
\end{align}
since $\bfv^\cals_{nn}=0$.\\
$\bullet$ Coding: $\gD^{\gs,\rma,\ell}_{nm}(\bfk)\to$ \verb=calDelta= 
and
$\gD^{\gs,\rma}_{nm}(\bfk)\to$ \verb=Delta= both in \verb=set_input_ascii.f90=

\subsection{Coding for $(\calv^{\lda,\rma,\ell}_{nm}(\bfk))_{;k^\rmb}$}
\begin{itemize}
\item $\gD^\rma_{nm}$ available in array \verb=Delta=, 
calculated in \verb=set_input_ascii.f90=,
 and contains the
  contribution from $\bfv^\nl_{nm}(\bfk)$ if the \verb=-n= option is
  chosen in \verb=all_responses.sh= 
\item $(r^\rma_{nm}(\bfk))_{;k^\rmb}$
 available in array
  \verb=derMatElem=,
calculated in \verb=set_input_ascii.f90= and \verb=functions.f90=,
 and contains the
  contribution from $\bfv^\nl_{nm}(\bfk)$ if the \verb=-n= option is
  chosen in \verb=all_responses.sh= 
\item With above two we compute $(v^{\lda,\rma}_{nm}(\bfk))_{;k^\rmb}$ 
in \verb=set_input_ascii.f90=  and store it in \verb=gdVlda= for
diagonal and off diagonal terms.
\item $\big(\calc^\ell_{nm}(\bfk)\big)_{;k^\rma}$ is coded in 
in \verb=set_input_ascii.f90=  and store it in \verb=gdf= for
diagonal and off diagonal terms. Darwin at work!
\item $(v^{\lda,\rma}_{nq})_{;k^\rmb}\to$ \verb=gdVlda=,
  $\calc^\ell_{qm}\to$ \verb=cfMatElem=,
$v^{\lda,\rma}_{nq}\to$ \verb=vldaMatElem=, 
$(\calc^\ell_{qm})_{;k^\rmb}\to$
\verb=gdf=\\
 $v^{\lda,\rma}_{nq}\to$ \verb=vldaMatElem=,

 \begin{align}\label{c-a.2nn}
\left(\calv^{\lda,\rma,\ell}_{nm}\right)_{;k^\rmb}
&=
\frac{1}{2}\sum_q\left(
(v^{\lda,\rma}_{nq})_{;k^\rmb}\calc^\ell_{qm}
+ 
v^{\lda,\rma}_{nq}(\calc^\ell_{qm})_{;k^\rmb}
+
(\calc^\ell_{nq})_{;k^\rmb} v^{\lda,\rma}_{qm}
+
\calc^\ell_{nq} (v^{\lda,\rma}_{qm})_{;k^\rmb}
\right)
\nonumber\\
\left(\calv^{\lda,\rma,\ell}_{mn}\right)_{;k^\rmb}&=\left(\left(\calv^{\lda,\rma,\ell}_{nm}\right)_{;k^\rmb}\right)^*
,
\end{align} 
$\left(\calv^{\lda,\rma,\ell}_{nm}\right)_{;k^\rmb}\to$
\verb=gdcalVlda= and coded in \verb=set_input_ascii.f90=
\end{itemize}

\subsection{Summary}
\begin{itemize}
\item
$\calv^{\gs,\rma,\ell}_{nm}(\bfk)=\calv^{\lda,\rma,\ell}_{nm}(\bfk)+\calv^{\cals,\rma,\ell}_{nm}(\bfk)\to$
\verb=calMomMatElem=

\item $\left(\calv^{\lda,\rma,\ell}_{nm}\right)_{;k^\rmb}\to$
\verb=gdcalVlda=

\item $\left(\calv^{\cals,\rma,\ell}_{nm}\right)_{;k^\rmb}\to$ 
\verb=gdcalVS=

\item $\left(\calv^{\gs,\rma,\ell}_{nm}\right)_{;k^\rmb}=
\left(\calv^{\lda,\rma,\ell}_{nm}\right)_{;k^\rmb}+\left(\calv^{\cals,\rma,\ell}_{nm}\right)_{;k^\rmb}\to$ 
\verb=gdcalVsig=
\end{itemize}

\subsection{Bulk expressions}

For a bulk $\calc^\ell_{nm}(\bfk)=\gd_{nm}$, then
$(\calc^\ell_{nm}(\bfk))_{;\bfk}=0$, and Eq.~\eqref{c-a.1} reduces to
\begin{align}\label{choni.9}
v^{\gs,\rma}_{nm}
&=
v^{\lda,\rma}_{nm}
+
v^{\cals,\rma}_{nm}
\nonumber\\
\bfv^\gs_{nm}(\bfk)
&=
\left(1+\frac{\gS}{\go_c(\bfk)-\go_v(\bfk)}\right)\bfv^\lda_{nm}(\bfk)
\quad\quad n\notin D_m
\nonumber\\
\bfv^\gs_{nn}(\bfk)
&=
\bfv^\lda_{nn}(\bfk)
,
\end{align}
where in \verb=$TINIBA/latm= the values are coded in the array
called
\verb=momMatElem=.  
If option \verb=-n= is given while running
\verb=all_resposnses.sh=, then $\bfv^\nl_{nm}(\bfk)$ are included in 
\verb=momMatElem=. 
Also,
\begin{align}\label{c-a.1nn}
\left(v^{\gs,\rma}_{nm}\right)_{;k^\rmb}
&=
\left(
v^{\lda,\rma}_{nm}
\right)_{;k^\rmb}
+
\left(
v^{\cals,\rma}_{nm}
\right)_{;k^\rmb}
\nonumber\\
&=
\left(
v^{\lda,\rma}_{nm}
\right)_{;k^\rmb}
+
i\gS f_{mn}(r^\rma_{nm})_{;k^\rmb}
\nonumber\\
\left(v^{\gs,\rma}_{mn}\right)_{;k^\rmb}
&=
\left(\left(v^{\gs,\rma}_{nm}\right)_{;k^\rmb}\right)^*
,
\end{align}
where with the r.h.s. expressions are given above.

$\bullet$ Coding:
$\bfv^\gs_{nm}(\bfk)\to$ \verb=momMatElem=,
$\left(v^{\lda,\rma}_{nm}\right)_{;k^\rmb}\to$ \verb=gdVlda=,
$\left(r^{\lda,\rma}_{nm}\right)_{;k^\rmb}\to$ \verb=derMatElem=, and
$\left(v^{\gs,\rma}_{nm}\right)_{;k^\rmb}\to$ \verb=gdVsig=

\subsection{Layer or Bulk calculation}
\begin{itemize}
\item Layer: The layer calculation is done by using
Eqs.~\eqref{c-calvimchiewn}, \eqref{c-calvimchie2wn}, \eqref{c-calvimchiwn}
 and \eqref{c-calvimchi2wn}. 
\item Bulk:
A bulk calculation can be performed by using the same
Eqs.~\eqref{c-calvimchiewn}, \eqref{c-calvimchie2wn}, \eqref{c-calvimchiwn}
 and \eqref{c-calvimchi2wn}, and by simply replacing
\begin{enumerate}
\item $\calbv^\gs_{nm}$ (\verb=calMomMatElem=) $\to \bfv^\gs_{nm}$ (\verb=momMatElem=) 
\item $(\calbv^\gs_{nm})_{;\bfk}$ (\verb=gdcalVsig=) $\to
  (\bfv^\gs_{nm})_{;\bfk}$ (\verb=gdVsig=) 
\end{enumerate}
\item Therefore: For the code to run either possibility we use the
  same arrays as for the layered response, where, if bulk is chosen, it
  simply copies the bulk matrix elements into the layer arrays, i.e.
\begin{itemize}
\item Layer: 
$\calbv^\gs_{nm}$ (\verb=calMomMatElem=) and
$(\calbv^\gs_{nm})_{;\bfk}$ (\verb=gdcalVsig=) 
\item Bulk:
$\bfv^\gs_{nm}$ (\verb=momMatElem=$\to$\verb=calVsig=) and
$(\bfv^\gs_{nm})_{;\bfk}$ (\verb=gdVsig=$\to$\verb=gdcalVsig=) \\
This change is done in \verb=set_input_ascii.f90= (look for
\verb=layer-to-bulk= tag)
\item ID: Notice that we have assigned\\
  \verb=calMomMatElem=$\to$ \verb=calVsig= (keeping
  \verb=calMomMatElem=), so it is easier to  code the responses.
Therefore, we have\\
$\calbv^\gs_{nm}\to$ \verb=calVsig= and
$(\calbv^\gs_{nm})_{;\bfk}\to$ \verb=gdcalVsig=\\
either for bulk or layered response. \\
If \verb=calMomMatElem= is not used, we should get rid of it (ID at work).
\end{itemize}
\end{itemize}

\subsection{$\calv$ vs ${\cal R}$}

Using 
$\mathrm{Re}[iz]=-\mathrm{Im}[z]$,
$\mathrm{Im}[iz]=\mathrm{Re}[z]$,
and
\begin{equation}\label{rnmenm91}
{\cal R}^\rma_{nm}=\frac{{\cal P}^\rma_{nm}}{im_e\go_{nm}}=\frac{\calv^\rma_{nm}}{i\go_{nm}}
\quad\quad n\ne m
,
\end{equation}
we can show the equivalence between the two formulations, i.e.
\input{appendices/chis-v-r}
\noindent If we take ${\cal R}^{\rma,\ell}_{nm}\to r^\rma_{nm}$, 
we
would recover the expressions for a bulk response.
We prefer to use the expressions in terms of $\calv^\ell$, since they are
more physically appealing, as the velocity is what gives the current
of a given layer, from which the polarization is computed and the
$\chi^\ell$ extracted.

\textcolor{red}{Remark:} We mention that above expressions with 
${\cal R}^{\rma,\ell}_{nm}\to r^\rma_{nm}$, are coded in
\verb=integrands.f90=, instead of Eq. 40 and 41 of Cabellos et
al.\cite{cabellosPRB09}, which were derived by using Eq. 19 of Aversa
and Sipe.\cite{aversaPRB95} To obtain above equations, we started
from Eq. 18 of Aversa and Sipe,\cite{aversaPRB95} which has the
advantage that applying the layer-by-layer formalism is very
transparent and straightforward. This coding is what constitutes the
{\it Length}-gauge implementation in \tiniba, which is, within a very
small numerical difference, equal to the {\it Velocity}-gauge
implementation of Eq. 35 of Cabellos et al.\cite{cabellosPRB09}, also
in \tiniba.

\subsection{Other responses}

\textcolor{red}{Warning: the layered responses MUST be looked at
  again, and modified according to the newly calculated 
$\calbv^\gs_{nm}$  and
$(\calbv^\gs_{nm})_{;\bfk}$. Linear response, current and
spin injection, should be revisited again!!}

\begin{itemize}
\item Injection Current\\
We need 
$\bfv^\gs_{nn}(\bfk)$
 or 
$\calbv^\gs_{nn}(\bfk)$, but $\bfv^\cals_{nn}(\bfk)=0$
and 
$\calbv^\cals_{nn}(\bfk)=0$ (proven numerically, would be nice to try
analytically), since the velocity of the electron in the conduction
bands should not depend on the scissors rigid ($\bfk$-independent) correction thus
\begin{align}\label{z.1}
\calbv^\gs_{nn}(\bfk)&=\calbv^\lda_{nn}(\bfk)
\nonumber\\ 
\bfv^\gs_{nn}(\bfk)&=\bfv^\lda_{nn}(\bfk) 
,
\end{align}
contained in  \verb=CalMomMatElem= and  \verb=momMatElem=,
respectively. Both would have the contribution from $\bfv^\nl$ if the
options \verb=(-v,-n)= are used. If $\bfv^\nl$ is neglected, the
option \verb=-l= for a layer calculation would be much faster as we
only need to calculate the diagonal elements of Eq.~\eqref{eni.2}, but
since the idea is to {\it always} include it, we are obliged to use
Eq.~\eqref{c-a.2}, where $\calc^\ell_{nm}(\bfk)$ is needed, and thus
we ought to use option \verb=-c=. 
Since  \verb=CalMomMatElem= is calculated for off-diagonal elements
only, we have added a \verb=do= loop in \verb=set_input_ascii.f90=
 to compute the diagonal  part, Eq.~\eqref{z.1}, which is
stored in \verb=calVsig=. In accordance to \ref{ccu4}, we have checked
that we obtain the same results by using Eq.~\eqref{eni.2} or
Eq.~\eqref{ccu.1}, in a layered injection current calculation, which
means that the results obtained thus far in our articles are correct,
of course, neglecting $\bfv^\nl$.

INCLUDE FIGURES.
\end{itemize}

\subsection{Consistency check-up 1}

To check that the layered expressions
Eqs.~\eqref{calvimchiewn}, \eqref{calvimchie2wn}, \eqref{calvimchiwn}
 and \eqref{calvimchi2wn}, agree with a bulk calculation, we must
 take 
$\calbv^\gs_{nm}\to\bfv^\gs_{nm}$
and
$\calbv^\gs_{nm;\bfk}\to\bfv^\gs_{nm;\bfk}$. To do this, proceed as
follows
\begin{enumerate}
\item Run bulk GaAs using  \verb=rlayer.sh= and \verb=chose_layers.sh=
as if it were a surface, even though it make no sense.
\item In \verb=$TINIBA/latm/SRC_1setinput/set_input_ascii.f90= look for\\
\verb=!########## MIMIC A BULK RESPONSE #######d=\\
and follow instructions given there.
\item Compile \verb=set_input_*= in \verb=$TINIBA/latm/SRC_1setinput=
\item run \verb=all_responses.sh= using\\ 
\verb=-w layer -r 44 ...=\\
\verb=-w total -r 21 ...=\\
and \\
\verb=-w total -r 42 ...=\\
thus obtaining a \verb=layer= calculation using bulk matrix elements, a
\verb=total= calculation for the length  and the velocity gauge, and plot the
three $\chi$'s, they ouught to be identical, if not CRY!. Try out to
reproduce Fig.~\ref{gaas}  
\end{enumerate}
\begin{figure}[b]
\centering
\includegraphics[scale=.7]{plots/shg-bulk}
\caption{Im[$\chi_{xyz}$] for GaAs, 10 Ha and 47 $\bfk$-points, using
  the layered formulation and mimicking a bulk. 
The correction
  due to $\bfv^\nl$, also agrees with the velocity and the layered
  approach (not shown in the figure for clarity).}
\label{gaas}
\end{figure}

\subsection{Consistency check-up 2}

In Fig.~\ref{si111as} we show
Im[$\chi_{xx}$] for a surface, where the 
The full-slab result is twice the half-slab
result, with or without $\bfv^\nl$,  as it must be. Also, the scissors
correction rigidly shifts the spectrum by $\hbar\gS$ as it should be.

\begin{figure}[b]
\centering
\includegraphics[scale=.7]{plots/surface-chi}
\caption{Im[$\chi_{xx}$] 
for a Si(111):As surface of 6-layers, 5 Ha and 14 $\bfk$-points using
the layered formulation. 
The full-slab result is twice the half-slab
result, as it must be.
}
\label{si111as}
\end{figure}

\subsection{Consistency check-up 3}

Check-of-Checks: 
A (100) $2\times 1$ surface has $\chi_{xxx}$
 different from zero,
whereas the ideally terminated (100) surface has $\chi_{xxx}=0$.
Clean Si(100) has the $2\times 1$ surface as a possible
reconstruction. Then, to calculate such a surface, one can use
a slab such that its front surface is the reconstructed 
Si(100)$2\times 1$ surface and its back surface is H-terminated.
Therefore,
 for the
layer-by-layer scheme one should expect that
\begin{align}\label{cc3}
\chi^{\mathrm{half-slab}}_{xxx}
\equiv
\chi^{\mathrm{full-slab}}_{xxx}
,
\end{align}
since the contribution from the back surface (H-terminated), would
have zero contribution, since this tensor component of $\chi$ is
symmetry forbidden. Fancy at
Fig.~\ref{si-2x1}, and notice that $\chi^\nl<\chi$. i.e. the
susceptibility with the inclusion of the non-local part of the
pseudopotential is smaller than that without it.\\
King-of-Kings: Rejoice at 
Fig.~\ref{si-2x1-n-vs-b}.
\begin{figure}[b]
\centering 
\includegraphics[scale=.7]{plots/shg-si-2x1-16}
\caption{$|\chi_{xxx}|$ 
for a Si(100)$2\times 1$ surface of 16 Si-layers and one H layer, 10 
Ha, 132 bands,  244 $\bfk$-points, and 1000 pwvs in \depe,
 using 
the layered formulation. We see that 
$\chi^{\mathrm{half-slab}}_{xxx}
\sim 
\chi^{\mathrm{full-slab}}_{xxx}$,
validating the layer-by-layer approach. 
}
\label{si-2x1}
\end{figure}
\begin{figure}[b]
\centering
\includegraphics[scale=.7]{plots/shg-si-2x1-n-vs-b}
\caption{$|\chi_{xxx}|$ 
for a Si(100)$2\times 1$ surface of 12 Si-layers and one H layer, 5
Ha, 100 bands and 244 $\bfk$-points for the CIO-\tiniba-coding
and 256 $\bfk$-point for the X-DP\copyr-coding.
Both broadened by 0.1 eV.   
}
\label{si-2x1-n-vs-b}
\end{figure}
\begin{figure}[b]
\centering 
\includegraphics[scale=1.5]{plots/shg-si-2x1-16-compa}
\caption{$|\chi_{xxx}|$ 
for a Si(100)$2\times 1$ surface of 16 Si-layers and one H layer, 10 
Ha, 132 bands, 244 $\bfk$-points and and 1000 pwvs in \depe, using 
the layered formulation. 
``Full'' uses full coding of 
$\calv^\cals_{nm}$ and 
$\calv^\cals_{nm:\bfk}$ through Eq.~\eqref{a.3b};
``only-$\calv^\cals$'' uses 
$\calv^\cals_{nm}$ through Eq.~\eqref{a.3b} and 
$\calv^\cals_{nm:\bfk}$ through Eq.~\eqref{ccu55};
``alternative'' uses $\calv^\cals_{nm}$ through Eq.~\eqref{temp.1} and 
$\calv^\cals_{nm:\bfk}$ through Eq.~\eqref{ccu55}.
Also, we show the results for 2000 pwvs.
Notice that all the curves are almost identical to each other. 
}
\label{si-2x1-compa}
\end{figure}

\subsection{Consistency check-up 4}\label{ccu4}

To check that the coding of 
$\calc^\ell_{nm}(\bfk)$ 
is correct, we can calculate $\calv^{\rma,\ell}_{nm}(\bfk)$ using
Eq.~\eqref{vcali} as follows
\begin{align}\label{ccu.1}
\calv^{\rma,\ell}_{nm}(\bfk)
&=
\frac{1}{2m_e}
\Big(
\calc^\ell(z)p^\rma
+
p^\rma\calc^\ell(z)
\Big)_{nm}
\nonumber\\
&=
\frac{1}{2m_e}
\sum_q
\Big(
\calc^\ell_{nq}p^\rma_{qm}
+
p^\rma_{nq}\calc^\ell_{qm}
\Big)
,
\end{align}
which must give the same results as those computed through
Eq.~\eqref{eni.2}.
Indeed, we have checked that this is the case. The
\verb=$TINIBA/util/consistency-of-cfmn.sh=
is used to check this.

\subsection{Consistency check-up 5}\label{ccu5}

When the \verb=-n= option is chosen, using \verb=all_responses.sh= as
coded above doesn't give consistent results, i.e. $\chi$
 with $\bfv^\nl$  
is not smaller than $\chi$ 
 without $\bfv^\nl$. Thus, we follow the bellow approach instead.

We use Eq.~\eqref{nmesn}
\begin{align}\label{nmesnn}
(\calv^{\lda,\rma}_{nm})_{;k^{\rmb}}&=
\frac{\hbar}{m_e}\gd_{\rma\rmb}
C^\ell_{nm} 
-i 
\sum_p 
[r^{\rmb},v^{\nl,\rma}]_{np}C^\ell_{pm} 
+i 
\sum_{\ell}
\bigg(
r^{\rmb}_{n\ell}  
\calv^{\lda,\rma}_{\ell m}
-
\calv^{\lda,\rma}_{n\ell}   
r^{\rmb}_{\ell m}
\bigg)  
+i  
r^{\rmb}_{nm}
\tilde\Delta^{\rma}_{mn}
,
\end{align}  
where 
\begin{eqnarray}\label{tdeln}
\tilde\Delta^{\rma}_{mn}
=
\calv^{\lda,\rma}_{nn}  
-
\calv^{\lda,\rma}_{mm}  
, 
\end{eqnarray}
which is coded instead of Eq.~\eqref{c-a.2nn}. 
 As mentioned before, the term $[r^{\rmb},v^{\nl,\rma}]_{nm}$
calculated in Appendix \ref{calt}, is small 
compared to the other terms, thus we neglect it throwout this work.\cite{valerie} 
The expression for $C^\ell_{nm}$ is calculated in Appendix \ref{calpcalc}.

Likewise, with the help
of Eq.~\eqref{a_gradw2} into
Eq.~\eqref{c-choni.1}, we obain
\begin{align}\label{ccu54}
(v^{\cals,\rma}_{nm})_{;k^\rmb}&=
i\gS f_{mn}(r^\rma_{nm})_{;k^\rmb}
=
i\gS f_{mn}\left(\frac{v^{\lda,\rma}_{nm}}{i\go^\lda_{nm}}\right)_{;k^\rmb}
\nonumber\\
&=
\gS\frac{f_{mn}}{\go^\lda_{nm}}\left[
\left(v^{\lda,a}_{nm}\right)_{;k^\rmb}
-
\frac{v^{\lda,a}_{nm}}{\go^\lda_{nm}}\left(\go^\lda_{nm}\right)_{;k^\rmb}
\right]
\nonumber\\
&=
\gS\frac{f_{mn}}{\go^\lda_{nm}}\left[
\left(v^{\lda,a}_{nm}\right)_{;k^\rmb}
-
\frac{\gD^{b}_{nm}}{\go^\lda_{nm}}v^{\lda,a}_{nm}
\right]
,
\end{align}
which is generalized as follows 
\begin{align}\label{ccu55}
(\calv^{\cals,\rma}_{nm})_{;k^\rmb}&=
\gS\frac{f_{mn}}{\go^\lda_{nm}}\left[
\left(\calv^{\lda,a}_{nm}\right)_{;k^\rmb}
-
\frac{\gD^{b}_{nm}}{\go^\lda_{nm}}\calv^{\lda,a}_{nm}
\right]
,
\end{align}
although, I haven't found a way to prove this rigorously, it gives 
very similar results to those obtained by Eq.~\eqref{c-a.3bnn}, which 
is coded. 
%which is coded instead of 
The following is also tempting,
\begin{align}\label{temp.1}
v^{\cals,a}_{nm}
&=\gS\frac{f_{mn}}{\go^\lda_{nm}}
v^{\lda,a}_{nm}
\nonumber\\
\calv^{\cals,a}_{nm}
&=\gS\frac{f_{mn}}{\go^\lda_{nm}}
\calv^{\lda,a}_{nm}
.
\end{align}
Again, I haven't found a way to prove this rigorously, but it gives 
very similar results to those obtained by Eq.~\eqref{a.3b}, which 
is coded. In Fig.~\ref{si-2x1-compa} we show the comparison between
the two alternatives, from where we see that they are basically equivalent.

\subsection{Subroutines}

The following subroutines/shells are involved in the coding,
and are documented between\\
\verb=#BMSd=\\
$\vdots$\\
\verb=#BMSu=\\
marks.
\begin{enumerate}
\item \verb=$TINIBA/utils/all_responses.sh=
\item \verb=$TINIBA/latm/SRC_1setinput/inparams.f90=\\
\textcolor{red}{Warning:} compile both\\
\verb=$TINIBA/latm/SRC_1setinput/= \\
and\\
\verb=$TINIBA/latm/SRC_2latm/= 
\item \verb=$TINIBA/latm/SRC_1setinput/set_input_ascii.f90=\\
\end{enumerate}

prescindir

### 2014-11-16 appendices/appendix-vc.tex
initial commit

\subsection{ Expressions for 
\texorpdfstring{${\cal V}^{a,\ell}_{nm}(\bfk)$, 
${\cal C}^{\ell}_{nm}(\bfk)$ 
and
$({\cal C}^{\ell}_{nm}(\bfk))_{;\bfk}$
}{Vnm and Cnm}}\label{calpcalc}

Expanding the wave function in plane waves we obtain
\begin{align}\label{eni.1}
\psi_{n\bfk}(\bfr)=\sum_\bfG A_{n\bfk}(\bfG)e^{i(\bfk+\bfG)\cdot\bfr}
,
\end{align}
where $\{\bfG\}$ are the reciprocal basis vectors satisfying
$e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
space, and $A_{n\bfk}(\bfG)$ are the expansion coefficients. Using
$m_e\bfv=-i\hbar\bfgnabla$ into the equivalent of Eq.~\eqref{intj}
 we obtain,\cite{mendozaPRB06}
\begin{align}\label{eni.2}
\calbv^\ell_{nm}(\bfk)=
\frac{\hbar}{2m_e}
\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
(2\bfk+\bfG+\bfG')
\gd_{\bfG_\parallel \bfG'_\parallel}  
f^\ell(G_\perp-G'_\perp)
,
\end{align}   
where
\begin{align}\label{vnl.9}
f^\ell(g)=\frac{1}{L}\int_{z_\ell-\gD^b_\ell}^{z_\ell+\gD^f_\ell} e^{igz}dz  
 ,
\end{align}
with $f^{\ell*}(g)=f^\ell(-g)$. 
the reciprocal lattice vectors $\bfG$ are decomposed into components
parallel to the surface $\bfG_\parallel$, and perpendicular to the
surface $G_\perp \hat z$, so
that $\bfG = \bfG_\parallel + G_\perp\hat z$.
The double summation over the $\bfG$ vectors can be efficiently done by  
creating a pointer array to identify all the plane-wave coefficients  
associated with the same $G_\parallel$.  
 Likewise we obtain that
\begin{align}\label{eni.4}
\calc^\ell_{nm}(\bfk)=
\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
\gd_{\bfG_\parallel \bfG'_\parallel} 
f_\ell(G_\perp-G'_\perp)
.
\end{align}  
If $\calc^\ell(z)=1$, $f_\ell(g)=\gd_{g0}$, and from Eqs.~\eqref{eni.2} and \eqref{eni.4}
we obtain the full-slab or bulk values
$\bfv_{nm}(\bfk)$ and
$\calc^\ell_{nm}(\bfk)=\gd_{nm}$.

Since for any function $F(\bfr)$, $[\bfr,F(\bfr)]=0$, using 
Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn}, we obtain
\begin{align} 
(\calc^\ell_{nm})_{;\bfk}
&=
i
\sum_{q} 
 \left(\bfr_{nq}
\calc^\ell_{qm}
-
\calc^\ell_{nq}
\bfr_{qm}
\right) 
+i\bfr_{nm}(\calc^\ell_{mm}-\calc^\ell_{nn}) 
,
\end{align} 
where we remind the reader that $\bfr_{nm}$ are calculated through
Eq.~\eqref{chon.10} for LDA.

### 2014-11-16 paper-v-0.tex
initial commit

\documentclass[floatfix,prb,aps,superscriptaddress,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\allowdisplaybreaks[1]%biutiful equation breaker!!!
\usepackage{ulem}
\usepackage{subfigure}
%\input{/Users/bms/util/definitions}
\input{definitions}
\usepackage{bm}
%%%%*****************%%%% fine hyperef 
\usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{showlabels}
%\usepackage{showkeys}

\begin{document}

\title{Length Gauge Theory of Surface Second Harmonic Generation}
\author{Sean M. Anderson}
\affiliation{Centro de Investigaciones en Optica, Le\'on, Guanajuato, M\'exico}
\author{Nicolas Tancogne-Dejean}
\affiliation{Laboratoire des Solides Irradi\'es, \'Ecole 
  Polytechnique, CNRS, CEA/DSM, 91128 Palaiseau, France}
 \affiliation{European Theoretical Spectroscopy Facility (ETSF), Palaiseau, France}
\author{Bernardo S. Mendoza}\email{bms@cio.mx}
\affiliation{Centro de Investigaciones en Optica, Le\'on, Guanajuato, M\'exico}
\author{Val\'erie V\'eniard}
\affiliation{Laboratoire des Solides Irradi\'es, \'Ecole 
  Polytechnique, CNRS, CEA/DSM, 91128 Palaiseau, France}
 \affiliation{European Theoretical Spectroscopy Facility (ETSF), Palaiseau, France}

\begin{abstract}
We present a theoretical review of surface second harmonic generation (SHG) 
from semiconductor surfaces based on the longitudinal gauge. This 
layer-by-layer analysis is carefully presented in order to show how a 
surface SHG calculation can be readily evaluated. The nonlinear susceptibility 
tensor for a surface, $\boldsymbol{\chi}^S(-2\go;\go,\go)$
 is split into two terms relating to inter-band 
and intra-band one-electron transitions.

\end{abstract}

\maketitle
%\tableofcontents

\section{Theory}

\label{theory}

In this section we present the scheme used to calculate the surface second-order
nonlinear response.
Our derivation includes new terms not included
before in the length gauge formalism for the surface nonlinear
response.

\subsection{Perturbation approach}

We use the independent particle approximation and neglect local field
and excitonic effects and
treat the electromagnetic field classically, while the matter is
described quantum-mechanically.
We can describe the system using 
%a scaled 
the one electron density
operator ${\rho}$, with which we can calculate the expectation value of a
single-particle observable $\mathcal{O}$ as 
$\langle{\cal O}\rangle=\mbox{Tr}({\rho} {\cal O}),  
\label{calo}$
with $\mathcal{O}$ the associated quantum mechanical operator and Tr
the trace. 
The density operator satisfies
$
i\hbar (d{\rho}/dt)=[H(t),{\rho}],  
\label{rho}
$
with $H(t)$ as the total single electron Hamiltonian, written as 
\begin{equation*}
H(t)=H_{0}+H_{I}(t),  
\label{ache}
\end{equation*}
where $H_{0}$ is the unperturbed time-independent Hamiltonian, and $H_{I}(t)$
is the time-dependent potential energy due to the interaction of the
electron with the electromagnetic field.
% ; $H_{0}$ has eigenvalues
% $\hbar \omega_{n}(\mathbf{k})$
% and eigenstates $| n\mathbf{k} \rangle$ (Bloch states) labeled by a band 
% index $n$ and crystal momentum $\mathbf{k}$.
To proceed with the solution of $\rho$  it is convenient to use the
interaction picture, where a unitary operator $U=\exp ({iH_{0}t/\hbar })$
transforms any operator $\mathcal{O}$ into 
$\tilde{\mathcal{O}}=U{\cal O}U^{\dagger }$. Even if $\mathcal{O}$ does not depend on $t$, $\tilde{
\mathcal{O}}$ does through the explicit time dependence of $U$. 
The dynamical 
equation for $\tilde{\rho}$ is 
given by
\begin{equation*}
i\hbar \frac{d\tilde{{\rho}}}{dt}=[\tilde{H}_{I}(t),\tilde{{\rho}}],  
\label{rho1}
\end{equation*}
with solution 
\begin{equation}
i\hbar \tilde{{\rho}}(t)=i\hbar \tilde{{\rho}}_{0}+\int_{-\infty }^{t}dt^{\prime }[
\tilde{H}_{I}(t^{\prime }),\tilde{\rho}(t^{\prime })],  
\label{trans}
\end{equation}
where $\tilde{\rho}_{0}=\tilde{\rho}(t=-\infty )$ is the unperturbed density matrix. We look
for the standard perturbation series solution, 
$\tilde{\rho}(t)=\tilde{\rho}^{(0)}+\tilde{\rho}^{(1)}+\tilde{\rho}^{(2)}+\cdots$,  
where the superscript denotes the order (power) with which each term depends
on the perturbation $H_{I}(t)$. From Eq.~\eqref{trans} the $N$-th order
term is 
\begin{equation}
\tilde{{\rho}}^{(N)}(t)=\frac{1}{i\hbar }\int_{-\infty }^{t}dt^{\prime }[\tilde{
H}_{I}(t^{\prime }),\tilde{\rho}^{(N-1)}(t^{\prime })].  
\label{rhop}
\end{equation}
The series is generated by the unperturbed density operator $\tilde{\rho}
^{(0)}\equiv \tilde{\rho}_{0}$, assumed to be the diagonal Fermi-Dirac distribution, 
$\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle=f(\hbar \omega_{n}(\mathbf{k}))\equiv f_{n}$. For a
clean, cold semiconductor $f_{n}=1$ for $n$ a valence ($v$) or
occupied band and zero for $n$ a conduction ($c$) or empty band. \ This we
assume throughout.
We remark that the expectation values satisfy
$
\langle{\cal O}\rangle=
\mbox{Tr}({\rho}{\cal O})
=
\mbox{Tr}(\tilde{{\rho}}\tilde{\cal O})
$.

We will look for the expectation value of the macroscopic current density, 
$\mathbf{J}$, given by 
\begin{equation}
\bfJ=\langle{\mathbf{J}}\rangle=\frac{e}{\Omega}\mbox{Tr}({\rho}\dot{\mathbf{r}})
,
\label{pe}
\end{equation}
where $\dot{\mathbf{r}}$ is the time derivative of the position operator of the
electron of charge $e$, 
\begin{equation}
\mathbf{v}\equiv \dot{\mathbf{r}}=\frac{1}{i\hbar }[\mathbf{r},H_0],  
\label{mv}
\end{equation}
with $\mathbf{v}$ the velocity operator of the electron, and $\Omega$ the
normalization volume. We calculate the macroscopic polarization density 
$\mathbf{P}$, related to $\bfJ$ by
$\bfJ=d\mathbf{P}/dt$. For a 
perturbing (Maxwell macroscopic) electromagnetic field, $\mathbf{E}(t)=
\mathbf{E}(\omega )e^{-i\tilde{\omega} t}+c.c.$,
where $\tilde\omega=\omega+i\eta $,
and $\eta >0$ is used to adiabatically turn on the interaction,
we write the second order nonlinear
polarization as, 
\begin{equation}
P^a(2\omega)=\chi ^{abc}(-2\omega;\omega,\omega)E^{b}(\omega)E^{c}(\omega),  
\label{pshg}
\end{equation}
where $\chi^{abc}(-2\omega ;\omega ,\omega )$ is the nonlinear
susceptibility responsible of surface second harmonic generation
(SSHG). 
The 
superscripts in Eq.~\eqref{pshg} denote Cartesian components, and if
repeated are to be summed over. Without loss of generality we can always
define $\chi^{abc}(-2\omega;\omega,\omega)$
 to satisfy intrinsic permutation
symmetry, 
$\chi^{abc}(-2\omega ;\omega ,\omega )=\chi ^{acb}(-2\omega ;\omega
,\omega )$.
% ;
% if it did not, the part that did not satisfy intrinsic permutation symmetric
% would make no contribution to the second order polarization. This part
% could be dropped since it would
% have no physical significance.

The unperturbed Hamiltonian 
is used to solve the Kohn-Sham equations\cite{kohnPR65} of Density  
Functional Theory (DFT), for convenience usually within the Local  
Density Approximation (LDA). Thus, we labeled the hamiltonian with the  
superscript LDA for convenience, although we could use any other
approximation, like GGA, and our derivation would remain the same. Then,
\begin{equation}\label{h0}
H^{\mathrm{LDA}}_{0}=\frac{p^{2}}{2m_e}+V
\end{equation}
with $m_e$ the mass of the electron, $\mathbf{p}$ its canonical momentum, and 
$V$
 the periodic crystal potential, where we neglect spin-orbit terms.
To be more general in our derivation of
$\chi^{abc}(-2\omega;\omega,\omega)$, we asume
that the crystal potential could have a spatially non-local
contribution, as is customary for most
pseudopotentials, and then we replace $V$ with
\begin{align}\label{ps}
V^{\mathrm{ps}}=V(\bfr)+V^{\mathrm{nl}}
,
\end{align}
where 
$V(\bfr)$ and $V^{\mathrm{nl}}$  are the local and 
non-local contributions to the pseudopotential $V^{\mathrm{ps}}$,
 respectively.
Now, the 
Schr\"odinger equation reads 
\begin{align}\label{ache.4} 
\left(
\frac{-\hbar^2}{2m_e}\nabla^2 
 +  V(\bfr)\right)\psi_{n\bfk}(\bfr) 
 + \int d\bfr' V^\nl(\bfr,\bfr')\psi_{n\bfk}(\bfr') 
=\hbar \omega_i\psi_{n\bfk}(\bfr) 
,
\end{align} 
where
$V^\nl(\bfr,\bfr') = \bra{\bfr} V^\nl\ket{\bfr'}$,
and
$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
is cell periodic. 
In case of
a local potential, i.e. $V= V(\bfr)$, 
like that of all-electron schemes, 
we simply omit the contribution coming
from the non-local part of the pseudopotential, $V^{\mathrm{nl}}$, in
the results that we will derive.

As is well known, the use of the solutions of Eq.~\eqref{ache.4} as single 
particle states leads to an underestimation of the band gap. A standard
procedure to correct for this is to
use the so-called ``scissors approximation'', by which one rigidly shifts the
conduction bands in energy so that the band gap corresponds to the accepted
experimental band gap; this is often in fairly good agreement with the GW
band gap based on a more sophisticated calculation.\cite{hybertsenPRB86}
Concurrently, one uses the LDA wave functions, since they produce band
structures with dispersion relations similar to those predicted by the GW
approximation. Mathematically, one adds the scissors (non-local) term 
$S$, to the unperturbed or unscissored Hamiltonian $H^{\mathrm{LDA}}_{0}$, i.e. 
\begin{equation*}
H^\sigma_{0}=H^{\mathrm{LDA}}_{0}+S
\end{equation*}
where 
\begin{equation}
S=\hbar \sigma\sum_{n}\int d^{3}k(1-f_{n})
|n\mathbf{k}\rangle\langle n\mathbf{k}|,
\label{hats}
\end{equation}
with $\hbar \sigma$  the rigid ($\mathbf{k}$-independent) energy correction to be
applied. 
%Several properties of $S(\mathbf{r},\mathbf{p})$ are shown in the
%appendix.
The unscissored and scissored Hamiltonians satisfy 
\begin{eqnarray*}
H^{\mathrm{LDA}}_{0}\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
\label{hamils} \\
H_{0}^\sigma\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\sigma
(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
\end{eqnarray*}
where the scissor-shifted energies, 
$\omega_{n}^\sigma(\mathbf{k})$ , are given by
\begin{equation}
\omega_{n}^\sigma(\mathbf{k})=\omega^\lda_{n}(\mathbf{k})+(1-f_{n})\sigma.
\label{wese}
\end{equation}
We emphasize that the scissored
Hamiltonian has the same eigenfunctions as the unscissored Hamiltonian.

\subsection{Length Gauge Formalism}
To calculate the optical response 
within the Length Gauge Formalism, the interaction Hamiltonian is given by 
\begin{equation}
H_{I}(t)=-e\mathbf{r}\cdot \mathbf{E}(t).  \label{rde}
\end{equation}
The 
treatment of the position operator
$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$, has been
described,\cite{adamsJCP53,blountSSP62} and in particular 
in Ref.~\onlinecite{aversaPRB95} the correct treatment of $\bfr$ for
the calculation of the non-linear optical response is
developed. Indeed, 
 $\bfr$ is split into the {\it intraband} 
($\bfr_i$) and {\it interband} ($\bfr_e$) parts , where 
$\bfr=\bfr_i+\bfr_e$. 
For $\bfr_e$ one uses
\begin{align}\label{rnminn}
\bra{n\bfk}\bfr_e\ket{m\bfk'} &=
(1-\gd_{nm})\gd(\bfk-\bfk')\bfgxi_{nm}(\bfk) 
,
\end{align}
such that $\bfr_{e,nm}=0$ for $n=m$, and thus
from Eq.~\eqref{mv}, with $H_0\to
H^\sigma_0$, we obtain
\begin{align}\label{pmnrmn}
\bfr_{e,nm}(\bfk) =
\bfgxi_{nm}(\bfk)\equiv 
\bfr_{nm}(\bfk) 
&=
\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
\quad\quad n\notin D_m 
,
\end{align}  
where we defined
$\go^\gs_{nm}(\bfk)\equiv\go^\gs_n(\bfk)-\go^\gs_m(\bfk)$, and
$D_m$ are all the possible degenerate $m$-states. 
%We recognize that for each instance in which 
%$n\notin D_m$ or $n\ne m$, $\bfr_{nm}(\bfk)$ are interband matrix
%elements.
For the intraband part, in turns out that $\bfr_i$ only appears in
commutators, as one does the derivation of
the optical response, then one uses,\cite{aversaPRB95}
\begin{equation}\label{conmri3n}
\bra{n\bfk}[\bfr_i,\calo]\ket{m\bfk'}
=i\gd(\bfk-\bfk')(\calo_{nm})_{;\bfk}
,
\end{equation}  
where
\begin{equation}\label{gendevnn}
(\calo_{nm})_{;\bfk}=
\nabla_\bfk 
\calo_{n m}(\bfk) 
- 
i 
\calo_{nm}(\bfk) 
\left(
\bfgxi_{nn}(\bfk) 
-
\bfgxi_{mm}(\bfk) 
\right) 
,
\end{equation} 
is
the generalized derivative 
of the operator $\calo$. 
The vectors $\bfgxi_{nn}(\bfk)$ are defined in 
Ref.~\onlinecite{aversaPRB95}, however they do not need to be 
calculated explicitly in what follows.

Before continue, 
we derive a key result for the length gauge formulation. Using again $H^\sigma_0$ in
Eq.~\eqref{mv} we obtain
\begin{align}\label{vop2}
\bfv^\gs&=
\bfv 
+
\bfv^\nl 
+\bfv^S
=
\bfv^\lda 
+\bfv^S 
,
\end{align}
where we have defined 
\begin{subequations}
\begin{align}\label{ve}
\bfv 
&=\frac{\bfp}{m_e},
\\\label{vnl}
\bfv^\nl 
&=
\frac{1}{i\hbar}[\bfr,V^\nl],
\\\label{vs}
\bfv^S
&=
\frac{1}{i\hbar}[\bfr, S],
\\\label{vlda}
\bfv^\lda 
&=
\bfv 
+\bfv^\nl
,
\end{align}  
\end{subequations}
with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator and used
$[r^a,p^b]=i\hbar\gd_{ab}$, with $\gd_{ab}$ Kronecker's delta.
Using Eq.~\eqref{hats} we obtain 
\begin{equation}\label{chon.2} 
\bfv^S_{nm}=i\sigma f_{mn}\bfr_{nm},
\end{equation}
with $f_{nm}\equiv f_n-f_m$,
where we see that $\bfv^S_{nn}=0$. Then, it follows that
\begin{align}\label{chon.9}
\bfv^\sigma_{nm}(\bfk) 
=\frac{\go^\sigma_{nm}(\bfk)}{\go^\lda_{nm}(\bfk)}
\bfv^\lda_{nm}(\bfk) 
,
\end{align}
and
 \begin{align}\label{chon.8}
\frac{\bfv^\gs_{nm}(\bfk)}{\go^\gs_{nm}(\bfk)}
&=
\frac{\bfv^\lda_{nm}(\bfk)}{\go^\lda_{nm}(\bfk)}
,
\end{align}
thus from Eq.~\eqref{pmnrmn} 
\begin{align}\label{chon.10}
\bfr_{nm}(\bfk) 
=
\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
=
\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
\quad\quad n\notin D_m 
. 
\end{align}
The matrix elements 
of $\bfr_{nm}(\bfk)$ are the same whether we use the LDA or the scissored 
Hamiltonian and thus there is no need to label them.
Of course, is more convenient to calculate them
through $\bfv^\lda_{nm}(\bfk)$, that 
include the contribution of 
the non-local part of the pseudopotentials through
$\bfv^\nl_{nm}(\bfk)$. These in turn, can be readily
calculated
for 
fully separable nonlocal pseudopotentials in the 
Kleinman-Bylander 
form,\cite{francesco,mottaCMS10,kleinman_efficacious_1982,adolphPRB96}
see Appendix \ref{vesnl}.
Thus, the advantage of the length gauge formalism for linear and 
non-linear calculations of optical responses resides in the ease with 
which one can introduce the scissors operator into the calculations as 
it is also discussed in Nastos et al.\cite{nastosPRB05} Indeed, the 
length-gauge formalism for the scissored Hamiltonian can be easily worked 
out by simply using the unscissored LDA Hamiltonian,
$H_0^{\mathrm{LDA}}$,
 for the unperturbed system 
with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction, and then at the end of the 
calculation we only need to replace $\omega^{\mathrm{LDA}}_{n}$ with 
$\omega_{n}^{\sigma}$
 to obtain the 
scissored results for any susceptibility expression, whether linear or 
nonlinear.

Therefore,
taking the matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of
Eq.~\eqref{rde}, we obtain 
$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\sigma_{nm}-\tilde\omega)t}$,
with 
\begin{equation}\label{j.1}
B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\sigma_{nm}-\tilde\omega},
\end{equation}
and 
\begin{eqnarray}\label{j.2}
(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\sigma_{nm}-2\tilde\omega}\bigg[
i\sum_{\ell }\Big(r_{n\ell }^{b}B_{\ell m}^{c}-B_{n\ell }^{c}r_{\ell m}^{b}
\Big)  
-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\sigma_{nm}-2\tilde\omega)t}
.
\end{eqnarray}
We have used the fact that for a cold semiconductor $\partial
f_{n}/\partial \mathbf{k}=0$ and thus the intraband contribution to the linear
term vanishes identically. 
%In Appendix \ref{gender} we show how the generalized derivative of 
%$(B^a_{nm})_{;k^b}$ can be obtained.
Notice that in Eq.~\eqref{j.2} all the indices are different from each
other since due to the $f_{nm}$ factor in Eq.~\eqref{j.1},
$B^a_{nn}=0$. The dependence of $\bfk$ of all quantities is implicitly understood from
here on. 
Above expression will be used in the next subsection.

\subsection{Layered Current Density}\label{cd}

% In this section, we derive the expressions for the microscopic current
% density of a given layer in the unit cell of the system.

The approach we use to study the surface of a semi-infinite
semiconductor crystal is as follows. Instead of using a
semi-infinite system, we replace it by a super slab that consist of a finite
slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
super slab, which constitutes the unit cell,
is repeated to form a full three dimensional crystalline structure.
The slab itself consists of a front and back surface and sub-surface regions, and in between these
a region that is equivalent to
bulk of the system. 
In general the surface of a crystal reconstructs or relaxes as the atoms
move to find equilibrium positions. This is due to the fact that
the otherwise
balanced forces are disrupted when the surface atoms do not find their 
partner atoms that are now absent at the surface of the slab. 
To take the reconstruction or relaxation into account, we take ``surface'' to mean
the true surface of the first layer of atoms, and
some of the atomic sub-layers adjacent to it.
Since the front and the back
surfaces of the slab are usually identical the total slab is
centrosymmetric. This would imply that 
$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=0$, and thus we must
find a scheme 
in order to have a finite $\chi_{\rma\rmb\rmc}$ representative of the
surface. Even if the front and back surfaces of the slab 
are different, breaking the centrosymmetry and therefore giving an
overall $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}\ne 0$, we still
need a procedure to extract the front surface $\chi^f_{\rma\rmb\rmc}$
and the back surface $\chi^b_{\rma\rmb\rmc}$ from the slab
susceptibility 
$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=\chi^f_{\rma\rmb\rmc}-\chi^b_{\rma\rmb\rmc}$.
We have omitted the frequency arguments of $\chi_{abc}$ for notation convenience.

A convenient way to accomplish the separation of the SH signal of
either surface is to introduce a ``cut function'', $\calc(z)$, which is 
usually taken to be unity over one half of the slab and zero over 
the other half.\cite{reining_microscopic_1994}
 In this case $\calc(z)$ will give the contribution of the 
side of the slab for which $\calc(z)=1$. 
As was done for the linear response,\cite{mendozaPRB06}
we can generalize this 
simple choice for $\calc(z)$ by a top-hat cut function
$\calc^\ell(z)$ that selects a given layer,
\begin{equation}
\label{sz}
\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{\rmb})  
            \Theta(z_\ell-z+\Delta_\ell^f),
\end{equation} 
where $\Theta$ is the Heaviside function. Here, $\Delta_\ell^{f/b}$
is the distance that the $\ell$-th layer extends towards the front
$(f)$ or back $(b)$ from its $z_\ell$ position. 
 We take $z_\ell$
 at the center of an atom that 
belongs to layer $\ell$, and thus above equation would give the $\ell$-th 
 atomic-layer 
contribution to the non-linear optical response.
$\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
(see Fig.~\ref{fslab}).
\begin{figure}[b]
\centering
%\includegraphics[height=5cm,width=7cm]{slab}
\includegraphics[scale=.7]{images/slab}
\caption{A sketch of a slab where the circles represent atoms.\label{fslab}}
\end{figure}

To introduce the
 cut function $\calc^\ell(z)$ in
the calculation of $\chi_{\rma\rmb\rmc}$, we start from 
the
operator for the electron current,
$\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gs\ket{\bfr}\bra{\bfr}
+\ket{\bfr}\bra{\bfr}\bfv^\gs\right)
$
, 
that leads to
\begin{align}\label{jmic}
\bfj^{(N)}(\bfr,t)=\mathrm{Tr}(\bfj(\bfr)\rho^{(N)}(t))
=
\int \frac{dk^3}{8\pi^3}
\sum_{nm}
\rho^{(N)}_{nm}(\bfk;t)\bfj_{mn}(\bfk;\bfr)
,
\end{align}
where 
\begin{equation}\label{jmic3}
\bfj_{mn}(\bfk;\bfr)=
\frac{e}{2}
\left(
\bra{m\bfk}\bfv^\gs\ket{\bfr}\braket{\bfr}{n\bfk}
+
\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gs\ket{n\bfk}
\right),
\end{equation}
are the matrix elements of the microscopic current operator.
% and we have used the fact that the matrix elements between states $\ket{n\bfk}$
% are diagonal in $\bfk$, i.e. proportional to $\gd(\bfk-\bfk')$.
Integrating the microscopic current $\mbf{j}(\mbf{r},t)$ over
the entire slab gives the averaged macroscopic current density, $\bfJ(t)$. 
If we want the contribution from only one region of the unit cell 
towards the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
over the desired region. Then, the contribution to the current density from the
$\ell$-th layer of the slab is given by
\begin{equation}\label{jsz}
\frac{1}{\Omega}\int d^3r\, \calc^\ell(z)\, \mathbf{j}^{(N)}(\mathbf{r},t)
 \equiv \mathbf{J}^{(N),\ell}(t),
\end{equation}
where $\mathbf{J}^{(N),\ell}(t)$ is the current of the
$\ell$-th layer.
Therefore we define
\begin{equation}\label{vcal}
e{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}(\mathbf{k})
\equiv
%\frac{1}{\Omega}
\int d^3r\, \calc^\ell(z)\,\bfj_{mn}({\bfk};\bfr),
\end{equation}
to write the Fourier transform of Eq.~\eqref{jmic} as
\begin{equation}\label{jmac2}
\bfJ^{(N),\ell}(2\go)=\frac{e}{\gO}
\int \frac{dk^3}{8\pi^3}
\sum_{mn}
\calbv^{\gs,\ell}_{mn}(\mathbf{k}) 
\rho^{(N)}_{nm}(\bfk;2\go) 
, 
\end{equation}
that gives the induced microscopic current of the $\ell$-th layer, to order $N$ 
in the external perturbation. 
%The matrix elements of the 
%density operator for $N=2$ is given by Eq.~\eqref{j.2}.
From
Eqs.~\eqref{vcal} and \eqref{jmic3} we obtain
\begin{align}\label{intj}
{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}({\mathbf k})
&=
\frac{1}{2}
\int \mathrm{d}^3 r\,
 \calc^\ell(z)
\bigg[
\langle m\mathbf{k}|\mathbf{v}^\gs | \mathbf{r}\rangle
\langle \mathbf{r} | n \mathbf k \rangle +
\langle m\mathbf{k} | \mathbf{r}\rangle
\langle \mathbf{r} | \mathbf{v}^\gs | n \mathbf k \rangle\bigg]
\nonumber\\
&=
\frac{1}{2}
\int \mathrm{d}^3 r\,
 \calc^\ell(z)
 \bigg[
\psi_{n\mathbf{k}}(\mathbf{r})
\bfv^{\gs *}\psi^*_{m\mathbf{k}}(\mathbf{r})
+ 
\psi^*_{m\mathbf{k}}(\mathbf{r})\bfv^\gs
\psi_{n\mathbf{k}}(\mathbf{r})
\bigg]
\nonumber\\
&=
\int \mathrm{d}^3 r\,
\psi^*_{m\mathbf{k}}(\mathbf{r})
\left[\frac{\calc^\ell(z) \bfv^\gs +
\bfv^\gs \calc^\ell(z)}{2}\right]
\psi_{n\mathbf{k}}(\mathbf{r})
\nonumber\\
&=
\int \mathrm{d}^3 r\,
\psi^*_{m\mathbf{k}}(\mathbf{r})
\calbv^{\gs,\ell}
\psi_{n\mathbf{k}}(\mathbf{r})
,
\end{align}
%Here an integration by parts is performed on the third term of the
%right-hand; since the $e^{-i\bfk\cdot\bfr}\psi_{n\bfk}(\bfr)$
%are periodic over the unit cell, the surface term vanishes. 
where, we used the hermitian property of $\bfv^\gs$ and defined
\begin{align}\label{nl.4}
\calbv^{\gs,\ell}
=
\frac{\calc^\ell(z) \bfv^\gs +
\bfv^\gs \calc^\ell(z)}{2}
,
\end{align} 
where the superscript $\ell$ is inherited from $\calc^\ell(z)$.
% and we
%supress the dependance on $z$ from the increasingly crowded notation.  
We see that the replacement
\begin{align}\label{vcali}
\bfV \to \calbv^{\ell}=\frac{\calc^\ell(z) \bfV +
\bfV \calc^\ell(z)}{2}
,
\end{align} 
is all that is needed to change any of the
velocity operators of the electron $\bfV$ to the new velocity
operator $\calbv^{\ell}$ that implicitly takes into account the
contribution of the region of the slab given by $\calc^\ell(z)$.
The operator $\bfV$ could be any of those given by Eq.~\eqref{vop2},
thus
\begin{align}\label{vopii}
\calbv^{\gs,\ell}
&=
\calbv^{\lda,\ell}
+
\calbv^{S,\ell}
\nonumber\\
\calbv^{\lda,\ell}
&=
\calbv^{\ell}
+
\calbv^{\nl,\ell}
=
\frac{1}{m_e}
\calbp^{\ell}
+
\calbv^{\nl,\ell}
.
\end{align}
We remark that the simple renormalization that gives 
$\bfv^{\sigma}_{nm}(\bfk)$ 
in terms of
$\bfv^{\lda}_{nm}(\bfk)$,
given in 
Eq.~\eqref{chon.9}, 
does not hold between
$\calbv^{\sigma,\ell}_{nm}(\bfk)$   
and 
$\calbv^{\lda,\ell}_{nm}(\bfk)$,
i.e.
$\calbv^{\sigma,\ell}_{nm}(\bfk)\ne
(\go^\gs_{nm}/\go^\lda_{nm})
\calbv^{\lda,\ell}_{nm}(\bfk)$,
and thus, to calculate
$\calbv^{\sigma,\ell}_{nm}(\bfk)$ 
we must calculate the matrix elements of 
$\calbv^{\lda,\ell}$ and $\calbv^{S,\ell}$
 (separately)
according to the expressions of
Appendices \ref{calpcalc}, \ref{vesnl} and \ref{calvs}.

To limit the SHG response to one surface, the equivalent of Eq.~\eqref{nl.4} 
for $\calbv^\ell=\calbp^\ell/m_e$ was proposed in 
Ref.~\onlinecite{reining_microscopic_1994} and later used in Refs.
\onlinecite{mendozaPRL98},
\onlinecite{mendoza_ab_2001},
\onlinecite{sanoPRB02},
 and \onlinecite{mejia_layer-by-layer_2004}. 
The layer-by-layer analysis of Refs. \onlinecite{hogan_optical_2003,castilloPRB03,mottaCMS10} 
used the equivalent Eq.~\eqref{sz}, 
%limiting the current response
%to a particular layer of the slab and used 
to obtain the
anisotropic linear optical response of semiconductor surfaces.
However, the first formal derivation
of this scheme 
for the linear response 
is presented in
Ref.~\onlinecite{mendozaPRB06}, 
and here in this 
article, for the second harmonic optical response of semiconductors.

Using
$\bfJ=d\bfP/dt$ 
and Eq.~\eqref{jmac2} 
we obtain the SH polarization of the $\ell$-th layer as
\begin{equation}\label{Pjikn}
\bfP^{(2),\ell}(2\go)
=\frac{ie}{2\gO\got}
\int \frac{dk^3}{8\pi^3}
\sum_{mn}
\calbv^{\gs,\ell}_{mn}(\mathbf{k})
\rho^{(2)}_{nm}(\bfk;2\go)
,
\end{equation}
and using Eqs.~\eqref{pshg} and \eqref{j.2} 
leads to
\begin{align}\label{Pjikn2}
\chi^{\ell,\rma\rmb\rmc}(-2\go;\go,\go) 
&=
\frac{e^2}{2A\hbar\got}
\int \frac{dk^3}{8\pi^3}
\sum_{mn}
\frac{\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})}
{\go^\gs_{nm\bfk}-2\got}
\bigg[
-(B_{nm}^{\rmc}(\bfk,\go))_{;k^{\rmb}}
\nonumber \\
&
+i\sum_\ell\left(r_{n\ell}^{\rmb}B_{\ell m}^{\rmc}(\bfk,\go) -
  B_{n\ell}^{\rmc}(\bfk,\go) 
  r_{\ell m}^{\rmb}\right)
\bigg]
,
\end{align}
which gives the susceptibility 
$\chi^{\ell}_{\rma\rmb\rmc}(-2\go;\go,\go)$ 
of the $\ell$-th layer of the slab, 
%where 
%$\calbv^\gs$ is given in Eq.~\eqref{vopii},
where $A=\gO/d$ is the area of the unit
cell that characterizes the surface of the system, and $d$
is the surface region, from which the non-linear susceptibility is different from zero.
We mention that the units of 
$\chi^{\ell,\rma\rmb\rmc}(-2\go;\go,\go)$
are m$^2$/V, as they should for a surface SH susceptibility.
Using Eq.~\eqref{j.1} we
split this equation into
two contributions from the first and second terms on the right hand side
of Eq.~\eqref{Pjikn2}, then,
\begin{equation}\label{chii}
\chi^{\ell,\rma\rmb\rmc}_i (-2\go;\go,\go)
=-\frac{e^3}{A\hbar^22\got}
\int \frac{dk^3}{8\pi^3}
\sum_{mn}
\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}
\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}}
,
\end{equation} 
that  is related to intraband transitions, and 
\begin{equation}\label{chie}
\chi^{\ell,\rma\rmb\rmc}_e (-2\go;\go,\go)
=\frac{ie^3}{A\hbar^22\got}
\int \frac{dk^3}{8\pi^3}
\sum_{\ell mn}
\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}
\left(
\frac{r_{n\ell}^{\rmc} r_{\ell m}^{\rmb} 
f_{m\ell}}{\go^\gs_{\ell m}-\got}
-\frac{r_{n\ell}^{\rmb} r_{\ell m}^{\rmc} 
f_{\ell n}}{\go^\gs_{n \ell}-\got}
\right),
\end{equation} 
that is related to interband transitions.
The generalized derivative in Eq.~\eqref{chii} is dealt with by using the chain rule 
\begin{equation}\label{gene2}
\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}}=
\frac{f_{mn}}{\go^\gs_{nm}-\got}\left(r_{nm}^\rmb\right)_{;k^{\rmc}}
-\frac{f_{mn}r_{nm}^{\rmb}\gD_{nm}^\rmc}{(\go^\gs_{nm}-\got)^2}
,
\end{equation}
where by using $H^\sigma_0$ into Eq.~\eqref{conmri3n} and then
Eq.~\eqref{chon.9}
 one can readily
show that
\begin{align}\label{eli.13}
\left(\go^\gs_{nm}\right)_{;k^{\rma}}
=
\left(\go^\lda_{nm}\right)_{;k^{\rma}}
= 
v_{nn}^{\lda,\rma}-v_{mm}^{\lda,\rma}\equiv\gD_{nm}^{\rma}
.
\end{align} 
The apparent divergence as $\got\to 0$
in Eqs. \eqref{chii} and \eqref{chie},  
is removed  by
 a partial fraction expansion over $\got$. 
Using time-reversal invariance, an integration by parts to 
remove the square in the denominator of the second term in Eq.~\eqref{gene2}, 
and taking the limit of $\eta\to 0$, 
we obtain the following expressions for the imaginary parts of 
Eqs. \eqref{chii} and \eqref{chie}, 
\input{appendices/chis}
\noindent where we have splited the interband and intraband $1\go$ and $2\go$
contributions and supres the $\go$ arguments for space sake.
 The real part of each contribution can be obtained through
a Kramers-Kronig transformation,\cite{nicolas} and then
$\chi^{\ell,\rma\rmb\rmc}=
\chi^{\ell,\rma\rmb\rmc}_{e,\go} 
+\chi^{\ell,\rma\rmb\rmc}_{e,2\go}
+\chi^{\ell,\rma\rmb\rmc}_{i,\go} 
+\chi^{\ell,\rma\rmb\rmc}_{i,2\go}
$.
To fulfill the required intrinsic permutation symmetry, %\cite{rashkeevPRB98} 
the $\{\}$ notation symmetrizes the $\rmb\rmc$ Cartesian indices, i.e. 
$\{u^{\rmb}s^{\rmc}\}=(u^{\rmb}s^{\rmc}+u^{\rmc}s^{\rmb})/2$,
and thus
$\chi^{\ell,\rma\rmb\rmc}=\chi^{\ell,\rma\rmc\rmb}$.
We calculate the nonlinear surface susceptibility as 
\begin{equation}\label{chiijksur}
\bfgchi(-2\go;\go,\go)
= \sum_{\{\ell\}}\bfgchi^{\ell}(-2\go;\go,\go),
\end{equation} 
where $\{\ell\}$
is meant to be set of layers that one choses. For instance, one go
from 
a layer that encompasses half of the slab  or each atomic layer till
the middle of the slab. In the first case, there is one single summand
in Eq.~\eqref{chiijksur}, in the second case a sum from
$\ell=1$ which denotes
the first layer right at the surface, 
till $\ell=N$ that corresponds to the layer at the middle of the salb, which ought to be
a bulk-like layer.
%that at a distance $d$ from the surface  as seen in Fig.~\ref{fslab}.  
%For a centrosymmetric slab, 
%$\bfgchi^{\ell=N}=0$;
We reamark that the value of 
$N$ is not universal, and  
the slab needs to have enough atomic layers for 
%$\bfgchi^{\ell=N}=0$;
%to be satisfied and to 
in order to give converged results for $\bfgchi (-2\go;\go,\go)$. 
We can use Eq.~\eqref{chiijksur} for 
either the front or the back surface. 
\textcolor{red}{Talk about the correct value of $d$}

\section{Results}

%%%%%%%%%%%%%
\appendix 
\section{}
We give explicit expressions for the quantities used in the evaluation 
of Eq.~\eqref{calvimchiewn}-\eqref{calvimchi2wn}; when appropriate, some 
intermediate steps are given for their derivation. 
\input{appendices/appendix-vc}%calpcalc G

\input{appendices/appendix-gen_der_r_nonlocal}%gdernl H

\input{appendices/appendix-vnl}%gdernl H

\input{appendices/appendix-tnm}%calt F

\input{appendices/appendix-gen_der_calv}%calvs C

%%%%%%%%%%%%%%
\bibliography{ref}% 
\end{document}  %
%%%%%%%%%%%%%%

\section{Introduction}\label{intro}

Second harmonic generation (SHG) is a powerful spectroscopic tool for 
studying the optical properties of surfaces and interfaces since it has 
the advantage of being surface sensitive. Within the dipole approximation, 
inversion symmetry forbids SHG from the bulk of controsymmetric materials. 
SHG is allowed at the surface of these materials where the inversion symmetry 
is broken and should necessarily come from the localized surface region. 
SHG allows the study of the structural atomic arrangement and phase 
transitions of clean and adsorbate covered surfaces. Since it is also an 
optical probe it can be used out of UHV conditions and is non-invasive 
and non-destructive. Experimentally, new tunable high intensity laser systems 
have made SHG spectroscopy readily accessible and applicable to a wide range 
of systems.\cite{downer_optical_2001,lupke_characterization_1999}

However, theoretical development of the field is still an ongoing 
subject of research. Some recent advances for the cases of semiconducting 
and metallic systems have appeared in the literature, where the use of 
theoretical models with experimental results have yielded correct 
physical interpretations for observed SHG spectra.
\cite{
downer_optical_2001,
mendoza_ab_2001,
lim_optical_2000,
gavrilenko_optical_2000,
mendoza_visible-infrared_1999,
mendozaPRL98,
mendozaPRB96,
mendoza_polarizable-bond_1997,
guyotPRB90}

In a previous article\cite{mendoza_epioptics_2001} we reviewed some 
of the recent results in the study of SHG using the velocity gauge 
for the coupling between the electromagnetic field and the electron. 
In particular, we demonstrated a method to systematically analyze the 
different contributions to the observed SHG peaks.\cite{arzatePRB01} 
This approach consists of separating the different contributions to 
the nonlinear susceptibility according to 1$\omega$ and 2$\omega$ 
transitions, and the surface or bulk nature of the states among 
which the transitions take place.

To compliment those results, in this article we review the calculation 
of the nonlinear susceptibility using the longitudinal gauge. We show 
that it is posible to clearly obtain the ``layer-by-layer'' contribution 
for a slab scheme used for surface calculations.

\section{Non-linear Surface Susceptibility}\label{nonchi}

In this section we outline the general procedure to obtain the 
surface susceptibility tensor for second harmonic generation.
We start with the 
non-linear polarization $\mbf{P}$ written as
\begin{align}\label{mshg}
P_{\rma}(2\go)&=\chi_{\rma\rmb\rmc}(-2\go;\go,\go)
E_{\rmb}(\go)
E_{\rmc}(\go)
\nonumber \\
&+
\chi_{\rma\rmb\rmc\rml}(-2\go;\go,\go)
E_{\rmb}(\go)\nabla_{\rmc} E_{\rml}(\go)
+\cdots,
\end{align}
where $\chi_{\rma\rmb\rmc}(-2\go;\go,\go)$ and 
$\chi_{\rma\rmb\rmc\rml}(-2\go;\go,\go)$
correspond to the dipolar and quadrupolar susceptibilities.
We drop  the $(-2\go;\go,\go)$ argument to ease on the notation.
The sum continues with higher multipolar terms.
If we consider a semi-infinite system with a centrosymmetric bulk, the
equation above can be separated into two contributions from symmetry 
considerations alone; one from the surface of the system and the other from
the bulk of the system. We take
\begin{equation}\label{mshg2}
P_{\rma}(\mbf{r})=\chi_{\rma\rmb\rmc}E_{\rmb}(\mbf{r})E_{\rmc}(\mbf{r})
+
\chi_{\rma\rmb\rmc\rml}E_{\rmb}(\mbf{r})\frac{\partial}{\partial
  \mbf{r}_{\rmc}} E_{\rml}(\mbf{r}) 
+\cdots,
\end{equation}
as the polarization with respect to the original coordinate system, and 
\begin{align}\label{mshg3}
P_{\rma}(-\mbf{r})&=\chi_{\rma\rmb\rmc}E_{\rmb}(-\mbf{r})E_{\rmc}(-\mbf{r})
\nonumber \\
&+
\chi_{\rma\rmb\rmc\rml}E_{\rmb}(-\mbf{r})\frac{\partial}{\partial(-
  \mbf{r}_{\rmc})} E_{\rml}(-\mbf{r}) 
+\cdots, 
\end{align}
as the polarization in the coordinate system where inversion is taken,
i.e. $\mbf{r} \to -\mbf{r}$. 
Note that we have kept the same susceptibility tensors, and they must be 
invariant under $\mbf{r} \to -\mbf{r}$ since the system is centrosymmetric.
Recalling that $\mbf{P}(\mbf{r})$ and $\mbf{E}(\mbf{r})$ are polar vectors 
\cite{jackson_classical_1975}, we have that Eq.~\eqref{mshg3} reduces to
\begin{align}\label{mshg4}
-P_{\rma}(\mbf{r})&=\chi_{\rma\rmb\rmc}(-E_{\rmb}(\mbf{r}))(-E_{\rmc}(\mbf{r}))
-
\chi_{\rma\rmb\rmc\rml}(-E_{\rmb}(\mbf{r}))(-\frac{\partial}{\partial\mbf{r}_{\rmc}})(-
E_{\rml}(\mbf{r})) 
+\cdots,
\nonumber \\
P_{\rma}(\mbf{r})&=-\chi_{\rma\rmb\rmc}E_{\rmb}(\mbf{r})E_{\rmc}(\mbf{r})
+
\chi_{\rma\rmb\rmc\rml}E_{\rmb}(\mbf{r})\frac{\partial}{\partial\mbf{r}_{\rmc}}
E_{\rml}(\mbf{r}) 
+\cdots,
\end{align}
that when compared with Eq.~\eqref{mshg2} leads to the conclusion that
\begin{equation}\label{sshg}
\chi_{\rma\rmb\rmc}=0
\end{equation}
for a centrosymmetric bulk.

\begin{figure}[t]
\centering
\includegraphics[scale=0.6]{images/system}
\caption{(Color Online) Sketch of the semi-infinite system with a 
centrosymmetric bulk. The surface region is of width $\sim d$. 
The incoming photon of frequency $\go$ is represented by
a downward red arrow, whereas both the surface and bulk created second
harmonic photons of frequency $2\go$ are represented by upward
green arrows. The red color suggests an incoming infrared photon with
a green second harmonic photon. The dipolar ($\chi_{\rma\rmb\rmc}$), and 
quadrupolar ($\chi_{\rma\rmb\rmc\rml}$) susceptibility tensors are shown 
in the regions where they are different from zero. The axis has $z$
perpendicular to the surface and $\mbf{R}$ parallel to it.\label{fsystem}}
\end{figure}

If we move to the surface of the semi-infinite system our assumption 
of centrosymmetry breaks down, and there is no restriction in 
$\chi_{\rma\rmb\rmc}$. We conclude that the leading term
of the polarization in a surface region is given by
\begin{align}\label{sshgp1}
\int dz 
P_{\rma}(\mbf{R},z)
\approx dP_{\rma}
\equiv
P_{\rma}^S
\equiv
\chi^S_{\rma\rmb\rmc}E_{\rmb}E_{\rmc},
\end{align}
where
% $\mbf{R}$ is a vector parallel to the surface which is
% perpendicular to $z$, $S$ is the surface area of the unit
% cell that characterizes the surface of the system, and 
$d$ is the
surface region from which the dipolar signal of $\bfP$ is
different from zero (see Fig.~\ref{fsystem}), and
$\mbf{P}^S\equiv d\mbf{P}$
 is the surface SH polarization. Then, from Eq.~\eqref{mshg} we obtain that
\begin{align}\label{sshgp2}
\chi_{\rma\rmb\rmc}^S=d\chi_{\rma\rmb\rmc}
\end{align}
is the
SH surface susceptibility. 
On the other hand, 
\begin{equation}\label{sshgp3}
P^b_{\rma}(\mbf{r})=\chi_{\rma\rmb\rmc\rml}E_{\rmb}(\mbf{r})\nabla_{\rmc}
E_{\rml}(\mbf{r}),  
\end{equation}
gives the bulk polarization. We immediately recognize that the surface
polarization is of dipolar order while the bulk polarization is of
quadrupolar order. The surface, $\chi^S_{\rma\rmb\rmc}$, and bulk, 
$\chi_{\rma\rmb\rmc\rml}$, susceptibility tensor ranks are three and 
four, respectively. We will only concentrate on surface SHG in this article 
even though bulk generated SH is also a very important optical phenomenon. 
Also, we leave out of this article other interesting surface SH phenomena like, electric field
induced second harmonic (EFISH), which would be represented by a
surface susceptibility  tensor of quadrupolar origin.
In centrosymmetric systems for which the quadrupolar bulk
response is much smaller than the dipolar surface response, SH is
readily used as a very useful and powerful 
optical surface probe.\cite{downer_optical_2001}

In the following sections we present the theoretical approach to
derive
the expressions for the 
surface susceptibility tensor $\chi^S_{\rma\rmb\rmc}$.

\section{Length Gauge}\label{longi}

We follow the article by Aversa and Sipe\cite{aversaPRB95} to calculate 
the optical properties of a given system within the longitudinal gauge. 
More recent derivations
can also be found in Refs. \onlinecite{sipe_second-order_2000,lambrechtPSSB00}.  
Assuming the long-wavelength approximation which implies 
a position independent electric field, 
$\bfE(t)$,  
the Hamiltonian in the length gauge approximation is given by
\begin{equation}\label{ache}
{H}={H}^\gs_0-e \bfr \cdot \bfE,
\end{equation}
with
\begin{align}\label{ache.1}
 H^\gs_0= H^\lda_0
 + S(\bfr,\bfp)
,
\end{align} 
as the unperturbed Hamiltonian.
The LDA Hamiltonian can be expressed as follows,
\begin{align}\label{ache.2}
 H^\lda_0&=\frac{ p^2}{2m_e}  +  V^\ps
\nonumber\\
 V^\ps
&= V^l(\bfr) +  V^\nl
,
\end{align}  
where $ V^l(\bfr)$ and $ V^\nl$ are the local and
the non-local parts of the  
crystal 
pseudopotential $ V^\ps$. For the latter, we have that
\begin{align}\label{ache.3n}
 V^\nl(\bfr,\bfr') 
\equiv \bra{\bfr} V^\nl\ket{\bfr'}\neq 0 \quad\mathrm{for}\quad\bfr\neq\bfr' 
,
\end{align}
where $V^\nl(\bfr,\bfr')$ is a function of $\bfr$ and $\bfr'$
representing the non-local contribution of the pseudopotential.
The 
Schr\"odinger equation reads  
\begin{align}\label{ache.4} 
\left(
\frac{-\hbar^2}{2m_e}\nabla^2  
 +  V^l(\bfr)\right)\psi_{n\bfk}(\bfr)  
 + \int d\bfr' V^\nl(\bfr,\bfr')\psi_{n\bfk}(\bfr')  
=E_i\psi_{n\bfk}(\bfr)  
,
\end{align}  
where
%$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=\sqrt{\gO/8\pi^3}e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
is cell periodic. 
$m_e$ is the bare mass of the electron and $\gO$ is the 
unit cell volume. The nonlocal scissors operator is given by 
\begin{equation}\label{chon.0}
S(\bfr,\bfp)=\hbar\gS\sum_n\int d^3k'(1-f_n(\bfk)) \ket{n\bfk'}\bra{n\bfk'}
,
\end{equation}
where $f_n(\bfk)$ is the occupation number, that for $T=0$ K, is
independent of $\bfk$, and is one for filled bands and zero for
unoccupied bands. For semiconductors the filled bands correspond to
valence
bands  ($n=v$) and the unoccupied bands to conduction bands
 ($n=c$). 
We have that
\begin{align}\label{chon.1}  
H^\lda_0\ket{n\bfk}&=\hbar\go^\lda_n(\bfk)\ket{n\bfk}
\nonumber\\
H^\gs_0\ket{n\bfk}&=\hbar\go^\gs_n(\bfk)\ket{n\bfk}
,
\end{align} 
where 
\begin{align}\label{chon.78}
\hbar\go^\gs_n(\bfk)=\hbar\go^\lda_n(\bfk)+\hbar\gS(1-f_n)
,
\end{align}
is the
 scissored
energy. Here, 
$\hbar\gS$ is the value by which the conduction bands are rigidly
($\bfk$-independent)
 shifted
upwards in energy, also known as the scissors shift. $\gS$ could be taken to be
$\bfk$ dependent, but for most calculations (like the ones presented
here), a rigid shift is sufficient. We can take 
$\hbar\gS=E_g-E_g^\lda$ where $E_g$ could be the experimental band gap 
or GW band gap taken at the $\Gamma$ point, i.e. $\bfk=0$.
We used the fact that 
$\ket{n\bfk}^\lda\approx\ket{n\bfk}^\gs$, thus negating the need to label
the Bloch states with the LDA or $\gs$ superscripts. 
The matrix elements of $\bfr$ are split between the {\it intraband} 
($\bfr_i$) and {\it interband} ($\bfr_e$) parts , where 
$\bfr=\bfr_i+\bfr_e$ and\cite{adamsJCP53,blountSSP62,aversaPRB95}
\begin{align}\label{rnminn}
\bra{n\bfk}\bfr_i\ket{m\bfk'} &= \gd_{nm}\left[\gd(\bfk-\bfk')\bfgxi_{nn}(\bfk)  
+i\nabla_\bfk\gd(\bfk-\bfk')\right], \\
\bra{n\bfk}\bfr_e\ket{m\bfk'} &=
(1-\gd_{nm})\gd(\bfk-\bfk')\bfgxi_{nm}(\bfk)  
,
\label{rnmenn}
\end{align} 
and
\begin{equation}\label{zetann}
\bfgxi_{nm}(\bfk) \equiv i\frac{(2\pi)^3}{\gO}\int_{\gO} d\bfr\, u^*_{n\bfk}(\bfr)\nabla_{\bfk}u_{m\bfk}(\bfr).
\end{equation}
The interband part $\bfr_e$ can
be obtained as follows. 
We start by introducing the 
velocity operator 
\begin{align}\label{vop}
\bfv^\gs&=
\frac{1}{i\hbar}[\bfr,{H}^\gs_0]
,
\end{align}
and calculating its matrix elements 
\begin{equation}\label{conhrnm}
i\hbar\bra{n\bfk}
\bfv^\gs 
\ket{m\bfk}
=
\bra{n\bfk}
[\bfr,{H}^\gs_0]
\ket{m\bfk}
=\bra{n\bfk}\bfr{H}^\gs_0-{H}^\gs_0\bfr\ket{m\bfk}
=(\hbar\go^\gs_m(\bfk)-\hbar\go^\gs_n(\bfk))\bra{n\bfk}\bfr\ket{m\bfk}
,
\end{equation}
thus defining $\go^\gs_{nm}(\bfk)=\go^\gs_n(\bfk)-\go^\gs_m(\bfk)$ we get 
\begin{align}\label{pmnrmn}
\bfr_{nm}(\bfk) 
&=
\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
\quad\quad n\notin D_m 
,
\end{align} 
which can be identified as 
$\bfr_{nm}=(1-\gd_{nm})\bfgxi_{nm}\to \bfr_{e,nm}$. 
Here, $D_m$ are all the possible degenerate $m$-states. 
When $\bfr_i$ appears in 
commutators we use\cite{aversaPRB95} 
\begin{equation}\label{conmri3n}
\bra{n\bfk}[\bfr_i,\calo]\ket{m\bfk'}
=i\gd(\bfk-\bfk')(\calo_{nm})_{;\bfk}
,
\end{equation}  
with 
\begin{equation}\label{gendevnn}
(\calo_{nm})_{;\bfk}=
\nabla_\bfk 
\calo_{n m}(\bfk) 
- 
i 
\calo_{nm}(\bfk) 
\left(
\bfgxi_{nn}(\bfk) 
-
\bfgxi_{mm}(\bfk) 
\right) 
,
\end{equation} 
where ``$;\bfk$'' denotes the generalized derivative (see Appendix \ref{reri}).

As can be seen from Eq.~\eqref{ache.1} and \eqref{ache.2},
both $ S$ and $ V^\nl$ are nonlocal potentials. Their contribution 
in the calculation of the optical response has to be taken in order to
get reliable results.\cite{ismailPRL01}
We proceed as follows; from Eqs.~\eqref{vop}, \eqref{ache.1} and
\eqref{ache.2}
 we find
\begin{align}\label{vop2}
\bfv^\gs&=
\frac{\bfp}{m_e}
+
\frac{1}{i\hbar}[\bfr, V^\nl(\bfr,\bfr')]
+
\frac{1}{i\hbar}[\bfr,S(\bfr,\bfp)]
\nonumber\\
&\equiv  
\bfv  
+
\bfv^\nl  
+\bfv^S  
=
\bfv^\lda  
+\bfv^S  
,
\end{align} 
where we have defined  
\begin{align}\label{ve}
\bfv  
&=\frac{\bfp}{m_e}
\\
\bfv^\nl  
&=
\frac{1}{i\hbar}[\bfr, V^\nl]
\\\label{vs}
\bfv^S  
&=
\frac{1}{i\hbar}[\bfr, S(\bfr,\bfp)]
\nonumber\\\label{vlda}
\bfv^\lda  
&=
\bfv  
+\bfv^\nl  
\end{align}   
with $\bfp=-i\hbar\bfgnabla$ the momentum operator.  
Using Eq.~\eqref{chon.0}, we obtain that the
matrix elements of $\bfv^S$ are given by
\begin{equation}\label{chon.2} 
\bfv^S_{nm}=i\gS f_{mn}\bfr_{nm},
%=\frac{\gS f_{mn}}{\go^\lda_{nm}}\bfv^\lda_{nm}
\end{equation}
with $f_{nm}=f_n-f_m$,
where we see that $\bfv^S_{nn}=0$, then 
\begin{align}\label{chon.8}
\bfv^\gs_{nm}
&=
\bfv^\lda_{nm}+i\gS f_{mn}\bfr_{nm}
\nonumber\\
&=
\bfv^\lda_{nm}+i\gS f_{mn}\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
\nonumber\\
\bfv^\gs_{nm}
\frac{\go^\gs_{nm}-\gS f_{mn}}{\go^\gs_{nm}}
&=
\bfv^\lda_{nm}
\nonumber\\
\bfv^\gs_{nm}
\frac{\go^\lda_{nm}}{\go^\gs_{nm}}
&=
\bfv^\lda_{nm}
\nonumber\\
\frac{\bfv^\gs_{nm}}{\go^\gs_{nm}}
&=
\frac{\bfv^\lda_{nm}}{\go^\lda_{nm}}
,
\end{align} 
since
$\go^\gs_{nm}-\gS f_{mn}=\go^\lda_{nm}$. Therefore, 
\begin{align}\label{chon.9}
\bfv^\gs_{nm}(\bfk)
&=
\frac{\go^\gs_{nm}}{\go^\lda_{nm}}\bfv^\lda_{nm}(\bfk)
=\left(1+\frac{\gS}{\go_c(\bfk)-\go_v(\bfk)}\right)\bfv^\lda_{nm}(\bfk)
\quad\quad n\notin D_m
\nonumber\\
\bfv^\gs_{nn}(\bfk)
&=
\bfv^\lda_{nn}(\bfk)
,
\end{align} 
and Eq.~\eqref{pmnrmn} gives
\begin{align}\label{chon.10}
\bfr_{nm}(\bfk) 
=
\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
=
\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
\quad\quad n\notin D_m 
. 
\end{align}
The matrix elements 
of $\bfr_e$ are the same whether we use the LDA or the scissored 
Hamiltonian and there is no need to label them with either LDA 
or S superscripts. Thus,
we can write 
\begin{equation}\label{chon.98}
\bfr_{e,nm}\to 
\bfr_{nm}(\bfk)=\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
\quad\quad n\notin D_m 
,
\end{equation}   
which gives the interband matrix elements of the position operator 
in terms of the matrix elements of $\bfv^\lda$. 
These matrix elements include the matrix elements of 
$\bfv^\nl_{nm}(\bfk)$ which can be readily calculated\cite{francesco} for 
fully separable nonlocal pseudopotentials in the 
Kleinman-Bylander 
form.\cite{mottaCMS10,kleinman_efficacious_1982,adolphPRB96}
In Appendix \ref{appvnl} we outline how this  
can be accomplished.

\section{Time-dependent Perturbation Theory}\label{tdpt}

In the independent particle approximation, we use the electron density
operator $\gr$ to obtain the expectation value of any observable
$\calo$ as
\begin{equation}\label{traza}
\calo=\mathrm{Tr}(\calo\rho)=\mathrm{Tr}(\rho\calo),
\end{equation}
where $\mathrm{Tr}$ is the trace and is invariant under cyclic permutations.
The dynamic equation of motion for $\gr$ is given by
\begin{equation}\label{eqrho}
i\hbar \frac{d\gr}{dt}=[{H},\gr]
,
\end{equation}
where it is more convenient to work in the interaction picture. We transform 
all operators according to 
\begin{equation}\label{ip}
{\calo}_I={U}{\calo}{U}^\dagger,
\end{equation}
where
\begin{equation}\label{ou}
{U}=e^{i{H}_0t/\hbar},
\end{equation}
is the unitary operator that shifts us to the interaction picture.
Note that $\calo_I$ depends on time even if $\calo$ does not.
Then, we transform Eq.~\eqref{eqrho} into
\begin{equation}\label{intrho}
i\hbar\frac{d\gr_I(t)}{dt}=[-e\bfr_I(t)\cdot\bfE(t),\gr_I(t)],
\end{equation}
that leads to
\begin{equation}\label{intrho2}
\gr_I(t)=\gr_I(t=-\infty)
+
\frac{ie}{\hbar}\int_{-\infty}^t dt'[\bfr_I(t')\cdot\bfE(t'),\gr_I(t')].
\end{equation}
We assume that the interaction is switched-on adiabatically and
choose a time-periodic perturbing field, to write
\begin{equation}\label{efield}
\bfE(t)=\bfE e^{-i\go t}e^{\eta t}=\bfE e^{-i\got t}
,
\end{equation}
with
\begin{equation}\label{got}
\got=\go+i\eta
,
\end{equation} 
where $\eta > 0$ assures
that at $t=-\infty$ the interaction is zero and has its full strength $\bfE$ 
at $t=0$. After computing the required time integrals one takes
$\eta\to 0$. 
Also, $\gr_I(t=-\infty)$ should be time independent and thus 
$[{H},\rho]_{t=-\infty}=0$, This implies that 
$\gr_I(t=-\infty)=\rho(t=-\infty)\equiv\rho_0$,
where $\rho_0$ is
the density matrix of the unperturbed ground state,
such that
\begin{equation}\label{nrhon}
\bra{n\bfk}\gr_0\ket{m\bfk'}=f_n(\hbar\go^\gs_n(\bfk))\gd_{nm}\gd(\bfk-\bfk')
,
\end{equation}
with $f_n(\hbar\go^\gs_n(\bfk))=f_{n\bfk}$ as the Fermi-Dirac distribution function.

We solve Eq.~\eqref{intrho2} using the standard iterative
solution, for which we write
\begin{equation}\label{rhop}
\rho_I=\rho_I^{(0)}+\rho_I^{(1)}+\rho_I^{(2)}+\cdots
,
\end{equation}
where $\rho_I^{(N)}$ is the density operator to order $N$ in $\bfE(t)$.
Then, Eq.~\eqref{intrho2} reads
\begin{equation}\label{intrho3}
\rho_I^{(0)}+\rho_I^{(1)}+\rho_I^{(2)}+\cdots
=\gr_0
+
\frac{ie}{\hbar}\int_{-\infty}^t dt'[\bfr_I(t')\cdot\bfE(t'),
\rho_I^{(0)}+\rho_I^{(1)}+\rho_I^{(2)}+\cdots
]
,
\end{equation}
where, by equating equal orders in the perturbation, we find
\begin{equation}\label{rho0}
\gr_I^{(0)}\equiv\gr_0
,
\end{equation}
and
\begin{equation}\label{rhoN}
\gr_I^{(N)}(t)=
\frac{ie}{\hbar}
\int_{-\infty}^t dt'[\bfr_I(t')\cdot\bfE(t'),\gr^{(N-1)}_I(t')].
\end{equation}
It is simple to show that matrix elements of Eq. (\ref{rhoN}) satisfy
$\bra{n\bfk}\gr_I^{(N+1)}(t)\ket{m\bfk'}=\gr^{(N+1)}_{I,nm}(\bfk)\gd(\bfk-\bfk')$,
with
\begin{equation}\label{rtilde}
\gr^{(N+1)}_{I,nm}(\bfk;t)
=\frac{ie}{\hbar}\int_{-\infty}^t dt'
\bra{n\bfk}
[\bfr_I(t'),\gr^{(N)}_I(t')]
\ket{m\bfk}
\cdot\bfE(t')
.
\end{equation}

We now work out the commutator of Eq.~\eqref{rtilde}. Then,
\begin{align}\label{conmu1}
\bra{n\bfk}
[\bfr_I(t),\gr^{(N)}_I(t)]
\ket{m\bfk}
&=
\bra{n\bfk}
[{U}\bfr{U}^\dagger,{U}\gr^{(N)}(t){U}^\dagger]
\ket{m\bfk}
\nonumber \\
&=
\bra{n\bfk}
{U}[\bfr,\gr^{(N)}(t)]{U}^\dagger
\ket{m\bfk}
\\
&=
e^{i\go^\gs_{nm\bfk}t}
\left(
\bra{n\bfk}
[\bfr_e,\gr^{(N)}(t)]
+
[\bfr_i,\gr^{(N)}(t)]
\ket{m\bfk}
\right)
\nonumber
.
\end{align}
% where the time dependence of operator's interaction picture is
% explicitly shown by the exponential factor, and the implicit
% dependence of $\rho^{(N)}$ inherited from Eq.~\eqref{eqrho} is
% shown by its $t$ argument.
We calculate the interband term first, so using Eq.~\eqref{chon.98} we obtain
\begin{align}\label{conmu2}
\bra{n\bfk}
[\bfr_e,\gr^{(N)}(t)]
\ket{m\bfk}
&=
\sum_{\ell}
\left(
\bra{n\bfk}
\bfr_e
\ket{\ell\bfk}
\bra{\ell\bfk}
\gr^{(N)}(t)
\ket{m\bfk}
\right.
\nonumber \\
&
\left.
-
\bra{n\bfk}
\gr^{(N)}(t)
\ket{\ell\bfk}
\bra{\ell\bfk}
\bfr_e
\ket{m\bfk}
\right)
\nonumber \\
&=
\sum_{\ell\ne n,m}
\left(
\bfr_{n\ell}(\bfk)
\gr^{(N)}_{\ell m}(\bfk;t)
-
\gr^{(N)}_{n\ell}(\bfk;t)
\bfr_{\ell m}(\bfk)
\right)
\nonumber\\
&\equiv
\bfR^{(N)}_e(\bfk;t)
,
\end{align}

and from Eq.~\eqref{conmri3n},
\begin{equation}\label{conmri4}
\bra{n\bfk}[\bfr_i,\rho^{(N)}(t)]\ket{m\bfk'}
=i \gd(\bfk-\bfk') (\rho^{(N)}_{nm}(t))_{;\bfk}
\equiv \gd(\bfk-\bfk')\bfR_i^{(N)}(\bfk;t)
.
\end{equation}
Then Eq.~\eqref{rtilde} becomes
\begin{equation}\label{rtilde2}
\gr^{(N+1)}_{I,nm}(\bfk;t)
=\frac{ie}{\hbar}\int_{-\infty}^t dt'
e^{i(\go^\gs_{nm\bfk}-\got)t'}
\left[R_e^{\rmb(N)}(\bfk;t')+R_i^{\rmb(N)}(\bfk;t')\right]E^{\rmb}
,
\end{equation}
where the roman superindices
$\rma,\rmb,\rmc$ denote Cartesian components that are summed over if repeated.
Starting from the linear response and proceeding from Eq.~\eqref{nrhon} and ~\eqref{conmu2},
\begin{align}\label{R0e}
R_e^{\rmb(0)}(\bfk;t)
&=
\sum_{\ell}
\left(
r^{\rmb}_{n\ell}(\bfk)
\gr^{(0)}_{\ell m}(\bfk)
-
\gr^{(0)}_{n\ell}(\bfk)
r^{\rmb}_{\ell m}(\bfk)
\right)
\nonumber \\
&=
\sum_{\ell}
\left(
r^{\rmb}_{n\ell}(\bfk)
\gd_{\ell m}f_m(\hbar\go^\gs_m(\bfk))
-
\gd_{n\ell}f_n(\hbar\go^\gs_n(\bfk))
r^{\rmb}_{\ell m}(\bfk)
\right)
\nonumber \\
&=
f_{mn\bfk}
r^{\rmb}_{nm}(\bfk)
,
\end{align}
where $f_{mn\bfk}=f_{m\bfk}-f_{n\bfk}$.
From now on,
it should be clear that the matrix elements of $\bfr_{nm}$ imply
$n\notin D_m$.
We also have from Eq.~\eqref{conmri4} and Eq.~\eqref{gendevnn} that
\begin{equation}\label{R0i}
R_i^{\rmb(0)}(\bfk)=i(\rho^{(0)}_{nm})_{;k^{\rmb}}=i\gd_{nm}(f_{n\bfk})_{;k^{\rmb}}=i\gd_{nm}\nabla_{k^{\rmb}} f_{n\bfk}
.
\end{equation}
For a semiconductor at $T=0$, $f_{n\bfk}$ is one if the state
$\ket{n\bfk}$ is a valence state and zero if it is a conduction state; 
thus $\nabla_\bfk f_{n\bfk}=0$ and $\bfR_i^{(0)}=0$ and
the linear response has no contribution from
intraband transitions.
 Then,
\begin{align}\label{rtilde2n}
\gr^{(1)}_{I,nm}(\bfk;t)
&=\frac{ie}{\hbar}
f_{mn\bfk}
r^{\rmb}_{nm}(\bfk)E^{\rmb}
\int_{-\infty}^t dt'
e^{i(\go^\gs_{nm\bfk}-\got)t'}
\nonumber \\
&=\frac{e}{\hbar}
f_{mn\bfk}
r^{\rmb}_{nm}(\bfk)E^{\rmb}
\frac{e^{i(\go^\gs_{nm\bfk}-\got)t}}
{\go^\gs_{nm\bfk}-\got}
\nonumber \\
&=
e^{i\go^\gs_{nm\bfk}t}
B^{\rmb}_{mn}(\bfk)E^{\rmb}(t)
\nonumber \\
&=
e^{i\go^\gs_{nm\bfk}t}
\rho^{(1)}_{nm}(\bfk;t)
,
\end{align}
with
\begin{equation}\label{rho1} 
B^{\rmb}_{nm}(\bfk,\go)=
\frac{e}{\hbar}\frac{f_{mn\bfk}r^{\rmb}_{nm}(\bfk)}{\go^\gs_{nm\bfk}-\got}
,
\end{equation} 
and
\begin{equation}\label{rhonoi1}
\rho^{(1)}_{nm}(\bfk;t)=B^{\rmb}_{mn}(\bfk,\go)E^{\rmb}(\go)e^{-i\got t}
.
\end{equation}

Now, we calculate the second-order response. Then, from Eq.~\eqref{conmu2}
\begin{align}\label{R1e}
R_e^{\rmb(1)}(\bfk;t)
&=
\sum_{\ell}
\left(
r^{\rmb}_{n\ell}(\bfk)
\gr^{(1)}_{\ell m}(\bfk;t)
-
\gr^{(1)}_{n\ell}(\bfk;t)
r^{\rmb}_{\ell m}(\bfk)
\right)
\nonumber \\
&=
\sum_{\ell}
\left(
r^{\rmb}_{n\ell}(\bfk)
B^{\rmc}_{\ell m}(\bfk,\go)
-
B^{\rmc}_{n\ell}(\bfk,\go)
r^{\rmb}_{\ell m}(\bfk)
\right)E^{\rmc}(t)
,
\end{align}
and from Eq.~\eqref{conmri4}
\begin{equation}\label{R1i}
R_i^{\rmb(1)}(\bfk;t)=
i(\rho^{(1)}_{nm}(t))_{;k^{\rmb}}=
iE^{\rmc}(t)(B^{\rmc}_{nm}(\bfk,\go))_{;k^{\rmb}}
.
\end{equation}

Using Eqs.~\eqref{R1e} and ~\eqref{R1i} in Eq. (\ref{rtilde2}),
we obtain
\begin{align}\label{rtilde33}
\gr^{(2)}_{I,nm}(\bfk;t)
&=
\frac{ie}{\hbar}
\bigg[
\sum_{\ell}
\Big(
r^{\rmb}_{n\ell}(\bfk)
B^{\rmc}_{\ell m}(\bfk,\go)
-
B^{\rmc}_{n\ell}(\bfk,\go)
r^{\rmb}_{\ell m}(\bfk)
\Big)
\nonumber \\
&+
i
(B^{\rmc}_{nm}(\bfk,\go))_{;k^{\rmb}}
\bigg]
E^{\rmb}_{\go}E_{\go}^{\rmc}
\int_{-\infty}^t dt'
e^{i(\go^\gs_{nm\bfk}-2\got)t'}
\nonumber \\
&=
\frac{e}{\hbar}
\bigg[
\sum_{\ell}
\Big(
r^{\rmb}_{n\ell}(\bfk)
B^{\rmc}_{\ell m}(\bfk,\go)
-
B^{\rmc}_{n\ell}(\bfk,\go)
r^{\rmb}_{\ell m}(\bfk)
\Big)
\nonumber \\
&+
i
(B^{\rmc}_{nm}(\bfk,\go))_{;k^{\rmb}}
\bigg]
E^{\rmb}_{\go}E_{\go}^{\rmc}
\frac{e^{i(\go^\gs_{nm\bfk}-2\got)t}}
{\go^\gs_{nm\bfk}-2\got}
\nonumber \\
&=
e^{i\go^\gs_{nm\bfk}t}
\rho_{nm}^{(2)}(\bfk;t)
.
\end{align}
Now, we write
$\rho_{nm}^{(2)}(\bfk;t)=\rho_{nm}^{(2)}(\bfk;2\go)e^{-i2\got t}$,
with
\begin{align}\label{rho2}
\gr_{nm}^{(2)}(\bfk;2\go)&=\frac{e}{i\hbar}\frac{1}{\go^\gs_{nm\bfk}-2\got}
\bigg[-(B_{nm}^{\rmc}(\bfk,\go)_{;k^{\rmb}}
\nonumber \\
&
+i\sum_\ell\Big(r_{n\ell}^{\rmb}B_{\ell m}^{\rmc}(\bfk,\go) - B_{n\ell}^{\rmc}(\bfk,\go)
  r_{\ell m}^{\rmb}\Big)
\bigg] 
E^{\rmb}(\go)E^{\rmc}(\go)
\end{align} 
where 
$B_{\ell m}^{\rma}(\bfk,\go)$ are given by Eq.~\eqref{rho1}. 
We remark that $\bfr_{nm}(\bfk)$  are
the same whether calculated with the LDA or the scissored Hamiltonian. 
We chose the former in this article.

\section{Layered Current Density}\label{cd}

In this section, we derive the expressions for the microscopic current
density of a given layer in the unit cell of the system.
The approach we use to study the surface of a semi-infinite
semiconductor crystal is as follows. Instead of using a
semi-infinite system, we replace it by a slab (see Fig.~\ref{fslab}).
The slab consists of a front and back surface, and in between these
two surfaces is the bulk of the system. 
In general the surface of a crystal reconstructs or relaxes as the atoms
move to find equilibrium positions. This is due to the fact that
the otherwise
balanced forces are disrupted when the surface atoms do not find their 
partner atoms that are now absent at the surface of the slab.

To take the reconstruction or relaxation into account, we take ``surface'' to mean
the true surface of the first layer of atoms, and
some of the atomic sub-layers adjacent to it.
Since the front and the back
surfaces of the slab are usually identical the total slab is
centrosymmetric. This implies that 
$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=0$, and thus we must
find a way to bypass this characteristic of a centrosymmetric slab 
in order to have a finite $\chi^s_{\rma\rmb\rmc}$ representative of the
surface. Even if the front and back surfaces of the slab 
are different, breaking the centrosymmetry and therefore giving an
overall $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}\ne 0$, we still
need a procedure to extract the front surface $\chi^f_{\rma\rmb\rmc}$
and the back surface $\chi^b_{\rma\rmb\rmc}$ from the non-linear
susceptibility 
$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=\chi^f_{\rma\rmb\rmc}-\chi^b_{\rma\rmb\rmc}$
 of the entire slab.

A convenient way to accomplish the separation of the SH signal of
either surface is to introduce a ``cut function'', $\calc(z)$, which is 
usually taken to be unity over one half of the slab and zero over 
the other half.\cite{reining_microscopic_1994}
 In this case $\calc(z)$ will give the contribution of the 
side of the slab for which $\calc(z)=1$. We can generalize this 
simple choice for $\calc(z)$ by a top-hat cut function
$\calc^\ell(z)$ that selects a given layer,
\begin{equation}
\label{sz}
\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{\rmb})
            \Theta(z_\ell-z+\Delta_\ell^f),
\end{equation}
where $\Theta$ is the Heaviside function. Here, $\Delta_\ell^{f/b}$
is the distance that the $\ell$-th layer extends towards the front
$(f)$ or back $(b)$ from its $z_\ell$ position. 
$\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
(see Fig.~\ref{fslab}).
\begin{figure}[b]
\centering
%\includegraphics[height=5cm,width=7cm]{slab}
\includegraphics[scale=.7]{images/slab}
\caption{A sketch of a slab where the circles represent atoms.\label{fslab}}
\end{figure}

Now, we show how this ``cut function'' $\calc^\ell(z)$ is introduced in
the calculation of $\chi_{\rma\rmb\rmc}$. 
The microscopic current density is given by
\begin{equation}\label{jmic}
\bfj(\bfr,t)=\mathrm{Tr}(\bfj(\bfr)\rho(t)),
\end{equation}
where the operator for the electron's current is
\begin{equation}\label{hatjmic}
\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gs\ket{\bfr}\bra{\bfr}
+\ket{\bfr}\bra{\bfr}\bfv^\gs\right), 
\end{equation}
where ${\mbf{v}}^\gs$ is the electron's velocity operator to be dealt
with below. We define
$\mu\equiv\ket{\bfr}\bra{\bfr}$ and use the cyclic invariance of
the trace to write
\begin{align}\label{jmic2}
\mathrm{Tr}(\bfj(\bfr)\rho(t)&=\mathrm{Tr}(\rho(t)\bfj(\bfr))=
\frac{e}{2}
\left(
\mathrm{Tr}(\rho\bfv^\gs\mu)
+
\mathrm{Tr}(\rho\mu\bfv^\gs)
\right)
\nonumber\\
&=
\frac{e}{2}
\sum_{n\bfk}
\left(
\bra{n\bfk}\rho\bfv^\gs\mu\ket{n\bfk}
+
\bra{n\bfk}\rho\mu\bfv^\gs\ket{n\bfk}
\right)
\nonumber\\
&=
\frac{e}{2}
\sum_{nm\bfk}
\bra{n\bfk}\rho\ket{m\bfk}
\left(
\bra{m\bfk}\bfv^\gs\ket{\bfr}\braket{\bfr}{n\bfk}
+
\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gs\ket{n\bfk}
\right)
\nonumber\\
\bfj(\bfr,t)&=
\sum_{nm\bfk}
\rho_{nm}(\bfk;t)\bfj_{mn}(\bfk;\bfr),
\end{align}
where
\begin{equation}\label{jmic3}
\bfj_{mn}(\bfk;\bfr)=
\frac{e}{2}
\left(
\bra{m\bfk}\bfv^\gs\ket{\bfr}\braket{\bfr}{n\bfk}
+
\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gs\ket{n\bfk}
\right),
\end{equation}
are the matrix elements of the microscopic current operator,
and we have used the fact that the matrix elements between states $\ket{n\bfk}$
are diagonal in $\bfk$, i.e. proportional to $\gd(\bfk-\bfk')$.

Integrating the microscopic current $\mbf{j}(\mbf{r},t)$ over
the entire slab gives the averaged microscopic current density. 
If we want the contribution from only one region of the unit cell 
towards the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
over the desired region. The contribution to the current density from the
$\ell$-th layer of the slab is given by
\begin{equation}\label{jsz}
\frac{1}{\Omega}\int d^3r\, \calc^\ell(z)\, \mathbf{j}(\mathbf{r},t)
 \equiv \mathbf{J}^{\ell}(t),
\end{equation}
where $\mathbf{J}^{\ell}(t)$ is the microscopic current in the
$\ell$-th layer.
Therefore we define
\begin{equation}\label{vcal}
e{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}(\mathbf{k})
\equiv
%\frac{1}{\Omega}
\int d^3r\, \calc^\ell(z)\,\bfj_{mn}({\bfk};\bfr),
\end{equation}
to write
\begin{equation}\label{jmac}
J_a^{(N,\ell)}(t)=\frac{e}{\gO}
\sum_{mn\bfk}
\mathcal{V}^{\gs,a,\ell}_{mn}(\mathbf{k})
\rho^{(N)}_{nm}(\bfk;t),
\end{equation}
as the induced microscopic current of the $\ell$-th layer, to order $N$ 
in the external perturbation. The matrix elements of the 
density operator for $N=1,2$ are given by Eqs.~\eqref{rho1} and
\eqref{rho2} respectively. 
The Fourier component of microscopic current of Eq.~\eqref{jmac} is given by
\begin{equation}\label{jmac2}
J_{\rma}^{(N,\ell)}(\go_3)=\frac{e}{\gO}
\sum_{mn\bfk}
\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})
\rho^{(N)}_{nm}(\bfk;\go_3)
.
\end{equation}
We proceed to give an explicit expression of
$\calbv^{\gs,\ell}_{mn}(\mathbf{k})$.
From
Eqs.~\eqref{vcal} and \eqref{jmic3} we obtain
\begin{equation}\label{intj}
{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}({\mathbf k})=
\frac{1}{2}
\int \mathrm{d}^3 r\,
 \calc^\ell(z)
\bigg[
\langle m\mathbf{k}|\mathbf{v}^\gs | \mathbf{r}\rangle
\langle \mathbf{r} | n \mathbf k \rangle +
\langle m\mathbf{k} | \mathbf{r}\rangle
\langle \mathbf{r} | \mathbf{v}^\gs | n \mathbf k \rangle\bigg]
,
\end{equation}  
and using the following property
\begin{equation}\label{nl.2}
\langle \mathbf{r} | \bfv^\gs(\bfr,\bfr')| n\mathbf{k} \rangle
=\int d^3 r'' \bra{\mbf{r}}\bfv^\gs(\bfr,\bfr')\ket{\mbf{r}''}
\braket{\mbf{r}''}{n\mbf{k}}
=\bfv^\gs(\bfr,\bfr'')
\int d^3 r'' \braket{\mbf{r}}{\mbf{r}''}
\braket{\mbf{r}''}{n\mbf{k}}
=\bfv^\gs(\bfr,\bfr')
\psi_{n\mathbf{k}}(\mathbf{r})
,
\end{equation}
that stems from the fact that the operator $\bfv^\gs(\bfr,\bfr')$ does not act on
$\bfr''$, we can write
\begin{align}\label{nl.3}
\calbv^{\gs,\ell}_{mn}({\mathbf k})
&=
\frac{1}{2}
\int \mathrm{d}^3 r\,
 \calc^\ell(z)
 \bigg[
\psi_{n\mathbf{k}}(\mathbf{r})
\bfv^{\gs *}\psi^*_{m\mathbf{k}}(\mathbf{r})
+ 
\psi^*_{m\mathbf{k}}(\mathbf{r})\bfv^\gs
\psi_{n\mathbf{k}}(\mathbf{r})
\bigg]
\nonumber\\
&=
\int \mathrm{d}^3 r\,
\psi^*_{m\mathbf{k}}(\mathbf{r})
\left[\frac{\calc^\ell(z) \bfv^\gs +
\bfv^\gs \calc^\ell(z)}{2}\right]
\psi_{n\mathbf{k}}(\mathbf{r})
\nonumber\\
&=
\int \mathrm{d}^3 r\,
\psi^*_{m\mathbf{k}}(\mathbf{r})
\calbv^{\gs,\ell}
\psi_{n\mathbf{k}}(\mathbf{r})
.
\end{align}
We used the hermitian property of $\bfv^\gs$ and defined
\begin{align}\label{nl.4}
\calbv^{\gs,\ell}
=
\frac{\calc^\ell(z) \bfv^\gs +
\bfv^\gs \calc^\ell(z)}{2}
,
\end{align} 
where the superscript $\ell$ is inherited from $\calc^\ell(z)$ and we
supress the dependance on $z$ from the increasingly crowded notation.  
We see that the replacement
\begin{align}\label{vcali}
\bfv^\gs \to \calbv^{\gs,\ell}=\left[\frac{\calc^\ell(z) \bfv^\gs +
\bfv^\gs \calc^\ell(z)}{2}\right]
,
\end{align} 
is all that is needed to change the
velocity operator of the electron $\bfv^\gs$ to the new velocity
operator $\calbv^{\gs,\ell}$ that implicitly takes into account the
contribution of the region of the slab given by $\calc^\ell(z)$.
From Eq.~\eqref{vop2},
\begin{align}\label{vopii}
\calbv^{\gs,\ell}
&=
\calbv^{\lda,\ell}
+
\calbv^{S,\ell}
\nonumber\\
\calbv^{\lda,\ell}
&=
\calbv^{\ell}
+
\calbv^{\nl,\ell}
=
\frac{1}{m_e}
\calbp^{\ell}
+
\calbv^{\nl,\ell}
.
\end{align}
We remark that the simple relationship between 
$\bfv^{\sigma}_{nm}(\bfk)$ 
and 
$\bfv^{\lda}_{nm}(\bfk)$,
given in 
Eq.~\eqref{chon.9}, 
does not hold between
$\calbv^{\sigma,\ell}_{nm}(\bfk)$   
and 
$\calbv^{\lda,\ell}_{nm}(\bfk)$,
i.e.
$\calbv^{\sigma,\ell}_{nm}(\bfk)\ne
(\go^\gs_{nm}/\go_{nm})
\calbv^{\lda,\ell}_{nm}(\bfk)$ 
and
$\calbv^{\sigma,\ell}_{nn}(\bfk)\ne
\calbv^{\lda,\ell}_{nn}(\bfk)$
,
and thus, to calculate
$\calbv^{\sigma,\ell}_{nm}(\bfk)$ 
we must calculate the matrix elements of $\calbv^{S,\ell}$ and
$\calbv^{\lda,\ell}$ (separately)
according to the expressions of
Appendix \ref{calvs}.
% Using Eq.~\eqref{chon.9}, the matrix elements of Eq.~\eqref{vcali}
% are given by
% \begin{align}\label{end.1}
% \calbv^{\sigma,\ell}_{nm}(\bfk)
% &=
% \frac{1}{2}
% \sum_q
% \Big(
% \calc^\ell_{nq}\bfv^\sigma_{qm}
% +
% \bfv^\sigma_{nq}\calc^\ell_{qm}
% \Big)
% \nonumber\\
% &=
% \frac{1}{2}
% \Big(
% \calc^\ell_{nn}\bfv^\sigma_{nm}
% +
% \bfv^\sigma_{nn}\calc^\ell_{nm}
% +
% \calc^\ell_{nm}\bfv^\sigma_{mm}
% +
% \bfv^\sigma_{nm}\calc^\ell_{mm}
% \Big)
% \nonumber\\
% &+
% \frac{1}{2}
% \sum_{q\neq(nm)}
% \Big(
% \frac{\go^\sigma_{qm}}{\go_{qm}}
% \calc^\ell_{nq}\bfv^\lda_{qm}
% +
% \frac{\go^\sigma_{nq}}{\go_{nq}}
% \bfv^\lda_{nq}\calc^\ell_{qm}
% \Big)
% ,
% \end{align}
% \begin{align}\label{end.2}
% \calbv^{\sigma,\ell}_{nn}(\bfk)
% &=
% \frac{1}{2}
% \Big(
% \calc^\ell_{nn}\bfv^\sigma_{nn}
% +
% \bfv^\sigma_{nn}\calc^\ell_{nn}
% \Big)
% \nonumber\\
% &+
% \frac{1}{2}
% \sum_{q\neq n}
% \Big(
% \frac{\go^\sigma_{qn}}{\go_{qn}}
% \calc^\ell_{nq}\bfv^\lda_{qn}
% +
% \frac{\go^\sigma_{nq}}{\go_{nq}}
% \bfv^\lda_{nq}\calc^\ell_{qn}
% \Big)
% \nonumber\\
% &=
% \calc^\ell_{nn}\bfv^\lda_{nn}
% \nonumber\\
% &+
% \frac{1}{2}
% \sum_{q\neq n}
% \frac{\go^\sigma_{qn}}{\go_{qn}}
% \Big(
% \calc^\ell_{nq}\bfv^\lda_{qn}
% +
% \bfv^\lda_{nq}\calc^\ell_{qn}
% \Big)
% ,
% \end{align}

To limit the response to one surface, the equivalent of Eq.~\eqref{nl.4} 
for $\calbv^\ell=\calbp^\ell/m_e$ was proposed in 
Ref.~\onlinecite{reining_microscopic_1994} and later used in Refs.
\onlinecite{mendozaPRL98},
\onlinecite{mendoza_ab_2001},
\onlinecite{sanoPRB02},
 and \onlinecite{mejia_layer-by-layer_2004} 
also in the context of SHG. 
The layer-by-layer analysis of Refs. \onlinecite{hogan_optical_2003} 
and \onlinecite{castilloPRB03} used Eq.~\eqref{sz}, 
limiting the current response
to a particular layer of the slab and used to obtain the
anisotropic linear optical response of semiconductor surfaces.
However, the first formal derivation of this scheme is presented in
Ref.~\onlinecite{mendozaPRB06} for the linear response, and here in this 
article, for the second harmonic optical response of semiconductors.

%??d
% debemos mencionar que v caligrafica con tijeras esta mal calculada
% para la respuesta lineal y la inyeccion de corriente en el
% tratamiento capa por capa. Tambien, habria que ver que pasa con la
% inyeccion de espin capa-por-capa
%??u

\section{Microscopic surface susceptibility}
In this section we obtain the expressions for the 
surface susceptibility tensor $\chi^S_{\rma\rmb\rmc}$.
We start with the basic relation $\bfJ=d\bfP/dt$ 
with $\bfJ$ the current calculated in Sec.~\ref{cd}. From Eq.~\eqref{jmac2} 
we obtain
\begin{equation}\label{Pjikn}
J_{\rma}^{(2,\ell)}(2\go)=-i2\got P_{\rma}(2\go)
=\frac{e}{\gO}
\sum_{mn\bfk}
\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})
\rho^{(2)}_{nm}(\bfk;2\go)
,
\end{equation}
and using Eqs.~\eqref{rho2} and \eqref{sshgp2} leads to
\begin{align}\label{Pjikn2}
\chi^{S,\ell}_{\rma\rmb\rmc}
&=
\frac{ie}{A E^{\rmb}_1E^{\rmc}_2 2\got}
\sum_{mn\bfk}
\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})
\rho^{(2)}_{nm}(\bfk;2\got)
\nonumber \\
&=
\frac{e^2}{A\hbar2\got}
\sum_{mn\bfk}
\frac{\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})}
{\go^\gs_{nm\bfk}-2\got}
\bigg[
-(B_{nm}^{\rmc}(\bfk,\go))_{;k^{\rmb}}
\nonumber \\
&
+i\sum_\ell\left(r_{n\ell}^{\rmb}B_{\ell m}^{\rmc}(\bfk,\go) -
  B_{n\ell}^{\rmc}(\bfk,\go) 
  r_{\ell m}^{\rmb}\right)
\bigg]
,
\end{align}
which gives the surface-like susceptibility of $\ell$-th layer, where 
$\calbv^\gs$ is given in Eq.~\eqref{vopii},
where $A=\gO/d$ is the surface area of the unit
cell that characterizes the surface of the system.
Using Eq.~\eqref{rho1} we
split this equation into
two contributions from the first and second terms on the right hand side,
\begin{equation}\label{chii}
\chi_{i,\rma\rmb\rmc}^{S,\ell}
=-\frac{e^3}{A\hbar^22\got}\sum_{mn\bfk}
\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}
\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}},
\end{equation} 
and 
\begin{equation}\label{chie}
\chi_{e,\rma\rmb\rmc}^{S,\ell}
=\frac{ie^3}{A\hbar^22\got}\sum_{\ell mn\bfk}
\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}
\left(
\frac{r_{n\ell}^{\rmc} r_{\ell m}^{\rmb} 
f_{m\ell}}{\go^\gs_{\ell m}-\got}
-\frac{r_{n\ell}^{\rmb} r_{\ell m}^{\rmc} 
f_{\ell n}}{\go^\gs_{n \ell}-\got}
\right),
\end{equation} 
where $\boldsymbol{\chi}^{S,\ell}_i$
 is related to intraband transitions and
$\boldsymbol{\chi}^{S,\ell}_e$
to interband transitions.
For the generalized derivative in Eq.~\eqref{chii} we use the chain rule 
\begin{equation}\label{gene2}
\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}}=
\frac{f_{mn}}{\go^\gs_{nm}-\got}\left(r_{nm}^\rmb\right)_{;k^{\rmc}}
-\frac{f_{mn}r_{nm}^{\rmb}\gD_{nm}^\rmc}{(\go^\gs_{nm}-\got)^2}
,
\end{equation}
and the following result
shown in Appendix \ref{gwk},
\begin{align}\label{eli.13}
\left(\go^\gs_{nm}\right)_{;k^{\rma}}
=
\left(\go^\lda_{nm}\right)_{;k^{\rma}}
= 
v_{nn}^{\lda,\rma}-v_{mm}^{\lda,\rma}\equiv\gD_{nm}^{\rma}
.
\end{align}

In order to calculate the nonlinear susceptibility of any given layer 
$\ell$ we simply add the above terms $\boldsymbol{\chi}^{S,\ell}=
\boldsymbol{\chi}_e^{S,\ell}+\boldsymbol{\chi}_i^{S,\ell}$ and 
then calculate the surface susceptibility as 
\begin{equation}\label{chiijksur}
\bfgchi^S\equiv \sum_{\ell=1}^{N}\bfgchi^{S,\ell},
\end{equation} 
where $\ell=1$ is the first layer right at the surface, 
and $\ell=N$ is the bulk-like layer (at a distance $\sim d$ from the
surface  as seen in
Fig.~\ref{fsystem}), such that 
\begin{equation}\label{chiijksur2}
\bfgchi^{S,\ell=N}=0,
\end{equation}
in accordance to Eq.~\eqref{sshg} valid for a centrosymmetric environment. 
We note that the value of
$N$ is not universal.
This means that the slab needs to have enough atomic layers for 
Eq.~\eqref{chiijksur2} 
to be satisfied and to give converged results for $\bfgchi^S$. 
We can use Eq.~\eqref{chiijksur} for
either the front or the back surface.

We can see from the prefactors of Eqs. \eqref{chii} and \eqref{chie} 
that they diverge as $\got\to 0$. To remove this apparent divergence of 
$\bfgchi^{S,\ell}$, we perform a partial fraction expansion over $\got$. 
As shown in Appendix \ref{appv}, we use time-reversal invariance to 
remove these divergences and obtain the following expressions for $\bfgchi^S$,
\input{appendices/chis}
\noindent where the limit of $\eta\to 0$ has been taken.
We have split the interband and intraband $1\go$ and $2\go$
contributions. The real part of each contribution can be obtained through
a Kramers-Kronig transformation,\cite{nicolas} and then
$\chi^{S,\ell}_{\rma\rmb\rmc}=\chi^{S,\ell}_{e,\rma\rmb\rmc,\go} 
+\chi^{S,\ell}_{e,\rma\rmb\rmc,2\go}+\chi^{S,\ell}_{i,\rma\rmb\rmc,\go}
+\chi^{S,\ell}_{i,\rma\rmb\rmc,2\go}
$.
To fulfill the required intrinsic permutation symmetry,\cite{rashkeevPRB98} 
the $\{\}$ notation symmetrizes the $\rmb\rmc$ Cartesian indices, i.e. 
$\{u^{\rmb}s^{\rmc}\}=(u^{\rmb}s^{\rmc}+u^{\rmc}s^{\rmb})/2$,
and thus
$\chi_{\rma\rmb\rmc}^{S,\ell}=\chi_{\rma\rmc\rmb}^{S,\ell}$.
In Appendices \ref{gdernl} and \ref{calvs} we demonstrate how to calculate  
the generalized derivatives of $\bfr_{nm;\bfk}$ and
$\calv^{\gs,\rma,\ell}_{nm;\bfk}$. 
We find that
\begin{align}\label{rgen.69}
(r^{\rmb}_{nm})_{;k^{\rma}}
&=
-i\calt^{\rma\rmb}_{nm}
+
\frac{
r^{\rma}_{nm}
\Delta^{\rmb}_{mn}
+r^{\rmb}_{nm}
\Delta^{\rma}_{mn}
}
{\go^\lda_{nm}}
+
\frac{i}{\go^\lda_{nm}}
\sum_{\ell}
\bigg(
\go^\lda_{\ell m}
r^{\rma}_{n\ell}
r^{\rmb}_{\ell m}
-
\go^\lda_{n\ell}
r^{\rmb}_{n\ell}
r^{\rma}_{\ell m}
\bigg)
,
\end{align}
where
\begin{align}\label{tau.1}
\calt_{nm}^{\rma\rmb}
=
[r^{\rma},v^{\lda,\rmb}]= 
\frac{i\hbar}{m_e}\gd_{ab}\gd_{nm}+
\call_{nm}^{\rma\rmb}
,
\end{align}  
and
\begin{align}\label{tau.2}
\call_{nm}^{\rma\rmb}
=
\frac{1}{i\hbar}[r^{\rma},v^{\nl,\rmb}]_{nm}
,
\end{align}
is the contribution to the generalized derivative of $\bfr_{nm}$
coming from the nonlocal part of the pseudopotential.
In Appendix \ref{calt} we calculate
$\call^{\rma\rmb}_{nm}$, that
is a term with very small numerical value but with a computational time 
at least an order of magnitude larger
than for all the other terms involved in the expressions for 
$\chi^{s,\ell}_{\mathrm{abc}}$.\cite{valerie}
Therefore, we neglect it throughout this article and take
\begin{align}\label{tau.69}
\calt_{nm}^{\rma\rmb}
\approx
\frac{i\hbar}{m_e}\gd_{ab}\gd_{nm}
.
\end{align} 
Finally,
we also need the following term (Eq.~\eqref{ntita})
\begin{align}\label{a.3c}
(v^{\lda,\rma}_{nn})_{;k^\rmb}
=
\nabla_{k^{\rma}}  
v^{\lda,\rmb}_{nn}(\bfk)
&=
-i\calt^{\rma\rmb}_{nn}
-
\sum_{\ell\ne n}
\go^\lda_{\ell n}
\bigg(  
r^{\rma}_{n\ell}  
r^\rmb_{\ell n}
+  
r^\rmb_{n\ell}  
r^{\rma}_{\ell n}
\bigg)
\nonumber\\
&\approx
\frac{\hbar}{m_e}\gd_{\rma\rmb}
-
\sum_{\ell\ne n}
\go^\lda_{\ell n}
\bigg(  
r^{\rma}_{n\ell}  
r^\rmb_{\ell n}
+  
r^\rmb_{n\ell}  
r^{\rma}_{\ell n}
\bigg)
,
\end{align}  
among other quantities for $\calv^{\gs,\rma,\ell}_{nm;\bfk}$, where we 
also use Eq.~\eqref{tau.69}. Above is the standard effective-mas sum rule.
\cite{ashcroft_solid_1976} 
%In Appendix \ref{code}, we list all the quantities that should be
%coded in order to calculate the previous expressions for $\bfgchi$.

\section{Conclusions}\label{con}

We have presented a complete derivation of the required elements to
calculate in the independent particle approach (IPA) the microscopic  
surface second harmonic susceptibility tensor $\bfgchi^S(-2\go;\go,\go)$ 
using a layer-by-layer approach. We have done so for semiconductors using 
the length gauge for the coupling of the external electric field to the 
electron. 
%Also,
%we calculated the radiated efficiency $R$ within the three layer
%model. 
%The combination of $\boldsymbol{\chi}$ and $R$ allow us
%to study this fascinating surface optical phenomena.

\appendix 
\input{appendices/appendix-re_ri}%reri A
\input{appendices/appendix-vnl}%appvnl B
\input{appendices/appendix-gen_der_calv}%calvs C 
\input{appendices/appendix-gen_der_omega}%gwk D
\input{appendices/appendix-calv}%appv E
\input{appendices/appendix-tnm}%calt F
\input{appendices/appendix-vc}%calpcalc G 
\input{appendices/appendix-gen_der_r_nonlocal}%gdernl H 
\input{appendices/appendix-code}%code I
%%%%
%\input{appendices/appendix-div_free_exp_chi}
%\input{appendices/appendix-dirac}
%\input{appendices/appendix-basic_relationships}
%\input{appendices/appendix-gen_der_r}
%\input{appendices/appendix-gen_der_calr}
%\input{appendices/appendix-odds_ends}

%\backmatter
%\bibliographystyle{prb}
%\nocite{*} 
\bibliography{ref}

\end{document}

* smoother C(z)

### 2014-11-16 paper-v-0.tex
bms-1

@@
  respectively.
-Now, the 
-Schr\"odinger equation reads 
-\begin{align}\label{ache.4} 
-\left(
-\frac{-\hbar^2}{2m_e}\nabla^2 
- +  V(\bfr)\right)\psi_{n\bfk}(\bfr) 
- + \int d\bfr' V^\nl(\bfr,\bfr')\psi_{n\bfk}(\bfr') 
-=\hbar \omega_i\psi_{n\bfk}(\bfr) 
-,
-\end{align} 
-where
-$V^\nl(\bfr,\bfr') = \bra{\bfr} V^\nl\ket{\bfr'}$,
-and
-$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
-are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
-by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
-is cell periodic. 
 In case of
@@
 
-As is well known, the use of the solutions of Eq.~\eqref{ache.4} as single 
-particle states leads to an underestimation of the band gap. A standard
+As is well known, the use of the LDA scheme 
+leads to an underestimation of the band gap. A standard
 procedure to correct for this is to
@@
 We emphasize that the scissored
-Hamiltonian has the same eigenfunctions as the unscissored Hamiltonian.
+Hamiltonian has the same eigenfunctions as the unscissored
+Hamiltonian,
+where
+$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
+are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
+by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
+is cell periodic.
 
@@
 (\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\sigma_{nm}-2\tilde\omega}\bigg[
-i\sum_{\ell }\Big(r_{n\ell }^{b}B_{\ell m}^{c}-B_{n\ell }^{c}r_{\ell m}^{b}
+i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
 \Big)  
@@
 semiconductor crystal is as follows. Instead of using a
-semi-infinite system, we replace it by a super slab that consist of a finite
+semi-infinite system, we replace it by a super cell that consist of a finite
 slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
-super slab, which constitutes the unit cell,
+super cell, which constitutes the unit cell,
 is repeated to form a full three dimensional crystalline structure.
@@
 \label{sz}
-\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{\rmb})  
+\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{b})  
             \Theta(z_\ell-z+\Delta_\ell^f),
@@
 \includegraphics[scale=.7]{images/slab}
-\caption{A sketch of a slab where the circles represent atoms.\label{fslab}}
+\caption{A sketch of the super cell, where the slab corresponds to the
+  circles representing the atoms of the system.\label{fslab}} 
 \end{figure}
@@
 \frac{1}{\Omega}\int d^3r\, \calc^\ell(z)\, \mathbf{j}^{(N)}(\mathbf{r},t)
- \equiv \mathbf{J}^{(N),\ell}(t),
-\end{equation}
-where $\mathbf{J}^{(N),\ell}(t)$ is the current of the
+ \equiv \mathbf{J}^{\ell,(N)}(t),
+\end{equation}
+where $\mathbf{J}^{\ell,(N)}(t)$ is the current of the
 $\ell$-th layer.
@@
 \begin{equation}\label{vcal}
-e{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}(\mathbf{k})
+e{\boldsymbol{\mathcal{V}}}^{\ell,\gs}_{mn}(\mathbf{k})
 \equiv
@@
 \begin{equation}\label{jmac2}
-\bfJ^{(N),\ell}(2\go)=\frac{e}{\gO}
+\bfJ^{\ell,(N)}(2\go)=\frac{e}{\gO}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\calbv^{\gs,\ell}_{mn}(\mathbf{k}) 
+\calbv^{\ell,\gs}_{mn}(\mathbf{k}) 
 \rho^{(N)}_{nm}(\bfk;2\go) 
@@
 \begin{align}\label{intj}
-{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}({\mathbf k})
+{\boldsymbol{\mathcal{V}}}^{\ell,\gs}_{mn}({\mathbf k})
 &=
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\calbv^{\gs,\ell}
+\calbv^{\ell,\gs}
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \begin{align}\label{nl.4}
-\calbv^{\gs,\ell}
+\calbv^{\ell,\gs}
 =
@@
 \begin{align}\label{vopii}
-\calbv^{\gs,\ell}
-&=
-\calbv^{\lda,\ell}
-+
-\calbv^{S,\ell}
+\calbv^{\ell,\gs}
+&=
+\calbv^{\ell,\lda}
++
+\calbv^{\ell,S}
 \nonumber\\
-\calbv^{\lda,\ell}
+\calbv^{\ell,\lda}
 &=
@@
 +
-\calbv^{\nl,\ell}
-=
-\frac{1}{m_e}
-\calbp^{\ell}
-+
-\calbv^{\nl,\ell}
+\calbv^{\ell,\nl}
 .
@@
 does not hold between
-$\calbv^{\sigma,\ell}_{nm}(\bfk)$   
+$\calbv^{\ell,\sigma}_{nm}(\bfk)$   
 and 
-$\calbv^{\lda,\ell}_{nm}(\bfk)$,
+$\calbv^{\ell,\lda}_{nm}(\bfk)$,
 i.e.
-$\calbv^{\sigma,\ell}_{nm}(\bfk)\ne
+$\calbv^{\ell,\sigma}_{nm}(\bfk)\ne
 (\go^\gs_{nm}/\go^\lda_{nm})
-\calbv^{\lda,\ell}_{nm}(\bfk)$,
+\calbv^{\ell,\lda}_{nm}(\bfk)$,
 and thus, to calculate
-$\calbv^{\sigma,\ell}_{nm}(\bfk)$ 
+$\calbv^{\ell,\sigma}_{nm}(\bfk)$ 
 we must calculate the matrix elements of 
-$\calbv^{\lda,\ell}$ and $\calbv^{S,\ell}$
+$\calbv^{\ell,\lda}$ and $\calbv^{\ell,S}$
  (separately)
@@
 To limit the SHG response to one surface, the equivalent of Eq.~\eqref{nl.4} 
-for $\calbv^\ell=\calbp^\ell/m_e$ was proposed in 
+for $\calbp^\ell$ was proposed in 
 Ref.~\onlinecite{reining_microscopic_1994} and later used in Refs.
@@
 \begin{equation}\label{Pjikn}
-\bfP^{(2),\ell}(2\go)
+\bfP^{\ell,(2)}(2\go)
 =\frac{ie}{2\gO\got}
@@
 \sum_{mn}
-\calbv^{\gs,\ell}_{mn}(\mathbf{k})
+\calbv^{\ell,\gs}_{mn}(\mathbf{k})
 \rho^{(2)}_{nm}(\bfk;2\go)
@@
 \sum_{mn}
-\frac{\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})}
+\frac{\calv^{\ell,\gs,\rma}_{mn}(\mathbf{k})}
 {\go^\gs_{nm\bfk}-2\got}
@@
 &
-+i\sum_\ell\left(r_{n\ell}^{\rmb}B_{\ell m}^{\rmc}(\bfk,\go) -
-  B_{n\ell}^{\rmc}(\bfk,\go) 
-  r_{\ell m}^{\rmb}\right)
++i\sum_q\left(r_{nq}^{\rmb}B_{qm}^{\rmc}(\bfk,\go) -
+  B_{nq}^{\rmc}(\bfk,\go) 
+  r_{qm}^{\rmb}\right)
 \bigg]
@@
 which gives the susceptibility 
-$\chi^{\ell}_{\rma\rmb\rmc}(-2\go;\go,\go)$ 
+$\chi^{\ell,\rma\rmb\rmc}(-2\go;\go,\go)$ 
 of the $\ell$-th layer of the slab, 
@@
 of Eq.~\eqref{Pjikn2}, then,
+%\end{document}  
 \begin{equation}\label{chii}
@@
 \sum_{mn}
-\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}
+\frac{\calv_{mn}^{\ell,\gs,\rma}}{\go^\gs_{nm}-2\got}
 \left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}}
@@
 \int \frac{dk^3}{8\pi^3}
-\sum_{\ell mn}
-\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}
+\sum_{qmn}
+\frac{\calv_{mn}^{\ell,\gs,\rma}}{\go^\gs_{nm}-2\got}
 \left(
-\frac{r_{n\ell}^{\rmc} r_{\ell m}^{\rmb} 
-f_{m\ell}}{\go^\gs_{\ell m}-\got}
--\frac{r_{n\ell}^{\rmb} r_{\ell m}^{\rmc} 
-f_{\ell n}}{\go^\gs_{n \ell}-\got}
+\frac{r_{nq}^{\rmc} r_{qm}^{\rmb} 
+f_{mq}}{\go^\gs_{qm}-\got}
+-\frac{r_{nq}^{\rmb} r_{qm}^{\rmc} 
+f_{qn}}{\go^\gs_{nq}-\got}
 \right),
@@
 that is related to interband transitions.
-The generalized derivative in Eq.~\eqref{chii} is dealt with by using the chain rule 
+The generalized derivative in Eq.~\eqref{chii} is dealt with by the chain rule 
 \begin{equation}\label{gene2}
@@
 Using time-reversal invariance, an integration by parts to 
-remove the square in the denominator of the second term in Eq.~\eqref{gene2}, 
+remove the square in the denominator of the second term of Eq.~\eqref{gene2}, 
 and taking the limit of $\eta\to 0$, 
@@
 Eqs. \eqref{chii} and \eqref{chie}, 
-\input{appendices/chis}
-\noindent where we have splited the interband and intraband $1\go$ and $2\go$
+\begin{subequations}\label{xis}
+\begin{equation}\label{calvimchiewn}
+\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{e,\go}]= 
+\frac{\pi |e|^3}{2\hbar^2}
+\int \frac{dk^3}{8\pi^3}
+\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\gs_{cv}}
+\left[
+\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{qc}\{r^{\rmb}_{cv}r^{\rmc}_{vq}\}]}
+{(2\go^\gs_{cv}-\go^\gs_{cq})} 
+-\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vq}\{r^{\rmc}_{qc}r^{\rmb}_{cv}\}]}
+{(2\go^\gs_{cv}-\go^\gs_{qv})}
+\right]\gd(\go^\gs_{cv}-\go),
+\end{equation}  
+\begin{equation}\label{calvimchiwn}
+\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{i,\go}]= 
+\frac{\pi\vert e\vert^3}{2\hbar^2}
+\int \frac{dk^3}{8\pi^3}
+\sum_{cv}\frac{1}{(\omega^\gs_{cv})^{2}}
+\left[
+\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
++\frac{\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gs_{cv}} 
+\right]\delta(\omega^\gs_{cv}-\omega),
+\end{equation}
+\begin{equation}\label{calvimchie2wn}
+\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{e,2\go}]= 
+-\frac{\pi |e|^3}{2\hbar^2}
+\int \frac{dk^3}{8\pi^3}
+\sum_{vc}\frac{4}{\omega^\gs_{cv}}
+\left[
+\sum_{v'\ne
+  v}\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\{r^{\rmb}_{cv'}r^{\rmc}_{v'v}\}]}
+{2\go^\gs_{cv'}-\go^\gs_{cv}}
+- \sum_{c'\ne
+  c}\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\{r^{\rmc}_{cc'}r^{\rmb}_{c'v}\}]}
+{2\go^\gs_{c'v}-\go^\gs_{cv}}
+\right]\gd(\go^\gs_{cv}-2\go),
+\end{equation}
+\begin{equation}\label{calvimchi2wn}
+\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{i,2\go}]= 
+ \frac{\pi \vert
+   e\vert^{3}}{2\hbar^2}
+\int \frac{dk^3}{8\pi^3}
+\sum_{vc}\frac{4}{(\omega^\gs_{cv})^{2}}
+\left[\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
+\right\}\right] -
+\frac{2\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gs_{cv}}\right]\delta(\omega^\gs_{cv}-2\omega)
+,
+\end{equation}
+\end{subequations}
+ where we have splited the interband and intraband $1\go$ and $2\go$
 contributions and supres the $\go$ arguments for space sake.
@@
 $\chi^{\ell,\rma\rmb\rmc}=\chi^{\ell,\rma\rmc\rmb}$.
-We calculate the nonlinear surface susceptibility as 
+The various quantities involved in Eqs.~\eqref{xis} are given in
+the Appendix \ref{appe}.
+
+Finally, we calculate the nonlinear surface susceptibility as 
 \begin{equation}\label{chiijksur}
@@
 either the front or the back surface. 
-\textcolor{red}{Talk about the correct value of $d$}
+%\mathrextcolor{red}{Talk about the correct value of $d$}
 
@@
 \appendix 
-\section{}
+\section{}\label{appe}
 We give explicit expressions for the quantities used in the evaluation 
-of Eq.~\eqref{calvimchiewn}-\eqref{calvimchi2wn}; when appropriate, some 
+of Eq.~\eqref{xis}; when appropriate, some 
 intermediate steps are given for their derivation. 
-\input{appendices/appendix-vc}%calpcalc G
-
-\input{appendices/appendix-gen_der_r_nonlocal}%gdernl H
-
-\input{appendices/appendix-vnl}%gdernl H
-
-\input{appendices/appendix-tnm}%calt F
-
-\input{appendices/appendix-gen_der_calv}%calvs C
+%%%%
+\subsection{ Expressions for 
+\texorpdfstring{${\cal V}^{\ell,a}_{nm}(\bfk)$, 
+${\cal C}^{\ell}_{nm}(\bfk)$ 
+and
+$({\cal C}^{\ell}_{nm}(\bfk))_{;\bfk}$
+}{Vnm and Cnm}}\label{calpcalc}
+
+Expanding the wave function in plane waves we obtain
+\begin{align}\label{eni.1}
+\psi_{n\bfk}(\bfr)=\sum_\bfG A_{n\bfk}(\bfG)e^{i(\bfk+\bfG)\cdot\bfr}
+,
+\end{align}
+where $\{\bfG\}$ are the reciprocal basis vectors satisfying
+$e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
+space, and $A_{n\bfk}(\bfG)$ are the expansion coefficients. Using
+$m_e\bfv=-i\hbar\bfgnabla$ into the equivalent of Eq.~\eqref{intj}
+ we obtain,\cite{mendozaPRB06}
+\begin{align}\label{eni.2}
+\calbv^\ell_{nm}(\bfk)=
+\frac{\hbar}{2m_e}
+\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
+(2\bfk+\bfG+\bfG')
+\gd_{\bfG_\parallel \bfG'_\parallel}  
+f^\ell(G_\perp-G'_\perp)
+,
+\end{align}   
+where
+\begin{align}\label{vnl.9}
+f^\ell(g)=\frac{1}{L}\int_{z_\ell-\gD^b_\ell}^{z_\ell+\gD^f_\ell} e^{igz}dz  
+ ,
+\end{align}
+with $f^{\ell*}(g)=f^\ell(-g)$. 
+the reciprocal lattice vectors $\bfG$ are decomposed into components
+parallel to the surface $\bfG_\parallel$, and perpendicular to the
+surface $G_\perp \hat z$, so
+that $\bfG = \bfG_\parallel + G_\perp\hat z$.
+The double summation over the $\bfG$ vectors can be efficiently done by  
+creating a pointer array to identify all the plane-wave coefficients  
+associated with the same $G_\parallel$.  
+ Likewise we obtain that
+\begin{align}\label{eni.4}
+\calc^\ell_{nm}(\bfk)=
+\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
+\gd_{\bfG_\parallel \bfG'_\parallel} 
+f^\ell(G_\perp-G'_\perp)
+.
+\end{align}  
+If $\calc^\ell(z)=1$, $f^\ell(g)=\gd_{g0}$, and from Eqs.~\eqref{eni.2} and \eqref{eni.4}
+we obtain the full-slab or bulk values
+$\bfv_{nm}(\bfk)$ and
+$\calc^\ell_{nm}(\bfk)=\gd_{nm}$.
+
+Since for any function $F(\bfr)$, $[\bfr,F(\bfr)]=0$, using 
+Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn}, we obtain
+\begin{align} 
+(\calc^\ell_{nm})_{;\bfk}
+&=
+i
+\sum_{q} 
+ \left(\bfr_{nq}
+\calc^\ell_{qm}
+-
+\calc^\ell_{nq}
+\bfr_{qm}
+\right) 
++i\bfr_{nm}(\calc^\ell_{mm}-\calc^\ell_{nn}) 
+,
+\end{align} 
+where we remind the reader that $\bfr_{nm}$ are calculated through
+Eq.~\eqref{chon.10} for LDA.
+
+%%%%
+\subsection{Expressions for 
+\texorpdfstring{$(\calv^{\ell,\lda,a}_{nm})_{;k^b}$}{Vnonlocal}
+and \texorpdfstring{$(r^a_{nm})_{;k^b}$}{Vnonlocal}
+for non-local potentials}\label{appvnl}
+Using Eqs.~\eqref{rnminn}, \eqref{conmri3n}, \eqref{gendevnn} and
+defining 
+$
+\calt^{\rma\rmb}\equiv[r^{\rmb},\calv^{\lda,\rma}]
+\equiv
+[r^{\rmb},\calv^\rma]
+$
+one can show that
+\begin{align}\label{nmesn}
+(\calv^{\ell,\lda,\rma}_{nm})_{;k^{\rmb}}&=
+\calt_{nm}^{\ell,\rma\rmb}
++i
+\sum_{q}
+\bigg(
+r^{\rmb}_{nq}  
+\calv^{\ell,\lda,\rma}_{q m}
+-
+\calv^{\ell,\lda,\rma}_{nq}   
+r^{\rmb}_{q m}
+\bigg)  
++i  
+r^{\rmb}_{nm}
+\Delta^{ell,\rma}_{mn}
+,
+\end{align}  
+where
+\begin{eqnarray}\label{tdel}
+\Delta^{\ell,\rma}_{mn}
+=
+\calv^{\ell,\lda,\rma}_{nn}  
+-
+\calv^{\ell,\lda,\rma}_{mm}  
+,
+\end{eqnarray} 
+and
+\begin{align}\label{tau.1}
+\calt_{nm}^{\ell,\rma\rmb}
+=
+\frac{\hbar}{m_e}\gd_{\rma\rmb}
+C^\ell_{nm} 
+-i 
+\sum_q 
+[r^{\rmb},v^{\nl,\rma}]_{nq}C^\ell_{qm} 
+.
+\end{align}  
+The matrix elements $[r^{\rmb},v^{\nl,\rma}]_{nm}$
+are calculated in Appendix \ref{calt}.
+% is small
+%compared to the other terms, thus we neglect it throwout this work.\cite{valerie}
+
+To obtain $(r^{\rma}_{nm})_{;k^{\rmb}}$ we use Eq.~\eqref{chon.10} to
+write
+$(r^{\rma}_{nm})_{;k^{\rmb}}
+=(v^{\lda,\rma}_{nm}/i\go^\lda_{nm})_{;k^{\rmb}}
+$, and apply the chain rule. 
+Then, 
+$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$ is obtained
+from Eq.~\eqref{nmesn} by
+taking
+$C^\ell(z)=1$ or $C^\ell_{nm}=\gd_{nm}$. Thus,
+\begin{align}\label{na_rgendevn}
+(r^{\rma}_{nm})_{;k^{\rmb}}
+&=
+t^{\rma\rmb}_{nm}
++
+\frac{ 
+r^{\rmb}_{nm}
+\Delta^{\rma}_{mn}
++r^{\rma}_{nm}
+\Delta^{\rmb}_{mn}
+}
+{\go^\lda_{nm}}
++
+\frac{i}{\go^\lda_{nm}}
+\sum_{q}
+\bigg(
+\go^\lda_{q m} 
+r^{\rmb}_{nq} 
+r^{\rma}_{q m}
+-
+\go^\lda_{nq} 
+r^{\rma}_{nq} 
+r^{\rmb}_{q m}
+\bigg)
+,
+\end{align} 
+where 
+\begin{eqnarray}\label{del}
+\Delta^{\rma}_{mn}
+=
+v^{\lda,\rma}_{nn}  
+-
+v^{\lda,\rma}_{mm}  
+,
+\end{eqnarray}
+and
+\begin{align}\label{tau.1n} 
+t_{nm}^{\rma\rmb}
+=\frac{\hbar}{m_e}\gd_{ab}\gd_{nm} 
+-i [r^{\rmb},v^{\nl,\rma}]_{nm} 
+.
+\end{align}   
+generalizes the usual expresion of
+$(r^a_{nm})_{;k^b}$ for local 
+Hamiltonians,\cite{aversaPRB95,nastosPRB05,cabellosPRB09,rashkeevPRB98}
+to
+the case of a
+nonlocal Hamiltonian.
+Note that the layered term
+$\calt^{\ell,\rma\rmb}_{nm}$ reduces to $t^{\rma\rmb}_{nm}$, for the
+full-slab or bulk case.
+
+%%%%
+\subsection{Matrix elements of 
+\texorpdfstring{$\calbv^{\ell,\nl}$}{Vnonlocal},
+\texorpdfstring{$\bfv^\nl$}{Vnonlocal}
+and $[\bfr,\bfv^\nl]$}\label{vesnl}
+
+We take Eq.~\eqref{vcali}
+and use
+$\sum_{\bfG}\ket{\bfk+\bfG}\bra{\bfk+\bfG}=1$, 
+with
+$\braket{\bfr}{\bfk+\bfG}=(1/\sqrt{\gO})
+\mathrm{exp}(i(\bfk+\bfG)\cdot\bfr)$,
+to obtain,\cite{mottaCMS10}
+\begin{align}\label{vnl.5}
+\calbv^{\ell,\nl}_{nm}(\bfk)&=\frac{1}{2}
+\sum_{\bfG}
+\Big(
+\bra{n\bfk} C^\ell(z) 
+\ket{\bfk+\bfG}\bra{\bfk+\bfG}
+\bfv^\nl \ket{m\bfk}
++
+\bra{n\bfk}
+\bfv^\nl  
+\ket{\bfk+\bfG}\bra{\bfk+\bfG}
+C^\ell(z) \ket{m\bfk}
+\Big)
+\nonumber\\
+&=\frac{1}{2 \hbar}
+\sum_{\bfG}
+\left(
+\calf^{\ell*}_{n\bfk}(\bfG) 
+\calh_{m\bfk}(\bfG) 
++
+\calh^*_{n\bfk}(\bfG) 
+\calf^\ell_{m\bfk}(\bfG) 
+\right) 
+,
+\end{align}  
+where 
+we have defined  
+\begin{align}\label{vnl.10}
+\calf^\ell_{n\bfk}(\bfG) 
+=
+\sum_{\bfG'} 
+A_{n\bfk}(\bfG') 
+\gd_{\bfG_\parallel \bfG'_\parallel}f^\ell(\bfG'_\perp-\bfG_\perp) 
+,
+\end{align} 
+\begin{align}\label{vnl.11}
+\calh_{n\bfk}(\bfG)&=
+\sum_{\bfG'} 
+A_{n\bfk}(\bfG') 
+(\nabla_{\bfK}+\nabla_{\bfK'})  
+V^\nl(\bfK,\bfK') 
+,
+\end{align}
+and $\bfK=\bfk+\bfG$, $\bfK'=\bfk+\bfG'$.
+For fully  separable pseudopotentials in the   
+Kleinman-Bylander (KB) form,\cite{mottaCMS10,kleinman_efficacious_1982,adolphPRB96}  
+the  
+matrix elements 
+ $\bra{\bfK}  
+V^\nl  
+\ket{\bfK'}
+=V^\nl(\bfK,\bfK')  
+$, 
+along with their $\bfK$ and $\bfK'$ gradient 
+can be readily calculated.\cite{mottaCMS10,adolphPRB96,gordienkoRPJ04,fuchsCPC99} 
+We have 
+implemented 
+the calculation of $\calbv^{\ell,\nl}_{nm}(\bfk)$ with the help pf the  \depe~code.\cite{francesco}
+Taking $C^\ell(z)=1$, which implies $f^\ell(g)=\gd_{g0}$, form above three expressions we obtain 
+$\bfv^{\nl}_{nm}(\bfk)$.
+
+%%%%
+%\subsection{Matrix elements of $[\bfr,\bfv^\nl]$}\label{calt}
+\label{calt}
+Using Eq.~\eqref{vnl} we define
+\begin{align}\label{3.1}
+\call^{\rma\rmb}_{nm}(\bfk) 
+\equiv\frac{1}{i\hbar}
+\bra{n\bfk}
+[ r^{\rma}, v^{\nl,\rmb}]
+\ket{m\bfk}
+=
+\frac{1}{\hbar^2}
+\bra{n\bfk}
+\big[ r^{\rma},[ V^\nl, r^\rmb]\big]
+\ket{m\bfk}
+,
+\end{align} 
+that uppon expanding the triple commutator, leads to
+\begin{align}\label{3.5}
+\call^{\rma\rmb}_{nm}(\bfk) 
+&=
+\frac{1}{\hbar^2\gO}
+\sum_{\bfG,\bfG'} 
+A^*_{n\bfk}(\bfK) 
+A_{m\bfk}(\bfK')
+\int
+d\bfr d\bfr'
+ e^{-i\bfK\cdot\bfr}
+\Big(
+r^{\rma}
+V^\nl(\bfr,\bfr')
+r'^\rmb
+-
+V^\nl(\bfr,\bfr')
+r'^\rma
+r'^{\rmb}
+\nonumber\\
+&-
+r^\rmb
+r^{\rma}
+V^\nl(\bfr,\bfr')
++
+ r^\rmb
+V^\nl(\bfr,\bfr')
+r'^{\rma}
+\Big) 
+ e^{i\bfK'\cdot\bfr'}
+,
+\end{align} 
+where 
+$V^\nl(\bfr,\bfr') = \bra{\bfr} V^\nl\ket{\bfr'}$.
+We use the following identity
+\begin{align}\label{3.4}
+&
+\Big(
+\frac{\partial^2}{\partial K^\rma\partial K'^\rmb}
++
+\frac{\partial^2}{\partial K'^\rma\partial K'^\rmb}
++
+\frac{\partial^2}{\partial K^\rma\partial K^\rmb}
++
+\frac{\partial^2}{\partial K^\rmb\partial K'^\rma}
+\Big)
+\int 
+d\bfr d\bfr' 
+ e^{-i\bfK\cdot\bfr}
+V^\nl(\bfr,\bfr') 
+e^{i\bfK'\cdot\bfr'}
+\nonumber\\
+&=
+\int d\bfr d\bfr'
+ e^{-i\bfK\cdot\bfr}
+\Big( 
+r^{\rma} 
+V^\nl(\bfr,\bfr') 
+r'^\rmb
+- 
+V^\nl(\bfr,\bfr') 
+r'^\rma 
+r'^{\rmb}
+- 
+r^\rmb 
+r^{\rma} 
+V^\nl(\bfr,\bfr')
++
+ r^\rmb 
+V^\nl(\bfr,\bfr') 
+r'^{\rma}
+\Big)  
+e^{i\bfK'\cdot\bfr'}
+,
+\end{align}
+to write
+\begin{align}\label{3.7}
+\call^{\rma\rmb}_{nm}(\bfk)
+&=
+\frac{1}{\hbar^2\gO}
+\sum_{\bfG,\bfG'} 
+A^*_{n\bfk}(\bfK) 
+A_{m\bfk}(\bfK')
+\Big(
+\frac{\partial^2}{\partial K^\rma\partial K'^\rmb}
++
+\frac{\partial^2}{\partial K'^\rma\partial K'^\rmb}
++
+\frac{\partial^2}{\partial K^\rma\partial K^\rmb}
++
+\frac{\partial^2}{\partial K^\rmb\partial K'^\rma}
+\Big)
+\bra{\bfK} 
+V^\nl 
+\ket{\bfK'} 
+\nonumber\\
+,
+\end{align} 
+where
+\begin{align}\label{vkk}
+\bra{\bfK} 
+V^\nl 
+\ket{\bfK'} 
+=
+\int 
+d\bfr d\bfr' 
+ e^{-i\bfK\cdot\bfr}
+V^\nl(\bfr,\bfr') 
+e^{i\bfK'\cdot\bfr'}
+,
+\end{align}
+The double derivatives with respect to $\bfK$ and $\bfK'$ 
+can be worked out explicitly, and thus ${\cal L}^{\rma\rmb}_{nm}(\bfk)$
+could be calculated.\cite{valerie}
+
+%%%%
+\subsection{Expressions for  \texorpdfstring{$\bfv^{\ell,S}_{nm}$}{Vnm}
+and
+\texorpdfstring{$\Big({\calbv}^{\ell,S}_{nm}\Big)_{;\bfk}$}{(Vnm);kb}
+}\label{calvs} 
+From Eq.~\eqref{vcali}
+\begin{align}\label{a.3b}
+\calbv^{\ell,S}_{nm}
+&=
+\frac{1}{2}\sum_q\left(   
+\bfv^S_{nq}\calc^\ell_{qm}+\calc^\ell_{nq}\bfv^S_{qm}
+\right)  
+,
+\end{align}    
+where $\sum_q\ket{q\bfk}\bra{q\bfk}=1$ was used,
+and $\bfv^S_{nm}$ are given in Eq.~\eqref{chon.2}.
+Taking the generalized derivative of Eq.~\eqref{a.3b}, and applying
+the chain rule, we obtain
+\begin{align}\label{a.3bn}
+\left(\calbv^{\ell,S}_{nm}\right)_{;\bfk}
+&=
+\frac{1}{2}\sum_q\left(
+(\bfv^S_{nq})_{;\bfk}\calc^\ell_{qm}
++    
+\bfv^S_{nq}(\calc^\ell_{qm})_{;\bfk}
++
+(\calc^\ell_{nq})_{;\bfk} \bfv^S_{qm}
++
+\calc^\ell_{nq} (\bfv^S_{qm})_{;\bfk}
+\right)  
+,
+\end{align}    
+where from 
+Eq.~\eqref{chon.2}, 
+\begin{align}\label{choni.1}
+(\bfv^S_{nm})_{;\bfk}=i\gs f_{mn}
+(\bfr_{nm})_{;\bfk}
+,
+\end{align}
+in agreement with Eq. A(6) of Ref.~\onlinecite{cabellosPRB09}.
 
@@
 for a slab scheme used for surface calculations.
-
-\section{Non-linear Surface Susceptibility}\label{nonchi}
-
-In this section we outline the general procedure to obtain the 
-surface susceptibility tensor for second harmonic generation.
-We start with the 
-non-linear polarization $\mbf{P}$ written as
-\begin{align}\label{mshg}
-P_{\rma}(2\go)&=\chi_{\rma\rmb\rmc}(-2\go;\go,\go)
-E_{\rmb}(\go)
-E_{\rmc}(\go)
-\nonumber \\
-&+
-\chi_{\rma\rmb\rmc\rml}(-2\go;\go,\go)
-E_{\rmb}(\go)\nabla_{\rmc} E_{\rml}(\go)
-+\cdots,
-\end{align}
-where $\chi_{\rma\rmb\rmc}(-2\go;\go,\go)$ and 
-$\chi_{\rma\rmb\rmc\rml}(-2\go;\go,\go)$
-correspond to the dipolar and quadrupolar susceptibilities.
-We drop  the $(-2\go;\go,\go)$ argument to ease on the notation.
-The sum continues with higher multipolar terms.
-If we consider a semi-infinite system with a centrosymmetric bulk, the
-equation above can be separated into two contributions from symmetry 
-considerations alone; one from the surface of the system and the other from
-the bulk of the system. We take
-\begin{equation}\label{mshg2}
-P_{\rma}(\mbf{r})=\chi_{\rma\rmb\rmc}E_{\rmb}(\mbf{r})E_{\rmc}(\mbf{r})
-+
-\chi_{\rma\rmb\rmc\rml}E_{\rmb}(\mbf{r})\frac{\partial}{\partial
-  \mbf{r}_{\rmc}} E_{\rml}(\mbf{r}) 
-+\cdots,
-\end{equation}
-as the polarization with respect to the original coordinate system, and 
-\begin{align}\label{mshg3}
-P_{\rma}(-\mbf{r})&=\chi_{\rma\rmb\rmc}E_{\rmb}(-\mbf{r})E_{\rmc}(-\mbf{r})
-\nonumber \\
-&+
-\chi_{\rma\rmb\rmc\rml}E_{\rmb}(-\mbf{r})\frac{\partial}{\partial(-
-  \mbf{r}_{\rmc})} E_{\rml}(-\mbf{r}) 
-+\cdots, 
-\end{align}
-as the polarization in the coordinate system where inversion is taken,
-i.e. $\mbf{r} \to -\mbf{r}$. 
-Note that we have kept the same susceptibility tensors, and they must be 
-invariant under $\mbf{r} \to -\mbf{r}$ since the system is centrosymmetric.
-Recalling that $\mbf{P}(\mbf{r})$ and $\mbf{E}(\mbf{r})$ are polar vectors 
-\cite{jackson_classical_1975}, we have that Eq.~\eqref{mshg3} reduces to
-\begin{align}\label{mshg4}
--P_{\rma}(\mbf{r})&=\chi_{\rma\rmb\rmc}(-E_{\rmb}(\mbf{r}))(-E_{\rmc}(\mbf{r}))
--
-\chi_{\rma\rmb\rmc\rml}(-E_{\rmb}(\mbf{r}))(-\frac{\partial}{\partial\mbf{r}_{\rmc}})(-
-E_{\rml}(\mbf{r})) 
-+\cdots,
-\nonumber \\
-P_{\rma}(\mbf{r})&=-\chi_{\rma\rmb\rmc}E_{\rmb}(\mbf{r})E_{\rmc}(\mbf{r})
-+
-\chi_{\rma\rmb\rmc\rml}E_{\rmb}(\mbf{r})\frac{\partial}{\partial\mbf{r}_{\rmc}}
-E_{\rml}(\mbf{r}) 
-+\cdots,
-\end{align}
-that when compared with Eq.~\eqref{mshg2} leads to the conclusion that
-\begin{equation}\label{sshg}
-\chi_{\rma\rmb\rmc}=0
-\end{equation}
-for a centrosymmetric bulk.
-
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.6]{images/system}
-\caption{(Color Online) Sketch of the semi-infinite system with a 
-centrosymmetric bulk. The surface region is of width $\sim d$. 
-The incoming photon of frequency $\go$ is represented by
-a downward red arrow, whereas both the surface and bulk created second
-harmonic photons of frequency $2\go$ are represented by upward
-green arrows. The red color suggests an incoming infrared photon with
-a green second harmonic photon. The dipolar ($\chi_{\rma\rmb\rmc}$), and 
-quadrupolar ($\chi_{\rma\rmb\rmc\rml}$) susceptibility tensors are shown 
-in the regions where they are different from zero. The axis has $z$
-perpendicular to the surface and $\mbf{R}$ parallel to it.\label{fsystem}}
-\end{figure}
-
-If we move to the surface of the semi-infinite system our assumption 
-of centrosymmetry breaks down, and there is no restriction in 
-$\chi_{\rma\rmb\rmc}$. We conclude that the leading term
-of the polarization in a surface region is given by
-\begin{align}\label{sshgp1}
-\int dz 
-P_{\rma}(\mbf{R},z)
-\approx dP_{\rma}
-\equiv
-P_{\rma}^S
-\equiv
-\chi^S_{\rma\rmb\rmc}E_{\rmb}E_{\rmc},
-\end{align}
-where
-% $\mbf{R}$ is a vector parallel to the surface which is
-% perpendicular to $z$, $S$ is the surface area of the unit
-% cell that characterizes the surface of the system, and 
-$d$ is the
-surface region from which the dipolar signal of $\bfP$ is
-different from zero (see Fig.~\ref{fsystem}), and
-$\mbf{P}^S\equiv d\mbf{P}$
- is the surface SH polarization. Then, from Eq.~\eqref{mshg} we obtain that
-\begin{align}\label{sshgp2}
-\chi_{\rma\rmb\rmc}^S=d\chi_{\rma\rmb\rmc}
-\end{align}
-is the
-SH surface susceptibility. 
-On the other hand, 
-\begin{equation}\label{sshgp3}
-P^b_{\rma}(\mbf{r})=\chi_{\rma\rmb\rmc\rml}E_{\rmb}(\mbf{r})\nabla_{\rmc}
-E_{\rml}(\mbf{r}),  
-\end{equation}
-gives the bulk polarization. We immediately recognize that the surface
-polarization is of dipolar order while the bulk polarization is of
-quadrupolar order. The surface, $\chi^S_{\rma\rmb\rmc}$, and bulk, 
-$\chi_{\rma\rmb\rmc\rml}$, susceptibility tensor ranks are three and 
-four, respectively. We will only concentrate on surface SHG in this article 
-even though bulk generated SH is also a very important optical phenomenon. 
-Also, we leave out of this article other interesting surface SH phenomena like, electric field
-induced second harmonic (EFISH), which would be represented by a
-surface susceptibility  tensor of quadrupolar origin.
-In centrosymmetric systems for which the quadrupolar bulk
-response is much smaller than the dipolar surface response, SH is
-readily used as a very useful and powerful 
-optical surface probe.\cite{downer_optical_2001}
-
-In the following sections we present the theoretical approach to
-derive
-the expressions for the 
-surface susceptibility tensor $\chi^S_{\rma\rmb\rmc}$.
-
-\section{Length Gauge}\label{longi}
-
-We follow the article by Aversa and Sipe\cite{aversaPRB95} to calculate 
-the optical properties of a given system within the longitudinal gauge. 
-More recent derivations
-can also be found in Refs. \onlinecite{sipe_second-order_2000,lambrechtPSSB00}.  
-Assuming the long-wavelength approximation which implies 
-a position independent electric field, 
-$\bfE(t)$,  
-the Hamiltonian in the length gauge approximation is given by
-\begin{equation}\label{ache}
-{H}={H}^\gs_0-e \bfr \cdot \bfE,
-\end{equation}
-with
-\begin{align}\label{ache.1}
- H^\gs_0= H^\lda_0
- + S(\bfr,\bfp)
-,
-\end{align} 
-as the unperturbed Hamiltonian.
-The LDA Hamiltonian can be expressed as follows,
-\begin{align}\label{ache.2}
- H^\lda_0&=\frac{ p^2}{2m_e}  +  V^\ps
-\nonumber\\
- V^\ps
-&= V^l(\bfr) +  V^\nl
-,
-\end{align}  
-where $ V^l(\bfr)$ and $ V^\nl$ are the local and
-the non-local parts of the  
-crystal 
-pseudopotential $ V^\ps$. For the latter, we have that
-\begin{align}\label{ache.3n}
- V^\nl(\bfr,\bfr') 
-\equiv \bra{\bfr} V^\nl\ket{\bfr'}\neq 0 \quad\mathrm{for}\quad\bfr\neq\bfr' 
-,
-\end{align}
-where $V^\nl(\bfr,\bfr')$ is a function of $\bfr$ and $\bfr'$
-representing the non-local contribution of the pseudopotential.
-The 
-Schr\"odinger equation reads  
-\begin{align}\label{ache.4} 
-\left(
-\frac{-\hbar^2}{2m_e}\nabla^2  
- +  V^l(\bfr)\right)\psi_{n\bfk}(\bfr)  
- + \int d\bfr' V^\nl(\bfr,\bfr')\psi_{n\bfk}(\bfr')  
-=E_i\psi_{n\bfk}(\bfr)  
-,
-\end{align}  
-where
-%$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=\sqrt{\gO/8\pi^3}e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
-$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
-are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
-by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
-is cell periodic. 
-$m_e$ is the bare mass of the electron and $\gO$ is the 
-unit cell volume. The nonlocal scissors operator is given by 
-\begin{equation}\label{chon.0}
-S(\bfr,\bfp)=\hbar\gS\sum_n\int d^3k'(1-f_n(\bfk)) \ket{n\bfk'}\bra{n\bfk'}
-,
-\end{equation}
-where $f_n(\bfk)$ is the occupation number, that for $T=0$ K, is
-independent of $\bfk$, and is one for filled bands and zero for
-unoccupied bands. For semiconductors the filled bands correspond to
-valence
-bands  ($n=v$) and the unoccupied bands to conduction bands
- ($n=c$). 
-We have that
-\begin{align}\label{chon.1}  
-H^\lda_0\ket{n\bfk}&=\hbar\go^\lda_n(\bfk)\ket{n\bfk}
-\nonumber\\
-H^\gs_0\ket{n\bfk}&=\hbar\go^\gs_n(\bfk)\ket{n\bfk}
-,
-\end{align} 
-where 
-\begin{align}\label{chon.78}
-\hbar\go^\gs_n(\bfk)=\hbar\go^\lda_n(\bfk)+\hbar\gS(1-f_n)
-,
-\end{align}
-is the
- scissored
-energy. Here, 
-$\hbar\gS$ is the value by which the conduction bands are rigidly
-($\bfk$-independent)
- shifted
-upwards in energy, also known as the scissors shift. $\gS$ could be taken to be
-$\bfk$ dependent, but for most calculations (like the ones presented
-here), a rigid shift is sufficient. We can take 
-$\hbar\gS=E_g-E_g^\lda$ where $E_g$ could be the experimental band gap 
-or GW band gap taken at the $\Gamma$ point, i.e. $\bfk=0$.
-We used the fact that 
-$\ket{n\bfk}^\lda\approx\ket{n\bfk}^\gs$, thus negating the need to label
-the Bloch states with the LDA or $\gs$ superscripts. 
-The matrix elements of $\bfr$ are split between the {\it intraband} 
-($\bfr_i$) and {\it interband} ($\bfr_e$) parts , where 
-$\bfr=\bfr_i+\bfr_e$ and\cite{adamsJCP53,blountSSP62,aversaPRB95}
-\begin{align}\label{rnminn}
-\bra{n\bfk}\bfr_i\ket{m\bfk'} &= \gd_{nm}\left[\gd(\bfk-\bfk')\bfgxi_{nn}(\bfk)  
-+i\nabla_\bfk\gd(\bfk-\bfk')\right], \\
-\bra{n\bfk}\bfr_e\ket{m\bfk'} &=
-(1-\gd_{nm})\gd(\bfk-\bfk')\bfgxi_{nm}(\bfk)  
-,
-\label{rnmenn}
-\end{align} 
-and
-\begin{equation}\label{zetann}
-\bfgxi_{nm}(\bfk) \equiv i\frac{(2\pi)^3}{\gO}\int_{\gO} d\bfr\, u^*_{n\bfk}(\bfr)\nabla_{\bfk}u_{m\bfk}(\bfr).
-\end{equation}
-The interband part $\bfr_e$ can
-be obtained as follows. 
-We start by introducing the 
-velocity operator 
-\begin{align}\label{vop}
-\bfv^\gs&=
-\frac{1}{i\hbar}[\bfr,{H}^\gs_0]
-,
-\end{align}
-and calculating its matrix elements 
-\begin{equation}\label{conhrnm}
-i\hbar\bra{n\bfk}
-\bfv^\gs 
-\ket{m\bfk}
-=
-\bra{n\bfk}
-[\bfr,{H}^\gs_0]
-\ket{m\bfk}
-=\bra{n\bfk}\bfr{H}^\gs_0-{H}^\gs_0\bfr\ket{m\bfk}
-=(\hbar\go^\gs_m(\bfk)-\hbar\go^\gs_n(\bfk))\bra{n\bfk}\bfr\ket{m\bfk}
-,
-\end{equation}
-thus defining $\go^\gs_{nm}(\bfk)=\go^\gs_n(\bfk)-\go^\gs_m(\bfk)$ we get 
-\begin{align}\label{pmnrmn}
-\bfr_{nm}(\bfk) 
-&=
-\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
-\quad\quad n\notin D_m 
-,
-\end{align} 
-which can be identified as 
-$\bfr_{nm}=(1-\gd_{nm})\bfgxi_{nm}\to \bfr_{e,nm}$. 
-Here, $D_m$ are all the possible degenerate $m$-states. 
-When $\bfr_i$ appears in 
-commutators we use\cite{aversaPRB95} 
-\begin{equation}\label{conmri3n}
-\bra{n\bfk}[\bfr_i,\calo]\ket{m\bfk'}
-=i\gd(\bfk-\bfk')(\calo_{nm})_{;\bfk}
-,
-\end{equation}  
-with 
-\begin{equation}\label{gendevnn}
-(\calo_{nm})_{;\bfk}=
-\nabla_\bfk 
-\calo_{n m}(\bfk) 
-- 
-i 
-\calo_{nm}(\bfk) 
-\left(
-\bfgxi_{nn}(\bfk) 
--
-\bfgxi_{mm}(\bfk) 
-\right) 
-,
-\end{equation} 
-where ``$;\bfk$'' denotes the generalized derivative (see Appendix \ref{reri}).
-
-As can be seen from Eq.~\eqref{ache.1} and \eqref{ache.2},
-both $ S$ and $ V^\nl$ are nonlocal potentials. Their contribution 
-in the calculation of the optical response has to be taken in order to
-get reliable results.\cite{ismailPRL01}
-We proceed as follows; from Eqs.~\eqref{vop}, \eqref{ache.1} and
-\eqref{ache.2}
- we find
-\begin{align}\label{vop2}
-\bfv^\gs&=
-\frac{\bfp}{m_e}
-+
-\frac{1}{i\hbar}[\bfr, V^\nl(\bfr,\bfr')]
-+
-\frac{1}{i\hbar}[\bfr,S(\bfr,\bfp)]
-\nonumber\\
-&\equiv  
-\bfv  
-+
-\bfv^\nl  
-+\bfv^S  
-=
-\bfv^\lda  
-+\bfv^S  
-,
-\end{align} 
-where we have defined  
-\begin{align}\label{ve}
-\bfv  
-&=\frac{\bfp}{m_e}
-\\
-\bfv^\nl  
-&=
-\frac{1}{i\hbar}[\bfr, V^\nl]
-\\\label{vs}
-\bfv^S  
-&=
-\frac{1}{i\hbar}[\bfr, S(\bfr,\bfp)]
-\nonumber\\\label{vlda}
-\bfv^\lda  
-&=
-\bfv  
-+\bfv^\nl  
-\end{align}   
-with $\bfp=-i\hbar\bfgnabla$ the momentum operator.  
-Using Eq.~\eqref{chon.0}, we obtain that the
-matrix elements of $\bfv^S$ are given by
-\begin{equation}\label{chon.2} 
-\bfv^S_{nm}=i\gS f_{mn}\bfr_{nm},
-%=\frac{\gS f_{mn}}{\go^\lda_{nm}}\bfv^\lda_{nm}
-\end{equation}
-with $f_{nm}=f_n-f_m$,
-where we see that $\bfv^S_{nn}=0$, then 
-\begin{align}\label{chon.8}
-\bfv^\gs_{nm}
-&=
-\bfv^\lda_{nm}+i\gS f_{mn}\bfr_{nm}
-\nonumber\\
-&=
-\bfv^\lda_{nm}+i\gS f_{mn}\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
-\nonumber\\
-\bfv^\gs_{nm}
-\frac{\go^\gs_{nm}-\gS f_{mn}}{\go^\gs_{nm}}
-&=
-\bfv^\lda_{nm}
-\nonumber\\
-\bfv^\gs_{nm}
-\frac{\go^\lda_{nm}}{\go^\gs_{nm}}
-&=
-\bfv^\lda_{nm}
-\nonumber\\
-\frac{\bfv^\gs_{nm}}{\go^\gs_{nm}}
-&=
-\frac{\bfv^\lda_{nm}}{\go^\lda_{nm}}
-,
-\end{align} 
-since
-$\go^\gs_{nm}-\gS f_{mn}=\go^\lda_{nm}$. Therefore, 
-\begin{align}\label{chon.9}
-\bfv^\gs_{nm}(\bfk)
-&=
-\frac{\go^\gs_{nm}}{\go^\lda_{nm}}\bfv^\lda_{nm}(\bfk)
-=\left(1+\frac{\gS}{\go_c(\bfk)-\go_v(\bfk)}\right)\bfv^\lda_{nm}(\bfk)
-\quad\quad n\notin D_m
-\nonumber\\
-\bfv^\gs_{nn}(\bfk)
-&=
-\bfv^\lda_{nn}(\bfk)
-,
-\end{align} 
-and Eq.~\eqref{pmnrmn} gives
-\begin{align}\label{chon.10}
-\bfr_{nm}(\bfk) 
-=
-\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
-=
-\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
-\quad\quad n\notin D_m 
-. 
-\end{align}
-The matrix elements 
-of $\bfr_e$ are the same whether we use the LDA or the scissored 
-Hamiltonian and there is no need to label them with either LDA 
-or S superscripts. Thus,
-we can write 
-\begin{equation}\label{chon.98}
-\bfr_{e,nm}\to 
-\bfr_{nm}(\bfk)=\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
-\quad\quad n\notin D_m 
-,
-\end{equation}   
-which gives the interband matrix elements of the position operator 
-in terms of the matrix elements of $\bfv^\lda$. 
-These matrix elements include the matrix elements of 
-$\bfv^\nl_{nm}(\bfk)$ which can be readily calculated\cite{francesco} for 
-fully separable nonlocal pseudopotentials in the 
-Kleinman-Bylander 
-form.\cite{mottaCMS10,kleinman_efficacious_1982,adolphPRB96}
-In Appendix \ref{appvnl} we outline how this  
-can be accomplished.
-
-\section{Time-dependent Perturbation Theory}\label{tdpt}
-
-In the independent particle approximation, we use the electron density
-operator $\gr$ to obtain the expectation value of any observable
-$\calo$ as
-\begin{equation}\label{traza}
-\calo=\mathrm{Tr}(\calo\rho)=\mathrm{Tr}(\rho\calo),
-\end{equation}
-where $\mathrm{Tr}$ is the trace and is invariant under cyclic permutations.
-The dynamic equation of motion for $\gr$ is given by
-\begin{equation}\label{eqrho}
-i\hbar \frac{d\gr}{dt}=[{H},\gr]
-,
-\end{equation}
-where it is more convenient to work in the interaction picture. We transform 
-all operators according to 
-\begin{equation}\label{ip}
-{\calo}_I={U}{\calo}{U}^\dagger,
-\end{equation}
-where
-\begin{equation}\label{ou}
-{U}=e^{i{H}_0t/\hbar},
-\end{equation}
-is the unitary operator that shifts us to the interaction picture.
-Note that $\calo_I$ depends on time even if $\calo$ does not.
-Then, we transform Eq.~\eqref{eqrho} into
-\begin{equation}\label{intrho}
-i\hbar\frac{d\gr_I(t)}{dt}=[-e\bfr_I(t)\cdot\bfE(t),\gr_I(t)],
-\end{equation}
-that leads to
-\begin{equation}\label{intrho2}
-\gr_I(t)=\gr_I(t=-\infty)
-+
-\frac{ie}{\hbar}\int_{-\infty}^t dt'[\bfr_I(t')\cdot\bfE(t'),\gr_I(t')].
-\end{equation}
-We assume that the interaction is switched-on adiabatically and
-choose a time-periodic perturbing field, to write
-\begin{equation}\label{efield}
-\bfE(t)=\bfE e^{-i\go t}e^{\eta t}=\bfE e^{-i\got t}
-,
-\end{equation}
-with
-\begin{equation}\label{got}
-\got=\go+i\eta
-,
-\end{equation} 
-where $\eta > 0$ assures
-that at $t=-\infty$ the interaction is zero and has its full strength $\bfE$ 
-at $t=0$. After computing the required time integrals one takes
-$\eta\to 0$. 
-Also, $\gr_I(t=-\infty)$ should be time independent and thus 
-$[{H},\rho]_{t=-\infty}=0$, This implies that 
-$\gr_I(t=-\infty)=\rho(t=-\infty)\equiv\rho_0$,
-where $\rho_0$ is
-the density matrix of the unperturbed ground state,
-such that
-\begin{equation}\label{nrhon}
-\bra{n\bfk}\gr_0\ket{m\bfk'}=f_n(\hbar\go^\gs_n(\bfk))\gd_{nm}\gd(\bfk-\bfk')
-,
-\end{equation}
-with $f_n(\hbar\go^\gs_n(\bfk))=f_{n\bfk}$ as the Fermi-Dirac distribution function.
-
-We solve Eq.~\eqref{intrho2} using the standard iterative
-solution, for which we write
-\begin{equation}\label{rhop}
-\rho_I=\rho_I^{(0)}+\rho_I^{(1)}+\rho_I^{(2)}+\cdots
-,
-\end{equation}
-where $\rho_I^{(N)}$ is the density operator to order $N$ in $\bfE(t)$.
-Then, Eq.~\eqref{intrho2} reads
-\begin{equation}\label{intrho3}
-\rho_I^{(0)}+\rho_I^{(1)}+\rho_I^{(2)}+\cdots
-=\gr_0
-+
-\frac{ie}{\hbar}\int_{-\infty}^t dt'[\bfr_I(t')\cdot\bfE(t'),
-\rho_I^{(0)}+\rho_I^{(1)}+\rho_I^{(2)}+\cdots
-]
-,
-\end{equation}
-where, by equating equal orders in the perturbation, we find
-\begin{equation}\label{rho0}
-\gr_I^{(0)}\equiv\gr_0
-,
-\end{equation}
-and
-\begin{equation}\label{rhoN}
-\gr_I^{(N)}(t)=
-\frac{ie}{\hbar}
-\int_{-\infty}^t dt'[\bfr_I(t')\cdot\bfE(t'),\gr^{(N-1)}_I(t')].
-\end{equation}
-It is simple to show that matrix elements of Eq. (\ref{rhoN}) satisfy
-$\bra{n\bfk}\gr_I^{(N+1)}(t)\ket{m\bfk'}=\gr^{(N+1)}_{I,nm}(\bfk)\gd(\bfk-\bfk')$,
-with
-\begin{equation}\label{rtilde}
-\gr^{(N+1)}_{I,nm}(\bfk;t)
-=\frac{ie}{\hbar}\int_{-\infty}^t dt'
-\bra{n\bfk}
-[\bfr_I(t'),\gr^{(N)}_I(t')]
-\ket{m\bfk}
-\cdot\bfE(t')
-.
-\end{equation}
-
-We now work out the commutator of Eq.~\eqref{rtilde}. Then,
-\begin{align}\label{conmu1}
-\bra{n\bfk}
-[\bfr_I(t),\gr^{(N)}_I(t)]
-\ket{m\bfk}
-&=
-\bra{n\bfk}
-[{U}\bfr{U}^\dagger,{U}\gr^{(N)}(t){U}^\dagger]
-\ket{m\bfk}
-\nonumber \\
-&=
-\bra{n\bfk}
-{U}[\bfr,\gr^{(N)}(t)]{U}^\dagger
-\ket{m\bfk}
-\\
-&=
-e^{i\go^\gs_{nm\bfk}t}
-\left(
-\bra{n\bfk}
-[\bfr_e,\gr^{(N)}(t)]
-+
-[\bfr_i,\gr^{(N)}(t)]
-\ket{m\bfk}
-\right)
-\nonumber
-.
-\end{align}
-% where the time dependence of operator's interaction picture is
-% explicitly shown by the exponential factor, and the implicit
-% dependence of $\rho^{(N)}$ inherited from Eq.~\eqref{eqrho} is
-% shown by its $t$ argument.
-We calculate the interband term first, so using Eq.~\eqref{chon.98} we obtain
-\begin{align}\label{conmu2}
-\bra{n\bfk}
-[\bfr_e,\gr^{(N)}(t)]
-\ket{m\bfk}
-&=
-\sum_{\ell}
-\left(
-\bra{n\bfk}
-\bfr_e
-\ket{\ell\bfk}
-\bra{\ell\bfk}
-\gr^{(N)}(t)
-\ket{m\bfk}
-\right.
-\nonumber \\
-&
-\left.
--
-\bra{n\bfk}
-\gr^{(N)}(t)
-\ket{\ell\bfk}
-\bra{\ell\bfk}
-\bfr_e
-\ket{m\bfk}
-\right)
-\nonumber \\
-&=
-\sum_{\ell\ne n,m}
-\left(
-\bfr_{n\ell}(\bfk)
-\gr^{(N)}_{\ell m}(\bfk;t)
--
-\gr^{(N)}_{n\ell}(\bfk;t)
-\bfr_{\ell m}(\bfk)
-\right)
-\nonumber\\
-&\equiv
-\bfR^{(N)}_e(\bfk;t)
-,
-\end{align}
-
-and from Eq.~\eqref{conmri3n},
-\begin{equation}\label{conmri4}
-\bra{n\bfk}[\bfr_i,\rho^{(N)}(t)]\ket{m\bfk'}
-=i \gd(\bfk-\bfk') (\rho^{(N)}_{nm}(t))_{;\bfk}
-\equiv \gd(\bfk-\bfk')\bfR_i^{(N)}(\bfk;t)
-.
-\end{equation}
-Then Eq.~\eqref{rtilde} becomes
-\begin{equation}\label{rtilde2}
-\gr^{(N+1)}_{I,nm}(\bfk;t)
-=\frac{ie}{\hbar}\int_{-\infty}^t dt'
-e^{i(\go^\gs_{nm\bfk}-\got)t'}
-\left[R_e^{\rmb(N)}(\bfk;t')+R_i^{\rmb(N)}(\bfk;t')\right]E^{\rmb}
-,
-\end{equation}
-where the roman superindices
-$\rma,\rmb,\rmc$ denote Cartesian components that are summed over if repeated.
-Starting from the linear response and proceeding from Eq.~\eqref{nrhon} and ~\eqref{conmu2},
-\begin{align}\label{R0e}
-R_e^{\rmb(0)}(\bfk;t)
-&=
-\sum_{\ell}
-\left(
-r^{\rmb}_{n\ell}(\bfk)
-\gr^{(0)}_{\ell m}(\bfk)
--
-\gr^{(0)}_{n\ell}(\bfk)
-r^{\rmb}_{\ell m}(\bfk)
-\right)
-\nonumber \\
-&=
-\sum_{\ell}
-\left(
-r^{\rmb}_{n\ell}(\bfk)
-\gd_{\ell m}f_m(\hbar\go^\gs_m(\bfk))
--
-\gd_{n\ell}f_n(\hbar\go^\gs_n(\bfk))
-r^{\rmb}_{\ell m}(\bfk)
-\right)
-\nonumber \\
-&=
-f_{mn\bfk}
-r^{\rmb}_{nm}(\bfk)
-,
-\end{align}
-where $f_{mn\bfk}=f_{m\bfk}-f_{n\bfk}$.
-From now on,
-it should be clear that the matrix elements of $\bfr_{nm}$ imply
-$n\notin D_m$.
-We also have from Eq.~\eqref{conmri4} and Eq.~\eqref{gendevnn} that
-\begin{equation}\label{R0i}
-R_i^{\rmb(0)}(\bfk)=i(\rho^{(0)}_{nm})_{;k^{\rmb}}=i\gd_{nm}(f_{n\bfk})_{;k^{\rmb}}=i\gd_{nm}\nabla_{k^{\rmb}} f_{n\bfk}
-.
-\end{equation}
-For a semiconductor at $T=0$, $f_{n\bfk}$ is one if the state
-$\ket{n\bfk}$ is a valence state and zero if it is a conduction state; 
-thus $\nabla_\bfk f_{n\bfk}=0$ and $\bfR_i^{(0)}=0$ and
-the linear response has no contribution from
-intraband transitions.
- Then,
-\begin{align}\label{rtilde2n}
-\gr^{(1)}_{I,nm}(\bfk;t)
-&=\frac{ie}{\hbar}
-f_{mn\bfk}
-r^{\rmb}_{nm}(\bfk)E^{\rmb}
-\int_{-\infty}^t dt'
-e^{i(\go^\gs_{nm\bfk}-\got)t'}
-\nonumber \\
-&=\frac{e}{\hbar}
-f_{mn\bfk}
-r^{\rmb}_{nm}(\bfk)E^{\rmb}
-\frac{e^{i(\go^\gs_{nm\bfk}-\got)t}}
-{\go^\gs_{nm\bfk}-\got}
-\nonumber \\
-&=
-e^{i\go^\gs_{nm\bfk}t}
-B^{\rmb}_{mn}(\bfk)E^{\rmb}(t)
-\nonumber \\
-&=
-e^{i\go^\gs_{nm\bfk}t}
-\rho^{(1)}_{nm}(\bfk;t)
-,
-\end{align}
-with
-\begin{equation}\label{rho1} 
-B^{\rmb}_{nm}(\bfk,\go)=
-\frac{e}{\hbar}\frac{f_{mn\bfk}r^{\rmb}_{nm}(\bfk)}{\go^\gs_{nm\bfk}-\got}
-,
-\end{equation} 
-and
-\begin{equation}\label{rhonoi1}
-\rho^{(1)}_{nm}(\bfk;t)=B^{\rmb}_{mn}(\bfk,\go)E^{\rmb}(\go)e^{-i\got t}
-.
-\end{equation}
-
-Now, we calculate the second-order response. Then, from Eq.~\eqref{conmu2}
-\begin{align}\label{R1e}
-R_e^{\rmb(1)}(\bfk;t)
-&=
-\sum_{\ell}
-\left(
-r^{\rmb}_{n\ell}(\bfk)
-\gr^{(1)}_{\ell m}(\bfk;t)
--
-\gr^{(1)}_{n\ell}(\bfk;t)
-r^{\rmb}_{\ell m}(\bfk)
-\right)
-\nonumber \\
-&=
-\sum_{\ell}
-\left(
-r^{\rmb}_{n\ell}(\bfk)
-B^{\rmc}_{\ell m}(\bfk,\go)
--
-B^{\rmc}_{n\ell}(\bfk,\go)
-r^{\rmb}_{\ell m}(\bfk)
-\right)E^{\rmc}(t)
-,
-\end{align}
-and from Eq.~\eqref{conmri4}
-\begin{equation}\label{R1i}
-R_i^{\rmb(1)}(\bfk;t)=
-i(\rho^{(1)}_{nm}(t))_{;k^{\rmb}}=
-iE^{\rmc}(t)(B^{\rmc}_{nm}(\bfk,\go))_{;k^{\rmb}}
-.
-\end{equation}
-
-Using Eqs.~\eqref{R1e} and ~\eqref{R1i} in Eq. (\ref{rtilde2}),
-we obtain
-\begin{align}\label{rtilde33}
-\gr^{(2)}_{I,nm}(\bfk;t)
-&=
-\frac{ie}{\hbar}
-\bigg[
-\sum_{\ell}
-\Big(
-r^{\rmb}_{n\ell}(\bfk)
-B^{\rmc}_{\ell m}(\bfk,\go)
--
-B^{\rmc}_{n\ell}(\bfk,\go)
-r^{\rmb}_{\ell m}(\bfk)
-\Big)
-\nonumber \\
-&+
-i
-(B^{\rmc}_{nm}(\bfk,\go))_{;k^{\rmb}}
-\bigg]
-E^{\rmb}_{\go}E_{\go}^{\rmc}
-\int_{-\infty}^t dt'
-e^{i(\go^\gs_{nm\bfk}-2\got)t'}
-\nonumber \\
-&=
-\frac{e}{\hbar}
-\bigg[
-\sum_{\ell}
-\Big(
-r^{\rmb}_{n\ell}(\bfk)
-B^{\rmc}_{\ell m}(\bfk,\go)
--
-B^{\rmc}_{n\ell}(\bfk,\go)
-r^{\rmb}_{\ell m}(\bfk)
-\Big)
-\nonumber \\
-&+
-i
-(B^{\rmc}_{nm}(\bfk,\go))_{;k^{\rmb}}
-\bigg]
-E^{\rmb}_{\go}E_{\go}^{\rmc}
-\frac{e^{i(\go^\gs_{nm\bfk}-2\got)t}}
-{\go^\gs_{nm\bfk}-2\got}
-\nonumber \\
-&=
-e^{i\go^\gs_{nm\bfk}t}
-\rho_{nm}^{(2)}(\bfk;t)
-.
-\end{align}
-Now, we write
-$\rho_{nm}^{(2)}(\bfk;t)=\rho_{nm}^{(2)}(\bfk;2\go)e^{-i2\got t}$,
-with
-\begin{align}\label{rho2}
-\gr_{nm}^{(2)}(\bfk;2\go)&=\frac{e}{i\hbar}\frac{1}{\go^\gs_{nm\bfk}-2\got}
-\bigg[-(B_{nm}^{\rmc}(\bfk,\go)_{;k^{\rmb}}
-\nonumber \\
-&
-+i\sum_\ell\Big(r_{n\ell}^{\rmb}B_{\ell m}^{\rmc}(\bfk,\go) - B_{n\ell}^{\rmc}(\bfk,\go)
-  r_{\ell m}^{\rmb}\Big)
-\bigg] 
-E^{\rmb}(\go)E^{\rmc}(\go)
-\end{align} 
-where 
-$B_{\ell m}^{\rma}(\bfk,\go)$ are given by Eq.~\eqref{rho1}. 
-We remark that $\bfr_{nm}(\bfk)$  are
-the same whether calculated with the LDA or the scissored Hamiltonian. 
-We chose the former in this article.
-
-\section{Layered Current Density}\label{cd}
-
-In this section, we derive the expressions for the microscopic current
-density of a given layer in the unit cell of the system.
-The approach we use to study the surface of a semi-infinite
-semiconductor crystal is as follows. Instead of using a
-semi-infinite system, we replace it by a slab (see Fig.~\ref{fslab}).
-The slab consists of a front and back surface, and in between these
-two surfaces is the bulk of the system. 
-In general the surface of a crystal reconstructs or relaxes as the atoms
-move to find equilibrium positions. This is due to the fact that
-the otherwise
-balanced forces are disrupted when the surface atoms do not find their 
-partner atoms that are now absent at the surface of the slab.
-
-To take the reconstruction or relaxation into account, we take ``surface'' to mean
-the true surface of the first layer of atoms, and
-some of the atomic sub-layers adjacent to it.
-Since the front and the back
-surfaces of the slab are usually identical the total slab is
-centrosymmetric. This implies that 
-$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=0$, and thus we must
-find a way to bypass this characteristic of a centrosymmetric slab 
-in order to have a finite $\chi^s_{\rma\rmb\rmc}$ representative of the
-surface. Even if the front and back surfaces of the slab 
-are different, breaking the centrosymmetry and therefore giving an
-overall $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}\ne 0$, we still
-need a procedure to extract the front surface $\chi^f_{\rma\rmb\rmc}$
-and the back surface $\chi^b_{\rma\rmb\rmc}$ from the non-linear
-susceptibility 
-$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=\chi^f_{\rma\rmb\rmc}-\chi^b_{\rma\rmb\rmc}$
- of the entire slab.
-
-A convenient way to accomplish the separation of the SH signal of
-either surface is to introduce a ``cut function'', $\calc(z)$, which is 
-usually taken to be unity over one half of the slab and zero over 
-the other half.\cite{reining_microscopic_1994}
- In this case $\calc(z)$ will give the contribution of the 
-side of the slab for which $\calc(z)=1$. We can generalize this 
-simple choice for $\calc(z)$ by a top-hat cut function
-$\calc^\ell(z)$ that selects a given layer,
-\begin{equation}
-\label{sz}
-\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{\rmb})
-            \Theta(z_\ell-z+\Delta_\ell^f),
-\end{equation}
-where $\Theta$ is the Heaviside function. Here, $\Delta_\ell^{f/b}$
-is the distance that the $\ell$-th layer extends towards the front
-$(f)$ or back $(b)$ from its $z_\ell$ position. 
-$\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
-(see Fig.~\ref{fslab}).
-\begin{figure}[b]
-\centering
-%\includegraphics[height=5cm,width=7cm]{slab}
-\includegraphics[scale=.7]{images/slab}
-\caption{A sketch of a slab where the circles represent atoms.\label{fslab}}
-\end{figure}
-
-Now, we show how this ``cut function'' $\calc^\ell(z)$ is introduced in
-the calculation of $\chi_{\rma\rmb\rmc}$. 
-The microscopic current density is given by
-\begin{equation}\label{jmic}
-\bfj(\bfr,t)=\mathrm{Tr}(\bfj(\bfr)\rho(t)),
-\end{equation}
-where the operator for the electron's current is
-\begin{equation}\label{hatjmic}
-\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gs\ket{\bfr}\bra{\bfr}
-+\ket{\bfr}\bra{\bfr}\bfv^\gs\right), 
-\end{equation}
-where ${\mbf{v}}^\gs$ is the electron's velocity operator to be dealt
-with below. We define
-$\mu\equiv\ket{\bfr}\bra{\bfr}$ and use the cyclic invariance of
-the trace to write
-\begin{align}\label{jmic2}
-\mathrm{Tr}(\bfj(\bfr)\rho(t)&=\mathrm{Tr}(\rho(t)\bfj(\bfr))=
-\frac{e}{2}
-\left(
-\mathrm{Tr}(\rho\bfv^\gs\mu)
-+
-\mathrm{Tr}(\rho\mu\bfv^\gs)
-\right)
-\nonumber\\
-&=
-\frac{e}{2}
-\sum_{n\bfk}
-\left(
-\bra{n\bfk}\rho\bfv^\gs\mu\ket{n\bfk}
-+
-\bra{n\bfk}\rho\mu\bfv^\gs\ket{n\bfk}
-\right)
-\nonumber\\
-&=
-\frac{e}{2}
-\sum_{nm\bfk}
-\bra{n\bfk}\rho\ket{m\bfk}
-\left(
-\bra{m\bfk}\bfv^\gs\ket{\bfr}\braket{\bfr}{n\bfk}
-+
-\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gs\ket{n\bfk}
-\right)
-\nonumber\\
-\bfj(\bfr,t)&=
-\sum_{nm\bfk}
-\rho_{nm}(\bfk;t)\bfj_{mn}(\bfk;\bfr),
-\end{align}
-where
-\begin{equation}\label{jmic3}
-\bfj_{mn}(\bfk;\bfr)=
-\frac{e}{2}
-\left(
-\bra{m\bfk}\bfv^\gs\ket{\bfr}\braket{\bfr}{n\bfk}
-+
-\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gs\ket{n\bfk}
-\right),
-\end{equation}
-are the matrix elements of the microscopic current operator,
-and we have used the fact that the matrix elements between states $\ket{n\bfk}$
-are diagonal in $\bfk$, i.e. proportional to $\gd(\bfk-\bfk')$.
-
-Integrating the microscopic current $\mbf{j}(\mbf{r},t)$ over
-the entire slab gives the averaged microscopic current density. 
-If we want the contribution from only one region of the unit cell 
-towards the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
-over the desired region. The contribution to the current density from the
-$\ell$-th layer of the slab is given by
-\begin{equation}\label{jsz}
-\frac{1}{\Omega}\int d^3r\, \calc^\ell(z)\, \mathbf{j}(\mathbf{r},t)
- \equiv \mathbf{J}^{\ell}(t),
-\end{equation}
-where $\mathbf{J}^{\ell}(t)$ is the microscopic current in the
-$\ell$-th layer.
-Therefore we define
-\begin{equation}\label{vcal}
-e{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}(\mathbf{k})
-\equiv
-%\frac{1}{\Omega}
-\int d^3r\, \calc^\ell(z)\,\bfj_{mn}({\bfk};\bfr),
-\end{equation}
-to write
-\begin{equation}\label{jmac}
-J_a^{(N,\ell)}(t)=\frac{e}{\gO}
-\sum_{mn\bfk}
-\mathcal{V}^{\gs,a,\ell}_{mn}(\mathbf{k})
-\rho^{(N)}_{nm}(\bfk;t),
-\end{equation}
-as the induced microscopic current of the $\ell$-th layer, to order $N$ 
-in the external perturbation. The matrix elements of the 
-density operator for $N=1,2$ are given by Eqs.~\eqref{rho1} and
-\eqref{rho2} respectively. 
-The Fourier component of microscopic current of Eq.~\eqref{jmac} is given by
-\begin{equation}\label{jmac2}
-J_{\rma}^{(N,\ell)}(\go_3)=\frac{e}{\gO}
-\sum_{mn\bfk}
-\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})
-\rho^{(N)}_{nm}(\bfk;\go_3)
-.
-\end{equation}
-We proceed to give an explicit expression of
-$\calbv^{\gs,\ell}_{mn}(\mathbf{k})$.
-From
-Eqs.~\eqref{vcal} and \eqref{jmic3} we obtain
-\begin{equation}\label{intj}
-{\boldsymbol{\mathcal{V}}}^{\gs,\ell}_{mn}({\mathbf k})=
-\frac{1}{2}
-\int \mathrm{d}^3 r\,
- \calc^\ell(z)
-\bigg[
-\langle m\mathbf{k}|\mathbf{v}^\gs | \mathbf{r}\rangle
-\langle \mathbf{r} | n \mathbf k \rangle +
-\langle m\mathbf{k} | \mathbf{r}\rangle
-\langle \mathbf{r} | \mathbf{v}^\gs | n \mathbf k \rangle\bigg]
-,
-\end{equation}  
-and using the following property
-\begin{equation}\label{nl.2}
-\langle \mathbf{r} | \bfv^\gs(\bfr,\bfr')| n\mathbf{k} \rangle
-=\int d^3 r'' \bra{\mbf{r}}\bfv^\gs(\bfr,\bfr')\ket{\mbf{r}''}
-\braket{\mbf{r}''}{n\mbf{k}}
-=\bfv^\gs(\bfr,\bfr'')
-\int d^3 r'' \braket{\mbf{r}}{\mbf{r}''}
-\braket{\mbf{r}''}{n\mbf{k}}
-=\bfv^\gs(\bfr,\bfr')
-\psi_{n\mathbf{k}}(\mathbf{r})
-,
-\end{equation}
-that stems from the fact that the operator $\bfv^\gs(\bfr,\bfr')$ does not act on
-$\bfr''$, we can write
-\begin{align}\label{nl.3}
-\calbv^{\gs,\ell}_{mn}({\mathbf k})
-&=
-\frac{1}{2}
-\int \mathrm{d}^3 r\,
- \calc^\ell(z)
- \bigg[
-\psi_{n\mathbf{k}}(\mathbf{r})
-\bfv^{\gs *}\psi^*_{m\mathbf{k}}(\mathbf{r})
-+ 
-\psi^*_{m\mathbf{k}}(\mathbf{r})\bfv^\gs
-\psi_{n\mathbf{k}}(\mathbf{r})
-\bigg]
-\nonumber\\
-&=
-\int \mathrm{d}^3 r\,
-\psi^*_{m\mathbf{k}}(\mathbf{r})
-\left[\frac{\calc^\ell(z) \bfv^\gs +
-\bfv^\gs \calc^\ell(z)}{2}\right]
-\psi_{n\mathbf{k}}(\mathbf{r})
-\nonumber\\
-&=
-\int \mathrm{d}^3 r\,
-\psi^*_{m\mathbf{k}}(\mathbf{r})
-\calbv^{\gs,\ell}
-\psi_{n\mathbf{k}}(\mathbf{r})
-.
-\end{align}
-We used the hermitian property of $\bfv^\gs$ and defined
-\begin{align}\label{nl.4}
-\calbv^{\gs,\ell}
-=
-\frac{\calc^\ell(z) \bfv^\gs +
-\bfv^\gs \calc^\ell(z)}{2}
-,
-\end{align} 
-where the superscript $\ell$ is inherited from $\calc^\ell(z)$ and we
-supress the dependance on $z$ from the increasingly crowded notation.  
-We see that the replacement
-\begin{align}\label{vcali}
-\bfv^\gs \to \calbv^{\gs,\ell}=\left[\frac{\calc^\ell(z) \bfv^\gs +
-\bfv^\gs \calc^\ell(z)}{2}\right]
-,
-\end{align} 
-is all that is needed to change the
-velocity operator of the electron $\bfv^\gs$ to the new velocity
-operator $\calbv^{\gs,\ell}$ that implicitly takes into account the
-contribution of the region of the slab given by $\calc^\ell(z)$.
-From Eq.~\eqref{vop2},
-\begin{align}\label{vopii}
-\calbv^{\gs,\ell}
-&=
-\calbv^{\lda,\ell}
-+
-\calbv^{S,\ell}
-\nonumber\\
-\calbv^{\lda,\ell}
-&=
-\calbv^{\ell}
-+
-\calbv^{\nl,\ell}
-=
-\frac{1}{m_e}
-\calbp^{\ell}
-+
-\calbv^{\nl,\ell}
-.
-\end{align}
-We remark that the simple relationship between 
-$\bfv^{\sigma}_{nm}(\bfk)$ 
-and 
-$\bfv^{\lda}_{nm}(\bfk)$,
-given in 
-Eq.~\eqref{chon.9}, 
-does not hold between
-$\calbv^{\sigma,\ell}_{nm}(\bfk)$   
-and 
-$\calbv^{\lda,\ell}_{nm}(\bfk)$,
-i.e.
-$\calbv^{\sigma,\ell}_{nm}(\bfk)\ne
-(\go^\gs_{nm}/\go_{nm})
-\calbv^{\lda,\ell}_{nm}(\bfk)$ 
-and
-$\calbv^{\sigma,\ell}_{nn}(\bfk)\ne
-\calbv^{\lda,\ell}_{nn}(\bfk)$
-,
-and thus, to calculate
-$\calbv^{\sigma,\ell}_{nm}(\bfk)$ 
-we must calculate the matrix elements of $\calbv^{S,\ell}$ and
-$\calbv^{\lda,\ell}$ (separately)
-according to the expressions of
-Appendix \ref{calvs}.
-% Using Eq.~\eqref{chon.9}, the matrix elements of Eq.~\eqref{vcali}
-% are given by
-% \begin{align}\label{end.1}
-% \calbv^{\sigma,\ell}_{nm}(\bfk)
-% &=
-% \frac{1}{2}
-% \sum_q
-% \Big(
-% \calc^\ell_{nq}\bfv^\sigma_{qm}
-% +
-% \bfv^\sigma_{nq}\calc^\ell_{qm}
-% \Big)
-% \nonumber\\
-% &=
-% \frac{1}{2}
-% \Big(
-% \calc^\ell_{nn}\bfv^\sigma_{nm}
-% +
-% \bfv^\sigma_{nn}\calc^\ell_{nm}
-% +
-% \calc^\ell_{nm}\bfv^\sigma_{mm}
-% +
-% \bfv^\sigma_{nm}\calc^\ell_{mm}
-% \Big)
-% \nonumber\\
-% &+
-% \frac{1}{2}
-% \sum_{q\neq(nm)}
-% \Big(
-% \frac{\go^\sigma_{qm}}{\go_{qm}}
-% \calc^\ell_{nq}\bfv^\lda_{qm}
-% +
-% \frac{\go^\sigma_{nq}}{\go_{nq}}
-% \bfv^\lda_{nq}\calc^\ell_{qm}
-% \Big)
-% ,
-% \end{align}
-% \begin{align}\label{end.2}
-% \calbv^{\sigma,\ell}_{nn}(\bfk)
-% &=
-% \frac{1}{2}
-% \Big(
-% \calc^\ell_{nn}\bfv^\sigma_{nn}
-% +
-% \bfv^\sigma_{nn}\calc^\ell_{nn}
-% \Big)
-% \nonumber\\
-% &+
-% \frac{1}{2}
-% \sum_{q\neq n}
-% \Big(
-% \frac{\go^\sigma_{qn}}{\go_{qn}}
-% \calc^\ell_{nq}\bfv^\lda_{qn}
-% +
-% \frac{\go^\sigma_{nq}}{\go_{nq}}
-% \bfv^\lda_{nq}\calc^\ell_{qn}
-% \Big)
-% \nonumber\\
-% &=
-% \calc^\ell_{nn}\bfv^\lda_{nn}
-% \nonumber\\
-% &+
-% \frac{1}{2}
-% \sum_{q\neq n}
-% \frac{\go^\sigma_{qn}}{\go_{qn}}
-% \Big(
-% \calc^\ell_{nq}\bfv^\lda_{qn}
-% +
-% \bfv^\lda_{nq}\calc^\ell_{qn}
-% \Big)
-% ,
-% \end{align}
-
-To limit the response to one surface, the equivalent of Eq.~\eqref{nl.4} 
-for $\calbv^\ell=\calbp^\ell/m_e$ was proposed in 
-Ref.~\onlinecite{reining_microscopic_1994} and later used in Refs.
-\onlinecite{mendozaPRL98},
-\onlinecite{mendoza_ab_2001},
-\onlinecite{sanoPRB02},
- and \onlinecite{mejia_layer-by-layer_2004} 
-also in the context of SHG. 
-The layer-by-layer analysis of Refs. \onlinecite{hogan_optical_2003} 
-and \onlinecite{castilloPRB03} used Eq.~\eqref{sz}, 
-limiting the current response
-to a particular layer of the slab and used to obtain the
-anisotropic linear optical response of semiconductor surfaces.
-However, the first formal derivation of this scheme is presented in
-Ref.~\onlinecite{mendozaPRB06} for the linear response, and here in this 
-article, for the second harmonic optical response of semiconductors.
-
-%??d
-% debemos mencionar que v caligrafica con tijeras esta mal calculada
-% para la respuesta lineal y la inyeccion de corriente en el
-% tratamiento capa por capa. Tambien, habria que ver que pasa con la
-% inyeccion de espin capa-por-capa
-%??u
-
-\section{Microscopic surface susceptibility}
-In this section we obtain the expressions for the 
-surface susceptibility tensor $\chi^S_{\rma\rmb\rmc}$.
-We start with the basic relation $\bfJ=d\bfP/dt$ 
-with $\bfJ$ the current calculated in Sec.~\ref{cd}. From Eq.~\eqref{jmac2} 
-we obtain
-\begin{equation}\label{Pjikn}
-J_{\rma}^{(2,\ell)}(2\go)=-i2\got P_{\rma}(2\go)
-=\frac{e}{\gO}
-\sum_{mn\bfk}
-\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})
-\rho^{(2)}_{nm}(\bfk;2\go)
-,
-\end{equation}
-and using Eqs.~\eqref{rho2} and \eqref{sshgp2} leads to
-\begin{align}\label{Pjikn2}
-\chi^{S,\ell}_{\rma\rmb\rmc}
-&=
-\frac{ie}{A E^{\rmb}_1E^{\rmc}_2 2\got}
-\sum_{mn\bfk}
-\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})
-\rho^{(2)}_{nm}(\bfk;2\got)
-\nonumber \\
-&=
-\frac{e^2}{A\hbar2\got}
-\sum_{mn\bfk}
-\frac{\calv^{\gs,\rma,\ell}_{mn}(\mathbf{k})}
-{\go^\gs_{nm\bfk}-2\got}
-\bigg[
--(B_{nm}^{\rmc}(\bfk,\go))_{;k^{\rmb}}
-\nonumber \\
-&
-+i\sum_\ell\left(r_{n\ell}^{\rmb}B_{\ell m}^{\rmc}(\bfk,\go) -
-  B_{n\ell}^{\rmc}(\bfk,\go) 
-  r_{\ell m}^{\rmb}\right)
-\bigg]
-,
-\end{align}
-which gives the surface-like susceptibility of $\ell$-th layer, where 
-$\calbv^\gs$ is given in Eq.~\eqref{vopii},
-where $A=\gO/d$ is the surface area of the unit
-cell that characterizes the surface of the system.
-Using Eq.~\eqref{rho1} we
-split this equation into
-two contributions from the first and second terms on the right hand side,
-\begin{equation}\label{chii}
-\chi_{i,\rma\rmb\rmc}^{S,\ell}
-=-\frac{e^3}{A\hbar^22\got}\sum_{mn\bfk}
-\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}
-\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}},
-\end{equation} 
-and 
-\begin{equation}\label{chie}
-\chi_{e,\rma\rmb\rmc}^{S,\ell}
-=\frac{ie^3}{A\hbar^22\got}\sum_{\ell mn\bfk}
-\frac{\calv_{mn}^{\gs,\rma,\ell}}{\go^\gs_{nm}-2\got}
-\left(
-\frac{r_{n\ell}^{\rmc} r_{\ell m}^{\rmb} 
-f_{m\ell}}{\go^\gs_{\ell m}-\got}
--\frac{r_{n\ell}^{\rmb} r_{\ell m}^{\rmc} 
-f_{\ell n}}{\go^\gs_{n \ell}-\got}
-\right),
-\end{equation} 
-where $\boldsymbol{\chi}^{S,\ell}_i$
- is related to intraband transitions and
-$\boldsymbol{\chi}^{S,\ell}_e$
-to interband transitions.
-For the generalized derivative in Eq.~\eqref{chii} we use the chain rule 
-\begin{equation}\label{gene2}
-\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}}=
-\frac{f_{mn}}{\go^\gs_{nm}-\got}\left(r_{nm}^\rmb\right)_{;k^{\rmc}}
--\frac{f_{mn}r_{nm}^{\rmb}\gD_{nm}^\rmc}{(\go^\gs_{nm}-\got)^2}
-,
-\end{equation}
-and the following result
-shown in Appendix \ref{gwk},
-\begin{align}\label{eli.13}
-\left(\go^\gs_{nm}\right)_{;k^{\rma}}
-=
-\left(\go^\lda_{nm}\right)_{;k^{\rma}}
-= 
-v_{nn}^{\lda,\rma}-v_{mm}^{\lda,\rma}\equiv\gD_{nm}^{\rma}
-.
-\end{align}
-
-In order to calculate the nonlinear susceptibility of any given layer 
-$\ell$ we simply add the above terms $\boldsymbol{\chi}^{S,\ell}=
-\boldsymbol{\chi}_e^{S,\ell}+\boldsymbol{\chi}_i^{S,\ell}$ and 
-then calculate the surface susceptibility as 
-\begin{equation}\label{chiijksur}
-\bfgchi^S\equiv \sum_{\ell=1}^{N}\bfgchi^{S,\ell},
-\end{equation} 
-where $\ell=1$ is the first layer right at the surface, 
-and $\ell=N$ is the bulk-like layer (at a distance $\sim d$ from the
-surface  as seen in
-Fig.~\ref{fsystem}), such that 
-\begin{equation}\label{chiijksur2}
-\bfgchi^{S,\ell=N}=0,
-\end{equation}
-in accordance to Eq.~\eqref{sshg} valid for a centrosymmetric environment. 
-We note that the value of
-$N$ is not universal.
-This means that the slab needs to have enough atomic layers for 
-Eq.~\eqref{chiijksur2} 
-to be satisfied and to give converged results for $\bfgchi^S$. 
-We can use Eq.~\eqref{chiijksur} for
-either the front or the back surface.
-
-We can see from the prefactors of Eqs. \eqref{chii} and \eqref{chie} 
-that they diverge as $\got\to 0$. To remove this apparent divergence of 
-$\bfgchi^{S,\ell}$, we perform a partial fraction expansion over $\got$. 
-As shown in Appendix \ref{appv}, we use time-reversal invariance to 
-remove these divergences and obtain the following expressions for $\bfgchi^S$,
-\input{appendices/chis}
-\noindent where the limit of $\eta\to 0$ has been taken.
-We have split the interband and intraband $1\go$ and $2\go$
-contributions. The real part of each contribution can be obtained through
-a Kramers-Kronig transformation,\cite{nicolas} and then
-$\chi^{S,\ell}_{\rma\rmb\rmc}=\chi^{S,\ell}_{e,\rma\rmb\rmc,\go} 
-+\chi^{S,\ell}_{e,\rma\rmb\rmc,2\go}+\chi^{S,\ell}_{i,\rma\rmb\rmc,\go}
-+\chi^{S,\ell}_{i,\rma\rmb\rmc,2\go}
-$.
-To fulfill the required intrinsic permutation symmetry,\cite{rashkeevPRB98} 
-the $\{\}$ notation symmetrizes the $\rmb\rmc$ Cartesian indices, i.e. 
-$\{u^{\rmb}s^{\rmc}\}=(u^{\rmb}s^{\rmc}+u^{\rmc}s^{\rmb})/2$,
-and thus
-$\chi_{\rma\rmb\rmc}^{S,\ell}=\chi_{\rma\rmc\rmb}^{S,\ell}$.
-In Appendices \ref{gdernl} and \ref{calvs} we demonstrate how to calculate  
-the generalized derivatives of $\bfr_{nm;\bfk}$ and
-$\calv^{\gs,\rma,\ell}_{nm;\bfk}$. 
-We find that
-\begin{align}\label{rgen.69}
-(r^{\rmb}_{nm})_{;k^{\rma}}
-&=
--i\calt^{\rma\rmb}_{nm}
-+
-\frac{
-r^{\rma}_{nm}
-\Delta^{\rmb}_{mn}
-+r^{\rmb}_{nm}
-\Delta^{\rma}_{mn}
-}
-{\go^\lda_{nm}}
-+
-\frac{i}{\go^\lda_{nm}}
-\sum_{\ell}
-\bigg(
-\go^\lda_{\ell m}
-r^{\rma}_{n\ell}
-r^{\rmb}_{\ell m}
--
-\go^\lda_{n\ell}
-r^{\rmb}_{n\ell}
-r^{\rma}_{\ell m}
-\bigg)
-,
-\end{align}
-where
-\begin{align}\label{tau.1}
-\calt_{nm}^{\rma\rmb}
-=
-[r^{\rma},v^{\lda,\rmb}]= 
-\frac{i\hbar}{m_e}\gd_{ab}\gd_{nm}+
-\call_{nm}^{\rma\rmb}
-,
-\end{align}  
-and
-\begin{align}\label{tau.2}
-\call_{nm}^{\rma\rmb}
-=
-\frac{1}{i\hbar}[r^{\rma},v^{\nl,\rmb}]_{nm}
-,
-\end{align}
-is the contribution to the generalized derivative of $\bfr_{nm}$
-coming from the nonlocal part of the pseudopotential.
-In Appendix \ref{calt} we calculate
-$\call^{\rma\rmb}_{nm}$, that
-is a term with very small numerical value but with a computational time 
-at least an order of magnitude larger
-than for all the other terms involved in the expressions for 
-$\chi^{s,\ell}_{\mathrm{abc}}$.\cite{valerie}
-Therefore, we neglect it throughout this article and take
-\begin{align}\label{tau.69}
-\calt_{nm}^{\rma\rmb}
-\approx
-\frac{i\hbar}{m_e}\gd_{ab}\gd_{nm}
-.
-\end{align} 
-Finally,
-we also need the following term (Eq.~\eqref{ntita})
-\begin{align}\label{a.3c}
-(v^{\lda,\rma}_{nn})_{;k^\rmb}
-=
-\nabla_{k^{\rma}}  
-v^{\lda,\rmb}_{nn}(\bfk)
-&=
--i\calt^{\rma\rmb}_{nn}
--
-\sum_{\ell\ne n}
-\go^\lda_{\ell n}
-\bigg(  
-r^{\rma}_{n\ell}  
-r^\rmb_{\ell n}
-+  
-r^\rmb_{n\ell}  
-r^{\rma}_{\ell n}
-\bigg)
-\nonumber\\
-&\approx
-\frac{\hbar}{m_e}\gd_{\rma\rmb}
--
-\sum_{\ell\ne n}
-\go^\lda_{\ell n}
-\bigg(  
-r^{\rma}_{n\ell}  
-r^\rmb_{\ell n}
-+  
-r^\rmb_{n\ell}  
-r^{\rma}_{\ell n}
-\bigg)
-,
-\end{align}  
-among other quantities for $\calv^{\gs,\rma,\ell}_{nm;\bfk}$, where we 
-also use Eq.~\eqref{tau.69}. Above is the standard effective-mas sum rule.
-\cite{ashcroft_solid_1976} 
-%In Appendix \ref{code}, we list all the quantities that should be
-%coded in order to calculate the previous expressions for $\bfgchi$.
-
-\section{Conclusions}\label{con}
-
-We have presented a complete derivation of the required elements to
-calculate in the independent particle approach (IPA) the microscopic  
-surface second harmonic susceptibility tensor $\bfgchi^S(-2\go;\go,\go)$ 
-using a layer-by-layer approach. We have done so for semiconductors using 
-the length gauge for the coupling of the external electric field to the 
-electron. 
-%Also,
-%we calculated the radiated efficiency $R$ within the three layer
-%model. 
-%The combination of $\boldsymbol{\chi}$ and $R$ allow us
-%to study this fascinating surface optical phenomena.
-
-\appendix 
-\input{appendices/appendix-re_ri}%reri A
-\input{appendices/appendix-vnl}%appvnl B
-\input{appendices/appendix-gen_der_calv}%calvs C 
-\input{appendices/appendix-gen_der_omega}%gwk D
-\input{appendices/appendix-calv}%appv E
-\input{appendices/appendix-tnm}%calt F
-\input{appendices/appendix-vc}%calpcalc G 
-\input{appendices/appendix-gen_der_r_nonlocal}%gdernl H 
-\input{appendices/appendix-code}%code I
-%%%%
-%\input{appendices/appendix-div_free_exp_chi}
-%\input{appendices/appendix-dirac}
-%\input{appendices/appendix-basic_relationships}
-%\input{appendices/appendix-gen_der_r}
-%\input{appendices/appendix-gen_der_calr}
-%\input{appendices/appendix-odds_ends}
-
-%\backmatter
-%\bibliographystyle{prb}
-%\nocite{*} 
-\bibliography{ref}
-
-\end{document}
-
-* smoother C(z)

### 2014-11-16 paper-v-0.tex
english revisions done!

@@
 We present a theoretical review of surface second harmonic generation (SHG) 
-from semiconductor surfaces based on the longitudinal gauge. This 
+from semiconductor surfaces based on the {\color{red} length} gauge. This 
 layer-by-layer analysis is carefully presented in order to show how a 
@@
 We use the independent particle approximation and neglect local field
-and excitonic effects and
-treat the electromagnetic field classically, while the matter is
-described quantum-mechanically.
+and excitonic effects{\color{red}. We assume a classical} electromagnetic field
+{\color{red} and quantum mechanical matter.}
 We can describe the system using 
@@
 transforms any operator $\mathcal{O}$ into 
-$\tilde{\mathcal{O}}=U{\cal O}U^{\dagger }$. Even if $\mathcal{O}$ does not depend on $t$, $\tilde{
-\mathcal{O}}$ does through the explicit time dependence of $U$. 
+$\tilde{\mathcal{O}}=U{\cal O}U^{\dagger }$. Even if $\mathcal{O}$ 
+{\color{red} is time-independent}, 
+$\tilde{\mathcal{O}}$ {\color{red} is time-dependent} through the explicit
+time dependence of $U$. 
 The dynamical 
@@
 $\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle=f(\hbar \omega_{n}(\mathbf{k}))\equiv f_{n}$. For a
-clean, cold semiconductor $f_{n}=1$ for $n$ a valence ($v$) or
-occupied band and zero for $n$ a conduction ($c$) or empty band. \ This we
-assume throughout.
-We remark that the expectation values satisfy
+clean, cold semiconductor $f_{n}=1$ {\color{red} when} $n$ {\color{red} is} a valence ($v$) or
+occupied band, and zero {\color{red} when} $n$ {\color{red} is} a conduction ($c$) or empty band.
+{\color{red} We assume this for the remainder of the article.}
+{\color{red} T}he expectation values must satisfy
 $
@@
 where $\dot{\mathbf{r}}$ is the time derivative of the position operator of the
-electron of charge $e$, 
+electron {\color{red} with} charge $e$, {\color{red} defined as}
 \begin{equation}
@@
 where $\tilde\omega=\omega+i\eta $,
-and $\eta >0$ is used to adiabatically turn on the interaction,
+and $\eta >0$ {\color{red} adiabatically switches on} the interaction,
 we write the second order nonlinear
@@
 where $\chi^{abc}(-2\omega ;\omega ,\omega )$ is the nonlinear
-susceptibility responsible of surface second harmonic generation
+susceptibility responsible {\color{red} for} surface second harmonic generation
 (SSHG). 
@@
 is used to solve the Kohn-Sham equations\cite{kohnPR65} of Density  
-Functional Theory (DFT), for convenience usually within the Local  
-Density Approximation (LDA). Thus, we labeled the hamiltonian with the  
-superscript LDA for convenience, although we could use any other
-approximation, like GGA, and our derivation would remain the same. Then,
+Functional Theory (DFT) {\color{red}. It is convenient to work within the Local 
+Density Approximation (LDA), so we label the hamiltonian with the corresponding  
+LDA superscript. Any other approximation can be used (like GGA) and our 
+derivation remains the same.} Then,
 \begin{equation}\label{h0}
@@
 $\chi^{abc}(-2\omega;\omega,\omega)$, we asume
-that the crystal potential could have a spatially non-local
-contribution, as is customary for most
+that the crystal potential could have a {\color{red} non-local spatial}
+contribution {\color{red} as} is customary for most
 pseudopotentials, and then we replace $V$ with
@@
 from the non-local part of the pseudopotential, $V^{\mathrm{nl}}$, in
-the results that we will derive.
-
-As is well known, the use of the LDA scheme 
+the results that we {\color{red} have derived.}
+
+{\color{red} It} is well known that the use of the {\color{red} LDA} 
 leads to an underestimation of the band gap. A standard
 procedure to correct for this is to
-use the so-called ``scissors approximation'', by which one rigidly shifts the
-conduction bands in energy so that the band gap corresponds to the accepted
-experimental band gap; this is often in fairly good agreement with the GW
+use {\color{red} the} ``scissors approximation'', {\color{red} where the 
+conduction bands are rigidly shifted in energy} so that the band gap 
+corresponds to the accepted experimental band gap{\color{red} .
+This} is often in fairly good agreement with the GW
 band gap based on a more sophisticated calculation.\cite{hybertsenPRB86}
-Concurrently, one uses the LDA wave functions, since they produce band
+{\color{red} T}he LDA wave functions are used since they produce band
 structures with dispersion relations similar to those predicted by the GW
-approximation. Mathematically, one adds the scissors (non-local) term 
-$S$, to the unperturbed or unscissored Hamiltonian $H^{\mathrm{LDA}}_{0}$, i.e. 
+approximation. Mathematically, {\color{red} the scissors (non-local) term 
+$S$ is added} to the unperturbed or unscissored Hamiltonian {\color{red} $H^{\mathrm{LDA}}_{0}$ ,}
 \begin{equation*}
@@
 \end{equation}
-We emphasize that the scissored
-Hamiltonian has the same eigenfunctions as the unscissored
-Hamiltonian,
-where
+We emphasize that the {\color{red} scissored and unscissored Hamiltonian 
+have the same eigenfunctions,} where
 $\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
@@
 To calculate the optical response 
-within the Length Gauge Formalism, the interaction Hamiltonian is given by 
+within the Length Gauge {\color{red} formalism}, we take the 
+interaction Hamiltonian {\color{red} as} 
 \begin{equation}
@@
 treatment of the position operator
-$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$, has been
-described,\cite{adamsJCP53,blountSSP62} and in particular 
-in Ref.~\onlinecite{aversaPRB95} the correct treatment of $\bfr$ for
-the calculation of the non-linear optical response is
-developed. Indeed, 
+$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$ {\color{red} has} 
+been described {\color{red} in detail}.\cite{adamsJCP53,blountSSP62} 
+{\color{red} In particular, Ref.~\onlinecite{aversaPRB95} develops} 
+the correct treatment of $\bfr$ for
+the calculation of the {\color{red} nonlinear} optical {\color{red} response.} 
 $\bfr$ is split into the {\it intraband} 
-($\bfr_i$) and {\it interband} ($\bfr_e$) parts , where 
+($\bfr_i$) and {\it interband} ($\bfr_e$) parts, where 
 $\bfr=\bfr_i+\bfr_e$. 
@@
 \end{align}
-such that $\bfr_{e,nm}=0$ for $n=m$, and thus
-from Eq.~\eqref{mv}, with $H_0\to
+such that $\bfr_{e,nm}=0$ for $n=m${\color{red} .
+From Eq.~\eqref{mv}} with $H_0\to
 H^\sigma_0$, we obtain
@@
 %elements.
-For the intraband part, in turns out that $\bfr_i$ only appears in
-commutators, as one does the derivation of
-the optical response, then one uses,\cite{aversaPRB95}
+For the intraband part, {\color{red} $\bfr_i$ only appears in
+commutators during the derivation of
+the optical response. We use},\cite{aversaPRB95}
 \begin{equation}\label{conmri3n}
@@
 The vectors $\bfgxi_{nn}(\bfk)$ are defined in 
-Ref.~\onlinecite{aversaPRB95}, however they do not need to be 
+Ref.~\onlinecite{aversaPRB95} {\color{red} though} they do not need to be 
 calculated explicitly in what follows.
 
-Before continue, 
-we derive a key result for the length gauge formulation. Using again $H^\sigma_0$ in
+Before {\color{red} continuing,} 
+we derive a key result for the length gauge formulation. 
+{\color{red} Again, using} $H^\sigma_0$ in
 Eq.~\eqref{mv} we obtain
@@
 \end{subequations}
-with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator and used
-$[r^a,p^b]=i\hbar\gd_{ab}$, with $\gd_{ab}$ Kronecker's delta.
+with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator{\color{red}, 
+and using $[r^a,p^b]=i\hbar\gd_{ab}$, where $\gd_{ab}$ is the Kronecker} delta.
 Using Eq.~\eqref{hats} we obtain 
@@
 with $f_{nm}\equiv f_n-f_m$,
-where we see that $\bfv^S_{nn}=0$. Then, it follows that
+where we see that $\bfv^S_{nn}=0$. {\color{red} It} follows that
 \begin{align}\label{chon.9}
@@
 \end{align}
-thus from Eq.~\eqref{pmnrmn} 
+{\color{red} and from Eq.~\eqref{pmnrmn},} 
 \begin{align}\label{chon.10}
@@
 The matrix elements 
-of $\bfr_{nm}(\bfk)$ are the same whether we use the LDA or the scissored 
-Hamiltonian and thus there is no need to label them.
-Of course, is more convenient to calculate them
-through $\bfv^\lda_{nm}(\bfk)$, that 
-include the contribution of 
+of $\bfr_{nm}(\bfk)$ are {\color{red} identical using either} 
+the LDA or {\color{red} scissored} 
+Hamiltonian, {\color{red} thus negating the} need to label them.
+Of course, {\color{red} it} is more convenient to calculate them
+through $\bfv^\lda_{nm}(\bfk)$ {\color{red} which 
+includes} the contribution of 
 the non-local part of the pseudopotentials through
-$\bfv^\nl_{nm}(\bfk)$. These in turn, can be readily
+$\bfv^\nl_{nm}(\bfk)$. These {\color{red} can} be readily
 calculated
@@
 Kleinman-Bylander 
-form,\cite{francesco,mottaCMS10,kleinman_efficacious_1982,adolphPRB96}
-see Appendix \ref{vesnl}.
-Thus, the advantage of the length gauge formalism for linear and 
-non-linear calculations of optical responses resides in the ease with 
-which one can introduce the scissors operator into the calculations as 
-it is also discussed in Nastos et al.\cite{nastosPRB05} Indeed, the 
-length-gauge formalism for the scissored Hamiltonian can be easily worked 
+form,\cite{francesco,mottaCMS10,kleinmanPRL82,adolphPRB96}
+{\color{red} as shown in Appendix \ref{vesnl}.}
+{\color{red} The} advantage of the length gauge formalism for 
+{\color{red} calculating linear and nonlinear} optical responses 
+resides in the ease with {\color{red} which the scissors operator
+can be introduced into the calculation, also described in detail in} 
+Nastos et al.\cite{nastosPRB05} {\color{red} The 
+length gauge} formalism for the scissored Hamiltonian can be easily worked 
 out by simply using the unscissored LDA Hamiltonian,
@@
 for the unperturbed system 
-with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction, and then at the end of the 
-calculation we only need to replace $\omega^{\mathrm{LDA}}_{n}$ with 
-$\omega_{n}^{\sigma}$
- to obtain the 
-scissored results for any susceptibility expression, whether linear or 
-nonlinear.
-
-Therefore,
-taking the matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of
+with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction{\color{red}. 
+We need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
+$\omega_{n}^{\sigma}$ at the end of the derivation}
+to obtain the scissored results for any 
+susceptibility expression, whether linear or nonlinear.
+
+{\color{red} Taking} the matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of
 Eq.~\eqref{rde}, we obtain 
@@
 %$(B^a_{nm})_{;k^b}$ can be obtained.
-Notice that in Eq.~\eqref{j.2} all the indices are different from each
-other since due to the $f_{nm}$ factor in Eq.~\eqref{j.1},
-$B^a_{nn}=0$. The dependence of $\bfk$ of all quantities is implicitly understood from
-here on. 
-Above expression will be used in the next subsection.
+{\color{red} Note that the indices in Eq.~\eqref{j.2} are all} different from each
+other. {\color{red} This is due to the $f_{nm}$ factor in Eq.~\eqref{j.1}, 
+and therefore $B^a_{nn}=0$}. The dependence {\color{red} on} $\bfk$ 
+of all quantities is implicitly understood from 
+{\color{red} this point forward.}
 
@@
 semiconductor crystal is as follows. Instead of using a
-semi-infinite system, we replace it by a super cell that consist of a finite
+semi-infinite system, we replace it by a super cell that 
+{\color{red} consists} of a finite
 slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
-super cell, which constitutes the unit cell,
+super cell {\color{red} constitutes the unit cell, and}
 is repeated to form a full three dimensional crystalline structure.
-The slab itself consists of a front and back surface and sub-surface regions, and in between these
-a region that is equivalent to
+The slab itself consists of {\color{red} front, back, and 
+sub-surface} regions, and in between these
+a region that is equivalent to {\color{red} the}
 bulk of the system. 
@@
 balanced forces are disrupted when the surface atoms do not find their 
-partner atoms that are now absent at the surface of the slab. 
-To take the reconstruction or relaxation into account, we take ``surface'' to mean
-the true surface of the first layer of atoms, and
+partner atoms that are now absent at the {\color{red} slab surface.} 
+To take the reconstruction or relaxation into account, 
+we take ``surface'' to mean
+the true surface of the first layer of {\color{red} atoms} and
 some of the atomic sub-layers adjacent to it.
@@
 centrosymmetric. This would imply that 
-$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=0$, and thus we must
+$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=0$ {\color{red} so} we must
 find a scheme 
@@
 are different, breaking the centrosymmetry and therefore giving an
-overall $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}\ne 0$, we still
+overall $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}\ne 0${\color{red} ;} we still
 need a procedure to extract the front surface $\chi^f_{\rma\rmb\rmc}$
@@
 $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=\chi^f_{\rma\rmb\rmc}-\chi^b_{\rma\rmb\rmc}$.
-We have omitted the frequency arguments of $\chi_{abc}$ for notation convenience.
+We have omitted the frequency arguments of $\chi_{abc}$ for 
+{\color{red} convenience of} notation.
 
@@
 usually taken to be unity over one half of the slab and zero over 
-the other half.\cite{reining_microscopic_1994}
+the other half.\cite{reiningPRB94}
 In this case $\calc(z)$ will give the contribution of the 
@@
 $(f)$ or back $(b)$ from its $z_\ell$ position. 
- We take $z_\ell$
- at the center of an atom that 
-belongs to layer $\ell$, and thus above equation would give the $\ell$-th 
- atomic-layer 
-contribution to the non-linear optical response.
+We take $z_\ell$ {\color{red} to be}at the center of an atom that 
+belongs to layer $\ell${\color{red} , so the previous} 
+equation would give the $\ell$-th atomic-layer 
+contribution to the {\color{red} nonlinear} optical response.
 $\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
@@
 \includegraphics[scale=.7]{images/slab}
-\caption{A sketch of the super cell, where the slab corresponds to the
+\caption{A sketch of the super {\color{red} cell. 
+The} slab corresponds to the
 circles representing the atoms of the system.\label{fslab}} 
@@
 the calculation of $\chi_{\rma\rmb\rmc}$, we start from 
-the
-operator for the electron current,
+the operator for the electron current,
 $\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gs\ket{\bfr}\bra{\bfr}
-+\ket{\bfr}\bra{\bfr}\bfv^\gs\right)
-$
-, 
-that leads to
++\ket{\bfr}\bra{\bfr}\bfv^\gs\right)${\color{red} , that} leads to
 \begin{align}\label{jmic}
@@
 If we want the contribution from only one region of the unit cell 
-towards the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
-over the desired region. Then, the contribution to the current density from the
+{\color{red} towards??} the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
+over the desired region. {\color{red} Then the} contribution 
+to the current density from the
 $\ell$-th layer of the slab is given by
@@
 is all that is needed to change any of the
-velocity operators of the electron $\bfV$ to the new velocity
+{\color{red} electron} velocity operators $\bfV$ to the new velocity
 operator $\calbv^{\ell}$ that implicitly takes into account the
@@
 (\go^\gs_{nm}/\go^\lda_{nm})
-\calbv^{\ell,\lda}_{nm}(\bfk)$,
-and thus, to calculate
+\calbv^{\ell,\lda}_{nm}(\bfk)${\color{red} .
+To calculate}
 $\calbv^{\ell,\sigma}_{nm}(\bfk)$ 
@@
 for $\calbp^\ell$ was proposed in 
-Ref.~\onlinecite{reining_microscopic_1994} and later used in Refs.
+Ref.~\onlinecite{reiningPRB94} and later used in Refs.
 \onlinecite{mendozaPRL98},
-\onlinecite{mendoza_ab_2001},
+\onlinecite{mendozaPRB01},
 \onlinecite{sanoPRB02},
- and \onlinecite{mejia_layer-by-layer_2004}. 
-The layer-by-layer analysis of Refs. \onlinecite{hogan_optical_2003,castilloPRB03,mottaCMS10} 
+ and \onlinecite{mejiaRMF04}. 
+The layer-by-layer analysis of Refs. \onlinecite{hoganPRB03,castilloPRB03,mottaCMS10} 
 used the equivalent Eq.~\eqref{sz}, 
@@
 cell that characterizes the surface of the system, and $d$
-is the surface region, from which the non-linear susceptibility is different from zero.
+is the surface region {\color{red} from} which the {\color{red} nonlinear} 
+susceptibility is different from zero.
 We mention that the units of 
 $\chi^{\ell,\rma\rmb\rmc}(-2\go;\go,\go)$
-are m$^2$/V, as they should for a surface SH susceptibility.
+are m$^2$/V, as they {\color{red} should be} for a surface SH susceptibility.
 Using Eq.~\eqref{j.1} we
@@
 two contributions from the first and second terms on the right hand side
-of Eq.~\eqref{Pjikn2}, then,
+{\color{red} of Eq.~\eqref{Pjikn2}:}
 %\end{document}  
@@
 \end{equation} 
-that  is related to intraband transitions, and 
+{\color{red} related} to intraband transitions, and 
 \begin{equation}\label{chie}
@@
 \end{equation} 
-that is related to interband transitions.
+{\color{red} related} to interband transitions.
 The generalized derivative in Eq.~\eqref{chii} is dealt with by the chain rule 
@@
 \end{equation}
-where by using $H^\sigma_0$ into Eq.~\eqref{conmri3n} and then
+where {\color{red} substituting} $H^\sigma_0$ 
+into Eq.~\eqref{conmri3n} and then
 Eq.~\eqref{chon.9}
- one can readily
-show that
+{\color{red} we obtain}
 \begin{align}\label{eli.13}
@@
 \end{subequations}
- where we have splited the interband and intraband $1\go$ and $2\go$
-contributions and supres the $\go$ arguments for space sake.
+where we have {\color{red} split} the interband and intraband $1\go$ and $2\go$
+contributions and {\color{red} supressed} the $\go$ arguments for 
+{\color{red} convenience of notation}.
 The real part of each contribution can be obtained through
-a Kramers-Kronig transformation,\cite{nicolas} and then
+a Kramers-Kronig transformation\cite{nicolas} {\color{red} and}
 $\chi^{\ell,\rma\rmb\rmc}=
@@
 where $\{\ell\}$
-is meant to be set of layers that one choses. For instance, one go
-from 
-a layer that encompasses half of the slab  or each atomic layer till
-the middle of the slab. In the first case, there is one single summand
-in Eq.~\eqref{chiijksur}, in the second case a sum from
-$\ell=1$ which denotes
-the first layer right at the surface, 
-till $\ell=N$ that corresponds to the layer at the middle of the salb, which ought to be
-a bulk-like layer.
+is meant to be {\color{red} a chosen set of layers.} For instance, 
+one {\color{red} can take a single layer 
+encompassing} half of the slab, or {\color{red} take each 
+atomic layer individually to the middle} 
+of the slab. {\color{red} For} the first case there is 
+{\color{red} a} single summand
+in Eq.~\eqref{chiijksur}{\color{red} . For the second case
+there is} a sum from $\ell=1${\color{red} , denoting the first layer 
+right at the surface, to $\ell=N$, denoting the layer at the middle of the slab 
+that behaves like a bulk layer.}
 %that at a distance $d$ from the surface  as seen in Fig.~\ref{fslab}.  
@@
 %$\bfgchi^{\ell=N}=0$;
-We reamark that the value of 
-$N$ is not universal, and  
-the slab needs to have enough atomic layers for 
+We {\color{red} remark} that the value of 
+$N$ is not universal {\color{red} and} 
+the slab needs to have enough atomic layers 
 %$\bfgchi^{\ell=N}=0$;
 %to be satisfied and to 
-in order to give converged results for $\bfgchi (-2\go;\go,\go)$. 
+{\color{red} in} order to give converged results for 
+$\bfgchi (-2\go;\go,\go)$. 
 We can use Eq.~\eqref{chiijksur} for 
@@
 $e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
-space, and $A_{n\bfk}(\bfG)$ are the expansion coefficients. Using
-$m_e\bfv=-i\hbar\bfgnabla$ into the equivalent of Eq.~\eqref{intj}
+space, and $A_{n\bfk}(\bfG)$ {\color{red} the} expansion coefficients. Using
+$m_e\bfv=-i\hbar\bfgnabla$ into the {\color{red} equivalent??} 
+of Eq.~\eqref{intj}
 we obtain,\cite{mendozaPRB06}
@@
 with $f^{\ell*}(g)=f^\ell(-g)$. 
-the reciprocal lattice vectors $\bfG$ are decomposed into components
-parallel to the surface $\bfG_\parallel$, and perpendicular to the
-surface $G_\perp \hat z$, so
+{\color{red} The} reciprocal lattice vectors $\bfG$ are 
+decomposed into components
+{\color{red} parallel ($\bfG_\parallel$), and perpendicular ($G_\perp \hat z$)} 
+to the surface, so
 that $\bfG = \bfG_\parallel + G_\perp\hat z$.
-The double summation over the $\bfG$ vectors can be efficiently done by  
+The double summation over the $\bfG$ vectors can be 
+{\color{red} calculated efficiently} by  
 creating a pointer array to identify all the plane-wave coefficients  
@@
 \end{align}  
-If $\calc^\ell(z)=1$, $f^\ell(g)=\gd_{g0}$, and from Eqs.~\eqref{eni.2} and \eqref{eni.4}
-we obtain the full-slab or bulk values
-$\bfv_{nm}(\bfk)$ and
-$\calc^\ell_{nm}(\bfk)=\gd_{nm}$.
-
-Since for any function $F(\bfr)$, $[\bfr,F(\bfr)]=0$, using 
-Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn}, we obtain
+If $\calc^\ell(z)=1$, {\color{red} then $f^\ell(g)=\gd_{g0}$ and we 
+obtain the full-slab/bulk values, 
+$\bfv_{nm}(\bfk)$ and $\calc^\ell_{nm}(\bfk)=\gd_{nm}$,
+from Eqs.~\eqref{eni.2} and \eqref{eni.4}}.
+
+{\color{red} For} any function $F(\bfr)$, $[\bfr,F(\bfr)]=0$, 
+{\color{red} we can use}
+Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn} 
+{\color{red} to }obtain
 \begin{align} 
@@
 \end{align} 
-where we remind the reader that $\bfr_{nm}$ are calculated through
+where we remind the reader that $\bfr_{nm}$ 
+{\color{red} is} calculated through
 Eq.~\eqref{chon.10} for LDA.
@@
 for non-local potentials}\label{appvnl}
-Using Eqs.~\eqref{rnminn}, \eqref{conmri3n}, \eqref{gendevnn} and
+
+Using Eqs.~\eqref{rnminn}, \eqref{conmri3n}, \eqref{gendevnn}, and
 defining 
@@
 $
-one can show that
+one can show that{\color{red} 
 \begin{align}\label{nmesn}
@@
 r^{\rmb}_{nm}
-\Delta^{ell,\rma}_{mn}
-,
-\end{align}  
+\Delta^{\ell,\rma}_{mn}
+,
+\end{align}}
 where
@@
 =(v^{\lda,\rma}_{nm}/i\go^\lda_{nm})_{;k^{\rmb}}
-$, and apply the chain rule. 
-Then, 
-$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$ is obtained
+$ {\color{red} and} apply the chain rule. 
+{\color{red} 
+$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$} is obtained
 from Eq.~\eqref{nmesn} by
 taking
-$C^\ell(z)=1$ or $C^\ell_{nm}=\gd_{nm}$. Thus,
+{\color{red} $C^\ell(z)=1$ or $C^\ell_{nm}=\gd_{nm}$,}
 \begin{align}\label{na_rgendevn}
@@
 \end{eqnarray}
-and
+and{\color{red} 
 \begin{align}\label{tau.1n} 
@@
 -i [r^{\rmb},v^{\nl,\rma}]_{nm} 
-.
-\end{align}   
+,
+\end{align}}
 generalizes the usual expresion of
-$(r^a_{nm})_{;k^b}$ for local 
-Hamiltonians,\cite{aversaPRB95,nastosPRB05,cabellosPRB09,rashkeevPRB98}
+$(r^a_{nm})_{;k^b}$ for {\color{red} a} local 
+{\color{red} Hamiltonian}
+\cite{aversaPRB95,nastosPRB05,cabellosPRB09,rashkeevPRB98}
 to
@@
 Note that the layered term
-$\calt^{\ell,\rma\rmb}_{nm}$ reduces to $t^{\rma\rmb}_{nm}$, for the
-full-slab or bulk case.
+$\calt^{\ell,\rma\rmb}_{nm}$ reduces to $t^{\rma\rmb}_{nm}$
+{\color{red} for the full-slab/bulk case.}
 
@@
 \texorpdfstring{$\calbv^{\ell,\nl}$}{Vnonlocal},
-\texorpdfstring{$\bfv^\nl$}{Vnonlocal}
-and $[\bfr,\bfv^\nl]$}\label{vesnl}
+\texorpdfstring{$\bfv^\nl$}{Vnonlocal}, and 
+\texorpdfstring{$[\bfr,\bfv^\nl]$}{[r,vnl]}}
+\label{vesnl}
 
@@
 $\braket{\bfr}{\bfk+\bfG}=(1/\sqrt{\gO})
-\mathrm{exp}(i(\bfk+\bfG)\cdot\bfr)$,
-to obtain,\cite{mottaCMS10}
+\mathrm{exp}(i(\bfk+\bfG)\cdot\bfr)$
+{\color{red} to} obtain,\cite{mottaCMS10}
 \begin{align}\label{vnl.5}
@@
 For fully  separable pseudopotentials in the   
-Kleinman-Bylander (KB) form,\cite{mottaCMS10,kleinman_efficacious_1982,adolphPRB96}  
+Kleinman-Bylander (KB) form,\cite{mottaCMS10,kleinmanPRL82,adolphPRB96}  
 the  
@@
 =V^\nl(\bfK,\bfK')  
-$, 
-along with their $\bfK$ and $\bfK'$ gradient 
+$ 
+{\color{red} and} their $\bfK$ and $\bfK'$ gradient 
 can be readily calculated.\cite{mottaCMS10,adolphPRB96,gordienkoRPJ04,fuchsCPC99} 
@@
 implemented 
-the calculation of $\calbv^{\ell,\nl}_{nm}(\bfk)$ with the help pf the  \depe~code.\cite{francesco}
-Taking $C^\ell(z)=1$, which implies $f^\ell(g)=\gd_{g0}$, form above three expressions we obtain 
-$\bfv^{\nl}_{nm}(\bfk)$.
+the calculation of $\calbv^{\ell,\nl}_{nm}(\bfk)$ with the help {\color{red} of} 
+the \depe~code.\cite{francesco}
+{\color{red} We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
+expressions by t}aking $C^\ell(z)=1$, which implies 
+{\color{red} that} $f^\ell(g)=\gd_{g0}$.
 
@@
 \end{align} 
-that uppon expanding the triple commutator, leads to
+{\color{red} and we expand the triple commutator,}
 \begin{align}\label{3.5}
@@
 The double derivatives with respect to $\bfK$ and $\bfK'$ 
-can be worked out explicitly, and thus ${\cal L}^{\rma\rmb}_{nm}(\bfk)$
-could be calculated.\cite{valerie}
+can be worked out explicitly {\color{red} and} 
+${\cal L}^{\rma\rmb}_{nm}(\bfk)$
+{\color{red} can} be finally calculated.\cite{valerie}
 
@@
 }\label{calvs}
+
 From Eq.~\eqref{vcali}
@@
 \end{align}    
-where $\sum_q\ket{q\bfk}\bra{q\bfk}=1$ was used,
-and $\bfv^S_{nm}$ are given in Eq.~\eqref{chon.2}.
-Taking the generalized derivative of Eq.~\eqref{a.3b}, and applying
+where $\sum_q\ket{q\bfk}\bra{q\bfk}=1$ was used
+{\color{red} and $\bfv^S_{nm}$ is} given in Eq.~\eqref{chon.2}.
+Taking the generalized derivative of Eq.~\eqref{a.3b}
+{\color{red} and} applying
 the chain rule, we obtain
@@
 \end{align}    
-where from 
+{\color{red} and considering} 
 Eq.~\eqref{chon.2}, 
@@
 \end{align}
-in agreement with Eq. A(6) of Ref.~\onlinecite{cabellosPRB09}.
+{\color{red} is in total} 
+agreement with Eq. A(6) of Ref.~\onlinecite{cabellosPRB09}.
 
@@
 downer_optical_2001,
-mendoza_ab_2001,
+mendozaPRB01,
 lim_optical_2000,
 gavrilenko_optical_2000,
-mendoza_visible-infrared_1999,
+mendozaPRB99,
 mendozaPRL98,
 mendozaPRB96,
-mendoza_polarizable-bond_1997,
+mendozaPRB97,
 guyotPRB90}

### 2014-11-17 res.tex

In this section we present a test case in order to check the
consistency of the approach presented above. To do this end,  we have chosen
a clean Si(111) surface  with a $2\times 1$ surface reconstruction.
The slab for such a surface could be chosen to be centrosymmetric 
by having the fron and the back surface with the same $2\times 1$
reconstruction, however we terminate with hydrogen one of the
surfaces, thus having it as an ideally bulk terminated Si
surface. Indeed, the H atoms simply saturate the dangling bonds of the
bulk-like Si atoms at the surface, as seen in Fig.~\ref{si2x1}. \textcolor{red}{details of the
 relaxation and some references}
\begin{figure}
\centering 
\includegraphics[scale=.3]{images/si2x1}
\caption{The slab shows a front clean Si(111)$2\times 1$ surface,
  while the back surface is ideally terminated Si bulk, where the
  dangling bonds are H (small balls) saturated. This picture shows 12
  Si layers and one H layer. 
\label{si2x1}} 
\end{figure}
The idea for such an slab is that the cristalline symmetry of the
H terminated surface imposes that $\chi^{\mathrm{H}}_{xxx}=0$, while
for the $2\times 1$ surface there is not such symmetry restriction and
then, $\chi^{2\times 1}_{xxx}\ne 0$.
Therefore, due to this fact, calculating $\chi_{xxx}$ for the full-slab
or the half-salb that contains the $2\times 1$ surface,
 ought to give the same result, since the contribution from the H
 saturated  surface is zero any way. Then, for this slab one must check
 whether or not this relation is satisfied,
\begin{align}\label{hs}
\chi^{\mathrm{half-slab}}_{xxx}(-2\go;\go,\go)
=
\chi^{\mathrm{full-slab}}_{xxx}(-2\go;\go,\go)
.
\end{align}
Indeed, in what follows we show the results of such a comparison.

The $2\times 1$ and the H surfaces are relaxed\textcolor{red}{Nicolas
  some details}, where we find agreement with previous results for
both surfaces.\cite{relax} For instance the Si-H bond distance is 1.48 $\AA$.
The calculation of the LDA is done using the ABINIT code with ???
pseudopotentials, that are of the BK form, and thus the $\bfv^\nl$
contribution can be calculated. We use the DP code for this end.

### 2014-11-17 intro.tex
added intro.tex

\section{Introduction}\label{intro}

%check past tense
%nonlocal, not non-local

Nonlinear optics has been a relevant topic in physics research for the past 
fifty years and has gained a wide range of applications in 
diverse areas of technology and research. Second-harmonic generation (SHG) is
one of the most fundamental nonlinear processes, and is of particular 
importance in material science where it is has been proven to be very 
effective as a surface probe. \cite{hughesPRB96, guyot-sionnestPRB88, 
lupkeSSR99, downerPSSA01, shenAPB99, shenNAT89, chenPRL81, 
maytorenaPRB98} Surface spectroscopy is a crucial tool for research in 
microelectronics, semiconductors, nanomaterials, and many more recent 
topics of great scientific and commercial interest.

Numerous theoretical models have been developed to study the nonlinear optical
response of materials -- each with various disadvantages. We recognize two 
formalisms in this context. The first, called the ``velocity gauge'' or 
``$\mathbf{v}\cdot\mathbf{A}$ gauge,'' features in early research in the field. 
Butcher and McLean,\cite{butcherPOPS63}, and Aspnes\cite{aspnesPRB72} laid the 
groundwork for the theory. They derived expressions for the nonlinear 
susceptibility tensor that were either diverged at low frequencies or were 
limited to specific crystal classes. Two decades later, 
Ghahramani \emph{et al}.\cite{ghahramaniPRB91} realized a full-band-structure 
calculation for Si/Ge superlattices and found significant surface SHG 
while providing further proof that the expressions do not necessarily diverge 
at low frequencies for certain materials. Sipe \emph{et al}.\cite{sipePRB93}
distinguish both interband and intraband contributions to the susceptibility 
tensor and further verify the non-divergent behavior. Levine\cite{levinePRB94} 
introduces the scissors operator within the LDA and finds good agreement 
with experimental results. The second formalism, called the ``length gauge'' 
or ``$\mathbf{r}\cdot\mathbf{E}$ gauge'' presents various advantages over 
the velocity gauge. Aversa and Sipe\cite{aversaPRB95} presented divergence 
free expressions for the susceptibility tensor using the length gauge that 
are  significantly easier to derive and calculate. Cabellos \emph{et al}.
\cite{cabellosPRB09} provide a detailed derivation of the length gauge 
formalism and a comparison with the velocity gauge for bulk GaAs. They 
prove that the results are indeed gauge-invariant. This presents a strong
case for using the length over the velocity gauge in order to more 
efficiently calculate the nonlinear response of a system.

We are also concerned with the optical response of surfaces and interfaces.
Mendoza and Moch\'an\cite{mendozaPRB96} produced an early review of 
theoretical SHG spectra from surfaces. Arzate and Mendoza\cite{arzatePRB01}
went into a detailed analysis of the different contributions to the SHG spectra
of a surface and the nuanced relationship between bulk, surface, interband, 
intraband, $1\omega$ and $2\omega$ terms. Calculated surface SHG spectra are 
in good agreement with experiment using both the velocity gauge
\cite{mendozaPRB01,mendozaPRL98,reiningPRB94} and the length-gauge 
approach\cite{sharmaPRB03}. Perhaps of more relevance to this 
work, Mendoza \emph{et al}.\cite{mendozaPRB06} and 
Mejía \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis of 
the linear and nonlinear responses of semiconductor systems. This model
allows for obtaining results from selected regions within a slab or bulk,
including the surface.

The inclusion of nonlocal effects, such as the scissors correction, has been
correctly introduced in both the velocity and length guage schemes. Both 
Rashkeev \emph{et al}.,\cite{rashkeevPRB98} and Hughes and 
Sipe\cite{hughesPRB96} effectively introduce the scissors correction when 
calculating the nonlinear susceptibility tensor. Leitsmann 
\emph{et al}.\cite{leitsmannPRB05} includes excitonic and local field 
interactions within the velocity gauge scheme. 
Lastly, Nastos \emph{et al}.\cite{nastosPRB05} demonstrated the correct 
way to calculate the nonlinear response using the scissored Hamiltonian 
providing better agreement with experiment.

The most complete approach will be a combination of the different techniques 
that lead to more efficient and accurate calculations. Therefore,
the goal of this paper is to derive new expressions for the nonlinear
susceptibility tensor $\chi^{abc}(-2\omega;\omega,\omega)$ in the length
gauge formalism, that includes nonlocal terms from the scissors approximation 
and the nonlocal spatial contribution from the pseudopotentials, and a 
layer-by-layer approach convenient for surface SHG calculations. We are also
able to divide the resulting tensor in interband, intraband, $1\omega$, and 
$2\omega$ contributions. This approach is unprecedented and yet produces 
results that are in total agreement with the literature.

The paper is organized as follows. In Sec. \ref{theory} we present the 
relevant steps for deriving the surface second-order susceptibility tensor 
within the length gauge formalism. This derivation includes the addition of 
nonlocal terms and a layer-by-layer analysis that have been omitted 
from previous works. In Sec. \ref{results} we show the results of this 
reformulation with a layer-by-layer study of a clean Si(111) surface with a 
$2 \times 1$ reconstruction. These calculations are done with ab initio 
programs based on density-functional theory (DFT) within the local-density 
approximation (LDA), using Troullier-Martins pseudopotentials. We compare 
results from before and after adding the different nonlocal contributions.
Finally, in Sec. \ref{conclusions} we present our conclusions.

### 2014-11-17 intro.tex
finished first draft of intro

@@
 \section{Introduction}\label{intro}
-
-%check past tense
 %nonlocal, not non-local
@@
 fifty years and has gained a wide range of applications in 
-diverse areas of technology and research. Second-harmonic generation (SHG) is
-one of the most fundamental nonlinear processes, and is of particular 
-importance in material science where it is has been proven to be very 
-effective as a surface probe. \cite{hughesPRB96, guyot-sionnestPRB88, 
-lupkeSSR99, downerPSSA01, shenAPB99, shenNAT89, chenPRL81, 
-maytorenaPRB98} Surface spectroscopy is a crucial tool for research in 
-microelectronics, semiconductors, nanomaterials, and many more recent 
-topics of great scientific and commercial interest.
+many areas of technology and research. In material science,
+Second-harmonic generation (SHG) is of particular importance.
+It was one of the first nonlinear processes to be observed and is now one
+of the most used. It has proven to be very effective as an 
+optical probe for material surface and interfaces. 
+\cite{hughesPRB96, 
+     guyot-sionnestPRB88, 
+     lupkeSSR99, 
+     downerPSSA01, 
+     shenAPB99, 
+     shenNAT89, 
+     chenPRL81, 
+     maytorenaPRB98}
+SHG surface spectroscopy is a crucial tool for research and development
+in microelectronics, semiconductors, nanomaterials, and many more recent 
+areas of great scientific and commercial interest.
 
-Numerous theoretical models have been developed to study the nonlinear optical
-response of materials -- each with various disadvantages. We recognize two 
-formalisms in this context. The first, called the ``velocity gauge'' or 
-``$\mathbf{v}\cdot\mathbf{A}$ gauge,'' features in early research in the field. 
-Butcher and McLean,\cite{butcherPOPS63}, and Aspnes\cite{aspnesPRB72} laid the 
-groundwork for the theory. They derived expressions for the nonlinear 
-susceptibility tensor that were either diverged at low frequencies or were 
-limited to specific crystal classes. Two decades later, 
-Ghahramani \emph{et al}.\cite{ghahramaniPRB91} realized a full-band-structure 
-calculation for Si/Ge superlattices and found significant surface SHG 
-while providing further proof that the expressions do not necessarily diverge 
-at low frequencies for certain materials. Sipe \emph{et al}.\cite{sipePRB93}
-distinguish both interband and intraband contributions to the susceptibility 
-tensor and further verify the non-divergent behavior. Levine\cite{levinePRB94} 
-introduces the scissors operator within the LDA and finds good agreement 
-with experimental results. The second formalism, called the ``length gauge'' 
-or ``$\mathbf{r}\cdot\mathbf{E}$ gauge'' presents various advantages over 
-the velocity gauge. Aversa and Sipe\cite{aversaPRB95} presented divergence 
-free expressions for the susceptibility tensor using the length gauge that 
-are  significantly easier to derive and calculate. Cabellos \emph{et al}.
-\cite{cabellosPRB09} provide a detailed derivation of the length gauge 
+Numerous mathematical frameworks have been developed to study the 
+nonlinear optical response of materials; 
+in the context of this paper, we note two in particular. 
+The first, called the ``velocity gauge'' or 
+``$\mathbf{v}\cdot\mathbf{A}$ gauge,'' features in early 
+research in this field. 
+Butcher and McLean\cite{butcherPOPS63}, and 
+Aspnes\cite{aspnesPRB72} 
+laid the initial groundwork by deriving explicit expressions for the 
+nonlinear susceptibility tensor. However, these either diverged at 
+low frequencies or were limited to zincblende crystal structures. 
+A few decades later, 
+Ghahramani \emph{et al}.\cite{ghahramaniPRB91} 
+realized a full-band-structure calculation for Si/Ge superlattices and 
+found significant surface SHG while providing proof that the expressions 
+do not diverge at low frequencies for clean, cold semiconductors. 
+Sipe \emph{et al}.\cite{sipePRB93}
+separated the interband and intraband one-electron contributions 
+of the susceptibility tensor and further verified the non-divergent behavior. 
+Levine\cite{levinePRB94} 
+introduced the scissors operator within the Local Density Approximation
+(LDA) and found good agreement with experimental results. 
+The second, called the ``length gauge'' or 
+``$\mathbf{r}\cdot\mathbf{E}$ gauge'' has various advantages over 
+the velocity gauge. 
+Aversa and Sipe\cite{aversaPRB95} 
+presented divergence free expressions for the susceptibility tensor using 
+the length gauge that are significantly easier to derive and calculate. 
+Cabellos \emph{et al}.\cite{cabellosPRB09} 
+elaborated a detailed derivation of the length gauge 
 formalism and a comparison with the velocity gauge for bulk GaAs. They 
-prove that the results are indeed gauge-invariant. This presents a strong
-case for using the length over the velocity gauge in order to more 
-efficiently calculate the nonlinear response of a system.
+proved that the results are indeed gauge-invariant. 
+This presents a strong case for using the length over the velocity gauge 
+for more generalized, and less expensive calculations of the nonlinear
+response of a system.
 
-We are also concerned with the optical response of surfaces and interfaces.
+The inclusion of nonlocal effects, such as the scissors correction, has been
+correctly included in both the velocity and length guage schemes. 
+In the former, both 
+Rashkeev \emph{et al}.\cite{rashkeevPRB98} and 
+Hughes and Sipe\cite{hughesPRB96} 
+accounted for the scissors correction when calculating the nonlinear 
+susceptibility tensor.
+Leitsmann \emph{et al}.\cite{leitsmannPRB05} 
+included excitonic and local field interactions working with GaAs.
+Nastos \emph{et al}.\cite{nastosPRB05} demonstrated the correct 
+way to calculate the nonlinear response using the scissored Hamiltonian 
+in the length gauge, providing better agreement with experiment.
+
+Concerning the optical response of surfaces and interfaces,
 Mendoza and Moch\'an\cite{mendozaPRB96} produced an early review of 
-theoretical SHG spectra from surfaces. Arzate and Mendoza\cite{arzatePRB01}
-went into a detailed analysis of the different contributions to the SHG spectra
-of a surface and the nuanced relationship between bulk, surface, interband, 
-intraband, $1\omega$ and $2\omega$ terms. Calculated surface SHG spectra are 
-in good agreement with experiment using both the velocity gauge
-\cite{mendozaPRB01,mendozaPRL98,reiningPRB94} and the length-gauge 
-approach\cite{sharmaPRB03}. Perhaps of more relevance to this 
-work, Mendoza \emph{et al}.\cite{mendozaPRB06} and 
-Mejía \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis of 
+theoretical SHG spectra from surfaces. 
+Arzate and Mendoza\cite{arzatePRB01} improved upon that work and
+went into a detailed analysis of the different contributions to the 
+SHG spectra of a surface and the nuanced relationship between bulk, 
+surface, interband, intraband, $1\omega$ and $2\omega$ terms. 
+Calculated surface SHG spectra using these schemes coincide with experimental
+results in both the velocity gauge
+\cite{mendozaPRB01,mendozaPRL98,reiningPRB94} and the length gauge approach
+\cite{sharmaPRB03}. 
+Perhaps of more relevance to this work, 
+Mendoza \emph{et al}.\cite{mendozaPRB06} and 
+Mejía \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
 the linear and nonlinear responses of semiconductor systems. This model
@@
 including the surface.
-
-The inclusion of nonlocal effects, such as the scissors correction, has been
-correctly introduced in both the velocity and length guage schemes. Both 
-Rashkeev \emph{et al}.,\cite{rashkeevPRB98} and Hughes and 
-Sipe\cite{hughesPRB96} effectively introduce the scissors correction when 
-calculating the nonlinear susceptibility tensor. Leitsmann 
-\emph{et al}.\cite{leitsmannPRB05} includes excitonic and local field 
-interactions within the velocity gauge scheme. 
-Lastly, Nastos \emph{et al}.\cite{nastosPRB05} demonstrated the correct 
-way to calculate the nonlinear response using the scissored Hamiltonian 
-providing better agreement with experiment.
+More recently, 
+Trolle \emph{et al}.\cite{trollePRB14} 
+elaborate on a new theoretical model for surface SHG of 
+mono-layer MoS$_{2}$. Their approach is based on the tight-binding model 
+fitted to GW results and include excitonic effects from the 
+Bethe-Salpeter exciton picture. However, this new model is not 
+\emph{ab initio} and has no consideration for a layered approach.
 
@@
 that lead to more efficient and accurate calculations. Therefore,
-the goal of this paper is to derive new expressions for the nonlinear
-susceptibility tensor $\chi^{abc}(-2\omega;\omega,\omega)$ in the length
-gauge formalism, that includes nonlocal terms from the scissors approximation 
-and the nonlocal spatial contribution from the pseudopotentials, and a 
-layer-by-layer approach convenient for surface SHG calculations. We are also
+the goal of this paper is to present newly derived expressions for the 
+nonlinear susceptibility tensor $\chi^{abc}(-2\omega;\omega,\omega)$ 
+in the length gauge formalism, 
+that include nonlocal terms from the scissors approximation 
+and the spatial contribution from the pseudopotentials, and a 
+layer-by-layer approach convenient for surface SHG calculations. We are
 able to divide the resulting tensor in interband, intraband, $1\omega$, and 
@@
 
-The paper is organized as follows. In Sec. \ref{theory} we present the 
+The paper is organized as follows. 
+In Sec. \ref{theory} we present the 
 relevant steps for deriving the surface second-order susceptibility tensor 
 within the length gauge formalism. This derivation includes the addition of 
-nonlocal terms and a layer-by-layer analysis that have been omitted 
-from previous works. In Sec. \ref{results} we show the results of this 
+the nonlocal terms mentioned above and a layer-by-layer analysis that have 
+been absent in previous works. 
+In Sec. \ref{results} we show the results of this 
 reformulation with a layer-by-layer study of a clean Si(111) surface with a 
-$2 \times 1$ reconstruction. These calculations are done with ab initio 
+$2 \times 1$ reconstruction. These calculations are done with \emph{ab initio}
 programs based on density-functional theory (DFT) within the local-density 
@@
 results from before and after adding the different nonlocal contributions.
-Finally, in Sec. \ref{conclusions} we present our conclusions.
+Finally, in Sec. \ref{conclusions} we submit our conclusions.

### 2014-11-17 intro.tex
corrected accent on Mejia in intro

@@
 Mendoza \emph{et al}.\cite{mendozaPRB06} and 
-Mejía \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
+Mej\'ia \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
 the linear and nonlinear responses of semiconductor systems. This model

### 2014-11-17 intro.tex
removed commas per abhishek

@@
 This presents a strong case for using the length over the velocity gauge 
-for more generalized, and less expensive calculations of the nonlinear
+for more generalized and less expensive calculations of the nonlinear
 response of a system.
@@
 way to calculate the nonlinear response using the scissored Hamiltonian 
-in the length gauge, providing better agreement with experiment.
+in the length gauge providing better agreement with experiment.
 
@@
 nonlinear susceptibility tensor $\chi^{abc}(-2\omega;\omega,\omega)$ 
-in the length gauge formalism, 
+in the length gauge formalism
 that include nonlocal terms from the scissors approximation 
-and the spatial contribution from the pseudopotentials, and a 
+and the spatial contribution from the pseudopotentials and a 
 layer-by-layer approach convenient for surface SHG calculations. We are
@@
 $2\omega$ contributions. This approach is unprecedented and yet produces 
-results that are in total agreement with the literature.
+results that are in complete agreement with the literature.
 
@@
 programs based on density-functional theory (DFT) within the local-density 
-approximation (LDA), using Troullier-Martins pseudopotentials. We compare 
+approximation (LDA) using Troullier-Martins pseudopotentials. We compare 
 results from before and after adding the different nonlocal contributions.

### 2014-11-17 paper-v-0.tex
graficas
added intro.tex

@@
 %\input{/Users/bms/util/definitions}
+\def\chon{black}
 \input{definitions}
@@
 We present a theoretical review of surface second harmonic generation (SHG) 
-from semiconductor surfaces based on the {\color{red} length} gauge. This 
+from semiconductor surfaces based on the {\color{\chon} length} gauge. This 
 layer-by-layer analysis is carefully presented in order to show how a 
@@
 
+\input{intro}
+
 \section{Theory}
@@
 We use the independent particle approximation and neglect local field
-and excitonic effects{\color{red}. We assume a classical} electromagnetic field
-{\color{red} and quantum mechanical matter.}
+and excitonic effects{\color{\chon}. We assume a classical} electromagnetic field
+{\color{\chon} and quantum mechanical matter.}
 We can describe the system using 
@@
 $\tilde{\mathcal{O}}=U{\cal O}U^{\dagger }$. Even if $\mathcal{O}$ 
-{\color{red} is time-independent}, 
-$\tilde{\mathcal{O}}$ {\color{red} is time-dependent} through the explicit
+{\color{\chon} is time-independent}, 
+$\tilde{\mathcal{O}}$ {\color{\chon} is time-dependent} through the explicit
 time dependence of $U$. 
@@
 $\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle=f(\hbar \omega_{n}(\mathbf{k}))\equiv f_{n}$. For a
-clean, cold semiconductor $f_{n}=1$ {\color{red} when} $n$ {\color{red} is} a valence ($v$) or
-occupied band, and zero {\color{red} when} $n$ {\color{red} is} a conduction ($c$) or empty band.
-{\color{red} We assume this for the remainder of the article.}
-{\color{red} T}he expectation values must satisfy
+clean, cold semiconductor $f_{n}=1$ {\color{\chon} when} $n$ {\color{\chon} is} a valence ($v$) or
+occupied band, and zero {\color{\chon} when} $n$ {\color{\chon} is} a conduction ($c$) or empty band.
+{\color{\chon} We assume this for the remainder of the article.}
+{\color{\chon} T}he expectation values must satisfy
 $
@@
 where $\dot{\mathbf{r}}$ is the time derivative of the position operator of the
-electron {\color{red} with} charge $e$, {\color{red} defined as}
+electron {\color{\chon} with} charge $e$, {\color{\chon} defined as}
 \begin{equation}
@@
 where $\tilde\omega=\omega+i\eta $,
-and $\eta >0$ {\color{red} adiabatically switches on} the interaction,
+and $\eta >0$ {\color{\chon} adiabatically switches on} the interaction,
 we write the second order nonlinear
@@
 where $\chi^{abc}(-2\omega ;\omega ,\omega )$ is the nonlinear
-susceptibility responsible {\color{red} for} surface second harmonic generation
+susceptibility responsible {\color{\chon} for} surface second harmonic generation
 (SSHG). 
@@
 is used to solve the Kohn-Sham equations\cite{kohnPR65} of Density  
-Functional Theory (DFT) {\color{red}. It is convenient to work within the Local 
+Functional Theory (DFT) {\color{\chon}. It is convenient to work within the Local 
 Density Approximation (LDA), so we label the hamiltonian with the corresponding  
@@
 $\chi^{abc}(-2\omega;\omega,\omega)$, we asume
-that the crystal potential could have a {\color{red} non-local spatial}
-contribution {\color{red} as} is customary for most
+that the crystal potential could have a {\color{\chon} non-local spatial}
+contribution {\color{\chon} as} is customary for most
 pseudopotentials, and then we replace $V$ with
@@
 from the non-local part of the pseudopotential, $V^{\mathrm{nl}}$, in
-the results that we {\color{red} have derived.}
-
-{\color{red} It} is well known that the use of the {\color{red} LDA} 
+the results that we {\color{\chon} have derived.}
+
+{\color{\chon} It} is well known that the use of the {\color{\chon} LDA} 
 leads to an underestimation of the band gap. A standard
 procedure to correct for this is to
-use {\color{red} the} ``scissors approximation'', {\color{red} where the 
+use {\color{\chon} the} ``scissors approximation'', {\color{\chon} where the 
 conduction bands are rigidly shifted in energy} so that the band gap 
-corresponds to the accepted experimental band gap{\color{red} .
+corresponds to the accepted experimental band gap{\color{\chon} .
 This} is often in fairly good agreement with the GW
 band gap based on a more sophisticated calculation.\cite{hybertsenPRB86}
-{\color{red} T}he LDA wave functions are used since they produce band
+{\color{\chon} T}he LDA wave functions are used since they produce band
 structures with dispersion relations similar to those predicted by the GW
-approximation. Mathematically, {\color{red} the scissors (non-local) term 
-$S$ is added} to the unperturbed or unscissored Hamiltonian {\color{red} $H^{\mathrm{LDA}}_{0}$ ,}
+approximation. Mathematically, {\color{\chon} the scissors (non-local) operator 
+$S$ is added} to the unperturbed or unscissored Hamiltonian {\color{\chon} $H^{\mathrm{LDA}}_{0}$ ,}
 \begin{equation*}
@@
 \end{equation}
-We emphasize that the {\color{red} scissored and unscissored Hamiltonian 
+We emphasize that the {\color{\chon} scissored and unscissored Hamiltonian 
 have the same eigenfunctions,} where
@@
 To calculate the optical response 
-within the Length Gauge {\color{red} formalism}, we take the 
-interaction Hamiltonian {\color{red} as} 
+within the Length Gauge {\color{\chon} formalism}, we take the 
+interaction Hamiltonian {\color{\chon} as} 
 \begin{equation}
@@
 treatment of the position operator
-$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$ {\color{red} has} 
-been described {\color{red} in detail}.\cite{adamsJCP53,blountSSP62} 
-{\color{red} In particular, Ref.~\onlinecite{aversaPRB95} develops} 
+$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$ {\color{\chon} has} 
+been described {\color{\chon} in detail}.\cite{adamsJCP53,blountSSP62} 
+{\color{\chon} In particular, Ref.~\onlinecite{aversaPRB95} develops} 
 the correct treatment of $\bfr$ for
-the calculation of the {\color{red} nonlinear} optical {\color{red} response.} 
+the calculation of the {\color{\chon} nonlinear} optical {\color{\chon} response.} 
 $\bfr$ is split into the {\it intraband} 
@@
 \end{align}
-such that $\bfr_{e,nm}=0$ for $n=m${\color{red} .
+such that $\bfr_{e,nm}=0$ for $n=m${\color{\chon} .
 From Eq.~\eqref{mv}} with $H_0\to
@@
 %elements.
-For the intraband part, {\color{red} $\bfr_i$ only appears in
+For the intraband part, {\color{\chon} $\bfr_i$ only appears in
 commutators during the derivation of
@@
 The vectors $\bfgxi_{nn}(\bfk)$ are defined in 
-Ref.~\onlinecite{aversaPRB95} {\color{red} though} they do not need to be 
+Ref.~\onlinecite{aversaPRB95} {\color{\chon} though} they do not need to be 
 calculated explicitly in what follows.
 
-Before {\color{red} continuing,} 
+Before {\color{\chon} continuing,} 
 we derive a key result for the length gauge formulation. 
-{\color{red} Again, using} $H^\sigma_0$ in
+{\color{\chon} Again, using} $H^\sigma_0$ in
 Eq.~\eqref{mv} we obtain
@@
 \end{subequations}
-with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator{\color{red}, 
+with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator{\color{\chon}, 
 and using $[r^a,p^b]=i\hbar\gd_{ab}$, where $\gd_{ab}$ is the Kronecker} delta.
@@
 with $f_{nm}\equiv f_n-f_m$,
-where we see that $\bfv^S_{nn}=0$. {\color{red} It} follows that
+where we see that $\bfv^S_{nn}=0$. {\color{\chon} It} follows that
 \begin{align}\label{chon.9}
@@
 \end{align}
-{\color{red} and from Eq.~\eqref{pmnrmn},} 
+{\color{\chon} and from Eq.~\eqref{pmnrmn},} 
 \begin{align}\label{chon.10}
@@
 The matrix elements 
-of $\bfr_{nm}(\bfk)$ are {\color{red} identical using either} 
-the LDA or {\color{red} scissored} 
-Hamiltonian, {\color{red} thus negating the} need to label them.
-Of course, {\color{red} it} is more convenient to calculate them
-through $\bfv^\lda_{nm}(\bfk)$ {\color{red} which 
+of $\bfr_{nm}(\bfk)$ are {\color{\chon} identical using either} 
+the LDA or {\color{\chon} scissored} 
+Hamiltonian, {\color{\chon} thus negating the} need to label them.
+Of course, {\color{\chon} it} is more convenient to calculate them
+through $\bfv^\lda_{nm}(\bfk)$ {\color{\chon} which 
 includes} the contribution of 
 the non-local part of the pseudopotentials through
-$\bfv^\nl_{nm}(\bfk)$. These {\color{red} can} be readily
+$\bfv^\nl_{nm}(\bfk)$. These {\color{\chon} can} be readily
 calculated
@@
 form,\cite{francesco,mottaCMS10,kleinmanPRL82,adolphPRB96}
-{\color{red} as shown in Appendix \ref{vesnl}.}
-{\color{red} The} advantage of the length gauge formalism for 
-{\color{red} calculating linear and nonlinear} optical responses 
-resides in the ease with {\color{red} which the scissors operator
+{\color{\chon} as shown in Appendix \ref{vesnl}.}
+{\color{\chon} The} advantage of the length gauge formalism for 
+{\color{\chon} calculating linear and nonlinear} optical responses 
+resides in the ease with {\color{\chon} which the scissors operator
 can be introduced into the calculation, also described in detail in} 
-Nastos et al.\cite{nastosPRB05} {\color{red} The 
+Nastos et al.\cite{nastosPRB05} {\color{\chon} The 
 length gauge} formalism for the scissored Hamiltonian can be easily worked 
@@
 for the unperturbed system 
-with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction{\color{red}. 
+with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction{\color{\chon}. 
 We need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
@@
 
-{\color{red} Taking} the matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of
+{\color{\chon} Taking} the matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of
 Eq.~\eqref{rde}, we obtain 
@@
 %$(B^a_{nm})_{;k^b}$ can be obtained.
-{\color{red} Note that the indices in Eq.~\eqref{j.2} are all} different from each
-other. {\color{red} This is due to the $f_{nm}$ factor in Eq.~\eqref{j.1}, 
-and therefore $B^a_{nn}=0$}. The dependence {\color{red} on} $\bfk$ 
+{\color{\chon} Note that the indices in Eq.~\eqref{j.2} are all} different from each
+other. {\color{\chon} This is due to the $f_{nm}$ factor in Eq.~\eqref{j.1}, 
+and therefore $B^a_{nn}=0$}. The dependence {\color{\chon} on} $\bfk$ 
 of all quantities is implicitly understood from 
-{\color{red} this point forward.}
+{\color{\chon} this point forward.}
 
@@
 semi-infinite system, we replace it by a super cell that 
-{\color{red} consists} of a finite
+{\color{\chon} consists} of a finite
 slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
-super cell {\color{red} constitutes the unit cell, and}
-is repeated to form a full three dimensional crystalline structure.
-The slab itself consists of {\color{red} front, back, and 
+super cell {\color{\chon} constitutes the unit cell, and}
+is repeated to form a full three dimensional crystalline structure,
+i.e. a repeated super cell scheme. 
+The slab itself consists of {\color{\chon} front, back, and 
 sub-surface} regions, and in between these
-a region that is equivalent to {\color{red} the}
+a region that is equivalent to {\color{\chon} the}
 bulk of the system. 
@@
 balanced forces are disrupted when the surface atoms do not find their 
-partner atoms that are now absent at the {\color{red} slab surface.} 
+partner atoms that are now absent at the {\color{\chon} slab surface.} 
 To take the reconstruction or relaxation into account, 
 we take ``surface'' to mean
-the true surface of the first layer of {\color{red} atoms} and
+the true surface of the first layer of {\color{\chon} atoms} and
 some of the atomic sub-layers adjacent to it.
@@
 centrosymmetric. This would imply that 
-$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=0$ {\color{red} so} we must
+$\chi^{\mathrm{slab},\rma\rmb\rmc}=0$ {\color{\chon} so} we must
 find a scheme 
-in order to have a finite $\chi_{\rma\rmb\rmc}$ representative of the
+in order to have a finite $\chi^{\rma\rmb\rmc}$ representative of the
 surface. Even if the front and back surfaces of the slab 
 are different, breaking the centrosymmetry and therefore giving an
-overall $\chi^{\mathrm{slab}}_{\rma\rmb\rmc}\ne 0${\color{red} ;} we still
-need a procedure to extract the front surface $\chi^f_{\rma\rmb\rmc}$
-and the back surface $\chi^b_{\rma\rmb\rmc}$ from the slab
+overall $\chi^{\mathrm{slab},\rma\rmb\rmc}\ne 0${\color{\chon} ;} we still
+need a procedure to extract the front surface $\chi^{f,\rma\rmb\rmc}$
+and the back surface $\chi^{b,\rma\rmb\rmc}$ from the slab
 susceptibility 
-$\chi^{\mathrm{slab}}_{\rma\rmb\rmc}=\chi^f_{\rma\rmb\rmc}-\chi^b_{\rma\rmb\rmc}$.
-We have omitted the frequency arguments of $\chi_{abc}$ for 
-{\color{red} convenience of} notation.
+$\chi^{\mathrm{slab},\rma\rmb\rmc}=\chi^{f,\rma\rmb\rmc}-\chi^{b,\rma\rmb\rmc}$.
+We have omitted the frequency arguments of $\chi^{\rma\rmb\rmc}$ for 
+{\color{\chon} convenience of} notation ({\color{blue}shall we??}).
 
@@
 $(f)$ or back $(b)$ from its $z_\ell$ position. 
-We take $z_\ell$ {\color{red} to be}at the center of an atom that 
-belongs to layer $\ell${\color{red} , so the previous} 
+We take $z_\ell$ {\color{\chon} to be} at the center of an atom that 
+belongs to layer $\ell${\color{\chon} , so the previous} 
 equation would give the $\ell$-th atomic-layer 
-contribution to the {\color{red} nonlinear} optical response.
+contribution to the {\color{\chon} nonlinear} optical response.
 $\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
 (see Fig.~\ref{fslab}).
-\begin{figure}[b]
+\begin{figure}
 \centering
@@
 \includegraphics[scale=.7]{images/slab}
-\caption{A sketch of the super {\color{red} cell. 
+\caption{(color on line) A sketch of the super {\color{\chon} cell. 
 The} slab corresponds to the
@@
 $\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gs\ket{\bfr}\bra{\bfr}
-+\ket{\bfr}\bra{\bfr}\bfv^\gs\right)${\color{red} , that} leads to
++\ket{\bfr}\bra{\bfr}\bfv^\gs\right)${\color{\chon} , that} leads to
 \begin{align}\label{jmic}
@@
 If we want the contribution from only one region of the unit cell 
-{\color{red} towards??} the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
-over the desired region. {\color{red} Then the} contribution 
+{\color{\chon} towards} the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
+over the desired region. {\color{\chon} Then the} contribution 
 to the current density from the
@@
 is all that is needed to change any of the
-{\color{red} electron} velocity operators $\bfV$ to the new velocity
+{\color{\chon} electron} velocity operators $\bfV$ to the new velocity
 operator $\calbv^{\ell}$ that implicitly takes into account the
@@
 (\go^\gs_{nm}/\go^\lda_{nm})
-\calbv^{\ell,\lda}_{nm}(\bfk)${\color{red} .
+\calbv^{\ell,\lda}_{nm}(\bfk)${\color{\chon} .
 To calculate}
@@
 
-To limit the SHG response to one surface, the equivalent of Eq.~\eqref{nl.4} 
-for $\calbp^\ell$ was proposed in 
+To limit the SHG response to one surface, Eq.~\eqref{vcali} 
+for $\calbv^\ell$ was proposed in 
 Ref.~\onlinecite{reiningPRB94} and later used in Refs.
@@
 cell that characterizes the surface of the system, and $d$
-is the surface region {\color{red} from} which the {\color{red} nonlinear} 
+is the surface region {\color{\chon} from} which the {\color{\chon} nonlinear} 
 susceptibility is different from zero.
@@
 $\chi^{\ell,\rma\rmb\rmc}(-2\go;\go,\go)$
-are m$^2$/V, as they {\color{red} should be} for a surface SH susceptibility.
+are m$^2$/V, as they {\color{\chon} should be} for a surface SH susceptibility.
 Using Eq.~\eqref{j.1} we
@@
 two contributions from the first and second terms on the right hand side
-{\color{red} of Eq.~\eqref{Pjikn2}:}
+{\color{\chon} of Eq.~\eqref{Pjikn2}:}
 %\end{document}  
@@
 \end{equation} 
-{\color{red} related} to intraband transitions, and 
+{\color{\chon} related} to intraband transitions, and 
 \begin{equation}\label{chie}
@@
 \end{equation} 
-{\color{red} related} to interband transitions.
+{\color{\chon} related} to interband transitions.
 The generalized derivative in Eq.~\eqref{chii} is dealt with by the chain rule 
@@
 \end{equation}
-where {\color{red} substituting} $H^\sigma_0$ 
+where {\color{\chon} substituting} $H^\sigma_0$ 
 into Eq.~\eqref{conmri3n} and then
 Eq.~\eqref{chon.9}
-{\color{red} we obtain}
+{\color{\chon} we obtain}
 \begin{align}\label{eli.13}
@@
 Eqs. \eqref{chii} and \eqref{chie}, 
-\begin{subequations}\label{xis}
+\begin{subequations}\label{chis}
 \begin{equation}\label{calvimchiewn}
@@
 \end{subequations}
-where we have {\color{red} split} the interband and intraband $1\go$ and $2\go$
-contributions and {\color{red} supressed} the $\go$ arguments for 
-{\color{red} convenience of notation}.
+where we have {\color{\chon} split} the interband and intraband $1\go$ and $2\go$
+contributions and {\color{\chon} supressed} the $\go$ arguments for 
+{\color{\chon} convenience of notation}.
 The real part of each contribution can be obtained through
-a Kramers-Kronig transformation\cite{nicolas} {\color{red} and}
+a Kramers-Kronig transformation\cite{nicolas} {\color{\chon} and}
 $\chi^{\ell,\rma\rmb\rmc}=
@@
 $\chi^{\ell,\rma\rmb\rmc}=\chi^{\ell,\rma\rmc\rmb}$.
-The various quantities involved in Eqs.~\eqref{xis} are given in
+The various quantities involved in Eqs.~\eqref{chis} are given in
 the Appendix \ref{appe}. 
+We mention that if we take $C^\ell(z)=1$ through out, the layered
+matrix elements $\calbv^{\ell,\sigma}_{nm}$ become standard bulk-like
+$\bfv^{\ell,\sigma}_{nm}$ matrix elements. We mention that in this
+case, Eq.~\eqref{chis} is equivalent to the expressions of
+Ref.~\onlinecite{cabellosPRB09}, valid for bulk semiconductors.
 
@@
 where $\{\ell\}$
-is meant to be {\color{red} a chosen set of layers.} For instance, 
-one {\color{red} can take a single layer 
-encompassing} half of the slab, or {\color{red} take each 
+is meant to be {\color{\chon} a chosen set of layers.} For instance, 
+one {\color{\chon} can take a single layer 
+encompassing} half of the slab, or {\color{\chon} take each 
 atomic layer individually to the middle} 
-of the slab. {\color{red} For} the first case there is 
-{\color{red} a} single summand
-in Eq.~\eqref{chiijksur}{\color{red} . For the second case
-there is} a sum from $\ell=1${\color{red} , denoting the first layer 
+of the slab. {\color{\chon} For} the first case there is 
+{\color{\chon} a} single summand
+in Eq.~\eqref{chiijksur}{\color{\chon} . For the second case
+there is} a sum from $\ell=1${\color{\chon} , denoting the first layer 
 right at the surface, to $\ell=N$, denoting the layer at the middle of the slab 
@@
 %$\bfgchi^{\ell=N}=0$;
-We {\color{red} remark} that the value of 
-$N$ is not universal {\color{red} and} 
+We {\color{\chon} remark} that the value of 
+$N$ is not universal {\color{\chon} and} 
 the slab needs to have enough atomic layers 
@@
 %to be satisfied and to 
-{\color{red} in} order to give converged results for 
+{\color{\chon} in} order to give converged results for 
 $\bfgchi (-2\go;\go,\go)$. 
@@
 either the front or the back surface. 
-%\mathrextcolor{red}{Talk about the correct value of $d$}
-
-\section{Results}
+%\mathrextcolor{\chon}{Talk about the correct value of $d$}
+
+\section{Test Case}
+
+In this section we present a test case in order to check the
+consistency of the approach presented above. To do this end,  we have chosen
+a clean Si(001) surface  with a $2\times 1$ surface reconstruction.
+The slab for such a surface could be chosen to be centrosymmetric 
+by having the fron and the back surface with the same $2\times 1$
+reconstruction, however we terminate with hydrogen one of the
+surfaces, thus having it as an ideally bulk terminated Si
+surface. Indeed, the H atoms simply saturate the dangling bonds of the
+bulk-like Si atoms at the surface, as seen in Fig.~\ref{si2x1}. \textcolor{red}{details of the
+ relaxation and some references}
+\begin{figure}
+\centering 
+\includegraphics[scale=.8]{images/si2x1-crop}
+\caption{(color on line) The slab shows a front clean Si(001)$2\times 1$ surface,
+  while the back surface is ideally terminated Si bulk, where the  
+  dangling bonds are H (small balls) saturated. This picture shows 12  
+  Si atomic layers and one H atomic layer. 
+\label{si2x1}} 
+\end{figure} 
+The idea for such an slab is that the cristalline symmetry of the
+H terminated surface imposes that $\chi_{\mathrm{H}}^{xxx}=0$, while
+for the $2\times 1$ surface there is not such symmetry restriction and
+then, $\chi_{2\times 1}^{xxx}\ne 0$.
+Therefore, due to this fact, calculating $\chi^{xxx}$ for the full-slab
+or the half-salb that contains the $2\times 1$ surface,\cite{nota1}
+ ought to give the same result, since the contribution from the H
+ saturated  surface is zero any way. Then, for this slab one must check
+ whether or not this relation is satisfied,
+\begin{align}\label{hs}
+\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go) 
+=
+\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)
+,
+\end{align}
+where
+$\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
+$C^\ell(z)=1$ un the upper half of the salb, that contains the
+$2\times 1$ surface reconstruction, as seen in
+Fig.~\ref{si2x1},
+whereas $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
+$C^\ell(z)=1$ through the full slab.
+Indeed, in what follows we show the results of such a comparison.
+
+The self-consistent ground state and their Kohn-Sham states were
+calculated in the frame-work of DFT-LDA,
+with the use of the plane-wave ABINIT
+code.\cite{abinit}
+We used
+Troullier-Martins
+pseudopotentials,\cite{troullierPRB91} 
+that are 
+fully separable nonlocal pseudopotentials in the 
+Kleinman-Bylander 
+form;\cite{kleinmanPRL82}
+the contribution of $\bfv^\nl$ to Eq.~\eqref{chis},
+can be calculated, as explained in Appendix \ref{vesnl}; 
+we use the DP code for this end.\cite{francesco}
+The $2\times 1$ and the H surfaces are relaxed \textcolor{red}{Nicolas 
+  some details}, where we find agreement with previous results for 
+both surfaces.\cite{relax} For instance the Si-H bond distance is 1.48 
+$\AA$ ??. The unit cell has two atoms, making the calculation
+computational very heavy. The vacuum size is equivalent to one quarter
+of the
+size of the slab, thus avoiding spurious effects from the possible
+tunneling of the wave function of the contiguous surfaces of the full
+crystal formed by the repeated super cell scheme.\cite{mendozaPRB06}
+
+Spin-orbit effects, local field effects, and the consequences of the
+electron-hole
+attraction\cite{leitsmannPRB05,trollePRB14} on the SHG process are neglected. 
+Although all these effects are important for the optical response of a
+semiconductor, 
+their calculation is still an open question and a numerical challenge
+that ought to be pursued. 
+However this endeavor is beyond the scope of this paper. 
+For a given size  
+of the slab,  we find converged spectra
+for all the quantities of interest in this
+work, 
+and the most important parameters are
+an energy cut-off of 10 Ha,
+a number 
+of conduction bands equal to the number of valence 
+bands,
+and a set of 244 $\bfk$-points
+is used for
+the linear analytic tetrahedron method used in the evaluation
+the Brillouin zone integrals for of Eq.~\eqref{chis}, where special care
+was taken to examine the double resonances.\cite{nastosPRB05}
+Double resonances occur if for a given frequency $\omega$
+there can be resonant
+transitions at both frequencies $\omega$ and 2$\omega$,
+ that is, if there is a
+region in the Brillouin zone such that
+$\omega_{cv}(\bfk)=2 \omega_{c'v}(\bfk)$
+For these $\bfk$ points the perturbation theory used to calculate
+the spectrum breaks down, since there is real population excited,
+which in a correct calculation must be taken into account. These
+points introduce sharp spikes in the spectrum that can, in
+principle, affect the low-frequency results, since the response at
+frequencies below the band gap is computed from the Kramers-Kronig
+relation. However, in agreement with Ref.~\onlinecite{nastosPRB05}
+ we find here
+that the double resonances affect the low-frequency results by less
+than 2\%. 
+Finally, all the spectra are calculated with an energy Gaussian smearing of  0.15 eV {\color{red}Chon??} 
+\begin{figure}
+\centering 
+\includegraphics[scale=.8]{plots/fig1}
+\caption{(color on line) 
+$|\chi_{\mathrm{half-slab}}^{xxx}|$ vs $\hbar\omega$ for different slabs with 16, 24, 32 
+and 40 atomic Si layers. The fron surface of the slab presents a 
+$2\times 1$ reconstruction, whereas the back sufrace is ideally bulk 
+terminated where the dangling bonds are H-saturated. 
+\label{fig1}} 
+\end{figure}
+\begin{figure}
+\centering 
+\includegraphics[scale=.8]{plots/fig2}
+\caption{(color on line) 
+$\chi^{xxx}_{\mathrm{half-slab}}$
+and 
+$\chi^{xxx}_{\mathrm{full-slab}}$
+vs $\hbar\omega$ for a slab of 32 
+atomic Si layers plus one layer of H. 
+\label{fig2}} 
+\end{figure}
+
+To obtain Eqs.~\eqref{tau.1}  and \eqref{tau.1n}, required in
+Eq.~\eqref{chis},
+we see that the
+evaluation of the commutator $[\bfr,\bfv^\nl]$ is required as it is
+outlined in Appendix \ref{appvnl}; second order derivatives
+are required to do such evaluation, making the numerical
+procedure very
+time consuming. Indeed this time is much higher than the already
+lengthy time needed for the calculation of the contribution of
+$\bfv^\nl$, proportional to only first order derivatives, as seen in
+Appendix \ref{vesnl}. 
+On top of this, the memory required is also very demanding for both
+$\bfv^\nl$ and $[\bfr,\bfv^\nl]$.
+However, the contribution coming from 
+$[\bfr,\bfv^\nl]$, turns out to be very small,\cite{valerie} and
+therefore we neglect it in this work.
+
+In Fig.~\ref{fig1} we show $|\chi_{\mathrm{half-slab}}^{xxx}|$
+ for a
+slab of 16, 24, 32, and 40 Si atomic layers. We see that 
+$|\chi_{\mathrm{half-slab}}^{xxx}|$ for
+16 and 24
+layers are different from each other, and different from that of 32
+and 40 layers,
+that are very similar to each other.
+As explained above,
+the calculation of the contribution given by
+$\bfv^\nl$ is computational very demanding. A good
+compromise between the accuracy in the convergence of
+$\chi^{xxx}_{\mathrm{half-slab}}$,
+ as a function of the number
+of atomic layers in the slab, and the computational burden, is to take
+the slab with 32 Si atomic layers, as representative of our system
+in question.
+In Fig.~\ref{fig2}
+we compare 
+$\chi^{xxx}_{\mathrm{half-slab}}$  
+vs. 
+$\chi^{xxx}_{\mathrm{full-slab}}$, 
+for the four different possibilities of including or not including the
+effects of $\bfv^\nl$ and having or not having a scissors correction
+$\hbar\sigma$.
+We see that for all four instances the comparison between
+$\chi^{xxx}_{\mathrm{half-slab}}$ 
+and
+$\chi^{xxx}_{\mathrm{full-slab}}$,
+is very good. Indeed, where the value of $|\chi^{xxx}|$ is large the
+difference is very small, and when its value is small, the difference
+are just slightly larger, but the spectra is so close to zero that the
+differences are not very meaning full; this diferences would become
+smaller as the number of atomic layers is increased. However, with 32
+layers is more than enough to 
+confirm that the extraction of the surface second
+harmonic susceptibility from the $2\times 1$ surface is 
+readily possible using the formalism contained in Eq.~\eqref{chis}.
+
+Although in a forthcoming publication we will present a study of SSHG
+from several Si surfaces where experimental results are available, we
+proceed to explain, without confronting with experiment ({\color{red}
+Nicolas are there any for this particular component?}), some of the
+features seen in 
+$|\chi^{xxx}_{2\times 1}|$,  
+that is of coures given by
+$|\chi^{xxx}_{\mathrm{half-slab}}|$.
+First of all, from Fig.~\ref{fig2} 
+we see a series of resonances 
+that come from the 1$\omega$ and 2$\omega$ terms in
+Eq.~\eqref{chis}. The 2$\omega$ resonances naively would start below
+$E_g/2$ where $E_g=3.4$ eV is
+the band gap of bulk Si. However we see that there are several
+resonances well below 1.7 eV, and of course, these resonances are
+coming from the electronic states of the $2\times 1$ surface, that lie
+inside the gap, that are the well known surface states.
+We see that the effect of $\bfv^\nl$ reduces the value of   
+$|\chi^{xxx}_{2\times 1}|$,\cite{note2}  
+without changing the form of the spectra, as can be confirmed from
+both cases of zero and non-zero scissors correction.
+Also, we see that the scissors correction shifts the spectra to higher
+energies, as expected, however, contrary to the case of linear optics,\cite{cabellosPRB09}
+the shifting is not rigid. This is because, the second harmonic optical
+response mixes 1$\omega$ and 2$\omega$ transitions, as can be seen from
+Eq.~\eqref{chis}, and thus the spectra is not rigidly shifted. 
+\begin{figure}
+\centering 
+\includegraphics[scale=.8]{plots/fig3}
+\caption{(color on line) 
+$\chi^{xxx}_{2\times 1}$
+vs $\hbar\omega$ for a slab of 32 
+atomic Si layers plus one layer of H, for two different values of 
+the scissors correction $\hbar\sigma$.
+\label{fig3}} 
+\end{figure}
+
+Finally, in Fig.~\ref{fig3} we compare 
+$|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$. Indeed,
+the value of $\hbar\sigma=0.5$ eV, used above, is the ``average'' GW gap taken from 
+Refs.~\onlinecite{rholfingPRB95} which is in agreement with \onlinecite{garciaCPC01}, and  
+the value of $\hbar\sigma=0.63$ eV is the ``average'' GW gap taken from 
+Refs.~\onlinecite{asahiPRB00}, where more $\bfk$ points in the
+Brillouin zone were used.
+We see that the fist two peaks are almost rigidly shifted with a small
+difference in height, while the rest of the peaks are modified
+substantially. This behavior, comes from the fact that the first two
+peaks are almost related to the 1$\omega$ resonances of
+Eq.~\eqref{chis}, and the other peaks are a combination of 1$\omega$
+and 2$\omega$ resonances, thus giving a more different spectrum. This
+way we see that small changes in the value of the scissors shift can affect the
+$|\chi^{xxx}_{2\times 1}|$ quite dramatically.
 
@@
 We give explicit expressions for the quantities used in the evaluation 
-of Eq.~\eqref{xis}; when appropriate, some 
+of Eq.~\eqref{chis}; when appropriate, some 
 intermediate steps are given for their derivation. 
@@
 $e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
-space, and $A_{n\bfk}(\bfG)$ {\color{red} the} expansion coefficients. Using
-$m_e\bfv=-i\hbar\bfgnabla$ into the {\color{red} equivalent??} 
-of Eq.~\eqref{intj}
+space, and $A_{n\bfk}(\bfG)$ {\color{\chon} the} expansion coefficients. Using
+$m_e\bfv=-i\hbar\bfgnabla$ into 
+Eq.~\eqref{vcali}
 we obtain,\cite{mendozaPRB06}
@@
 with $f^{\ell*}(g)=f^\ell(-g)$. 
-{\color{red} The} reciprocal lattice vectors $\bfG$ are 
+{\color{\chon} The} reciprocal lattice vectors $\bfG$ are 
 decomposed into components
-{\color{red} parallel ($\bfG_\parallel$), and perpendicular ($G_\perp \hat z$)} 
+{\color{\chon} parallel ($\bfG_\parallel$), and perpendicular ($G_\perp \hat z$)} 
 to the surface, so
@@
 The double summation over the $\bfG$ vectors can be 
-{\color{red} calculated efficiently} by  
+{\color{\chon} calculated efficiently} by  
 creating a pointer array to identify all the plane-wave coefficients  
@@
 \end{align}  
-If $\calc^\ell(z)=1$, {\color{red} then $f^\ell(g)=\gd_{g0}$ and we 
+If $\calc^\ell(z)=1$, {\color{\chon} then $f^\ell(g)=\gd_{g0}$ and we 
 obtain the full-slab/bulk values, 
@@
 
-{\color{red} For} any function $F(\bfr)$, $[\bfr,F(\bfr)]=0$, 
-{\color{red} we can use}
-Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn} 
-{\color{red} to }obtain
+We use Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn},
+along with $[\bfr,F(\bfr)]=0$, valid for 
+and any function $F(\bfr)$, 
+{\color{\chon} to }obtain
 \begin{align} 
@@
 where we remind the reader that $\bfr_{nm}$ 
-{\color{red} is} calculated through
+{\color{\chon} is} calculated through
 Eq.~\eqref{chon.10} for LDA.
@@
 $
-one can show that{\color{red} 
+one can show that{\color{\chon} 
 \begin{align}\label{nmesn}
@@
 %compared to the other terms, thus we neglect it throwout this work.\cite{valerie} 
+Notice that
+{\color{\chon} 
+$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$} is obtained 
+from Eq.~\eqref{nmesn} by 
+taking 
+{\color{\chon} $C^\ell(z)=1$ or $C^\ell_{nm}=\gd_{nm}$.}
 
@@
 =(v^{\lda,\rma}_{nm}/i\go^\lda_{nm})_{;k^{\rmb}}
-$ {\color{red} and} apply the chain rule. 
-{\color{red} 
-$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$} is obtained
-from Eq.~\eqref{nmesn} by
-taking
-{\color{red} $C^\ell(z)=1$ or $C^\ell_{nm}=\gd_{nm}$,}
-\begin{align}\label{na_rgendevn}
+$ {\color{\chon} and} simply apply the chain rule,
+\begin{align}\label{rka}
 (r^{\rma}_{nm})_{;k^{\rmb}}
@@
 \end{eqnarray}
-and{\color{red} 
+and{\color{\chon} 
 \begin{align}\label{tau.1n} 
@@
 -i [r^{\rmb},v^{\nl,\rma}]_{nm} 
-,
+.
 \end{align}}
-generalizes the usual expresion of
-$(r^a_{nm})_{;k^b}$ for {\color{red} a} local 
-{\color{red} Hamiltonian}
+Eq.~\eqref{rka} generalizes the usual expresion of
+$(r^a_{nm})_{;k^b}$ for {\color{\chon} a} local 
+{\color{\chon} Hamiltonian}
 \cite{aversaPRB95,nastosPRB05,cabellosPRB09,rashkeevPRB98}
@@
 $\calt^{\ell,\rma\rmb}_{nm}$ reduces to $t^{\rma\rmb}_{nm}$
-{\color{red} for the full-slab/bulk case.}
+{\color{\chon} for the full-slab/bulk case.}
 
@@
 \mathrm{exp}(i(\bfk+\bfG)\cdot\bfr)$
-{\color{red} to} obtain,\cite{mottaCMS10}
+{\color{\chon} to} obtain,\cite{mottaCMS10}
 \begin{align}\label{vnl.5}
@@
 $ 
-{\color{red} and} their $\bfK$ and $\bfK'$ gradient 
+{\color{\chon} and} their $\bfK$ and $\bfK'$ gradient 
 can be readily calculated.\cite{mottaCMS10,adolphPRB96,gordienkoRPJ04,fuchsCPC99} 
@@
 implemented 
-the calculation of $\calbv^{\ell,\nl}_{nm}(\bfk)$ with the help {\color{red} of} 
+the calculation of $\calbv^{\ell,\nl}_{nm}(\bfk)$ with the help {\color{\chon} of} 
 the \depe~code.\cite{francesco}
-{\color{red} We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
+{\color{\chon} We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
 expressions by t}aking $C^\ell(z)=1$, which implies 
-{\color{red} that} $f^\ell(g)=\gd_{g0}$.
+{\color{\chon} that} $f^\ell(g)=\gd_{g0}$.
 
@@
 \end{align} 
-{\color{red} and we expand the triple commutator,}
+{\color{\chon} and we expand the triple commutator,}
 \begin{align}\label{3.5}
@@
 V^\nl 
-\ket{\bfK'} 
+\ket{\bfK'}, 
 \nonumber\\
-,
 \end{align} 
@@
 The double derivatives with respect to $\bfK$ and $\bfK'$ 
-can be worked out explicitly {\color{red} and} 
+can be worked out explicitly {\color{\chon} and} 
 ${\cal L}^{\rma\rmb}_{nm}(\bfk)$
-{\color{red} can} be finally calculated.\cite{valerie}
+{\color{\chon} can} be finally calculated.\cite{valerie}
 
@@
 where $\sum_q\ket{q\bfk}\bra{q\bfk}=1$ was used
-{\color{red} and $\bfv^S_{nm}$ is} given in Eq.~\eqref{chon.2}.
+{\color{\chon} and $\bfv^S_{nm}$ is} given in Eq.~\eqref{chon.2}.
 Taking the generalized derivative of Eq.~\eqref{a.3b}
-{\color{red} and} applying
+{\color{\chon} and} applying
 the chain rule, we obtain
@@
 \end{align}    
-{\color{red} and considering} 
+{\color{\chon} and considering} 
 Eq.~\eqref{chon.2}, 
@@
 \end{align}
-{\color{red} is in total} 
+{\color{\chon} is in total} 
 agreement with Eq. A(6) of Ref.~\onlinecite{cabellosPRB09}.

### 2014-11-17 res.tex

@@
 The calculation of the LDA is done using the ABINIT code with ???
-pseudopotentials, that are of the BK form, and thus the $\bfv^\nl$
-contribution can be calculated. We use the DP code for this end.
+pseudopotentials, that are of the Kleinman-Bylander form, and thus the $\bfv^\nl$
+contribution can be calculated, as explained in Appendix \ref{vesnl}; 
+we use the DP code for this end.\cite{francesco}

### 2014-11-17 paper-v-0.tex
added intro to main document

@@
 
+%%%
+% notes:
+% nonlocal or non-local
+%%%
+
 \begin{document}
@@
 
-\input{intro}
+\section{Introduction}\label{intro}
+
+Nonlinear optics has been a relevant topic in physics research for the past 
+fifty years and has gained a wide range of applications in 
+many areas of technology and research. In material science,
+Second-harmonic generation (SHG) is of particular importance.
+It was one of the first nonlinear processes to be observed and is now one
+of the most used. It has proven to be very effective as an 
+optical probe for material surface and interfaces. 
+\cite{hughesPRB96, 
+     guyot-sionnestPRB88, 
+     lupkeSSR99, 
+     downerPSSA01, 
+     shenAPB99, 
+     shenNAT89, 
+     chenPRL81, 
+     maytorenaPRB98}
+SHG surface spectroscopy is a crucial tool for research and development
+in microelectronics, semiconductors, nanomaterials, and many more recent 
+areas of great scientific and commercial interest.
+
+Numerous mathematical frameworks have been developed to study the 
+nonlinear optical response of materials; 
+in the context of this paper, we note two in particular. 
+The first, called the ``velocity gauge'' or 
+``$\mathbf{v}\cdot\mathbf{A}$ gauge,'' features in early 
+research in this field. 
+Butcher and McLean\cite{butcherPOPS63}, and 
+Aspnes\cite{aspnesPRB72} 
+laid the initial groundwork by deriving explicit expressions for the 
+nonlinear susceptibility tensor. However, these either diverged at 
+low frequencies or were limited to zincblende crystal structures. 
+A few decades later, 
+Ghahramani \emph{et al}.\cite{ghahramaniPRB91} 
+realized a full-band-structure calculation for Si/Ge superlattices and 
+found significant surface SHG while providing proof that the expressions 
+do not diverge at low frequencies for clean, cold semiconductors. 
+Sipe \emph{et al}.\cite{sipePRB93}
+separated the interband and intraband one-electron contributions 
+of the susceptibility tensor and further verified the non-divergent behavior. 
+Levine\cite{levinePRB94} 
+introduced the scissors operator within the Local Density Approximation
+(LDA) and found good agreement with experimental results. 
+The second, called the ``length gauge'' or 
+``$\mathbf{r}\cdot\mathbf{E}$ gauge'' has various advantages over 
+the velocity gauge. 
+Aversa and Sipe\cite{aversaPRB95} 
+presented divergence free expressions for the susceptibility tensor using 
+the length gauge that are significantly easier to derive and calculate. 
+Cabellos \emph{et al}.\cite{cabellosPRB09} 
+elaborated a detailed derivation of the length gauge 
+formalism and a comparison with the velocity gauge for bulk GaAs. They 
+proved that the results are indeed gauge-invariant. 
+This presents a strong case for using the length over the velocity gauge 
+for more generalized and less expensive calculations of the nonlinear
+response of a system.
+
+The inclusion of nonlocal effects, such as the scissors correction, has been
+correctly included in both the velocity and length guage schemes. 
+In the former, both 
+Rashkeev \emph{et al}.\cite{rashkeevPRB98} and 
+Hughes and Sipe\cite{hughesPRB96} 
+accounted for the scissors correction when calculating the nonlinear 
+susceptibility tensor.
+Leitsmann \emph{et al}.\cite{leitsmannPRB05} 
+included excitonic and local field interactions working with GaAs.
+Nastos \emph{et al}.\cite{nastosPRB05} demonstrated the correct 
+way to calculate the nonlinear response using the scissored Hamiltonian 
+in the length gauge providing better agreement with experiment.
+
+Concerning the optical response of surfaces and interfaces,
+Mendoza and Moch\'an\cite{mendozaPRB96} produced an early review of 
+theoretical SHG spectra from surfaces. 
+Arzate and Mendoza\cite{arzatePRB01} improved upon that work and
+went into a detailed analysis of the different contributions to the 
+SHG spectra of a surface and the nuanced relationship between bulk, 
+surface, interband, intraband, $1\omega$ and $2\omega$ terms. 
+Calculated surface SHG spectra using these schemes coincide with experimental
+results in both the velocity gauge
+\cite{mendozaPRB01,mendozaPRL98,reiningPRB94} and the length gauge approach
+\cite{sharmaPRB03}. 
+Perhaps of more relevance to this work, 
+Mendoza \emph{et al}.\cite{mendozaPRB06} and 
+Mej\'ia \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
+the linear and nonlinear responses of semiconductor systems. This model
+allows for obtaining results from selected regions within a slab or bulk,
+including the surface.
+More recently, 
+Trolle \emph{et al}.\cite{trollePRB14} 
+elaborate on a new theoretical model for surface SHG of 
+mono-layer MoS$_{2}$. Their approach is based on the tight-binding model 
+fitted to GW results and include excitonic effects from the 
+Bethe-Salpeter exciton picture. However, this new model is not 
+\emph{ab initio} and has no consideration for a layered approach.
+
+The most complete approach will be a combination of the different techniques 
+that lead to more efficient and accurate calculations. Therefore,
+the goal of this paper is to present newly derived expressions for the 
+nonlinear susceptibility tensor $\chi^{abc}(-2\omega;\omega,\omega)$ 
+in the length gauge formalism
+that include nonlocal terms from the scissors approximation 
+and the spatial contribution from the pseudopotentials and a 
+layer-by-layer approach convenient for surface SHG calculations. We are
+able to divide the resulting tensor in interband, intraband, $1\omega$, and 
+$2\omega$ contributions. This approach is unprecedented and yet produces 
+results that are in complete agreement with the literature.
+
+The paper is organized as follows. 
+In Sec. \ref{theory} we present the 
+relevant steps for deriving the surface second-order susceptibility tensor 
+within the length gauge formalism. This derivation includes the addition of 
+the nonlocal terms mentioned above and a layer-by-layer analysis that have 
+been absent in previous works. 
+In Sec. \ref{results} we show the results of this 
+reformulation with a layer-by-layer study of a clean Si(111) surface with a 
+$2 \times 1$ reconstruction. These calculations are done with \emph{ab initio}
+programs based on density-functional theory (DFT) within the local-density 
+approximation (LDA) using Troullier-Martins pseudopotentials. We compare 
+results from before and after adding the different nonlocal contributions.
+Finally, in Sec. \ref{conclusions} we submit our conclusions.
 
@@
 
-\section{Test Case}
+\section{Test Case}\label{results}
 

### 2014-11-18 paper-v-0.tex
results
first round of corrections to results

@@
 %\input{/Users/bms/util/definitions}
-\def\chon{black}
+%\def\chon[1]{\color{red} #1}
+\def\chon{red}
 \input{definitions}
@@
 % notes:
-% nonlocal or non-local
+% nonlocal, not non-local
+% check for tense consistency
+% surface SHG = SSHG
+% $\bfk$-points, not $\bfk$ points
 %%%
@@
 Concerning the optical response of surfaces and interfaces,
-Mendoza and Moch\'an\cite{mendozaPRB96} produced an early review of 
+Mendoza and Moch\'an\cite{mendozaPRB96a} produced an early review of 
 theoretical SHG spectra from surfaces. 
@@
 
-In this section we present a test case in order to check the
-consistency of the approach presented above. To do this end,  we have chosen
+In this section we present a 
+{\color{\chon} relevant test case} to check the
+consistency of {\color{\chon} our approach. We have selected}
 a clean Si(001) surface with a $2\times 1$ surface reconstruction.
 The slab for such a surface could be chosen to be centrosymmetric 
-by having the fron and the back surface with the same $2\times 1$
-reconstruction, however we terminate with hydrogen one of the
-surfaces, thus having it as an ideally bulk terminated Si
-surface. Indeed, the H atoms simply saturate the dangling bonds of the
-bulk-like Si atoms at the surface, as seen in Fig.~\ref{si2x1}. \textcolor{red}{details of the
- relaxation and some references}
+by {\color{\chon} creating} the front {\color{\chon} and back} 
+surfaces with the same $2\times 1$
+reconstruction{\color{\chon} . However, we choose to} terminate one 
+of the surfaces with hydrogen {\color{\chon} producing an ideal  
+terminated} bulk Si surface. {\color{\chon} The} H atoms simply saturate the 
+dangling bonds of the bulk-like Si atoms at the surface, as seen in 
+Fig.~\ref{si2x1}.
+\textcolor{red}{details of the relaxation and some references}
 \begin{figure}
@@
 \includegraphics[scale=.8]{images/si2x1-crop}
-\caption{(color on line) The slab shows a front clean Si(001)$2\times 1$ surface,
-  while the back surface is ideally terminated Si bulk, where the  
-  dangling bonds are H (small balls) saturated. This picture shows 12  
-  Si atomic layers and one H atomic layer. 
+\caption{(color on line) The slab shows a 
+    {\color{\chon} clean Si(001)$2\times 1$ front surface 
+    with an ideal terminated Si bulk back surface. The}
+    dangling bonds are H (small balls) saturated. 
+    This {\color{\chon} image depicts} 12 Si atomic layers
+    {\color{\chon} with one} H atomic layer. 
 \label{si2x1}} 
 \end{figure} 
-The idea for such an slab is that the cristalline symmetry of the
-H terminated surface imposes that $\chi_{\mathrm{H}}^{xxx}=0$, while
-for the $2\times 1$ surface there is not such symmetry restriction and
-then, $\chi_{2\times 1}^{xxx}\ne 0$.
-Therefore, due to this fact, calculating $\chi^{xxx}$ for the full-slab
-or the half-salb that contains the $2\times 1$ surface,\cite{nota1}
- ought to give the same result, since the contribution from the H
- saturated  surface is zero any way. Then, for this slab one must check
- whether or not this relation is satisfied,
+The idea {\color{\chon} behind this slab configuration} is that the 
+cristalline symmetry of the H terminated surface imposes that 
+$\chi_{\mathrm{H}}^{xxx}=0${\color{\chon} . The $2\times 1$ surface has no 
+such restrictions, so} $\chi_{2\times 1}^{xxx}\ne 0$.
+Thus, calculating $\chi^{xxx}$ for the full-slab, or the 
+{\color{\chon} half-slab containing} the $2\times 1$ surface\cite{nota1}
+{\color{\chon} should yield the} same result since the contribution from the H
+saturated surface is zero {\color{\chon} regardless. 
+We must check that the following 
+relationship is satisfied for this particular slab}
 \begin{align}\label{hs}
@@
 $\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
-$C^\ell(z)=1$ un the upper half of the salb, that contains the
-$2\times 1$ surface reconstruction, as seen in
-Fig.~\ref{si2x1},
-whereas $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
+$C^\ell(z)=1$ {\color{\chon} from the upper half containing the }$2\times 1$ 
+surface reconstruction, as seen in Fig.~\ref{si2x1},
+and $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
 $C^\ell(z)=1$ through the full slab.
-Indeed, in what follows we show the results of such a comparison.
+{\color{\chon} We show the results for this comparison in the remainder 
+of this section}
 
 The self-consistent ground state and their Kohn-Sham states were
-calculated in the frame-work of DFT-LDA,
-with the use of the plane-wave ABINIT
-code.\cite{abinit}
-We used
-Troullier-Martins
-pseudopotentials,\cite{troullierPRB91} 
-that are 
-fully separable nonlocal pseudopotentials in the 
-Kleinman-Bylander 
-form;\cite{kleinmanPRL82}
-the contribution of $\bfv^\nl$ to Eq.~\eqref{chis},
-can be calculated, as explained in Appendix \ref{vesnl}; 
-we use the DP code for this end.\cite{francesco}
-The $2\times 1$ and the H surfaces are relaxed \textcolor{red}{Nicolas 
-  some details}, where we find agreement with previous results for 
-both surfaces.\cite{relax} For instance the Si-H bond distance is 1.48 
-$\AA$ ??. The unit cell has two atoms, making the calculation
-computational very heavy. The vacuum size is equivalent to one quarter
-of the
-size of the slab, thus avoiding spurious effects from the possible
-tunneling of the wave function of the contiguous surfaces of the full
+calculated in the {\color{\chon} DFT-LDA framework using} the plane-wave 
+ABINIT code.\cite{abinit}
+We used Troullier-Martins pseudopotentials\cite{troullierPRB91} that are 
+fully separable nonlocal pseudopotentials in the Kleinman-Bylander 
+{\color{\chon} form.\cite{kleinmanPRL82}
+The contribution of $\bfv^\nl$ to Eq.~\eqref{chis} can be carried out 
+as detailed in Appendix \ref{vesnl} using the DP code for 
+the actual calculation}.\cite{francesco}
+The $2\times 1$ and {\color{\chon} H terminated} surfaces are relaxed 
+\textcolor{red}{Nicolas some details}, 
+{\color{\chon} and we} find agreement with previous results for 
+both surfaces.\cite{relax} 
+{\color{\chon} For instance the Si-H bond distance is 1.48 \AA ??.} 
+The unit cell has two atoms {\color{\chon} making for a computationally
+expensive calculation.} 
+The vacuum size is equivalent to one {\color{\chon} quarter the} 
+size of the slab, {\color{\chon} avoiding the effects produced by possible 
+wave-function tunneling} from the contiguous surfaces of the full
 crystal formed by the repeated super cell scheme.\cite{mendozaPRB06}
 
-Spin-orbit effects, local field effects, and the consequences of the
-electron-hole
-attraction\cite{leitsmannPRB05,trollePRB14} on the SHG process are neglected. 
-Although all these effects are important for the optical response of a
-semiconductor, 
-their calculation is still an open question and a numerical challenge
-that ought to be pursued. 
-However this endeavor is beyond the scope of this paper. 
-For a given size  
-of the slab,  we find converged spectra
-for all the quantities of interest in this
-work, 
-and the most important parameters are
-an energy cut-off of 10 Ha,
-a number 
-of conduction bands equal to the number of valence 
-bands,
-and a set of 244 $\bfk$-points
-is used for
-the linear analytic tetrahedron method used in the evaluation
-the Brillouin zone integrals for of Eq.~\eqref{chis}, where special care
-was taken to examine the double resonances.\cite{nastosPRB05}
-Double resonances occur if for a given frequency $\omega$
-there can be resonant
-transitions at both frequencies $\omega$ and 2$\omega$,
- that is, if there is a
-region in the Brillouin zone such that
-$\omega_{cv}(\bfk)=2 \omega_{c'v}(\bfk)$
-For these $\bfk$ points the perturbation theory used to calculate
-the spectrum breaks down, since there is real population excited,
-which in a correct calculation must be taken into account. These
-points introduce sharp spikes in the spectrum that can, in
-principle, affect the low-frequency results, since the response at
-frequencies below the band gap is computed from the Kramers-Kronig
-relation. However, in agreement with Ref.~\onlinecite{nastosPRB05}
- we find here
-that the double resonances affect the low-frequency results by less
-than 2\%. 
-Finally, all the spectra are calculated with an energy Gaussian smearing of  0.15 eV {\color{red}Chon??} 
+{\color{\chon} Spin-orbit, local field, and electron-hole 
+attraction\cite{leitsmannPRB05,trollePRB14} 
+effects on the SHG process are all neglected.}
+Although these are important factors in the optical response of a semiconductor,
+{\color{\chon} their efficient calculation is still numerically challenging and 
+under debate. This merits further study but is beyond the scope of this paper.}
+For a given {\color{\chon} slab size, we find the} converged spectra 
+{\color{\chon} to obtain the relevant parameters. The most important of 
+these are: an energy cut-off of 10 Ha, an equal number of conduction and 
+valence bands, and a set of 244 $\bfk$-points.} 
+{\color{\chon} The $\bfk$-points are used} for the linear analytic 
+tetrahedron method {\color{\chon} for evaluating the}
+Brillouin zone integrals in Eq.~\eqref{chis}, 
+where special care was taken to examine the double resonances.\cite{nastosPRB05}
+{\color{\chon}
+Double resonances occur at a given frequency $\omega$ if there can be 
+resonant transitions at both $\omega$ and 2$\omega$ frequencies. This 
+implies that there is a region in the Brillouin zone where 
+$\omega_{cv}(\bfk)=2 \omega_{c'v}(\bfk)$. Since there is a real population 
+excitation, the perturbation theory used to calculate the spectrum breaks 
+down at these $\bfk$-points. This must be taken into account in a more 
+complete calculation. These points introduce sharp spikes in the spectrum. 
+In principle, these can affect the low frequency results since the response 
+is computed from the Kramers-Kronig relation for frequencies below the band 
+gap. However, in agreement with Ref.~\onlinecite{nastosPRB05} we find that 
+the double resonances affect the low frequency results by less than 2\%. 
+All spectra were calculated with a Gaussian smearing of 0.15 eV.}
 \begin{figure}
@@
 \caption{(color on line) 
-$|\chi_{\mathrm{half-slab}}^{xxx}|$ vs $\hbar\omega$ for different slabs with 16, 24, 32 
-and 40 atomic Si layers. The fron surface of the slab presents a 
-$2\times 1$ reconstruction, whereas the back sufrace is ideally bulk 
-terminated where the dangling bonds are H-saturated. 
+$|\chi_{\mathrm{half-slab}}^{xxx}|$ vs $\hbar\omega$ 
+{\color{\chon} for the slab} 
+with 16, 24, 32, and 40 atomic Si layers. The front surface 
+{\color{\chon} is in a clean $2\times 1$ reconstruction and the back 
+surface is an ideal terminated bulk H-saturated dangling bonds.}
 \label{fig1}} 
@@
 $\chi^{xxx}_{\mathrm{full-slab}}$
-vs $\hbar\omega$ for a slab of 32 
-atomic Si layers plus one layer of H. 
+vs $\hbar\omega$ for a slab {\color{\chon} with} 32 
+atomic Si layers plus one {\color{\chon} H layer}. 
 \label{fig2}} 
@@
 
-To obtain Eqs.~\eqref{tau.1}  and \eqref{tau.1n}, required in
-Eq.~\eqref{chis},
-we see that the
-evaluation of the commutator $[\bfr,\bfv^\nl]$ is required as it is
-outlined in Appendix \ref{appvnl}; second order derivatives
-are required to do such evaluation, making the numerical
-procedure very
-time consuming. Indeed this time is much higher than the already
-lengthy time needed for the calculation of the contribution of
-$\bfv^\nl$, proportional to only first order derivatives, as seen in
-Appendix \ref{vesnl}. 
-On top of this, the memory required is also very demanding for both
-$\bfv^\nl$ and $[\bfr,\bfv^\nl]$.
-However, the contribution coming from 
-$[\bfr,\bfv^\nl]$, turns out to be very small,\cite{valerie} and
-therefore we neglect it in this work.
+{\color{\chon} 
+We must evaluate the $[\bfr,\bfv^\nl]$ commutator in order to obtain 
+Eqs.~\eqref{tau.1} and \eqref{tau.1n} that are required for 
+Eq.~\eqref{chis}, as outlined in Appendix \ref{appvnl}. Computing second 
+order derivatives is required thus making the numerical procedure very 
+time consuming. This adds significantly to the already lengthy time needed 
+for the calculation of the $\bfv^\nl$ contribution and is proportional to 
+the first order derivatives. Details can be found in Appendix \ref{vesnl}. 
+Memory requirements are also increased for both $\bfv^\nl$ and 
+$[\bfr,\bfv^\nl]$. However, the contribution from $[\bfr,\bfv^\nl]$ 
+is very small\cite{valerie} and therefore we neglect it in this work.}
 
 In Fig.~\ref{fig1} we show $|\chi_{\mathrm{half-slab}}^{xxx}|$
- for a
-slab of 16, 24, 32, and 40 Si atomic layers. We see that 
-$|\chi_{\mathrm{half-slab}}^{xxx}|$ for
-16 and 24
-layers are different from each other, and different from that of 32
-and 40 layers,
-that are very similar to each other.
+for {\color{\chon} the} slab {\color{\chon} with} 
+16, 24, 32, and 40 Si atomic layers. 
+{\color{\chon} We note that the response differs substantially 
+for 16 and 24 layers but is quite similar for 32 and 40 layers.}
 As explained above,
-the calculation of the contribution given by
-$\bfv^\nl$ is computational very demanding. A good
-compromise between the accuracy in the convergence of
-$\chi^{xxx}_{\mathrm{half-slab}}$,
- as a function of the number
-of atomic layers in the slab, and the computational burden, is to take
-the slab with 32 Si atomic layers, as representative of our system
-in question.
+the calculation of {\color{\chon} the $\bfv^\nl$ contribution is 
+computationally expensive.} 
+A good compromise between the accuracy in the convergence of
+$\chi^{xxx}_{\mathrm{half-slab}}$ as a function of the number
+of layers in the slab, and the computational 
+{\color{\chon} expense is to consider} 
+the slab with 32 Si atomic layers as 
+{\color{\chon} an accurate representation} of our 
+{\color{\chon} system.}
 In Fig.~\ref{fig2}
@@
 vs. 
-$\chi^{xxx}_{\mathrm{full-slab}}$, 
-for the four different possibilities of including or not including the
-effects of $\bfv^\nl$ and having or not having a scissors correction
+$\chi^{xxx}_{\mathrm{full-slab}}$ 
+{\color{\chon} for} the four different possibilities 
+{\color{\chon} between} including or not including the
+effects of $\bfv^\nl$ {\color{\chon} or the} scissors correction
 $\hbar\sigma$.
-We see that for all four instances the comparison between
-$\chi^{xxx}_{\mathrm{half-slab}}$ 
-and
-$\chi^{xxx}_{\mathrm{full-slab}}$,
-is very good. Indeed, where the value of $|\chi^{xxx}|$ is large the
-difference is very small, and when its value is small, the difference
-are just slightly larger, but the spectra is so close to zero that the
-differences are not very meaning full; this diferences would become
-smaller as the number of atomic layers is increased. However, with 32
-layers is more than enough to 
-confirm that the extraction of the surface second
-harmonic susceptibility from the $2\times 1$ surface is 
-readily possible using the formalism contained in Eq.~\eqref{chis}.
-
-Although in a forthcoming publication we will present a study of SSHG
-from several Si surfaces where experimental results are available, we
-proceed to explain, without confronting with experiment ({\color{red}
-Nicolas are there any for this particular component?}), some of the
-features seen in 
-$|\chi^{xxx}_{2\times 1}|$,  
-that is of coures given by
-$|\chi^{xxx}_{\mathrm{half-slab}}|$.
-First of all, from Fig.~\ref{fig2} 
-we see a series of resonances 
-that come from the 1$\omega$ and 2$\omega$ terms in
-Eq.~\eqref{chis}. The 2$\omega$ resonances naively would start below
-$E_g/2$ where $E_g=3.4$ eV is
-the band gap of bulk Si. However we see that there are several
-resonances well below 1.7 eV, and of course, these resonances are
-coming from the electronic states of the $2\times 1$ surface, that lie
-inside the gap, that are the well known surface states.
+We see that for all four instances the 
+{\color{\chon} difference between responses is quite small}. 
+{\color{\chon} Indeed, when the value $|\chi^{xxx}|$ 
+is large the difference between the two is very small; 
+when the value is small the difference increases only slightly, 
+but the spectra is so close to zero that these it is negligible. 
+These differences would decrease as the number of atomic layers 
+increases. We note that 32 layers in the slab} is more than enough 
+to confirm that the extraction of the surface second harmonic 
+susceptibility from the $2\times 1$ surface is readily possible 
+using the formalism contained in Eq.~\eqref{chis}.
+
+{\color{\chon} We will present a study of SSHG from several 
+Si surfaces in a forthcoming publication with comparison to experimental 
+results. However, without referring to experiment, we proceed to explain 
+some of the features seen in $|\chi^{xxx}_{2\times 1}|$ that are obtained 
+by calculating $|\chi^{xxx}_{\mathrm{half-slab}}|$.}
+({\color{red} Nicolas are there any for this particular component?})
+{\color{\chon} First,} from Fig.~\ref{fig2} we note a series of resonances 
+that {\color{\chon} derive from} 1$\omega$ and 2$\omega$ terms in
+Eq.~\eqref{chis}. {\color{\chon} We could naively assume that the
+2$\omega$ resonances} start below $E_g/2$ where $E_g=3.4$ eV is
+the band gap of bulk Si. However, we also {\color{\chon} observe} 
+that there are several resonances well below 1.7 eV
+{\color{\chon} that come} from the electronic states of the 
+$2\times 1$ surface, that lie inside the gap and are the 
+well known surface states.
 We see that the effect of $\bfv^\nl$ reduces the value of   
 $|\chi^{xxx}_{2\times 1}|$,\cite{note2}  
-without changing the form of the spectra, as can be confirmed from
+without changing the 
+{\color{\chon} spectral shape}, as can be confirmed from
 both cases of zero and non-zero scissors correction.
-Also, we see that the scissors correction shifts the spectra to higher
-energies, as expected, however, contrary to the case of linear optics,\cite{cabellosPRB09}
-the shifting is not rigid. This is because, the second harmonic optical
-response mixes 1$\omega$ and 2$\omega$ transitions, as can be seen from
-Eq.~\eqref{chis}, and thus the spectra is not rigidly shifted. 
+{\color{\chon} We note that the scissors correction 
+shifts the spectra to higher energies as expected. However, contrary 
+to the case of linear optics\cite{cabellosPRB09} this shit is not 
+rigid. This is because the second harmonic optical response mixes 
+$1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
+for the non-rigid shift.}
 \begin{figure}
@@
 $\chi^{xxx}_{2\times 1}$
-vs $\hbar\omega$ for a slab of 32 
-atomic Si layers plus one layer of H, for two different values of 
+vs $\hbar\omega$ for a slab {\color{\chon} with} 32 
+atomic Si layers plus one {\color{\chon} H layer}, 
+for two different values of 
 the scissors correction $\hbar\sigma$.
@@
 Finally, in Fig.~\ref{fig3} we compare 
-$|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$. Indeed,
-the value of $\hbar\sigma=0.5$ eV, used above, is the ``average'' GW gap taken from 
-Refs.~\onlinecite{rholfingPRB95} which is in agreement with \onlinecite{garciaCPC01}, and  
-the value of $\hbar\sigma=0.63$ eV is the ``average'' GW gap taken from 
-Refs.~\onlinecite{asahiPRB00}, where more $\bfk$ points in the
-Brillouin zone were used.
-We see that the fist two peaks are almost rigidly shifted with a small
-difference in height, while the rest of the peaks are modified
-substantially. This behavior, comes from the fact that the first two
-peaks are almost related to the 1$\omega$ resonances of
-Eq.~\eqref{chis}, and the other peaks are a combination of 1$\omega$
-and 2$\omega$ resonances, thus giving a more different spectrum. This
-way we see that small changes in the value of the scissors shift can affect the
-$|\chi^{xxx}_{2\times 1}|$ quite dramatically.
+$|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$. 
+{\color{\chon} The} value of $\hbar\sigma=0.5$ eV, used above, 
+is the ``average'' GW gap taken from 
+Refs.~\onlinecite{rholfingPRB95} 
+{\color{\chon} and} is in agreement with \onlinecite{garciaCPC01}. 
+{\color{\chon} The} value of $\hbar\sigma=0.63$ eV is the ``average'' 
+GW gap taken from Refs.~\onlinecite{asahiPRB00}, 
+where more {\color{\chon}$\bfk$-points} in the Brillouin zone were used.
+We see that the {\color{\chon} first} two peaks are almost rigidly 
+shifted with a small difference in {\color{\chon} height while} 
+the rest of the peaks are modified substantially. 
+This {\color{\chon} behavior comes} from the fact that the first two
+peaks are {\color{\chon} almost exclusively related} to the 
+1$\omega$ resonances of
+Eq.~\eqref{chis}{\color{\chon} . The} other peaks are a combination 
+of 1$\omega$ and 2$\omega$ resonances 
+{\color{\chon} and yield a more varied} spectrum. 
+This way we see that small changes in the value of the 
+scissors shift can affect the
+$|\chi^{xxx}_{2\times 1}|$ {\color{\chon} spectrum} quite dramatically.
 
@@
 mendozaPRL98,
-mendozaPRB96,
+mendozaPRB96a,
 mendozaPRB97,

### 2014-11-19 paper-v-0.tex

@@
 %\def\chon[1]{\color{red} #1}
-\def\chon{red}
+\def\chon{black}
+\def\chonn{red}
+\def\nicolas{red}
 \input{definitions}
@@
 % $\bfk$-points, not $\bfk$ points
+% Nature Materials de valerie
+% Local fields?
 %%%
@@
 
-Nonlinear optics has been a relevant topic in physics research for the past 
-fifty years and has gained a wide range of applications in 
-many areas of technology and research. In material science,
-Second-harmonic generation (SHG) is of particular importance.
-It was one of the first nonlinear processes to be observed and is now one
-of the most used. It has proven to be very effective as an 
-optical probe for material surface and interfaces. 
-\cite{hughesPRB96, 
+In recent years, surface nonlinear optical spectroscopies and in 
+particular, surface second-harmonic generation SSHG have evolved as useful 
+nondestructive and noninvasive tool to study properties of surfaces 
+and interfaces: atomic structure, phase transitions, and adsorption of 
+atoms,1–11{\color{\chonn} [incluye estas de mejiaPRB02 y combinalas con
+  las tuyas y tambien pon la de salazarPRB14 y busca referencia para
+  ??, checando q sean relevantes con SSHG]} 
+among others. \cite{hughesPRB96, 
      guyot-sionnestPRB88, 
@@
      chenPRL81, 
-     maytorenaPRB98}
-SHG surface spectroscopy is a crucial tool for research and development
-in microelectronics, semiconductors, nanomaterials, and many more recent 
-areas of great scientific and commercial interest.
-
-Numerous mathematical frameworks have been developed to study the 
-nonlinear optical response of materials; 
-in the context of this paper, we note two in particular. 
-The first, called the ``velocity gauge'' or 
-``$\mathbf{v}\cdot\mathbf{A}$ gauge,'' features in early 
-research in this field. 
-Butcher and McLean\cite{butcherPOPS63}, and 
-Aspnes\cite{aspnesPRB72} 
-laid the initial groundwork by deriving explicit expressions for the 
-nonlinear susceptibility tensor. However, these either diverged at 
-low frequencies or were limited to zincblende crystal structures. 
-A few decades later, 
-Ghahramani \emph{et al}.\cite{ghahramaniPRB91} 
-realized a full-band-structure calculation for Si/Ge superlattices and 
-found significant surface SHG while providing proof that the expressions 
-do not diverge at low frequencies for clean, cold semiconductors. 
-Sipe \emph{et al}.\cite{sipePRB93}
-separated the interband and intraband one-electron contributions 
-of the susceptibility tensor and further verified the non-divergent behavior. 
-Levine\cite{levinePRB94} 
-introduced the scissors operator within the Local Density Approximation
-(LDA) and found good agreement with experimental results. 
-The second, called the ``length gauge'' or 
-``$\mathbf{r}\cdot\mathbf{E}$ gauge'' has various advantages over 
-the velocity gauge. 
-Aversa and Sipe\cite{aversaPRB95} 
-presented divergence free expressions for the susceptibility tensor using 
-the length gauge that are significantly easier to derive and calculate. 
-Cabellos \emph{et al}.\cite{cabellosPRB09} 
-elaborated a detailed derivation of the length gauge 
-formalism and a comparison with the velocity gauge for bulk GaAs. They 
-proved that the results are indeed gauge-invariant. 
-This presents a strong case for using the length over the velocity gauge 
-for more generalized and less expensive calculations of the nonlinear
-response of a system.
-
-The inclusion of nonlocal effects, such as the scissors correction, has been
-correctly included in both the velocity and length guage schemes. 
-In the former, both 
-Rashkeev \emph{et al}.\cite{rashkeevPRB98} and 
-Hughes and Sipe\cite{hughesPRB96} 
-accounted for the scissors correction when calculating the nonlinear 
+    mendozaPRL98,downerSIA01,cazzanelliNM14}
+Nowadays, SSHG spectroscopy is a crucial tool for research and development
+in microelectronics,\cite{??} semiconductors,\cite{??} nanomaterials,\cite{??} and many more recent 
+areas of great scientific and commercial interest.\cite{??}
+The high surface sensitivity of SSHG spectroscopy is due to the
+fact that, within the dipole approximation, the bulk SHG signal of
+centrosymmetric materials is identically zero, and thus only the
+surface, where the inversion symmetry is broken, can radiate. 
+Much of the foundation of surface science has been built from
+experiments involving emission or scattering of electrons from
+surfaces. These require ultrahigh vacuum (UHV) environments and
+provide no access to buried interfaces. 
+However, SSHG
+is compatibile with non-UHV ambients and has access to
+interfaces buried beneath transparent overlayers. Even when applied to
+surfaces in UHV, the light source and detectors can be aligned and
+used outside the vacuum chamber.
+
+The usefulness of SSHG could be limited by
+the lack of microscopic theoretical
+understanding of the nonlinear spectra. The macroscopic
+phenomenological theory of SSHG, which relates the intensity,
+phase and polarization of detected fields to nonlinear and linear
+susceptibilities at the material interface, is now fully developed. 
+However, microscopic theory, which relates atomic-level structure to
+the nonlinear source polarization, is still being developed.
+In this article we put forward an approach to calculate the
+microscopic second harmonic surface susceptibility that encompasses several
+theoretical features not taken into account before. 
+Within the independent particle approximation (IPA),  
+some frameworks for bulk SHG have been developed to study the 
+nonlinear optical response of bulk materials, where several degrees of
+sophistication were 
+introduced.\cite{butcherPOPS63,aspnesPRB72,sipePRB93,levinePRB94,aversaPRB95,hughesPRB96,rashkeevPRB98}
+
+As it is well known
+Density Functional Theory (DFT) within the Local Density Approximation
+(LDA),\cite{??} the most 
+used framework for \textit{ab initio} calculations,
+underestimates the energy band gap of semiconductors. It is also well
+understood that one has to include the many-body interaction to
+correct for the underestimation of the gap. In this context the so
+called GW approximation\cite{onidaRMP01} is known to correct the band
+gap of most semiconductors\cite{cual}. However this could be a numerically very
+expensive calculation, and thus one uses the much simpler scissors
+operator scheme.\cite{?} This allows to ``open'' the DFT-LDA gap to
+its correct experimental or GW value, when experiments are not available. 
+In this context,
+to correct for the underestimation of the energy band gap of semiconductors, 
+Nastos et al.\cite{nastosPRB05} 
+using the 
+``length gauge'' or 
+``$\mathbf{r}\cdot\mathbf{E}$ gauge'' 
+showed how to correctly include in the SH susceptibility the many-body corrections through the
+scissors operator.
+Later,
+Cabellos \textit{et al}.\cite{cabellosPRB09} 
+elaborated a derivation of the ``velocity gauge'' or
+``$\bfA\cdot\bfv$'' gauge properly including the scissors
+operator, and proved  gauge 
+invariance with respect to the length gauge.
+From these works, it is clear the length gauge is a much better staring
+point to obtain the surface SH (SSH) susceptibility, as it will be
+elaborated in this article. Above only being valid for bulk semiconductors.
+
+Concerning the optical response of surfaces and interfaces,
+Reining \textit{et al.}\cite{reiningPRB94}  
+introduced the concept of a 
+cut function in order to obtain the surface SH 
 susceptibility tensor.
-Leitsmann \emph{et al}.\cite{leitsmannPRB05} 
-included excitonic and local field interactions working with GaAs.
-Nastos \emph{et al}.\cite{nastosPRB05} demonstrated the correct 
-way to calculate the nonlinear response using the scissored Hamiltonian 
-in the length gauge providing better agreement with experiment.
-
-Concerning the optical response of surfaces and interfaces,
-Mendoza and Moch\'an\cite{mendozaPRB96a} produced an early review of 
-theoretical SHG spectra from surfaces. 
-Arzate and Mendoza\cite{arzatePRB01} improved upon that work and
+This cut function is required since one usually uses a slab approach
+when treating semi-infinite surface systems; if the slab is
+centrosymmetric the susceptibility tensor will be identically zero.
+The cut function is such that it separates the nonlinear response
+ the two surfaces of the slab, thus avoiding the destructive
+ interference between them, giving a finite value of the tensor, that
+ one identifies with the SSH susceptibility tensor; if the slab is not
+ centrosymmetric the cut function can be used to separate the
+ different signals coming from either surface of the slab.
+Indeed, one of the main results of this article is to show that the SSH 
+susceptibility tensor obtained by using the cut function is 
+correctly extracted from the slab.   
+After Reining \textit{et al.}\cite{reiningPRB94}, 
+Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02}
+followed upon this work and in particular Ref.~\onlinecite{arzatePRB01}
 went into a detailed analysis of the different contributions to the 
@@
 surface, interband, intraband, $1\omega$ and $2\omega$ terms. 
-Calculated surface SHG spectra using these schemes coincide with experimental
-results in both the velocity gauge
-\cite{mendozaPRB01,mendozaPRL98,reiningPRB94} and the length gauge approach
-\cite{sharmaPRB03}. 
 Perhaps of more relevance to this work, 
-Mendoza \emph{et al}.\cite{mendozaPRB06} and 
-Mej\'ia \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
-the linear and nonlinear responses of semiconductor systems. This model
-allows for obtaining results from selected regions within a slab or bulk,
+Mej\'ia \textit{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
+the nonlinear responses of semiconductor systems, within a 
+Tight-Binding framework. This model 
+allows for obtaining results from selected regions of a system,
 including the surface. 
-More recently, 
-Trolle \emph{et al}.\cite{trollePRB14} 
-elaborate on a new theoretical model for surface SHG of 
-mono-layer MoS$_{2}$. Their approach is based on the tight-binding model 
-fitted to GW results and include excitonic effects from the 
-Bethe-Salpeter exciton picture. However, this new model is not 
-\emph{ab initio} and has no consideration for a layered approach.
-
-The most complete approach will be a combination of the different techniques 
-that lead to more efficient and accurate calculations. Therefore,
-the goal of this paper is to present newly derived expressions for the 
-nonlinear susceptibility tensor $\chi^{abc}(-2\omega;\omega,\omega)$ 
-in the length gauge formalism
-that include nonlocal terms from the scissors approximation 
-and the spatial contribution from the pseudopotentials and a 
-layer-by-layer approach convenient for surface SHG calculations. We are
-able to divide the resulting tensor in interband, intraband, $1\omega$, and 
-$2\omega$ contributions. This approach is unprecedented and yet produces 
-results that are in complete agreement with the literature.
+However, in these references the scissors is either not implemented or
+wrongly implemented.
+
+Finally,  DFT-LDA calculations are often based on the use of pseudopotentials
+that usually contain spatial non-localities to include the interaction
+of the valence electrons with the core electrons.
+As it will be discussed in this article, 
+ this non-local
+dependence on the position vector, introduces corrections to
+the velocity operator of the electron, that have to be included with
+care in the SSH susceptibility. So far, this contribution has been
+neglected in the literature.
+
+Therefore, within the IPA, 
+the most complete approach 
+for the calculation of the SSH susceptibility is one which includes
+i) the scissors correction, ii) the cut function, and iii) the  
+contribution of the non-local part of the pseudopotentials.  
+Therefore,
+the goal of this paper is to derive, within the length gauge, 
+new expressions for the 
+SSH susceptibility tensor 
+$\chi^{\rma\rmb\rmc}(-2\omega;\omega,\omega)$ 
+that include the aforementioned contributions.
+The inclusion of these three contributions 
+makes 
+our scheme unprecedented and opens the possibility to study 
+surface SHG with more versatility and providing 
+more accurate results than previous approaches.
 
@@
 In Sec. \ref{theory} we present the 
-relevant steps for deriving the surface second-order susceptibility tensor 
+relevant steps for deriving the surface second-order susceptibility
+tensor
+$\chi^{\rma\rmb\rmc}(-2\omega;\omega,\omega)$  
 within the length gauge formalism. This derivation includes the addition of 
-the nonlocal terms mentioned above and a layer-by-layer analysis that have 
-been absent in previous works. 
+the terms mentioned above
+that have 
+been absent in previous works. Also,
+our $\chi^{\rma\rmb\rmc}(-2\omega;\omega,\omega)$  
+can be used for a layer-by-layer
+analysis if wanted.
 In Sec. \ref{results} we show the results of this 
-reformulation with a layer-by-layer study of a clean Si(111) surface with a 
-$2 \times 1$ reconstruction. These calculations are done with \emph{ab initio}
-programs based on density-functional theory (DFT) within the local-density 
-approximation (LDA) using Troullier-Martins pseudopotentials. We compare 
+reformulation with a study of a clean Si(001) surface with a 
+$2 \times 1$ reconstruction. 
+We compare 
 results from before and after adding the different nonlocal contributions.
-Finally, in Sec. \ref{conclusions} we submit our conclusions.
+Finally, in Sec. \ref{conc} we give our conclusions.
 
@@
 
-We use the independent particle approximation and neglect local field
-and excitonic effects{\color{\chon}. We assume a classical} electromagnetic field
+We use the IPA and neglect local field  
+and excitonic effects{\color{\chon}.  
+We assume a classical} electromagnetic field
 {\color{\chon} and quantum mechanical matter.}
@@
 
-We will look for the expectation value of the macroscopic current density, 
+We will look for the expectation value of the microscopic current density, 
 $\mathbf{J}$, given by 
 \begin{equation}
-\bfJ=\langle{\mathbf{J}}\rangle=\frac{e}{\Omega}\mbox{Tr}({\rho}\dot{\mathbf{r}})
+\bfJ=\langle{\mathbf{J}}\rangle=\frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}})
 ,
@@
 \end{equation}
-with $\mathbf{v}$ the velocity operator of the electron, and $\Omega$ the
-normalization volume. We calculate the macroscopic polarization density 
+with $\mathbf{v}$ the velocity operator of the electron, and $A$ the
+area of the unit cell. We calculate the macroscopic polarization density 
 $\mathbf{P}$, related to $\bfJ$ by
@@
 \begin{equation}
-P^a(2\omega)=\chi ^{abc}(-2\omega;\omega,\omega)E^{b}(\omega)E^{c}(\omega),  
+\calp^a(2\omega)=\chi ^{abc}(-2\omega;\omega,\omega)E^{b}(\omega)E^{c}(\omega),  
 \label{pshg}
@@
 \begin{equation}\label{h0}
-H^{\mathrm{LDA}}_{0}=\frac{p^{2}}{2m_e}+V
+H^{\mathrm{LDA}}_{0}(\bfr,\bfp)
+=\frac{p^{2}}{2m_e}+V(\bfr,\bfp)
 \end{equation}
@@
 \begin{align}\label{ps}
-V^{\mathrm{ps}}=V(\bfr)+V^{\mathrm{nl}}
+V^{\mathrm{ps}}(\bfr,\bfp)=V(\bfr)+V^{\mathrm{nl}}(\bfr,\bfp)
 ,
@@
 where 
-$V(\bfr)$ and $V^{\mathrm{nl}}$  are the local and 
+$V(\bfr)$ and $V^{\mathrm{nl}}(\bfr,\bfp)$  are the local and 
 non-local contributions to the pseudopotential $V^{\mathrm{ps}}$,
@@
 \begin{equation*}
-H^\sigma_{0}=H^{\mathrm{LDA}}_{0}+S
+H^\sigma_{0}(\bfr,\bfp)=H^{\mathrm{LDA}}_{0}(\bfr,\bfp)+S(\bfr,\bfp)
 \end{equation*}
@@
 \begin{equation}
-S=\hbar \sigma\sum_{n}\int d^{3}k(1-f_{n})
+S(\bfr,\bfp)=\hbar \sigma\sum_{n}\int d^{3}k(1-f_{n})
 |n\mathbf{k}\rangle\langle n\mathbf{k}|,
@@
 \begin{eqnarray*}
-H^{\mathrm{LDA}}_{0}\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
+H^{\mathrm{LDA}}_{0}(\bfr,\bfp)\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
 \label{hamils} \\
-H_{0}^\sigma\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\sigma
+H_{0}^\sigma (\bfr,\bfp)\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\sigma
 (\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
@@
 through $\bfv^\lda_{nm}(\bfk)$ {\color{\chon} which 
-includes} the contribution of 
+includes} only the contribution of 
 the non-local part of the pseudopotentials through
@@
 with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction{\color{\chon}. 
-We need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
+We stress that within the length gauge,
+we need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
 $\omega_{n}^{\sigma}$ at the end of the derivation}
@@
 
-{\color{\chon} Taking} the matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of
+Following Ref.~\onlinecite{sipePRB00} {\color{\chon} we take} the
+matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of 
 Eq.~\eqref{rde}, we obtain 
@@
 We have omitted the frequency arguments of $\chi^{\rma\rmb\rmc}$ for 
-{\color{\chon} convenience of} notation ({\color{blue}shall we??}).
+{\color{\chon} convenience of} notation.
 
@@
 To introduce the
-cut function $\calc^\ell(z)$ in
-the calculation of $\chi_{\rma\rmb\rmc}$, we start from 
+cut function $\calc(z)$ in
+the calculation of $\chi^{\rma\rmb\rmc}$, we start from 
 the operator for the electron current,
@@
 Integrating the microscopic current $\mbf{j}(\mbf{r},t)$ over
-the entire slab gives the averaged macroscopic current density, $\bfJ(t)$. 
+the entire slab gives the averaged microscopic current density, $\bfJ(t)$. 
 If we want the contribution from only one region of the unit cell 
@@
 to the current density from the
-$\ell$-th layer of the slab is given by
+chosen layer of the slab is given by
 \begin{equation}\label{jsz}
-\frac{1}{\Omega}\int d^3r\, \calc^\ell(z)\, \mathbf{j}^{(N)}(\mathbf{r},t)
- \equiv \mathbf{J}^{\ell,(N)}(t),
-\end{equation}
-where $\mathbf{J}^{\ell,(N)}(t)$ is the current of the
-$\ell$-th layer.
+\frac{1}{A}\int d^3r\, \calc(z)\, 
+\mathbf{J} (\mathbf{r},t)
+ \equiv \calbj^{(N)}(t),
+\end{equation}
+where $\calbj^{(N)}(t)$ is the current of the
+layer specified by $\calc(z)$.
 Therefore we define
 \begin{equation}\label{vcal}
-e{\boldsymbol{\mathcal{V}}}^{\ell,\gs}_{mn}(\mathbf{k})
+e{\boldsymbol{\mathcal{V}}}^{\gs}_{mn}(\mathbf{k})
 \equiv
-%\frac{1}{\Omega}
-\int d^3r\, \calc^\ell(z)\,\bfj_{mn}({\bfk};\bfr),
+\int d^3r\, \calc(z)\,\bfj_{mn}({\bfk};\bfr),
 \end{equation}
@@
 \begin{equation}\label{jmac2}
-\bfJ^{\ell,(N)}(2\go)=\frac{e}{\gO}
+\calbj^{(N)}(2\go)=\frac{e}{A}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\calbv^{\ell,\gs}_{mn}(\mathbf{k}) 
+\calbv^{\gs}_{mn}(\mathbf{k}) 
 \rho^{(N)}_{nm}(\bfk;2\go) 
@@
 \end{equation}
-that gives the induced microscopic current of the $\ell$-th layer, to order $N$ 
+that gives the induced microscopic current of the chosen layer, to order $N$ 
 in the external perturbation. 
@@
 \begin{align}\label{intj}
-{\boldsymbol{\mathcal{V}}}^{\ell,\gs}_{mn}({\mathbf k})
+{\boldsymbol{\mathcal{V}}}^{\gs}_{mn}({\mathbf k})
 &=
@@
 \int \mathrm{d}^3 r\,
- \calc^\ell(z)
+ \calc(z)
 \bigg[
@@
 \int \mathrm{d}^3 r\,
- \calc^\ell(z)
+ \calc(z)
  \bigg[
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\left[\frac{\calc^\ell(z) \bfv^\gs +
-\bfv^\gs \calc^\ell(z)}{2}\right]
+\left[\frac{\calc(z) \bfv^\gs +
+\bfv^\gs \calc(z)}{2}\right]
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\calbv^{\ell,\gs}
+\calbv^{\gs}
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \begin{align}\label{nl.4}
-\calbv^{\ell,\gs}
+\calbv^{\gs}
 =
-\frac{\calc^\ell(z) \bfv^\gs +
-\bfv^\gs \calc^\ell(z)}{2}
-,
-\end{align} 
-where the superscript $\ell$ is inherited from $\calc^\ell(z)$.
+\frac{\calc(z) \bfv^\gs +
+\bfv^\gs \calc(z)}{2}
+.
+\end{align} 
+%where the superscript $\ell$ is inherited from $\calc(z)$.
 % and we
@@
 \begin{align}\label{vcali}
-\bfV \to \calbv^{\ell}=\frac{\calc^\ell(z) \bfV +
-\bfV \calc^\ell(z)}{2}
+\bfV \to \calbv=\frac{\calc(z) \bfV +
+\bfV \calc(z)}{2}
 ,
@@
 {\color{\chon} electron} velocity operators $\bfV$ to the new velocity
-operator $\calbv^{\ell}$ that implicitly takes into account the
-contribution of the region of the slab given by $\calc^\ell(z)$.
+operator $\calbv$ that implicitly takes into account the
+contribution of the region of the slab given by $\calc(z)$.
+We note that this modified operator is hermitian as it must.\cite{nota3}
 The operator $\bfV$ could be any of those given by Eq.~\eqref{vop2},
@@
 \begin{align}\label{vopii}
-\calbv^{\ell,\gs}
+\calbv^{\gs}
 &=
-\calbv^{\ell,\lda}
+\calbv^{\lda}
 +
-\calbv^{\ell,S}
+\calbv^{S}
 \nonumber\\
-\calbv^{\ell,\lda}
+\calbv^{\lda}
 &=
-\calbv^{\ell}
+\calbv
 +
-\calbv^{\ell,\nl}
+\calbv^{\nl}
 .
 \end{align}
-We remark that the simple renormalization that gives 
-$\bfv^{\sigma}_{nm}(\bfk)$ 
-in terms of
-$\bfv^{\lda}_{nm}(\bfk)$,
-given in 
-Eq.~\eqref{chon.9}, 
-does not hold between
-$\calbv^{\ell,\sigma}_{nm}(\bfk)$   
-and 
-$\calbv^{\ell,\lda}_{nm}(\bfk)$,
-i.e.
-$\calbv^{\ell,\sigma}_{nm}(\bfk)\ne
-(\go^\gs_{nm}/\go^\lda_{nm})
-\calbv^{\ell,\lda}_{nm}(\bfk)${\color{\chon} .
-To calculate}
-$\calbv^{\ell,\sigma}_{nm}(\bfk)$ 
-we must calculate the matrix elements of 
-$\calbv^{\ell,\lda}$ and $\calbv^{\ell,S}$
+To calculate
+$\calbv^{\sigma}_{nm}(\bfk)$ 
+we calculate the matrix elements of 
+$\calbv^{\lda}$ and $\calbv^{S}$
  (separately)
@@
 Appendices \ref{calpcalc}, \ref{vesnl} and \ref{calvs}.
-
+If not stated differently, calligraphic letters correspond to layer quantities.
+
+The layer-by-layer analysis of Refs. \onlinecite{hoganPRB03,castilloPRB03,mottaCMS10}  
+used $\calc(z)$
+%limiting the current response  
+%to a particular layer of the slab and used  
+to obtain the  
+anisotropic linear optical response of semiconductor surfaces, 
+where, the first formal derivation  
+of this scheme  
+was presented in  
+Ref.~\onlinecite{mendozaPRB06}. 
 To limit the SHG response to one surface, Eq.~\eqref{vcali} 
-for $\calbv^\ell$ was proposed in 
+for $\calbv$ was proposed in 
 Ref.~\onlinecite{reiningPRB94} and later used in Refs.
@@
  and \onlinecite{mejiaRMF04}. 
-The layer-by-layer analysis of Refs. \onlinecite{hoganPRB03,castilloPRB03,mottaCMS10} 
-used the equivalent Eq.~\eqref{sz}, 
-%limiting the current response
-%to a particular layer of the slab and used 
-to obtain the
-anisotropic linear optical response of semiconductor surfaces.
-However, the first formal derivation
-of this scheme 
-for the linear response 
-is presented in
-Ref.~\onlinecite{mendozaPRB06}, 
-and here in this 
-article, for the second harmonic optical response of semiconductors.
+In this article, for the first time, we formally introduce the cut function $C(z)$ 
+for the second harmonic optical response of semiconductors.
 
 Using
-$\bfJ=d\bfP/dt$ 
+$\calbj=d\calbp/dt$ 
 and Eq.~\eqref{jmac2} 
-we obtain the SH polarization of the $\ell$-th layer as
+we obtain the SH polarization of a given layer as
 \begin{equation}\label{Pjikn}
-\bfP^{\ell,(2)}(2\go)
-=\frac{ie}{2\gO\got}
+\calbp^{(2)}(2\go)
+=\frac{ie}{2A\got}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\calbv^{\ell,\gs}_{mn}(\mathbf{k})
+\calbv^{\gs}_{mn}(\mathbf{k})
 \rho^{(2)}_{nm}(\bfk;2\go)
@@
 \begin{align}\label{Pjikn2}
-\chi^{\ell,\rma\rmb\rmc}(-2\go;\go,\go) 
+\chi^{\rma\rmb\rmc}(-2\go;\go,\go) 
 &=
@@
 \sum_{mn}
-\frac{\calv^{\ell,\gs,\rma}_{mn}(\mathbf{k})}
+\frac{\calv^{\gs,\rma}_{mn}(\mathbf{k})}
 {\go^\gs_{nm\bfk}-2\got}
@@
 which gives the susceptibility 
-$\chi^{\ell,\rma\rmb\rmc}(-2\go;\go,\go)$ 
-of the $\ell$-th layer of the slab, 
-%where 
-%$\calbv^\gs$ is given in Eq.~\eqref{vopii},
-where $A=\gO/d$ is the area of the unit
-cell that characterizes the surface of the system, and $d$
-is the surface region {\color{\chon} from} which the {\color{\chon} nonlinear} 
-susceptibility is different from zero.
+$\chi^{\rma\rmb\rmc}(-2\go;\go,\go)$ 
+of the layer of the slab specified by $\calc(z)$. 
 We mention that the units of 
-$\chi^{\ell,\rma\rmb\rmc}(-2\go;\go,\go)$
+$\chi^{\rma\rmb\rmc}(-2\go;\go,\go)$
 are m$^2$/V, as they {\color{\chon} should be} for a surface SH susceptibility.
@@
 \begin{equation}\label{chii}
-\chi^{\ell,\rma\rmb\rmc}_i (-2\go;\go,\go)
+\chi^{\rma\rmb\rmc}_i (-2\go;\go,\go)
 =-\frac{e^3}{A\hbar^22\got}
@@
 \sum_{mn}
-\frac{\calv_{mn}^{\ell,\gs,\rma}}{\go^\gs_{nm}-2\got}
+\frac{\calv_{mn}^{\gs,\rma}}{\go^\gs_{nm}-2\got}
 \left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}}
@@
 \begin{equation}\label{chie}
-\chi^{\ell,\rma\rmb\rmc}_e (-2\go;\go,\go)
+\chi^{\rma\rmb\rmc}_e (-2\go;\go,\go)
 =\frac{ie^3}{A\hbar^22\got}
@@
 \sum_{qmn}
-\frac{\calv_{mn}^{\ell,\gs,\rma}}{\go^\gs_{nm}-2\got}
+\frac{\calv_{mn}^{\gs,\rma}}{\go^\gs_{nm}-2\got}
 \left(
@@
 \begin{equation}\label{calvimchiewn}
-\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{e,\go}]= 
+\mathrm{Im}[\chi^{\rma\rmb\rmc}_{e,\go}]= 
 \frac{\pi |e|^3}{2\hbar^2}
@@
 \left[
-\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{qc}\{r^{\rmb}_{cv}r^{\rmc}_{vq}\}]}
+\frac{\mathrm{Im}[\mathcal{V}^{\gs,\mathrm{a}}_{qc}\{r^{\rmb}_{cv}r^{\rmc}_{vq}\}]}
 {(2\go^\gs_{cv}-\go^\gs_{cq})} 
--\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vq}\{r^{\rmc}_{qc}r^{\rmb}_{cv}\}]}
+-\frac{\mathrm{Im}[\mathcal{V}^{\gs,\mathrm{a}}_{vq}\{r^{\rmc}_{qc}r^{\rmb}_{cv}\}]}
 {(2\go^\gs_{cv}-\go^\gs_{qv})}
@@
 \begin{equation}\label{calvimchiwn}
-\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{i,\go}]= 
+\mathrm{Im}[\chi^{\rma\rmb\rmc}_{i,\go}]= 
 \frac{\pi\vert e\vert^3}{2\hbar^2}
@@
 \left[
-\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
-+\frac{\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\gs,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
++\frac{\mathrm{Re}\left[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
 \Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gs_{cv}} 
@@
 \begin{equation}\label{calvimchie2wn}
-\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{e,2\go}]= 
+\mathrm{Im}[\chi^{\rma\rmb\rmc}_{e,2\go}]= 
 -\frac{\pi |e|^3}{2\hbar^2}
@@
 \sum_{v'\ne
-  v}\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\{r^{\rmb}_{cv'}r^{\rmc}_{v'v}\}]}
+  v}\frac{\mathrm{Im}[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\{r^{\rmb}_{cv'}r^{\rmc}_{v'v}\}]}
 {2\go^\gs_{cv'}-\go^\gs_{cv}}
 - \sum_{c'\ne
-  c}\frac{\mathrm{Im}[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\{r^{\rmc}_{cc'}r^{\rmb}_{c'v}\}]}
+  c}\frac{\mathrm{Im}[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\{r^{\rmc}_{cc'}r^{\rmb}_{c'v}\}]}
 {2\go^\gs_{c'v}-\go^\gs_{cv}}
@@
 \begin{equation}\label{calvimchi2wn}
-\mathrm{Im}[\chi^{\ell,\rma\rmb\rmc}_{i,2\go}]= 
+\mathrm{Im}[\chi^{\rma\rmb\rmc}_{i,2\go}]= 
  \frac{\pi \vert
@@
 \sum_{vc}\frac{4}{(\omega^\gs_{cv})^{2}}
-\left[\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
+\left[\mathrm{Re}\left[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
 \right\}\right] -
-\frac{2\mathrm{Re}\left[\mathcal{V}^{\ell,\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\frac{2\mathrm{Re}\left[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
 \Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gs_{cv}}\right]\delta(\omega^\gs_{cv}-2\omega)
@@
 The real part of each contribution can be obtained through
-a Kramers-Kronig transformation\cite{nicolas} {\color{\chon} and}
-$\chi^{\ell,\rma\rmb\rmc}=
-\chi^{\ell,\rma\rmb\rmc}_{e,\go} 
-+\chi^{\ell,\rma\rmb\rmc}_{e,2\go}
-+\chi^{\ell,\rma\rmb\rmc}_{i,\go} 
-+\chi^{\ell,\rma\rmb\rmc}_{i,2\go}
+a Kramers-Kronig transformation\cite{tacongePRB14} {\color{\chon} and}
+$\chi^{\rma\rmb\rmc}=
+\chi^{\rma\rmb\rmc}_{e,\go} 
++\chi^{\rma\rmb\rmc}_{e,2\go}
++\chi^{\rma\rmb\rmc}_{i,\go} 
++\chi^{\rma\rmb\rmc}_{i,2\go}
 $.
@@
 and thus
-$\chi^{\ell,\rma\rmb\rmc}=\chi^{\ell,\rma\rmc\rmb}$.
+$\chi^{\rma\rmb\rmc}=\chi^{\rma\rmc\rmb}$.
 The various quantities involved in Eqs.~\eqref{chis} are given in
 the Appendix \ref{appe}. 
-We mention that if we take $C^\ell(z)=1$ through out, the layered
-matrix elements $\calbv^{\ell,\sigma}_{nm}$ become standard bulk-like
-$\bfv^{\ell,\sigma}_{nm}$ matrix elements. We mention that in this
+We mention that if we take $\calc(z)=1$ through out, the layered
+matrix elements $\calbv^{\sigma}_{nm}$ become standard bulk-like
+$\bfv^{\sigma}_{nm}$ matrix elements. We mention that in this
 case, Eq.~\eqref{chis} is equivalent to the expressions of
@@
 
-Finally, we calculate the nonlinear surface susceptibility as 
+Finally, we could also calculate the nonlinear surface susceptibility as 
 \begin{equation}\label{chiijksur}
 \bfgchi(-2\go;\go,\go)
-= \sum_{\{\ell\}}\bfgchi^{\ell}(-2\go;\go,\go),
-\end{equation} 
-where $\{\ell\}$
+= \sum_{\{\ell\}}\bfgchi^\ell(-2\go;\go,\go),
+\end{equation} 
+where $\ell$ would denote a particular layer chosen through
+$\calc^\ell(z)$ of Eq.~\eqref{sz} and
+$\{\ell\}$
 is meant to be {\color{\chon} a chosen set of layers.} For instance, 
@@
 either the front or the back surface. 
+% This decomposition of $\bfgchi(-2\go;\go,\go)$ into
+% $\bfgchi^\ell(-2\go;\go,\go)$ would allow to explore the contribution
+% to SH from any particular layer of the system,
+% and answers questions like: which part of the system has the strongest
+% contribution? or  how deep does the surface region needs to be to
+% recover centrosymmetry?, etc.
+\cite{mejiaRMF04}
 %\mathrextcolor{\chon}{Talk about the correct value of $d$}
 
-\section{Test Case}\label{results}
+\section{Results}\label{results}
 
@@
 Fig.~\ref{si2x1}.
-\textcolor{red}{details of the relaxation and some references}
+We take the $z$ coordinate pointing out of the surface and the $x$
+coordinate, along the crystallographic [011] direction, is parallel to the dimmers. 
 \begin{figure}
@@
 cristalline symmetry of the H terminated surface imposes that 
-$\chi_{\mathrm{H}}^{xxx}=0${\color{\chon} . The $2\times 1$ surface has no 
+$\chi_{\mathrm{H}}^{xxx}=0${\color{\chon}.
+ The $2\times 1$ surface has no 
 such restrictions, so} $\chi_{2\times 1}^{xxx}\ne 0$.
+This is due to the fact that along the $y$ direction there is a mirror plane for the
+H-saturated surface, whereas 
+for the $2\times 1$ surface this mirror is lost as the dimers are
+asymmetric along $x$. 
 Thus, calculating $\chi^{xxx}$ for the full-slab, or the 
@@
 $\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
-$C^\ell(z)=1$ {\color{\chon} from the upper half containing the }$2\times 1$ 
+$C(z)=1$ {\color{\chon} from the upper half containing the }$2\times 1$ 
 surface reconstruction, as seen in Fig.~\ref{si2x1},
 and $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
-$C^\ell(z)=1$ through the full slab.
+$C(z)=1$ through the full slab.
 {\color{\chon} We show the results for this comparison in the remainder 
@@
 the actual calculation}.\cite{francesco}
-The $2\times 1$ and {\color{\chon} H terminated} surfaces are relaxed 
-\textcolor{red}{Nicolas some details}, 
-{\color{\chon} and we} find agreement with previous results for 
-both surfaces.\cite{relax} 
-{\color{\chon} For instance the Si-H bond distance is 1.48 \AA ??.} 
+%
+The surfaces have been studied with a lattice constant of 5.43 \AA. 
+Structural optimizations were performed using DFT-LDA
+with the ABINIT code.\cite{abinit}  
+The geometry optimization as been carried out in slabs of 12
+atomic layers where the central four layers where fixed at the bulk
+positions.  
+The structures were relaxed until the Cartesian force components were less and 5 meV/\AA. 
+The geometry optimization for the clean surface gives
+a dimer buckling of 0.721 \AA, and a dimer length of 2.301 \AA.  
+For the  Si(001)$1\times 1$:2H dihydride surface, we have obtained a Si-H bond distance of 1.48 \AA. 
+This results are in good agreement with previous 
+theoretical studies.\cite{caramellaPRB09,mendozaPRB06}
+%
 The unit cell has two atoms {\color{\chon} making for a computationally
@@
 Although these are important factors in the optical response of a semiconductor,
-{\color{\chon} their efficient calculation is still numerically challenging and 
+{\color{\chon} their efficient calculation is still theoretically and  
+  numerically challenging and  
 under debate. This merits further study but is beyond the scope of this paper.} 
@@
 {\color{\chon} 
-We must evaluate the $[\bfr,\bfv^\nl]$ commutator in order to obtain 
+We must evaluate 
+$T^{\rma\rmb}_{nm}=(i/\hbar)[r^\rmb,v^{\nl,\rma}]_{nm}$,
+as outlined in Appendix \ref{vesnl},
+in order to obtain 
 Eqs.~\eqref{tau.1} and \eqref{tau.1n} that are required for 
-Eq.~\eqref{chis}, as outlined in Appendix \ref{appvnl}. Computing second 
+Eq.~\eqref{chis}.
+Computing second 
 order derivatives is required thus making the numerical procedure very 
 time consuming. This adds significantly to the already lengthy time needed 
-for the calculation of the $\bfv^\nl$ contribution and is proportional to 
-the first order derivatives. Details can be found in Appendix \ref{vesnl}. 
+for the calculation of the $\bfv^\nl$ contribution that is
+proportional only to 
+the first order derivatives. 
 Memory requirements are also increased for both $\bfv^\nl$ and 
@@
 
-In Fig.~\ref{fig1} we show $|\chi_{\mathrm{half-slab}}^{xxx}|$
+\subsection{Full-slab results}\label{results}
+
+In Fig.~\ref{fig1} we show $|\chi_{\mathrm{full-slab}}^{xxx}|$
 for {\color{\chon} the} slab {\color{\chon} with} 
@@
 A good compromise between the accuracy in the convergence of
-$\chi^{xxx}_{\mathrm{half-slab}}$ as a function of the number
+$\chi^{xxx}_{\mathrm{full-slab}}$ as a function of the number
 of layers in the slab, and the computational 
@@
 {\color{\chon} system.}
+
+\subsection{Half-slab vs. full-slab}
+
 In Fig.~\ref{fig2}
@@
 $\hbar\sigma$.   
+For these results we chose
+$\hbar\sigma=0.5$ eV, that is the GW gap reported in
+Ref.~\onlinecite{rholfingPRB95,garciaCPC01}. 
 We see that for all four instances the 
@@
 These differences would decrease as the number of atomic layers 
-increases. We note that 32 layers in the slab} is more than enough 
+increases. We remark that 32 layers in the slab} is more than enough 
 to confirm that the extraction of the surface second harmonic 
@@
 using the formalism contained in Eq.~\eqref{chis}.
-
-{\color{\chon} We will present a study of SSHG from several 
-Si surfaces in a forthcoming publication with comparison to experimental 
-results. However, without referring to experiment, we proceed to explain 
-some of the features seen in $|\chi^{xxx}_{2\times 1}|$ that are obtained 
-by calculating $|\chi^{xxx}_{\mathrm{half-slab}}|$.}
-({\color{red} Nicolas are there any for this particular component?})
+This confirms the validity of our theory and is the main result of
+this article; through the proposed layer formalism we can calculate the surface SH
+susceptibility including the effect of the non-local part of the
+pseudopotentials
+and the many-body effects through the scissors correction.
+Our scheme ought to work for any slab.
+
+\subsection{Results for $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$}
+
+We proceed to explain 
+some of the features seen in $|\chi^{xxx}_{2\times 1}|$ that, as
+explained above, are obtained 
+by calculating $|\chi^{xxx}_{\mathrm{half-slab}}|$.
 {\color{\chon} First,} from Fig.~\ref{fig2} we note a series of resonances 
 that {\color{\chon} derive from} 1$\omega$ and 2$\omega$ terms in
-Eq.~\eqref{chis}. {\color{\chon} We could naively assume that the
-2$\omega$ resonances} start below $E_g/2$ where $E_g=3.4$ eV is
-the band gap of bulk Si. However, we also {\color{\chon} observe} 
-that there are several resonances well below 1.7 eV
-{\color{\chon} that come} from the electronic states of the 
-$2\times 1$ surface, that lie inside the gap and are the 
-well known surface states.
+Eq.~\eqref{chis}. 
+Notice that the 
+2$\omega$ resonances start below $E_g/2$ where $E_g=??$ eV is the LDA
+band gap of this surface, and $E_g=??$ eV is the scissor shifted gap for
+$\hbar\sigma=0.5$ eV.
+These resonances
+{\color{\chon} come} from the electronic states of the 
+$2\times 1$ surface, that lie inside the bulk band gap of Si and are the 
+well known electronic surface states.
+In Fig.~\ref{fig3}
 We see that the effect of $\bfv^\nl$ reduces the value of   
-$|\chi^{xxx}_{2\times 1}|$,\cite{note2}  
+$|\chi^{xxx}_{2\times 1}|$,
 without changing the 
-{\color{\chon} spectral shape}, as can be confirmed from
-both cases of zero and non-zero scissors correction.
-{\color{\chon} We note that the scissors correction 
+{\color{\chon} spectral shape}; this also can be confirmed from
+the cases of zero scissors correction from Fig.~\ref{fig2}.
+%
+From Fig.~\ref{fig4}
+{\color{\chon} we note that the scissors correction 
 shifts the spectra to higher energies as expected. However, contrary 
-to the case of linear optics\cite{cabellosPRB09} this shit is not 
-rigid. This is because the second harmonic optical response mixes 
+to the case of linear optics\cite{cabellosPRB09} the shift introduced
+by the scissors correction is not 
+rigid, as pointed out in Ref.~\onlinecite{nastosPRB05}.
+ This is because the second harmonic optical response mixes 
 $1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
@@
 atomic Si layers plus one {\color{\chon} H layer}, 
+with and without the contribution from $\bfv^\nl$.
+\label{fig3}} 
+\end{figure}
+\begin{figure}
+\centering 
+\includegraphics[scale=.8]{plots/fig4}
+\caption{(color on line) 
+$\chi^{xxx}_{2\times 1}$
+vs $\hbar\omega$ for a slab {\color{\chon} with} 32 
+atomic Si layers plus one {\color{\chon} H layer}, 
 for two different values of 
 the scissors correction $\hbar\sigma$. 
-\label{fig3}} 
+\label{fig4}} 
 \end{figure}
-
-Finally, in Fig.~\ref{fig3} we compare 
-$|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$. 
-{\color{\chon} The} value of $\hbar\sigma=0.5$ eV, used above, 
-is the ``average'' GW gap taken from 
-Refs.~\onlinecite{rholfingPRB95} 
-{\color{\chon} and} is in agreement with \onlinecite{garciaCPC01}. 
-{\color{\chon} The} value of $\hbar\sigma=0.63$ eV is the ``average'' 
-GW gap taken from Refs.~\onlinecite{asahiPRB00}, 
-where more {\color{\chon}$\bfk$-points} in the Brillouin zone were used.
-We see that the {\color{\chon} first} two peaks are almost rigidly 
+Also, when we compare 
+$|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$,
+we see that the {\color{\chon} first} two peaks are almost rigidly 
 shifted with a small difference in {\color{\chon} height while} 
@@
 peaks are {\color{\chon} almost exclusively related} to the 
-1$\omega$ resonances of
+2$\omega$ resonances of
 Eq.~\eqref{chis}{\color{\chon} . The} other peaks are a combination 
@@
 This way we see that small changes in the value of the 
-scissors shift can affect the
-$|\chi^{xxx}_{2\times 1}|$ {\color{\chon} spectrum} quite dramatically.
+scissors shift can in general affect the SSH susceptibility 
+spectrum quite dramatically.
+We mention that
+{\color{\chon} the} value of $\hbar\sigma=0.5$ eV, used above, 
+is the ``average'' GW gap taken from 
+Ref.~\onlinecite{rholfingPRB95} 
+{\color{\chon} that} is in agreement with Ref.~\onlinecite{garciaCPC01}. 
+{\color{\chon} The} value of $\hbar\sigma=0.63$ eV is the ``average'' 
+gap taken from Refs.~\onlinecite{asahiPRB00}, 
+where more {\color{\chon}$\bfk$-points} in the Brillouin zone were
+used to calculate its GW value.
+
+Although, local fields are neglected, we expect that parallel to an
+interface they should be in principle 
+small, as the electric field is continuous. So, we would expect that the $xxx$ component 
+of 
+$\bfgchi(-2\go;\go,\go)$, would have an small influence from the local fields.
+Unfortunately the experimental measurement of the $xxx$
+component of $\bfgchi(-2\go;\go,\go)$ is not possible, as the SH 
+radiated intensity would be proportional not only to this component 
+but also to other components of $\bfgchi(-2\go;\go,\go)$,  
+however,
+in a forthcoming publication 
+we will present a study of SSHG from several 
+Si surfaces 
+with comparison to experimental 
+results.
+
+\section{Conclusions}\label{conc}
+
+We have presented a formulation to calculate surface second harmonic
+(SSH) susceptibility tensor $\bfgchi(-2\go;\go,\go)$,  using a length gauge
+formalism and within the independent particle approximation (IPA).
+It includes on the same footing and for the first time in the literature 
+i) the scissors correction, ii) the cut function, and iii) the 
+contribution of the non-local part of the pseudopotentials. 
+We have used a Si(001)$2\times 1$ surface to confirm that our scheme
+correctly obtains the surface response as we confirm
+that 
+$\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go) 
+\approx
+\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go) 
+. 
+$
+Although, one can in principle increase the number of atomic layers,
+$\bfk$-points, etc.  to
+improve even further on the similarity of the half-slab and full-slab results, we
+hava chosen a good compromise between accuracy and the burden and time
+of the computational scheme. 
+We describe the effect of the independent inclusion of the three
+effects mentioned above in the calculation of  
+$\bfgchi(-2\go;\go,\go)$. To wit,
+the scissors correction shifts the spectrum to higher energies, though
+the shifting is not rigid and mixes the 1$\omega$ and 2$\omega$
+resonances, the cut functions
+allow us
+to extract unequivocally   
+$\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$, 
+and the effects of the nonlocal part of the pseudopotentials,
+reduces the value of   
+$|\chi^{xxx}_{2\times 1}(-2\go;\go,\go)|$, keeping the same lineshape. 
+The $xxx$ component of  
+$\bfgchi_{2\times 1}(-2\go;\go,\go)$, 
+could not be
+experimentally  isolated,
+however in a forthcoming publication 
+we will compare our formulation 
+against
+experimental 
+results. 
+We have neglected
+local field 
+and excitonic effects.
+Although these are important factors in the optical response of a semiconductor,
+their efficient calculation is theoretically and 
+numerically challenging and still 
+under debate.\cite{beyond}  
+This merits further study but is beyond the scope of this paper. 
+Nevertheless, the inclusion of aforementioned contributions 
+in
+our scheme opens the unprecedented possibility to study 
+surface SHG with more versatility and providing 
+more accurate results than previous approaches.
 
@@
 \subsection{ Expressions for 
-\texorpdfstring{${\cal V}^{\ell,a}_{nm}(\bfk)$, 
-${\cal C}^{\ell}_{nm}(\bfk)$ 
+\texorpdfstring{${\calbv}_{nm}(\bfk)$}{que} 
 and 
-$({\cal C}^{\ell}_{nm}(\bfk))_{;\bfk}$
-}{Vnm and Cnm}}\label{calpcalc}
+\texorpdfstring{${\cal C}_{nm}(\bfk)$}{que}
+}\label{calpcalc}
 
@@
 space, and $A_{n\bfk}(\bfG)$ {\color{\chon} the} expansion coefficients. Using
-$m_e\bfv=-i\hbar\bfgnabla$ into 
+$m_e\bfv=\bfp=-i\hbar\bfgnabla$ into 
 Eq.~\eqref{vcali}
@@
 \begin{align}\label{eni.2}
-\calbv^\ell_{nm}(\bfk)=
+\calbv_{nm}(\bfk)=
 \frac{\hbar}{2m_e}
@@
 \gd_{\bfG_\parallel \bfG'_\parallel}  
-f^\ell(G_\perp-G'_\perp)
+f(G_\perp-G'_\perp)
 ,
@@
 \begin{align}\label{vnl.9}
-f^\ell(g)=\frac{1}{L}\int_{z_\ell-\gD^b_\ell}^{z_\ell+\gD^f_\ell} e^{igz}dz  
- ,
-\end{align}
-with $f^{\ell*}(g)=f^\ell(-g)$. 
+f(G_\perp)=\frac{1}{L}\int\calc(z) e^{iG_\perp z}dz  
+ ,
+\end{align}
+with $f^*(G_\perp)=f(-G_\perp)$,
+and $L$
+is the length of the super cell. 
 {\color{\chon} The} reciprocal lattice vectors $\bfG$ are 
@@
 associated with the same $G_\parallel$.
+
 Likewise we obtain that
 \begin{align}\label{eni.4}
-\calc^\ell_{nm}(\bfk)=
+\calc_{nm}(\bfk)=
 \sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
 \gd_{\bfG_\parallel \bfG'_\parallel} 
-f^\ell(G_\perp-G'_\perp)
+f(G_\perp-G'_\perp)
 .
 \end{align}  
-If $\calc^\ell(z)=1$, {\color{\chon} then $f^\ell(g)=\gd_{g0}$ and we 
+If $\calc(z)=1$, {\color{\chon} then $f(G_\perp)=\gd_{G_\perp 0}$ and we 
 obtain the full-slab/bulk values, 
-$\bfv_{nm}(\bfk)$ and $\calc^\ell_{nm}(\bfk)=\gd_{nm}$,
+$\bfv_{nm}(\bfk)$ and $\calc_{nm}(\bfk)=\gd_{nm}$,
 from Eqs.~\eqref{eni.2} and \eqref{eni.4}}.
-
-We use Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn},
-along with $[\bfr,F(\bfr)]=0$, valid for 
-and any function $F(\bfr)$, 
-{\color{\chon} to }obtain
-\begin{align} 
-(\calc^\ell_{nm})_{;\bfk}
-&=
-i
-\sum_{q} 
- \left(\bfr_{nq}
-\calc^\ell_{qm}
--
-\calc^\ell_{nq}
-\bfr_{qm}
-\right) 
-+i\bfr_{nm}(\calc^\ell_{mm}-\calc^\ell_{nn}) 
-,
-\end{align} 
-where we remind the reader that $\bfr_{nm}$ 
-{\color{\chon} is} calculated through
-Eq.~\eqref{chon.10} for LDA.
 
@@
 \subsection{Expressions for 
-\texorpdfstring{$(\calv^{\ell,\lda,a}_{nm})_{;k^b}$}{Vnonlocal}
-and \texorpdfstring{$(r^a_{nm})_{;k^b}$}{Vnonlocal}
-for non-local potentials}\label{appvnl}
+\texorpdfstring{$(\calv^{\lda,a}_{nm})_{;k^b}$}{Vnonlocal},
+\texorpdfstring{$(r^a_{nm})_{;k^b}$}{Vnonlocal}
+and \texorpdfstring{$(\calc_{nm})_{;\bfk}$}{Vnonlocal}
+}\label{appvnl}
 
@@
 \begin{align}\label{nmesn}
-(\calv^{\ell,\lda,\rma}_{nm})_{;k^{\rmb}}&=
-\calt_{nm}^{\ell,\rma\rmb}
+(\calv^{\lda,\rma}_{nm})_{;k^{\rmb}}&=
+\calt_{nm}^{\rma\rmb}
 +i
@@
 r^{\rmb}_{nq}  
-\calv^{\ell,\lda,\rma}_{q m}
+\calv^{\lda,\rma}_{q m}
 -
-\calv^{\ell,\lda,\rma}_{nq}   
+\calv^{\lda,\rma}_{nq}   
 r^{\rmb}_{q m}
@@
 r^{\rmb}_{nm}
-\Delta^{\ell,\rma}_{mn}
+\Delta^{\rma}_{mn}
 ,
@@
 \begin{eqnarray}\label{tdel}
-\Delta^{\ell,\rma}_{mn}
+\Delta^{\rma}_{mn}
 =
-\calv^{\ell,\lda,\rma}_{nn}  
+\calv^{\lda,\rma}_{nn}  
 -
-\calv^{\ell,\lda,\rma}_{mm}  
+\calv^{\lda,\rma}_{mm}  
 ,
 \end{eqnarray} 
-and
 \begin{align}\label{tau.1}
-\calt_{nm}^{\ell,\rma\rmb}
+\calt_{nm}^{\rma\rmb}
 =
 \frac{\hbar}{m_e}\gd_{\rma\rmb} 
-C^\ell_{nm} 
--i 
+C_{nm} 
+-\hbar 
 \sum_q  
-[r^{\rmb},v^{\nl,\rma}]_{nq}C^\ell_{qm} 
+L^{\rma\rmb}_{nq} 
+C_{qm} 
+,
+\end{align}   
+and
+\begin{align}\label{lab}
+L_{nm}^{\rma\rmb}
+=\frac{i}{\hbar}[r^{\rmb},v^{\nl,\rma}]_{nm}
 .
 \end{align}  
-The matrix elements $[r^{\rmb},v^{\nl,\rma}]_{nm}$
-are calculated in Appendix \ref{calt}.
+The matrix elements $L^{\rma\rmb}_{nm}$
+are calculated in Appendix \ref{vesnl}.
 % is small
@@
 taking 
-{\color{\chon} $C^\ell(z)=1$ or $C^\ell_{nm}=\gd_{nm}$.}
+{\color{\chon} $C(z)=1$ or $C_{nm}=\gd_{nm}$.}
 
@@
 &=
-t^{\rma\rmb}_{nm}
+T^{\rma\rmb}_{nm}
 +
@@
 \begin{align}\label{tau.1n} 
-t_{nm}^{\rma\rmb}
+T_{nm}^{\rma\rmb}
 =\frac{\hbar}{m_e}\gd_{ab}\gd_{nm} 
--i [r^{\rmb},v^{\nl,\rma}]_{nm} 
+-\hbar L^{\rma\rmb}_{nm} 
 .
@@
 Note that the layered term
-$\calt^{\ell,\rma\rmb}_{nm}$ reduces to $t^{\rma\rmb}_{nm}$
+$\calt^{\rma\rmb}_{nm}$ reduces to $T^{\rma\rmb}_{nm}$
 {\color{\chon} for the full-slab/bulk case.}
 
+Again, we use Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn},
+along with $[\bfr,F(\bfr)]=0$, valid for 
+and any function $F(\bfr)$, 
+{\color{\chon} to }obtain 
+\begin{align} 
+(\calc_{nm})_{;\bfk}
+&=
+i 
+\sum_{q} 
+ \left(\bfr_{nq}
+\calc_{qm}
+-
+\calc_{nq}
+\bfr_{qm}
+\right) 
++i\bfr_{nm}(\calc_{mm}-\calc_{nn}) 
+,
+\end{align} 
+where we remind the reader that $\bfr_{nm}$ 
+{\color{\chon} is} calculated through 
+Eq.~\eqref{chon.10} for LDA.
+
 %%%%
-\subsection{Matrix elements of 
-\texorpdfstring{$\calbv^{\ell,\nl}$}{Vnonlocal},
-\texorpdfstring{$\bfv^\nl$}{Vnonlocal}, and 
-\texorpdfstring{$[\bfr,\bfv^\nl]$}{[r,vnl]}}
+\subsection{
+\texorpdfstring{$\calbv^{\nl}_{nm}$}{Vnonlocal},
+\texorpdfstring{$\bfv^\nl_{nm}$}{Vnonlocal}, and 
+\texorpdfstring{$T^{\rma\rmb}_{nm}$}{[r,vnl]}}
 \label{vesnl}
 
-We take Eq.~\eqref{vcali}
+We take Eq.~\eqref{vnl} into Eq.~\eqref{vcali}
 and use
@@
 \mathrm{exp}(i(\bfk+\bfG)\cdot\bfr)$
-{\color{\chon} to} obtain,\cite{mottaCMS10}
+{\color{\chon} to} obtain,
 \begin{align}\label{vnl.5}
-\calbv^{\ell,\nl}_{nm}(\bfk)&=\frac{1}{2}
+\calbv^{\nl}_{nm}(\bfk)&=\frac{1}{2}
 \sum_{\bfG}
 \Big(
-\bra{n\bfk} C^\ell(z) 
+\bra{n\bfk} C(z) 
 \ket{\bfk+\bfG}\bra{\bfk+\bfG}
@@
 \ket{\bfk+\bfG}\bra{\bfk+\bfG}
-C^\ell(z) \ket{m\bfk}
+C(z) \ket{m\bfk}
 \Big)
@@
 \calf^{\ell*}_{n\bfk}(\bfG) 
-\calh_{m\bfk}(\bfG) 
+H_{m\bfk}(\bfG) 
 +
-\calh^*_{n\bfk}(\bfG) 
-\calf^\ell_{m\bfk}(\bfG) 
+H^*_{n\bfk}(\bfG) 
+\calf_{m\bfk}(\bfG) 
 \right) 
@@
 \begin{align}\label{vnl.10}
-\calf^\ell_{n\bfk}(\bfG) 
+\calf_{n\bfk}(\bfG) 
 =
@@
 A_{n\bfk}(\bfG') 
-\gd_{\bfG_\parallel \bfG'_\parallel}f^\ell(\bfG'_\perp-\bfG_\perp) 
+\gd_{\bfG_\parallel \bfG'_\parallel}f(G'_\perp-G_\perp) 
 ,
@@
 \begin{align}\label{vnl.11}
-\calh_{n\bfk}(\bfG)&=
+H_{n\bfk}(\bfG)&=
 \sum_{\bfG'} 
@@
 implemented 
-the calculation of $\calbv^{\ell,\nl}_{nm}(\bfk)$ with the help {\color{\chon} of} 
+the calculation of $\calbv^{\nl}_{nm}(\bfk)$ with the help {\color{\chon} of} 
 the \depe~code.\cite{francesco}
 {\color{\chon} We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
-expressions by t}aking $C^\ell(z)=1$, which implies 
-{\color{\chon} that} $f^\ell(g)=\gd_{g0}$.
-
-%%%%
-%\subsection{Matrix elements of $[\bfr,\bfv^\nl]$}\label{calt}
-\label{calt}
-Using Eq.~\eqref{vnl} we define
+expressions by t}aking $C(z)=1$, which implies 
+{\color{\chon} that} $f(G_\perp)=\gd_{G_\perp 0}$, 
+and 
+$\calf_{n\bfk}(\bfG) \to F_{n\bfk}(\bfG)=A_{n\bfk}(\bfG)$.
+
+From Eqs.~\eqref{lab} and \eqref{vnl} 
 \begin{align}\label{3.1}
-\call^{\rma\rmb}_{nm}(\bfk) 
-\equiv\frac{1}{i\hbar}
-\bra{n\bfk}
-[ r^{\rma}, v^{\nl,\rmb}]
-\ket{m\bfk}
+L^{\rma\rmb}_{nm}(\bfk) 
 =
@@
 \end{align} 
-{\color{\chon} and we expand the triple commutator,}
+{\color{\chon} and we expand the triple commutator to obtain,}
 \begin{align}\label{3.5}
-\call^{\rma\rmb}_{nm}(\bfk) 
+L^{\rma\rmb}_{nm}(\bfk) 
 &=
@@
 \begin{align}\label{3.7}
-\call^{\rma\rmb}_{nm}(\bfk)
+L^{\rma\rmb}_{nm}(\bfk)
 &=
-\frac{1}{\hbar^2\gO}
+\frac{1}{\hbar^2}
 \sum_{\bfG,\bfG'} 
@@
 =
+\frac{1}{\gO}
 \int 
@@
 can be worked out explicitly {\color{\chon} and} 
-${\cal L}^{\rma\rmb}_{nm}(\bfk)$
+$L^{\rma\rmb}_{nm}(\bfk)$
 {\color{\chon} can} be finally calculated.\cite{valerie}
@@
 %%%%
-\subsection{Expressions for  \texorpdfstring{$\bfv^{\ell,S}_{nm}$}{Vnm}
+\subsection{Expressions for  \texorpdfstring{$\calbv^{S}_{nm}$}{Vnm}
 and
-\texorpdfstring{$\Big({\calbv}^{\ell,S}_{nm}\Big)_{;\bfk}$}{(Vnm);kb}
+\texorpdfstring{$\Big({\calbv}^{S}_{nm}\Big)_{;\bfk}$}{(Vnm);kb}
 }\label{calvs}
@@
 \begin{align}\label{a.3b}
-\calbv^{\ell,S}_{nm}
+\calbv^{S}_{nm}
 &=
 \frac{1}{2}\sum_q\left(   
-\bfv^S_{nq}\calc^\ell_{qm}+\calc^\ell_{nq}\bfv^S_{qm}
+\bfv^S_{nq}\calc_{qm}+\calc_{nq}\bfv^S_{qm}
 \right)  
@@
 \begin{align}\label{a.3bn}
-\left(\calbv^{\ell,S}_{nm}\right)_{;\bfk}
+\left(\calbv^{S}_{nm}\right)_{;\bfk}
 &=
 \frac{1}{2}\sum_q\left(
-(\bfv^S_{nq})_{;\bfk}\calc^\ell_{qm}
+(\bfv^S_{nq})_{;\bfk}\calc_{qm}
 +    
-\bfv^S_{nq}(\calc^\ell_{qm})_{;\bfk}
+\bfv^S_{nq}(\calc_{qm})_{;\bfk}
 +
-(\calc^\ell_{nq})_{;\bfk} \bfv^S_{qm}
+(\calc_{nq})_{;\bfk} \bfv^S_{qm}
 +
-\calc^\ell_{nq} (\bfv^S_{qm})_{;\bfk}
+\calc_{nq} (\bfv^S_{qm})_{;\bfk}
 \right)  
-,
-\end{align}    
-{\color{\chon} and considering} 
+.
+\end{align}    
+Again, from
 Eq.~\eqref{chon.2}, 
@@
 \end{align}
-{\color{\chon} is in total} 
-agreement with Eq. A(6) of Ref.~\onlinecite{cabellosPRB09}.
+a result that is in agreement with Eq. A(6) of Ref.~\onlinecite{cabellosPRB09}.
 
 %%%%%%%%%%%%%%
+\newpage
 \bibliography{ref}% 
@@
 
-Second harmonic generation (SHG) is a powerful spectroscopic tool for 
-studying the optical properties of surfaces and interfaces since it has 
-the advantage of being surface sensitive. Within the dipole approximation, 
-inversion symmetry forbids SHG from the bulk of controsymmetric materials. 
-SHG is allowed at the surface of these materials where the inversion symmetry 
-is broken and should necessarily come from the localized surface region. 
-SHG allows the study of the structural atomic arrangement and phase 
-transitions of clean and adsorbate covered surfaces. Since it is also an 
-optical probe it can be used out of UHV conditions and is non-invasive 
-and non-destructive. Experimentally, new tunable high intensity laser systems 
-have made SHG spectroscopy readily accessible and applicable to a wide range 
-of systems.\cite{downer_optical_2001,lupke_characterization_1999}
-
-However, theoretical development of the field is still an ongoing 
-subject of research. Some recent advances for the cases of semiconducting 
-and metallic systems have appeared in the literature, where the use of 
-theoretical models with experimental results have yielded correct 
-physical interpretations for observed SHG spectra.
-\cite{
-downer_optical_2001,
-mendozaPRB01,
-lim_optical_2000,
-gavrilenko_optical_2000,
-mendozaPRB99,
-mendozaPRL98,
-mendozaPRB96a,
-mendozaPRB97,
-guyotPRB90}
-
-In a previous article\cite{mendoza_epioptics_2001} we reviewed some 
-of the recent results in the study of SHG using the velocity gauge 
-for the coupling between the electromagnetic field and the electron. 
-In particular, we demonstrated a method to systematically analyze the 
-different contributions to the observed SHG peaks.\cite{arzatePRB01} 
-This approach consists of separating the different contributions to 
-the nonlinear susceptibility according to 1$\omega$ and 2$\omega$ 
-transitions, and the surface or bulk nature of the states among 
-which the transitions take place.
-
-To compliment those results, in this article we review the calculation 
-of the nonlinear susceptibility using the longitudinal gauge. We show 
-that it is posible to clearly obtain the ``layer-by-layer'' contribution 
-for a slab scheme used for surface calculations.
+The first attempt to go beyond the independent particle approximation 
+could be found in 
+Leitsmann \textit{et al}.\cite{leitsmannPRB05} 
+that included excitonic and local field interactions. 
+More recently, Luppi \textit{et al.}\cite{luppiPRB10,luppiJCP10} and 
+H\"ubner \textit{at al.}\cite{hubenerPRB11}, studied 
+the influence of crystal local fields and excitonic effects on 
+the spectrum of the second harmonic generation of several bulk 
+systems 
+by using time-dependent density-functional theory 
+including many-body effects, namely, quasiparticle corrections through 
+the scissors operator, crystal local field effects, and excitonic 
+interaction.  
+Also,
+Trolle \textit{et al}.\cite{trollePRB14} 
+elaborate on a theoretical model for SHG of 
+mono-layer MoS$_{2}$. Their approach is based on the tight-binding model 
+fitted to GW results and include excitonic effects from the 
+Bethe-Salpeter exciton picture. However, neither of these models 
+has a consideration for surfaces.

### 2014-11-19 paper-v-0.tex
added new references and updated bib file

@@
 and interfaces: atomic structure, phase transitions, and adsorption of 
-atoms,1–11{\color{\chonn} [incluye estas de mejiaPRB02 y combinalas con
-  las tuyas y tambien pon la de salazarPRB14 y busca referencia para
-  ??, checando q sean relevantes con SSHG]} 
-among others. \cite{hughesPRB96, 
+atoms, among others.\cite{
+    daumPRL93, 
+    mcgilpOE94, 
+    meyerPRL95, 
+    powerPRL95, 
+    godefroyAPL96, 
+    hoferAPA96, 
+    dadapPRB97, 
+    bloembergenAPB99, 
+    mcgilpSRL99, 
+    suzukiAPB99, 
+    mitchellSS01,
+    hughesPRB96, 
     guyot-sionnestPRB88, 
-     lupkeSSR99, 
     downerPSSA01, 
@@
     chenPRL81, 
-    mendozaPRL98,downerSIA01,cazzanelliNM14}
+    mendozaPRL98,
+    downerSIA01,
+    cazzanelliNM14}
 Nowadays, SSHG spectroscopy is a crucial tool for research and development
-in microelectronics,\cite{??} semiconductors,\cite{??} nanomaterials,\cite{??} and many more recent 
-areas of great scientific and commercial interest.\cite{??}
+in microelectronics,\cite{zheltikovLP00} semiconductors,
+\cite{lupkeSSR99}, nanomaterials,\cite{salazar-aparicioPRB14} 
+and many more recent 
+areas of great scientific and commercial interest.\cite{cazzanelliNM14}
 The high surface sensitivity of SSHG spectroscopy is due to the
@@
 Density Functional Theory (DFT) within the Local Density Approximation
-(LDA),\cite{??} the most 
+(LDA),\cite{kohnPR65} the most 
 used framework for \textit{ab initio} calculations,
@@
 correct for the underestimation of the gap. In this context the so
-called GW approximation\cite{onidaRMP01} is known to correct the band
-gap of most semiconductors\cite{cual}. However this could be a numerically very
+called GW approximation\cite{onidaRMP02} is known to correct the band
+gap of most semiconductors\cite{luceroJPCM12}. However this could be a numerically very
 expensive calculation, and thus one uses the much simpler scissors
-operator scheme.\cite{?} This allows to ``open'' the DFT-LDA gap to
+operator scheme.%%\cite{?} find this
+This allows to ``open'' the DFT-LDA gap to
 its correct experimental or GW value, when experiments are not available. 
@@
 The real part of each contribution can be obtained through
-a Kramers-Kronig transformation\cite{tacongePRB14} {\color{\chon} and}
+a Kramers-Kronig transformation\cite{nicolasPRB14} {\color{\chon} and}
 $\chi^{\rma\rmb\rmc}=
@@
 
-\subsection{Full-slab results}\label{results}
+\subsection{Full-slab results}\label{fsresults}
 

### 2014-11-20 paper-v-0.tex
fixing bib file and cleaning up paper

@@
 %\def\chon[1]{\color{red} #1}
-\def\chon{black}
+\def\chon{red}
 \def\chonn{red}
@@
 We present a theoretical review of surface second harmonic generation (SHG) 
-from semiconductor surfaces based on the {\color{\chon} length} gauge. This 
+from semiconductor surfaces based on the length gauge. This 
 layer-by-layer analysis is carefully presented in order to show how a 
@@
     hughesPRB96, 
-    guyot-sionnestPRB88, 
+    guyotPRB88, 
     downerPSSA01, 
@@
 expensive calculation, and thus one uses the much simpler scissors
-operator scheme.%%\cite{?} find this
+operator scheme.\cite{levinePRL89, levinePRL91}
 This allows to ``open'' the DFT-LDA gap to
@@
 We use the IPA and neglect local field  
-and excitonic effects{\color{\chon}.  
-We assume a classical} electromagnetic field
-{\color{\chon} and quantum mechanical matter.}
+and excitonic effects.  
+We assume a classical electromagnetic field
+and quantum mechanical matter.
 We can describe the system using 
@@
 $\tilde{\mathcal{O}}=U{\cal O}U^{\dagger }$. Even if $\mathcal{O}$ 
-{\color{\chon} is time-independent}, 
-$\tilde{\mathcal{O}}$ {\color{\chon} is time-dependent} through the explicit
+is time-independent, 
+$\tilde{\mathcal{O}}$ is time-dependent through the explicit
 time dependence of $U$. 
@@
 $\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle=f(\hbar \omega_{n}(\mathbf{k}))\equiv f_{n}$. For a
-clean, cold semiconductor $f_{n}=1$ {\color{\chon} when} $n$ {\color{\chon} is} a valence ($v$) or
-occupied band, and zero {\color{\chon} when} $n$ {\color{\chon} is} a conduction ($c$) or empty band.
-{\color{\chon} We assume this for the remainder of the article.}
-{\color{\chon} T}he expectation values must satisfy
+clean, cold semiconductor $f_{n}=1$ when $n$ is a valence ($v$) or
+occupied band, and zero when $n$ is a conduction ($c$) or empty band.
+We assume this for the remainder of the article.
+The expectation values must satisfy
 $
@@
 where $\dot{\mathbf{r}}$ is the time derivative of the position operator of the
-electron {\color{\chon} with} charge $e$, {\color{\chon} defined as}
+electron with charge $e$, defined as
 \begin{equation}
@@
 where $\tilde\omega=\omega+i\eta $,
-and $\eta >0$ {\color{\chon} adiabatically switches on} the interaction,
+and $\eta >0$ adiabatically switches on the interaction,
 we write the second order nonlinear
@@
 where $\chi^{abc}(-2\omega ;\omega ,\omega )$ is the nonlinear
-susceptibility responsible {\color{\chon} for} surface second harmonic generation
+susceptibility responsible for surface second harmonic generation
 (SSHG). 
@@
 is used to solve the Kohn-Sham equations\cite{kohnPR65} of Density  
-Functional Theory (DFT) {\color{\chon}. It is convenient to work within the Local 
+Functional Theory (DFT). It is convenient to work within the Local 
 Density Approximation (LDA), so we label the hamiltonian with the corresponding  
 LDA superscript. Any other approximation can be used (like GGA) and our 
-derivation remains the same.} Then,
+derivation remains the same. Then,
 \begin{equation}\label{h0}
@@
 $\chi^{abc}(-2\omega;\omega,\omega)$, we asume
-that the crystal potential could have a {\color{\chon} non-local spatial}
-contribution {\color{\chon} as} is customary for most
+that the crystal potential could have a non-local spatial
+contribution as is customary for most
 pseudopotentials, and then we replace $V$ with
@@
 from the non-local part of the pseudopotential, $V^{\mathrm{nl}}$, in
-the results that we {\color{\chon} have derived.}
-
-{\color{\chon} It} is well known that the use of the {\color{\chon} LDA} 
+the results that we have derived.
+
+It is well known that the use of the LDA
 leads to an underestimation of the band gap. A standard
 procedure to correct for this is to
-use {\color{\chon} the} ``scissors approximation'', {\color{\chon} where the 
-conduction bands are rigidly shifted in energy} so that the band gap 
-corresponds to the accepted experimental band gap{\color{\chon} .
-This} is often in fairly good agreement with the GW
+use the ``scissors approximation'', where the 
+conduction bands are rigidly shifted in energy so that the band gap 
+corresponds to the accepted experimental band gap.
+This is often in fairly good agreement with the GW
 band gap based on a more sophisticated calculation.\cite{hybertsenPRB86}
-{\color{\chon} T}he LDA wave functions are used since they produce band
+The LDA wave functions are used since they produce band
 structures with dispersion relations similar to those predicted by the GW
-approximation. Mathematically, {\color{\chon} the scissors (non-local) operator 
-$S$ is added} to the unperturbed or unscissored Hamiltonian {\color{\chon} $H^{\mathrm{LDA}}_{0}$ ,}
+approximation. Mathematically, the scissors (non-local) operator 
+$S$ is added to the unperturbed or unscissored Hamiltonian $H^{\mathrm{LDA}}_{0}$ ,
 \begin{equation*}
@@
 \end{equation}
-We emphasize that the {\color{\chon} scissored and unscissored Hamiltonian 
-have the same eigenfunctions,} where
+We emphasize that the scissored and unscissored Hamiltonian 
+have the same eigenfunctions, where
 $\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
@@
 To calculate the optical response 
-within the Length Gauge {\color{\chon} formalism}, we take the 
-interaction Hamiltonian {\color{\chon} as} 
+within the Length Gauge formalism, we take the 
+interaction Hamiltonian as
 \begin{equation}
@@
 treatment of the position operator
-$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$ {\color{\chon} has} 
-been described {\color{\chon} in detail}.\cite{adamsJCP53,blountSSP62} 
-{\color{\chon} In particular, Ref.~\onlinecite{aversaPRB95} develops} 
+$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$ has 
+been described in detail.\cite{adamsJCP53,blountSSP62} 
+In particular, Ref.~\onlinecite{aversaPRB95} develops
 the correct treatment of $\bfr$ for
-the calculation of the {\color{\chon} nonlinear} optical {\color{\chon} response.} 
+the calculation of the nonlinear optical response.
 $\bfr$ is split into the {\it intraband} 
@@
 \end{align}
-such that $\bfr_{e,nm}=0$ for $n=m${\color{\chon} .
-From Eq.~\eqref{mv}} with $H_0\to
+such that $\bfr_{e,nm}=0$ for $n=m$.
+From Eq.~\eqref{mv} with $H_0\to
 H^\sigma_0$, we obtain
@@
 %elements.
-For the intraband part, {\color{\chon} $\bfr_i$ only appears in
+For the intraband part, $\bfr_i$ only appears in
 commutators during the derivation of
-the optical response. We use},\cite{aversaPRB95}
+the optical response. We use,\cite{aversaPRB95}
 \begin{equation}\label{conmri3n}
@@
 The vectors $\bfgxi_{nn}(\bfk)$ are defined in 
-Ref.~\onlinecite{aversaPRB95} {\color{\chon} though} they do not need to be 
+Ref.~\onlinecite{aversaPRB95} though they do not need to be 
 calculated explicitly in what follows.
 
-Before {\color{\chon} continuing,} 
+Before continuing,
 we derive a key result for the length gauge formulation. 
-{\color{\chon} Again, using} $H^\sigma_0$ in
+Again, using $H^\sigma_0$ in
 Eq.~\eqref{mv} we obtain
@@
 \end{subequations}
-with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator{\color{\chon}, 
-and using $[r^a,p^b]=i\hbar\gd_{ab}$, where $\gd_{ab}$ is the Kronecker} delta.
+with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator, 
+and using $[r^a,p^b]=i\hbar\gd_{ab}$, where $\gd_{ab}$ is the Kronecker delta.
 Using Eq.~\eqref{hats} we obtain 
@@
 with $f_{nm}\equiv f_n-f_m$,
-where we see that $\bfv^S_{nn}=0$. {\color{\chon} It} follows that
+where we see that $\bfv^S_{nn}=0$. It follows that
 \begin{align}\label{chon.9}
@@
 \end{align}
-{\color{\chon} and from Eq.~\eqref{pmnrmn},} 
+and from Eq.~\eqref{pmnrmn},
 \begin{align}\label{chon.10}
@@
 The matrix elements 
-of $\bfr_{nm}(\bfk)$ are {\color{\chon} identical using either} 
-the LDA or {\color{\chon} scissored} 
-Hamiltonian, {\color{\chon} thus negating the} need to label them.
-Of course, {\color{\chon} it} is more convenient to calculate them
-through $\bfv^\lda_{nm}(\bfk)$ {\color{\chon} which 
-includes} only the contribution of 
+of $\bfr_{nm}(\bfk)$ are identical using either
+the LDA or scissored
+Hamiltonian, thus negating the need to label them.
+Of course, it is more convenient to calculate them
+through $\bfv^\lda_{nm}(\bfk)$ which 
+includes only the contribution of 
 the non-local part of the pseudopotentials through
-$\bfv^\nl_{nm}(\bfk)$. These {\color{\chon} can} be readily
+$\bfv^\nl_{nm}(\bfk)$. These can be readily
 calculated
@@
 form,\cite{francesco,mottaCMS10,kleinmanPRL82,adolphPRB96}
-{\color{\chon} as shown in Appendix \ref{vesnl}.}
-{\color{\chon} The} advantage of the length gauge formalism for 
-{\color{\chon} calculating linear and nonlinear} optical responses 
-resides in the ease with {\color{\chon} which the scissors operator
-can be introduced into the calculation, also described in detail in} 
-Nastos et al.\cite{nastosPRB05} {\color{\chon} The 
-length gauge} formalism for the scissored Hamiltonian can be easily worked 
+as shown in Appendix \ref{vesnl}.
+The advantage of the length gauge formalism for 
+calculating linear and nonlinear optical responses 
+resides in the ease with which the scissors operator
+can be introduced into the calculation, also described in detail in
+Nastos et al.\cite{nastosPRB05} The 
+length gauge formalism for the scissored Hamiltonian can be easily worked 
 out by simply using the unscissored LDA Hamiltonian,
@@
 for the unperturbed system 
-with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction{\color{\chon}. 
+with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction. 
 We stress that within the length gauge,
 we need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
-$\omega_{n}^{\sigma}$ at the end of the derivation}
+$\omega_{n}^{\sigma}$ at the end of the derivation
 to obtain the scissored results for any 
@@
 
-Following Ref.~\onlinecite{sipePRB00} {\color{\chon} we take} the
+Following Ref.~\onlinecite{sipePRB00} we take the
 matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of 
@@
 %$(B^a_{nm})_{;k^b}$ can be obtained.
-{\color{\chon} Note that the indices in Eq.~\eqref{j.2} are all} different from each
-other. {\color{\chon} This is due to the $f_{nm}$ factor in Eq.~\eqref{j.1}, 
-and therefore $B^a_{nn}=0$}. The dependence {\color{\chon} on} $\bfk$ 
+Note that the indices in Eq.~\eqref{j.2} are all different from each
+other. This is due to the $f_{nm}$ factor in Eq.~\eqref{j.1}, 
+and therefore $B^a_{nn}=0$. The dependence on $\bfk$ 
 of all quantities is implicitly understood from 
-{\color{\chon} this point forward.}
+this point forward.
 
@@
 semi-infinite system, we replace it by a super cell that 
-{\color{\chon} consists} of a finite
+consists of a finite
 slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
-super cell {\color{\chon} constitutes the unit cell, and}
+super cell constitutes the unit cell, and
 is repeated to form a full three dimensional crystalline structure,
 i.e. a repeated super cell scheme. 
-The slab itself consists of {\color{\chon} front, back, and 
-sub-surface} regions, and in between these
-a region that is equivalent to {\color{\chon} the}
+The slab itself consists of front, back, and 
+sub-surface regions, and in between these
+a region that is equivalent to the
 bulk of the system. 
@@
 balanced forces are disrupted when the surface atoms do not find their 
-partner atoms that are now absent at the {\color{\chon} slab surface.} 
+partner atoms that are now absent at the slab surface.
 To take the reconstruction or relaxation into account, 
 we take ``surface'' to mean
-the true surface of the first layer of {\color{\chon} atoms} and
+the true surface of the first layer of atoms and
 some of the atomic sub-layers adjacent to it.
@@
 centrosymmetric. This would imply that 
-$\chi^{\mathrm{slab},\rma\rmb\rmc}=0$ {\color{\chon} so} we must
+$\chi^{\mathrm{slab},\rma\rmb\rmc}=0$ so we must
 find a scheme 
@@
 are different, breaking the centrosymmetry and therefore giving an
-overall $\chi^{\mathrm{slab},\rma\rmb\rmc}\ne 0${\color{\chon} ;} we still
+overall $\chi^{\mathrm{slab},\rma\rmb\rmc}\ne 0$; we still
 need a procedure to extract the front surface $\chi^{f,\rma\rmb\rmc}$
@@
 We have omitted the frequency arguments of $\chi^{\rma\rmb\rmc}$ for 
-{\color{\chon} convenience of} notation.
+convenience of notation.
 
@@
 $(f)$ or back $(b)$ from its $z_\ell$ position. 
-We take $z_\ell$ {\color{\chon} to be} at the center of an atom that 
-belongs to layer $\ell${\color{\chon} , so the previous} 
+We take $z_\ell$ to be at the center of an atom that 
+belongs to layer $\ell$, so the previous
 equation would give the $\ell$-th atomic-layer 
-contribution to the {\color{\chon} nonlinear} optical response.
+contribution to the nonlinear optical response.
 $\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
@@
 \includegraphics[scale=.7]{images/slab}
-\caption{(color on line) A sketch of the super {\color{\chon} cell. 
-The} slab corresponds to the
+\caption{(color on line) A sketch of the super cell. 
+The slab corresponds to the
 circles representing the atoms of the system.\label{fslab}} 
@@
 $\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gs\ket{\bfr}\bra{\bfr}
-+\ket{\bfr}\bra{\bfr}\bfv^\gs\right)${\color{\chon} , that} leads to
++\ket{\bfr}\bra{\bfr}\bfv^\gs\right)$, that leads to
 \begin{align}\label{jmic}
@@
 If we want the contribution from only one region of the unit cell 
-{\color{\chon} towards} the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
-over the desired region. {\color{\chon} Then the} contribution 
+towards the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
+over the desired region. Then the contribution 
 to the current density from the
@@
 is all that is needed to change any of the
-{\color{\chon} electron} velocity operators $\bfV$ to the new velocity
+electron velocity operators $\bfV$ to the new velocity
 operator $\calbv$ that implicitly takes into account the
@@
 $\chi^{\rma\rmb\rmc}(-2\go;\go,\go)$
-are m$^2$/V, as they {\color{\chon} should be} for a surface SH susceptibility.
+are m$^2$/V, as they should be for a surface SH susceptibility.
 Using Eq.~\eqref{j.1} we
@@
 two contributions from the first and second terms on the right hand side
-{\color{\chon} of Eq.~\eqref{Pjikn2}:}
+of Eq.~\eqref{Pjikn2}:
 %\end{document}  
@@
 \end{equation} 
-{\color{\chon} related} to intraband transitions, and 
+related to intraband transitions, and 
 \begin{equation}\label{chie}
@@
 \end{equation} 
-{\color{\chon} related} to interband transitions.
+related to interband transitions.
 The generalized derivative in Eq.~\eqref{chii} is dealt with by the chain rule 
@@
 \end{equation}
-where {\color{\chon} substituting} $H^\sigma_0$ 
+where substituting $H^\sigma_0$ 
 into Eq.~\eqref{conmri3n} and then
 Eq.~\eqref{chon.9}
-{\color{\chon} we obtain}
+we obtain
 \begin{align}\label{eli.13}
@@
 \end{subequations}
-where we have {\color{\chon} split} the interband and intraband $1\go$ and $2\go$
-contributions and {\color{\chon} supressed} the $\go$ arguments for 
-{\color{\chon} convenience of notation}.
+where we have split the interband and intraband $1\go$ and $2\go$
+contributions and supressed the $\go$ arguments for 
+convenience of notation.
 The real part of each contribution can be obtained through
-a Kramers-Kronig transformation\cite{nicolasPRB14} {\color{\chon} and}
+a Kramers-Kronig transformation\cite{nicolasPRB14} and
 $\chi^{\rma\rmb\rmc}=
@@
 $\{\ell\}$
-is meant to be {\color{\chon} a chosen set of layers.} For instance, 
-one {\color{\chon} can take a single layer 
-encompassing} half of the slab, or {\color{\chon} take each 
-atomic layer individually to the middle} 
-of the slab. {\color{\chon} For} the first case there is 
-{\color{\chon} a} single summand
-in Eq.~\eqref{chiijksur}{\color{\chon} . For the second case
-there is} a sum from $\ell=1${\color{\chon} , denoting the first layer 
+is meant to be a chosen set of layers. For instance, 
+one can take a single layer 
+encompassing half of the slab, or take each 
+atomic layer individually to the middle
+of the slab. For the first case there is 
+a single summand
+in Eq.~\eqref{chiijksur}. For the second case
+there is a sum from $\ell=1$, denoting the first layer 
 right at the surface, to $\ell=N$, denoting the layer at the middle of the slab 
-that behaves like a bulk layer.}
+that behaves like a bulk layer.
 %that at a distance $d$ from the surface  as seen in Fig.~\ref{fslab}.  
@@
 %$\bfgchi^{\ell=N}=0$;
-We {\color{\chon} remark} that the value of 
-$N$ is not universal {\color{\chon} and} 
+We remark that the value of 
+$N$ is not universal and
 the slab needs to have enough atomic layers 
@@
 %to be satisfied and to 
-{\color{\chon} in} order to give converged results for 
+in order to give converged results for 
 $\bfgchi (-2\go;\go,\go)$. 
@@
 In this section we present a 
-{\color{\chon} relevant test case} to check the
-consistency of {\color{\chon} our approach. We have selected}
+relevant test case to check the
+consistency of our approach. We have selected
 a clean Si(001) surface with a $2\times 1$ surface reconstruction.
 The slab for such a surface could be chosen to be centrosymmetric 
-by {\color{\chon} creating} the front {\color{\chon} and back} 
+by creating the front and back
 surfaces with the same $2\times 1$
-reconstruction{\color{\chon} . However, we choose to} terminate one 
-of the surfaces with hydrogen {\color{\chon} producing an ideal  
-terminated} bulk Si surface. {\color{\chon} The} H atoms simply saturate the 
+reconstruction. However, we choose to terminate one 
+of the surfaces with hydrogen producing an ideal  
+terminated bulk Si surface. The H atoms simply saturate the 
 dangling bonds of the bulk-like Si atoms at the surface, as seen in 
@@
 \caption{(color on line) The slab shows a 
-    {\color{\chon} clean Si(001)$2\times 1$ front surface 
-    with an ideal terminated Si bulk back surface. The}
+    clean Si(001)$2\times 1$ front surface 
+    with an ideal terminated Si bulk back surface. The
     dangling bonds are H (small balls) saturated. 
-    This {\color{\chon} image depicts} 12 Si atomic layers
-    {\color{\chon} with one} H atomic layer. 
+    This image depicts 12 Si atomic layers
+    with one H atomic layer. 
 \label{si2x1}} 
 \end{figure} 
-The idea {\color{\chon} behind this slab configuration} is that the 
+The idea behind this slab configuration is that the 
 cristalline symmetry of the H terminated surface imposes that 
-$\chi_{\mathrm{H}}^{xxx}=0${\color{\chon}.
+$\chi_{\mathrm{H}}^{xxx}=0$.
  The $2\times 1$ surface has no 
-such restrictions, so} $\chi_{2\times 1}^{xxx}\ne 0$.
+such restrictions, so $\chi_{2\times 1}^{xxx}\ne 0$.
 This is due to the fact that along the $y$ direction there is a mirror plane for the
@@
 Thus, calculating $\chi^{xxx}$ for the full-slab, or the 
-{\color{\chon} half-slab containing} the $2\times 1$ surface\cite{nota1}
-{\color{\chon} should yield the} same result since the contribution from the H
-saturated surface is zero {\color{\chon} regardless. 
+half-slab containing the $2\times 1$ surface\cite{nota1}
+should yield the same result since the contribution from the H
+saturated surface is zero regardless. 
 We must check that the following 
-relationship is satisfied for this particular slab}
+relationship is satisfied for this particular slab
 \begin{align}\label{hs}
@@
 $\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
-$C(z)=1$ {\color{\chon} from the upper half containing the }$2\times 1$ 
+$C(z)=1$ from the upper half containing the $2\times 1$ 
 surface reconstruction, as seen in Fig.~\ref{si2x1},
@@
 $C(z)=1$ through the full slab.
-{\color{\chon} We show the results for this comparison in the remainder 
-of this section}
+We show the results for this comparison in the remainder 
+of this section
 
 The self-consistent ground state and their Kohn-Sham states were
-calculated in the {\color{\chon} DFT-LDA framework using} the plane-wave 
+calculated in the DFT-LDA framework using the plane-wave 
 ABINIT code.\cite{abinit}
@@
 fully separable nonlocal pseudopotentials in the Kleinman-Bylander 
-{\color{\chon} form.\cite{kleinmanPRL82}
+form.\cite{kleinmanPRL82}
 The contribution of $\bfv^\nl$ to Eq.~\eqref{chis} can be carried out 
 as detailed in Appendix \ref{vesnl} using the DP code for 
-the actual calculation}.\cite{francesco}
+the actual calculation.\cite{francesco}
 %
@@
 %
-The unit cell has two atoms {\color{\chon} making for a computationally
-expensive calculation.} 
-The vacuum size is equivalent to one {\color{\chon} quarter the} 
-size of the slab, {\color{\chon} avoiding the effects produced by possible 
-wave-function tunneling} from the contiguous surfaces of the full
+The unit cell has two atoms making for a computationally
+expensive calculation.
+The vacuum size is equivalent to one quarter the
+size of the slab, avoiding the effects produced by possible 
+wave-function tunneling from the contiguous surfaces of the full
 crystal formed by the repeated super cell scheme.\cite{mendozaPRB06}
 
-{\color{\chon} Spin-orbit, local field, and electron-hole 
+Spin-orbit, local field, and electron-hole 
 attraction\cite{leitsmannPRB05,trollePRB14} 
-effects on the SHG process are all neglected.}
+effects on the SHG process are all neglected.
 Although these are important factors in the optical response of a semiconductor,
-{\color{\chon} their efficient calculation is still theoretically and  
+their efficient calculation is still theoretically and  
 numerically challenging and  
-under debate. This merits further study but is beyond the scope of this paper.} 
-For a given {\color{\chon} slab size, we find the} converged spectra 
-{\color{\chon} to obtain the relevant parameters. The most important of 
+under debate. This merits further study but is beyond the scope of this paper.
+For a given slab size, we find the converged spectra 
+to obtain the relevant parameters. The most important of 
 these are: an energy cut-off of 10 Ha, an equal number of conduction and 
-valence bands, and a set of 244 $\bfk$-points.} 
-{\color{\chon} The $\bfk$-points are used} for the linear analytic 
-tetrahedron method {\color{\chon} for evaluating the}
+valence bands, and a set of 244 $\bfk$-points.
+The $\bfk$-points are used for the linear analytic 
+tetrahedron method for evaluating the
 Brillouin zone integrals in Eq.~\eqref{chis}, 
 where special care was taken to examine the double resonances.\cite{nastosPRB05}
-{\color{\chon}
 Double resonances occur at a given frequency $\omega$ if there can be 
@@
 the double resonances affect the low frequency results by less than 2\%. 
-All spectra were calculated with a Gaussian smearing of 0.15 eV.}
+All spectra were calculated with a Gaussian smearing of 0.15 eV.
 \begin{figure}
@@
 $|\chi_{\mathrm{half-slab}}^{xxx}|$ vs $\hbar\omega$ 
-{\color{\chon} for the slab} 
+for the slab
 with 16, 24, 32, and 40 atomic Si layers. The front surface 
-{\color{\chon} is in a clean $2\times 1$ reconstruction and the back 
-surface is an ideal terminated bulk H-saturated dangling bonds.}
+is in a clean $2\times 1$ reconstruction and the back 
+surface is an ideal terminated bulk H-saturated dangling bonds.
 \label{fig1}} 
@@
 $\chi^{xxx}_{\mathrm{full-slab}}$
-vs $\hbar\omega$ for a slab {\color{\chon} with} 32 
-atomic Si layers plus one {\color{\chon} H layer}. 
+vs $\hbar\omega$ for a slab with 32 
+atomic Si layers plus one H layer. 
 \label{fig2}} 
@@
 
-{\color{\chon} 
 We must evaluate 
@@
 $[\bfr,\bfv^\nl]$. However, the contribution from $[\bfr,\bfv^\nl]$ 
-is very small\cite{valerie} and therefore we neglect it in this work.}
+is very small\cite{valerie} and therefore we neglect it in this work.
 
@@
 In Fig.~\ref{fig1} we show $|\chi_{\mathrm{full-slab}}^{xxx}|$
-for {\color{\chon} the} slab {\color{\chon} with} 
+for the slab with
 16, 24, 32, and 40 Si atomic layers. 
-{\color{\chon} We note that the response differs substantially 
-for 16 and 24 layers but is quite similar for 32 and 40 layers.}
+We note that the response differs substantially 
+for 16 and 24 layers but is quite similar for 32 and 40 layers.
 As explained above,
-the calculation of {\color{\chon} the $\bfv^\nl$ contribution is 
-computationally expensive.} 
+the calculation of the $\bfv^\nl$ contribution is 
+computationally expensive.
 A good compromise between the accuracy in the convergence of
@@
 of layers in the slab, and the computational 
-{\color{\chon} expense is to consider} 
+expense is to consider
 the slab with 32 Si atomic layers as 
-{\color{\chon} an accurate representation} of our 
-{\color{\chon} system.}
+an accurate representation of our 
+system.
 
@@
 $\chi^{xxx}_{\mathrm{full-slab}}$ 
-{\color{\chon} for} the four different possibilities 
-{\color{\chon} between} including or not including the
-effects of $\bfv^\nl$ {\color{\chon} or the} scissors correction
+for the four different possibilities 
+between including or not including the
+effects of $\bfv^\nl$ or the scissors correction
 $\hbar\sigma$.   
@@
 We see that for all four instances the 
-{\color{\chon} difference between responses is quite small}. 
-{\color{\chon} Indeed, when the value $|\chi^{xxx}|$ 
+difference between responses is quite small.
+Indeed, when the value $|\chi^{xxx}|$ 
 is large the difference between the two is very small; 
@@
 These differences would decrease as the number of atomic layers 
-increases. We remark that 32 layers in the slab} is more than enough 
+increases. We remark that 32 layers in the slab is more than enough 
 to confirm that the extraction of the surface second harmonic 
@@
 
-\subsection{Results for $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$}
+\subsection{\texorpdfstring{Results for $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$}{Results for Xxxx(2x1)(-2w;w,w)}}
 
@@
 by calculating $|\chi^{xxx}_{\mathrm{half-slab}}|$.
-{\color{\chon} First,} from Fig.~\ref{fig2} we note a series of resonances 
-that {\color{\chon} derive from} 1$\omega$ and 2$\omega$ terms in
+First, from Fig.~\ref{fig2} we note a series of resonances 
+that derive from 1$\omega$ and 2$\omega$ terms in
 Eq.~\eqref{chis}. 
@@
 These resonances
-{\color{\chon} come} from the electronic states of the 
+come from the electronic states of the 
 $2\times 1$ surface, that lie inside the bulk band gap of Si and are the 
@@
 without changing the 
-{\color{\chon} spectral shape}; this also can be confirmed from
+spectral shape; this also can be confirmed from
 the cases of zero scissors correction from Fig.~\ref{fig2}.
@@
 From Fig.~\ref{fig4}
-{\color{\chon} we note that the scissors correction 
+we note that the scissors correction 
 shifts the spectra to higher energies as expected. However, contrary 
@@
 $1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
-for the non-rigid shift.}
+for the non-rigid shift.
 \begin{figure}
@@
 $\chi^{xxx}_{2\times 1}$
-vs $\hbar\omega$ for a slab {\color{\chon} with} 32 
-atomic Si layers plus one {\color{\chon} H layer}, 
+vs $\hbar\omega$ for a slab with 32 
+atomic Si layers plus one H layer, 
 with and without the contribution from $\bfv^\nl$.
@@
 $\chi^{xxx}_{2\times 1}$
-vs $\hbar\omega$ for a slab {\color{\chon} with} 32 
-atomic Si layers plus one {\color{\chon} H layer}, 
+vs $\hbar\omega$ for a slab with 32 
+atomic Si layers plus one H layer, 
 for two different values of 
@@
 $|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$,
-we see that the {\color{\chon} first} two peaks are almost rigidly 
-shifted with a small difference in {\color{\chon} height while} 
+we see that the first two peaks are almost rigidly 
+shifted with a small difference in height while
 the rest of the peaks are modified substantially. 
-This {\color{\chon} behavior comes} from the fact that the first two
-peaks are {\color{\chon} almost exclusively related} to the 
+This behavior comes from the fact that the first two
+peaks are almost exclusively related to the 
 2$\omega$ resonances of
-Eq.~\eqref{chis}{\color{\chon} . The} other peaks are a combination 
+Eq.~\eqref{chis}. The other peaks are a combination 
 of 1$\omega$ and 2$\omega$ resonances 
-{\color{\chon} and yield a more varied} spectrum. 
+and yield a more varied spectrum. 
 This way we see that small changes in the value of the 
@@
 We mention that
-{\color{\chon} the} value of $\hbar\sigma=0.5$ eV, used above, 
+the value of $\hbar\sigma=0.5$ eV, used above, 
 is the ``average'' GW gap taken from 
 Ref.~\onlinecite{rholfingPRB95} 
-{\color{\chon} that} is in agreement with Ref.~\onlinecite{garciaCPC01}. 
-{\color{\chon} The} value of $\hbar\sigma=0.63$ eV is the ``average'' 
+that is in agreement with Ref.~\onlinecite{garciaCPC01}. 
+The value of $\hbar\sigma=0.63$ eV is the ``average'' 
 gap taken from Refs.~\onlinecite{asahiPRB00}, 
-where more {\color{\chon}$\bfk$-points} in the Brillouin zone were
+where more $\bfk$-points in the Brillouin zone were
 used to calculate its GW value.
@@
 $e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
-space, and $A_{n\bfk}(\bfG)$ {\color{\chon} the} expansion coefficients. Using
+space, and $A_{n\bfk}(\bfG)$ the expansion coefficients. Using
 $m_e\bfv=\bfp=-i\hbar\bfgnabla$ into 
@@
 is the length of the super cell. 
-{\color{\chon} The} reciprocal lattice vectors $\bfG$ are 
+The reciprocal lattice vectors $\bfG$ are 
 decomposed into components
-{\color{\chon} parallel ($\bfG_\parallel$), and perpendicular ($G_\perp \hat z$)} 
+parallel ($\bfG_\parallel$), and perpendicular ($G_\perp \hat z$)
 to the surface, so
@@
 The double summation over the $\bfG$ vectors can be 
-{\color{\chon} calculated efficiently} by  
+calculated efficiently by  
 creating a pointer array to identify all the plane-wave coefficients  
@@
 \end{align}  
-If $\calc(z)=1$, {\color{\chon} then $f(G_\perp)=\gd_{G_\perp 0}$ and we 
+If $\calc(z)=1$, then $f(G_\perp)=\gd_{G_\perp 0}$ and we 
 obtain the full-slab/bulk values, 
 $\bfv_{nm}(\bfk)$ and $\calc_{nm}(\bfk)=\gd_{nm}$,
-from Eqs.~\eqref{eni.2} and \eqref{eni.4}}.
+from Eqs.~\eqref{eni.2} and \eqref{eni.4}.
 
@@
 $
-one can show that{\color{\chon} 
+one can show that
 \begin{align}\label{nmesn}
@@
 ,
-\end{align}}
+\end{align}
 where
@@
 Notice that
-{\color{\chon} 
-$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$} is obtained 
+$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$ is obtained 
 from Eq.~\eqref{nmesn} by 
 taking 
-{\color{\chon} $C(z)=1$ or $C_{nm}=\gd_{nm}$.}
+$C(z)=1$ or $C_{nm}=\gd_{nm}$.
 
@@
 =(v^{\lda,\rma}_{nm}/i\go^\lda_{nm})_{;k^{\rmb}}
-$ {\color{\chon} and} simply apply the chain rule,
+$ and simply apply the chain rule,
 \begin{align}\label{rka}
@@
 \end{eqnarray}
-and{\color{\chon} 
+and
 \begin{align}\label{tau.1n} 
@@
 .
-\end{align}}
+\end{align}
 Eq.~\eqref{rka} generalizes the usual expresion of
-$(r^a_{nm})_{;k^b}$ for {\color{\chon} a} local 
-{\color{\chon} Hamiltonian}
+$(r^a_{nm})_{;k^b}$ for a local 
+Hamiltonian
 \cite{aversaPRB95,nastosPRB05,cabellosPRB09,rashkeevPRB98}
@@
 $\calt^{\rma\rmb}_{nm}$ reduces to $T^{\rma\rmb}_{nm}$
-{\color{\chon} for the full-slab/bulk case.}
+for the full-slab/bulk case.
 
@@
 and any function $F(\bfr)$, 
-{\color{\chon} to }obtain 
+to obtain 
 \begin{align} 
@@
 where we remind the reader that $\bfr_{nm}$ 
-{\color{\chon} is} calculated through 
+is calculated through 
 Eq.~\eqref{chon.10} for LDA.
@@
 \mathrm{exp}(i(\bfk+\bfG)\cdot\bfr)$
-{\color{\chon} to} obtain,
+to obtain,
 \begin{align}\label{vnl.5}
@@
 $ 
-{\color{\chon} and} their $\bfK$ and $\bfK'$ gradient 
+and their $\bfK$ and $\bfK'$ gradient 
 can be readily calculated.\cite{mottaCMS10,adolphPRB96,gordienkoRPJ04,fuchsCPC99} 
@@
 implemented 
-the calculation of $\calbv^{\nl}_{nm}(\bfk)$ with the help {\color{\chon} of} 
+the calculation of $\calbv^{\nl}_{nm}(\bfk)$ with the help of 
 the \depe~code.\cite{francesco}
-{\color{\chon} We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
-expressions by t}aking $C(z)=1$, which implies 
-{\color{\chon} that} $f(G_\perp)=\gd_{G_\perp 0}$, 
+We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
+expressions by taking $C(z)=1$, which implies 
+that $f(G_\perp)=\gd_{G_\perp 0}$, 
 and 
@@
 \end{align} 
-{\color{\chon} and we expand the triple commutator to obtain,}
+and we expand the triple commutator to obtain,
 \begin{align}\label{3.5}
@@
 The double derivatives with respect to $\bfK$ and $\bfK'$ 
-can be worked out explicitly {\color{\chon} and} 
+can be worked out explicitly and 
 $L^{\rma\rmb}_{nm}(\bfk)$
-{\color{\chon} can} be finally calculated.\cite{valerie}
+can be finally calculated.\cite{valerie}
 
@@
 where $\sum_q\ket{q\bfk}\bra{q\bfk}=1$ was used
-{\color{\chon} and $\bfv^S_{nm}$ is} given in Eq.~\eqref{chon.2}.
+and $\bfv^S_{nm}$ is given in Eq.~\eqref{chon.2}.
 Taking the generalized derivative of Eq.~\eqref{a.3b}
-{\color{\chon} and} applying
+and applying
 the chain rule, we obtain

### 2014-11-20 paper-v-0.tex
cleaned up color changes and finished intro

@@
 %\input{/Users/bms/util/definitions}
-%\def\chon[1]{\color{red} #1}
 \def\chon{red}
-\def\chonn{red}
 \def\nicolas{red}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{showlabels}
+%\usepackage{showlabels}
 %\usepackage{showkeys}
@@
 % notes:
-% nonlocal, not non-local
 % check for tense consistency
-% surface SHG = SSHG
-% $\bfk$-points, not $\bfk$ points
-% Nature Materials de valerie
 % Local fields?
@@
 
-In recent years, surface nonlinear optical spectroscopies and in 
-particular, surface second-harmonic generation SSHG have evolved as useful 
-nondestructive and noninvasive tool to study properties of surfaces 
-and interfaces: atomic structure, phase transitions, and adsorption of 
-atoms, among others.\cite{
-    daumPRL93, 
-    mcgilpOE94, 
-    meyerPRL95, 
-    powerPRL95, 
-    godefroyAPL96, 
-    hoferAPA96, 
-    dadapPRB97, 
-    bloembergenAPB99, 
-    mcgilpSRL99, 
-    suzukiAPB99, 
-    mitchellSS01,
-    hughesPRB96, 
-    guyotPRB88, 
-    downerPSSA01, 
-    shenAPB99, 
-    shenNAT89, 
-    chenPRL81, 
-    mendozaPRL98,
-    downerSIA01,
-    cazzanelliNM14}
-Nowadays, SSHG spectroscopy is a crucial tool for research and development
-in microelectronics,\cite{zheltikovLP00} semiconductors,
-\cite{lupkeSSR99}, nanomaterials,\cite{salazar-aparicioPRB14} 
-and many more recent 
-areas of great scientific and commercial interest.\cite{cazzanelliNM14}
-The high surface sensitivity of SSHG spectroscopy is due to the
-fact that, within the dipole approximation, the bulk SHG signal of
-centrosymmetric materials is identically zero, and thus only the
-surface, where the inversion symmetry is broken, can radiate. 
-Much of the foundation of surface science has been built from
-experiments involving emission or scattering of electrons from
-surfaces. These require ultrahigh vacuum (UHV) environments and
-provide no access to buried interfaces. 
-However, SSHG
-is compatibile with non-UHV ambients and has access to
-interfaces buried beneath transparent overlayers. Even when applied to
-surfaces in UHV, the light source and detectors can be aligned and
+In recent years surface nonlinear optical spectroscopies, particulary surface 
+second-harmonic generation (SSHG), have evolved as useful nondestructive 
+and noninvasive tools to study surface and interface properties. These 
+properties include atomic structure, phase transitions, adsorption of 
+atoms, and many others.\cite{daumPRL93, mcgilpOE94, meyerPRL95, powerPRL95, 
+godefroyAPL96, hoferAPA96, dadapPRB97, bloembergenAPB99, mcgilpSRL99, 
+suzukiAPB99, mitchellSS01, hughesPRB96, guyotPRB88, downerPSSA01, shenAPB99, 
+shenNAT89, chenPRL81, mendozaPRL98, downerSIA01} Nowadays, SSHG spectroscopy 
+is a crucial tool for research and development in microelectronics,
+\cite{zheltikovLP00} semiconductors, \cite{lupkeSSR99} nanomaterials,
+\cite{salazar-aparicioPRB14} and many more recent areas of great scientific 
+and commercial interest.\cite{cazzanelliNM14} The high surface sensitivity 
+of SSHG spectroscopy is due to the fact that within the dipole approximation 
+the bulk SHG signal of centrosymmetric materials is identically zero. 
+Radiation can occur only at the surface where the inversion symmetry is 
+broken. Much of the foundation of surface science has been built from 
+experiments involving emission or scattering of electrons from surfaces. 
+These require ultrahigh vacuum (UHV) environments and provide no access to 
+buried interfaces. However, SSHG is compatibile with non-UHV conditions and 
+has access to interfaces buried beneath transparent overlayers. Even when 
+applied to surfaces in UHV, the light source and detectors can be aligned and
 used outside the vacuum chamber.
 
-The usefulness of SSHG could be limited by
-the lack of microscopic theoretical
-understanding of the nonlinear spectra. The macroscopic
-phenomenological theory of SSHG, which relates the intensity,
-phase and polarization of detected fields to nonlinear and linear
-susceptibilities at the material interface, is now fully developed. 
-However, microscopic theory, which relates atomic-level structure to
-the nonlinear source polarization, is still being developed.
-In this article we put forward an approach to calculate the
+The usefulness of SSHG could be limited by the lack of microscopic theoretical
+understanding of the nonlinear spectra. The macroscopic phenomenological 
+theory of SSHG, which relates the intensity, phase, and polarization of 
+detected fields to nonlinear and linear susceptibilities at the material 
+interface is now fully developed. However, microscopic theory that relates 
+atomic-level structure to the nonlinear source polarization is still being 
+developed. In this article we put forward an approach to calculate the
 microscopic second harmonic surface susceptibility that encompasses several
-theoretical features not taken into account before. 
-Within the independent particle approximation (IPA),  
-some frameworks for bulk SHG have been developed to study the 
-nonlinear optical response of bulk materials, where several degrees of
-sophistication were 
-introduced.\cite{butcherPOPS63,aspnesPRB72,sipePRB93,levinePRB94,aversaPRB95,hughesPRB96,rashkeevPRB98}
-
-As it is well known
-Density Functional Theory (DFT) within the Local Density Approximation
-(LDA),\cite{kohnPR65} the most 
-used framework for \textit{ab initio} calculations,
-underestimates the energy band gap of semiconductors. It is also well
-understood that one has to include the many-body interaction to
-correct for the underestimation of the gap. In this context the so
-called GW approximation\cite{onidaRMP02} is known to correct the band
-gap of most semiconductors\cite{luceroJPCM12}. However this could be a numerically very
-expensive calculation, and thus one uses the much simpler scissors
-operator scheme.\cite{levinePRL89, levinePRL91}
-This allows to ``open'' the DFT-LDA gap to
-its correct experimental or GW value, when experiments are not available. 
-In this context,
-to correct for the underestimation of the energy band gap of semiconductors, 
-Nastos et al.\cite{nastosPRB05} 
-using the 
-``length gauge'' or 
-``$\mathbf{r}\cdot\mathbf{E}$ gauge'' 
-showed how to correctly include in the SH susceptibility the many-body corrections through the
-scissors operator.
-Later,
-Cabellos \textit{et al}.\cite{cabellosPRB09} 
-elaborated a derivation of the ``velocity gauge'' or
-``$\bfA\cdot\bfv$'' gauge properly including the scissors
-operator, and proved  gauge 
-invariance with respect to the length gauge.
-From these works, it is clear the length gauge is a much better staring
-point to obtain the surface SH (SSH) susceptibility, as it will be
-elaborated in this article. Above only being valid for bulk semiconductors.
-
-Concerning the optical response of surfaces and interfaces,
-Reining \textit{et al.}\cite{reiningPRB94}  
-introduced the concept of a 
-cut function in order to obtain the surface SH 
-susceptibility tensor.
-This cut function is required since one usually uses a slab approach
-when treating semi-infinite surface systems; if the slab is
-centrosymmetric the susceptibility tensor will be identically zero.
-The cut function is such that it separates the nonlinear response
-the two surfaces of the slab, thus avoiding the destructive
-interference between them, giving a finite value of the tensor, that
-one identifies with the SSH susceptibility tensor; if the slab is not
-centrosymmetric the cut function can be used to separate the
-different signals coming from either surface of the slab.
-Indeed, one of the main results of this article is to show that the SSH 
-susceptibility tensor obtained by using the cut function is 
-correctly extracted from the slab.   
-After Reining \textit{et al.}\cite{reiningPRB94}, 
-Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02}
-followed upon this work and in particular Ref.~\onlinecite{arzatePRB01}
-went into a detailed analysis of the different contributions to the 
-SHG spectra of a surface and the nuanced relationship between bulk, 
-surface, interband, intraband, $1\omega$ and $2\omega$ terms. 
-Perhaps of more relevance to this work, 
-Mej\'ia \textit{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
-the nonlinear responses of semiconductor systems, within a 
-Tight-Binding framework. This model 
-allows for obtaining results from selected regions of a system,
-including the surface. 
-However, in these references the scissors is either not implemented or
-wrongly implemented.
+theoretical features not taken into account before. Within the independent 
+particle approximation (IPA) some frameworks for bulk SHG have been developed 
+to study the nonlinear optical response of bulk materials with varying degrees 
+of sophistication.\cite{butcherPOPS63, aspnesPRB72, sipePRB93, levinePRB94,
+aversaPRB95, hughesPRB96, rashkeevPRB98}
+
+The most used framework for \textit{ab initio} calculations, Density 
+Functional Theory (DFT) within the Local Density Approximation (LDA),
+\cite{kohnPR65} underestimates the energy band gap of semiconductors. It is 
+well understood that one has to include the many-body interaction to correct 
+for this underestimation of the gap. In this context, the so-called GW 
+approximation\cite{onidaRMP02} is known to correct the band gap of most 
+semiconductors\cite{luceroJPCM12}. However this can be a very expensive
+calculation and thus one uses the much simpler scissors operator scheme.
+\cite{levinePRL89, levinePRL91} This allows us to ``open'' the DFT-LDA gap to 
+its correct experimental or GW value even when experiments are not available. 
+In this context, to correct for the underestimation of the energy band gap of 
+semiconductors Nastos et al.\cite{nastosPRB05} used the ``length gauge'' or 
+``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to correctly include the 
+many-body corrections through the scissors operator in the SH susceptibility.
+Later, Cabellos \textit{et al}.\cite{cabellosPRB09} elaborated a derivation 
+of the ``velocity gauge'' or ``$\bfA\cdot\bfv$'' gauge properly including the 
+scissors operator and proved gauge invariance with respect to the length 
+gauge. From these works it is clear the length gauge is a much better staring
+point to obtain the surface SH (SSH) susceptibility, as will be elaborated 
+in this article. However, these considerations are only valid for bulk 
+semiconductors.
+
+Concerning the optical response of surfaces and interfaces, Reining 
+\textit{et al.}\cite{reiningPRB94} introduced the concept of a cut function 
+in order to obtain the surface SH susceptibility tensor. This cut function 
+is required since one usually uses a slab approach when treating semi-infinite 
+surface systems. If the slab is centrosymmetric the susceptibility tensor will 
+be identically zero. The cut function is such that it separates the nonlinear 
+response for the two surfaces of the slab avoiding the destructive 
+interference between them giving a finite value that one identifies with the 
+SSH susceptibility tensor. If the slab is not centrosymmetric the cut function 
+can be used to separate the different signals coming from either surface of 
+the slab. Indeed, one of the main results of this article is to show that the 
+SSH  susceptibility tensor obtained by using the cut function is correctly 
+extracted from the slab. After Reining \textit{et al.},\cite{reiningPRB94}
+Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02} followed 
+upon this work and in particular Ref.~\onlinecite{arzatePRB01} went into a 
+detailed analysis of the different contributions to the SHG spectra of a 
+surface and the nuanced relationship between bulk, surface, interband, 
+intraband, $1\omega$ and $2\omega$ terms. Perhaps of more relevance to this 
+work, Mej\'ia \textit{et al}.\cite{mejiaRMF04} developed a layer-by-layer 
+analysis for the nonlinear responses of semiconductor systems, within a 
+tight-binding framework. This model allows for obtaining results from 
+selected regions of a system including the surface. However, in these 
+references the scissors operator is either excluded or incorrectly 
+implemented.
 
 Finally, DFT-LDA calculations are often based on the use of pseudopotentials
-that usually contain spatial non-localities to include the interaction
-of the valence electrons with the core electrons.
-As it will be discussed in this article, 
- this non-local
-dependence on the position vector, introduces corrections to
-the velocity operator of the electron, that have to be included with
-care in the SSH susceptibility. So far, this contribution has been
-neglected in the literature.
-
-Therefore, within the IPA, 
-the most complete approach 
-for the calculation of the SSH susceptibility is one which includes
-i) the scissors correction, ii) the cut function, and iii) the  
-contribution of the non-local part of the pseudopotentials.  
-Therefore,
-the goal of this paper is to derive, within the length gauge, 
-new expressions for the 
-SSH susceptibility tensor 
-$\chi^{\rma\rmb\rmc}(-2\omega;\omega,\omega)$ 
-that include the aforementioned contributions.
-The inclusion of these three contributions 
-makes 
-our scheme unprecedented and opens the possibility to study 
-surface SHG with more versatility and providing 
-more accurate results than previous approaches.
-
-The paper is organized as follows. 
-In Sec. \ref{theory} we present the 
-relevant steps for deriving the surface second-order susceptibility
-tensor
-$\chi^{\rma\rmb\rmc}(-2\omega;\omega,\omega)$  
-within the length gauge formalism. This derivation includes the addition of 
-the terms mentioned above
-that have 
-been absent in previous works. Also,
-our $\chi^{\rma\rmb\rmc}(-2\omega;\omega,\omega)$  
-can be used for a layer-by-layer
-analysis if wanted.
-In Sec. \ref{results} we show the results of this 
-reformulation with a study of a clean Si(001) surface with a 
-$2 \times 1$ reconstruction. 
-We compare 
-results from before and after adding the different nonlocal contributions.
-Finally, in Sec. \ref{conc} we give our conclusions.
+of the valence electrons with the core electrons. As it will be discussed 
+in this article, dependence on the position vector introduces corrections to
+the velocity operator of the electron that have to be included with care in 
+the SSH susceptibility. So far, this contribution has been neglected in the 
+literature.
+
+Therefore, within the IPA the most complete approach for the calculation of 
+the SSH susceptibility is one which includes i) the scissors correction, 
+ii) the cut function, and iii) the contribution of the nonlocal part of the 
+pseudopotentials. Therefore, the goal of this paper is to derive new 
+expressions within the length gauge for the SSH susceptibility tensor 
+$\chi^{\textrm{a}\textrm{b}\textrm{c}}(-2\omega;\omega,\omega)$ 
+that include the aforementioned contributions. The inclusion of these three 
+contributions makes our scheme unprecedented and opens the possibility to 
+study surface SHG with more versatility and providing more accurate results 
+than previous approaches.
+
+The paper is organized as follows. In Sec. \ref{theory} we present the 
+relevant steps for deriving the surface second-order susceptibility tensor
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
+length gauge formalism. This derivation includes the addition of the terms 
+mentioned above that have been absent in previous works. Also, our 
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ can be used 
+for a layer-by-layer analysis if desired. In Sec. \ref{results} we show the 
+results of this reformulation with a study of a clean Si(001) surface with a 
+$2 \times 1$ reconstruction. We compare results from before and after adding 
+the different nonlocal contributions. Finally, in Sec. \ref{conc} we give 
+our conclusions.
 
@@
 $\chi^{abc}(-2\omega;\omega,\omega)$, we asume
-that the crystal potential could have a non-local spatial
 contribution as is customary for most
@@
 $V(\bfr)$ and $V^{\mathrm{nl}}(\bfr,\bfp)$  are the local and 
-non-local contributions to the pseudopotential $V^{\mathrm{ps}}$,
  respectively.
@@
 we simply omit the contribution coming
-from the non-local part of the pseudopotential, $V^{\mathrm{nl}}$, in
 the results that we have derived.
@@
 structures with dispersion relations similar to those predicted by the GW
-approximation. Mathematically, the scissors (non-local) operator 
 $S$ is added to the unperturbed or unscissored Hamiltonian $H^{\mathrm{LDA}}_{0}$ ,
@@
 includes only the contribution of 
-the non-local part of the pseudopotentials through
 $\bfv^\nl_{nm}(\bfk)$. These can be readily
@@
 contribution of the region of the slab given by $\calc(z)$.
-We note that this modified operator is hermitian as it must.\cite{nota3}
+We note that this modified operator is hermitian as it must.\cite{note2}
 The operator $\bfV$ could be any of those given by Eq.~\eqref{vop2},
@@
 Thus, calculating $\chi^{xxx}$ for the full-slab, or the 
-half-slab containing the $2\times 1$ surface\cite{nota1}
+half-slab containing the $2\times 1$ surface\cite{note1}
 should yield the same result since the contribution from the H
@@
 $\hbar\sigma=0.5$ eV, that is the GW gap reported in
-Ref.~\onlinecite{rholfingPRB95,garciaCPC01}. 
+Ref.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
 We see that for all four instances the 
@@
 this article; through the proposed layer formalism we can calculate the surface SH
-susceptibility including the effect of the non-local part of the
 pseudopotentials
@@
 
-\subsection{\texorpdfstring{Results for $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$}{Results for Xxxx(2x1)(-2w;w,w)}}
+\subsection{\texorpdfstring{Results for $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$}
+{Results for Xxxx(2x1)(-2w;w,w)}}
 
@@
 is the ``average'' GW gap taken from 
-Ref.~\onlinecite{rholfingPRB95} 
+Ref.~\onlinecite{rohlfingPRB95} 
 that is in agreement with Ref.~\onlinecite{garciaCPC01}. 
@@
 
-We have presented a formulation to calculate surface second harmonic
-(SSH) susceptibility tensor $\bfgchi(-2\go;\go,\go)$,  using a length gauge
+We have presented a formulation to calculate the surface second harmonic
+(SSH) susceptibility tensor $\bfgchi(-2\go;\go,\go)$, using the length gauge
 formalism and within the independent particle approximation (IPA).
@@
 i) the scissors correction, ii) the cut function, and iii) the 
-contribution of the non-local part of the pseudopotentials. 
 We have used a Si(001)$2\times 1$ surface to confirm that our scheme
@@
 
-%%%%%%%%%%%%%%
 \newpage
-\bibliography{ref}% 
-\end{document}  %
-%%%%%%%%%%%%%%
-
-\section{Introduction}\label{intro}
-
-The first attempt to go beyond the independent particle approximation 
-could be found in 
-Leitsmann \textit{et al}.\cite{leitsmannPRB05} 
-that included excitonic and local field interactions. 
-More recently, Luppi \textit{et al.}\cite{luppiPRB10,luppiJCP10} and 
-H\"ubner \textit{at al.}\cite{hubenerPRB11}, studied 
-the influence of crystal local fields and excitonic effects on 
-the spectrum of the second harmonic generation of several bulk 
-systems 
-by using time-dependent density-functional theory 
-including many-body effects, namely, quasiparticle corrections through 
-the scissors operator, crystal local field effects, and excitonic 
-interaction.  
-Also,
-Trolle \textit{et al}.\cite{trollePRB14} 
-elaborate on a theoretical model for SHG of 
-mono-layer MoS$_{2}$. Their approach is based on the tight-binding model 
-fitted to GW results and include excitonic effects from the 
-Bethe-Salpeter exciton picture. However, neither of these models 
-has a consideration for surfaces.
+\bibliography{ref}
+\end{document}

### 2014-11-20 paper-v-0.tex
finished conclusions

@@
 We take the $z$ coordinate pointing out of the surface and the $x$
-coordinate, along the crystallographic [011] direction, is parallel to the dimmers. 
+coordinate along the crystallographic [011] direction is parallel to the dimmers. 
 \begin{figure}
@@
 We show the results for this comparison in the remainder 
-of this section
+of this section.
 
@@
 $\hbar\sigma=0.5$ eV, that is the GW gap reported in
-Ref.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
+Refs.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
 We see that for all four instances the 
@@
 and the many-body effects through the scissors correction.
-Our scheme ought to work for any slab.
+Our scheme should to work for any slab.
 
@@
 Notice that the 
-2$\omega$ resonances start below $E_g/2$ where $E_g=??$ eV is the LDA
-band gap of this surface, and $E_g=??$ eV is the scissor shifted gap for
+2$\omega$ resonances start below $E_g/2$ where $E_g=0.53$ eV is the LDA
+band gap of this surface, and $E_g=1.035$ eV is the scissor shifted gap for
 $\hbar\sigma=0.5$ eV.
@@
 The value of $\hbar\sigma=0.63$ eV is the ``average'' 
-gap taken from Refs.~\onlinecite{asahiPRB00}, 
+gap taken from Ref.~\onlinecite{asahiPRB00}, 
 where more $\bfk$-points in the Brillouin zone were
@@
 
-Although, local fields are neglected, we expect that parallel to an
-interface they should be in principle 
-small, as the electric field is continuous. So, we would expect that the $xxx$ component 
-of 
-$\bfgchi(-2\go;\go,\go)$, would have an small influence from the local fields.
+Although local fields are neglected they should, in principle, 
+be small parallel to the interface as the electric field is continuous.
+So, we would expect that the $xxx$ component of 
+$\bfgchi(-2\go;\go,\go)$ would have a small influence from the local fields.
 Unfortunately the experimental measurement of the $xxx$
-component of $\bfgchi(-2\go;\go,\go)$ is not possible, as the SH 
+component of $\bfgchi(-2\go;\go,\go)$ is not possible as the SH 
 radiated intensity would be proportional not only to this component 
-but also to other components of $\bfgchi(-2\go;\go,\go)$,  
-however,
+but also to the other components of $\bfgchi(-2\go;\go,\go)$.  
+However,
 in a forthcoming publication 
@@
 formalism and within the independent particle approximation (IPA).
-It includes on the same footing and for the first time in the literature 
+It includes on equal footing and for the first time in the literature: 
 i) the scissors correction, ii) the cut function, and iii) the
+contribution of the non-local part of the pseudopotentials.
 We have used a Si(001)$2\times 1$ surface to confirm that our scheme
@@
 $
-Although, one can in principle increase the number of atomic layers,
+Although one can in principle increase the number of atomic layers,
 $\bfk$-points, etc. to
@@
 hava chosen a good compromise between accuracy and the burden and time
-of the computational scheme. 
+of the computations. 
 We describe the effect of the independent inclusion of the three
 effects mentioned above in the calculation of  
-$\bfgchi(-2\go;\go,\go)$. To wit,
-the scissors correction shifts the spectrum to higher energies, though
-the shifting is not rigid and mixes the 1$\omega$ and 2$\omega$
-resonances, the cut functions
-allow us
-to extract unequivocally   
-$\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$, 
-and the effects of the nonlocal part of the pseudopotentials,
+$\bfgchi(-2\go;\go,\go)$. The scissors correction shifts the spectrum to 
+higher energies though the shifting is not rigid and mixes the 
+$1\omega$ and $2\omega$ resonances, the cut function allows us
+to extract unequivocally $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$, 
+and the effects of the nonlocal part of the pseudopotentials
 reduces the value of   
-$|\chi^{xxx}_{2\times 1}(-2\go;\go,\go)|$, keeping the same lineshape. 
+$|\chi^{xxx}_{2\times 1}(-2\go;\go,\go)|$ keeping the same lineshape. 
 The $xxx$ component of  
 $\bfgchi_{2\times 1}(-2\go;\go,\go)$, 
-could not be
+can not be
 experimentally isolated,
@@
 our scheme opens the unprecedented possibility to study 
-surface SHG with more versatility and providing 
+surface SHG with more versatility and provides 
 more accurate results than previous approaches.

### 2014-11-20 paper-v-0.tex
ready to convert to real LaTeX

@@
-\documentclass[floatfix,prb,aps,superscriptaddress,11pt,preprint,letterpaper]{revtex4}
+\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
 \usepackage{amsfonts}
@@
 
+\pacs{78.68.+m, 42.65.An, 71.15.Mb, 42.65.Ky, 78.66.-w}
+
 \begin{document}
 
-\title{Length Gauge Theory of Surface Second Harmonic Generation}
+\title{Theory of surface second-harmonic generation for semiconductors
+including effects of nonlocal operators}
 \author{Sean M. Anderson}
@@
 \begin{abstract}
-We present a theoretical review of surface second harmonic generation (SHG) 
-from semiconductor surfaces based on the length gauge. This 
-layer-by-layer analysis is carefully presented in order to show how a 
-surface SHG calculation can be readily evaluated. The nonlinear susceptibility 
-tensor for a surface, $\boldsymbol{\chi}^S(-2\go;\go,\go)$
-is split into two terms relating to inter-band 
-and intra-band one-electron transitions.
-
+We formulate a theoretical approach of surface second-harmonic generation 
+from semiconductor surfaces based on the length gauge. 
+Within the independent particle approximation 
+the nonlinear second-order surface susceptibility tensor
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$   
+is calculated, including for the first time in one unique formulation
+i) the scissors correction, needed to have the correct value of the
+energy band gap,
+ii) the derivation for the inclusion of the cut function, 
+used to extract the surface nonlinear response,
+and iii) the contribution of the nonlocal part of the 
+pseudopotentials, routinely  used in \textit{ab initio} band structure calculations. 
+First and third terms are described by spatially nonlocal quantum
+mechanical operators and are fully taken into account in the present formulation. 
+As a test-case of the approach we calculate 
+$\chi^{xxx}(-2\omega;\omega,\omega)$ for the clean Si(001)$2\times 1$
+reconstructed surface and find that our scheme to extract the surface
+through the cut function
+readily works and is exact. The effects of the scissors correction and
+of the nonlocal part of the pseudopotentials are discussed for the
+first time in surface nonlinear optics. 
+The scissors correction shifts the spectrum to 
+higher energies though the shifting is not rigid and mixes the 
+$1\omega$ and $2\omega$ resonances, 
+and has a strong influence in the line-shape. 
+The 
+effects of the nonlocal part of the pseudopotentials 
+keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\go;\go,\go)|$, but 
+reduces its value
+by 15-20\%.  
+Therefore, the inclusion of the three aforementioned 
+contributions is very important and 
+makes our scheme unprecedented and opens the possibility to 
+study SSHG with more versatility and providing more accurate results 
+than previous approaches. 
 \end{abstract}
@@
 developed. In this article we put forward an approach to calculate the
-microscopic second harmonic surface susceptibility that encompasses several
+microscopic second-harmonic surface susceptibility that encompasses several
 theoretical features not taken into account before. Within the independent 
@@
 calculation and thus one uses the much simpler scissors operator scheme.
-\cite{levinePRL89, levinePRL91} This allows us to ``open'' the DFT-LDA gap to 
+\cite{levinePRL89,levinePRL91,delsolePRB93}  
+This allows us to ``open'' the DFT-LDA gap to 
 its correct experimental or GW value even when experiments are not available. 
@@
 extracted from the slab. After Reining \textit{et al.},\cite{reiningPRB94}
-Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02} followed 
+Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02,sanoPRB02}
+followed 
 upon this work and in particular Ref.~\onlinecite{arzatePRB01} went into a 
@@
 surface and the nuanced relationship between bulk, surface, interband, 
-intraband, $1\omega$ and $2\omega$ terms. Perhaps of more relevance to this 
-work, Mej\'ia \textit{et al}.\cite{mejiaRMF04} developed a layer-by-layer 
+intraband, $1\omega$ and $2\omega$ terms, and
+Ref. \onlinecite{mejiaRMF04} developed a layer-by-layer 
 analysis for the nonlinear responses of semiconductor systems, within a 
@@
 
-Finally, DFT-LDA calculations are often based on the use of pseudopotentials
-of the valence electrons with the core electrons. As it will be discussed 
-in this article, dependence on the position vector introduces corrections to
-the velocity operator of the electron that have to be included with care in 
+Finally, DFT-LDA calculations are often based on the use of pseudopotentials.
+As it will be discussed 
+in this article, the presence of a nonlocal part of the
+pseudopotential introduces
+corrections to
+the momentum operator of the electron that have to be included with care in 
 the SSH susceptibility. So far, this contribution has been neglected in the 
@@
 ii) the cut function, and iii) the contribution of the nonlocal part of the 
-pseudopotentials. Therefore, the goal of this paper is to derive new 
-expressions within the length gauge for the SSH susceptibility tensor 
-$\chi^{\textrm{a}\textrm{b}\textrm{c}}(-2\omega;\omega,\omega)$ 
+pseudopotential. Therefore, the goal of this paper is to derive a new 
+expression within the length gauge for the SSH susceptibility tensor 
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ 
 that include the aforementioned contributions. The inclusion of these three 
@@
 results of this reformulation with a study of a clean Si(001) surface with a 
-$2 \times 1$ reconstruction. We compare results from before and after adding 
+$2 \times 1$ reconstruction, 
+proving the correctness of our approach with a special test-case,
+never exploited before.
+We compare results from before and after adding 
 the different nonlocal contributions. Finally, in Sec. \ref{conc} we give 
@@
 In this section we present the scheme used to calculate the surface second-order
-nonlinear response.
-Our derivation includes new terms not included
-before in the length gauge formalism for the surface nonlinear
+nonlinear response using the length gauge formalism.
+Our derivation includes the aforementioned terms not included
+before for the surface nonlinear
 response.
@@
 
-We use the IPA and neglect local field  
+We assume the IPA and neglect local field  
 and excitonic effects.  
@@
 with $\mathbf{v}$ the velocity operator of the electron, and $A$ the
-area of the unit cell. We calculate the macroscopic polarization density 
+area of the unit cell. We calculate the polarization density 
 $\mathbf{P}$, related to $\bfJ$ by
 $\bfJ=d\mathbf{P}/dt$. For a 
-perturbing (Maxwell macroscopic) electromagnetic field, $\mathbf{E}(t)=
+perturbing electromagnetic field, $\mathbf{E}(t)=
 \mathbf{E}(\omega )e^{-i\tilde{\omega} t}+c.c.$,
@@
 and $\eta >0$ adiabatically switches on the interaction,
-we write the second order nonlinear
+we write the second-order nonlinear
 polarization as, 
@@
 where $\chi^{abc}(-2\omega ;\omega ,\omega )$ is the nonlinear
-susceptibility responsible for surface second harmonic generation
+susceptibility responsible for surface second-harmonic generation
 (SSHG). 
@@
 where 
-$V(\bfr)$ and $V^{\mathrm{nl}}(\bfr,\bfp)$  are the local and 
+$V(\bfr)$ and $V^{\mathrm{nl}}(\bfr,\bfp)$  are the local and nonlocal parts, 
  respectively.
@@
 conduction bands are rigidly shifted in energy so that the band gap 
-corresponds to the accepted experimental band gap.
+corresponds to the accepted 
+experimental band gap.\cite{levinePRL89,levinePRL91,delsolePRB93} 
 This is often in fairly good agreement with the GW
@@
 The LDA wave functions are used since they produce band
-structures with dispersion relations similar to those predicted by the GW
+structures with dispersion relations similar to those predicted by the GW.
+Mathematically, the scissors (non-local) operator 
 $S$ is added to the unperturbed or unscissored Hamiltonian $H^{\mathrm{LDA}}_{0}$ ,
 \begin{equation*}
-H^\sigma_{0}(\bfr,\bfp)=H^{\mathrm{LDA}}_{0}(\bfr,\bfp)+S(\bfr,\bfp)
+H^\gS_{0}(\bfr,\bfp)=H^{\mathrm{LDA}}_{0}(\bfr,\bfp)+S(\bfr,\bfp)
 \end{equation*}
@@
 \begin{equation}
-S(\bfr,\bfp)=\hbar \sigma\sum_{n}\int d^{3}k(1-f_{n})
+S(\bfr,\bfp)=\hbar \gD\sum_{n}\int d^{3}k(1-f_{n})
 |n\mathbf{k}\rangle\langle n\mathbf{k}|,
@@
 \end{equation}
-with $\hbar \sigma$  the rigid ($\mathbf{k}$-independent) energy correction to be
+with $\hbar \gD$  the rigid ($\mathbf{k}$-independent) energy correction to be
 applied. 
@@
 \label{hamils} \\
-H_{0}^\sigma (\bfr,\bfp)\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\sigma
+H_{0}^\gS (\bfr,\bfp)\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\gS
 (\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
@@
 where the scissor-shifted energies, 
-$\omega_{n}^\sigma(\mathbf{k})$ , are given by
+$\omega_{n}^\gS(\mathbf{k})$ , are given by
 \begin{equation}
-\omega_{n}^\sigma(\mathbf{k})=\omega^\lda_{n}(\mathbf{k})+(1-f_{n})\sigma.
+\omega_{n}^\gS(\mathbf{k})=\omega^\lda_{n}(\mathbf{k})+(1-f_{n})\gD.
 \label{wese}
@@
 From Eq.~\eqref{mv} with $H_0\to
-H^\sigma_0$, we obtain
+H^\gS_0$, we obtain
 \begin{align}\label{pmnrmn}
@@
 &=
-\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
+\frac{\bfv^\gS_{nm}(\bfk)}{i\go^\gS_{nm}(\bfk)}
 \quad\quad n\notin D_m 
@@
 where we defined
-$\go^\gs_{nm}(\bfk)\equiv\go^\gs_n(\bfk)-\go^\gs_m(\bfk)$, and
+$\go^\gS_{nm}(\bfk)\equiv\go^\gS_n(\bfk)-\go^\gS_m(\bfk)$, and
 $D_m$ are all the possible degenerate $m$-states. 
@@
 we derive a key result for the length gauge formulation. 
-Again, using $H^\sigma_0$ in
+Again, using $H^\gS_0$ in
 Eq.~\eqref{mv} we obtain
 \begin{align}\label{vop2}
-\bfv^\gs&=
+\bfv^\gS&=
 \bfv 
@@
 \begin{equation}\label{chon.2} 
-\bfv^S_{nm}=i\sigma f_{mn}\bfr_{nm},
+\bfv^S_{nm}=i\gD f_{mn}\bfr_{nm},
 \end{equation}
@@
 \begin{align}\label{chon.9}
-\bfv^\sigma_{nm}(\bfk) 
-=\frac{\go^\sigma_{nm}(\bfk)}{\go^\lda_{nm}(\bfk)}
+\bfv^\gS_{nm}(\bfk) 
+=\frac{\go^\gS_{nm}(\bfk)}{\go^\lda_{nm}(\bfk)}
 \bfv^\lda_{nm}(\bfk) 
@@
  \begin{align}\label{chon.8}
-\frac{\bfv^\gs_{nm}(\bfk)}{\go^\gs_{nm}(\bfk)}
+\frac{\bfv^\gS_{nm}(\bfk)}{\go^\gS_{nm}(\bfk)}
 &=
@@
 =
-\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
+\frac{\bfv^\gS_{nm}(\bfk)}{i\go^\gS_{nm}(\bfk)}
 =
@@
 we need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
-$\omega_{n}^{\sigma}$ at the end of the derivation
+$\omega_{n}^{\gS}$ at the end of the derivation
 to obtain the scissored results for any 
@@
 Eq.~\eqref{rde}, we obtain 
-$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\sigma_{nm}-\tilde\omega)t}$,
+$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\gS_{nm}-\tilde\omega)t}$,
 with 
 \begin{equation}\label{j.1}
-B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\sigma_{nm}-\tilde\omega},
+B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\gS_{nm}-\tilde\omega},
 \end{equation}
@@
 \begin{eqnarray}\label{j.2}
-(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\sigma_{nm}-2\tilde\omega}\bigg[
+(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\gS_{nm}-2\tilde\omega}\bigg[
 i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
 \Big)  
--(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\sigma_{nm}-2\tilde\omega)t}
+-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\gS_{nm}-2\tilde\omega)t}
 .
@@
 semiconductor crystal is as follows. Instead of using a
-semi-infinite system, we replace it by a super cell that 
+semi-infinite system, we replace it by a super-cell that 
 consists of a finite
 slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
-super cell constitutes the unit cell, and
+super-cell constitutes the unit cell, and
 is repeated to form a full three dimensional crystalline structure,
-i.e. a repeated super cell scheme. 
+i.e. a repeated super-cell scheme. 
 The slab itself consists of front, back, and 
@@
 \centering
-%\includegraphics[height=5cm,width=7cm]{slab}
 \includegraphics[scale=.7]{images/slab}
-\caption{(color on line) A sketch of the super cell. 
+\caption{(color on line) A sketch of the super-cell. 
 The slab corresponds to the
@@
 the operator for the electron current,
-$\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gs\ket{\bfr}\bra{\bfr}
-+\ket{\bfr}\bra{\bfr}\bfv^\gs\right)$, that leads to
+$\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gS\ket{\bfr}\bra{\bfr}
++\ket{\bfr}\bra{\bfr}\bfv^\gS\right)$, that leads to
 \begin{align}\label{jmic}
@@
 \left(
-\bra{m\bfk}\bfv^\gs\ket{\bfr}\braket{\bfr}{n\bfk}
-+
-\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gs\ket{n\bfk}
+\bra{m\bfk}\bfv^\gS\ket{\bfr}\braket{\bfr}{n\bfk}
++
+\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gS\ket{n\bfk}
 \right),
@@
 to the current density from the
-chosen layer of the slab is given by
+chosen region of the slab is given by
 \begin{equation}\label{jsz}
@@
 \end{equation}
-where $\calbj^{(N)}(t)$ is the current of the
-layer specified by $\calc(z)$.
+where $\calbj^{(N)}(t)$ is the $N$-th order current induced in the
+region specified by $\calc(z)$.
 Therefore we define
 \begin{equation}\label{vcal}
-e{\boldsymbol{\mathcal{V}}}^{\gs}_{mn}(\mathbf{k})
+e{\boldsymbol{\mathcal{V}}}^{\gS}_{mn}(\mathbf{k})
 \equiv
@@
 \sum_{mn}
-\calbv^{\gs}_{mn}(\mathbf{k}) 
+\calbv^{\gS}_{mn}(\mathbf{k}) 
 \rho^{(N)}_{nm}(\bfk;2\go) 
@@
 \end{equation}
-that gives the induced microscopic current of the chosen layer, to order $N$ 
+that gives the induced microscopic current of the chosen region, to order $N$ 
 in the external perturbation. 
-%The matrix elements of the 
-%density operator for $N=2$ is given by Eq.~\eqref{j.2}.
 From
@@
 \begin{align}\label{intj}
-{\boldsymbol{\mathcal{V}}}^{\gs}_{mn}({\mathbf k})
+{\boldsymbol{\mathcal{V}}}^{\gS}_{mn}({\mathbf k})
 &=
@@
 \bigg[
-\langle m\mathbf{k}|\mathbf{v}^\gs | \mathbf{r}\rangle
+\langle m\mathbf{k}|\mathbf{v}^\gS | \mathbf{r}\rangle
 \langle \mathbf{r} | n \mathbf k \rangle +
 \langle m\mathbf{k} | \mathbf{r}\rangle
-\langle \mathbf{r} | \mathbf{v}^\gs | n \mathbf k \rangle\bigg]
+\langle \mathbf{r} | \mathbf{v}^\gS | n \mathbf k \rangle\bigg]
 \nonumber\\
@@
 \psi_{n\mathbf{k}}(\mathbf{r})
-\bfv^{\gs *}\psi^*_{m\mathbf{k}}(\mathbf{r})
-+ 
-\psi^*_{m\mathbf{k}}(\mathbf{r})\bfv^\gs
+\bfv^{\gS *}\psi^*_{m\mathbf{k}}(\mathbf{r})
++ 
+\psi^*_{m\mathbf{k}}(\mathbf{r})\bfv^\gS
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\left[\frac{\calc(z) \bfv^\gs +
-\bfv^\gs \calc(z)}{2}\right]
+\left[\frac{\calc(z) \bfv^\gS +
+\bfv^\gS \calc(z)}{2}\right]
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\calbv^{\gs}
+\calbv^{\gS}
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 %are periodic over the unit cell, the surface term vanishes. 
-where, we used the hermitian property of $\bfv^\gs$ and defined
+where, we used the hermitian property of $\bfv^\gS$ and defined
 \begin{align}\label{nl.4}
-\calbv^{\gs}
+\calbv^{\gS}
 =
-\frac{\calc(z) \bfv^\gs +
-\bfv^\gs \calc(z)}{2}
+\frac{\calc(z) \bfv^\gS +
+\bfv^\gS \calc(z)}{2}
 .
@@
 \begin{align}\label{vopii}
-\calbv^{\gs}
+\calbv^{\gS}
 &=
@@
 To calculate
-$\calbv^{\sigma}_{nm}(\bfk)$ 
+$\calbv^{\gS}_{nm}(\bfk)$ 
 we calculate the matrix elements of 
@@
  and \onlinecite{mejiaRMF04}. 
-In this article, for the first time, we formally introduce the cut function $C(z)$ 
-for the second harmonic optical response of semiconductors.
+In this article, for the first time, we formally introduce the cut function $\calc(z)$ 
+for the second-harmonic optical response of semiconductor surfaces,
+from an average of the second order polarization over the region of interest.
 
@@
 and Eq.~\eqref{jmac2} 
-we obtain the SH polarization of a given layer as
+we obtain the SH polarization of a given region as
 \begin{equation}\label{Pjikn}
@@
 \sum_{mn}
-\calbv^{\gs}_{mn}(\mathbf{k})
+\calbv^{\gS}_{mn}(\mathbf{k})
 \rho^{(2)}_{nm}(\bfk;2\go)
@@
 \sum_{mn}
-\frac{\calv^{\gs,\rma}_{mn}(\mathbf{k})}
-{\go^\gs_{nm\bfk}-2\got}
+\frac{\calv^{\gS,\rma}_{mn}(\mathbf{k})}
+{\go^\gS_{nm\bfk}-2\got}
 \bigg[
@@
 $\chi^{\rma\rmb\rmc}(-2\go;\go,\go)$ 
-of the layer of the slab specified by $\calc(z)$. 
+of the layers of the slab specified by $\calc(z)$. 
 We mention that the units of 
@@
 \sum_{mn}
-\frac{\calv_{mn}^{\gs,\rma}}{\go^\gs_{nm}-2\got}
-\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}}
+\frac{\calv_{mn}^{\gS,\rma}}{\go^\gS_{nm}-2\got}
+\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gS_{nm}-\got}\right)_{;k^{\rmc}}
 ,
@@
 \sum_{qmn}
-\frac{\calv_{mn}^{\gs,\rma}}{\go^\gs_{nm}-2\got}
+\frac{\calv_{mn}^{\gS,\rma}}{\go^\gS_{nm}-2\got}
 \left(
 \frac{r_{nq}^{\rmc} r_{qm}^{\rmb} 
-f_{mq}}{\go^\gs_{qm}-\got}
+f_{mq}}{\go^\gS_{qm}-\got}
 -\frac{r_{nq}^{\rmb} r_{qm}^{\rmc} 
-f_{qn}}{\go^\gs_{nq}-\got}
+f_{qn}}{\go^\gS_{nq}-\got}
 \right),
@@
 \begin{equation}\label{gene2}
-\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gs_{nm}-\got}\right)_{;k^{\rmc}}=
-\frac{f_{mn}}{\go^\gs_{nm}-\got}\left(r_{nm}^\rmb\right)_{;k^{\rmc}}
--\frac{f_{mn}r_{nm}^{\rmb}\gD_{nm}^\rmc}{(\go^\gs_{nm}-\got)^2}
-,
-\end{equation}
-where substituting $H^\sigma_0$ 
+\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gS_{nm}-\got}\right)_{;k^{\rmc}}=
+\frac{f_{mn}}{\go^\gS_{nm}-\got}\left(r_{nm}^\rmb\right)_{;k^{\rmc}}
+-\frac{f_{mn}r_{nm}^{\rmb}\gD_{nm}^\rmc}{(\go^\gS_{nm}-\got)^2}
+,
+\end{equation}
+where substituting $H^\gS_0$ 
 into Eq.~\eqref{conmri3n} and then
@@
 \begin{align}\label{eli.13}
-\left(\go^\gs_{nm}\right)_{;k^{\rma}}
+\left(\go^\gS_{nm}\right)_{;k^{\rma}}
 =
@@
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\gs_{cv}}
+\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\gS_{cv}}
 \left[
-\frac{\mathrm{Im}[\mathcal{V}^{\gs,\mathrm{a}}_{qc}\{r^{\rmb}_{cv}r^{\rmc}_{vq}\}]}
-{(2\go^\gs_{cv}-\go^\gs_{cq})} 
--\frac{\mathrm{Im}[\mathcal{V}^{\gs,\mathrm{a}}_{vq}\{r^{\rmc}_{qc}r^{\rmb}_{cv}\}]}
-{(2\go^\gs_{cv}-\go^\gs_{qv})}
-\right]\gd(\go^\gs_{cv}-\go),
+\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{qc}\{r^{\rmb}_{cv}r^{\rmc}_{vq}\}]}
+{(2\go^\gS_{cv}-\go^\gS_{cq})} 
+-\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vq}\{r^{\rmc}_{qc}r^{\rmb}_{cv}\}]}
+{(2\go^\gS_{cv}-\go^\gS_{qv})}
+\right]\gd(\go^\gS_{cv}-\go),
 \end{equation}  
@@
 \int \frac{dk^3}{8\pi^3}
-\sum_{cv}\frac{1}{(\omega^\gs_{cv})^{2}}
+\sum_{cv}\frac{1}{(\omega^\gS_{cv})^{2}}
 \left[
-\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\gs,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
-+\frac{\mathrm{Re}\left[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
-\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gs_{cv}} 
-\right]\delta(\omega^\gs_{cv}-\omega),
+\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\gS,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
++\frac{\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gS_{cv}} 
+\right]\delta(\omega^\gS_{cv}-\omega),
 \end{equation}
@@
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\frac{4}{\omega^\gs_{cv}}
+\sum_{vc}\frac{4}{\omega^\gS_{cv}}
 \left[
 \sum_{v'\ne
-  v}\frac{\mathrm{Im}[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\{r^{\rmb}_{cv'}r^{\rmc}_{v'v}\}]}
-{2\go^\gs_{cv'}-\go^\gs_{cv}}
+  v}\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\{r^{\rmb}_{cv'}r^{\rmc}_{v'v}\}]}
+{2\go^\gS_{cv'}-\go^\gS_{cv}}
 - \sum_{c'\ne
-  c}\frac{\mathrm{Im}[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\{r^{\rmc}_{cc'}r^{\rmb}_{c'v}\}]}
-{2\go^\gs_{c'v}-\go^\gs_{cv}}
-\right]\gd(\go^\gs_{cv}-2\go),
+  c}\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\{r^{\rmc}_{cc'}r^{\rmb}_{c'v}\}]}
+{2\go^\gS_{c'v}-\go^\gS_{cv}}
+\right]\gd(\go^\gS_{cv}-2\go),
 \end{equation}
@@
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\frac{4}{(\omega^\gs_{cv})^{2}}
-\left[\mathrm{Re}\left[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
+\sum_{vc}\frac{4}{(\omega^\gS_{cv})^{2}}
+\left[\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
 \right\}\right] -
-\frac{2\mathrm{Re}\left[\mathcal{V}^{\gs,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
-\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gs_{cv}}\right]\delta(\omega^\gs_{cv}-2\omega)
+\frac{2\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gS_{cv}}\right]\delta(\omega^\gS_{cv}-2\omega)
 ,
@@
 We mention that if we take $\calc(z)=1$ through out, the layered
-matrix elements $\calbv^{\sigma}_{nm}$ become standard bulk-like
-$\bfv^{\sigma}_{nm}$ matrix elements. We mention that in this
+matrix elements $\calbv^{\gS}_{nm}$ become standard bulk-like
+$\bfv^{\gS}_{nm}$ matrix elements. We mention that in this
 case, Eq.~\eqref{chis} is equivalent to the expressions of
@@
 $\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
-$C(z)=1$ from the upper half containing the $2\times 1$ 
+$\calc(z)=1$ from the upper half containing the $2\times 1$ 
 surface reconstruction, as seen in Fig.~\ref{si2x1},
 and $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
-$C(z)=1$ through the full slab.
+$\calc(z)=1$ through the full slab.
 We show the results for this comparison in the remainder 
@@
 positions.  
-The structures were relaxed until the Cartesian force components were less and 5 meV/\AA. 
+The structures were relaxed until the Cartesian force components were less than 5 meV/\AA. 
 The geometry optimization for the clean surface gives
@@
 theoretical studies.\cite{caramellaPRB09,mendozaPRB06}
-%
-The unit cell has two atoms making for a computationally
-expensive calculation.
 The vacuum size is equivalent to one quarter the
@@
 wave-function tunneling from the contiguous surfaces of the full
-crystal formed by the repeated super cell scheme.\cite{mendozaPRB06}
+crystal formed by the repeated super-cell scheme.\cite{mendozaPRB06}
 
 Spin-orbit, local field, and electron-hole 
-attraction\cite{leitsmannPRB05,trollePRB14} 
+attraction\cite{beyond}
 effects on the SHG process are all neglected.
@@
 is in a clean $2\times 1$ reconstruction and the back 
-surface is an ideal terminated bulk H-saturated dangling bonds.
+surface is an ideal terminated bulk H-saturated dangling bonds (see Fig.~\ref{si2x1}).
 \label{fig1}} 
@@
 Eq.~\eqref{chis}.
-Computing second 
-order derivatives is required thus making the numerical procedure very 
+Computing second-order 
+derivatives is required thus making the numerical procedure very 
 time consuming. This adds significantly to the already lengthy time needed 
@@
 16, 24, 32, and 40 Si atomic layers. 
+In making the slabs larger, we add steps of 8 layers of bulk-like atomic positions. 
 We note that the response differs substantially 
@@
 effects of $\bfv^\nl$ or the scissors correction
-$\hbar\sigma$.   
+$\hbar\gD$.   
 For these results we chose
-$\hbar\sigma=0.5$ eV, that is the GW gap reported in
+$\hbar\gD=0.5$ eV, that is the GW gap reported in
 Refs.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
@@
 increases. We remark that 32 layers in the slab is more than enough 
-to confirm that the extraction of the surface second harmonic 
+to confirm that the extraction of the surface second-harmonic 
 susceptibility from the $2\times 1$ surface is readily possible 
@@
 2$\omega$ resonances start below $E_g/2$ where $E_g=0.53$ eV is the LDA
-band gap of this surface, and $E_g=1.035$ eV is the scissor shifted gap for
-$\hbar\sigma=0.5$ eV.
+band gap of this surface, and $E_g=1.03$ eV is the scissor shifted gap for
+$\hbar\gD=0.5$ eV.
 These resonances
@@
 In Fig.~\ref{fig3}
-We see that the effect of $\bfv^\nl$ reduces the value of   
-$|\chi^{xxx}_{2\times 1}|$,
-without changing the 
-spectral shape; this also can be confirmed from
+We see that the effect of $\bfv^\nl$
+ reduces the value of   
+$|\chi^{xxx}_{2\times 1}|$  
+by 15-20\% 
+showing the importance of this contribution for a correct calculation  
+of SSHG.  
+However, the inclusion of $\bfv^\nl$ does not
+changes the 
+spectral shape of $|\chi^{xxx}_{2\times 1}|$;
+ this also can be confirmed from
 the cases of zero scissors correction from Fig.~\ref{fig2}.
 %
-From Fig.~\ref{fig4}
-we note that the scissors correction 
-shifts the spectra to higher energies as expected. However, contrary 
-to the case of linear optics\cite{cabellosPRB09} the shift introduced
-by the scissors correction is not 
-rigid, as pointed out in Ref.~\onlinecite{nastosPRB05}.
- This is because the second harmonic optical response mixes 
-$1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
-for the non-rigid shift.
 \begin{figure}
@@
 \end{figure}
-\begin{figure}
-\centering 
-\includegraphics[scale=.8]{plots/fig4}
-\caption{(color on line) 
-$\chi^{xxx}_{2\times 1}$
-vs $\hbar\omega$ for a slab with 32 
-atomic Si layers plus one H layer, 
-for two different values of 
-the scissors correction $\hbar\sigma$. 
-\label{fig4}} 
-\end{figure}
-Also, when we compare 
-$|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$,
+
+To see the effect of the scissors correction, we take two different
+finite values for $\hbar\gD$. The first one
+with a value of $\hbar\gD=0.5$ eV, used in above results, 
+is the ``average'' GW gap taken from 
+Ref.~\onlinecite{rohlfingPRB95} 
+that is in agreement with Ref.~\onlinecite{garciaCPC01}. The second one
+with a value of $\hbar\gD=0.63$ eV is the ``average'' 
+gap taken from Ref.~\onlinecite{asahiPRB00}, 
+where more $\bfk$-points in the Brillouin zone were 
+used to calculate its GW value.
+From Fig.~\ref{fig4}
+we note that the scissors correction 
+shifts the spectra from it LDA value to higher energies as expected.
+ However, contrary 
+to the case of linear optics\cite{cabellosPRB09} the shift introduced 
+by the scissors correction is not 
+rigid, as pointed out in Ref.~\onlinecite{nastosPRB05}. 
+ This is because the second-harmonic optical response mixes 
+$1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
+for the non-rigid shift. 
+When we compare 
+$|\chi^{xxx}_{2\times 1}|$ for the two finite values of $\hbar\gD$,
 we see that the first two peaks are almost rigidly 
@@
 spectrum quite dramatically.
-We mention that
-the value of $\hbar\sigma=0.5$ eV, used above, 
-is the ``average'' GW gap taken from 
-Ref.~\onlinecite{rohlfingPRB95} 
-that is in agreement with Ref.~\onlinecite{garciaCPC01}. 
-The value of $\hbar\sigma=0.63$ eV is the ``average'' 
-gap taken from Ref.~\onlinecite{asahiPRB00}, 
-where more $\bfk$-points in the Brillouin zone were
-used to calculate its GW value.
+\begin{figure}
+\centering 
+\includegraphics[scale=.8]{plots/fig4}
+\caption{(color on line) 
+$\chi^{xxx}_{2\times 1}$
+vs $\hbar\omega$ for a slab with 32 
+atomic Si layers plus one H layer, 
+for two different values of 
+the scissors correction $\hbar\gD$. 
+\label{fig4}} 
+\end{figure}
 
@@
 
-We have presented a formulation to calculate the surface second harmonic
+We have presented a formulation to calculate the surface second-harmonic
 (SSH) susceptibility tensor $\bfgchi(-2\go;\go,\go)$, using the length gauge
@@
 improve even further on the similarity of the half-slab and full-slab results, we
-hava chosen a good compromise between accuracy and the burden and time
+have chosen a good compromise between accuracy and the burden and time
 of the computations. 
@@
 effects mentioned above in the calculation of  
-$\bfgchi(-2\go;\go,\go)$. The scissors correction shifts the spectrum to 
+$\bfgchi(-2\go;\go,\go)$. 
+The scissors correction shifts the spectrum to  
 higher energies though the shifting is not rigid and mixes the 
-$1\omega$ and $2\omega$ resonances, the cut function allows us
-to extract unequivocally $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$, 
-and the effects of the nonlocal part of the pseudopotentials
+$1\omega$ and $2\omega$ resonances,  
+and has a strong influence in the line-shape.
+ The cut function allows us  
+to extract unequivocally $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$.
+ The effects of the nonlocal part of the pseudopotentials  
+keeps the same line-shape of $|\chi^{xxx}_{2\times
+  1}(-2\go;\go,\go)|$,
+but  
 reduces the value of   
-$|\chi^{xxx}_{2\times 1}(-2\go;\go,\go)|$ keeping the same lineshape. 
+by 15-20\%. 
+Therefore,  
+the results show the importance of these three contributions for a correct calculation 
+of SSHG.
 The $xxx$ component of  
@@
 
+\section{Acknowledgments}
+
+B.S.M. acknowledges the Laboratoire des Solides Irradi\'es (Ecole
+Polytechnique, Palaiseau, France) for the support and hospitality
+during a sabbatical year. B.S.M. acknowledges partial support from
+CONACYT-M\'exico Grant 153930.
+
 %%%%%%%%%%%%%
@@
 and $L$
-is the length of the super cell. 
+is the length of the super-cell. 
 The reciprocal lattice vectors $\bfG$ are 
@@
 taking 
-$C(z)=1$ or $C_{nm}=\gd_{nm}$.
+$\calc(z)=1$ or $C_{nm}=\gd_{nm}$.
 
@@
 \Big(
-\bra{n\bfk} C(z) 
+\bra{n\bfk} \calc(z) 
 \ket{\bfk+\bfG}\bra{\bfk+\bfG}
@@
 \ket{\bfk+\bfG}\bra{\bfk+\bfG}
-C(z) \ket{m\bfk}
+\calc(z) \ket{m\bfk}
 \Big)
@@
 We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
-expressions by taking $C(z)=1$, which implies 
+expressions by taking $\calc(z)=1$, which implies 
 that $f(G_\perp)=\gd_{G_\perp 0}$, 
@@
 \begin{align}\label{choni.1}
-(\bfv^S_{nm})_{;\bfk}=i\gs f_{mn}
+(\bfv^S_{nm})_{;\bfk}=i\gD f_{mn}
 (\bfr_{nm})_{;\bfk}

### 2014-11-20 paper-v-0.tex
converted to real LaTeX, needs final polishing

@@
 \usepackage{graphicx}
-\allowdisplaybreaks[1]%biutiful equation breaker!!!
 \usepackage{ulem}
 \usepackage{subfigure}
-%\input{/Users/bms/util/definitions}
-\def\chon{red}
-\def\nicolas{red}
-\input{definitions}
 \usepackage{bm}
-%%%%*****************%%%% fine hyperef 
 \usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %\usepackage{showlabels}
 %\usepackage{showkeys}
+\allowdisplaybreaks[1]%biutiful equation breaker!!!
 
@@
 effects of the nonlocal part of the pseudopotentials 
-keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\go;\go,\go)|$, but 
+keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$, but 
 reduces its value
@@
 Later, Cabellos \textit{et al}.\cite{cabellosPRB09} elaborated a derivation 
-of the ``velocity gauge'' or ``$\bfA\cdot\bfv$'' gauge properly including the 
+of the ``velocity gauge'' or ``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the 
 scissors operator and proved gauge invariance with respect to the length 
@@
 \begin{equation}
-\bfJ=\langle{\mathbf{J}}\rangle=\frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}})
+\mathbf{J}=\langle{\mathbf{J}}\rangle=\frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}})
 ,
@@
 area of the unit cell. We calculate the polarization density 
-$\mathbf{P}$, related to $\bfJ$ by
-$\bfJ=d\mathbf{P}/dt$. For a 
+$\mathbf{P}$, related to $\mathbf{J}$ by
+$\mathbf{J}=d\mathbf{P}/dt$. For a 
 perturbing electromagnetic field, $\mathbf{E}(t)=
@@
 \begin{equation}
-\calp^a(2\omega)=\chi ^{abc}(-2\omega;\omega,\omega)E^{b}(\omega)E^{c}(\omega),  
+{\mathbf{\cal P}}^a(2\omega)=\chi ^{abc}(-2\omega;\omega,\omega)E^{b}(\omega)E^{c}(\omega),  
 \label{pshg}
@@
 \begin{equation}\label{h0}
-H^{\mathrm{LDA}}_{0}(\bfr,\bfp)
-=\frac{p^{2}}{2m_e}+V(\bfr,\bfp)
+H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})
+=\frac{p^{2}}{2m_e}+V(\mathbf{r},\mathbf{p})
 \end{equation}
@@
 \begin{align}\label{ps}
-V^{\mathrm{ps}}(\bfr,\bfp)=V(\bfr)+V^{\mathrm{nl}}(\bfr,\bfp)
+V^{\mathrm{ps}}(\mathbf{r},\mathbf{p})=V(\mathbf{r})+V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})
 ,
@@
 where 
-$V(\bfr)$ and $V^{\mathrm{nl}}(\bfr,\bfp)$  are the local and nonlocal parts, 
+$V(\mathbf{r})$ and $V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})$  are the local and nonlocal parts, 
  respectively.
 In case of
-a local potential, i.e. $V= V(\bfr)$, 
+a local potential, i.e. $V= V(\mathbf{r})$, 
 like that of all-electron schemes, 
@@
 \begin{equation*}
-H^\gS_{0}(\bfr,\bfp)=H^{\mathrm{LDA}}_{0}(\bfr,\bfp)+S(\bfr,\bfp)
+H^\Sigma_{0}(\mathbf{r},\mathbf{p})=H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})+S(\mathbf{r},\mathbf{p})
 \end{equation*}
@@
 \begin{equation}
-S(\bfr,\bfp)=\hbar \gD\sum_{n}\int d^{3}k(1-f_{n})
+S(\mathbf{r},\mathbf{p})=\hbar \Delta\sum_{n}\int d^{3}k(1-f_{n})
 |n\mathbf{k}\rangle\langle n\mathbf{k}|,
@@
 \end{equation}
-with $\hbar \gD$  the rigid ($\mathbf{k}$-independent) energy correction to be
+with $\hbar \Delta$  the rigid ($\mathbf{k}$-independent) energy correction to be
 applied. 
@@
 \begin{eqnarray*}
-H^{\mathrm{LDA}}_{0}(\bfr,\bfp)\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
+H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
 \label{hamils} \\
-H_{0}^\gS (\bfr,\bfp)\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\gS
+H_{0}^\Sigma (\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\Sigma
 (\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
@@
 where the scissor-shifted energies, 
-$\omega_{n}^\gS(\mathbf{k})$ , are given by
+$\omega_{n}^\Sigma(\mathbf{k})$ , are given by
 \begin{equation}
-\omega_{n}^\gS(\mathbf{k})=\omega^\lda_{n}(\mathbf{k})+(1-f_{n})\gD.
+\omega_{n}^\Sigma(\mathbf{k})=\omega^\mathrm{LDA}_{n}(\mathbf{k})+(1-f_{n})\Delta.
 \label{wese}
@@
 have the same eigenfunctions, where
-$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
-are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
-by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
+$\psi_{n\mathbf{k}}(\mathbf{r})=\langle\mathbf{r}|n\mathbf{k}\rangle=e^{i\mathbf{k}\cdot\mathbf{r}}u_{n\mathbf{k}}(\mathbf{r})$,
+are the real space representations of the Bloch states $| n\mathbf{k}\rangle$ labelled 
+by the band index $n$ and the crystal momentum $\mathbf{k}$, and $u_{n\mathbf{k}}(\mathbf{r})$
 is cell periodic.
@@
 treatment of the position operator
-$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$ has 
+$\mathbf{r}$ for extended Bloch states, $\langle\mathbf{r}|n\mathbf{k}\rangle$ has 
 been described in detail.\cite{adamsJCP53,blountSSP62} 
 In particular, Ref.~\onlinecite{aversaPRB95} develops
-the correct treatment of $\bfr$ for
+the correct treatment of $\mathbf{r}$ for
 the calculation of the nonlinear optical response.
-$\bfr$ is split into the {\it intraband} 
-($\bfr_i$) and {\it interband} ($\bfr_e$) parts, where 
-$\bfr=\bfr_i+\bfr_e$. 
-For $\bfr_e$ one uses
+$\mathbf{r}$ is split into the {\it intraband} 
+($\mathbf{r}_i$) and {\it interband} ($\mathbf{r}_e$) parts, where 
+$\mathbf{r}=\mathbf{r}_i+\mathbf{r}_e$. 
+For $\mathbf{r}_e$ one uses
 \begin{align}\label{rnminn}
-\bra{n\bfk}\bfr_e\ket{m\bfk'} &=
-(1-\gd_{nm})\gd(\bfk-\bfk')\bfgxi_{nm}(\bfk) 
-,
-\end{align}
-such that $\bfr_{e,nm}=0$ for $n=m$.
+\langle n\mathbf{k} | \mathbf{r}_e | m\mathbf{k}'\rangle &=
+(1-\delta_{nm})\delta(\mathbf{k}-\mathbf{k}')\boldsymbol{\xi}_{nm}(\mathbf{k})
+,
+\end{align}
+such that $\mathbf{r}_{e,nm}=0$ for $n=m$.
 From Eq.~\eqref{mv} with $H_0\to
-H^\gS_0$, we obtain
+H^\Sigma_0$, we obtain
 \begin{align}\label{pmnrmn}
-\bfr_{e,nm}(\bfk) =
-\bfgxi_{nm}(\bfk)\equiv 
-\bfr_{nm}(\bfk) 
+\mathbf{r}_{e,nm}(\mathbf{k}) =
+\boldsymbol{\xi}_{nm}(\mathbf{k})\equiv 
+\mathbf{r}_{nm}(\mathbf{k}) 
 &=
-\frac{\bfv^\gS_{nm}(\bfk)}{i\go^\gS_{nm}(\bfk)}
+\frac{\mathbf{v}^\Sigma_{nm}(\mathbf{k})}{i\omega^\Sigma_{nm}(\mathbf{k})}
 \quad\quad n\notin D_m 
@@
 where we defined
-$\go^\gS_{nm}(\bfk)\equiv\go^\gS_n(\bfk)-\go^\gS_m(\bfk)$, and
+$\omega^\Sigma_{nm}(\mathbf{k})\equiv\omega^\Sigma_n(\mathbf{k})-\omega^\Sigma_m(\mathbf{k})$, and
 $D_m$ are all the possible degenerate $m$-states. 
 %We recognize that for each instance in which 
-%$n\notin D_m$ or $n\ne m$, $\bfr_{nm}(\bfk)$ are interband matrix
+%$n\notin D_m$ or $n\ne m$, $\mathbf{r}_{nm}(\mathbf{k})$ are interband matrix
 %elements.
-For the intraband part, $\bfr_i$ only appears in
+For the intraband part, $\mathbf{r}_i$ only appears in
 commutators during the derivation of
@@
 \begin{equation}\label{conmri3n}
-\bra{n\bfk}[\bfr_i,\calo]\ket{m\bfk'}
-=i\gd(\bfk-\bfk')(\calo_{nm})_{;\bfk}
+\langle n\mathbf{k} | [\mathbf{r}_i,{\cal O}] | m\mathbf{k}'\rangle
+=i\delta(\mathbf{k}-\mathbf{k}')({\cal O}_{nm})_{;\mathbf{k}}
 ,
@@
 \begin{equation}\label{gendevnn}
-(\calo_{nm})_{;\bfk}=
-\nabla_\bfk 
-\calo_{n m}(\bfk) 
+({\cal O}_{nm})_{;\mathbf{k}}=
+\nabla_\mathbf{k} 
+{\cal O}_{n m}(\mathbf{k}) 
 - 
 i 
-\calo_{nm}(\bfk) 
+{\cal O}_{nm}(\mathbf{k}) 
 \left(
-\bfgxi_{nn}(\bfk) 
+\boldsymbol{\xi}_{nn}(\mathbf{k}) 
 -
-\bfgxi_{mm}(\bfk) 
+\boldsymbol{\xi}_{mm}(\mathbf{k}) 
 \right) 
@@
 the generalized derivative 
-of the operator $\calo$. 
-The vectors $\bfgxi_{nn}(\bfk)$ are defined in 
+of the operator ${\cal O}$. 
+The vectors $\boldsymbol{\xi}_{nn}(\mathbf{k})$ are defined in 
 Ref.~\onlinecite{aversaPRB95} though they do not need to be 
@@
 we derive a key result for the length gauge formulation. 
-Again, using $H^\gS_0$ in
+Again, using $H^\Sigma_0$ in
 Eq.~\eqref{mv} we obtain
 \begin{align}\label{vop2}
-\bfv^\gS&=
-\bfv 
-+
-\bfv^\nl 
-+\bfv^S
+\mathbf{v}^\Sigma&=
+\mathbf{v} 
++
+\mathbf{v}^\mathrm{nl} 
++\mathbf{v}^S
 =
-\bfv^\lda 
-+\bfv^S 
+\mathbf{v}^\mathrm{LDA} 
++\mathbf{v}^S 
 ,
@@
 \begin{align}\label{ve}
-\bfv 
-&=\frac{\bfp}{m_e},
+\mathbf{v} 
+&=\frac{\mathbf{p}}{m_e},
 \\\label{vnl}
-\bfv^\nl 
+\mathbf{v}^\mathrm{nl} 
 &=
-\frac{1}{i\hbar}[\bfr,V^\nl],
+\frac{1}{i\hbar}[\mathbf{r},V^\mathrm{nl}],
 \\\label{vs}
-\bfv^S
+\mathbf{v}^S
 &=
-\frac{1}{i\hbar}[\bfr, S],
+\frac{1}{i\hbar}[\mathbf{r}, S],
 \\\label{vlda}
-\bfv^\lda 
+\mathbf{v}^\mathrm{LDA} 
 &=
-\bfv 
-+\bfv^\nl
+\mathbf{v} 
++\mathbf{v}^\mathrm{nl}
 ,
@@
 \end{subequations}
-with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator, 
-and using $[r^a,p^b]=i\hbar\gd_{ab}$, where $\gd_{ab}$ is the Kronecker delta.
+with $\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ the canonical momentum operator, 
+and using $[r^a,p^b]=i\hbar\delta_{ab}$, where $\delta_{ab}$ is the Kronecker delta.
 Using Eq.~\eqref{hats} we obtain 
 \begin{equation}\label{chon.2} 
-\bfv^S_{nm}=i\gD f_{mn}\bfr_{nm},
+\mathbf{v}^S_{nm}=i\Delta f_{mn}\mathbf{r}_{nm},
 \end{equation}
 with $f_{nm}\equiv f_n-f_m$,
-where we see that $\bfv^S_{nn}=0$. It follows that
+where we see that $\mathbf{v}^S_{nn}=0$. It follows that
 \begin{align}\label{chon.9}
-\bfv^\gS_{nm}(\bfk) 
-=\frac{\go^\gS_{nm}(\bfk)}{\go^\lda_{nm}(\bfk)}
-\bfv^\lda_{nm}(\bfk) 
+\mathbf{v}^\Sigma_{nm}(\mathbf{k}) 
+=\frac{\omega^\Sigma_{nm}(\mathbf{k})}{\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
+\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k}) 
 ,
@@
  \begin{align}\label{chon.8}
-\frac{\bfv^\gS_{nm}(\bfk)}{\go^\gS_{nm}(\bfk)}
+\frac{\mathbf{v}^\Sigma_{nm}(\mathbf{k})}{\omega^\Sigma_{nm}(\mathbf{k})}
 &=
-\frac{\bfv^\lda_{nm}(\bfk)}{\go^\lda_{nm}(\bfk)}
+\frac{\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k})}{\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
 ,
@@
 \begin{align}\label{chon.10}
-\bfr_{nm}(\bfk) 
+\mathbf{r}_{nm}(\mathbf{k}) 
 =
-\frac{\bfv^\gS_{nm}(\bfk)}{i\go^\gS_{nm}(\bfk)}
+\frac{\mathbf{v}^\Sigma_{nm}(\mathbf{k})}{i\omega^\Sigma_{nm}(\mathbf{k})}
 =
-\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
+\frac{\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k})}{i\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
 \quad\quad n\notin D_m 
@@
 The matrix elements 
-of $\bfr_{nm}(\bfk)$ are identical using either
+of $\mathbf{r}_{nm}(\mathbf{k})$ are identical using either
 the LDA or scissored
@@
 Of course, it is more convenient to calculate them
-through $\bfv^\lda_{nm}(\bfk)$ which 
+through $\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k})$ which 
 includes only the contribution of 
-$\bfv^\nl_{nm}(\bfk)$. These can be readily
+$\mathbf{v}^\mathrm{nl}_{nm}(\mathbf{k})$. These can be readily
 calculated
@@
 we need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
-$\omega_{n}^{\gS}$ at the end of the derivation
+$\omega_{n}^{\Sigma}$ at the end of the derivation
 to obtain the scissored results for any 
@@
 Eq.~\eqref{rde}, we obtain 
-$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\gS_{nm}-\tilde\omega)t}$,
+$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\Sigma_{nm}-\tilde\omega)t}$,
 with 
 \begin{equation}\label{j.1}
-B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\gS_{nm}-\tilde\omega},
+B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\Sigma_{nm}-\tilde\omega},
 \end{equation}
@@
 \begin{eqnarray}\label{j.2}
-(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\gS_{nm}-2\tilde\omega}\bigg[
+(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
 i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
 \Big)  
--(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\gS_{nm}-2\tilde\omega)t}
+-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\Sigma_{nm}-2\tilde\omega)t}
 .
@@
 other. This is due to the $f_{nm}$ factor in Eq.~\eqref{j.1}, 
-and therefore $B^a_{nn}=0$. The dependence on $\bfk$ 
+and therefore $B^a_{nn}=0$. The dependence on $\mathbf{k}$ 
 of all quantities is implicitly understood from 
@@
 centrosymmetric. This would imply that 
-$\chi^{\mathrm{slab},\rma\rmb\rmc}=0$ so we must
+$\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}=0$ so we must
 find a scheme 
-in order to have a finite $\chi^{\rma\rmb\rmc}$ representative of the
+in order to have a finite $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$ representative of the
 surface. Even if the front and back surfaces of the slab 
 are different, breaking the centrosymmetry and therefore giving an
-overall $\chi^{\mathrm{slab},\rma\rmb\rmc}\ne 0$; we still
-need a procedure to extract the front surface $\chi^{f,\rma\rmb\rmc}$
-and the back surface $\chi^{b,\rma\rmb\rmc}$ from the slab
+overall $\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}\ne 0$; we still
+need a procedure to extract the front surface $\chi^{f,\mathrm{a}\mathrm{b}\mathrm{c}}$
+and the back surface $\chi^{b,\mathrm{a}\mathrm{b}\mathrm{c}}$ from the slab
 susceptibility 
-$\chi^{\mathrm{slab},\rma\rmb\rmc}=\chi^{f,\rma\rmb\rmc}-\chi^{b,\rma\rmb\rmc}$.
-We have omitted the frequency arguments of $\chi^{\rma\rmb\rmc}$ for 
+$\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}=\chi^{f,\mathrm{a}\mathrm{b}\mathrm{c}}-\chi^{b,\mathrm{a}\mathrm{b}\mathrm{c}}$.
+We have omitted the frequency arguments of $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$ for 
 convenience of notation.
@@
 A convenient way to accomplish the separation of the SH signal of
-either surface is to introduce a ``cut function'', $\calc(z)$, which is 
+either surface is to introduce a ``cut function'', ${\mathbf{\cal C}}(z)$, which is 
 usually taken to be unity over one half of the slab and zero over 
 the other half.\cite{reiningPRB94}
-In this case $\calc(z)$ will give the contribution of the 
-side of the slab for which $\calc(z)=1$. 
+In this case ${\mathbf{\cal C}}(z)$ will give the contribution of the 
+side of the slab for which ${\mathbf{\cal C}}(z)=1$. 
 As was done for the linear response,\cite{mendozaPRB06}
 we can generalize this 
-simple choice for $\calc(z)$ by a top-hat cut function
-$\calc^\ell(z)$ that selects a given layer,
+simple choice for ${\mathbf{\cal C}}(z)$ by a top-hat cut function
+${\mathbf{\cal C}}^\ell(z)$ that selects a given layer,
 \begin{equation}
 \label{sz}
-\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{b})  
+{\mathbf{\cal C}}^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{b})  
             \Theta(z_\ell-z+\Delta_\ell^f),
@@
 To introduce the
-cut function $\calc(z)$ in
-the calculation of $\chi^{\rma\rmb\rmc}$, we start from 
+cut function ${\mathbf{\cal C}}(z)$ in
+the calculation of $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$, we start from 
 the operator for the electron current,
-$\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gS\ket{\bfr}\bra{\bfr}
-+\ket{\bfr}\bra{\bfr}\bfv^\gS\right)$, that leads to
+$\mathbf{j}(\mathbf{r})=\frac{e}{2}\left(\mathbf{v}^\Sigma | \mathbf{r}\rangle\langle\mathbf{r} |
++ | \mathbf{r}\rangle\langle\mathbf{r} | \mathbf{v}^\Sigma\right)$, that leads to
 \begin{align}\label{jmic}
-\bfj^{(N)}(\bfr,t)=\mathrm{Tr}(\bfj(\bfr)\rho^{(N)}(t))
+\mathbf{j}^{(N)}(\mathbf{r},t)=\mathrm{Tr}(\mathbf{j}(\mathbf{r})\rho^{(N)}(t))
 =
@@
 \sum_{nm}
-\rho^{(N)}_{nm}(\bfk;t)\bfj_{mn}(\bfk;\bfr)
+\rho^{(N)}_{nm}(\mathbf{k};t)\mathbf{j}_{mn}(\mathbf{k};\mathbf{r})
 ,
@@
 \begin{equation}\label{jmic3}
-\bfj_{mn}(\bfk;\bfr)=
+\mathbf{j}_{mn}(\mathbf{k};\mathbf{r})=
 \frac{e}{2}
 \left(
-\bra{m\bfk}\bfv^\gS\ket{\bfr}\braket{\bfr}{n\bfk}
-+
-\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gS\ket{n\bfk}
+\langle m\mathbf{k} | \mathbf{v}^\Sigma | \mathbf{r}\rangle \langle\mathbf{r}|n\mathbf{k}\rangle
++
+\langle m\mathbf{k}|\mathbf{r}\rangle\langle\mathbf{r} | \mathbf{v}^\Sigma | n\mathbf{k}\rangle
 \right),
@@
 are the matrix elements of the microscopic current operator.
-% and we have used the fact that the matrix elements between states $\ket{n\bfk}$
-% are diagonal in $\bfk$, i.e. proportional to $\gd(\bfk-\bfk')$.
-Integrating the microscopic current $\mbf{j}(\mbf{r},t)$ over
-the entire slab gives the averaged microscopic current density, $\bfJ(t)$. 
+% and we have used the fact that the matrix elements between states $\ket{n\mathbf{k}}$
+% are diagonal in $\mathbf{k}$, i.e. proportional to $\delta(\mathbf{k}-\mathbf{k}')$.
+Integrating the microscopic current $\mathbf{j}(\mathbf{r},t)$ over
+the entire slab gives the averaged microscopic current density, $\mathbf{J}(t)$. 
 If we want the contribution from only one region of the unit cell 
@@
 \begin{equation}\label{jsz}
-\frac{1}{A}\int d^3r\, \calc(z)\, 
+\frac{1}{A}\int d^3r\, {\mathbf{\cal C}}(z)\, 
 \mathbf{J} (\mathbf{r},t)
- \equiv \calbj^{(N)}(t),
-\end{equation}
-where $\calbj^{(N)}(t)$ is the $N$-th order current induced in the
-region specified by $\calc(z)$.
+ \equiv \boldsymbol{\mathcal{\cal J}}^{(N)}(t),
+\end{equation}
+where $\boldsymbol{\mathcal{\cal J}}^{(N)}(t)$ is the $N$-th order current induced in the
+region specified by ${\mathbf{\cal C}}(z)$.
 Therefore we define
 \begin{equation}\label{vcal}
-e{\boldsymbol{\mathcal{V}}}^{\gS}_{mn}(\mathbf{k})
+e{\boldsymbol{\mathcal{V}}}^{\Sigma}_{mn}(\mathbf{k})
 \equiv
-\int d^3r\, \calc(z)\,\bfj_{mn}({\bfk};\bfr),
+\int d^3r\, {\mathbf{\cal C}}(z)\,\mathbf{j}_{mn}({\mathbf{k}};\mathbf{r}),
 \end{equation}
@@
 \begin{equation}\label{jmac2}
-\calbj^{(N)}(2\go)=\frac{e}{A}
+\boldsymbol{\mathcal{\cal J}}^{(N)}(2\omega)=\frac{e}{A}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\calbv^{\gS}_{mn}(\mathbf{k}) 
-\rho^{(N)}_{nm}(\bfk;2\go) 
+\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{mn}(\mathbf{k}) 
+\rho^{(N)}_{nm}(\mathbf{k};2\omega) 
 , 
@@
 \begin{align}\label{intj}
-{\boldsymbol{\mathcal{V}}}^{\gS}_{mn}({\mathbf k})
+{\boldsymbol{\mathcal{V}}}^{\Sigma}_{mn}({\mathbf k})
 &=
@@
 \int \mathrm{d}^3 r\,
- \calc(z)
+ {\mathbf{\cal C}}(z)
 \bigg[
-\langle m\mathbf{k}|\mathbf{v}^\gS | \mathbf{r}\rangle
+\langle m\mathbf{k}|\mathbf{v}^\Sigma | \mathbf{r}\rangle
 \langle \mathbf{r} | n \mathbf k \rangle +
 \langle m\mathbf{k} | \mathbf{r}\rangle
-\langle \mathbf{r} | \mathbf{v}^\gS | n \mathbf k \rangle\bigg]
+\langle \mathbf{r} | \mathbf{v}^\Sigma | n \mathbf k \rangle\bigg]
 \nonumber\\
@@
 \int \mathrm{d}^3 r\,
- \calc(z)
+ {\mathbf{\cal C}}(z)
  \bigg[
 \psi_{n\mathbf{k}}(\mathbf{r})
-\bfv^{\gS *}\psi^*_{m\mathbf{k}}(\mathbf{r})
-+ 
-\psi^*_{m\mathbf{k}}(\mathbf{r})\bfv^\gS
+\mathbf{v}^{\Sigma *}\psi^*_{m\mathbf{k}}(\mathbf{r})
++ 
+\psi^*_{m\mathbf{k}}(\mathbf{r})\mathbf{v}^\Sigma
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\left[\frac{\calc(z) \bfv^\gS +
-\bfv^\gS \calc(z)}{2}\right]
+\left[\frac{{\mathbf{\cal C}}(z) \mathbf{v}^\Sigma +
+\mathbf{v}^\Sigma {\mathbf{\cal C}}(z)}{2}\right]
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\calbv^{\gS}
+\boldsymbol{\mathcal{\cal V}}^{\Sigma}
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 %Here an integration by parts is performed on the third term of the
-%right-hand; since the $e^{-i\bfk\cdot\bfr}\psi_{n\bfk}(\bfr)$
+%right-hand; since the $e^{-i\mathbf{k}\cdot\mathbf{r}}\psi_{n\mathbf{k}}(\mathbf{r})$
 %are periodic over the unit cell, the surface term vanishes. 
-where, we used the hermitian property of $\bfv^\gS$ and defined
+where, we used the hermitian property of $\mathbf{v}^\Sigma$ and defined
 \begin{align}\label{nl.4}
-\calbv^{\gS}
+\boldsymbol{\mathcal{\cal V}}^{\Sigma}
 =
-\frac{\calc(z) \bfv^\gS +
-\bfv^\gS \calc(z)}{2}
+\frac{{\mathbf{\cal C}}(z) \mathbf{v}^\Sigma +
+\mathbf{v}^\Sigma {\mathbf{\cal C}}(z)}{2}
 .
 \end{align} 
-%where the superscript $\ell$ is inherited from $\calc(z)$.
+%where the superscript $\ell$ is inherited from ${\mathbf{\cal C}}(z)$.
 % and we
@@
 \begin{align}\label{vcali}
-\bfV \to \calbv=\frac{\calc(z) \bfV +
-\bfV \calc(z)}{2}
+\mathbf{V} \to \boldsymbol{\mathcal{\cal V}}=\frac{{\mathbf{\cal C}}(z) \mathbf{V} +
+\mathbf{V} {\mathbf{\cal C}}(z)}{2}
 ,
@@
 is all that is needed to change any of the
-electron velocity operators $\bfV$ to the new velocity
-operator $\calbv$ that implicitly takes into account the
-contribution of the region of the slab given by $\calc(z)$.
+electron velocity operators $\mathbf{V}$ to the new velocity
+operator $\boldsymbol{\mathcal{\cal V}}$ that implicitly takes into account the
+contribution of the region of the slab given by ${\mathbf{\cal C}}(z)$.
 We note that this modified operator is hermitian as it must.\cite{note2}
-The operator $\bfV$ could be any of those given by Eq.~\eqref{vop2},
+The operator $\mathbf{V}$ could be any of those given by Eq.~\eqref{vop2},
 thus
 \begin{align}\label{vopii}
-\calbv^{\gS}
+\boldsymbol{\mathcal{\cal V}}^{\Sigma}
 &=
-\calbv^{\lda}
-+
-\calbv^{S}
+\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}
++
+\boldsymbol{\mathcal{\cal V}}^{S}
 \nonumber\\
-\calbv^{\lda}
+\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}
 &=
-\calbv
-+
-\calbv^{\nl}
+\boldsymbol{\mathcal{\cal V}}
++
+\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}
 .
@@
 To calculate
-$\calbv^{\gS}_{nm}(\bfk)$ 
+$\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{nm}(\mathbf{k})$ 
 we calculate the matrix elements of 
-$\calbv^{\lda}$ and $\calbv^{S}$
+$\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}$ and $\boldsymbol{\mathcal{\cal V}}^{S}$
  (separately)
@@
 The layer-by-layer analysis of Refs. \onlinecite{hoganPRB03,castilloPRB03,mottaCMS10}  
-used $\calc(z)$
+used ${\mathbf{\cal C}}(z)$
 %limiting the current response  
@@
 To limit the SHG response to one surface, Eq.~\eqref{vcali} 
-for $\calbv$ was proposed in 
+for $\boldsymbol{\mathcal{\cal V}}$ was proposed in 
 Ref.~\onlinecite{reiningPRB94} and later used in Refs.
@@
  and \onlinecite{mejiaRMF04}. 
-In this article, for the first time, we formally introduce the cut function $\calc(z)$ 
+In this article, for the first time, we formally introduce the cut function ${\mathbf{\cal C}}(z)$ 
 for the second-harmonic optical response of semiconductor surfaces,
@@
 Using
-$\calbj=d\calbp/dt$ 
+$\boldsymbol{\mathcal{\cal J}}=d\boldsymbol{\mathcal{\cal P}}/dt$ 
 and Eq.~\eqref{jmac2} 
@@
 \begin{equation}\label{Pjikn}
-\calbp^{(2)}(2\go)
-=\frac{ie}{2A\got}
+\boldsymbol{\mathcal{\cal P}}^{(2)}(2\omega)
+=\frac{ie}{2A\tilde\omega}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\calbv^{\gS}_{mn}(\mathbf{k})
-\rho^{(2)}_{nm}(\bfk;2\go)
+\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{mn}(\mathbf{k})
+\rho^{(2)}_{nm}(\mathbf{k};2\omega)
 ,
@@
 \begin{align}\label{Pjikn2}
-\chi^{\rma\rmb\rmc}(-2\go;\go,\go) 
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega) 
 &=
-\frac{e^2}{2A\hbar\got}
+\frac{e^2}{2A\hbar\tilde\omega}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\frac{\calv^{\gS,\rma}_{mn}(\mathbf{k})}
-{\go^\gS_{nm\bfk}-2\got}
+\frac{{\mathbf{\cal V}}^{\Sigma,\mathrm{a}}_{mn}(\mathbf{k})}
+{\omega^\Sigma_{nm\mathbf{k}}-2\tilde\omega}
 \bigg[
--(B_{nm}^{\rmc}(\bfk,\go))_{;k^{\rmb}}
+-(B_{nm}^{\mathrm{c}}(\mathbf{k},\omega))_{;k^{\mathrm{b}}}
 \nonumber \\
 &
-+i\sum_q\left(r_{nq}^{\rmb}B_{qm}^{\rmc}(\bfk,\go) -
-  B_{nq}^{\rmc}(\bfk,\go) 
-  r_{qm}^{\rmb}\right)
++i\sum_q\left(r_{nq}^{\mathrm{b}}B_{qm}^{\mathrm{c}}(\mathbf{k},\omega) -
+  B_{nq}^{\mathrm{c}}(\mathbf{k},\omega) 
+  r_{qm}^{\mathrm{b}}\right)
 \bigg]
@@
 which gives the susceptibility 
-$\chi^{\rma\rmb\rmc}(-2\go;\go,\go)$ 
-of the layers of the slab specified by $\calc(z)$. 
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ 
+of the layers of the slab specified by ${\mathbf{\cal C}}(z)$. 
 We mention that the units of 
-$\chi^{\rma\rmb\rmc}(-2\go;\go,\go)$
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$
 are m$^2$/V, as they should be for a surface SH susceptibility.
@@
 \begin{equation}\label{chii}
-\chi^{\rma\rmb\rmc}_i (-2\go;\go,\go)
-=-\frac{e^3}{A\hbar^22\got}
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_i (-2\omega;\omega,\omega)
+=-\frac{e^3}{A\hbar^22\tilde\omega}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\frac{\calv_{mn}^{\gS,\rma}}{\go^\gS_{nm}-2\got}
-\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gS_{nm}-\got}\right)_{;k^{\rmc}}
+\frac{{\mathbf{\cal V}}_{mn}^{\Sigma,\mathrm{a}}}{\omega^\Sigma_{nm}-2\tilde\omega}
+\left(\frac{f_{mn}r_{nm}^{\mathrm{b}}}{\omega^\Sigma_{nm}-\tilde\omega}\right)_{;k^{\mathrm{c}}}
 ,
@@
 \begin{equation}\label{chie}
-\chi^{\rma\rmb\rmc}_e (-2\go;\go,\go)
-=\frac{ie^3}{A\hbar^22\got}
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_e (-2\omega;\omega,\omega)
+=\frac{ie^3}{A\hbar^22\tilde\omega}
 \int \frac{dk^3}{8\pi^3}
 \sum_{qmn}
-\frac{\calv_{mn}^{\gS,\rma}}{\go^\gS_{nm}-2\got}
+\frac{{\mathbf{\cal V}}_{mn}^{\Sigma,\mathrm{a}}}{\omega^\Sigma_{nm}-2\tilde\omega}
 \left(
-\frac{r_{nq}^{\rmc} r_{qm}^{\rmb} 
-f_{mq}}{\go^\gS_{qm}-\got}
--\frac{r_{nq}^{\rmb} r_{qm}^{\rmc} 
-f_{qn}}{\go^\gS_{nq}-\got}
+\frac{r_{nq}^{\mathrm{c}} r_{qm}^{\mathrm{b}} 
+f_{mq}}{\omega^\Sigma_{qm}-\tilde\omega}
+-\frac{r_{nq}^{\mathrm{b}} r_{qm}^{\mathrm{c}} 
+f_{qn}}{\omega^\Sigma_{nq}-\tilde\omega}
 \right),
@@
 \begin{equation}\label{gene2}
-\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gS_{nm}-\got}\right)_{;k^{\rmc}}=
-\frac{f_{mn}}{\go^\gS_{nm}-\got}\left(r_{nm}^\rmb\right)_{;k^{\rmc}}
--\frac{f_{mn}r_{nm}^{\rmb}\gD_{nm}^\rmc}{(\go^\gS_{nm}-\got)^2}
-,
-\end{equation}
-where substituting $H^\gS_0$ 
+\left(\frac{f_{mn}r_{nm}^{\mathrm{b}}}{\omega^\Sigma_{nm}-\tilde\omega}\right)_{;k^{\mathrm{c}}}=
+\frac{f_{mn}}{\omega^\Sigma_{nm}-\tilde\omega}\left(r_{nm}^\mathrm{b}\right)_{;k^{\mathrm{c}}}
+-\frac{f_{mn}r_{nm}^{\mathrm{b}}\Delta_{nm}^\mathrm{c}}{(\omega^\Sigma_{nm}-\tilde\omega)^2}
+,
+\end{equation}
+where substituting $H^\Sigma_0$ 
 into Eq.~\eqref{conmri3n} and then
@@
 \begin{align}\label{eli.13}
-\left(\go^\gS_{nm}\right)_{;k^{\rma}}
+\left(\omega^\Sigma_{nm}\right)_{;k^{\mathrm{a}}}
 =
-\left(\go^\lda_{nm}\right)_{;k^{\rma}}
+\left(\omega^\mathrm{LDA}_{nm}\right)_{;k^{\mathrm{a}}}
 = 
-v_{nn}^{\lda,\rma}-v_{mm}^{\lda,\rma}\equiv\gD_{nm}^{\rma}
+v_{nn}^{\mathrm{LDA},\mathrm{a}}-v_{mm}^{\mathrm{LDA},\mathrm{a}}\equiv\Delta_{nm}^{\mathrm{a}}
 .
 \end{align} 
-The apparent divergence as $\got\to 0$
+The apparent divergence as $\tilde\omega\to 0$
 in Eqs. \eqref{chii} and \eqref{chie},  
 is removed  by
- a partial fraction expansion over $\got$. 
+ a partial fraction expansion over $\tilde\omega$. 
 Using time-reversal invariance, an integration by parts to 
@@
 \begin{equation}\label{calvimchiewn}
-\mathrm{Im}[\chi^{\rma\rmb\rmc}_{e,\go}]= 
+\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}]= 
 \frac{\pi |e|^3}{2\hbar^2}
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\gS_{cv}}
+\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\Sigma_{cv}}
 \left[
-\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{qc}\{r^{\rmb}_{cv}r^{\rmc}_{vq}\}]}
-{(2\go^\gS_{cv}-\go^\gS_{cq})} 
--\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vq}\{r^{\rmc}_{qc}r^{\rmb}_{cv}\}]}
-{(2\go^\gS_{cv}-\go^\gS_{qv})}
-\right]\gd(\go^\gS_{cv}-\go),
+\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{qc}\{r^{\mathrm{b}}_{cv}r^{\mathrm{c}}_{vq}\}]}
+{(2\omega^\Sigma_{cv}-\omega^\Sigma_{cq})} 
+-\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vq}\{r^{\mathrm{c}}_{qc}r^{\mathrm{b}}_{cv}\}]}
+{(2\omega^\Sigma_{cv}-\omega^\Sigma_{qv})}
+\right]\delta(\omega^\Sigma_{cv}-\omega),
 \end{equation}  
 \begin{equation}\label{calvimchiwn}
-\mathrm{Im}[\chi^{\rma\rmb\rmc}_{i,\go}]= 
+\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}]= 
 \frac{\pi\vert e\vert^3}{2\hbar^2}
 \int \frac{dk^3}{8\pi^3}
-\sum_{cv}\frac{1}{(\omega^\gS_{cv})^{2}}
+\sum_{cv}\frac{1}{(\omega^\Sigma_{cv})^{2}}
 \left[
-\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\gS,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
-+\frac{\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
-\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gS_{cv}} 
-\right]\delta(\omega^\gS_{cv}-\omega),
+\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
++\frac{\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\Sigma_{cv}} 
+\right]\delta(\omega^\Sigma_{cv}-\omega),
 \end{equation}
 \begin{equation}\label{calvimchie2wn}
-\mathrm{Im}[\chi^{\rma\rmb\rmc}_{e,2\go}]= 
+\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}]= 
 -\frac{\pi |e|^3}{2\hbar^2}
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\frac{4}{\omega^\gS_{cv}}
+\sum_{vc}\frac{4}{\omega^\Sigma_{cv}}
 \left[
 \sum_{v'\ne
-  v}\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\{r^{\rmb}_{cv'}r^{\rmc}_{v'v}\}]}
-{2\go^\gS_{cv'}-\go^\gS_{cv}}
+  v}\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\{r^{\mathrm{b}}_{cv'}r^{\mathrm{c}}_{v'v}\}]}
+{2\omega^\Sigma_{cv'}-\omega^\Sigma_{cv}}
 - \sum_{c'\ne
-  c}\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\{r^{\rmc}_{cc'}r^{\rmb}_{c'v}\}]}
-{2\go^\gS_{c'v}-\go^\gS_{cv}}
-\right]\gd(\go^\gS_{cv}-2\go),
+  c}\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\{r^{\mathrm{c}}_{cc'}r^{\mathrm{b}}_{c'v}\}]}
+{2\omega^\Sigma_{c'v}-\omega^\Sigma_{cv}}
+\right]\delta(\omega^\Sigma_{cv}-2\omega),
 \end{equation}
 \begin{equation}\label{calvimchi2wn}
-\mathrm{Im}[\chi^{\rma\rmb\rmc}_{i,2\go}]= 
+\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}]= 
  \frac{\pi \vert
@@
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\frac{4}{(\omega^\gS_{cv})^{2}}
-\left[\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
+\sum_{vc}\frac{4}{(\omega^\Sigma_{cv})^{2}}
+\left[\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
 \right\}\right] -
-\frac{2\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
-\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gS_{cv}}\right]\delta(\omega^\gS_{cv}-2\omega)
+\frac{2\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\Sigma_{cv}}\right]\delta(\omega^\Sigma_{cv}-2\omega)
 ,
@@
 \end{subequations}
-where we have split the interband and intraband $1\go$ and $2\go$
-contributions and supressed the $\go$ arguments for 
+where we have split the interband and intraband $1\omega$ and $2\omega$
+contributions and supressed the $\omega$ arguments for 
 convenience of notation.
@@
 a Kramers-Kronig transformation\cite{nicolasPRB14} and
-$\chi^{\rma\rmb\rmc}=
-\chi^{\rma\rmb\rmc}_{e,\go} 
-+\chi^{\rma\rmb\rmc}_{e,2\go}
-+\chi^{\rma\rmb\rmc}_{i,\go} 
-+\chi^{\rma\rmb\rmc}_{i,2\go}
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}=
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega} 
++\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}
++\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega} 
++\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}
 $.
 To fulfill the required intrinsic permutation symmetry, %\cite{rashkeevPRB98} 
-the $\{\}$ notation symmetrizes the $\rmb\rmc$ Cartesian indices, i.e. 
-$\{u^{\rmb}s^{\rmc}\}=(u^{\rmb}s^{\rmc}+u^{\rmc}s^{\rmb})/2$,
+the $\{\}$ notation symmetrizes the $\mathrm{b}\mathrm{c}$ Cartesian indices, i.e. 
+$\{u^{\mathrm{b}}s^{\mathrm{c}}\}=(u^{\mathrm{b}}s^{\mathrm{c}}+u^{\mathrm{c}}s^{\mathrm{b}})/2$,
 and thus
-$\chi^{\rma\rmb\rmc}=\chi^{\rma\rmc\rmb}$.
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}=\chi^{\mathrm{a}\mathrm{c}\mathrm{b}}$.
 The various quantities involved in Eqs.~\eqref{chis} are given in
 the Appendix \ref{appe}. 
-We mention that if we take $\calc(z)=1$ through out, the layered
-matrix elements $\calbv^{\gS}_{nm}$ become standard bulk-like
-$\bfv^{\gS}_{nm}$ matrix elements. We mention that in this
+We mention that if we take ${\mathbf{\cal C}}(z)=1$ through out, the layered
+matrix elements $\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{nm}$ become standard bulk-like
+$\mathbf{v}^{\Sigma}_{nm}$ matrix elements. We mention that in this
 case, Eq.~\eqref{chis} is equivalent to the expressions of
@@
 \begin{equation}\label{chiijksur}
-\bfgchi(-2\go;\go,\go)
-= \sum_{\{\ell\}}\bfgchi^\ell(-2\go;\go,\go),
+\boldsymbol{\chi}(-2\omega;\omega,\omega)
+= \sum_{\{\ell\}}\boldsymbol{\chi}^\ell(-2\omega;\omega,\omega),
 \end{equation} 
 where $\ell$ would denote a particular layer chosen through
-$\calc^\ell(z)$ of Eq.~\eqref{sz} and
+${\mathbf{\cal C}}^\ell(z)$ of Eq.~\eqref{sz} and
 $\{\ell\}$
@@
 %For a centrosymmetric slab, 
-%$\bfgchi^{\ell=N}=0$;
+%$\boldsymbol{\chi}^{\ell=N}=0$;
 We remark that the value of 
@@
 the slab needs to have enough atomic layers 
-%$\bfgchi^{\ell=N}=0$;
+%$\boldsymbol{\chi}^{\ell=N}=0$;
 %to be satisfied and to 
 in order to give converged results for 
-$\bfgchi (-2\go;\go,\go)$. 
+$\boldsymbol{\chi} (-2\omega;\omega,\omega)$. 
 We can use Eq.~\eqref{chiijksur} for 
 either the front or the back surface. 
-% This decomposition of $\bfgchi(-2\go;\go,\go)$ into
-% $\bfgchi^\ell(-2\go;\go,\go)$ would allow to explore the contribution
+% This decomposition of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$ into
+% $\boldsymbol{\chi}^\ell(-2\omega;\omega,\omega)$ would allow to explore the contribution
 % to SH from any particular layer of the system,
@@
 \begin{align}\label{hs}
-\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go) 
+\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega) 
 =
-\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go) 
+\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega) 
 ,
@@
 where
-$\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
-$\calc(z)=1$ from the upper half containing the $2\times 1$ 
+$\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega)$ is calculated using
+${\mathbf{\cal C}}(z)=1$ from the upper half containing the $2\times 1$ 
 surface reconstruction, as seen in Fig.~\ref{si2x1},
-and $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
-$\calc(z)=1$ through the full slab.
+and $\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega)$ is calculated using
+${\mathbf{\cal C}}(z)=1$ through the full slab.
 We show the results for this comparison in the remainder 
@@
 form.\cite{kleinmanPRL82}
-The contribution of $\bfv^\nl$ to Eq.~\eqref{chis} can be carried out 
+The contribution of $\mathbf{v}^\mathrm{nl}$ to Eq.~\eqref{chis} can be carried out 
 as detailed in Appendix \ref{vesnl} using the DP code for 
@@
 these are: an energy cut-off of 10 Ha, an equal number of conduction and 
-valence bands, and a set of 244 $\bfk$-points.
-The $\bfk$-points are used for the linear analytic 
+valence bands, and a set of 244 $\mathbf{k}$-points.
+The $\mathbf{k}$-points are used for the linear analytic 
 tetrahedron method for evaluating the
@@
 implies that there is a region in the Brillouin zone where 
-$\omega_{cv}(\bfk)=2 \omega_{c'v}(\bfk)$. Since there is a real population 
+$\omega_{cv}(\mathbf{k})=2 \omega_{c'v}(\mathbf{k})$. Since there is a real population 
 excitation, the perturbation theory used to calculate the spectrum breaks 
-down at these $\bfk$-points. This must be taken into account in a more 
+down at these $\mathbf{k}$-points. This must be taken into account in a more 
 complete calculation. These points introduce sharp spikes in the spectrum. 
@@
 We must evaluate 
-$T^{\rma\rmb}_{nm}=(i/\hbar)[r^\rmb,v^{\nl,\rma}]_{nm}$,
+$T^{\mathrm{a}\mathrm{b}}_{nm}=(i/\hbar)[r^\mathrm{b},v^{\mathrm{nl},\mathrm{a}}]_{nm}$,
 as outlined in Appendix \ref{vesnl},
@@
 time consuming. This adds significantly to the already lengthy time needed 
-for the calculation of the $\bfv^\nl$ contribution that is
+for the calculation of the $\mathbf{v}^\mathrm{nl}$ contribution that is
 proportional only to 
 the first order derivatives. 
-Memory requirements are also increased for both $\bfv^\nl$ and 
-$[\bfr,\bfv^\nl]$. However, the contribution from $[\bfr,\bfv^\nl]$ 
+Memory requirements are also increased for both $\mathbf{v}^\mathrm{nl}$ and 
+$[\mathbf{r},\mathbf{v}^\mathrm{nl}]$. However, the contribution from $[\mathbf{r},\mathbf{v}^\mathrm{nl}]$ 
 is very small\cite{valerie} and therefore we neglect it in this work.
@@
 As explained above,
-the calculation of the $\bfv^\nl$ contribution is 
+the calculation of the $\mathbf{v}^\mathrm{nl}$ contribution is 
 computationally expensive.
@@
 between including or not including the
-effects of $\bfv^\nl$ or the scissors correction
-$\hbar\gD$.   
+effects of $\mathbf{v}^\mathrm{nl}$ or the scissors correction
+$\hbar\Delta$.   
 For these results we chose
-$\hbar\gD=0.5$ eV, that is the GW gap reported in
+$\hbar\Delta=0.5$ eV, that is the GW gap reported in
 Refs.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
@@
 
-\subsection{\texorpdfstring{Results for $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$}
+\subsection{\texorpdfstring{Results for $\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)$}
 {Results for Xxxx(2x1)(-2w;w,w)}}
@@
 band gap of this surface, and $E_g=1.03$ eV is the scissor shifted gap for
-$\hbar\gD=0.5$ eV.
+$\hbar\Delta=0.5$ eV.
 These resonances
@@
 In Fig.~\ref{fig3}
-We see that the effect of $\bfv^\nl$
+We see that the effect of $\mathbf{v}^\mathrm{nl}$
  reduces the value of   
@@
 of SSHG.  
-However, the inclusion of $\bfv^\nl$ does not
+However, the inclusion of $\mathbf{v}^\mathrm{nl}$ does not
 changes the 
@@
 atomic Si layers plus one H layer, 
-with and without the contribution from $\bfv^\nl$.
+with and without the contribution from $\mathbf{v}^\mathrm{nl}$.
 \label{fig3}} 
@@
 To see the effect of the scissors correction, we take two different
-finite values for $\hbar\gD$. The first one
-with a value of $\hbar\gD=0.5$ eV, used in above results, 
+finite values for $\hbar\Delta$. The first one
+with a value of $\hbar\Delta=0.5$ eV, used in above results, 
 is the ``average'' GW gap taken from 
@@
 that is in agreement with Ref.~\onlinecite{garciaCPC01}. The second one
-with a value of $\hbar\gD=0.63$ eV is the ``average'' 
+with a value of $\hbar\Delta=0.63$ eV is the ``average'' 
 gap taken from Ref.~\onlinecite{asahiPRB00}, 
-where more $\bfk$-points in the Brillouin zone were 
+where more $\mathbf{k}$-points in the Brillouin zone were 
 used to calculate its GW value.
@@
 When we compare 
-$|\chi^{xxx}_{2\times 1}|$ for the two finite values of $\hbar\gD$,
+$|\chi^{xxx}_{2\times 1}|$ for the two finite values of $\hbar\Delta$,
 we see that the first two peaks are almost rigidly 
@@
 for two different values of 
-the scissors correction $\hbar\gD$. 
+the scissors correction $\hbar\Delta$. 
 \label{fig4}} 
@@
 So, we would expect that the $xxx$ component of 
-$\bfgchi(-2\go;\go,\go)$ would have a small influence from the local fields.
+$\boldsymbol{\chi}(-2\omega;\omega,\omega)$ would have a small influence from the local fields.
 Unfortunately the experimental measurement of the $xxx$
-component of $\bfgchi(-2\go;\go,\go)$ is not possible as the SH 
+component of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$ is not possible as the SH 
 radiated intensity would be proportional not only to this component 
-but also to the other components of $\bfgchi(-2\go;\go,\go)$.  
+but also to the other components of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$.  
 However,
@@
 We have presented a formulation to calculate the surface second-harmonic
-(SSH) susceptibility tensor $\bfgchi(-2\go;\go,\go)$, using the length gauge
+(SSH) susceptibility tensor $\boldsymbol{\chi}(-2\omega;\omega,\omega)$, using the length gauge
 formalism and within the independent particle approximation (IPA).
@@
 that 
-$\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go) 
+$\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega) 
 \approx
-\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go) 
+\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega) 
 . 
@@
 Although one can in principle increase the number of atomic layers,
-$\bfk$-points, etc. to
+$\mathbf{k}$-points, etc. to
 improve even further on the similarity of the half-slab and full-slab results, we
@@
 effects mentioned above in the calculation of  
-$\bfgchi(-2\go;\go,\go)$. 
+$\boldsymbol{\chi}(-2\omega;\omega,\omega)$. 
 The scissors correction shifts the spectrum to  
@@
  The cut function allows us  
-to extract unequivocally $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$.
+to extract unequivocally $\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)$.
  The effects of the nonlocal part of the pseudopotentials  
 keeps the same line-shape of $|\chi^{xxx}_{2\times
-  1}(-2\go;\go,\go)|$,
+  1}(-2\omega;\omega,\omega)|$,
 but  
@@
 The $xxx$ component of  
-$\bfgchi_{2\times 1}(-2\go;\go,\go)$, 
+$\boldsymbol{\chi}_{2\times 1}(-2\omega;\omega,\omega)$, 
 can not be
@@
 \subsection{ Expressions for 
-\texorpdfstring{${\calbv}_{nm}(\bfk)$}{que} 
+\texorpdfstring{${\boldsymbol{\mathcal{\cal V}}}_{nm}(\mathbf{k})$}{que} 
 and 
-\texorpdfstring{${\cal C}_{nm}(\bfk)$}{que}
+\texorpdfstring{${\cal C}_{nm}(\mathbf{k})$}{que}
 }\label{calpcalc}
@@
 \begin{align}\label{eni.1}
-\psi_{n\bfk}(\bfr)=\sum_\bfG A_{n\bfk}(\bfG)e^{i(\bfk+\bfG)\cdot\bfr}
-,
-\end{align}
-where $\{\bfG\}$ are the reciprocal basis vectors satisfying
-$e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
-space, and $A_{n\bfk}(\bfG)$ the expansion coefficients. Using
-$m_e\bfv=\bfp=-i\hbar\bfgnabla$ into 
+\psi_{n\mathbf{k}}(\mathbf{r})=\sum_\mathbf{G} A_{n\mathbf{k}}(\mathbf{G})e^{i(\mathbf{k}+\mathbf{G})\cdot\mathbf{r}}
+,
+\end{align}
+where $\{\mathbf{G}\}$ are the reciprocal basis vectors satisfying
+$e^{\mathbf{R}\cdot\mathbf{G}}=1$, with $\{\mathbf{R}\}$ the translation vectors in real
+space, and $A_{n\mathbf{k}}(\mathbf{G})$ the expansion coefficients. Using
+$m_e\mathbf{v}=\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ into 
 Eq.~\eqref{vcali}
@@
 \begin{align}\label{eni.2}
-\calbv_{nm}(\bfk)=
+\boldsymbol{\mathcal{\cal V}}_{nm}(\mathbf{k})=
 \frac{\hbar}{2m_e}
-\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
-(2\bfk+\bfG+\bfG')
-\gd_{\bfG_\parallel \bfG'_\parallel}  
+\sum_{\mathbf{G},\mathbf{G}'} A^*_{n\mathbf{k}}(\mathbf{G}')  A_{m\mathbf{k}}(\mathbf{G})
+(2\mathbf{k}+\mathbf{G}+\mathbf{G}')
+\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel}  
 f(G_\perp-G'_\perp)
@@
 \begin{align}\label{vnl.9}
-f(G_\perp)=\frac{1}{L}\int\calc(z) e^{iG_\perp z}dz  
+f(G_\perp)=\frac{1}{L}\int{\mathbf{\cal C}}(z) e^{iG_\perp z}dz  
  ,
@@
 is the length of the super-cell. 
-The reciprocal lattice vectors $\bfG$ are 
+The reciprocal lattice vectors $\mathbf{G}$ are 
 decomposed into components
-parallel ($\bfG_\parallel$), and perpendicular ($G_\perp \hat z$)
+parallel ($\mathbf{G}_\parallel$), and perpendicular ($G_\perp \hat z$)
 to the surface, so
-that $\bfG = \bfG_\parallel + G_\perp\hat z$.
-The double summation over the $\bfG$ vectors can be 
+that $\mathbf{G} = \mathbf{G}_\parallel + G_\perp\hat z$.
+The double summation over the $\mathbf{G}$ vectors can be 
 calculated efficiently by  
@@
 \begin{align}\label{eni.4}
-\calc_{nm}(\bfk)=
-\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
-\gd_{\bfG_\parallel \bfG'_\parallel} 
+{\mathbf{\cal C}}_{nm}(\mathbf{k})=
+\sum_{\mathbf{G},\mathbf{G}'} A^*_{n\mathbf{k}}(\mathbf{G}')  A_{m\mathbf{k}}(\mathbf{G})
+\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel} 
 f(G_\perp-G'_\perp)
@@
 \end{align}  
-If $\calc(z)=1$, then $f(G_\perp)=\gd_{G_\perp 0}$ and we 
+If ${\mathbf{\cal C}}(z)=1$, then $f(G_\perp)=\delta_{G_\perp 0}$ and we 
 obtain the full-slab/bulk values, 
-$\bfv_{nm}(\bfk)$ and $\calc_{nm}(\bfk)=\gd_{nm}$,
+$\mathbf{v}_{nm}(\mathbf{k})$ and ${\mathbf{\cal C}}_{nm}(\mathbf{k})=\delta_{nm}$,
 from Eqs.~\eqref{eni.2} and \eqref{eni.4}.
@@
 \subsection{Expressions for 
-\texorpdfstring{$(\calv^{\lda,a}_{nm})_{;k^b}$}{Vnonlocal},
+\texorpdfstring{$({\mathbf{\cal V}}^{\mathrm{LDA},a}_{nm})_{;k^b}$}{Vnonlocal},
 \texorpdfstring{$(r^a_{nm})_{;k^b}$}{Vnonlocal}
-and \texorpdfstring{$(\calc_{nm})_{;\bfk}$}{Vnonlocal}
+and \texorpdfstring{$({\mathbf{\cal C}}_{nm})_{;\mathbf{k}}$}{Vnonlocal}
 }\label{appvnl}
@@
 $
-\calt^{\rma\rmb}\equiv[r^{\rmb},\calv^{\lda,\rma}]
+{\cal T}^{\mathrm{a}\mathrm{b}}\equiv[r^{\mathrm{b}},{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}]
 \equiv
-[r^{\rmb},\calv^\rma]
+[r^{\mathrm{b}},{\mathbf{\cal V}}^\mathrm{a}]
 $
@@
 \begin{align}\label{nmesn}
-(\calv^{\lda,\rma}_{nm})_{;k^{\rmb}}&=
-\calt_{nm}^{\rma\rmb}
+({\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}&=
+{\cal T}_{nm}^{\mathrm{a}\mathrm{b}}
 +i
@@
 \bigg(
-r^{\rmb}_{nq}  
-\calv^{\lda,\rma}_{q m}
+r^{\mathrm{b}}_{nq}  
+{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{q m}
 -
-\calv^{\lda,\rma}_{nq}   
-r^{\rmb}_{q m}
+{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nq}   
+r^{\mathrm{b}}_{q m}
 \bigg)  
 +i  
-r^{\rmb}_{nm}
-\Delta^{\rma}_{mn}
+r^{\mathrm{b}}_{nm}
+\Delta^{\mathrm{a}}_{mn}
 ,
@@
 \begin{eqnarray}\label{tdel}
-\Delta^{\rma}_{mn}
+\Delta^{\mathrm{a}}_{mn}
 =
-\calv^{\lda,\rma}_{nn}  
+{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nn}  
 -
-\calv^{\lda,\rma}_{mm}  
+{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{mm}  
 ,
@@
 \begin{align}\label{tau.1}
-\calt_{nm}^{\rma\rmb}
+{\cal T}_{nm}^{\mathrm{a}\mathrm{b}}
 =
-\frac{\hbar}{m_e}\gd_{\rma\rmb} 
+\frac{\hbar}{m_e}\delta_{\mathrm{a}\mathrm{b}} 
 C_{nm} 
@@
 \sum_q  
-L^{\rma\rmb}_{nq} 
+L^{\mathrm{a}\mathrm{b}}_{nq} 
 C_{qm} 
@@
 \begin{align}\label{lab}
-L_{nm}^{\rma\rmb}
-=\frac{i}{\hbar}[r^{\rmb},v^{\nl,\rma}]_{nm}
+L_{nm}^{\mathrm{a}\mathrm{b}}
+=\frac{i}{\hbar}[r^{\mathrm{b}},v^{\mathrm{nl},\mathrm{a}}]_{nm}
 .
 \end{align}  
-The matrix elements $L^{\rma\rmb}_{nm}$
+The matrix elements $L^{\mathrm{a}\mathrm{b}}_{nm}$
 are calculated in Appendix \ref{vesnl}.
@@
 Notice that
-$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$ is obtained 
+$(v^{\mathrm{LDA},\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}$ is obtained 
 from Eq.~\eqref{nmesn} by 
 taking 
-$\calc(z)=1$ or $C_{nm}=\gd_{nm}$.
-
-To obtain $(r^{\rma}_{nm})_{;k^{\rmb}}$ we use Eq.~\eqref{chon.10} to
+${\mathbf{\cal C}}(z)=1$ or $C_{nm}=\delta_{nm}$.
+
+To obtain $(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}$ we use Eq.~\eqref{chon.10} to
 write
-$(r^{\rma}_{nm})_{;k^{\rmb}}
-=(v^{\lda,\rma}_{nm}/i\go^\lda_{nm})_{;k^{\rmb}}
+$(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}
+=(v^{\mathrm{LDA},\mathrm{a}}_{nm}/i\omega^\mathrm{LDA}_{nm})_{;k^{\mathrm{b}}}
 $ and simply apply the chain rule,
 \begin{align}\label{rka}
-(r^{\rma}_{nm})_{;k^{\rmb}}
+(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}
 &=
-T^{\rma\rmb}_{nm}
+T^{\mathrm{a}\mathrm{b}}_{nm}
 +
 \frac{ 
-r^{\rmb}_{nm}
-\Delta^{\rma}_{mn}
-+r^{\rma}_{nm}
-\Delta^{\rmb}_{mn}
+r^{\mathrm{b}}_{nm}
+\Delta^{\mathrm{a}}_{mn}
++r^{\mathrm{a}}_{nm}
+\Delta^{\mathrm{b}}_{mn}
 }
-{\go^\lda_{nm}}
-+
-\frac{i}{\go^\lda_{nm}}
+{\omega^\mathrm{LDA}_{nm}}
++
+\frac{i}{\omega^\mathrm{LDA}_{nm}}
 \sum_{q}
 \bigg(
-\go^\lda_{q m} 
-r^{\rmb}_{nq} 
-r^{\rma}_{q m}
+\omega^\mathrm{LDA}_{q m} 
+r^{\mathrm{b}}_{nq} 
+r^{\mathrm{a}}_{q m}
 -
-\go^\lda_{nq} 
-r^{\rma}_{nq} 
-r^{\rmb}_{q m}
+\omega^\mathrm{LDA}_{nq} 
+r^{\mathrm{a}}_{nq} 
+r^{\mathrm{b}}_{q m}
 \bigg)
@@
 \begin{eqnarray}\label{del}
-\Delta^{\rma}_{mn}
+\Delta^{\mathrm{a}}_{mn}
 =
-v^{\lda,\rma}_{nn}  
+v^{\mathrm{LDA},\mathrm{a}}_{nn}  
 -
-v^{\lda,\rma}_{mm}  
+v^{\mathrm{LDA},\mathrm{a}}_{mm}  
 ,
@@
 \begin{align}\label{tau.1n} 
-T_{nm}^{\rma\rmb}
-=\frac{\hbar}{m_e}\gd_{ab}\gd_{nm} 
--\hbar L^{\rma\rmb}_{nm} 
+T_{nm}^{\mathrm{a}\mathrm{b}}
+=\frac{\hbar}{m_e}\delta_{ab}\delta_{nm} 
+-\hbar L^{\mathrm{a}\mathrm{b}}_{nm} 
 .
@@
 Note that the layered term
-$\calt^{\rma\rmb}_{nm}$ reduces to $T^{\rma\rmb}_{nm}$
+${\cal T}^{\mathrm{a}\mathrm{b}}_{nm}$ reduces to $T^{\mathrm{a}\mathrm{b}}_{nm}$
 for the full-slab/bulk case.
@@
 Again, we use Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn},
-along with $[\bfr,F(\bfr)]=0$, valid for 
-and any function $F(\bfr)$, 
+along with $[\mathbf{r},F(\mathbf{r})]=0$, valid for 
+and any function $F(\mathbf{r})$, 
 to obtain 
 \begin{align} 
-(\calc_{nm})_{;\bfk}
+({\mathbf{\cal C}}_{nm})_{;\mathbf{k}}
 &=
@@
 \sum_{q} 
- \left(\bfr_{nq}
-\calc_{qm}
+ \left(\mathbf{r}_{nq}
+{\mathbf{\cal C}}_{qm}
 -
-\calc_{nq}
-\bfr_{qm}
+{\mathbf{\cal C}}_{nq}
+\mathbf{r}_{qm}
 \right) 
-+i\bfr_{nm}(\calc_{mm}-\calc_{nn}) 
-,
-\end{align} 
-where we remind the reader that $\bfr_{nm}$ 
++i\mathbf{r}_{nm}({\mathbf{\cal C}}_{mm}-{\mathbf{\cal C}}_{nn}) 
+,
+\end{align} 
+where we remind the reader that $\mathbf{r}_{nm}$ 
 is calculated through 
@@
 \subsection{
-\texorpdfstring{$\calbv^{\nl}_{nm}$}{Vnonlocal},
-\texorpdfstring{$\bfv^\nl_{nm}$}{Vnonlocal}, and 
-\texorpdfstring{$T^{\rma\rmb}_{nm}$}{[r,vnl]}}
+\texorpdfstring{$\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}$}{Vnonlocal},
+\texorpdfstring{$\mathbf{v}^\mathrm{nl}_{nm}$}{Vnonlocal}, and 
+\texorpdfstring{$T^{\mathrm{a}\mathrm{b}}_{nm}$}{[r,vnl]}}
 \label{vesnl}
@@
 and use
-$\sum_{\bfG}\ket{\bfk+\bfG}\bra{\bfk+\bfG}=1$, 
+$\sum_{\mathbf{G}} | \mathbf{k}+\mathbf{G}\rangle \langle\mathbf{k}+\mathbf{G} | =1$, 
 with
-$\braket{\bfr}{\bfk+\bfG}=(1/\sqrt{\gO})
-\mathrm{exp}(i(\bfk+\bfG)\cdot\bfr)$
+$\langle\mathbf{r}|\mathbf{k}+\mathbf{G}=(1/\sqrt{\Omega})
+\mathrm{exp}(i(\mathbf{k}+\mathbf{G})\cdot\mathbf{r})$
 to obtain,
 \begin{align}\label{vnl.5}
-\calbv^{\nl}_{nm}(\bfk)&=\frac{1}{2}
-\sum_{\bfG}
+\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}(\mathbf{k})&=\frac{1}{2}
+\sum_{\mathbf{G}}
 \Big(
-\bra{n\bfk} \calc(z) 
-\ket{\bfk+\bfG}\bra{\bfk+\bfG}
-\bfv^\nl \ket{m\bfk}
-+
-\bra{n\bfk}
-\bfv^\nl  
-\ket{\bfk+\bfG}\bra{\bfk+\bfG}
-\calc(z) \ket{m\bfk}
+\langle n\mathbf{k} | {\mathbf{\cal C}}(z) 
+| \mathbf{k}+\mathbf{G}\rangle\langle\mathbf{k}+\mathbf{G} |
+\mathbf{v}^\mathrm{nl} | m\mathbf{k}\rangle
++
+\langle n\mathbf{k} |
+\mathbf{v}^\mathrm{nl}  
+| \mathbf{k}+\mathbf{G}\rangle\langle\mathbf{k}+\mathbf{G} |
+{\mathbf{\cal C}}(z) | m\mathbf{k}\rangle
 \Big)
@@
 &=\frac{1}{2 \hbar}
-\sum_{\bfG}
+\sum_{\mathbf{G}}
 \left(
-\calf^{\ell*}_{n\bfk}(\bfG) 
-H_{m\bfk}(\bfG) 
-+
-H^*_{n\bfk}(\bfG) 
-\calf_{m\bfk}(\bfG) 
+{\mathbf{\cal F}}^{\ell*}_{n\mathbf{k}}(\mathbf{G}) 
+H_{m\mathbf{k}}(\mathbf{G}) 
++
+H^*_{n\mathbf{k}}(\mathbf{G}) 
+{\mathbf{\cal F}}_{m\mathbf{k}}(\mathbf{G}) 
 \right) 
@@
 \begin{align}\label{vnl.10}
-\calf_{n\bfk}(\bfG) 
+{\mathbf{\cal F}}_{n\mathbf{k}}(\mathbf{G}) 
 =
-\sum_{\bfG'} 
-A_{n\bfk}(\bfG') 
-\gd_{\bfG_\parallel \bfG'_\parallel}f(G'_\perp-G_\perp) 
+\sum_{\mathbf{G}'} 
+A_{n\mathbf{k}}(\mathbf{G}') 
+\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel}f(G'_\perp-G_\perp) 
 ,
@@
 \begin{align}\label{vnl.11}
-H_{n\bfk}(\bfG)&=
-\sum_{\bfG'} 
-A_{n\bfk}(\bfG') 
-(\nabla_{\bfK}+\nabla_{\bfK'})  
-V^\nl(\bfK,\bfK')
-,
-\end{align}
-and $\bfK=\bfk+\bfG$, $\bfK'=\bfk+\bfG'$.
+H_{n\mathbf{k}}(\mathbf{G})&=
+\sum_{\mathbf{G}'} 
+A_{n\mathbf{k}}(\mathbf{G}') 
+(\nabla_{\mathbf{K}}+\nabla_{\mathbf{K}'})  
+V^\mathrm{nl}(\mathbf{K},\mathbf{K}')
+,
+\end{align}
+and $\mathbf{K}=\mathbf{k}+\mathbf{G}$, $\mathbf{K}'=\mathbf{k}+\mathbf{G}'$.
 For fully  separable pseudopotentials in the   
@@
 matrix elements 
- $\bra{\bfK}  
-V^\nl  
-\ket{\bfK'}
-=V^\nl(\bfK,\bfK')  
+ $\langle\mathbf{K} |
+V^\mathrm{nl}  
+| \mathbf{K}'\rangle
+=V^\mathrm{nl}(\mathbf{K},\mathbf{K}')  
 $ 
-and their $\bfK$ and $\bfK'$ gradient 
+and their $\mathbf{K}$ and $\mathbf{K}'$ gradient 
 can be readily calculated.\cite{mottaCMS10,adolphPRB96,gordienkoRPJ04,fuchsCPC99} 
@@
 implemented 
-the calculation of $\calbv^{\nl}_{nm}(\bfk)$ with the help of 
-the \depe~code.\cite{francesco}
-We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
-expressions by taking $\calc(z)=1$, which implies 
-that $f(G_\perp)=\gd_{G_\perp 0}$, 
+the calculation of $\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}(\mathbf{k})$ with the help of 
+the DP${\texttrademark}$~code.\cite{francesco}
+We obtain $\mathbf{v}^{\mathrm{nl}}_{nm}(\mathbf{k})$ from the previous three 
+expressions by taking ${\mathbf{\cal C}}(z)=1$, which implies 
+that $f(G_\perp)=\delta_{G_\perp 0}$, 
 and 
-$\calf_{n\bfk}(\bfG) \to F_{n\bfk}(\bfG)=A_{n\bfk}(\bfG)$.
+${\mathbf{\cal F}}_{n\mathbf{k}}(\mathbf{G}) \to F_{n\mathbf{k}}(\mathbf{G})=A_{n\mathbf{k}}(\mathbf{G})$.
 
@@
 \begin{align}\label{3.1}
-L^{\rma\rmb}_{nm}(\bfk) 
+L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
 =
 \frac{1}{\hbar^2}
-\bra{n\bfk}
-\big[ r^{\rma},[ V^\nl, r^\rmb]\big]
-\ket{m\bfk}
+\langle n\mathbf{k} |
+\big[ r^{\mathrm{a}},[ V^\mathrm{nl}, r^\mathrm{b}]\big]
+| m\mathbf{k}\rangle
 ,
@@
 \begin{align}\label{3.5}
-L^{\rma\rmb}_{nm}(\bfk) 
+L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
 &=
-\frac{1}{\hbar^2\gO}
-\sum_{\bfG,\bfG'} 
-A^*_{n\bfk}(\bfK) 
-A_{m\bfk}(\bfK')
+\frac{1}{\hbar^2\Omega}
+\sum_{\mathbf{G},\mathbf{G}'} 
+A^*_{n\mathbf{k}}(\mathbf{K}) 
+A_{m\mathbf{k}}(\mathbf{K}')
 \int
-d\bfr d\bfr'
- e^{-i\bfK\cdot\bfr}
+d\mathbf{r} d\mathbf{r}'
+ e^{-i\mathbf{K}\cdot\mathbf{r}}
 \Big(
-r^{\rma}
-V^\nl(\bfr,\bfr')
-r'^\rmb
+r^{\mathrm{a}}
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
+r'^\mathrm{b}
 -
-V^\nl(\bfr,\bfr')
-r'^\rma
-r'^{\rmb}
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
+r'^\mathrm{a}
+r'^{\mathrm{b}}
 \nonumber\\
 &-
-r^\rmb
-r^{\rma}
-V^\nl(\bfr,\bfr')
-+
- r^\rmb
-V^\nl(\bfr,\bfr')
-r'^{\rma}
+r^\mathrm{b}
+r^{\mathrm{a}}
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
++
+ r^\mathrm{b}
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
+r'^{\mathrm{a}}
 \Big) 
- e^{i\bfK'\cdot\bfr'}
+ e^{i\mathbf{K}'\cdot\mathbf{r}'}
 ,
@@
 where 
-$V^\nl(\bfr,\bfr') = \bra{\bfr} V^\nl\ket{\bfr'}$.
+$V^\mathrm{nl}(\mathbf{r},\mathbf{r}') = \langle\mathbf{r} | V^\mathrm{nl} | \mathbf{r}'\rangle$.
 We use the following identity
@@
 \Big(
-\frac{\partial^2}{\partial K^\rma\partial K'^\rmb}
-+
-\frac{\partial^2}{\partial K'^\rma\partial K'^\rmb}
-+
-\frac{\partial^2}{\partial K^\rma\partial K^\rmb}
-+
-\frac{\partial^2}{\partial K^\rmb\partial K'^\rma}
+\frac{\partial^2}{\partial K^\mathrm{a}\partial K'^\mathrm{b}}
++
+\frac{\partial^2}{\partial K'^\mathrm{a}\partial K'^\mathrm{b}}
++
+\frac{\partial^2}{\partial K^\mathrm{a}\partial K^\mathrm{b}}
++
+\frac{\partial^2}{\partial K^\mathrm{b}\partial K'^\mathrm{a}}
 \Big)
 \int 
-d\bfr d\bfr' 
- e^{-i\bfK\cdot\bfr}
-V^\nl(\bfr,\bfr') 
-e^{i\bfK'\cdot\bfr'}
+d\mathbf{r} d\mathbf{r}' 
+ e^{-i\mathbf{K}\cdot\mathbf{r}}
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
+e^{i\mathbf{K}'\cdot\mathbf{r}'}
 \nonumber\\
 &=
-\int d\bfr d\bfr'
- e^{-i\bfK\cdot\bfr}
+\int d\mathbf{r} d\mathbf{r}'
+ e^{-i\mathbf{K}\cdot\mathbf{r}}
 \Big( 
-r^{\rma} 
-V^\nl(\bfr,\bfr') 
-r'^\rmb
+r^{\mathrm{a}} 
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
+r'^\mathrm{b}
 - 
-V^\nl(\bfr,\bfr') 
-r'^\rma 
-r'^{\rmb}
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
+r'^\mathrm{a} 
+r'^{\mathrm{b}}
 - 
-r^\rmb 
-r^{\rma} 
-V^\nl(\bfr,\bfr')
-+
- r^\rmb 
-V^\nl(\bfr,\bfr') 
-r'^{\rma}
+r^\mathrm{b} 
+r^{\mathrm{a}} 
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
++
+ r^\mathrm{b} 
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
+r'^{\mathrm{a}}
 \Big)  
-e^{i\bfK'\cdot\bfr'}
+e^{i\mathbf{K}'\cdot\mathbf{r}'}
 ,
@@
 \begin{align}\label{3.7}
-L^{\rma\rmb}_{nm}(\bfk)
+L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k})
 &=
 \frac{1}{\hbar^2}
-\sum_{\bfG,\bfG'} 
-A^*_{n\bfk}(\bfK) 
-A_{m\bfk}(\bfK')
+\sum_{\mathbf{G},\mathbf{G}'} 
+A^*_{n\mathbf{k}}(\mathbf{K}) 
+A_{m\mathbf{k}}(\mathbf{K}')
 \Big(
-\frac{\partial^2}{\partial K^\rma\partial K'^\rmb}
-+
-\frac{\partial^2}{\partial K'^\rma\partial K'^\rmb}
-+
-\frac{\partial^2}{\partial K^\rma\partial K^\rmb}
-+
-\frac{\partial^2}{\partial K^\rmb\partial K'^\rma}
+\frac{\partial^2}{\partial K^\mathrm{a}\partial K'^\mathrm{b}}
++
+\frac{\partial^2}{\partial K'^\mathrm{a}\partial K'^\mathrm{b}}
++
+\frac{\partial^2}{\partial K^\mathrm{a}\partial K^\mathrm{b}}
++
+\frac{\partial^2}{\partial K^\mathrm{b}\partial K'^\mathrm{a}}
 \Big)
-\bra{\bfK} 
-V^\nl 
-\ket{\bfK'}, 
+\langle \mathbf{K} |
+V^\mathrm{nl} 
+| \mathbf{K}'\rangle, 
 \nonumber\\
@@
 \begin{align}\label{vkk}
-\bra{\bfK} 
-V^\nl 
-\ket{\bfK'} 
+\langle\mathbf{K} |
+V^\mathrm{nl} 
+| \mathbf{K}'\rangle
 =
-\frac{1}{\gO}
+\frac{1}{\Omega}
 \int 
-d\bfr d\bfr' 
- e^{-i\bfK\cdot\bfr}
-V^\nl(\bfr,\bfr') 
-e^{i\bfK'\cdot\bfr'}
-,
-\end{align}
-The double derivatives with respect to $\bfK$ and $\bfK'$ 
+d\mathbf{r} d\mathbf{r}' 
+ e^{-i\mathbf{K}\cdot\mathbf{r}}
+V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
+e^{i\mathbf{K}'\cdot\mathbf{r}'}
+,
+\end{align}
+The double derivatives with respect to $\mathbf{K}$ and $\mathbf{K}'$ 
 can be worked out explicitly and 
-$L^{\rma\rmb}_{nm}(\bfk)$
+$L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k})$
 can be finally calculated.\cite{valerie}
@@
 %%%%
-\subsection{Expressions for  \texorpdfstring{$\calbv^{S}_{nm}$}{Vnm}
+\subsection{Expressions for  \texorpdfstring{$\boldsymbol{\mathcal{\cal V}}^{S}_{nm}$}{Vnm}
 and
-\texorpdfstring{$\Big({\calbv}^{S}_{nm}\Big)_{;\bfk}$}{(Vnm);kb}
+\texorpdfstring{$\Big({\boldsymbol{\mathcal{\cal V}}}^{S}_{nm}\Big)_{;\mathbf{k}}$}{(Vnm);kb}
 }\label{calvs}
@@
 \begin{align}\label{a.3b}
-\calbv^{S}_{nm}
+\boldsymbol{\mathcal{\cal V}}^{S}_{nm}
 &=
 \frac{1}{2}\sum_q\left(   
-\bfv^S_{nq}\calc_{qm}+\calc_{nq}\bfv^S_{qm}
+\mathbf{v}^S_{nq}{\mathbf{\cal C}}_{qm}+{\mathbf{\cal C}}_{nq}\mathbf{v}^S_{qm}
 \right)  
@@
 \end{align}    
-where $\sum_q\ket{q\bfk}\bra{q\bfk}=1$ was used
-and $\bfv^S_{nm}$ is given in Eq.~\eqref{chon.2}.
+where $\sum_q | q\mathbf{k}\rangle\langle q\mathbf{k} | =1$ was used
+and $\mathbf{v}^S_{nm}$ is given in Eq.~\eqref{chon.2}.
 Taking the generalized derivative of Eq.~\eqref{a.3b}
@@
 \begin{align}\label{a.3bn}
-\left(\calbv^{S}_{nm}\right)_{;\bfk}
+\left(\boldsymbol{\mathcal{\cal V}}^{S}_{nm}\right)_{;\mathbf{k}}
 &=
 \frac{1}{2}\sum_q\left(
-(\bfv^S_{nq})_{;\bfk}\calc_{qm}
-+    
-\bfv^S_{nq}(\calc_{qm})_{;\bfk}
-+
-(\calc_{nq})_{;\bfk} \bfv^S_{qm}
-+
-\calc_{nq} (\bfv^S_{qm})_{;\bfk}
+(\mathbf{v}^S_{nq})_{;\mathbf{k}}{\mathbf{\cal C}}_{qm}
++    
+\mathbf{v}^S_{nq}({\mathbf{\cal C}}_{qm})_{;\mathbf{k}}
++
+({\mathbf{\cal C}}_{nq})_{;\mathbf{k}} \mathbf{v}^S_{qm}
++
+{\mathbf{\cal C}}_{nq} (\mathbf{v}^S_{qm})_{;\mathbf{k}}
 \right)  
@@
 \begin{align}\label{choni.1}
-(\bfv^S_{nm})_{;\bfk}=i\gD f_{mn}
-(\bfr_{nm})_{;\bfk}
+(\mathbf{v}^S_{nm})_{;\mathbf{k}}=i\Delta f_{mn}
+(\mathbf{r}_{nm})_{;\mathbf{k}}
 ,

### 2014-11-20 paper-v-0.tex
no errors any more! ready for last format

@@
 
-\title{Theory of surface second-harmonic generation for semiconductors
-including effects of nonlocal operators}
+\title{Theory of surface second-harmonic generation for 
+       semiconductors including effects of nonlocal operators}
+
 \author{Sean M. Anderson}
-\affiliation{Centro de Investigaciones en Optica, Le\'on, Guanajuato, M\'exico}
+    \affiliation{Centro de Investigaciones en Optica, 
+                Le\'on, Guanajuato, M\'exico}
 \author{Nicolas Tancogne-Dejean}
-\affiliation{Laboratoire des Solides Irradi\'es, \'Ecole 
-  Polytechnique, CNRS, CEA/DSM, 91128 Palaiseau, France}
-\affiliation{European Theoretical Spectroscopy Facility (ETSF), Palaiseau, France}
+    \affiliation{Laboratoire des Solides Irradi\'es, \'Ecole Polytechnique, 
+                CNRS, CEA/DSM, 91128 Palaiseau, France}
+    \affiliation{European Theoretical Spectroscopy Facility (ETSF), 
+                Palaiseau, France}
 \author{Bernardo S. Mendoza}\email{bms@cio.mx}
-\affiliation{Centro de Investigaciones en Optica, Le\'on, Guanajuato, M\'exico}
+    \affiliation{Centro de Investigaciones en Optica, 
+                Le\'on, Guanajuato, M\'exico}
 \author{Val\'erie V\'eniard}
-\affiliation{Laboratoire des Solides Irradi\'es, \'Ecole 
-  Polytechnique, CNRS, CEA/DSM, 91128 Palaiseau, France}
- \affiliation{European Theoretical Spectroscopy Facility (ETSF), Palaiseau, France}
+    \affiliation{Laboratoire des Solides Irradi\'es, \'Ecole Polytechnique, 
+                CNRS, CEA/DSM, 91128 Palaiseau, France}
+    \affiliation{European Theoretical Spectroscopy Facility (ETSF), 
+                Palaiseau, France}
 
@@
 the calculation of $\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}(\mathbf{k})$ with the help of 
-the DP${\texttrademark}$~code.\cite{francesco}
+the DP\texttrademark~code.\cite{francesco}
 We obtain $\mathbf{v}^{\mathrm{nl}}_{nm}(\mathbf{k})$ from the previous three 

### 2014-11-21 paper-v-0.tex
done with abstract and intro

@@
-\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
+%\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
+\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,reprint,letterpaper]{revtex4}
 \usepackage{amsfonts}
@@
 \begin{abstract}
-We formulate a theoretical approach of surface second-harmonic generation 
-from semiconductor surfaces based on the length gauge. 
-Within the independent particle approximation 
-the nonlinear second-order surface susceptibility tensor
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$   
-is calculated, including for the first time in one unique formulation
-i) the scissors correction, needed to have the correct value of the
-energy band gap,
-ii) the derivation for the inclusion of the cut function, 
-used to extract the surface nonlinear response,
-and iii) the contribution of the nonlocal part of the 
-pseudopotentials, routinely  used in \textit{ab initio} band structure calculations. 
-First and third terms are described by spatially nonlocal quantum
-mechanical operators and are fully taken into account in the present formulation. 
-As a test-case of the approach we calculate 
-$\chi^{xxx}(-2\omega;\omega,\omega)$ for the clean Si(001)$2\times 1$
-reconstructed surface and find that our scheme to extract the surface
-through the cut function
-readily works and is exact. The effects of the scissors correction and
-of the nonlocal part of the pseudopotentials are discussed for the
-first time in surface nonlinear optics. 
-The scissors correction shifts the spectrum to 
-higher energies though the shifting is not rigid and mixes the 
-$1\omega$ and $2\omega$ resonances, 
-and has a strong influence in the line-shape. 
-The 
-effects of the nonlocal part of the pseudopotentials 
-keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$, but 
-reduces its value
-by 15-20\%.  
-Therefore, the inclusion of the three aforementioned 
-contributions is very important and 
-makes our scheme unprecedented and opens the possibility to 
-study SSHG with more versatility and providing more accurate results 
-than previous approaches. 
+We formulate a theoretical approach of surface second-harmonic generation from 
+semiconductor surfaces based on the length gauge. We calculate the nonlinear 
+second-order surface susceptibility tensor 
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
+independent particle approximation, and for the first time in one unique 
+formulation we include i) the scissors correction, needed to have the correct 
+value for the energy band gap, ii) the derivation for the inclusion of the cut 
+function used to extract the surface nonlinear response, and iii) the 
+contribution of the nonlocal part of the pseudopotentials that are routinely 
+used in \textit{ab initio} band structure calculations. The first and third 
+terms are described by spatially nonlocal quantum mechanical operators and 
+are fully taken into account in the present formulation. As a test-case for 
+this approach, we calculate $\chi^{xxx}(-2\omega;\omega,\omega)$ for a clean 
+Si(001)$2\times 1$ reconstructed surface and find that this scheme for 
+extracting the surface response through the cut function is exact and readily
+implemented. The effects of the scissors correction and the nonlocal part of 
+the pseudopotentials are discussed for the first time for surface nonlinear 
+optics. The scissors correction non-rigidly shifts the spectrum to higher 
+energies by mixing the $1\omega$ and $2\omega$ resonances and exerts a strong 
+influence on the line-shape. The effects of the nonlocal part of the 
+pseudopotentials keeps the same line-shape of 
+$|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$ but reduces its value by 
+15-20\%. Therefore, our scheme is unprecedented through the inclusion of 
+these three contributions and also opens the possibility to study SSHG with 
+more versatility and more accuracy than previous approaches. 
 \end{abstract}
@@
 calculation and thus one uses the much simpler scissors operator scheme.
-\cite{levinePRL89,levinePRL91,delsolePRB93}  
-This allows us to ``open'' the DFT-LDA gap to 
-its correct experimental or GW value even when experiments are not available. 
-In this context, to correct for the underestimation of the energy band gap of 
-semiconductors Nastos et al.\cite{nastosPRB05} used the ``length gauge'' or 
-``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to correctly include the 
-many-body corrections through the scissors operator in the SH susceptibility.
-Later, Cabellos \textit{et al}.\cite{cabellosPRB09} elaborated a derivation 
-of the ``velocity gauge'' or ``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the 
-scissors operator and proved gauge invariance with respect to the length 
-gauge. From these works it is clear the length gauge is a much better staring
-point to obtain the surface SH (SSH) susceptibility, as will be elaborated 
-in this article. However, these considerations are only valid for bulk 
-semiconductors.
+\cite{levinePRL89,levinePRL91,delsolePRB93} This allows us to ``open'' the 
+DFT-LDA gap to its correct experimental or GW value even when experiments 
+are not available. In this context, to correct for the underestimation of 
+the energy band gap of semiconductors Nastos et al.\cite{nastosPRB05} used 
+the ``length gauge'' or ``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to 
+correctly include the many-body corrections through the scissors operator in 
+the SH susceptibility. Later, Cabellos \textit{et al}.\cite{cabellosPRB09} 
+elaborated a derivation of the ``velocity gauge'' or 
+``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the scissors operator 
+and proved gauge invariance with respect to the length gauge. From these works 
+it is clear the length gauge is a much better staring point to obtain the 
+surface SH (SSH) susceptibility, as will be elaborated in this article. 
+However, these considerations are only valid for bulk semiconductors.
 
@@
 Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02,sanoPRB02}
-followed 
-upon this work and in particular Ref.~\onlinecite{arzatePRB01} went into a 
-detailed analysis of the different contributions to the SHG spectra of a 
-surface and the nuanced relationship between bulk, surface, interband, 
+followed upon this work and in particular Ref.~\onlinecite{arzatePRB01} went 
+into a  detailed analysis of the different contributions to the SHG spectra 
+of a surface and the nuanced relationship between bulk, surface, interband, 
 intraband, $1\omega$ and $2\omega$ terms, and
@@
 Finally, DFT-LDA calculations are often based on the use of pseudopotentials.
-As it will be discussed 
-in this article, the presence of a nonlocal part of the
-pseudopotential introduces
-corrections to
-the momentum operator of the electron that have to be included with care in 
-the SSH susceptibility. So far, this contribution has been neglected in the 
-literature.
+As we will discuss later in this article, the presence of the nonlocal part 
+of the pseudopotential introduces corrections to the momentum operator of 
+the electron that have to be included with care into the SSH susceptibility. 
+So far, this contribution has been neglected in the literature.
 
@@
 length gauge formalism. This derivation includes the addition of the terms 
-mentioned above that have been absent in previous works. Also, our 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ can be used 
-for a layer-by-layer analysis if desired. In Sec. \ref{results} we show the 
-results of this reformulation with a study of a clean Si(001) surface with a 
-$2 \times 1$ reconstruction, 
-proving the correctness of our approach with a special test-case,
-never exploited before.
-We compare results from before and after adding 
-the different nonlocal contributions. Finally, in Sec. \ref{conc} we give 
-our conclusions.
+mentioned above that have been absent in previous works. If desired, we can 
+use $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ for a 
+layer-by-layer analysis. In Sec. \ref{results} we show the results of this 
+reformulation with a study of a clean Si(001) surface with a $2 \times 1$ 
+reconstruction that proves the validity of our approach with a special 
+test-case that has never been implemented before. We compare results from 
+before and after adding the different nonlocal contributions. Finally, in 
+Sec. \ref{conc} we submit our conclusions.
 

### 2014-11-21 paper-v-0.tex
equation formatting

@@
 %\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
-\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,reprint,letterpaper]{revtex4}
+\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,letterpaper]{revtex4}
 \usepackage{amsfonts}
@@
 
-In this section we present the scheme used to calculate the surface second-order
-nonlinear response using the length gauge formalism.
-Our derivation includes the aforementioned terms not included
-before for the surface nonlinear
-response.
+In this section we present the scheme used to calculate the surface 
+second-order nonlinear response using the length gauge formalism. Our 
+derivation includes the aforementioned terms not included before for 
+this response.
 
@@
 
-We assume the IPA and neglect local field  
-and excitonic effects.  
-We assume a classical electromagnetic field
-and quantum mechanical matter.
-We can describe the system using 
+We assume the IPA, a classical electromagnetic field, and quantum mechanical 
+matter. We neglect local field and excitonic effects. We can describe the 
+system using 
 %a scaled 
-the one electron density
-operator ${\rho}$, with which we can calculate the expectation value of a
-single-particle observable $\mathcal{O}$ as 
-$\langle{\cal O}\rangle=\mbox{Tr}({\rho} {\cal O}),  
-\label{calo}$
-with $\mathcal{O}$ the associated quantum mechanical operator and Tr
-the trace. 
-The density operator satisfies
-$
-i\hbar (d{\rho}/dt)=[H(t),{\rho}],  
-\label{rho}
-$
+the one electron density operator ${\rho}$, with which we can calculate the 
+expectation value of a single-particle observable $\mathcal{O}$ as 
+$\langle{\cal O}\rangle = \mbox{Tr}({\rho}{\cal O}),\label{calo}$
+with $\mathcal{O}$ the associated quantum mechanical operator and Tr the 
+trace. The density operator satisfies
+$i\hbar(d{\rho}/dt) = [H(t), {\rho}], \label{rho}$
 with $H(t)$ as the total single electron Hamiltonian, written as 
 \begin{equation*}
-H(t)=H_{0}+H_{I}(t),  
+H(t) = H_{0} + H_{I}(t),  
 \label{ache}
@@
 where $H_{0}$ is the unperturbed time-independent Hamiltonian, and $H_{I}(t)$
-is the time-dependent potential energy due to the interaction of the
-electron with the electromagnetic field.
+is the time-dependent potential energy due to the interaction of the electron 
+with the electromagnetic field.
 % ; $H_{0}$ has eigenvalues
@@
 % index $n$ and crystal momentum $\mathbf{k}$.
-To proceed with the solution of $\rho$  it is convenient to use the
-interaction picture, where a unitary operator $U=\exp ({iH_{0}t/\hbar })$
-transforms any operator $\mathcal{O}$ into 
-$\tilde{\mathcal{O}}=U{\cal O}U^{\dagger }$. Even if $\mathcal{O}$ 
-is time-independent, 
-$\tilde{\mathcal{O}}$ is time-dependent through the explicit
-time dependence of $U$. 
-The dynamical 
-equation for $\tilde{\rho}$ is 
-given by
+To proceed with the solution of $\rho$ it is convenient to use the interaction 
+picture, where a unitary operator $U = \exp({iH_{0}t/\hbar})$ transforms any 
+operator $\mathcal{O}$ into $\tilde{\mathcal{O}} = U{\cal O}U^{\dagger}$. 
+Even if $\mathcal{O}$ is time-independent, $\tilde{\mathcal{O}}$ is 
+time-dependent through the explicit time dependence of $U$. The dynamical 
+equation for $\tilde{\rho}$ is given by
 \begin{equation*}
-i\hbar \frac{d\tilde{{\rho}}}{dt}=[\tilde{H}_{I}(t),\tilde{{\rho}}],  
+i\hbar \frac{d\tilde{\rho}}{dt} = [\tilde{H}_{I}(t), \tilde{\rho}],
 \label{rho1}
@@
 \begin{equation}
-i\hbar \tilde{{\rho}}(t)=i\hbar \tilde{{\rho}}_{0}+\int_{-\infty }^{t}dt^{\prime }[
-\tilde{H}_{I}(t^{\prime }),\tilde{\rho}(t^{\prime })],  
+i\hbar \tilde{\rho}(t) = 
+    i\hbar \tilde{\rho}_{0} + 
+    \int_{-\infty}^{t}dt^{\prime}
+    [\tilde{H}_{I}(t^{\prime}),\tilde{\rho}(t^{\prime})],  
 \label{trans}
 \end{equation}
-where $\tilde{\rho}_{0}=\tilde{\rho}(t=-\infty )$ is the unperturbed density matrix. We look
-for the standard perturbation series solution, 
-$\tilde{\rho}(t)=\tilde{\rho}^{(0)}+\tilde{\rho}^{(1)}+\tilde{\rho}^{(2)}+\cdots$,  
-where the superscript denotes the order (power) with which each term depends
-on the perturbation $H_{I}(t)$. From Eq.~\eqref{trans} the $N$-th order
-term is 
+where $\tilde{\rho}_{0} = \tilde{\rho}(t = -\infty)$ is the unperturbed 
+density matrix. We look for the standard perturbation series solution, 
+$\tilde{\rho}(t) = \tilde{\rho}^{(0)} + \tilde{\rho}^{(1)} + 
+\tilde{\rho}^{(2)} + \cdots$, where the superscript denotes the order (power) 
+with which each term depends on the perturbation $H_{I}(t)$. From 
+Eq.~\eqref{trans} the $N$-th order term is 
 \begin{equation}
-\tilde{{\rho}}^{(N)}(t)=\frac{1}{i\hbar }\int_{-\infty }^{t}dt^{\prime }[\tilde{
-H}_{I}(t^{\prime }),\tilde{\rho}^{(N-1)}(t^{\prime })].  
+\tilde{\rho}^{(N)}(t) = 
+    \frac{1}{i\hbar}\int_{-\infty }^{t}dt^{\prime}
+    [\tilde{H}_{I}(t^{\prime }), \tilde{\rho}^{(N - 1)}(t^{\prime})].
 \label{rhop}
 \end{equation}
-The series is generated by the unperturbed density operator $\tilde{\rho}
-^{(0)}\equiv \tilde{\rho}_{0}$, assumed to be the diagonal Fermi-Dirac distribution, 
-$\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle=f(\hbar \omega_{n}(\mathbf{k}))\equiv f_{n}$. For a
-clean, cold semiconductor $f_{n}=1$ when $n$ is a valence ($v$) or
-occupied band, and zero when $n$ is a conduction ($c$) or empty band.
-We assume this for the remainder of the article.
-The expectation values must satisfy
-$
-\langle{\cal O}\rangle=
-\mbox{Tr}({\rho}{\cal O})
-=
-\mbox{Tr}(\tilde{{\rho}}\tilde{\cal O})
-$.
+The series is generated by the unperturbed density operator $\tilde{\rho}^{(0)}
+\equiv \tilde{\rho}_{0}$, assumed to be the diagonal Fermi-Dirac distribution, 
+$\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle = 
+f(\hbar\omega_{n}(\mathbf{k}))\equiv f_{n}$. For a clean, cold semiconductor 
+$f_{n}=1$ when $n$ is a valence ($v$) or occupied band, and zero when $n$ is 
+a conduction ($c$) or empty band. We assume this for the remainder of the 
+article. The expectation values must satisfy
+$\langle{\cal O}\rangle = \mbox{Tr}({\rho}{\cal O}) = 
+\mbox{Tr}(\tilde{\rho}\tilde{\cal O})$.
 
@@
 \begin{equation}
-\mathbf{J}=\langle{\mathbf{J}}\rangle=\frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}})
-,
+\mathbf{J} = \langle{\mathbf{J}}\rangle 
+           = \frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}}),
 \label{pe}

### 2014-11-21 paper-v-0.tex
removed unused equation numbers

@@
 $\mathbf{J}$, given by 
-\begin{equation}
+\begin{equation*}
 \mathbf{J} = \langle{\mathbf{J}}\rangle 
            = \frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}}),
-\label{pe}
-\end{equation}
+\end{equation*}
 where $\dot{\mathbf{r}}$ is the time derivative of the position operator of the
@@
 derivation remains the same. Then,
-\begin{equation}\label{h0}
+\begin{equation*}
 H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})
 =\frac{p^{2}}{2m_e}+V(\mathbf{r},\mathbf{p})
-\end{equation}
+\end{equation*}
 with $m_e$ the mass of the electron, $\mathbf{p}$ its canonical momentum, and 
@@
 $\omega_{n}^\Sigma(\mathbf{k})$ , are given by
-\begin{equation}
+\begin{equation*}
 \omega_{n}^\Sigma(\mathbf{k})=\omega^\mathrm{LDA}_{n}(\mathbf{k})+(1-f_{n})\Delta.
-\label{wese}
-\end{equation}
+\end{equation*}
 We emphasize that the scissored and unscissored Hamiltonian 
@@
 \begin{equation}
-H_{I}(t)=-e\mathbf{r}\cdot \mathbf{E}(t).  \label{rde}
+H_{I}(t)=-e\mathbf{r}\cdot \mathbf{E}(t).\label{rde}
 \end{equation}
@@
 the optical response. We use,\cite{aversaPRB95}
-\begin{equation}\label{conmri3n}
+\begin{equation}
 \langle n\mathbf{k} | [\mathbf{r}_i,{\cal O}] | m\mathbf{k}'\rangle
-=i\delta(\mathbf{k}-\mathbf{k}')({\cal O}_{nm})_{;\mathbf{k}}
-,
+=i\delta(\mathbf{k}-\mathbf{k}')({\cal O}_{nm})_{;\mathbf{k}},
+\label{conmri3n}
 \end{equation}  
 where
-\begin{equation}\label{gendevnn}
+\begin{equation}
 ({\cal O}_{nm})_{;\mathbf{k}}=
@@
 ,
+\label{gendevnn}
 \end{equation} 
@@
 Using Eq.~\eqref{hats} we obtain 
-\begin{equation}\label{chon.2} 
+\begin{equation}
 \mathbf{v}^S_{nm}=i\Delta f_{mn}\mathbf{r}_{nm},
+\label{chon.2} 
 \end{equation}
@@
 with 
-\begin{equation}\label{j.1}
+\begin{equation}
 B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\Sigma_{nm}-\tilde\omega},
+\label{j.1}
 \end{equation}
@@
 \begin{equation}
-\label{sz}
 {\mathbf{\cal C}}^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{b})  
             \Theta(z_\ell-z+\Delta_\ell^f),
+\label{sz}
 \end{equation} 
@@
 where 
-\begin{equation}\label{jmic3}
+\begin{equation}
 \mathbf{j}_{mn}(\mathbf{k};\mathbf{r})=
@@
 \right),
+\label{jmic3}
 \end{equation}
@@
 chosen region of the slab is given by
-\begin{equation}\label{jsz}
+\begin{equation*}
 \frac{1}{A}\int d^3r\, {\mathbf{\cal C}}(z)\, 
@@
  \equiv \boldsymbol{\mathcal{\cal J}}^{(N)}(t),
-\end{equation}
+\end{equation*}
 where $\boldsymbol{\mathcal{\cal J}}^{(N)}(t)$ is the $N$-th order current induced in the
@@
 Therefore we define
-\begin{equation}\label{vcal}
+\begin{equation}
 e{\boldsymbol{\mathcal{V}}}^{\Sigma}_{mn}(\mathbf{k})
@@
 \int d^3r\, {\mathbf{\cal C}}(z)\,\mathbf{j}_{mn}({\mathbf{k}};\mathbf{r}),
+\label{vcal}
 \end{equation}
 to write the Fourier transform of Eq.~\eqref{jmic} as
-\begin{equation}\label{jmac2}
+\begin{equation}
 \boldsymbol{\mathcal{\cal J}}^{(N)}(2\omega)=\frac{e}{A}
@@
 , 
+\label{jmac2}
 \end{equation}
@@
 we obtain the SH polarization of a given region as
-\begin{equation}\label{Pjikn}
+\begin{equation}
 \boldsymbol{\mathcal{\cal P}}^{(2)}(2\omega)
@@
 ,
+\label{Pjikn}
 \end{equation}
@@
 %\end{document}  
-\begin{equation}\label{chii}
+\begin{equation}
 \chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_i (-2\omega;\omega,\omega)
@@
 ,
+\label{chii}
 \end{equation} 
 related to intraband transitions, and 
-\begin{equation}\label{chie}
+\begin{equation}
 \chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_e (-2\omega;\omega,\omega)
@@
 \right),
+\label{chie}
 \end{equation} 
@@
 The generalized derivative in Eq.~\eqref{chii} is dealt with by the chain rule 
-\begin{equation}\label{gene2}
+\begin{equation}
 \left(\frac{f_{mn}r_{nm}^{\mathrm{b}}}{\omega^\Sigma_{nm}-\tilde\omega}\right)_{;k^{\mathrm{c}}}=
@@
 ,
+\label{gene2}
 \end{equation}
@@
 \begin{subequations}\label{chis}
-\begin{equation}\label{calvimchiewn}
+\begin{equation*}
 \mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}]= 
@@
 \right]\delta(\omega^\Sigma_{cv}-\omega),
-\end{equation}  
-\begin{equation}\label{calvimchiwn}
+\end{equation*}  
+\begin{equation*}
 \mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}]= 
@@
 \right]\delta(\omega^\Sigma_{cv}-\omega),
-\end{equation}
-\begin{equation}\label{calvimchie2wn}
+\end{equation*}
+\begin{equation*}
 \mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}]= 
@@
 \right]\delta(\omega^\Sigma_{cv}-2\omega),
-\end{equation}
-\begin{equation}\label{calvimchi2wn}
+\end{equation*}
+\begin{equation*}
 \mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}]= 
@@
 ,
-\end{equation}
+\end{equation*}
 \end{subequations}
@@
 Finally, we could also calculate the nonlinear surface susceptibility as 
-\begin{equation}\label{chiijksur}
+\begin{equation}
 \boldsymbol{\chi}(-2\omega;\omega,\omega)
 = \sum_{\{\ell\}}\boldsymbol{\chi}^\ell(-2\omega;\omega,\omega),
+\label{chiijksur}
 \end{equation} 

### 2014-11-21 paper-v-0.tex
removed unused align and equation numbers
missing final review

@@
 pseudopotentials, and then we replace $V$ with
-\begin{align}\label{ps}
+\begin{align*}
 V^{\mathrm{ps}}(\mathbf{r},\mathbf{p})=V(\mathbf{r})+V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})
 ,
-\end{align}
+\end{align*}
 where 
@@
 The unscissored and scissored Hamiltonians satisfy 
-\begin{eqnarray*}
-H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
-\label{hamils} \\
-H_{0}^\Sigma (\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\Sigma
-(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
-\end{eqnarray*}
+\begin{align*}
+H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=
+\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),\label{hamils} \\
+H_{0}^\Sigma (\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=
+\hbar \omega_{n}^\Sigma(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
+\end{align*}
 where the scissor-shifted energies, 
@@
 For $\mathbf{r}_e$ one uses
-\begin{align}\label{rnminn}
+\begin{align}
 \langle n\mathbf{k} | \mathbf{r}_e | m\mathbf{k}'\rangle &=
@@
 ,
+\label{rnminn}
 \end{align}
@@
 H^\Sigma_0$, we obtain
-\begin{align}\label{pmnrmn}
+\begin{align}
 \mathbf{r}_{e,nm}(\mathbf{k}) =
@@
 ,
+\label{pmnrmn}
 \end{align}  
@@
 Eq.~\eqref{mv} we obtain
-\begin{align}\label{vop2}
+\begin{align}
 \mathbf{v}^\Sigma&=
@@
 ,
+\label{vop2}
 \end{align}
 where we have defined 
-\begin{subequations}
-\begin{align}\label{ve}
+\begin{align}
 \mathbf{v} 
 &=\frac{\mathbf{p}}{m_e},
-\\\label{vnl}
+\nonumber\\
 \mathbf{v}^\mathrm{nl} 
 &=
-\frac{1}{i\hbar}[\mathbf{r},V^\mathrm{nl}],
-\\\label{vs}
+\frac{1}{i\hbar}[\mathbf{r},V^\mathrm{nl}],\label{vnl}
+\\
 \mathbf{v}^S
@@
 \frac{1}{i\hbar}[\mathbf{r}, S],
-\\\label{vlda}
+\nonumber\\
 \mathbf{v}^\mathrm{LDA} 
@@
 +\mathbf{v}^\mathrm{nl}
-,
+,\nonumber
 \end{align}  
-\end{subequations}
 with $\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ the canonical momentum operator, 
@@
 where we see that $\mathbf{v}^S_{nn}=0$. It follows that
-\begin{align}\label{chon.9}
+\begin{equation}
 \mathbf{v}^\Sigma_{nm}(\mathbf{k}) 
@@
 ,
-\end{align}
+\label{chon.9}
+\end{equation}
 and
- \begin{align}\label{chon.8}
+\begin{equation*}
 \frac{\mathbf{v}^\Sigma_{nm}(\mathbf{k})}{\omega^\Sigma_{nm}(\mathbf{k})}
-&=
+=
 \frac{\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k})}{\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
 ,
-\end{align}
+\end{equation*}
 and from Eq.~\eqref{pmnrmn},
-\begin{align}\label{chon.10}
+\begin{equation}
 \mathbf{r}_{nm}(\mathbf{k}) 
@@
 . 
-\end{align}
+\label{chon.10}
+\end{equation}
 The matrix elements 
@@
 and 
-\begin{eqnarray}\label{j.2}
-(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
+\begin{equation}
+(\tilde{\rho}^{(2)}(t))_{nm} = \frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
 i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
@@
 .
-\end{eqnarray}
+\label{j.2}
+\end{equation}
 We have used the fact that for a cold semiconductor $\partial
@@
 + | \mathbf{r}\rangle\langle\mathbf{r} | \mathbf{v}^\Sigma\right)$, that leads to
-\begin{align}\label{jmic}
+\begin{equation}
 \mathbf{j}^{(N)}(\mathbf{r},t)=\mathrm{Tr}(\mathbf{j}(\mathbf{r})\rho^{(N)}(t))
@@
 ,
-\end{align}
+\label{jmic}
+\end{equation}
 where 
@@
 Eqs.~\eqref{vcal} and \eqref{jmic3} we obtain
-\begin{align}\label{intj}
+\begin{align*}
 {\boldsymbol{\mathcal{V}}}^{\Sigma}_{mn}({\mathbf k})
@@
 ,
-\end{align}
+\end{align*}
 %Here an integration by parts is performed on the third term of the
@@
 where, we used the hermitian property of $\mathbf{v}^\Sigma$ and defined
-\begin{align}\label{nl.4}
+\begin{equation*}
 \boldsymbol{\mathcal{\cal V}}^{\Sigma}
@@
 .
-\end{align} 
+\end{equation*} 
 %where the superscript $\ell$ is inherited from ${\mathbf{\cal C}}(z)$.
@@
 We see that the replacement
-\begin{align}\label{vcali}
+\begin{equation}
 \mathbf{V} \to \boldsymbol{\mathcal{\cal V}}=\frac{{\mathbf{\cal C}}(z) \mathbf{V} +
@@
 ,
-\end{align} 
+\label{vcali}
+\end{equation} 
 is all that is needed to change any of the
@@
 thus
-\begin{align}\label{vopii}
+\begin{align*}
 \boldsymbol{\mathcal{\cal V}}^{\Sigma}
@@
 .
-\end{align}
+\end{align*}
 To calculate
@@
 leads to
-\begin{align}\label{Pjikn2}
+\begin{align}
 \chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega) 
@@
 ,
+\label{Pjikn2}
 \end{align}
@@
 we obtain
-\begin{align}\label{eli.13}
+\begin{equation}
 \left(\omega^\Sigma_{nm}\right)_{;k^{\mathrm{a}}}
@@
 .
-\end{align} 
+\end{equation} 
 The apparent divergence as $\tilde\omega\to 0$
@@
 \begin{subequations}\label{chis}
-\begin{equation*}
+\begin{equation}
 \mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}]= 
@@
 \right]\delta(\omega^\Sigma_{cv}-\omega),
-\end{equation*}  
-\begin{equation*}
+\end{equation}  
+\begin{equation}
 \mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}]= 
@@
 \right]\delta(\omega^\Sigma_{cv}-\omega),
-\end{equation*}
-\begin{equation*}
+\end{equation}
+\begin{equation}
 \mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}]= 
@@
 \right]\delta(\omega^\Sigma_{cv}-2\omega),
-\end{equation*}
-\begin{equation*}
+\end{equation}
+\begin{equation}
 \mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}]= 
@@
 ,
-\end{equation*}
+\end{equation}
 \end{subequations}
@@
 relationship is satisfied for this particular slab
-\begin{align}\label{hs}
+\begin{equation}
 \chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega) 
@@
 ,
-\end{align}
+\end{equation}
 where
@@
 Expanding the wave function in plane waves we obtain
-\begin{align}\label{eni.1}
+\begin{equation*}
 \psi_{n\mathbf{k}}(\mathbf{r})=\sum_\mathbf{G} A_{n\mathbf{k}}(\mathbf{G})e^{i(\mathbf{k}+\mathbf{G})\cdot\mathbf{r}}
 ,
-\end{align}
+\end{equation*}
 where $\{\mathbf{G}\}$ are the reciprocal basis vectors satisfying
@@
 we obtain,\cite{mendozaPRB06}
-\begin{align}\label{eni.2}
+\begin{equation}
 \boldsymbol{\mathcal{\cal V}}_{nm}(\mathbf{k})=
@@
 ,
-\end{align}   
+\label{eni.2}
+\end{equation}   
 where
-\begin{align}\label{vnl.9}
-f(G_\perp)=\frac{1}{L}\int{\mathbf{\cal C}}(z) e^{iG_\perp z}dz  
- ,
-\end{align}
+\begin{equation*}
+f(G_\perp)=\frac{1}{L}\int{\mathbf{\cal C}}(z) e^{iG_\perp z}dz,
+\end{equation*}
 with $f^*(G_\perp)=f(-G_\perp)$,
@@
 Likewise we obtain that
-\begin{align}\label{eni.4}
+\begin{equation}
 {\mathbf{\cal C}}_{nm}(\mathbf{k})=
@@
 .
-\end{align}  
+\label{eni.4}
+\end{equation}  
 If ${\mathbf{\cal C}}(z)=1$, then $f(G_\perp)=\delta_{G_\perp 0}$ and we 
@@
 one can show that
-\begin{align}\label{nmesn}
-({\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}&=
+\begin{equation}
+({\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}=
 {\cal T}_{nm}^{\mathrm{a}\mathrm{b}}
@@
 ,
-\end{align}
+\label{nmesn}
+\end{equation}
 where
-\begin{eqnarray}\label{tdel}
+\begin{equation*}
 \Delta^{\mathrm{a}}_{mn}
@@
 ,
-\end{eqnarray} 
-\begin{align}\label{tau.1}
+\end{equation*} 
+\begin{equation}
 {\cal T}_{nm}^{\mathrm{a}\mathrm{b}}
@@
 ,
-\end{align}   
+\label{tau.1}
+\end{equation}   
 and
-\begin{align}\label{lab}
+\begin{equation}
 L_{nm}^{\mathrm{a}\mathrm{b}}
@@
 .
-\end{align}  
+\label{lab}
+\end{equation}  
 The matrix elements $L^{\mathrm{a}\mathrm{b}}_{nm}$
@@
 $ and simply apply the chain rule,
-\begin{align}\label{rka}
+\begin{equation}
 (r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}
-&=
+=
 T^{\mathrm{a}\mathrm{b}}_{nm}
@@
 ,
-\end{align} 
+\label{rka}
+\end{equation} 
 where 
-\begin{eqnarray}\label{del}
+\begin{equation*}
 \Delta^{\mathrm{a}}_{mn}
@@
 ,
-\end{eqnarray}
+\end{equation*}
 and
-\begin{align}\label{tau.1n} 
+\begin{equation}
 T_{nm}^{\mathrm{a}\mathrm{b}}
@@
 .
-\end{align}
+\label{tau.1n} 
+\end{equation}
 Eq.~\eqref{rka} generalizes the usual expresion of
@@
 to obtain 
-\begin{align} 
+\begin{equation*} 
 ({\mathbf{\cal C}}_{nm})_{;\mathbf{k}}
-&=
+=
 i 
@@
 ,
-\end{align} 
+\end{equation*} 
 where we remind the reader that $\mathbf{r}_{nm}$ 
@@
 to obtain,
-\begin{align}\label{vnl.5}
+\begin{align*}
 \boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}(\mathbf{k})&=\frac{1}{2}
@@
 ,
-\end{align}  
+\end{align*}  
 where 
 we have defined  
-\begin{align}\label{vnl.10}
+\begin{align*}
 {\mathbf{\cal F}}_{n\mathbf{k}}(\mathbf{G}) 
-=
+&=
 \sum_{\mathbf{G}'} 
@@
 ,
-\end{align} 
-\begin{align}\label{vnl.11}
-H_{n\mathbf{k}}(\mathbf{G})&=
+H_{n\mathbf{k}}(\mathbf{G})\\
+&=
 \sum_{\mathbf{G}'} 
@@
 ,
-\end{align}
+\end{align*}
 and $\mathbf{K}=\mathbf{k}+\mathbf{G}$, $\mathbf{K}'=\mathbf{k}+\mathbf{G}'$.
@@
 From Eqs.~\eqref{lab} and \eqref{vnl} 
-\begin{align}\label{3.1}
+\begin{equation}
 L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
@@
 ,
-\end{align} 
+\end{equation} 
 and we expand the triple commutator to obtain,
-\begin{align}\label{3.5}
+\begin{align*}
 L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
@@
 ,
-\end{align} 
+\end{align*} 
 where 
@@
 We use the following identity
-\begin{align}\label{3.4}
+\begin{align*}
 &
@@
 ,
-\end{align}
+\end{align*}
 to write
-\begin{align}\label{3.7}
+\begin{equation*}
 L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k})
-&=
+=
 \frac{1}{\hbar^2}
@@
 | \mathbf{K}'\rangle,
-\nonumber\\
-\end{align} 
+\end{equation*} 
 where
-\begin{align}\label{vkk}
+\begin{equation*}
 \langle\mathbf{K} |
@@
 ,
-\end{align}
+\end{equation*}
 The double derivatives with respect to $\mathbf{K}$ and $\mathbf{K}'$ 
@@
 From Eq.~\eqref{vcali}
-\begin{align}\label{a.3b}
+\begin{equation}
 \boldsymbol{\mathcal{\cal V}}^{S}_{nm}
-&=
+=
 \frac{1}{2}\sum_q\left(   
@@
 ,
-\end{align}    
+\label{a.3b}
+\end{equation}    
 where $\sum_q | q\mathbf{k}\rangle\langle q\mathbf{k} | =1$ was used
@@
 the chain rule, we obtain
-\begin{align}\label{a.3bn}
+\begin{equation*}
 \left(\boldsymbol{\mathcal{\cal V}}^{S}_{nm}\right)_{;\mathbf{k}}
-&=
+=
 \frac{1}{2}\sum_q\left(
@@
 .
-\end{align}    
+\end{equation*}    
 Again, from
 Eq.~\eqref{chon.2}, 
-\begin{align}\label{choni.1}
+\begin{equation*}
 (\mathbf{v}^S_{nm})_{;\mathbf{k}}=i\Delta f_{mn}
@@
 ,
-\end{align}
+\end{equation*}
 a result that is in agreement with Eq. A(6) of Ref.~\onlinecite{cabellosPRB09}.

### 2014-11-21 paper-v-0.tex
to valeri

--- 2014-11-21 paper-v-0.tex
@@
-%\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
-\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,letterpaper]{revtex4}
+\documentclass[floatfix,prb,aps,superscriptaddress,11pt,preprint,letterpaper]{revtex4}
 \usepackage{amsfonts}
@@
 \usepackage{graphicx}
+\allowdisplaybreaks[1]%biutiful equation breaker!!!
 \usepackage{ulem}
 \usepackage{subfigure}
+%\input{/Users/bms/util/definitions}
+\def\chon{red}
+\def\nicolas{red}
+\input{definitions}
 \usepackage{bm}
+%%%%*****************%%%% fine hyperef 
 \usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %\usepackage{showlabels}
 %\usepackage{showkeys}
-\allowdisplaybreaks[1]%biutiful equation breaker!!!
 
@@
 
-\pacs{78.68.+m, 42.65.An, 71.15.Mb, 42.65.Ky, 78.66.-w}
-
 \begin{document}
 
-\title{Theory of surface second-harmonic generation for 
-       semiconductors including effects of nonlocal operators}
-
+\title{Theory of surface second-harmonic generation for semiconductors
+including effects of nonlocal operators}
 \author{Sean M. Anderson}
-    \affiliation{Centro de Investigaciones en Optica, 
-                Le\'on, Guanajuato, M\'exico}
+\affiliation{Centro de Investigaciones en Optica, Le\'on, Guanajuato, M\'exico}
 \author{Nicolas Tancogne-Dejean}
-    \affiliation{Laboratoire des Solides Irradi\'es, \'Ecole Polytechnique, 
-                CNRS, CEA/DSM, 91128 Palaiseau, France}
-    \affiliation{European Theoretical Spectroscopy Facility (ETSF), 
-                Palaiseau, France}
+\affiliation{Laboratoire des Solides Irradi\'es, \'Ecole 
+  Polytechnique, CNRS, CEA/DSM, 91128 Palaiseau, France}
+\affiliation{European Theoretical Spectroscopy Facility (ETSF), Palaiseau, France}
 \author{Bernardo S. Mendoza}\email{bms@cio.mx}
-    \affiliation{Centro de Investigaciones en Optica, 
-                Le\'on, Guanajuato, M\'exico}
+\affiliation{Centro de Investigaciones en Optica, Le\'on, Guanajuato, M\'exico}
 \author{Val\'erie V\'eniard}
-    \affiliation{Laboratoire des Solides Irradi\'es, \'Ecole Polytechnique, 
-                CNRS, CEA/DSM, 91128 Palaiseau, France}
-    \affiliation{European Theoretical Spectroscopy Facility (ETSF), 
-                Palaiseau, France}
+\affiliation{Laboratoire des Solides Irradi\'es, \'Ecole 
+  Polytechnique, CNRS, CEA/DSM, 91128 Palaiseau, France}
+ \affiliation{European Theoretical Spectroscopy Facility (ETSF), Palaiseau, France}
 
 \begin{abstract}
-We formulate a theoretical approach of surface second-harmonic generation from 
-semiconductor surfaces based on the length gauge. We calculate the nonlinear 
-second-order surface susceptibility tensor 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
-independent particle approximation, and for the first time in one unique 
-formulation we include i) the scissors correction, needed to have the correct 
-value for the energy band gap, ii) the derivation for the inclusion of the cut 
-function used to extract the surface nonlinear response, and iii) the 
-contribution of the nonlocal part of the pseudopotentials that are routinely 
-used in \textit{ab initio} band structure calculations. The first and third 
-terms are described by spatially nonlocal quantum mechanical operators and 
-are fully taken into account in the present formulation. As a test-case for 
-this approach, we calculate $\chi^{xxx}(-2\omega;\omega,\omega)$ for a clean 
-Si(001)$2\times 1$ reconstructed surface and find that this scheme for 
-extracting the surface response through the cut function is exact and readily
-implemented. The effects of the scissors correction and the nonlocal part of 
-the pseudopotentials are discussed for the first time for surface nonlinear 
-optics. The scissors correction non-rigidly shifts the spectrum to higher 
-energies by mixing the $1\omega$ and $2\omega$ resonances and exerts a strong 
-influence on the line-shape. The effects of the nonlocal part of the 
-pseudopotentials keeps the same line-shape of 
-$|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$ but reduces its value by 
-15-20\%. Therefore, our scheme is unprecedented through the inclusion of 
-these three contributions and also opens the possibility to study SSHG with 
-more versatility and more accuracy than previous approaches. 
+We formulate a theoretical approach of surface second-harmonic generation 
+from semiconductor surfaces based on the length gauge. 
+Within the independent particle approximation 
+the nonlinear second-order surface susceptibility tensor
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$   
+is calculated, including for the first time in one unique formulation
+i) the scissors correction, needed to have the correct value of the
+energy band gap,
+ii) the derivation for the inclusion of the cut function, 
+used to extract the surface nonlinear response,
+and iii) the contribution of the nonlocal part of the 
+pseudopotentials, routinely  used in \textit{ab initio} band structure calculations. 
+First and third terms are described by spatially nonlocal quantum
+mechanical operators and are fully taken into account in the present formulation. 
+As a test-case of the approach we calculate 
+$\chi^{xxx}(-2\omega;\omega,\omega)$ for the clean Si(001)$2\times 1$
+reconstructed surface and find that our scheme to extract the surface
+through the cut function
+readily works and is exact. The effects of the scissors correction and
+of the nonlocal part of the pseudopotentials are discussed for the
+first time in surface nonlinear optics. 
+The scissors correction shifts the spectrum to 
+higher energies though the shifting is not rigid and mixes the 
+$1\omega$ and $2\omega$ resonances, 
+and has a strong influence in the line-shape. 
+The 
+effects of the nonlocal part of the pseudopotentials 
+keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\go;\go,\go)|$, but 
+reduces its value
+by 15-20\%.  
+Therefore, the inclusion of the three aforementioned 
+contributions is very important and 
+makes our scheme unprecedented and opens the possibility to 
+study SSHG with more versatility and providing more accurate results 
+than previous approaches. 
 \end{abstract}
@@
 calculation and thus one uses the much simpler scissors operator scheme.
-\cite{levinePRL89,levinePRL91,delsolePRB93} This allows us to ``open'' the 
-DFT-LDA gap to its correct experimental or GW value even when experiments 
-are not available. In this context, to correct for the underestimation of 
-the energy band gap of semiconductors Nastos et al.\cite{nastosPRB05} used 
-the ``length gauge'' or ``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to 
-correctly include the many-body corrections through the scissors operator in 
-the SH susceptibility. Later, Cabellos \textit{et al}.\cite{cabellosPRB09} 
-elaborated a derivation of the ``velocity gauge'' or 
-``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the scissors operator 
-and proved gauge invariance with respect to the length gauge. From these works 
-it is clear the length gauge is a much better staring point to obtain the 
-surface SH (SSH) susceptibility, as will be elaborated in this article. 
-However, these considerations are only valid for bulk semiconductors.
+\cite{levinePRL89,levinePRL91,delsolePRB93}  
+This allows us to ``open'' the DFT-LDA gap to 
+its correct experimental or GW value even when experiments are not available. 
+In this context, to correct for the underestimation of the energy band gap of 
+semiconductors Nastos et al.\cite{nastosPRB05} used the ``length gauge'' or 
+``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to correctly include the 
+many-body corrections through the scissors operator in the SH susceptibility.
+Later, Cabellos \textit{et al}.\cite{cabellosPRB09} elaborated a derivation 
+of the ``velocity gauge'' or ``$\bfA\cdot\bfv$'' gauge properly including the 
+scissors operator and proved gauge invariance with respect to the length 
+gauge. From these works it is clear the length gauge is a much better staring
+point to obtain the surface SH (SSH) susceptibility, as will be elaborated 
+in this article. However, these considerations are only valid for bulk 
+semiconductors.
 
@@
 Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02,sanoPRB02}
-followed upon this work and in particular Ref.~\onlinecite{arzatePRB01} went 
-into a  detailed analysis of the different contributions to the SHG spectra 
-of a surface and the nuanced relationship between bulk, surface, interband, 
+followed 
+upon this work and in particular Ref.~\onlinecite{arzatePRB01} went into a 
+detailed analysis of the different contributions to the SHG spectra of a 
+surface and the nuanced relationship between bulk, surface, interband, 
 intraband, $1\omega$ and $2\omega$ terms, and
@@
 Finally, DFT-LDA calculations are often based on the use of pseudopotentials.
-As we will discuss later in this article, the presence of the nonlocal part 
-of the pseudopotential introduces corrections to the momentum operator of 
-the electron that have to be included with care into the SSH susceptibility. 
-So far, this contribution has been neglected in the literature.
+As it will be discussed 
+in this article, the presence of a nonlocal part of the
+pseudopotential introduces
+corrections to
+the momentum operator of the electron that have to be included with care in 
+the SSH susceptibility. So far, this contribution has been neglected in the 
+literature.
 
@@
 length gauge formalism. This derivation includes the addition of the terms 
-mentioned above that have been absent in previous works. If desired, we can 
-use $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ for a 
-layer-by-layer analysis. In Sec. \ref{results} we show the results of this 
-reformulation with a study of a clean Si(001) surface with a $2 \times 1$ 
-reconstruction that proves the validity of our approach with a special 
-test-case that has never been implemented before. We compare results from 
-before and after adding the different nonlocal contributions. Finally, in 
-Sec. \ref{conc} we submit our conclusions.
+mentioned above that have been absent in previous works. Also, our 
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ can be used 
+for a layer-by-layer analysis if desired. In Sec. \ref{results} we show the 
+results of this reformulation with a study of a clean Si(001) surface with a 
+$2 \times 1$ reconstruction, 
+proving the correctness of our approach with a special test-case,
+never exploited before.
+We compare results from before and after adding 
+the different nonlocal contributions. Finally, in Sec. \ref{conc} we give 
+our conclusions.
 
@@
 
-In this section we present the scheme used to calculate the surface 
-second-order nonlinear response using the length gauge formalism. Our 
-derivation includes the aforementioned terms not included before for 
-this response.
+In this section we present the scheme used to calculate the surface second-order
+nonlinear response using the length gauge formalism.
+Our derivation includes the aforementioned terms not included
+before for the surface nonlinear
+response.
 
@@
 
-We assume the IPA, a classical electromagnetic field, and quantum mechanical 
-matter. We neglect local field and excitonic effects. We can describe the 
-system using 
+We assume the IPA and neglect local field  
+and excitonic effects.  
+We assume a classical electromagnetic field
+and quantum mechanical matter.
+We can describe the system using 
 %a scaled 
-the one electron density operator ${\rho}$, with which we can calculate the 
-expectation value of a single-particle observable $\mathcal{O}$ as 
-$\langle{\cal O}\rangle = \mbox{Tr}({\rho}{\cal O}),\label{calo}$
-with $\mathcal{O}$ the associated quantum mechanical operator and Tr the 
-trace. The density operator satisfies
-$i\hbar(d{\rho}/dt) = [H(t), {\rho}], \label{rho}$
+the one electron density
+operator ${\rho}$, with which we can calculate the expectation value of a
+single-particle observable $\mathcal{O}$ as 
+$\langle{\cal O}\rangle=\mbox{Tr}({\rho} {\cal O}),  
+\label{calo}$
+with $\mathcal{O}$ the associated quantum mechanical operator and Tr
+the trace. 
+The density operator satisfies
+$
+i\hbar (d{\rho}/dt)=[H(t),{\rho}],  
+\label{rho}
+$
 with $H(t)$ as the total single electron Hamiltonian, written as 
 \begin{equation*}
-H(t) = H_{0} + H_{I}(t),  
+H(t)=H_{0}+H_{I}(t),  
 \label{ache}
@@
 where $H_{0}$ is the unperturbed time-independent Hamiltonian, and $H_{I}(t)$
-is the time-dependent potential energy due to the interaction of the electron 
-with the electromagnetic field.
+is the time-dependent potential energy due to the interaction of the
+electron with the electromagnetic field.
 % ; $H_{0}$ has eigenvalues
@@
 % index $n$ and crystal momentum $\mathbf{k}$.
-To proceed with the solution of $\rho$ it is convenient to use the interaction 
-picture, where a unitary operator $U = \exp({iH_{0}t/\hbar})$ transforms any 
-operator $\mathcal{O}$ into $\tilde{\mathcal{O}} = U{\cal O}U^{\dagger}$. 
-Even if $\mathcal{O}$ is time-independent, $\tilde{\mathcal{O}}$ is 
-time-dependent through the explicit time dependence of $U$. The dynamical 
-equation for $\tilde{\rho}$ is given by
+To proceed with the solution of $\rho$  it is convenient to use the
+interaction picture, where a unitary operator $U=\exp ({iH_{0}t/\hbar })$
+transforms any operator $\mathcal{O}$ into 
+$\tilde{\mathcal{O}}=U{\cal O}U^{\dagger }$. Even if $\mathcal{O}$ 
+is time-independent, 
+$\tilde{\mathcal{O}}$ is time-dependent through the explicit
+time dependence of $U$. 
+The dynamical 
+equation for $\tilde{\rho}$ is 
+given by
 \begin{equation*}
-i\hbar \frac{d\tilde{\rho}}{dt} = [\tilde{H}_{I}(t), \tilde{\rho}],
+i\hbar \frac{d\tilde{{\rho}}}{dt}=[\tilde{H}_{I}(t),\tilde{{\rho}}],  
 \label{rho1}
@@
 \begin{equation}
-i\hbar \tilde{\rho}(t) = 
-    i\hbar \tilde{\rho}_{0} + 
-    \int_{-\infty}^{t}dt^{\prime}
-    [\tilde{H}_{I}(t^{\prime}),\tilde{\rho}(t^{\prime})],  
+i\hbar \tilde{{\rho}}(t)=i\hbar \tilde{{\rho}}_{0}+\int_{-\infty }^{t}dt^{\prime }[
+\tilde{H}_{I}(t^{\prime }),\tilde{\rho}(t^{\prime })],  
 \label{trans}
 \end{equation}
-where $\tilde{\rho}_{0} = \tilde{\rho}(t = -\infty)$ is the unperturbed 
-density matrix. We look for the standard perturbation series solution, 
-$\tilde{\rho}(t) = \tilde{\rho}^{(0)} + \tilde{\rho}^{(1)} + 
-\tilde{\rho}^{(2)} + \cdots$, where the superscript denotes the order (power) 
-with which each term depends on the perturbation $H_{I}(t)$. From 
-Eq.~\eqref{trans} the $N$-th order term is 
+where $\tilde{\rho}_{0}=\tilde{\rho}(t=-\infty )$ is the unperturbed density matrix. We look
+for the standard perturbation series solution, 
+$\tilde{\rho}(t)=\tilde{\rho}^{(0)}+\tilde{\rho}^{(1)}+\tilde{\rho}^{(2)}+\cdots$,  
+where the superscript denotes the order (power) with which each term depends
+on the perturbation $H_{I}(t)$. From Eq.~\eqref{trans} the $N$-th order
+term is 
 \begin{equation}
-\tilde{\rho}^{(N)}(t) = 
-    \frac{1}{i\hbar}\int_{-\infty }^{t}dt^{\prime}
-    [\tilde{H}_{I}(t^{\prime }), \tilde{\rho}^{(N - 1)}(t^{\prime})].
+\tilde{{\rho}}^{(N)}(t)=\frac{1}{i\hbar }\int_{-\infty }^{t}dt^{\prime }[\tilde{
+H}_{I}(t^{\prime }),\tilde{\rho}^{(N-1)}(t^{\prime })].  
 \label{rhop}
 \end{equation}
-The series is generated by the unperturbed density operator $\tilde{\rho}^{(0)}
-\equiv \tilde{\rho}_{0}$, assumed to be the diagonal Fermi-Dirac distribution, 
-$\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle = 
-f(\hbar\omega_{n}(\mathbf{k}))\equiv f_{n}$. For a clean, cold semiconductor 
-$f_{n}=1$ when $n$ is a valence ($v$) or occupied band, and zero when $n$ is 
-a conduction ($c$) or empty band. We assume this for the remainder of the 
-article. The expectation values must satisfy
-$\langle{\cal O}\rangle = \mbox{Tr}({\rho}{\cal O}) = 
-\mbox{Tr}(\tilde{\rho}\tilde{\cal O})$.
+The series is generated by the unperturbed density operator $\tilde{\rho}
+^{(0)}\equiv \tilde{\rho}_{0}$, assumed to be the diagonal Fermi-Dirac distribution, 
+$\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle=f(\hbar \omega_{n}(\mathbf{k}))\equiv f_{n}$. For a
+clean, cold semiconductor $f_{n}=1$ when $n$ is a valence ($v$) or
+occupied band, and zero when $n$ is a conduction ($c$) or empty band.
+We assume this for the remainder of the article.
+The expectation values must satisfy
+$
+\langle{\cal O}\rangle=
+\mbox{Tr}({\rho}{\cal O})
+=
+\mbox{Tr}(\tilde{{\rho}}\tilde{\cal O})
+$.
 
@@
 \begin{equation}
-\mathbf{J} = \langle{\mathbf{J}}\rangle 
-           = \frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}}),
+\bfJ=\langle{\mathbf{J}}\rangle=\frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}})
+,
 \label{pe}
@@
 area of the unit cell. We calculate the polarization density 
-$\mathbf{P}$, related to $\mathbf{J}$ by
-$\mathbf{J}=d\mathbf{P}/dt$. For a 
+$\mathbf{P}$, related to $\bfJ$ by
+$\bfJ=d\mathbf{P}/dt$. For a 
 perturbing electromagnetic field, $\mathbf{E}(t)=
@@
 \begin{equation}
-{\mathbf{\cal P}}^a(2\omega)=\chi ^{abc}(-2\omega;\omega,\omega)E^{b}(\omega)E^{c}(\omega),  
+\calp^a(2\omega)=\chi ^{abc}(-2\omega;\omega,\omega)E^{b}(\omega)E^{c}(\omega),  
 \label{pshg}
@@
 \begin{equation}\label{h0}
-H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})
-=\frac{p^{2}}{2m_e}+V(\mathbf{r},\mathbf{p})
+H^{\mathrm{LDA}}_{0}(\bfr,\bfp)
+=\frac{p^{2}}{2m_e}+V(\bfr,\bfp)
 \end{equation}
@@
 \begin{align}\label{ps}
-V^{\mathrm{ps}}(\mathbf{r},\mathbf{p})=V(\mathbf{r})+V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})
+V^{\mathrm{ps}}(\bfr,\bfp)=V(\bfr)+V^{\mathrm{nl}}(\bfr,\bfp)
 ,
@@
 where 
-$V(\mathbf{r})$ and $V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})$  are the local and nonlocal parts, 
+$V(\bfr)$ and $V^{\mathrm{nl}}(\bfr,\bfp)$  are the local and nonlocal parts, 
  respectively.
 In case of
-a local potential, i.e. $V= V(\mathbf{r})$, 
+a local potential, i.e. $V= V(\bfr)$, 
 like that of all-electron schemes, 
@@
 \begin{equation*}
-H^\Sigma_{0}(\mathbf{r},\mathbf{p})=H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})+S(\mathbf{r},\mathbf{p})
+H^\gS_{0}(\bfr,\bfp)=H^{\mathrm{LDA}}_{0}(\bfr,\bfp)+S(\bfr,\bfp)
 \end{equation*}
@@
 \begin{equation}
-S(\mathbf{r},\mathbf{p})=\hbar \Delta\sum_{n}\int d^{3}k(1-f_{n})
+S(\bfr,\bfp)=\hbar \gD\sum_{n}\int d^{3}k(1-f_{n})
 |n\mathbf{k}\rangle\langle n\mathbf{k}|,
@@
 \end{equation}
-with $\hbar \Delta$  the rigid ($\mathbf{k}$-independent) energy correction to be
+with $\hbar \gD$  the rigid ($\mathbf{k}$-independent) energy correction to be
 applied. 
@@
 \begin{eqnarray*}
-H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
+H^{\mathrm{LDA}}_{0}(\bfr,\bfp)\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
 \label{hamils} \\
-H_{0}^\Sigma (\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\Sigma
+H_{0}^\gS (\bfr,\bfp)\psi _{n\mathbf{k}}(\mathbf{r}) &=&\hbar \omega_{n}^\gS
 (\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
@@
 where the scissor-shifted energies, 
-$\omega_{n}^\Sigma(\mathbf{k})$ , are given by
+$\omega_{n}^\gS(\mathbf{k})$ , are given by
 \begin{equation}
-\omega_{n}^\Sigma(\mathbf{k})=\omega^\mathrm{LDA}_{n}(\mathbf{k})+(1-f_{n})\Delta.
+\omega_{n}^\gS(\mathbf{k})=\omega^\lda_{n}(\mathbf{k})+(1-f_{n})\gD.
 \label{wese}
@@
 have the same eigenfunctions, where
-$\psi_{n\mathbf{k}}(\mathbf{r})=\langle\mathbf{r}|n\mathbf{k}\rangle=e^{i\mathbf{k}\cdot\mathbf{r}}u_{n\mathbf{k}}(\mathbf{r})$,
-are the real space representations of the Bloch states $| n\mathbf{k}\rangle$ labelled 
-by the band index $n$ and the crystal momentum $\mathbf{k}$, and $u_{n\mathbf{k}}(\mathbf{r})$
+$\psi_{n\bfk}(\bfr)=\braket{\bfr}{n\bfk}=e^{i\bfk\cdot\bfr}u_{n\bfk}(\bfr)$,
+are the real space representations of the Bloch states $\ket{n\bfk}$ labelled 
+by the band index $n$ and the crystal momentum $\bfk$, and $u_{n\bfk}(\bfr)$
 is cell periodic.
@@
 treatment of the position operator
-$\mathbf{r}$ for extended Bloch states, $\langle\mathbf{r}|n\mathbf{k}\rangle$ has 
+$\bfr$ for extended Bloch states, $\braket{\bfr}{n\bfk}$ has 
 been described in detail.\cite{adamsJCP53,blountSSP62} 
 In particular, Ref.~\onlinecite{aversaPRB95} develops
-the correct treatment of $\mathbf{r}$ for
+the correct treatment of $\bfr$ for
 the calculation of the nonlinear optical response.
-$\mathbf{r}$ is split into the {\it intraband} 
-($\mathbf{r}_i$) and {\it interband} ($\mathbf{r}_e$) parts, where 
-$\mathbf{r}=\mathbf{r}_i+\mathbf{r}_e$. 
-For $\mathbf{r}_e$ one uses
+$\bfr$ is split into the {\it intraband} 
+($\bfr_i$) and {\it interband} ($\bfr_e$) parts, where 
+$\bfr=\bfr_i+\bfr_e$. 
+For $\bfr_e$ one uses
 \begin{align}\label{rnminn}
-\langle n\mathbf{k} | \mathbf{r}_e | m\mathbf{k}'\rangle &=
-(1-\delta_{nm})\delta(\mathbf{k}-\mathbf{k}')\boldsymbol{\xi}_{nm}(\mathbf{k})
-,
-\end{align}
-such that $\mathbf{r}_{e,nm}=0$ for $n=m$.
+\bra{n\bfk}\bfr_e\ket{m\bfk'} &=
+(1-\gd_{nm})\gd(\bfk-\bfk')\bfgxi_{nm}(\bfk) 
+,
+\end{align}
+such that $\bfr_{e,nm}=0$ for $n=m$.
 From Eq.~\eqref{mv} with $H_0\to
-H^\Sigma_0$, we obtain
+H^\gS_0$, we obtain
 \begin{align}\label{pmnrmn}
-\mathbf{r}_{e,nm}(\mathbf{k}) =
-\boldsymbol{\xi}_{nm}(\mathbf{k})\equiv 
-\mathbf{r}_{nm}(\mathbf{k}) 
+\bfr_{e,nm}(\bfk) =
+\bfgxi_{nm}(\bfk)\equiv 
+\bfr_{nm}(\bfk) 
 &=
-\frac{\mathbf{v}^\Sigma_{nm}(\mathbf{k})}{i\omega^\Sigma_{nm}(\mathbf{k})}
+\frac{\bfv^\gS_{nm}(\bfk)}{i\go^\gS_{nm}(\bfk)}
 \quad\quad n\notin D_m 
@@
 where we defined
-$\omega^\Sigma_{nm}(\mathbf{k})\equiv\omega^\Sigma_n(\mathbf{k})-\omega^\Sigma_m(\mathbf{k})$, and
+$\go^\gS_{nm}(\bfk)\equiv\go^\gS_n(\bfk)-\go^\gS_m(\bfk)$, and
 $D_m$ are all the possible degenerate $m$-states. 
 %We recognize that for each instance in which 
-%$n\notin D_m$ or $n\ne m$, $\mathbf{r}_{nm}(\mathbf{k})$ are interband matrix
+%$n\notin D_m$ or $n\ne m$, $\bfr_{nm}(\bfk)$ are interband matrix
 %elements.
-For the intraband part, $\mathbf{r}_i$ only appears in
+For the intraband part, $\bfr_i$ only appears in
 commutators during the derivation of
@@
 \begin{equation}\label{conmri3n}
-\langle n\mathbf{k} | [\mathbf{r}_i,{\cal O}] | m\mathbf{k}'\rangle
-=i\delta(\mathbf{k}-\mathbf{k}')({\cal O}_{nm})_{;\mathbf{k}}
+\bra{n\bfk}[\bfr_i,\calo]\ket{m\bfk'}
+=i\gd(\bfk-\bfk')(\calo_{nm})_{;\bfk}
 ,
@@
 \begin{equation}\label{gendevnn}
-({\cal O}_{nm})_{;\mathbf{k}}=
-\nabla_\mathbf{k} 
-{\cal O}_{n m}(\mathbf{k}) 
+(\calo_{nm})_{;\bfk}=
+\nabla_\bfk 
+\calo_{n m}(\bfk) 
 - 
 i 
-{\cal O}_{nm}(\mathbf{k}) 
+\calo_{nm}(\bfk) 
 \left(
-\boldsymbol{\xi}_{nn}(\mathbf{k}) 
+\bfgxi_{nn}(\bfk) 
 -
-\boldsymbol{\xi}_{mm}(\mathbf{k}) 
+\bfgxi_{mm}(\bfk) 
 \right) 
@@
 the generalized derivative 
-of the operator ${\cal O}$. 
-The vectors $\boldsymbol{\xi}_{nn}(\mathbf{k})$ are defined in 
+of the operator $\calo$. 
+The vectors $\bfgxi_{nn}(\bfk)$ are defined in 
 Ref.~\onlinecite{aversaPRB95} though they do not need to be 
@@
 we derive a key result for the length gauge formulation. 
-Again, using $H^\Sigma_0$ in
+Again, using $H^\gS_0$ in
 Eq.~\eqref{mv} we obtain
 \begin{align}\label{vop2}
-\mathbf{v}^\Sigma&=
-\mathbf{v} 
-+
-\mathbf{v}^\mathrm{nl} 
-+\mathbf{v}^S
+\bfv^\gS&=
+\bfv 
++
+\bfv^\nl 
++\bfv^S
 =
-\mathbf{v}^\mathrm{LDA} 
-+\mathbf{v}^S 
+\bfv^\lda 
++\bfv^S 
 ,
@@
 \begin{align}\label{ve}
-\mathbf{v} 
-&=\frac{\mathbf{p}}{m_e},
+\bfv 
+&=\frac{\bfp}{m_e},
 \\\label{vnl}
-\mathbf{v}^\mathrm{nl} 
+\bfv^\nl 
 &=
-\frac{1}{i\hbar}[\mathbf{r},V^\mathrm{nl}],
+\frac{1}{i\hbar}[\bfr,V^\nl],
 \\\label{vs}
-\mathbf{v}^S
+\bfv^S
 &=
-\frac{1}{i\hbar}[\mathbf{r}, S],
+\frac{1}{i\hbar}[\bfr, S],
 \\\label{vlda}
-\mathbf{v}^\mathrm{LDA} 
+\bfv^\lda 
 &=
-\mathbf{v} 
-+\mathbf{v}^\mathrm{nl}
+\bfv 
++\bfv^\nl
 ,
@@
 \end{subequations}
-with $\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ the canonical momentum operator, 
-and using $[r^a,p^b]=i\hbar\delta_{ab}$, where $\delta_{ab}$ is the Kronecker delta.
+with $\bfp=-i\hbar\bfgnabla$ the canonical momentum operator, 
+and using $[r^a,p^b]=i\hbar\gd_{ab}$, where $\gd_{ab}$ is the Kronecker delta.
 Using Eq.~\eqref{hats} we obtain 
 \begin{equation}\label{chon.2} 
-\mathbf{v}^S_{nm}=i\Delta f_{mn}\mathbf{r}_{nm},
+\bfv^S_{nm}=i\gD f_{mn}\bfr_{nm},
 \end{equation}
 with $f_{nm}\equiv f_n-f_m$,
-where we see that $\mathbf{v}^S_{nn}=0$. It follows that
+where we see that $\bfv^S_{nn}=0$. It follows that
 \begin{align}\label{chon.9}
-\mathbf{v}^\Sigma_{nm}(\mathbf{k}) 
-=\frac{\omega^\Sigma_{nm}(\mathbf{k})}{\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
-\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k}) 
+\bfv^\gS_{nm}(\bfk) 
+=\frac{\go^\gS_{nm}(\bfk)}{\go^\lda_{nm}(\bfk)}
+\bfv^\lda_{nm}(\bfk) 
 ,
@@
  \begin{align}\label{chon.8}
-\frac{\mathbf{v}^\Sigma_{nm}(\mathbf{k})}{\omega^\Sigma_{nm}(\mathbf{k})}
+\frac{\bfv^\gS_{nm}(\bfk)}{\go^\gS_{nm}(\bfk)}
 &=
-\frac{\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k})}{\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
+\frac{\bfv^\lda_{nm}(\bfk)}{\go^\lda_{nm}(\bfk)}
 ,
@@
 \begin{align}\label{chon.10}
-\mathbf{r}_{nm}(\mathbf{k}) 
+\bfr_{nm}(\bfk) 
 =
-\frac{\mathbf{v}^\Sigma_{nm}(\mathbf{k})}{i\omega^\Sigma_{nm}(\mathbf{k})}
+\frac{\bfv^\gS_{nm}(\bfk)}{i\go^\gS_{nm}(\bfk)}
 =
-\frac{\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k})}{i\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
+\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
 \quad\quad n\notin D_m 
@@
 The matrix elements 
-of $\mathbf{r}_{nm}(\mathbf{k})$ are identical using either
+of $\bfr_{nm}(\bfk)$ are identical using either
 the LDA or scissored
@@
 Of course, it is more convenient to calculate them
-through $\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k})$ which 
+through $\bfv^\lda_{nm}(\bfk)$ which 
 includes only the contribution of 
-$\mathbf{v}^\mathrm{nl}_{nm}(\mathbf{k})$. These can be readily
+$\bfv^\nl_{nm}(\bfk)$. These can be readily
 calculated
@@
 we need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
-$\omega_{n}^{\Sigma}$ at the end of the derivation
+$\omega_{n}^{\gS}$ at the end of the derivation
 to obtain the scissored results for any 
@@
 Eq.~\eqref{rde}, we obtain 
-$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\Sigma_{nm}-\tilde\omega)t}$,
+$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\gS_{nm}-\tilde\omega)t}$,
 with 
 \begin{equation}\label{j.1}
-B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\Sigma_{nm}-\tilde\omega},
+B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\gS_{nm}-\tilde\omega},
 \end{equation}
@@
 \begin{eqnarray}\label{j.2}
-(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
+(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\gS_{nm}-2\tilde\omega}\bigg[
 i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
 \Big)  
--(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\Sigma_{nm}-2\tilde\omega)t}
+-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\gS_{nm}-2\tilde\omega)t}
 .
@@
 other. This is due to the $f_{nm}$ factor in Eq.~\eqref{j.1}, 
-and therefore $B^a_{nn}=0$. The dependence on $\mathbf{k}$ 
+and therefore $B^a_{nn}=0$. The dependence on $\bfk$ 
 of all quantities is implicitly understood from 
@@
 centrosymmetric. This would imply that 
-$\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}=0$ so we must
+$\chi^{\mathrm{slab},\rma\rmb\rmc}=0$ so we must
 find a scheme 
-in order to have a finite $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$ representative of the
+in order to have a finite $\chi^{\rma\rmb\rmc}$ representative of the
 surface. Even if the front and back surfaces of the slab 
 are different, breaking the centrosymmetry and therefore giving an
-overall $\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}\ne 0$; we still
-need a procedure to extract the front surface $\chi^{f,\mathrm{a}\mathrm{b}\mathrm{c}}$
-and the back surface $\chi^{b,\mathrm{a}\mathrm{b}\mathrm{c}}$ from the slab
+overall $\chi^{\mathrm{slab},\rma\rmb\rmc}\ne 0$; we still
+need a procedure to extract the front surface $\chi^{f,\rma\rmb\rmc}$
+and the back surface $\chi^{b,\rma\rmb\rmc}$ from the slab
 susceptibility 
-$\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}=\chi^{f,\mathrm{a}\mathrm{b}\mathrm{c}}-\chi^{b,\mathrm{a}\mathrm{b}\mathrm{c}}$.
-We have omitted the frequency arguments of $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$ for 
+$\chi^{\mathrm{slab},\rma\rmb\rmc}=\chi^{f,\rma\rmb\rmc}-\chi^{b,\rma\rmb\rmc}$.
+We have omitted the frequency arguments of $\chi^{\rma\rmb\rmc}$ for 
 convenience of notation.
@@
 A convenient way to accomplish the separation of the SH signal of
-either surface is to introduce a ``cut function'', ${\mathbf{\cal C}}(z)$, which is 
+either surface is to introduce a ``cut function'', $\calc(z)$, which is 
 usually taken to be unity over one half of the slab and zero over 
 the other half.\cite{reiningPRB94}
-In this case ${\mathbf{\cal C}}(z)$ will give the contribution of the 
-side of the slab for which ${\mathbf{\cal C}}(z)=1$. 
+In this case $\calc(z)$ will give the contribution of the 
+side of the slab for which $\calc(z)=1$. 
 As was done for the linear response,\cite{mendozaPRB06}
 we can generalize this 
-simple choice for ${\mathbf{\cal C}}(z)$ by a top-hat cut function
-${\mathbf{\cal C}}^\ell(z)$ that selects a given layer,
+simple choice for $\calc(z)$ by a top-hat cut function
+$\calc^\ell(z)$ that selects a given layer,
 \begin{equation}
 \label{sz}
-{\mathbf{\cal C}}^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{b})  
+\calc^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{b})  
             \Theta(z_\ell-z+\Delta_\ell^f),
@@
 To introduce the
-cut function ${\mathbf{\cal C}}(z)$ in
-the calculation of $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$, we start from 
+cut function $\calc(z)$ in
+the calculation of $\chi^{\rma\rmb\rmc}$, we start from 
 the operator for the electron current,
-$\mathbf{j}(\mathbf{r})=\frac{e}{2}\left(\mathbf{v}^\Sigma | \mathbf{r}\rangle\langle\mathbf{r} |
-+ | \mathbf{r}\rangle\langle\mathbf{r} | \mathbf{v}^\Sigma\right)$, that leads to
+$\bfj(\bfr)=\frac{e}{2}\left(\bfv^\gS\ket{\bfr}\bra{\bfr}
++\ket{\bfr}\bra{\bfr}\bfv^\gS\right)$, that leads to
 \begin{align}\label{jmic}
-\mathbf{j}^{(N)}(\mathbf{r},t)=\mathrm{Tr}(\mathbf{j}(\mathbf{r})\rho^{(N)}(t))
+\bfj^{(N)}(\bfr,t)=\mathrm{Tr}(\bfj(\bfr)\rho^{(N)}(t))
 =
@@
 \sum_{nm}
-\rho^{(N)}_{nm}(\mathbf{k};t)\mathbf{j}_{mn}(\mathbf{k};\mathbf{r})
+\rho^{(N)}_{nm}(\bfk;t)\bfj_{mn}(\bfk;\bfr)
 ,
@@
 \begin{equation}\label{jmic3}
-\mathbf{j}_{mn}(\mathbf{k};\mathbf{r})=
+\bfj_{mn}(\bfk;\bfr)=
 \frac{e}{2}
 \left(
-\langle m\mathbf{k} | \mathbf{v}^\Sigma | \mathbf{r}\rangle \langle\mathbf{r}|n\mathbf{k}\rangle
-+
-\langle m\mathbf{k}|\mathbf{r}\rangle\langle\mathbf{r} | \mathbf{v}^\Sigma | n\mathbf{k}\rangle
+\bra{m\bfk}\bfv^\gS\ket{\bfr}\braket{\bfr}{n\bfk}
++
+\braket{m\bfk}{\bfr}\bra{\bfr}\bfv^\gS\ket{n\bfk}
 \right),
@@
 are the matrix elements of the microscopic current operator.
-% and we have used the fact that the matrix elements between states $\ket{n\mathbf{k}}$
-% are diagonal in $\mathbf{k}$, i.e. proportional to $\delta(\mathbf{k}-\mathbf{k}')$.
-Integrating the microscopic current $\mathbf{j}(\mathbf{r},t)$ over
-the entire slab gives the averaged microscopic current density, $\mathbf{J}(t)$. 
+% and we have used the fact that the matrix elements between states $\ket{n\bfk}$
+% are diagonal in $\bfk$, i.e. proportional to $\gd(\bfk-\bfk')$.
+Integrating the microscopic current $\mbf{j}(\mbf{r},t)$ over
+the entire slab gives the averaged microscopic current density, $\bfJ(t)$. 
 If we want the contribution from only one region of the unit cell 
@@
 \begin{equation}\label{jsz}
-\frac{1}{A}\int d^3r\, {\mathbf{\cal C}}(z)\, 
+\frac{1}{A}\int d^3r\, \calc(z)\, 
 \mathbf{J} (\mathbf{r},t)
- \equiv \boldsymbol{\mathcal{\cal J}}^{(N)}(t),
-\end{equation}
-where $\boldsymbol{\mathcal{\cal J}}^{(N)}(t)$ is the $N$-th order current induced in the
-region specified by ${\mathbf{\cal C}}(z)$.
+ \equiv \calbj^{(N)}(t),
+\end{equation}
+where $\calbj^{(N)}(t)$ is the $N$-th order current induced in the
+region specified by $\calc(z)$.
 Therefore we define
 \begin{equation}\label{vcal}
-e{\boldsymbol{\mathcal{V}}}^{\Sigma}_{mn}(\mathbf{k})
+e{\boldsymbol{\mathcal{V}}}^{\gS}_{mn}(\mathbf{k})
 \equiv
-\int d^3r\, {\mathbf{\cal C}}(z)\,\mathbf{j}_{mn}({\mathbf{k}};\mathbf{r}),
+\int d^3r\, \calc(z)\,\bfj_{mn}({\bfk};\bfr),
 \end{equation}
@@
 \begin{equation}\label{jmac2}
-\boldsymbol{\mathcal{\cal J}}^{(N)}(2\omega)=\frac{e}{A}
+\calbj^{(N)}(2\go)=\frac{e}{A}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{mn}(\mathbf{k}) 
-\rho^{(N)}_{nm}(\mathbf{k};2\omega) 
+\calbv^{\gS}_{mn}(\mathbf{k}) 
+\rho^{(N)}_{nm}(\bfk;2\go) 
 , 
@@
 \begin{align}\label{intj}
-{\boldsymbol{\mathcal{V}}}^{\Sigma}_{mn}({\mathbf k})
+{\boldsymbol{\mathcal{V}}}^{\gS}_{mn}({\mathbf k})
 &=
@@
 \int \mathrm{d}^3 r\,
- {\mathbf{\cal C}}(z)
+ \calc(z)
 \bigg[
-\langle m\mathbf{k}|\mathbf{v}^\Sigma | \mathbf{r}\rangle
+\langle m\mathbf{k}|\mathbf{v}^\gS | \mathbf{r}\rangle
 \langle \mathbf{r} | n \mathbf k \rangle +
 \langle m\mathbf{k} | \mathbf{r}\rangle
-\langle \mathbf{r} | \mathbf{v}^\Sigma | n \mathbf k \rangle\bigg]
+\langle \mathbf{r} | \mathbf{v}^\gS | n \mathbf k \rangle\bigg]
 \nonumber\\
@@
 \int \mathrm{d}^3 r\,
- {\mathbf{\cal C}}(z)
+ \calc(z)
  \bigg[
 \psi_{n\mathbf{k}}(\mathbf{r})
-\mathbf{v}^{\Sigma *}\psi^*_{m\mathbf{k}}(\mathbf{r})
-+ 
-\psi^*_{m\mathbf{k}}(\mathbf{r})\mathbf{v}^\Sigma
+\bfv^{\gS *}\psi^*_{m\mathbf{k}}(\mathbf{r})
++ 
+\psi^*_{m\mathbf{k}}(\mathbf{r})\bfv^\gS
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\left[\frac{{\mathbf{\cal C}}(z) \mathbf{v}^\Sigma +
-\mathbf{v}^\Sigma {\mathbf{\cal C}}(z)}{2}\right]
+\left[\frac{\calc(z) \bfv^\gS +
+\bfv^\gS \calc(z)}{2}\right]
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 \psi^*_{m\mathbf{k}}(\mathbf{r})
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}
+\calbv^{\gS}
 \psi_{n\mathbf{k}}(\mathbf{r})
@@
 %Here an integration by parts is performed on the third term of the
-%right-hand; since the $e^{-i\mathbf{k}\cdot\mathbf{r}}\psi_{n\mathbf{k}}(\mathbf{r})$
+%right-hand; since the $e^{-i\bfk\cdot\bfr}\psi_{n\bfk}(\bfr)$
 %are periodic over the unit cell, the surface term vanishes. 
-where, we used the hermitian property of $\mathbf{v}^\Sigma$ and defined
+where, we used the hermitian property of $\bfv^\gS$ and defined
 \begin{align}\label{nl.4}
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}
+\calbv^{\gS}
 =
-\frac{{\mathbf{\cal C}}(z) \mathbf{v}^\Sigma +
-\mathbf{v}^\Sigma {\mathbf{\cal C}}(z)}{2}
+\frac{\calc(z) \bfv^\gS +
+\bfv^\gS \calc(z)}{2}
 .
 \end{align} 
-%where the superscript $\ell$ is inherited from ${\mathbf{\cal C}}(z)$.
+%where the superscript $\ell$ is inherited from $\calc(z)$.
 % and we
@@
 \begin{align}\label{vcali}
-\mathbf{V} \to \boldsymbol{\mathcal{\cal V}}=\frac{{\mathbf{\cal C}}(z) \mathbf{V} +
-\mathbf{V} {\mathbf{\cal C}}(z)}{2}
+\bfV \to \calbv=\frac{\calc(z) \bfV +
+\bfV \calc(z)}{2}
 ,
@@
 is all that is needed to change any of the
-electron velocity operators $\mathbf{V}$ to the new velocity
-operator $\boldsymbol{\mathcal{\cal V}}$ that implicitly takes into account the
-contribution of the region of the slab given by ${\mathbf{\cal C}}(z)$.
+electron velocity operators $\bfV$ to the new velocity
+operator $\calbv$ that implicitly takes into account the
+contribution of the region of the slab given by $\calc(z)$.
 We note that this modified operator is hermitian as it must.\cite{note2}
-The operator $\mathbf{V}$ could be any of those given by Eq.~\eqref{vop2},
+The operator $\bfV$ could be any of those given by Eq.~\eqref{vop2},
 thus
 \begin{align}\label{vopii}
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}
+\calbv^{\gS}
 &=
-\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}
-+
-\boldsymbol{\mathcal{\cal V}}^{S}
+\calbv^{\lda}
++
+\calbv^{S}
 \nonumber\\
-\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}
+\calbv^{\lda}
 &=
-\boldsymbol{\mathcal{\cal V}}
-+
-\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}
+\calbv
++
+\calbv^{\nl}
 .
@@
 To calculate
-$\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{nm}(\mathbf{k})$ 
+$\calbv^{\gS}_{nm}(\bfk)$ 
 we calculate the matrix elements of 
-$\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}$ and $\boldsymbol{\mathcal{\cal V}}^{S}$
+$\calbv^{\lda}$ and $\calbv^{S}$
  (separately)
@@
 The layer-by-layer analysis of Refs. \onlinecite{hoganPRB03,castilloPRB03,mottaCMS10}  
-used ${\mathbf{\cal C}}(z)$
+used $\calc(z)$
 %limiting the current response  
@@
 To limit the SHG response to one surface, Eq.~\eqref{vcali} 
-for $\boldsymbol{\mathcal{\cal V}}$ was proposed in 
+for $\calbv$ was proposed in 
 Ref.~\onlinecite{reiningPRB94} and later used in Refs.
@@
  and \onlinecite{mejiaRMF04}. 
-In this article, for the first time, we formally introduce the cut function ${\mathbf{\cal C}}(z)$ 
+In this article, for the first time, we formally introduce the cut function $\calc(z)$ 
 for the second-harmonic optical response of semiconductor surfaces,
@@
 Using
-$\boldsymbol{\mathcal{\cal J}}=d\boldsymbol{\mathcal{\cal P}}/dt$ 
+$\calbj=d\calbp/dt$ 
 and Eq.~\eqref{jmac2} 
@@
 \begin{equation}\label{Pjikn}
-\boldsymbol{\mathcal{\cal P}}^{(2)}(2\omega)
-=\frac{ie}{2A\tilde\omega}
+\calbp^{(2)}(2\go)
+=\frac{ie}{2A\got}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{mn}(\mathbf{k})
-\rho^{(2)}_{nm}(\mathbf{k};2\omega)
+\calbv^{\gS}_{mn}(\mathbf{k})
+\rho^{(2)}_{nm}(\bfk;2\go)
 ,
@@
 \begin{align}\label{Pjikn2}
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega) 
+\chi^{\rma\rmb\rmc}(-2\go;\go,\go) 
 &=
-\frac{e^2}{2A\hbar\tilde\omega}
+\frac{e^2}{2A\hbar\got}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\frac{{\mathbf{\cal V}}^{\Sigma,\mathrm{a}}_{mn}(\mathbf{k})}
-{\omega^\Sigma_{nm\mathbf{k}}-2\tilde\omega}
+\frac{\calv^{\gS,\rma}_{mn}(\mathbf{k})}
+{\go^\gS_{nm\bfk}-2\got}
 \bigg[
--(B_{nm}^{\mathrm{c}}(\mathbf{k},\omega))_{;k^{\mathrm{b}}}
+-(B_{nm}^{\rmc}(\bfk,\go))_{;k^{\rmb}}
 \nonumber \\
 &
-+i\sum_q\left(r_{nq}^{\mathrm{b}}B_{qm}^{\mathrm{c}}(\mathbf{k},\omega) -
-  B_{nq}^{\mathrm{c}}(\mathbf{k},\omega) 
-  r_{qm}^{\mathrm{b}}\right)
++i\sum_q\left(r_{nq}^{\rmb}B_{qm}^{\rmc}(\bfk,\go) -
+  B_{nq}^{\rmc}(\bfk,\go) 
+  r_{qm}^{\rmb}\right)
 \bigg]
@@
 which gives the susceptibility 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ 
-of the layers of the slab specified by ${\mathbf{\cal C}}(z)$. 
+$\chi^{\rma\rmb\rmc}(-2\go;\go,\go)$ 
+of the layers of the slab specified by $\calc(z)$. 
 We mention that the units of 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$
+$\chi^{\rma\rmb\rmc}(-2\go;\go,\go)$
 are m$^2$/V, as they should be for a surface SH susceptibility.
@@
 \begin{equation}\label{chii}
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_i (-2\omega;\omega,\omega)
-=-\frac{e^3}{A\hbar^22\tilde\omega}
+\chi^{\rma\rmb\rmc}_i (-2\go;\go,\go)
+=-\frac{e^3}{A\hbar^22\got}
 \int \frac{dk^3}{8\pi^3}
 \sum_{mn}
-\frac{{\mathbf{\cal V}}_{mn}^{\Sigma,\mathrm{a}}}{\omega^\Sigma_{nm}-2\tilde\omega}
-\left(\frac{f_{mn}r_{nm}^{\mathrm{b}}}{\omega^\Sigma_{nm}-\tilde\omega}\right)_{;k^{\mathrm{c}}}
+\frac{\calv_{mn}^{\gS,\rma}}{\go^\gS_{nm}-2\got}
+\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gS_{nm}-\got}\right)_{;k^{\rmc}}
 ,
@@
 \begin{equation}\label{chie}
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_e (-2\omega;\omega,\omega)
-=\frac{ie^3}{A\hbar^22\tilde\omega}
+\chi^{\rma\rmb\rmc}_e (-2\go;\go,\go)
+=\frac{ie^3}{A\hbar^22\got}
 \int \frac{dk^3}{8\pi^3}
 \sum_{qmn}
-\frac{{\mathbf{\cal V}}_{mn}^{\Sigma,\mathrm{a}}}{\omega^\Sigma_{nm}-2\tilde\omega}
+\frac{\calv_{mn}^{\gS,\rma}}{\go^\gS_{nm}-2\got}
 \left(
-\frac{r_{nq}^{\mathrm{c}} r_{qm}^{\mathrm{b}} 
-f_{mq}}{\omega^\Sigma_{qm}-\tilde\omega}
--\frac{r_{nq}^{\mathrm{b}} r_{qm}^{\mathrm{c}} 
-f_{qn}}{\omega^\Sigma_{nq}-\tilde\omega}
+\frac{r_{nq}^{\rmc} r_{qm}^{\rmb} 
+f_{mq}}{\go^\gS_{qm}-\got}
+-\frac{r_{nq}^{\rmb} r_{qm}^{\rmc} 
+f_{qn}}{\go^\gS_{nq}-\got}
 \right),
@@
 \begin{equation}\label{gene2}
-\left(\frac{f_{mn}r_{nm}^{\mathrm{b}}}{\omega^\Sigma_{nm}-\tilde\omega}\right)_{;k^{\mathrm{c}}}=
-\frac{f_{mn}}{\omega^\Sigma_{nm}-\tilde\omega}\left(r_{nm}^\mathrm{b}\right)_{;k^{\mathrm{c}}}
--\frac{f_{mn}r_{nm}^{\mathrm{b}}\Delta_{nm}^\mathrm{c}}{(\omega^\Sigma_{nm}-\tilde\omega)^2}
-,
-\end{equation}
-where substituting $H^\Sigma_0$ 
+\left(\frac{f_{mn}r_{nm}^{\rmb}}{\go^\gS_{nm}-\got}\right)_{;k^{\rmc}}=
+\frac{f_{mn}}{\go^\gS_{nm}-\got}\left(r_{nm}^\rmb\right)_{;k^{\rmc}}
+-\frac{f_{mn}r_{nm}^{\rmb}\gD_{nm}^\rmc}{(\go^\gS_{nm}-\got)^2}
+,
+\end{equation}
+where substituting $H^\gS_0$ 
 into Eq.~\eqref{conmri3n} and then
@@
 \begin{align}\label{eli.13}
-\left(\omega^\Sigma_{nm}\right)_{;k^{\mathrm{a}}}
+\left(\go^\gS_{nm}\right)_{;k^{\rma}}
 =
-\left(\omega^\mathrm{LDA}_{nm}\right)_{;k^{\mathrm{a}}}
+\left(\go^\lda_{nm}\right)_{;k^{\rma}}
 = 
-v_{nn}^{\mathrm{LDA},\mathrm{a}}-v_{mm}^{\mathrm{LDA},\mathrm{a}}\equiv\Delta_{nm}^{\mathrm{a}}
+v_{nn}^{\lda,\rma}-v_{mm}^{\lda,\rma}\equiv\gD_{nm}^{\rma}
 .
 \end{align} 
-The apparent divergence as $\tilde\omega\to 0$
+The apparent divergence as $\got\to 0$
 in Eqs. \eqref{chii} and \eqref{chie},  
 is removed  by
- a partial fraction expansion over $\tilde\omega$. 
+ a partial fraction expansion over $\got$. 
 Using time-reversal invariance, an integration by parts to 
@@
 \begin{equation}\label{calvimchiewn}
-\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}]= 
+\mathrm{Im}[\chi^{\rma\rmb\rmc}_{e,\go}]= 
 \frac{\pi |e|^3}{2\hbar^2}
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\Sigma_{cv}}
+\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\gS_{cv}}
 \left[
-\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{qc}\{r^{\mathrm{b}}_{cv}r^{\mathrm{c}}_{vq}\}]}
-{(2\omega^\Sigma_{cv}-\omega^\Sigma_{cq})} 
--\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vq}\{r^{\mathrm{c}}_{qc}r^{\mathrm{b}}_{cv}\}]}
-{(2\omega^\Sigma_{cv}-\omega^\Sigma_{qv})}
-\right]\delta(\omega^\Sigma_{cv}-\omega),
+\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{qc}\{r^{\rmb}_{cv}r^{\rmc}_{vq}\}]}
+{(2\go^\gS_{cv}-\go^\gS_{cq})} 
+-\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vq}\{r^{\rmc}_{qc}r^{\rmb}_{cv}\}]}
+{(2\go^\gS_{cv}-\go^\gS_{qv})}
+\right]\gd(\go^\gS_{cv}-\go),
 \end{equation}  
 \begin{equation}\label{calvimchiwn}
-\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}]= 
+\mathrm{Im}[\chi^{\rma\rmb\rmc}_{i,\go}]= 
 \frac{\pi\vert e\vert^3}{2\hbar^2}
 \int \frac{dk^3}{8\pi^3}
-\sum_{cv}\frac{1}{(\omega^\Sigma_{cv})^{2}}
+\sum_{cv}\frac{1}{(\omega^\gS_{cv})^{2}}
 \left[
-\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
-+\frac{\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
-\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\Sigma_{cv}} 
-\right]\delta(\omega^\Sigma_{cv}-\omega),
+\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\gS,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
++\frac{\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gS_{cv}} 
+\right]\delta(\omega^\gS_{cv}-\omega),
 \end{equation}
 \begin{equation}\label{calvimchie2wn}
-\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}]= 
+\mathrm{Im}[\chi^{\rma\rmb\rmc}_{e,2\go}]= 
 -\frac{\pi |e|^3}{2\hbar^2}
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\frac{4}{\omega^\Sigma_{cv}}
+\sum_{vc}\frac{4}{\omega^\gS_{cv}}
 \left[
 \sum_{v'\ne
-  v}\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\{r^{\mathrm{b}}_{cv'}r^{\mathrm{c}}_{v'v}\}]}
-{2\omega^\Sigma_{cv'}-\omega^\Sigma_{cv}}
+  v}\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\{r^{\rmb}_{cv'}r^{\rmc}_{v'v}\}]}
+{2\go^\gS_{cv'}-\go^\gS_{cv}}
 - \sum_{c'\ne
-  c}\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\{r^{\mathrm{c}}_{cc'}r^{\mathrm{b}}_{c'v}\}]}
-{2\omega^\Sigma_{c'v}-\omega^\Sigma_{cv}}
-\right]\delta(\omega^\Sigma_{cv}-2\omega),
+  c}\frac{\mathrm{Im}[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\{r^{\rmc}_{cc'}r^{\rmb}_{c'v}\}]}
+{2\go^\gS_{c'v}-\go^\gS_{cv}}
+\right]\gd(\go^\gS_{cv}-2\go),
 \end{equation}
 \begin{equation}\label{calvimchi2wn}
-\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}]= 
+\mathrm{Im}[\chi^{\rma\rmb\rmc}_{i,2\go}]= 
  \frac{\pi \vert
@@
 \int \frac{dk^3}{8\pi^3}
-\sum_{vc}\frac{4}{(\omega^\Sigma_{cv})^{2}}
-\left[\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
+\sum_{vc}\frac{4}{(\omega^\gS_{cv})^{2}}
+\left[\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
 \right\}\right] -
-\frac{2\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
-\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\Sigma_{cv}}\right]\delta(\omega^\Sigma_{cv}-2\omega)
+\frac{2\mathrm{Re}\left[\mathcal{V}^{\gS,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
+\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\gS_{cv}}\right]\delta(\omega^\gS_{cv}-2\omega)
 ,
@@
 \end{subequations}
-where we have split the interband and intraband $1\omega$ and $2\omega$
-contributions and supressed the $\omega$ arguments for 
+where we have split the interband and intraband $1\go$ and $2\go$
+contributions and supressed the $\go$ arguments for 
 convenience of notation.
@@
 a Kramers-Kronig transformation\cite{nicolasPRB14} and
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}=
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega} 
-+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}
-+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega} 
-+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}
+$\chi^{\rma\rmb\rmc}=
+\chi^{\rma\rmb\rmc}_{e,\go} 
++\chi^{\rma\rmb\rmc}_{e,2\go}
++\chi^{\rma\rmb\rmc}_{i,\go} 
++\chi^{\rma\rmb\rmc}_{i,2\go}
 $.
 To fulfill the required intrinsic permutation symmetry, %\cite{rashkeevPRB98} 
-the $\{\}$ notation symmetrizes the $\mathrm{b}\mathrm{c}$ Cartesian indices, i.e. 
-$\{u^{\mathrm{b}}s^{\mathrm{c}}\}=(u^{\mathrm{b}}s^{\mathrm{c}}+u^{\mathrm{c}}s^{\mathrm{b}})/2$,
+the $\{\}$ notation symmetrizes the $\rmb\rmc$ Cartesian indices, i.e. 
+$\{u^{\rmb}s^{\rmc}\}=(u^{\rmb}s^{\rmc}+u^{\rmc}s^{\rmb})/2$,
 and thus
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}=\chi^{\mathrm{a}\mathrm{c}\mathrm{b}}$.
+$\chi^{\rma\rmb\rmc}=\chi^{\rma\rmc\rmb}$.
 The various quantities involved in Eqs.~\eqref{chis} are given in
 the Appendix \ref{appe}. 
-We mention that if we take ${\mathbf{\cal C}}(z)=1$ through out, the layered
-matrix elements $\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{nm}$ become standard bulk-like
-$\mathbf{v}^{\Sigma}_{nm}$ matrix elements. We mention that in this
+We mention that if we take $\calc(z)=1$ through out, the layered
+matrix elements $\calbv^{\gS}_{nm}$ become standard bulk-like
+$\bfv^{\gS}_{nm}$ matrix elements. We mention that in this
 case, Eq.~\eqref{chis} is equivalent to the expressions of
@@
 \begin{equation}\label{chiijksur}
-\boldsymbol{\chi}(-2\omega;\omega,\omega)
-= \sum_{\{\ell\}}\boldsymbol{\chi}^\ell(-2\omega;\omega,\omega),
+\bfgchi(-2\go;\go,\go)
+= \sum_{\{\ell\}}\bfgchi^\ell(-2\go;\go,\go),
 \end{equation} 
 where $\ell$ would denote a particular layer chosen through
-${\mathbf{\cal C}}^\ell(z)$ of Eq.~\eqref{sz} and
+$\calc^\ell(z)$ of Eq.~\eqref{sz} and
 $\{\ell\}$
@@
 %For a centrosymmetric slab, 
-%$\boldsymbol{\chi}^{\ell=N}=0$;
+%$\bfgchi^{\ell=N}=0$;
 We remark that the value of 
@@
 the slab needs to have enough atomic layers 
-%$\boldsymbol{\chi}^{\ell=N}=0$;
+%$\bfgchi^{\ell=N}=0$;
 %to be satisfied and to 
 in order to give converged results for 
-$\boldsymbol{\chi} (-2\omega;\omega,\omega)$. 
+$\bfgchi (-2\go;\go,\go)$. 
 We can use Eq.~\eqref{chiijksur} for 
 either the front or the back surface. 
-% This decomposition of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$ into
-% $\boldsymbol{\chi}^\ell(-2\omega;\omega,\omega)$ would allow to explore the contribution
+% This decomposition of $\bfgchi(-2\go;\go,\go)$ into
+% $\bfgchi^\ell(-2\go;\go,\go)$ would allow to explore the contribution
 % to SH from any particular layer of the system,
@@
 \begin{align}\label{hs}
-\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega) 
+\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go) 
 =
-\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega) 
+\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go) 
 ,
@@
 where
-$\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega)$ is calculated using
-${\mathbf{\cal C}}(z)=1$ from the upper half containing the $2\times 1$ 
+$\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
+$\calc(z)=1$ from the upper half containing the $2\times 1$ 
 surface reconstruction, as seen in Fig.~\ref{si2x1},
-and $\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega)$ is calculated using
-${\mathbf{\cal C}}(z)=1$ through the full slab.
+and $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
+$\calc(z)=1$ through the full slab.
 We show the results for this comparison in the remainder 
@@
 form.\cite{kleinmanPRL82}
-The contribution of $\mathbf{v}^\mathrm{nl}$ to Eq.~\eqref{chis} can be carried out 
+The contribution of $\bfv^\nl$ to Eq.~\eqref{chis} can be carried out 
 as detailed in Appendix \ref{vesnl} using the DP code for 
@@
 these are: an energy cut-off of 10 Ha, an equal number of conduction and 
-valence bands, and a set of 244 $\mathbf{k}$-points.
-The $\mathbf{k}$-points are used for the linear analytic 
+valence bands, and a set of 244 $\bfk$-points.
+The $\bfk$-points are used for the linear analytic 
 tetrahedron method for evaluating the
@@
 implies that there is a region in the Brillouin zone where 
-$\omega_{cv}(\mathbf{k})=2 \omega_{c'v}(\mathbf{k})$. Since there is a real population 
+$\omega_{cv}(\bfk)=2 \omega_{c'v}(\bfk)$. Since there is a real population 
 excitation, the perturbation theory used to calculate the spectrum breaks 
-down at these $\mathbf{k}$-points. This must be taken into account in a more 
+down at these $\bfk$-points. This must be taken into account in a more 
 complete calculation. These points introduce sharp spikes in the spectrum. 
@@
 We must evaluate 
-$T^{\mathrm{a}\mathrm{b}}_{nm}=(i/\hbar)[r^\mathrm{b},v^{\mathrm{nl},\mathrm{a}}]_{nm}$,
+$T^{\rma\rmb}_{nm}=(i/\hbar)[r^\rmb,v^{\nl,\rma}]_{nm}$,
 as outlined in Appendix \ref{vesnl},
@@
 time consuming. This adds significantly to the already lengthy time needed 
-for the calculation of the $\mathbf{v}^\mathrm{nl}$ contribution that is
+for the calculation of the $\bfv^\nl$ contribution that is
 proportional only to 
 the first order derivatives. 
-Memory requirements are also increased for both $\mathbf{v}^\mathrm{nl}$ and 
-$[\mathbf{r},\mathbf{v}^\mathrm{nl}]$. However, the contribution from $[\mathbf{r},\mathbf{v}^\mathrm{nl}]$ 
+Memory requirements are also increased for both $\bfv^\nl$ and 
+$[\bfr,\bfv^\nl]$. However, the contribution from $[\bfr,\bfv^\nl]$ 
 is very small\cite{valerie} and therefore we neglect it in this work.
@@
 As explained above,
-the calculation of the $\mathbf{v}^\mathrm{nl}$ contribution is 
+the calculation of the $\bfv^\nl$ contribution is 
 computationally expensive.
@@
 between including or not including the
-effects of $\mathbf{v}^\mathrm{nl}$ or the scissors correction
-$\hbar\Delta$.   
+effects of $\bfv^\nl$ or the scissors correction
+$\hbar\gD$.   
 For these results we chose
-$\hbar\Delta=0.5$ eV, that is the GW gap reported in
+$\hbar\gD=0.5$ eV, that is the GW gap reported in
 Refs.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
@@
 
-\subsection{\texorpdfstring{Results for $\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)$}
+\subsection{\texorpdfstring{Results for $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$}
 {Results for Xxxx(2x1)(-2w;w,w)}}
@@
 band gap of this surface, and $E_g=1.03$ eV is the scissor shifted gap for
-$\hbar\Delta=0.5$ eV.
+$\hbar\gD=0.5$ eV.
 These resonances
@@
 In Fig.~\ref{fig3}
-We see that the effect of $\mathbf{v}^\mathrm{nl}$
+We see that the effect of $\bfv^\nl$
  reduces the value of   
@@
 of SSHG.  
-However, the inclusion of $\mathbf{v}^\mathrm{nl}$ does not
+However, the inclusion of $\bfv^\nl$ does not
 changes the 
@@
 atomic Si layers plus one H layer, 
-with and without the contribution from $\mathbf{v}^\mathrm{nl}$.
+with and without the contribution from $\bfv^\nl$.
 \label{fig3}} 
@@
 To see the effect of the scissors correction, we take two different
-finite values for $\hbar\Delta$. The first one
-with a value of $\hbar\Delta=0.5$ eV, used in above results, 
+finite values for $\hbar\gD$. The first one
+with a value of $\hbar\gD=0.5$ eV, used in above results, 
 is the ``average'' GW gap taken from 
@@
 that is in agreement with Ref.~\onlinecite{garciaCPC01}. The second one
-with a value of $\hbar\Delta=0.63$ eV is the ``average'' 
+with a value of $\hbar\gD=0.63$ eV is the ``average'' 
 gap taken from Ref.~\onlinecite{asahiPRB00}, 
-where more $\mathbf{k}$-points in the Brillouin zone were 
+where more $\bfk$-points in the Brillouin zone were 
 used to calculate its GW value.
@@
 When we compare 
-$|\chi^{xxx}_{2\times 1}|$ for the two finite values of $\hbar\Delta$,
+$|\chi^{xxx}_{2\times 1}|$ for the two finite values of $\hbar\gD$,
 we see that the first two peaks are almost rigidly 
@@
 for two different values of 
-the scissors correction $\hbar\Delta$. 
+the scissors correction $\hbar\gD$. 
 \label{fig4}} 
@@
 So, we would expect that the $xxx$ component of 
-$\boldsymbol{\chi}(-2\omega;\omega,\omega)$ would have a small influence from the local fields.
+$\bfgchi(-2\go;\go,\go)$ would have a small influence from the local fields.
 Unfortunately the experimental measurement of the $xxx$
-component of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$ is not possible as the SH 
+component of $\bfgchi(-2\go;\go,\go)$ is not possible as the SH 
 radiated intensity would be proportional not only to this component 
-but also to the other components of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$.  
+but also to the other components of $\bfgchi(-2\go;\go,\go)$.  
 However,
@@
 We have presented a formulation to calculate the surface second-harmonic
-(SSH) susceptibility tensor $\boldsymbol{\chi}(-2\omega;\omega,\omega)$, using the length gauge
+(SSH) susceptibility tensor $\bfgchi(-2\go;\go,\go)$, using the length gauge
 formalism and within the independent particle approximation (IPA).
@@
 that 
-$\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega) 
+$\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go) 
 \approx
-\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega) 
+\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go) 
 . 
@@
 Although one can in principle increase the number of atomic layers,
-$\mathbf{k}$-points, etc. to
+$\bfk$-points, etc. to
 improve even further on the similarity of the half-slab and full-slab results, we
@@
 effects mentioned above in the calculation of  
-$\boldsymbol{\chi}(-2\omega;\omega,\omega)$. 
+$\bfgchi(-2\go;\go,\go)$. 
 The scissors correction shifts the spectrum to  
@@
  The cut function allows us  
-to extract unequivocally $\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)$.
+to extract unequivocally $\chi^{xxx}_{2\times 1}(-2\go;\go,\go)$.
  The effects of the nonlocal part of the pseudopotentials  
 keeps the same line-shape of $|\chi^{xxx}_{2\times
-  1}(-2\omega;\omega,\omega)|$,
+  1}(-2\go;\go,\go)|$,
 but  
@@
 The $xxx$ component of  
-$\boldsymbol{\chi}_{2\times 1}(-2\omega;\omega,\omega)$, 
+$\bfgchi_{2\times 1}(-2\go;\go,\go)$, 
 can not be
@@
 \subsection{ Expressions for 
-\texorpdfstring{${\boldsymbol{\mathcal{\cal V}}}_{nm}(\mathbf{k})$}{que} 
+\texorpdfstring{${\calbv}_{nm}(\bfk)$}{que} 
 and 
-\texorpdfstring{${\cal C}_{nm}(\mathbf{k})$}{que}
+\texorpdfstring{${\cal C}_{nm}(\bfk)$}{que}
 }\label{calpcalc}
@@
 \begin{align}\label{eni.1}
-\psi_{n\mathbf{k}}(\mathbf{r})=\sum_\mathbf{G} A_{n\mathbf{k}}(\mathbf{G})e^{i(\mathbf{k}+\mathbf{G})\cdot\mathbf{r}}
-,
-\end{align}
-where $\{\mathbf{G}\}$ are the reciprocal basis vectors satisfying
-$e^{\mathbf{R}\cdot\mathbf{G}}=1$, with $\{\mathbf{R}\}$ the translation vectors in real
-space, and $A_{n\mathbf{k}}(\mathbf{G})$ the expansion coefficients. Using
-$m_e\mathbf{v}=\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ into 
+\psi_{n\bfk}(\bfr)=\sum_\bfG A_{n\bfk}(\bfG)e^{i(\bfk+\bfG)\cdot\bfr}
+,
+\end{align}
+where $\{\bfG\}$ are the reciprocal basis vectors satisfying
+$e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
+space, and $A_{n\bfk}(\bfG)$ the expansion coefficients. Using
+$m_e\bfv=\bfp=-i\hbar\bfgnabla$ into 
 Eq.~\eqref{vcali}
@@
 \begin{align}\label{eni.2}
-\boldsymbol{\mathcal{\cal V}}_{nm}(\mathbf{k})=
+\calbv_{nm}(\bfk)=
 \frac{\hbar}{2m_e}
-\sum_{\mathbf{G},\mathbf{G}'} A^*_{n\mathbf{k}}(\mathbf{G}')  A_{m\mathbf{k}}(\mathbf{G})
-(2\mathbf{k}+\mathbf{G}+\mathbf{G}')
-\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel}  
+\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
+(2\bfk+\bfG+\bfG')
+\gd_{\bfG_\parallel \bfG'_\parallel}  
 f(G_\perp-G'_\perp)
@@
 \begin{align}\label{vnl.9}
-f(G_\perp)=\frac{1}{L}\int{\mathbf{\cal C}}(z) e^{iG_\perp z}dz  
+f(G_\perp)=\frac{1}{L}\int\calc(z) e^{iG_\perp z}dz  
  ,
@@
 is the length of the super-cell. 
-The reciprocal lattice vectors $\mathbf{G}$ are 
+The reciprocal lattice vectors $\bfG$ are 
 decomposed into components
-parallel ($\mathbf{G}_\parallel$), and perpendicular ($G_\perp \hat z$)
+parallel ($\bfG_\parallel$), and perpendicular ($G_\perp \hat z$)
 to the surface, so
-that $\mathbf{G} = \mathbf{G}_\parallel + G_\perp\hat z$.
-The double summation over the $\mathbf{G}$ vectors can be 
+that $\bfG = \bfG_\parallel + G_\perp\hat z$.
+The double summation over the $\bfG$ vectors can be 
 calculated efficiently by  
@@
 \begin{align}\label{eni.4}
-{\mathbf{\cal C}}_{nm}(\mathbf{k})=
-\sum_{\mathbf{G},\mathbf{G}'} A^*_{n\mathbf{k}}(\mathbf{G}')  A_{m\mathbf{k}}(\mathbf{G})
-\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel} 
+\calc_{nm}(\bfk)=
+\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
+\gd_{\bfG_\parallel \bfG'_\parallel} 
 f(G_\perp-G'_\perp)
@@
 \end{align}  
-If ${\mathbf{\cal C}}(z)=1$, then $f(G_\perp)=\delta_{G_\perp 0}$ and we 
+If $\calc(z)=1$, then $f(G_\perp)=\gd_{G_\perp 0}$ and we 
 obtain the full-slab/bulk values, 
-$\mathbf{v}_{nm}(\mathbf{k})$ and ${\mathbf{\cal C}}_{nm}(\mathbf{k})=\delta_{nm}$,
+$\bfv_{nm}(\bfk)$ and $\calc_{nm}(\bfk)=\gd_{nm}$,
 from Eqs.~\eqref{eni.2} and \eqref{eni.4}.
@@
 \subsection{Expressions for 
-\texorpdfstring{$({\mathbf{\cal V}}^{\mathrm{LDA},a}_{nm})_{;k^b}$}{Vnonlocal},
+\texorpdfstring{$(\calv^{\lda,a}_{nm})_{;k^b}$}{Vnonlocal},
 \texorpdfstring{$(r^a_{nm})_{;k^b}$}{Vnonlocal}
-and \texorpdfstring{$({\mathbf{\cal C}}_{nm})_{;\mathbf{k}}$}{Vnonlocal}
+and \texorpdfstring{$(\calc_{nm})_{;\bfk}$}{Vnonlocal}
 }\label{appvnl}
@@
 $
-{\cal T}^{\mathrm{a}\mathrm{b}}\equiv[r^{\mathrm{b}},{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}]
+\calt^{\rma\rmb}\equiv[r^{\rmb},\calv^{\lda,\rma}]
 \equiv
-[r^{\mathrm{b}},{\mathbf{\cal V}}^\mathrm{a}]
+[r^{\rmb},\calv^\rma]
 $
@@
 \begin{align}\label{nmesn}
-({\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}&=
-{\cal T}_{nm}^{\mathrm{a}\mathrm{b}}
+(\calv^{\lda,\rma}_{nm})_{;k^{\rmb}}&=
+\calt_{nm}^{\rma\rmb}
 +i
@@
 \bigg(
-r^{\mathrm{b}}_{nq}  
-{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{q m}
+r^{\rmb}_{nq}  
+\calv^{\lda,\rma}_{q m}
 -
-{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nq}   
-r^{\mathrm{b}}_{q m}
+\calv^{\lda,\rma}_{nq}   
+r^{\rmb}_{q m}
 \bigg)  
 +i  
-r^{\mathrm{b}}_{nm}
-\Delta^{\mathrm{a}}_{mn}
+r^{\rmb}_{nm}
+\Delta^{\rma}_{mn}
 ,
@@
 \begin{eqnarray}\label{tdel}
-\Delta^{\mathrm{a}}_{mn}
+\Delta^{\rma}_{mn}
 =
-{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nn}  
+\calv^{\lda,\rma}_{nn}  
 -
-{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{mm}  
+\calv^{\lda,\rma}_{mm}  
 ,
@@
 \begin{align}\label{tau.1}
-{\cal T}_{nm}^{\mathrm{a}\mathrm{b}}
+\calt_{nm}^{\rma\rmb}
 =
-\frac{\hbar}{m_e}\delta_{\mathrm{a}\mathrm{b}} 
+\frac{\hbar}{m_e}\gd_{\rma\rmb} 
 C_{nm} 
@@
 \sum_q  
-L^{\mathrm{a}\mathrm{b}}_{nq} 
+L^{\rma\rmb}_{nq} 
 C_{qm} 
@@
 \begin{align}\label{lab}
-L_{nm}^{\mathrm{a}\mathrm{b}}
-=\frac{i}{\hbar}[r^{\mathrm{b}},v^{\mathrm{nl},\mathrm{a}}]_{nm}
+L_{nm}^{\rma\rmb}
+=\frac{i}{\hbar}[r^{\rmb},v^{\nl,\rma}]_{nm}
 .
 \end{align}  
-The matrix elements $L^{\mathrm{a}\mathrm{b}}_{nm}$
+The matrix elements $L^{\rma\rmb}_{nm}$
 are calculated in Appendix \ref{vesnl}.
@@
 Notice that
-$(v^{\mathrm{LDA},\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}$ is obtained 
+$(v^{\lda,\rma}_{nm})_{;k^{\rmb}}$ is obtained 
 from Eq.~\eqref{nmesn} by 
 taking 
-${\mathbf{\cal C}}(z)=1$ or $C_{nm}=\delta_{nm}$.
-
-To obtain $(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}$ we use Eq.~\eqref{chon.10} to
+$\calc(z)=1$ or $C_{nm}=\gd_{nm}$.
+
+To obtain $(r^{\rma}_{nm})_{;k^{\rmb}}$ we use Eq.~\eqref{chon.10} to
 write
-$(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}
-=(v^{\mathrm{LDA},\mathrm{a}}_{nm}/i\omega^\mathrm{LDA}_{nm})_{;k^{\mathrm{b}}}
+$(r^{\rma}_{nm})_{;k^{\rmb}}
+=(v^{\lda,\rma}_{nm}/i\go^\lda_{nm})_{;k^{\rmb}}
 $ and simply apply the chain rule,
 \begin{align}\label{rka}
-(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}
+(r^{\rma}_{nm})_{;k^{\rmb}}
 &=
-T^{\mathrm{a}\mathrm{b}}_{nm}
+T^{\rma\rmb}_{nm}
 +
 \frac{ 
-r^{\mathrm{b}}_{nm}
-\Delta^{\mathrm{a}}_{mn}
-+r^{\mathrm{a}}_{nm}
-\Delta^{\mathrm{b}}_{mn}
+r^{\rmb}_{nm}
+\Delta^{\rma}_{mn}
++r^{\rma}_{nm}
+\Delta^{\rmb}_{mn}
 }
-{\omega^\mathrm{LDA}_{nm}}
-+
-\frac{i}{\omega^\mathrm{LDA}_{nm}}
+{\go^\lda_{nm}}
++
+\frac{i}{\go^\lda_{nm}}
 \sum_{q}
 \bigg(
-\omega^\mathrm{LDA}_{q m} 
-r^{\mathrm{b}}_{nq} 
-r^{\mathrm{a}}_{q m}
+\go^\lda_{q m} 
+r^{\rmb}_{nq} 
+r^{\rma}_{q m}
 -
-\omega^\mathrm{LDA}_{nq} 
-r^{\mathrm{a}}_{nq} 
-r^{\mathrm{b}}_{q m}
+\go^\lda_{nq} 
+r^{\rma}_{nq} 
+r^{\rmb}_{q m}
 \bigg)
@@
 \begin{eqnarray}\label{del}
-\Delta^{\mathrm{a}}_{mn}
+\Delta^{\rma}_{mn}
 =
-v^{\mathrm{LDA},\mathrm{a}}_{nn}  
+v^{\lda,\rma}_{nn}  
 -
-v^{\mathrm{LDA},\mathrm{a}}_{mm}  
+v^{\lda,\rma}_{mm}  
 ,
@@
 \begin{align}\label{tau.1n} 
-T_{nm}^{\mathrm{a}\mathrm{b}}
-=\frac{\hbar}{m_e}\delta_{ab}\delta_{nm} 
--\hbar L^{\mathrm{a}\mathrm{b}}_{nm} 
+T_{nm}^{\rma\rmb}
+=\frac{\hbar}{m_e}\gd_{ab}\gd_{nm} 
+-\hbar L^{\rma\rmb}_{nm} 
 .
@@
 Note that the layered term
-${\cal T}^{\mathrm{a}\mathrm{b}}_{nm}$ reduces to $T^{\mathrm{a}\mathrm{b}}_{nm}$
+$\calt^{\rma\rmb}_{nm}$ reduces to $T^{\rma\rmb}_{nm}$
 for the full-slab/bulk case.
@@
 Again, we use Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn},
-along with $[\mathbf{r},F(\mathbf{r})]=0$, valid for 
-and any function $F(\mathbf{r})$, 
+along with $[\bfr,F(\bfr)]=0$, valid for 
+and any function $F(\bfr)$, 
 to obtain 
 \begin{align} 
-({\mathbf{\cal C}}_{nm})_{;\mathbf{k}}
+(\calc_{nm})_{;\bfk}
 &=
@@
 \sum_{q} 
- \left(\mathbf{r}_{nq}
-{\mathbf{\cal C}}_{qm}
+ \left(\bfr_{nq}
+\calc_{qm}
 -
-{\mathbf{\cal C}}_{nq}
-\mathbf{r}_{qm}
+\calc_{nq}
+\bfr_{qm}
 \right) 
-+i\mathbf{r}_{nm}({\mathbf{\cal C}}_{mm}-{\mathbf{\cal C}}_{nn}) 
-,
-\end{align} 
-where we remind the reader that $\mathbf{r}_{nm}$ 
++i\bfr_{nm}(\calc_{mm}-\calc_{nn}) 
+,
+\end{align} 
+where we remind the reader that $\bfr_{nm}$ 
 is calculated through 
@@
 \subsection{
-\texorpdfstring{$\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}$}{Vnonlocal},
-\texorpdfstring{$\mathbf{v}^\mathrm{nl}_{nm}$}{Vnonlocal}, and 
-\texorpdfstring{$T^{\mathrm{a}\mathrm{b}}_{nm}$}{[r,vnl]}}
+\texorpdfstring{$\calbv^{\nl}_{nm}$}{Vnonlocal},
+\texorpdfstring{$\bfv^\nl_{nm}$}{Vnonlocal}, and 
+\texorpdfstring{$T^{\rma\rmb}_{nm}$}{[r,vnl]}}
 \label{vesnl}
@@
 and use
-$\sum_{\mathbf{G}} | \mathbf{k}+\mathbf{G}\rangle \langle\mathbf{k}+\mathbf{G} | =1$, 
+$\sum_{\bfG}\ket{\bfk+\bfG}\bra{\bfk+\bfG}=1$, 
 with
-$\langle\mathbf{r}|\mathbf{k}+\mathbf{G}=(1/\sqrt{\Omega})
-\mathrm{exp}(i(\mathbf{k}+\mathbf{G})\cdot\mathbf{r})$
+$\braket{\bfr}{\bfk+\bfG}=(1/\sqrt{\gO})
+\mathrm{exp}(i(\bfk+\bfG)\cdot\bfr)$
 to obtain,
 \begin{align}\label{vnl.5}
-\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}(\mathbf{k})&=\frac{1}{2}
-\sum_{\mathbf{G}}
+\calbv^{\nl}_{nm}(\bfk)&=\frac{1}{2}
+\sum_{\bfG}
 \Big(
-\langle n\mathbf{k} | {\mathbf{\cal C}}(z) 
-| \mathbf{k}+\mathbf{G}\rangle\langle\mathbf{k}+\mathbf{G} |
-\mathbf{v}^\mathrm{nl} | m\mathbf{k}\rangle
-+
-\langle n\mathbf{k} |
-\mathbf{v}^\mathrm{nl}  
-| \mathbf{k}+\mathbf{G}\rangle\langle\mathbf{k}+\mathbf{G} |
-{\mathbf{\cal C}}(z) | m\mathbf{k}\rangle
+\bra{n\bfk} \calc(z) 
+\ket{\bfk+\bfG}\bra{\bfk+\bfG}
+\bfv^\nl \ket{m\bfk}
++
+\bra{n\bfk}
+\bfv^\nl  
+\ket{\bfk+\bfG}\bra{\bfk+\bfG}
+\calc(z) \ket{m\bfk}
 \Big)
@@
 &=\frac{1}{2 \hbar}
-\sum_{\mathbf{G}}
+\sum_{\bfG}
 \left(
-{\mathbf{\cal F}}^{\ell*}_{n\mathbf{k}}(\mathbf{G}) 
-H_{m\mathbf{k}}(\mathbf{G}) 
-+
-H^*_{n\mathbf{k}}(\mathbf{G}) 
-{\mathbf{\cal F}}_{m\mathbf{k}}(\mathbf{G}) 
+\calf^{\ell*}_{n\bfk}(\bfG) 
+H_{m\bfk}(\bfG) 
++
+H^*_{n\bfk}(\bfG) 
+\calf_{m\bfk}(\bfG) 
 \right) 
@@
 \begin{align}\label{vnl.10}
-{\mathbf{\cal F}}_{n\mathbf{k}}(\mathbf{G}) 
+\calf_{n\bfk}(\bfG) 
 =
-\sum_{\mathbf{G}'} 
-A_{n\mathbf{k}}(\mathbf{G}') 
-\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel}f(G'_\perp-G_\perp) 
+\sum_{\bfG'} 
+A_{n\bfk}(\bfG') 
+\gd_{\bfG_\parallel \bfG'_\parallel}f(G'_\perp-G_\perp) 
 ,
@@
 \begin{align}\label{vnl.11}
-H_{n\mathbf{k}}(\mathbf{G})&=
-\sum_{\mathbf{G}'} 
-A_{n\mathbf{k}}(\mathbf{G}') 
-(\nabla_{\mathbf{K}}+\nabla_{\mathbf{K}'})  
-V^\mathrm{nl}(\mathbf{K},\mathbf{K}')
-,
-\end{align}
-and $\mathbf{K}=\mathbf{k}+\mathbf{G}$, $\mathbf{K}'=\mathbf{k}+\mathbf{G}'$.
+H_{n\bfk}(\bfG)&=
+\sum_{\bfG'} 
+A_{n\bfk}(\bfG') 
+(\nabla_{\bfK}+\nabla_{\bfK'})  
+V^\nl(\bfK,\bfK')
+,
+\end{align}
+and $\bfK=\bfk+\bfG$, $\bfK'=\bfk+\bfG'$.
 For fully  separable pseudopotentials in the   
@@
 matrix elements 
- $\langle\mathbf{K} |
-V^\mathrm{nl}  
-| \mathbf{K}'\rangle
-=V^\mathrm{nl}(\mathbf{K},\mathbf{K}')  
+ $\bra{\bfK}  
+V^\nl  
+\ket{\bfK'}
+=V^\nl(\bfK,\bfK')  
 $ 
-and their $\mathbf{K}$ and $\mathbf{K}'$ gradient 
+and their $\bfK$ and $\bfK'$ gradient 
 can be readily calculated.\cite{mottaCMS10,adolphPRB96,gordienkoRPJ04,fuchsCPC99} 
@@
 implemented 
-the calculation of $\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}(\mathbf{k})$ with the help of 
-the DP\texttrademark~code.\cite{francesco}
-We obtain $\mathbf{v}^{\mathrm{nl}}_{nm}(\mathbf{k})$ from the previous three 
-expressions by taking ${\mathbf{\cal C}}(z)=1$, which implies 
-that $f(G_\perp)=\delta_{G_\perp 0}$, 
+the calculation of $\calbv^{\nl}_{nm}(\bfk)$ with the help of 
+the \depe~code.\cite{francesco}
+We obtain $\bfv^{\nl}_{nm}(\bfk)$ from the previous three 
+expressions by taking $\calc(z)=1$, which implies 
+that $f(G_\perp)=\gd_{G_\perp 0}$, 
 and 
-${\mathbf{\cal F}}_{n\mathbf{k}}(\mathbf{G}) \to F_{n\mathbf{k}}(\mathbf{G})=A_{n\mathbf{k}}(\mathbf{G})$.
+$\calf_{n\bfk}(\bfG) \to F_{n\bfk}(\bfG)=A_{n\bfk}(\bfG)$.
 
@@
 \begin{align}\label{3.1}
-L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
+L^{\rma\rmb}_{nm}(\bfk) 
 =
 \frac{1}{\hbar^2}
-\langle n\mathbf{k} |
-\big[ r^{\mathrm{a}},[ V^\mathrm{nl}, r^\mathrm{b}]\big]
-| m\mathbf{k}\rangle
+\bra{n\bfk}
+\big[ r^{\rma},[ V^\nl, r^\rmb]\big]
+\ket{m\bfk}
 ,
@@
 \begin{align}\label{3.5}
-L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
+L^{\rma\rmb}_{nm}(\bfk) 
 &=
-\frac{1}{\hbar^2\Omega}
-\sum_{\mathbf{G},\mathbf{G}'} 
-A^*_{n\mathbf{k}}(\mathbf{K}) 
-A_{m\mathbf{k}}(\mathbf{K}')
+\frac{1}{\hbar^2\gO}
+\sum_{\bfG,\bfG'} 
+A^*_{n\bfk}(\bfK) 
+A_{m\bfk}(\bfK')
 \int
-d\mathbf{r} d\mathbf{r}'
- e^{-i\mathbf{K}\cdot\mathbf{r}}
+d\bfr d\bfr'
+ e^{-i\bfK\cdot\bfr}
 \Big(
-r^{\mathrm{a}}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-r'^\mathrm{b}
+r^{\rma}
+V^\nl(\bfr,\bfr')
+r'^\rmb
 -
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-r'^\mathrm{a}
-r'^{\mathrm{b}}
+V^\nl(\bfr,\bfr')
+r'^\rma
+r'^{\rmb}
 \nonumber\\
 &-
-r^\mathrm{b}
-r^{\mathrm{a}}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-+
- r^\mathrm{b}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-r'^{\mathrm{a}}
+r^\rmb
+r^{\rma}
+V^\nl(\bfr,\bfr')
++
+ r^\rmb
+V^\nl(\bfr,\bfr')
+r'^{\rma}
 \Big) 
- e^{i\mathbf{K}'\cdot\mathbf{r}'}
+ e^{i\bfK'\cdot\bfr'}
 ,
@@
 where 
-$V^\mathrm{nl}(\mathbf{r},\mathbf{r}') = \langle\mathbf{r} | V^\mathrm{nl} | \mathbf{r}'\rangle$.
+$V^\nl(\bfr,\bfr') = \bra{\bfr} V^\nl\ket{\bfr'}$.
 We use the following identity
@@
 \Big(
-\frac{\partial^2}{\partial K^\mathrm{a}\partial K'^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K'^\mathrm{a}\partial K'^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K^\mathrm{a}\partial K^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K^\mathrm{b}\partial K'^\mathrm{a}}
+\frac{\partial^2}{\partial K^\rma\partial K'^\rmb}
++
+\frac{\partial^2}{\partial K'^\rma\partial K'^\rmb}
++
+\frac{\partial^2}{\partial K^\rma\partial K^\rmb}
++
+\frac{\partial^2}{\partial K^\rmb\partial K'^\rma}
 \Big)
 \int 
-d\mathbf{r} d\mathbf{r}' 
- e^{-i\mathbf{K}\cdot\mathbf{r}}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-e^{i\mathbf{K}'\cdot\mathbf{r}'}
+d\bfr d\bfr' 
+ e^{-i\bfK\cdot\bfr}
+V^\nl(\bfr,\bfr') 
+e^{i\bfK'\cdot\bfr'}
 \nonumber\\
 &=
-\int d\mathbf{r} d\mathbf{r}'
- e^{-i\mathbf{K}\cdot\mathbf{r}}
+\int d\bfr d\bfr'
+ e^{-i\bfK\cdot\bfr}
 \Big( 
-r^{\mathrm{a}} 
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-r'^\mathrm{b}
+r^{\rma} 
+V^\nl(\bfr,\bfr') 
+r'^\rmb
 - 
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-r'^\mathrm{a} 
-r'^{\mathrm{b}}
+V^\nl(\bfr,\bfr') 
+r'^\rma 
+r'^{\rmb}
 - 
-r^\mathrm{b} 
-r^{\mathrm{a}} 
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-+
- r^\mathrm{b} 
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-r'^{\mathrm{a}}
+r^\rmb 
+r^{\rma} 
+V^\nl(\bfr,\bfr')
++
+ r^\rmb 
+V^\nl(\bfr,\bfr') 
+r'^{\rma}
 \Big)  
-e^{i\mathbf{K}'\cdot\mathbf{r}'}
+e^{i\bfK'\cdot\bfr'}
 ,
@@
 \begin{align}\label{3.7}
-L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k})
+L^{\rma\rmb}_{nm}(\bfk)
 &=
 \frac{1}{\hbar^2}
-\sum_{\mathbf{G},\mathbf{G}'} 
-A^*_{n\mathbf{k}}(\mathbf{K}) 
-A_{m\mathbf{k}}(\mathbf{K}')
+\sum_{\bfG,\bfG'} 
+A^*_{n\bfk}(\bfK) 
+A_{m\bfk}(\bfK')
 \Big(
-\frac{\partial^2}{\partial K^\mathrm{a}\partial K'^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K'^\mathrm{a}\partial K'^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K^\mathrm{a}\partial K^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K^\mathrm{b}\partial K'^\mathrm{a}}
+\frac{\partial^2}{\partial K^\rma\partial K'^\rmb}
++
+\frac{\partial^2}{\partial K'^\rma\partial K'^\rmb}
++
+\frac{\partial^2}{\partial K^\rma\partial K^\rmb}
++
+\frac{\partial^2}{\partial K^\rmb\partial K'^\rma}
 \Big)
-\langle \mathbf{K} |
-V^\mathrm{nl} 
-| \mathbf{K}'\rangle, 
+\bra{\bfK} 
+V^\nl 
+\ket{\bfK'}, 
 \nonumber\\
@@
 \begin{align}\label{vkk}
-\langle\mathbf{K} |
-V^\mathrm{nl} 
-| \mathbf{K}'\rangle
+\bra{\bfK} 
+V^\nl 
+\ket{\bfK'} 
 =
-\frac{1}{\Omega}
+\frac{1}{\gO}
 \int 
-d\mathbf{r} d\mathbf{r}' 
- e^{-i\mathbf{K}\cdot\mathbf{r}}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-e^{i\mathbf{K}'\cdot\mathbf{r}'}
-,
-\end{align}
-The double derivatives with respect to $\mathbf{K}$ and $\mathbf{K}'$ 
+d\bfr d\bfr' 
+ e^{-i\bfK\cdot\bfr}
+V^\nl(\bfr,\bfr') 
+e^{i\bfK'\cdot\bfr'}
+,
+\end{align}
+The double derivatives with respect to $\bfK$ and $\bfK'$ 
 can be worked out explicitly and 
-$L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k})$
+$L^{\rma\rmb}_{nm}(\bfk)$
 can be finally calculated.\cite{valerie}
@@
 %%%%
-\subsection{Expressions for  \texorpdfstring{$\boldsymbol{\mathcal{\cal V}}^{S}_{nm}$}{Vnm}
+\subsection{Expressions for  \texorpdfstring{$\calbv^{S}_{nm}$}{Vnm}
 and
-\texorpdfstring{$\Big({\boldsymbol{\mathcal{\cal V}}}^{S}_{nm}\Big)_{;\mathbf{k}}$}{(Vnm);kb}
+\texorpdfstring{$\Big({\calbv}^{S}_{nm}\Big)_{;\bfk}$}{(Vnm);kb}
 }\label{calvs}
@@
 \begin{align}\label{a.3b}
-\boldsymbol{\mathcal{\cal V}}^{S}_{nm}
+\calbv^{S}_{nm}
 &=
 \frac{1}{2}\sum_q\left(   
-\mathbf{v}^S_{nq}{\mathbf{\cal C}}_{qm}+{\mathbf{\cal C}}_{nq}\mathbf{v}^S_{qm}
+\bfv^S_{nq}\calc_{qm}+\calc_{nq}\bfv^S_{qm}
 \right)  
@@
 \end{align}    
-where $\sum_q | q\mathbf{k}\rangle\langle q\mathbf{k} | =1$ was used
-and $\mathbf{v}^S_{nm}$ is given in Eq.~\eqref{chon.2}.
+where $\sum_q\ket{q\bfk}\bra{q\bfk}=1$ was used
+and $\bfv^S_{nm}$ is given in Eq.~\eqref{chon.2}.
 Taking the generalized derivative of Eq.~\eqref{a.3b}
@@
 \begin{align}\label{a.3bn}
-\left(\boldsymbol{\mathcal{\cal V}}^{S}_{nm}\right)_{;\mathbf{k}}
+\left(\calbv^{S}_{nm}\right)_{;\bfk}
 &=
 \frac{1}{2}\sum_q\left(
-(\mathbf{v}^S_{nq})_{;\mathbf{k}}{\mathbf{\cal C}}_{qm}
-+    
-\mathbf{v}^S_{nq}({\mathbf{\cal C}}_{qm})_{;\mathbf{k}}
-+
-({\mathbf{\cal C}}_{nq})_{;\mathbf{k}} \mathbf{v}^S_{qm}
-+
-{\mathbf{\cal C}}_{nq} (\mathbf{v}^S_{qm})_{;\mathbf{k}}
+(\bfv^S_{nq})_{;\bfk}\calc_{qm}
++    
+\bfv^S_{nq}(\calc_{qm})_{;\bfk}
++
+(\calc_{nq})_{;\bfk} \bfv^S_{qm}
++
+\calc_{nq} (\bfv^S_{qm})_{;\bfk}
 \right)  
@@
 \begin{align}\label{choni.1}
-(\mathbf{v}^S_{nm})_{;\mathbf{k}}=i\Delta f_{mn}
-(\mathbf{r}_{nm})_{;\mathbf{k}}
+(\bfv^S_{nm})_{;\bfk}=i\gD f_{mn}
+(\bfr_{nm})_{;\bfk}
 ,

### 2014-11-21 paper-v-0.tex
finished correcting equations

--- 2014-11-21 paper-v-0.tex
@@
 \begin{equation}
-H_{I}(t)=-e\mathbf{r}\cdot \mathbf{E}(t).\label{rde}
+H_{I}(t)=-e\mathbf{r}\cdot \mathbf{E}(t).
+\label{rde}
 \end{equation}
@@
 For $\mathbf{r}_e$ one uses
-\begin{align}
-\langle n\mathbf{k} | \mathbf{r}_e | m\mathbf{k}'\rangle &=
+\begin{equation}
+\langle n\mathbf{k} | \mathbf{r}_e | m\mathbf{k}'\rangle =
 (1-\delta_{nm})\delta(\mathbf{k}-\mathbf{k}')\boldsymbol{\xi}_{nm}(\mathbf{k})
@@
 \label{rnminn}
-\end{align}
+\end{equation}
 such that $\mathbf{r}_{e,nm}=0$ for $n=m$.
@@
 we obtain
-\begin{equation}
+\begin{equation*}
 \left(\omega^\Sigma_{nm}\right)_{;k^{\mathrm{a}}}
@@
 .
-\end{equation} 
+\end{equation*} 
 The apparent divergence as $\tilde\omega\to 0$
@@
 relationship is satisfied for this particular slab
-\begin{equation}
+\begin{equation*}
 \chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega) 
@@
 ,
-\end{equation}
+\end{equation*}
 where
@@
 From Eqs.~\eqref{lab} and \eqref{vnl} 
-\begin{equation}
+\begin{equation*}
 L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
@@
 ,
-\end{equation} 
+\end{equation*} 
 and we expand the triple commutator to obtain,

### 2014-11-25 paper-v-0.tex
reverted intro and abstract to previous version

@@
 \begin{abstract}
-We formulate a theoretical approach of surface second-harmonic generation from 
-semiconductor surfaces based on the length gauge. We calculate the nonlinear 
-second-order surface susceptibility tensor 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
-independent particle approximation, and for the first time in one unique 
-formulation we include i) the scissors correction, needed to have the correct 
-value for the energy band gap, ii) the derivation for the inclusion of the cut 
-function used to extract the surface nonlinear response, and iii) the 
-contribution of the nonlocal part of the pseudopotentials that are routinely 
-used in \textit{ab initio} band structure calculations. The first and third 
-terms are described by spatially nonlocal quantum mechanical operators and 
-are fully taken into account in the present formulation. As a test-case for 
-this approach, we calculate $\chi^{xxx}(-2\omega;\omega,\omega)$ for a clean 
-Si(001)$2\times 1$ reconstructed surface and find that this scheme for 
-extracting the surface response through the cut function is exact and readily
-implemented. The effects of the scissors correction and the nonlocal part of 
-the pseudopotentials are discussed for the first time for surface nonlinear 
-optics. The scissors correction non-rigidly shifts the spectrum to higher 
-energies by mixing the $1\omega$ and $2\omega$ resonances and exerts a strong 
-influence on the line-shape. The effects of the nonlocal part of the 
-pseudopotentials keeps the same line-shape of 
-$|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$ but reduces its value by 
-15-20\%. Therefore, our scheme is unprecedented through the inclusion of 
-these three contributions and also opens the possibility to study SSHG with 
-more versatility and more accuracy than previous approaches. 
+We formulate a theoretical approach of surface second-harmonic generation 
+from semiconductor surfaces based on the length gauge. 
+Within the independent particle approximation 
+the nonlinear second-order surface susceptibility tensor
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$   
+is calculated, including for the first time in one unique formulation
+i) the scissors correction, needed to have the correct value of the
+energy band gap,
+ii) the derivation for the inclusion of the cut function, 
+used to extract the surface nonlinear response,
+and iii) the contribution of the nonlocal part of the 
+pseudopotentials, routinely  used in \textit{ab initio} band structure calculations. 
+First and third terms are described by spatially nonlocal quantum
+mechanical operators and are fully taken into account in the present formulation. 
+As a test-case of the approach we calculate 
+$\chi^{xxx}(-2\omega;\omega,\omega)$ for the clean Si(001)$2\times 1$
+reconstructed surface and find that our scheme to extract the surface
+through the cut function
+readily works and is exact. The effects of the scissors correction and
+of the nonlocal part of the pseudopotentials are discussed for the
+first time in surface nonlinear optics. 
+The scissors correction shifts the spectrum to 
+higher energies though the shifting is not rigid and mixes the 
+$1\omega$ and $2\omega$ resonances, 
+and has a strong influence in the line-shape. 
+The 
+effects of the nonlocal part of the pseudopotentials 
+keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$, but 
+reduces its value
+by 15-20\%.  
+Therefore, the inclusion of the three aforementioned 
+contributions is very important and 
+makes our scheme unprecedented and opens the possibility to 
+study SSHG with more versatility and providing more accurate results 
+than previous approaches. 
 \end{abstract}
+
+% \begin{abstract}
+% We formulate a theoretical approach of surface second-harmonic generation from 
+% semiconductor surfaces based on the length gauge. We calculate the nonlinear 
+% second-order surface susceptibility tensor 
+% $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
+% independent particle approximation, and for the first time in one unique 
+% formulation we include i) the scissors correction, needed to have the correct 
+% value for the energy band gap, ii) the derivation for the inclusion of the cut 
+% function used to extract the surface nonlinear response, and iii) the 
+% contribution of the nonlocal part of the pseudopotentials that are routinely 
+% used in \textit{ab initio} band structure calculations. The first and third 
+% terms are described by spatially nonlocal quantum mechanical operators and 
+% are fully taken into account in the present formulation. As a test-case for 
+% this approach, we calculate $\chi^{xxx}(-2\omega;\omega,\omega)$ for a clean 
+% Si(001)$2\times 1$ reconstructed surface and find that this scheme for 
+% extracting the surface response through the cut function is exact and readily
+% implemented. The effects of the scissors correction and the nonlocal part of 
+% the pseudopotentials are discussed for the first time for surface nonlinear 
+% optics. The scissors correction non-rigidly shifts the spectrum to higher 
+% energies by mixing the $1\omega$ and $2\omega$ resonances and exerts a strong 
+% influence on the line-shape. The effects of the nonlocal part of the 
+% pseudopotentials keeps the same line-shape of 
+% $|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$ but reduces its value by 
+% 15-20\%. Therefore, our scheme is unprecedented through the inclusion of 
+% these three contributions and also opens the possibility to study SSHG with 
+% more versatility and more accuracy than previous approaches. 
+% \end{abstract}
 
@@
 calculation and thus one uses the much simpler scissors operator scheme.
-\cite{levinePRL89,levinePRL91,delsolePRB93} This allows us to ``open'' the 
-DFT-LDA gap to its correct experimental or GW value even when experiments 
-are not available. In this context, to correct for the underestimation of 
-the energy band gap of semiconductors Nastos et al.\cite{nastosPRB05} used 
-the ``length gauge'' or ``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to 
-correctly include the many-body corrections through the scissors operator in 
-the SH susceptibility. Later, Cabellos \textit{et al}.\cite{cabellosPRB09} 
-elaborated a derivation of the ``velocity gauge'' or 
-``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the scissors operator 
-and proved gauge invariance with respect to the length gauge. From these works 
-it is clear the length gauge is a much better staring point to obtain the 
-surface SH (SSH) susceptibility, as will be elaborated in this article. 
-However, these considerations are only valid for bulk semiconductors.
+\cite{levinePRL89,levinePRL91,delsolePRB93}  
+This allows us to ``open'' the DFT-LDA gap to 
+its correct experimental or GW value even when experiments are not available. 
+In this context, to correct for the underestimation of the energy band gap of 
+semiconductors Nastos et al.\cite{nastosPRB05} used the ``length gauge'' or 
+``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to correctly include the 
+many-body corrections through the scissors operator in the SH susceptibility.
+Later, Cabellos \textit{et al}.\cite{cabellosPRB09} elaborated a derivation 
+of the ``velocity gauge'' or ``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the 
+scissors operator and proved gauge invariance with respect to the length 
+gauge. From these works it is clear the length gauge is a much better staring
+point to obtain the surface SH (SSH) susceptibility, as will be elaborated 
+in this article. However, these considerations are only valid for bulk 
+semiconductors.
 
@@
 length gauge formalism. This derivation includes the addition of the terms 
-mentioned above that have been absent in previous works. If desired, we can 
-use $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ for a 
-layer-by-layer analysis. In Sec. \ref{results} we show the results of this 
-reformulation with a study of a clean Si(001) surface with a $2 \times 1$ 
-reconstruction that proves the validity of our approach with a special 
-test-case that has never been implemented before. We compare results from 
-before and after adding the different nonlocal contributions. Finally, in 
-Sec. \ref{conc} we submit our conclusions.
+mentioned above that have been absent in previous works. Also, our 
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ can be used 
+for a layer-by-layer analysis if desired. In Sec. \ref{results} we show the 
+results of this reformulation with a study of a clean Si(001) surface with a 
+$2 \times 1$ reconstruction, 
+proving the correctness of our approach with a special test-case,
+never exploited before.
+We compare results from before and after adding 
+the different nonlocal contributions. Finally, in Sec. \ref{conc} we give 
+our conclusions.
+
+% \section{Introduction}\label{intro}
+
+% In recent years surface nonlinear optical spectroscopies, particulary surface 
+% second-harmonic generation (SSHG), have evolved as useful nondestructive 
+% and noninvasive tools to study surface and interface properties. These 
+% properties include atomic structure, phase transitions, adsorption of 
+% atoms, and many others.\cite{daumPRL93, mcgilpOE94, meyerPRL95, powerPRL95, 
+% godefroyAPL96, hoferAPA96, dadapPRB97, bloembergenAPB99, mcgilpSRL99, 
+% suzukiAPB99, mitchellSS01, hughesPRB96, guyotPRB88, downerPSSA01, shenAPB99, 
+% shenNAT89, chenPRL81, mendozaPRL98, downerSIA01} Nowadays, SSHG spectroscopy 
+% is a crucial tool for research and development in microelectronics,
+% \cite{zheltikovLP00} semiconductors, \cite{lupkeSSR99} nanomaterials,
+% \cite{salazar-aparicioPRB14} and many more recent areas of great scientific 
+% and commercial interest.\cite{cazzanelliNM14} The high surface sensitivity 
+% of SSHG spectroscopy is due to the fact that within the dipole approximation 
+% the bulk SHG signal of centrosymmetric materials is identically zero. 
+% Radiation can occur only at the surface where the inversion symmetry is 
+% broken. Much of the foundation of surface science has been built from 
+% experiments involving emission or scattering of electrons from surfaces. 
+% These require ultrahigh vacuum (UHV) environments and provide no access to 
+% buried interfaces. However, SSHG is compatibile with non-UHV conditions and 
+% has access to interfaces buried beneath transparent overlayers. Even when 
+% applied to surfaces in UHV, the light source and detectors can be aligned and
+% used outside the vacuum chamber.
+
+% The usefulness of SSHG could be limited by the lack of microscopic theoretical
+% understanding of the nonlinear spectra. The macroscopic phenomenological 
+% theory of SSHG, which relates the intensity, phase, and polarization of 
+% detected fields to nonlinear and linear susceptibilities at the material 
+% interface is now fully developed. However, microscopic theory that relates 
+% atomic-level structure to the nonlinear source polarization is still being 
+% developed. In this article we put forward an approach to calculate the
+% microscopic second-harmonic surface susceptibility that encompasses several
+% theoretical features not taken into account before. Within the independent 
+% particle approximation (IPA) some frameworks for bulk SHG have been developed 
+% to study the nonlinear optical response of bulk materials with varying degrees 
+% of sophistication.\cite{butcherPOPS63, aspnesPRB72, sipePRB93, levinePRB94,
+% aversaPRB95, hughesPRB96, rashkeevPRB98}
+
+% The most used framework for \textit{ab initio} calculations, Density 
+% Functional Theory (DFT) within the Local Density Approximation (LDA),
+% \cite{kohnPR65} underestimates the energy band gap of semiconductors. It is 
+% well understood that one has to include the many-body interaction to correct 
+% for this underestimation of the gap. In this context, the so-called GW 
+% approximation\cite{onidaRMP02} is known to correct the band gap of most 
+% semiconductors\cite{luceroJPCM12}. However this can be a very expensive
+% calculation and thus one uses the much simpler scissors operator scheme.
+% \cite{levinePRL89,levinePRL91,delsolePRB93} This allows us to ``open'' the 
+% DFT-LDA gap to its correct experimental or GW value even when experiments 
+% are not available. In this context, to correct for the underestimation of 
+% the energy band gap of semiconductors Nastos et al.\cite{nastosPRB05} used 
+% the ``length gauge'' or ``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to 
+% correctly include the many-body corrections through the scissors operator in 
+% the SH susceptibility. Later, Cabellos \textit{et al}.\cite{cabellosPRB09} 
+% elaborated a derivation of the ``velocity gauge'' or 
+% ``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the scissors operator 
+% and proved gauge invariance with respect to the length gauge. From these works 
+% it is clear the length gauge is a much better staring point to obtain the 
+% surface SH (SSH) susceptibility, as will be elaborated in this article. 
+% However, these considerations are only valid for bulk semiconductors.
+
+% Concerning the optical response of surfaces and interfaces, Reining 
+% \textit{et al.}\cite{reiningPRB94} introduced the concept of a cut function 
+% in order to obtain the surface SH susceptibility tensor. This cut function 
+% is required since one usually uses a slab approach when treating semi-infinite 
+% surface systems. If the slab is centrosymmetric the susceptibility tensor will 
+% be identically zero. The cut function is such that it separates the nonlinear 
+% response for the two surfaces of the slab avoiding the destructive 
+% interference between them giving a finite value that one identifies with the 
+% SSH susceptibility tensor. If the slab is not centrosymmetric the cut function 
+% can be used to separate the different signals coming from either surface of 
+% the slab. Indeed, one of the main results of this article is to show that the 
+% SSH  susceptibility tensor obtained by using the cut function is correctly 
+% extracted from the slab. After Reining \textit{et al.},\cite{reiningPRB94}
+% Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02,sanoPRB02}
+% followed upon this work and in particular Ref.~\onlinecite{arzatePRB01} went 
+% into a  detailed analysis of the different contributions to the SHG spectra 
+% of a surface and the nuanced relationship between bulk, surface, interband, 
+% intraband, $1\omega$ and $2\omega$ terms, and
+% Ref. \onlinecite{mejiaRMF04} developed a layer-by-layer 
+% analysis for the nonlinear responses of semiconductor systems, within a 
+% tight-binding framework. This model allows for obtaining results from 
+% selected regions of a system including the surface. However, in these 
+% references the scissors operator is either excluded or incorrectly 
+% implemented.
+
+% Finally, DFT-LDA calculations are often based on the use of pseudopotentials.
+% As we will discuss later in this article, the presence of the nonlocal part 
+% of the pseudopotential introduces corrections to the momentum operator of 
+% the electron that have to be included with care into the SSH susceptibility. 
+% So far, this contribution has been neglected in the literature.
+
+% Therefore, within the IPA the most complete approach for the calculation of 
+% the SSH susceptibility is one which includes i) the scissors correction, 
+% ii) the cut function, and iii) the contribution of the nonlocal part of the 
+% pseudopotential. Therefore, the goal of this paper is to derive a new 
+% expression within the length gauge for the SSH susceptibility tensor 
+% $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ 
+% that include the aforementioned contributions. The inclusion of these three 
+% contributions makes our scheme unprecedented and opens the possibility to 
+% study surface SHG with more versatility and providing more accurate results 
+% than previous approaches.
+
+% The paper is organized as follows. In Sec. \ref{theory} we present the 
+% relevant steps for deriving the surface second-order susceptibility tensor
+% $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
+% length gauge formalism. This derivation includes the addition of the terms 
+% mentioned above that have been absent in previous works. If desired, we can 
+% use $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ for a 
+% layer-by-layer analysis. In Sec. \ref{results} we show the results of this 
+% reformulation with a study of a clean Si(001) surface with a $2 \times 1$ 
+% reconstruction that proves the validity of our approach with a special 
+% test-case that has never been implemented before. We compare results from 
+% before and after adding the different nonlocal contributions. Finally, in 
+% Sec. \ref{conc} we submit our conclusions.
 

### 2014-12-03 cpfigures.sh
almost ready

@@ -1,0 +1,7 @@
+pfix=pdf
+cp images/slab.$pfix sent/fig1.$pfix 
+cp images/si2x1-crop.$pfix sent/fig2.$pfix 
+cp plots/fig1.$pfix sent/fig3.$pfix 
+cp plots/fig2.$pfix sent/fig4.$pfix 
+cp plots/fig3.$pfix sent/fig5.$pfix 
+cp plots/fig4.$pfix sent/fig6.$pfix 

### 2014-12-03 manuscript.tex

--- 2014-11-21 paper-v-0.tex
@@
-%\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
-\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,letterpaper]{revtex4}
-\usepackage{amsfonts}
+\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
 \usepackage{amsmath}
 \usepackage{graphicx}
-\usepackage{ulem}
-\usepackage{subfigure}
-\usepackage{bm}
 \usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
 %\usepackage{showlabels}
-%\usepackage{showkeys}
-\allowdisplaybreaks[1]%biutiful equation breaker!!!
-
-%%%
-% notes:
-% check for tense consistency
-% Local fields?
-%%%
 
@@
 \begin{abstract}
-We formulate a theoretical approach of surface second-harmonic generation from 
-semiconductor surfaces based on the length gauge. We calculate the nonlinear 
-second-order surface susceptibility tensor 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
-independent particle approximation, and for the first time in one unique 
-formulation we include i) the scissors correction, needed to have the correct 
-value for the energy band gap, ii) the derivation for the inclusion of the cut 
-function used to extract the surface nonlinear response, and iii) the 
-contribution of the nonlocal part of the pseudopotentials that are routinely 
-used in \textit{ab initio} band structure calculations. The first and third 
-terms are described by spatially nonlocal quantum mechanical operators and 
-are fully taken into account in the present formulation. As a test-case for 
-this approach, we calculate $\chi^{xxx}(-2\omega;\omega,\omega)$ for a clean 
-Si(001)$2\times 1$ reconstructed surface and find that this scheme for 
-extracting the surface response through the cut function is exact and readily
-implemented. The effects of the scissors correction and the nonlocal part of 
-the pseudopotentials are discussed for the first time for surface nonlinear 
-optics. The scissors correction non-rigidly shifts the spectrum to higher 
-energies by mixing the $1\omega$ and $2\omega$ resonances and exerts a strong 
-influence on the line-shape. The effects of the nonlocal part of the 
-pseudopotentials keeps the same line-shape of 
-$|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$ but reduces its value by 
-15-20\%. Therefore, our scheme is unprecedented through the inclusion of 
-these three contributions and also opens the possibility to study SSHG with 
-more versatility and more accuracy than previous approaches. 
+We formulate a theoretical approach of surface second-harmonic generation 
+from semiconductor surfaces based on the length gauge and the 
+electron  
+density operator.  
+Within the independent particle approximation 
+the nonlinear second-order surface susceptibility tensor
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$    
+is calculated, including for the first time in one unique formulation
+(i) the scissors correction, needed to have the correct value of the
+energy band gap,
+(ii) the contribution of the nonlocal part of the  
+pseudopotentials, routinely  used in \textit{ab initio} band structure calculations, 
+and (iii) the derivation for the inclusion of the cut function, 
+used to extract the surface response.
+The first two contributions
+are described by spatially nonlocal quantum
+mechanical operators and are fully taken into account in the present formulation. 
+As a test-case of the approach we calculate 
+$\chi^{xxx}(-2\omega;\omega,\omega)$ for the clean Si(001)$2\times 1$
+reconstructed surface and find that our scheme to extract the surface
+through the cut function
+readily works and is exact. The effects of the scissors correction and
+of the nonlocal part of the pseudopotentials are discussed 
+in surface nonlinear optics. 
+The scissors correction shifts the spectrum to 
+higher energies though the shifting is not rigid and mixes the 
+$1\omega$ and $2\omega$ resonances, 
+and has a strong influence in the line-shape. 
+The 
+effects of the nonlocal part of the pseudopotentials 
+keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$, but 
+reduces its value
+by 15-20\%.  
+Therefore, the inclusion of the three aforementioned 
+contributions is very important and 
+makes our scheme unprecedented and opens the possibility to 
+study surface second-harmonic generation
+with more versatility and providing more accurate results.
 \end{abstract}
@@
 \maketitle
-%\tableofcontents
 
@@
 Radiation can occur only at the surface where the inversion symmetry is 
-broken. Much of the foundation of surface science has been built from 
+broken. 
+The bulk quadrupole contribution for centrosymmetric materials is
+different from zero, but usually it is very small,\cite{downerSIA01} and we neglect it. 
+Much of the foundation of surface science has been built from 
 experiments involving emission or scattering of electrons from surfaces. 
@@
 detected fields to nonlinear and linear susceptibilities at the material 
-interface is now fully developed. However, microscopic theory that relates 
-atomic-level structure to the nonlinear source polarization is still being 
-developed. In this article we put forward an approach to calculate the
+interface is now fully developed.\cite{downerSIA01}
+ However, microscopic theory that relates 
+electronic-level structure to the nonlinear source polarization is still being 
+developed. \cite{butcherPOPS63, aspnesPRB72, sipePRB93,
+  levinePRB94,aversaPRB95, hughesPRB96, rashkeevPRB98,beyond}  
+ Within the independent 
+particle approximation (IPA) some frameworks for bulk SHG have been developed 
+to study the nonlinear optical response of bulk materials. \cite{butcherPOPS63, aspnesPRB72, sipePRB93, levinePRB94,aversaPRB95, hughesPRB96, rashkeevPRB98} 
+In this article we put forward an approach to calculate the
 microscopic second-harmonic surface susceptibility that encompasses several
-theoretical features not taken into account before. Within the independent 
-particle approximation (IPA) some frameworks for bulk SHG have been developed 
-to study the nonlinear optical response of bulk materials with varying degrees 
-of sophistication.\cite{butcherPOPS63, aspnesPRB72, sipePRB93, levinePRB94,
-aversaPRB95, hughesPRB96, rashkeevPRB98}
+theoretical features not taken into account before 
+in the case of a surface.
 
@@
 for this underestimation of the gap. In this context, the so-called GW 
-approximation\cite{onidaRMP02} is known to correct the band gap of most 
+approximation\cite{onidaRMP02} is known to correct the electronic gap of most 
 semiconductors\cite{luceroJPCM12}. However this can be a very expensive
 calculation and thus one uses the much simpler scissors operator scheme.
-\cite{levinePRL89,levinePRL91,delsolePRB93} This allows us to ``open'' the 
-DFT-LDA gap to its correct experimental or GW value even when experiments 
-are not available. In this context, to correct for the underestimation of 
-the energy band gap of semiconductors Nastos et al.\cite{nastosPRB05} used 
-the ``length gauge'' or ``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to 
-correctly include the many-body corrections through the scissors operator in 
-the SH susceptibility. Later, Cabellos \textit{et al}.\cite{cabellosPRB09} 
-elaborated a derivation of the ``velocity gauge'' or 
-``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the scissors operator 
-and proved gauge invariance with respect to the length gauge. From these works 
-it is clear the length gauge is a much better staring point to obtain the 
-surface SH (SSH) susceptibility, as will be elaborated in this article. 
-However, these considerations are only valid for bulk semiconductors.
+\cite{levinePRL89,levinePRL91,delsolePRB93}  
+This allows us to ``open'' the DFT-LDA gap to 
+its correct experimental or GW value even when experiments are not available. 
+In this context, to correct for the underestimation of the energy band gap of 
+semiconductors Nastos et al.\cite{nastosPRB05} used the ``length gauge'' or 
+``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to correctly include the 
+many-body corrections through the scissors operator in the SH susceptibility.
+Later, Cabellos \textit{et al}.\cite{cabellosPRB09} elaborated a derivation 
+of the ``velocity gauge'' or ``$\mathbf{A}\cdot\mathbf{v}$'' gauge properly including the 
+scissors operator and proved gauge invariance with respect to the length 
+gauge. From these works it is clear the length gauge is a much better starting
+point to obtain the surface SH (SSH) susceptibility, as will be elaborated 
+in this article. However, these considerations are only valid for bulk 
+semiconductors.
 
@@
 is required since one usually uses a slab approach when treating semi-infinite 
-surface systems. If the slab is centrosymmetric the susceptibility tensor will 
+surface systems.\cite{reiningPRB94} If the slab is centrosymmetric the susceptibility tensor will 
 be identically zero. The cut function is such that it separates the nonlinear 
@@
 Finally, DFT-LDA calculations are often based on the use of pseudopotentials.
-As we will discuss later in this article, the presence of the nonlocal part 
-of the pseudopotential introduces corrections to the momentum operator of 
-the electron that have to be included with care into the SSH susceptibility. 
-So far, this contribution has been neglected in the literature.
+As it will be discussed 
+in this article, the presence of a nonlocal part of the
+pseudopotential introduces
+corrections to
+the momentum operator of the electron that have to be included with care in 
+the SSH susceptibility. 
+For the bulk counterpart see for instance Refs.~\onlinecite{ismailPRL01,luppiPRB08}.
 
 Therefore, within the IPA the most complete approach for the calculation of 
-the SSH susceptibility is one which includes i) the scissors correction, 
-ii) the cut function, and iii) the contribution of the nonlocal part of the 
-pseudopotential. Therefore, the goal of this paper is to derive a new 
+the SSH susceptibility is one which includes (i) the scissors correction, 
+(ii) the contribution of the nonlocal part of the 
+pseudopotential, and
+(iii) the cut function.
+ Therefore, the goal of this paper is to derive a new 
 expression within the length gauge for the SSH susceptibility tensor 
 $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ 
-that include the aforementioned contributions. The inclusion of these three 
+that includes the aforementioned contributions. The inclusion of these three 
 contributions makes our scheme unprecedented and opens the possibility to 
-study surface SHG with more versatility and providing more accurate results 
-than previous approaches.
+study surface SHG with more versatility and providing accurate results.
 
@@
 length gauge formalism. This derivation includes the addition of the terms 
-mentioned above that have been absent in previous works. If desired, we can 
-use $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ for a 
-layer-by-layer analysis. In Sec. \ref{results} we show the results of this 
-reformulation with a study of a clean Si(001) surface with a $2 \times 1$ 
-reconstruction that proves the validity of our approach with a special 
-test-case that has never been implemented before. We compare results from 
-before and after adding the different nonlocal contributions. Finally, in 
-Sec. \ref{conc} we submit our conclusions.
+mentioned above that have been absent in previous works. Also, our 
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ can be used 
+for a layer-by-layer analysis if desired. 
+Nevertheless, such an analysis could depend on the choice of the basis
+used to expand the wavefunctions.
+In Sec. \ref{results} we show the 
+results of this reformulation with a study of a clean Si(001) surface with a 
+$2 \times 1$ reconstruction, 
+proving the correctness of our approach with a special test-case,
+never exploited before.
+We compare results from before and after adding 
+the different nonlocal contributions. Finally, in Sec. \ref{conc} we give 
+our conclusions.
 
@@
 In this section we present the scheme used to calculate the surface 
-second-order nonlinear response using the length gauge formalism. Our 
-derivation includes the aforementioned terms not included before for 
-this response.
-
-\subsection{Perturbation approach}
+second-order nonlinear response using the length gauge formalism and
+the electron 
+density operator. 
+Some of the results presented  in Subsections \ref{pa} and \ref{lgf}
+have already been discussed in earlier
+studies and can be applied to bulk and surface studies.\cite{sipePRB00,cabellosPRB09}
+We present them in order to have a self-contained derivation.
+The terms presented in these Subsections are used in Subsection \ref{cd}, 
+where we derive the expression 
+for the surface $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$.
+
+\subsection{Perturbation approach}\label{pa}
 
@@
 system using 
-%a scaled 
 the one electron density operator ${\rho}$, with which we can calculate the 
@@
 with the electromagnetic field.
-% ; $H_{0}$ has eigenvalues
-% $\hbar \omega_{n}(\mathbf{k})$
-% and eigenstates $| n\mathbf{k} \rangle$ (Bloch states) labeled by a band 
-% index $n$ and crystal momentum $\mathbf{k}$.
 To proceed with the solution of $\rho$ it is convenient to use the interaction 
@@
 $f_{n}=1$ when $n$ is a valence ($v$) or occupied band, and zero when $n$ is 
-a conduction ($c$) or empty band. We assume this for the remainder of the 
-article. The expectation values must satisfy
+a conduction ($c$) or empty band. 
+We assume this for the remainder of the  
+article.  
+As we neglect spin-orbit coupling, the
+final expression for
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$   
+has to be multiplied by a factor of 2
+to account for spin-degeneracy.
+The expectation values must satisfy
 $\langle{\cal O}\rangle = \mbox{Tr}({\rho}{\cal O}) = 
@@
 \begin{equation}
-{\mathbf{\cal P}}^a(2\omega)=\chi ^{abc}(-2\omega;\omega,\omega)E^{b}(\omega)E^{c}(\omega),  
+{\mathbf{\cal P}}^{\mathrm{a}}(2\omega)=
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)
+E^{\mathrm{b}}(\omega)E^{\mathrm{c}}(\omega),  
 \label{pshg}
 \end{equation}
-where $\chi^{abc}(-2\omega ;\omega ,\omega )$ is the nonlinear
+where $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega ;\omega ,\omega )$ is the nonlinear
 susceptibility responsible for surface second-harmonic generation
@@
 superscripts in Eq.~\eqref{pshg} denote Cartesian components, and if
-repeated are to be summed over. Without loss of generality we can always
-define $\chi^{abc}(-2\omega;\omega,\omega)$
+repeated are to be summed over. Without loss of generality we will
+define $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$
  to satisfy intrinsic permutation
 symmetry, 
-$\chi^{abc}(-2\omega ;\omega ,\omega )=\chi ^{acb}(-2\omega ;\omega
-,\omega )$.
-% ;
-% if it did not, the part that did not satisfy intrinsic permutation symmetric
-% would make no contribution to the second order polarization. This part
-% could be dropped since it would
-% have no physical significance.
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega ;\omega ,\omega )
+=\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega ;\omega,\omega )$.
 
@@
 To be more general in our derivation of
-$\chi^{abc}(-2\omega;\omega,\omega)$, we asume
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$, we asume
 contribution as is customary for most
@@
 where 
-$V(\mathbf{r})$ and $V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})$  are the local and nonlocal parts, 
+$V(\mathbf{r})$ and 
+$V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})$  
+are the local and nonlocal parts, 
  respectively.
+The argument $(\mathbf{r},\mathbf{p})$
+  is equivalente to the
+explicit $(\mathbf{r},\mathbf{r}')$ nonlocal notation.\cite{ismailPRL01}
 In case of
@@
 like that of all-electron schemes, 
-we simply omit the contribution coming
-the results that we have derived.
+we simply omit the contribution of $V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})$
+from the results that we have derived.
 
@@
 corresponds to the accepted 
-experimental band gap.\cite{levinePRL89,levinePRL91,delsolePRB93} 
+experimental electronic band gap.\cite{levinePRL89,levinePRL91,delsolePRB93} 
 This is often in fairly good agreement with the GW
@@
 applied. 
-%Several properties of $S(\mathbf{r},\mathbf{p})$ are shown in the
-%appendix.
 The unscissored and scissored Hamiltonians satisfy 
@@
 by the band index $n$ and the crystal momentum $\mathbf{k}$, and $u_{n\mathbf{k}}(\mathbf{r})$
-is cell periodic.
-
-\subsection{Length Gauge Formalism}
+are cell periodic.
+
+\subsection{Length Gauge Formalism}\label{lgf}
+
 To calculate the optical response 
@@
 In particular, Ref.~\onlinecite{aversaPRB95} develops
-the correct treatment of $\mathbf{r}$ for
+the treatment of $\mathbf{r}$ for
 the calculation of the nonlinear optical response.
@@
 $D_m$ are all the possible degenerate $m$-states. 
-%We recognize that for each instance in which 
-%$n\notin D_m$ or $n\ne m$, $\mathbf{r}_{nm}(\mathbf{k})$ are interband matrix
-%elements.
 For the intraband part, $\mathbf{r}_i$ only appears in
@@
 with $f_{nm}\equiv f_n-f_m$,
-where we see that $\mathbf{v}^S_{nn}=0$. It follows that
-\begin{equation}
-\mathbf{v}^\Sigma_{nm}(\mathbf{k}) 
-=\frac{\omega^\Sigma_{nm}(\mathbf{k})}{\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
-\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k}) 
-,
-\label{chon.9}
-\end{equation}
-and
-\begin{equation*}
-\frac{\mathbf{v}^\Sigma_{nm}(\mathbf{k})}{\omega^\Sigma_{nm}(\mathbf{k})}
-=
-\frac{\mathbf{v}^\mathrm{LDA}_{nm}(\mathbf{k})}{\omega^\mathrm{LDA}_{nm}(\mathbf{k})}
-,
-\end{equation*}
-and from Eq.~\eqref{pmnrmn},
+where we see that $\mathbf{v}^S_{nn}=0$. From Eqs.~\eqref{pmnrmn} and
+\eqref{vop2}
+ it follows that
 \begin{equation}
@@
 Kleinman-Bylander 
-form,\cite{francesco,mottaCMS10,kleinmanPRL82,adolphPRB96}
-as shown in Appendix \ref{vesnl}.
-The advantage of the length gauge formalism for 
-calculating linear and nonlinear optical responses 
+form.\cite{olevanoPC,mottaCMS10,kleinmanPRL82,adolphPRB96}
+The advantage of using the electron density operator along with 
+the length gauge formalism for 
+calculating linear and nonlinear optical responses,
+for the scissored Hamiltonian,
 resides in the ease with which the scissors operator
-can be introduced into the calculation, also described in detail in
-Nastos et al.\cite{nastosPRB05} The 
-length gauge formalism for the scissored Hamiltonian can be easily worked 
-out by simply using the unscissored LDA Hamiltonian,
+can be introduced into the calculation
+by simply using the unscissored LDA Hamiltonian,
 $H_0^{\mathrm{LDA}}$,
@@
 to obtain the scissored results for any 
-susceptibility expression, whether linear or nonlinear.
+susceptibility expression, whether linear or nonlinear.\cite{nastosPRB05}
 
@@
 term vanishes identically. 
-%In Appendix \ref{gender} we show how the generalized derivative of 
-%$(B^a_{nm})_{;k^b}$ can be obtained.
 Note that the indices in Eq.~\eqref{j.2} are all different from each
@@
 
-% In this section, we derive the expressions for the microscopic current
-% density of a given layer in the unit cell of the system.
-
 The approach we use to study the surface of a semi-infinite
@@
 slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
-super-cell constitutes the unit cell, and
-is repeated to form a full three dimensional crystalline structure,
-i.e. a repeated super-cell scheme. 
+super-cell
+is repeated to form a full three dimensional crystalline structure.
 The slab itself consists of front, back, and 
@@
 and the back surface $\chi^{b,\mathrm{a}\mathrm{b}\mathrm{c}}$ from the slab
-susceptibility 
-$\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}=\chi^{f,\mathrm{a}\mathrm{b}\mathrm{c}}-\chi^{b,\mathrm{a}\mathrm{b}\mathrm{c}}$.
-We have omitted the frequency arguments of $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$ for 
+susceptibility. 
+We have omitted the frequency dependence of $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$ for 
 convenience of notation.
@@
 \caption{(color on line) A sketch of the super-cell. 
-The slab corresponds to the
-circles representing the atoms of the system.\label{fslab}} 
+The atomic slab corresponds to the
+circles representing the atoms of the system.
+\label{fslab}} 
 \end{figure}
@@
 \langle m\mathbf{k}|\mathbf{r}\rangle\langle\mathbf{r} | \mathbf{v}^\Sigma | n\mathbf{k}\rangle
-\right),
+\right).
 \label{jmic3}
 \end{equation}
-are the matrix elements of the microscopic current operator.
-% and we have used the fact that the matrix elements between states $\ket{n\mathbf{k}}$
-% are diagonal in $\mathbf{k}$, i.e. proportional to $\delta(\mathbf{k}-\mathbf{k}')$.
 Integrating the microscopic current $\mathbf{j}(\mathbf{r},t)$ over
@@
 \frac{1}{A}\int d^3r\, {\mathbf{\cal C}}(z)\, 
-\mathbf{J} (\mathbf{r},t)
+\mathbf{J}^{(N)} (\mathbf{r},t)
  \equiv \boldsymbol{\mathcal{\cal J}}^{(N)}(t),
@@
 \end{align*}
-%Here an integration by parts is performed on the third term of the
-%right-hand; since the $e^{-i\mathbf{k}\cdot\mathbf{r}}\psi_{n\mathbf{k}}(\mathbf{r})$
-%are periodic over the unit cell, the surface term vanishes. 
 where, we used the hermitian property of $\mathbf{v}^\Sigma$ and defined
@@
 \end{equation*} 
-%where the superscript $\ell$ is inherited from ${\mathbf{\cal C}}(z)$.
-% and we
-%supress the dependance on $z$ from the increasingly crowded notation.  
 We see that the replacement
@@
 contribution of the region of the slab given by ${\mathbf{\cal C}}(z)$.
-We note that this modified operator is hermitian as it must.\cite{note2}
+We note that this modified operator is hermitian as it should.\cite{note2}
 The operator $\mathbf{V}$ could be any of those given by Eq.~\eqref{vop2},
@@
 according to the expressions of
-Appendices \ref{calpcalc}, \ref{vesnl} and \ref{calvs}.
+Appendices \ref{calpcalc} and \ref{calvs}.
 If not stated differently, calligraphic letters correspond to layer quantities.
 
-The layer-by-layer analysis of Refs. \onlinecite{hoganPRB03,castilloPRB03,mottaCMS10}  
-used ${\mathbf{\cal C}}(z)$
-%limiting the current response  
-%to a particular layer of the slab and used  
-to obtain the  
-anisotropic linear optical response of semiconductor surfaces, 
-where, the first formal derivation  
-of this scheme  
-was presented in  
-Ref.~\onlinecite{mendozaPRB06}. 
 To limit the SHG response to one surface, Eq.~\eqref{vcali} 
@@
 of Eq.~\eqref{Pjikn2}:
-%\end{document}  
 \begin{equation}
@@
 into Eq.~\eqref{conmri3n} and then
-Eq.~\eqref{chon.9}
+Eq.~\eqref{chon.10}
 we obtain
@@
  a partial fraction expansion over $\tilde\omega$. 
-Using time-reversal invariance, an integration by parts to 
+Using time-reversal symmetry, an integration by parts to 
 remove the square in the denominator of the second term of Eq.~\eqref{gene2}, 
@@
 convenience of notation.
+The factor of 2 for spin degeneracy is not included in Eq.~\eqref{chis}. 
 The real part of each contribution can be obtained through
@@
 $.
-To fulfill the required intrinsic permutation symmetry, %\cite{rashkeevPRB98} 
+To fulfill the required intrinsic permutation symmetry, 
 the $\{\}$ notation symmetrizes the $\mathrm{b}\mathrm{c}$ Cartesian indices, i.e. 
@@
 $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}=\chi^{\mathrm{a}\mathrm{c}\mathrm{b}}$.
-The various quantities involved in Eqs.~\eqref{chis} are given in
+The full expressions for
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$, along 
+with the various quantities involved in Eq.~\eqref{chis} are given in
 the Appendix \ref{appe}. 
@@
 that behaves like a bulk layer.
-%that at a distance $d$ from the surface  as seen in Fig.~\ref{fslab}.  
-%For a centrosymmetric slab, 
-%$\boldsymbol{\chi}^{\ell=N}=0$;
 We remark that the value of 
@@
 the slab needs to have enough atomic layers 
-%$\boldsymbol{\chi}^{\ell=N}=0$;
-%to be satisfied and to 
 in order to give converged results for 
@@
 either the front or the back surface.
-% This decomposition of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$ into
-% $\boldsymbol{\chi}^\ell(-2\omega;\omega,\omega)$ would allow to explore the contribution
-% to SH from any particular layer of the system,
-% and answers questions like: which part of the system has the strongest
-% contribution? or  how deep does the surface region needs to be to
-% recover centrosymmetry?, etc.
-\cite{mejiaRMF04}
-%\mathrextcolor{\chon}{Talk about the correct value of $d$}
 
@@
 of this section.  
-
-The self-consistent ground state and their Kohn-Sham states were
+Also, we checked that 
+for the
+dihydride surface
+$\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega)=0$.
+
+The self-consistent ground state and the Kohn-Sham states were
 calculated in the DFT-LDA framework using the plane-wave 
@@
 form.\cite{kleinmanPRL82}
-The contribution of $\mathbf{v}^\mathrm{nl}$ to Eq.~\eqref{chis} can be carried out 
-as detailed in Appendix \ref{vesnl} using the DP code for 
-the actual calculation.\cite{francesco}
-%
-The surfaces have been studied with a lattice constant of 5.43 \AA. 
-Structural optimizations were performed using DFT-LDA
+The contribution of $\mathbf{v}^\mathrm{nl}$ 
+and 
+$\boldsymbol{\mathcal{\cal V}}^\mathrm{nl}$ 
+to Eq.~\eqref{chis} is carried out 
+using the DP code.\cite{olevanoDP}
+The surfaces have been studied with the experimental lattice constant of 5.43 \AA. 
+Structural optimizations were performed 
 with the ABINIT code.\cite{abinit}  
-The geometry optimization as been carried out in slabs of 12
+The geometry optimization has been carried out in slabs of 12
 atomic layers where the central four layers where fixed at the bulk
@@
 tetrahedron method for evaluating the
-Brillouin zone integrals in Eq.~\eqref{chis}, 
-where special care was taken to examine the double resonances.\cite{nastosPRB05}
-Double resonances occur at a given frequency $\omega$ if there can be 
-resonant transitions at both $\omega$ and 2$\omega$ frequencies. This 
-implies that there is a region in the Brillouin zone where 
-$\omega_{cv}(\mathbf{k})=2 \omega_{c'v}(\mathbf{k})$. Since there is a real population 
-excitation, the perturbation theory used to calculate the spectrum breaks 
-down at these $\mathbf{k}$-points. This must be taken into account in a more 
-complete calculation. These points introduce sharp spikes in the spectrum. 
-In principle, these can affect the low frequency results since the response 
-is computed from the Kramers-Kronig relation for frequencies below the band 
-gap. However, in agreement with Ref.~\onlinecite{nastosPRB05} we find that 
-the double resonances affect the low frequency results by less than 2\%. 
+Brillouin zone integrals 
+where special care was taken to examine the double resonances
+of Eq.~\eqref{chis}. \cite{nastosPRB05}
 All spectra were calculated with a Gaussian smearing of 0.15 eV.
@@
 We must evaluate 
-$T^{\mathrm{a}\mathrm{b}}_{nm}=(i/\hbar)[r^\mathrm{b},v^{\mathrm{nl},\mathrm{a}}]_{nm}$,
-as outlined in Appendix \ref{vesnl},
+$T^{\mathrm{a}\mathrm{b}}_{nm}=(i/\hbar)[r^\mathrm{b},v^{\mathrm{nl},\mathrm{a}}]_{nm}$
 in order to obtain 
@@
 for the slab with
-16, 24, 32, and 40 Si atomic layers. 
+16, 24, 32, and 40 Si atomic layers, without the contribution of
+$\mathbf{v}^{\mathrm{nl}}$ and with no scissors correction.
+Since the clean Si(001) surface is $2\times 1$,
+there are two atoms per atomic layer, thus the total number of atoms per slab is 
+twice the number of atomic layers of the slab. 
 In making the slabs larger, we add steps of 8 layers of bulk-like atomic positions. 
@@
 when the value is small the difference increases only slightly, 
-but the spectra is so close to zero that these it is negligible. 
+but the spectra is so close to zero that it is negligible. 
 These differences would decrease as the number of atomic layers 
@@
 using the formalism contained in Eq.~\eqref{chis}.
+We have confirmed that for the dihydride surface
+$|\chi^{xxx}_{\mathrm{half-slab}}|\approx 0$ (not shown).
 This confirms the validity of our theory and is the main result of
 this article; through the proposed layer formalism we can calculate the surface SH
-pseudopotentials
-and the many-body effects through the scissors correction.
-Our scheme should to work for any slab.
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$    
+including
+the
+contribution of the nonlocal part of the pseudopotentials 
+and the part of the many-body effects through the scissors correction.
+Our scheme should work for any slab.
 
@@
 Notice that the 
-2$\omega$ resonances start below $E_g/2$ where $E_g=0.53$ eV is the LDA
-band gap of this surface, and $E_g=1.03$ eV is the scissor shifted gap for
-$\hbar\Delta=0.5$ eV.
+2$\omega$ resonances start below $E_g/2$ where $E_g$ is the band gap
+(0.53 eV  for LDA and
+1.03 eV if the scissor is used with
+$\hbar\Delta=0.5$ eV).
 These resonances
@@
 $2\times 1$ surface, that lie inside the bulk band gap of Si and are the 
-well known electronic surface states.
+well known electronic surface states.\cite{rohlfingPRB95} 
 In Fig.~\ref{fig3}
-We see that the effect of $\mathbf{v}^\mathrm{nl}$
+we see that the effect of $\mathbf{v}^\mathrm{nl}$
  reduces the value of   
@@
 showing the importance of this contribution for a correct calculation  
-of SSHG.  
+of SSHG, in agreement with the analysis for bulk semiconductors.\cite{luppiPRB08}
 However, the inclusion of $\mathbf{v}^\mathrm{nl}$ does not
@@
 the cases of zero scissors correction from Fig.~\ref{fig2}.
-%
 \begin{figure}
@@
 finite values for $\hbar\Delta$. The first one
-with a value of $\hbar\Delta=0.5$ eV, used in above results, 
+with a value of $\hbar\Delta=0.5$ eV, used in the above results, 
 is the ``average'' GW gap taken from 
@@
 and yield a more varied spectrum.
+We mention that for large gap materials, the 1$\omega$ and 2$\omega$ would be splited
+showing a small interference effect, but still the 2$\omega$ would
+strongly depend on the surface states.
  This way we see that small changes in the value of the 
@@
 spectrum quite dramatically.
+In Ref.~\onlinecite{adolphPRB00}, the authors already remarked
+that nonlinear optical response of bulk materials
+is more influenced by the electronic structure
+of the material than the linear case. 
+In the case of semiconducting surfaces the
+problem is even more intricate due to the presence of electronic
+surface states.
+The high sensitivity of SSHG to the energy position  of surface
+states, as seen in Fig.~\ref{fig4}, makes SSHG a good benchmark
+spectroscopical tool for testing the validity of the inclusion of many-body effects,
+and in particular the quasi-particle correction to the electronic states. 
 \begin{figure}
@@
 So, we would expect that the $xxx$ component of 
-$\boldsymbol{\chi}(-2\omega;\omega,\omega)$ would have a small influence from the local fields.
+$\boldsymbol{\chi}(-2\omega;\omega,\omega)$ 
+would have a small influence from the local fields.
+Also, the excitonic effects ought to be explored, but
+their efficient calculation is theoretically and 
+numerically challenging\cite{beyond}   and beyond the scope 
+of this article.
 Unfortunately the experimental measurement of the $xxx$
@@
 formalism and within the independent particle approximation (IPA).
-It includes on equal footing and for the first time in the literature: 
-i) the scissors correction, ii) the cut function, and iii) the
-contribution of the non-local part of the pseudopotentials.
+It includes on equal footing:
+(i) the scissors correction, 
+(ii) the  
+contribution of the non-local part of the pseudopotentials, and  
+(iii) the cut function.
 We have used a Si(001)$2\times 1$ surface to confirm that our scheme
@@
 $1\omega$ and $2\omega$ resonances,  
-and has a strong influence in the line-shape.
+and has a strong influence in the line-shape,
+as for the case of bulk semiconductors.\cite{luppiJCP10,luppiPRB10,leitsmannPRB05}
  The cut function allows us  
@@
 by 15-20\%. 
-Therefore,  
-the results show the importance of these three contributions for a correct calculation 
-of SSHG.
 The $xxx$ component of  
@@
 our scheme opens the unprecedented possibility to study 
-surface SHG with more versatility and provides 
-more accurate results than previous approaches.
+surface SHG with more versatility
+and 
+more accurate results.
 
@@
 intermediate steps are given for their derivation. 
+\subsection{ Expressions for 
+\texorpdfstring{$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$}{que} 
+}\label{fullchis}
+Omitting the frequency dependence for convenience of notation,
+$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}
+=
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}
++
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}
++
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}
++
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}
+$, where 
+\begin{align*}
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}
+&= \frac{|e|^3}{2\hbar^2}
+\int \frac{dk^3}{8\pi^3}
+\sum_{l\ne mn}\frac{f_{mn}}{\omega^\Sigma_{nm}}
+\left[
+  \frac{\mathrm{Im}\left[\mathcal{V}^{\Sigma,\text{a}}_{ml}
+        \{r^{\text{c}}_{ln}
+        r^{\text{b}}_{nm}\}\right]}
+        {2\omega^\Sigma_{nm}-\omega^\Sigma_{lm}}
+ -\frac{\mathrm{Im}\left[\mathcal{V}^{\Sigma,\text{a}}_{ln}
+        \{r^{\text{b}}_{nm}
+        r^{\text{c}}_{ml}\}\right]}
+        {2\omega^\Sigma_{nm}-\omega^\Sigma_{nl}}
+\right]\frac{1}{\omega^\Sigma_{nm}-\tilde\omega}
+,
+\nonumber\\
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}
+ &=   \frac{|e|^3}{2\hbar^2}\int \frac{dk^3}{8\pi^3}
+\sum_{mn}
+    \frac{f_{mn}}{(\omega^{\Sigma}_{nm})^{2}}
+\Biggl[
+    \mathrm{Re}\left[r^{\text{b}}_{nm}
+    \left(\mathcal{V}^{\Sigma,\text{a}}_{mn}\right)_{;k^{\text{c}}}\right]
++   \frac{\mathrm{Re}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
+    r^{\text{b}}_{nm}\right]\Delta^{\text{c}}_{nm}}{\omega^{\Sigma}_{nm}} 
+\Biggr]\frac{1}{\omega^{\Sigma}_{nm}-\tilde\omega}
+,
+\nonumber\\
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}
+&= \frac{|e|^3}{2\hbar^2}\int \frac{dk^3}{8\pi^3}
+\sum_{l\ne mn}
+4\left[
+  \frac{f_{ln}\,\mathrm{Im}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
+        \{r^{\text{b}}_{nl}
+        r^{\text{c}}_{lm}\}\right]}
+        {2\omega^\Sigma_{nl}-\omega^\Sigma_{nm}}
+ -\frac{f_{ml}\,\mathrm{Im}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
+        \{r^{\text{c}}_{nl}
+        r^{\text{b}}_{lm}\}\right]}
+        {2\omega^\Sigma_{lm}-\omega^\Sigma_{nm}}
+\right]\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}
+,
+\nonumber\\
+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}
+&=  \frac{|e|^3}{2\hbar^2}\int \frac{dk^3}{8\pi^3}
+\sum_{mn}
+    \frac{4f_{mn}}{(\omega^{\Sigma}_{nm})^{2}}
+    \left[\mathrm{Re}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
+    \left(r^{\text{b}}_{nm}\right)_{;k^{\text{c}}}\right] 
+-   \frac{2\,\mathrm{Re}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
+    r^{\text{b}}_{nm}\right]\Delta^{\text{c}}_{nm}}{\omega^{\Sigma}_{nm}}\right]
+    \frac{1}{\omega^{\Sigma}_{nm}-2\tilde\omega}
+. 
+\end{align*}
+
 %%%%
@@
 where $\{\mathbf{G}\}$ are the reciprocal basis vectors satisfying
-$e^{\mathbf{R}\cdot\mathbf{G}}=1$, with $\{\mathbf{R}\}$ the translation vectors in real
+$e^{i\mathbf{R}\cdot\mathbf{G}}=1$, with $\{\mathbf{R}\}$ the translation vectors in real
 space, and $A_{n\mathbf{k}}(\mathbf{G})$ the expansion coefficients. Using
@@
 \frac{\hbar}{m_e}\delta_{\mathrm{a}\mathrm{b}} 
-C_{nm} 
+{\cal C}_{nm} 
 -\hbar 
@@
 L^{\mathrm{a}\mathrm{b}}_{nq} 
-C_{qm} 
+{\cal C}_{qm} 
 ,
@@
 The matrix elements $L^{\mathrm{a}\mathrm{b}}_{nm}$
-are calculated in Appendix \ref{vesnl}.
-% is small
-%compared to the other terms, thus we neglect it throwout this work.\cite{valerie} 
+are small as 
+compared to the other terms, thus we neglect it throwout this work.\cite{valerie} 
 Notice that
@@
 taking 
-${\mathbf{\cal C}}(z)=1$ or $C_{nm}=\delta_{nm}$.
+${\mathbf{\cal C}}(z)=1$ or ${\cal C}_{nm}=\delta_{nm}$.
 
@@
 
-%%%%
-\subsection{
-\texorpdfstring{$\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}$}{Vnonlocal},
-\texorpdfstring{$\mathbf{v}^\mathrm{nl}_{nm}$}{Vnonlocal}, and 
-\texorpdfstring{$T^{\mathrm{a}\mathrm{b}}_{nm}$}{[r,vnl]}}
-\label{vesnl}
-
-We take Eq.~\eqref{vnl} into Eq.~\eqref{vcali}
-and use
-$\sum_{\mathbf{G}} | \mathbf{k}+\mathbf{G}\rangle \langle\mathbf{k}+\mathbf{G} | =1$, 
-with
-$\langle\mathbf{r}|\mathbf{k}+\mathbf{G}=(1/\sqrt{\Omega})
-\mathrm{exp}(i(\mathbf{k}+\mathbf{G})\cdot\mathbf{r})$
-to obtain,
-\begin{align*}
-\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}(\mathbf{k})&=\frac{1}{2}
-\sum_{\mathbf{G}}
-\Big(
-\langle n\mathbf{k} | {\mathbf{\cal C}}(z) 
-| \mathbf{k}+\mathbf{G}\rangle\langle\mathbf{k}+\mathbf{G} |
-\mathbf{v}^\mathrm{nl} | m\mathbf{k}\rangle
-+
-\langle n\mathbf{k} |
-\mathbf{v}^\mathrm{nl}  
-| \mathbf{k}+\mathbf{G}\rangle\langle\mathbf{k}+\mathbf{G} |
-{\mathbf{\cal C}}(z) | m\mathbf{k}\rangle
-\Big)
-\nonumber\\
-&=\frac{1}{2 \hbar}
-\sum_{\mathbf{G}}
-\left(
-{\mathbf{\cal F}}^{\ell*}_{n\mathbf{k}}(\mathbf{G}) 
-H_{m\mathbf{k}}(\mathbf{G}) 
-+
-H^*_{n\mathbf{k}}(\mathbf{G}) 
-{\mathbf{\cal F}}_{m\mathbf{k}}(\mathbf{G}) 
-\right) 
-,
-\end{align*}  
-where 
-we have defined  
-\begin{align*}
-{\mathbf{\cal F}}_{n\mathbf{k}}(\mathbf{G}) 
-&=
-\sum_{\mathbf{G}'} 
-A_{n\mathbf{k}}(\mathbf{G}') 
-\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel}f(G'_\perp-G_\perp) 
-,
-H_{n\mathbf{k}}(\mathbf{G})\\
-&=
-\sum_{\mathbf{G}'} 
-A_{n\mathbf{k}}(\mathbf{G}') 
-(\nabla_{\mathbf{K}}+\nabla_{\mathbf{K}'})  
-V^\mathrm{nl}(\mathbf{K},\mathbf{K}')
-,
-\end{align*}
-and $\mathbf{K}=\mathbf{k}+\mathbf{G}$, $\mathbf{K}'=\mathbf{k}+\mathbf{G}'$.
-For fully  separable pseudopotentials in the   
-Kleinman-Bylander (KB) form,\cite{mottaCMS10,kleinmanPRL82,adolphPRB96}  
-the  
-matrix elements 
- $\langle\mathbf{K} |
-V^\mathrm{nl}  
-| \mathbf{K}'\rangle
-=V^\mathrm{nl}(\mathbf{K},\mathbf{K}')  
-$ 
-and their $\mathbf{K}$ and $\mathbf{K}'$ gradient 
-can be readily calculated.\cite{mottaCMS10,adolphPRB96,gordienkoRPJ04,fuchsCPC99} 
-We have 
-implemented 
-the calculation of $\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}_{nm}(\mathbf{k})$ with the help of 
-the DP\texttrademark~code.\cite{francesco}
-We obtain $\mathbf{v}^{\mathrm{nl}}_{nm}(\mathbf{k})$ from the previous three 
-expressions by taking ${\mathbf{\cal C}}(z)=1$, which implies 
-that $f(G_\perp)=\delta_{G_\perp 0}$, 
-and 
-${\mathbf{\cal F}}_{n\mathbf{k}}(\mathbf{G}) \to F_{n\mathbf{k}}(\mathbf{G})=A_{n\mathbf{k}}(\mathbf{G})$.
-
-From Eqs.~\eqref{lab} and \eqref{vnl} 
-\begin{equation*}
-L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
-=
-\frac{1}{\hbar^2}
-\langle n\mathbf{k} |
-\big[ r^{\mathrm{a}},[ V^\mathrm{nl}, r^\mathrm{b}]\big]
-| m\mathbf{k}\rangle
-,
-\end{equation*} 
-and we expand the triple commutator to obtain,
-\begin{align*}
-L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k}) 
-&=
-\frac{1}{\hbar^2\Omega}
-\sum_{\mathbf{G},\mathbf{G}'} 
-A^*_{n\mathbf{k}}(\mathbf{K}) 
-A_{m\mathbf{k}}(\mathbf{K}')
-\int
-d\mathbf{r} d\mathbf{r}'
- e^{-i\mathbf{K}\cdot\mathbf{r}}
-\Big(
-r^{\mathrm{a}}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-r'^\mathrm{b}
--
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-r'^\mathrm{a}
-r'^{\mathrm{b}}
-\nonumber\\
-&-
-r^\mathrm{b}
-r^{\mathrm{a}}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-+
- r^\mathrm{b}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-r'^{\mathrm{a}}
-\Big) 
- e^{i\mathbf{K}'\cdot\mathbf{r}'}
-,
-\end{align*} 
-where 
-$V^\mathrm{nl}(\mathbf{r},\mathbf{r}') = \langle\mathbf{r} | V^\mathrm{nl} | \mathbf{r}'\rangle$.
-We use the following identity
-\begin{align*}
-&
-\Big(
-\frac{\partial^2}{\partial K^\mathrm{a}\partial K'^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K'^\mathrm{a}\partial K'^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K^\mathrm{a}\partial K^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K^\mathrm{b}\partial K'^\mathrm{a}}
-\Big)
-\int 
-d\mathbf{r} d\mathbf{r}' 
- e^{-i\mathbf{K}\cdot\mathbf{r}}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-e^{i\mathbf{K}'\cdot\mathbf{r}'}
-\nonumber\\
-&=
-\int d\mathbf{r} d\mathbf{r}'
- e^{-i\mathbf{K}\cdot\mathbf{r}}
-\Big( 
-r^{\mathrm{a}} 
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-r'^\mathrm{b}
-- 
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-r'^\mathrm{a} 
-r'^{\mathrm{b}}
-- 
-r^\mathrm{b} 
-r^{\mathrm{a}} 
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}')
-+
- r^\mathrm{b} 
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-r'^{\mathrm{a}}
-\Big)  
-e^{i\mathbf{K}'\cdot\mathbf{r}'}
-,
-\end{align*}
-to write
-\begin{equation*}
-L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k})
-=
-\frac{1}{\hbar^2}
-\sum_{\mathbf{G},\mathbf{G}'} 
-A^*_{n\mathbf{k}}(\mathbf{K}) 
-A_{m\mathbf{k}}(\mathbf{K}')
-\Big(
-\frac{\partial^2}{\partial K^\mathrm{a}\partial K'^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K'^\mathrm{a}\partial K'^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K^\mathrm{a}\partial K^\mathrm{b}}
-+
-\frac{\partial^2}{\partial K^\mathrm{b}\partial K'^\mathrm{a}}
-\Big)
-\langle \mathbf{K} |
-V^\mathrm{nl} 
-| \mathbf{K}'\rangle,
-\end{equation*} 
-where
-\begin{equation*}
-\langle\mathbf{K} |
-V^\mathrm{nl} 
-| \mathbf{K}'\rangle
-=
-\frac{1}{\Omega}
-\int 
-d\mathbf{r} d\mathbf{r}' 
- e^{-i\mathbf{K}\cdot\mathbf{r}}
-V^\mathrm{nl}(\mathbf{r},\mathbf{r}') 
-e^{i\mathbf{K}'\cdot\mathbf{r}'}
-,
-\end{equation*}
-The double derivatives with respect to $\mathbf{K}$ and $\mathbf{K}'$ 
-can be worked out explicitly and 
-$L^{\mathrm{a}\mathrm{b}}_{nm}(\mathbf{k})$
-can be finally calculated.\cite{valerie}
-
-%%%%
 \subsection{Expressions for  \texorpdfstring{$\boldsymbol{\mathcal{\cal V}}^{S}_{nm}$}{Vnm}
@@
 
-\newpage
+%%%%%%%%%%%%%
 \bibliography{ref}
 \end{document}
+%%%%%%%%%%%%%
+
+Figures log:
+fig1 -> images/slab
+fig2 -> images/si2x1-crop
+fig3 -> plots/fig1 
+fig4 -> plots/fig2
+fig5 -> plots/fig3
+fig6 -> plots/fig4

### 2014-12-03 sent/submission_letter.tex

In this article we formulate a theoretical approach of surface
second-harmonic generation from semiconductor surfaces based on the
length gauge and the one electron density operator.  We calculate the
nonlinear second-order surface susceptibility tensor within the
independent particle approximation including for the first time in one
unique formulation (i) the scissors correction, needed to have the
correct value of the energy band gap, (ii) the contribution of the
nonlocal part of the pseudopotentials, routinely used in ab initio
band structure calculations, and (iii) the derivation for the
inclusion of the cut function, used to extract the surface response.
We test our scheme with a clean Si(001)2x1 reconstructed surface with
excellent results.  Therefore, the inclusion of the three
aforementioned contributions is very important and makes our scheme
unprecedented and opens the possibility to study surface
second-harmonic generation with more versatility and providing more
accurate results.  Most of the cited works appear in the Physical
Review Letters and Physical Review B, for which we think that this
article would have a good visibility in Physical Review B, as well.

### 2014-12-04 sent/manuscript.tex
sent to prb

--- 2014-12-03 manuscript.tex
@@
 \centering
-\includegraphics[scale=.7]{images/slab}
+\includegraphics[scale=.7]{fig1}
 \caption{(color on line) A sketch of the super-cell. 
@@
 \centering 
-\includegraphics[scale=.8]{images/si2x1-crop}
+\includegraphics[scale=.8]{fig2}
 \caption{(color on line) The slab shows a 
@@
 to obtain the relevant parameters. The most important of 
-these are: an energy cut-off of 10 Ha, an equal number of conduction and 
+these are: an energy cut-off of 10 Ha for the 16, 24, and 32 layered
+slabs
+and 13 Ha for the 40 layer slab,
+an equal number of conduction and 
 valence bands, and a set of 244 $\mathbf{k}$-points.
@@
 \centering 
-\includegraphics[scale=.8]{plots/fig1}
+\includegraphics[scale=.8]{fig3}
 \caption{(color on line) 
@@
 \centering 
-\includegraphics[scale=.8]{plots/fig2}
+\includegraphics[scale=.8]{fig4}
 \caption{(color on line) 
@@
 \centering 
-\includegraphics[scale=.8]{plots/fig3}
+\includegraphics[scale=.8]{fig5}
 \caption{(color on line) 
@@
 \centering 
-\includegraphics[scale=.8]{plots/fig4}
+\includegraphics[scale=.8]{fig6}
 \caption{(color on line) 

### 2014-12-04 sent/submission_letter.tex
sent to prb

@@
 nonlinear second-order surface susceptibility tensor within the
-independent particle approximation including for the first time in one
-unique formulation (i) the scissors correction, needed to have the
+independent particle approximation. We include for the first time in one
+unique formulation: (i) the scissors correction, needed to have the
 correct value of the energy band gap, (ii) the contribution of the
@@
 unprecedented and opens the possibility to study surface
-second-harmonic generation with more versatility and providing more
+second-harmonic generation with more versatility while providing more
 accurate results.  Most of the cited works appear in the Physical
 Review Letters and Physical Review B, for which we think that this
-article would have a good visibility in Physical Review B, as well.
+article would have good visibility in Physical Review B as well.

### 2014-12-22 corrections/corrections.tex
preparing corrections by valerie and nicolas

--- 2014-12-03 manuscript.tex
@@
-\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
-%\usepackage{showlabels}
-
-\pacs{78.68.+m, 42.65.An, 71.15.Mb, 42.65.Ky, 78.66.-w}
-
-\begin{document}
-
-\title{Theory of surface second-harmonic generation for 
-       semiconductors including effects of nonlocal operators}
-
-\author{Sean M. Anderson}
-    \affiliation{Centro de Investigaciones en Optica, 
-                Le\'on, Guanajuato, M\'exico}
-\author{Nicolas Tancogne-Dejean}
-    \affiliation{Laboratoire des Solides Irradi\'es, \'Ecole Polytechnique, 
-                CNRS, CEA/DSM, 91128 Palaiseau, France}
-    \affiliation{European Theoretical Spectroscopy Facility (ETSF), 
-                Palaiseau, France}
-\author{Bernardo S. Mendoza}\email{bms@cio.mx}
-    \affiliation{Centro de Investigaciones en Optica, 
-                Le\'on, Guanajuato, M\'exico}
-\author{Val\'erie V\'eniard}
-    \affiliation{Laboratoire des Solides Irradi\'es, \'Ecole Polytechnique, 
-                CNRS, CEA/DSM, 91128 Palaiseau, France}
-    \affiliation{European Theoretical Spectroscopy Facility (ETSF), 
-                Palaiseau, France}
-
-\begin{abstract}
-We formulate a theoretical approach of surface second-harmonic generation 
-from semiconductor surfaces based on the length gauge and the 
-electron  
-density operator.  
-Within the independent particle approximation 
-the nonlinear second-order surface susceptibility tensor
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$    
-is calculated, including for the first time in one unique formulation
-(i) the scissors correction, needed to have the correct value of the
-energy band gap,
-(ii) the contribution of the nonlocal part of the  
-pseudopotentials, routinely  used in \textit{ab initio} band structure calculations, 
-and (iii) the derivation for the inclusion of the cut function, 
-used to extract the surface response.
-The first two contributions
-are described by spatially nonlocal quantum
-mechanical operators and are fully taken into account in the present formulation. 
-As a test-case of the approach we calculate 
-$\chi^{xxx}(-2\omega;\omega,\omega)$ for the clean Si(001)$2\times 1$
-reconstructed surface and find that our scheme to extract the surface
-through the cut function
-readily works and is exact. The effects of the scissors correction and
-of the nonlocal part of the pseudopotentials are discussed 
-in surface nonlinear optics. 
-The scissors correction shifts the spectrum to 
-higher energies though the shifting is not rigid and mixes the 
-$1\omega$ and $2\omega$ resonances, 
-and has a strong influence in the line-shape. 
-The 
-effects of the nonlocal part of the pseudopotentials 
-keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$, but 
-reduces its value
-by 15-20\%.  
-Therefore, the inclusion of the three aforementioned 
-contributions is very important and 
-makes our scheme unprecedented and opens the possibility to 
-study surface second-harmonic generation
-with more versatility and providing more accurate results.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}\label{intro}
-
-In recent years surface nonlinear optical spectroscopies, particulary surface 
-second-harmonic generation (SSHG), have evolved as useful nondestructive 
-and noninvasive tools to study surface and interface properties. These 
-properties include atomic structure, phase transitions, adsorption of 
-atoms, and many others.\cite{daumPRL93, mcgilpOE94, meyerPRL95, powerPRL95, 
-godefroyAPL96, hoferAPA96, dadapPRB97, bloembergenAPB99, mcgilpSRL99, 
-suzukiAPB99, mitchellSS01, hughesPRB96, guyotPRB88, downerPSSA01, shenAPB99, 
-shenNAT89, chenPRL81, mendozaPRL98, downerSIA01} Nowadays, SSHG spectroscopy 
-is a crucial tool for research and development in microelectronics,
-\cite{zheltikovLP00} semiconductors, \cite{lupkeSSR99} nanomaterials,
-\cite{salazar-aparicioPRB14} and many more recent areas of great scientific 
-and commercial interest.\cite{cazzanelliNM14} The high surface sensitivity 
-of SSHG spectroscopy is due to the fact that within the dipole approximation 
-the bulk SHG signal of centrosymmetric materials is identically zero.
-Radiation can occur only at the surface where the inversion symmetry is 
-broken. 
-The bulk quadrupole contribution for centrosymmetric materials is
-different from zero, but usually it is very small,\cite{downerSIA01} and we neglect it. 
-Much of the foundation of surface science has been built from 
-experiments involving emission or scattering of electrons from surfaces. 
-These require ultrahigh vacuum (UHV) environments and provide no access to 
-buried interfaces. However, SSHG is compatibile with non-UHV conditions and 
-has access to interfaces buried beneath transparent overlayers. Even when 
-applied to surfaces in UHV, the light source and detectors can be aligned and
-used outside the vacuum chamber.
-
-The usefulness of SSHG could be limited by the lack of microscopic theoretical
-understanding of the nonlinear spectra. The macroscopic phenomenological 
-theory of SSHG, which relates the intensity, phase, and polarization of 
-detected fields to nonlinear and linear susceptibilities at the material 
-interface is now fully developed.\cite{downerSIA01}
- However, microscopic theory that relates 
-electronic-level structure to the nonlinear source polarization is still being 
-developed. \cite{butcherPOPS63, aspnesPRB72, sipePRB93,
-  levinePRB94,aversaPRB95, hughesPRB96, rashkeevPRB98,beyond}  
- Within the independent 
-particle approximation (IPA) some frameworks for bulk SHG have been developed 
-to study the nonlinear optical response of bulk materials. \cite{butcherPOPS63, aspnesPRB72, sipePRB93, levinePRB94,aversaPRB95, hughesPRB96, rashkeevPRB98} 
-In this article we put forward an approach to calculate the
-microscopic second-harmonic surface susceptibility that encompasses several
-theoretical features not taken into account before 
-in the case of a surface.
-
+% @119: Introduction
+% =====================================================
 The most used framework for \textit{ab initio} calculations, Density 
@@
 This allows us to ``open'' the DFT-LDA gap to 
-its correct experimental or GW value even when experiments are not available. 
+its correct experimental or GW value for most bulk semiconductors.
+%%%% Valerie and Nicolas remove these lines
+% even when experiments are not available. 
+%%%%
+This approximation has already been used in linear optical calculations for 
+surfaces, see \cite{kippPRL96}, thus improving the agreement with 
+experimental results. 
 In this context, to correct for the underestimation of the energy band gap of 
@@
 
-Concerning the optical response of surfaces and interfaces, Reining 
-\textit{et al.}\cite{reiningPRB94} introduced the concept of a cut function 
-in order to obtain the surface SH susceptibility tensor. This cut function 
-is required since one usually uses a slab approach when treating semi-infinite 
-surface systems.\cite{reiningPRB94} If the slab is centrosymmetric the susceptibility tensor will 
-be identically zero. The cut function is such that it separates the nonlinear 
-response for the two surfaces of the slab avoiding the destructive 
-interference between them giving a finite value that one identifies with the 
-SSH susceptibility tensor. If the slab is not centrosymmetric the cut function 
-can be used to separate the different signals coming from either surface of 
-the slab. Indeed, one of the main results of this article is to show that the 
-SSH  susceptibility tensor obtained by using the cut function is correctly 
-extracted from the slab. After Reining \textit{et al.},\cite{reiningPRB94}
-Refs.~\onlinecite{mendozaPRL98,arzatePRB01,mendozaPRB01,mejiaPRB02,sanoPRB02}
-followed 
-upon this work and in particular Ref.~\onlinecite{arzatePRB01} went into a 
-detailed analysis of the different contributions to the SHG spectra of a 
-surface and the nuanced relationship between bulk, surface, interband, 
-intraband, $1\omega$ and $2\omega$ terms, and
-Ref. \onlinecite{mejiaRMF04} developed a layer-by-layer 
-analysis for the nonlinear responses of semiconductor systems, within a 
-tight-binding framework. This model allows for obtaining results from 
-selected regions of a system including the surface. However, in these 
-references the scissors operator is either excluded or incorrectly 
-implemented.
+% =====================================================
 
-Finally, DFT-LDA calculations are often based on the use of pseudopotentials.
-As it will be discussed 
-in this article, the presence of a nonlocal part of the
-pseudopotential introduces
-corrections to
-the momentum operator of the electron that have to be included with care in 
-the SSH susceptibility. 
-For the bulk counterpart see for instance Refs.~\onlinecite{ismailPRL01,luppiPRB08}.
+% @1192: Half-slab vs. full-slab
+% =====================================================
 
-Therefore, within the IPA the most complete approach for the calculation of 
-the SSH susceptibility is one which includes (i) the scissors correction, 
-(ii) the contribution of the nonlocal part of the 
-pseudopotential, and
-(iii) the cut function.
- Therefore, the goal of this paper is to derive a new 
-expression within the length gauge for the SSH susceptibility tensor 
+In Fig.~\ref{fig2}
+we compare 
+$\chi^{xxx}_{\mathrm{half-slab}}$  
+vs. 
+$\chi^{xxx}_{\mathrm{full-slab}}$ 
+for the four different possibilities 
+between including or not including the
+effects of $\mathbf{v}^\mathrm{nl}$ or the scissors correction
+$\hbar\Delta$.   
+For these results we chose
+$\hbar\Delta=0.5$ eV, that is the GW gap reported in
+Refs.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
+This is justified by the fact that the surface states of the clean 
+Si(001) surface investigated are found, from GW calculations 
+\cite{rohlfingPRB95}, to shift rigidly and to keep their dispersion with 
+respect to LDA.
+We see that for all four instances the 
+difference between responses is quite small.
+Indeed, when the value $|\chi^{xxx}|$ 
+is large the difference between the two is very small; 
+when the value is small the difference increases only slightly, 
+but the spectra is so close to zero that it is negligible. 
+These differences would decrease as the number of atomic layers 
+increases. We remark that 32 layers in the slab is more than enough 
+to confirm that the extraction of the surface second-harmonic 
+susceptibility from the $2\times 1$ surface is readily possible 
+using the formalism contained in Eq.~\eqref{chis}.
+We have confirmed that for the dihydride surface
+$|\chi^{xxx}_{\mathrm{half-slab}}|\approx 0$ (not shown).
+This confirms the validity of our theory and is the main result of
+this article; through the proposed layer formalism we can calculate the surface SH
 $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$    
-that includes the aforementioned contributions. The inclusion of these three 
-contributions makes our scheme unprecedented and opens the possibility to 
-study surface SHG with more versatility and providing accurate results.
+including
+the
+contribution of the nonlocal part of the pseudopotentials 
+and the part of the many-body effects through the scissors correction.
+Our scheme should work for any slab.
 
-The paper is organized as follows. In Sec. \ref{theory} we present the 
-relevant steps for deriving the surface second-order susceptibility tensor
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
-length gauge formalism. This derivation includes the addition of the terms 
-mentioned above that have been absent in previous works. Also, our 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ can be used 
-for a layer-by-layer analysis if desired. 
-Nevertheless, such an analysis could depend on the choice of the basis
-used to expand the wavefunctions.
-In Sec. \ref{results} we show the 
-results of this reformulation with a study of a clean Si(001) surface with a 
-$2 \times 1$ reconstruction, 
-proving the correctness of our approach with a special test-case,
-never exploited before.
-We compare results from before and after adding 
-the different nonlocal contributions. Finally, in Sec. \ref{conc} we give 
-our conclusions.
+% =====================================================
 
-\section{Theory}
+% @409: Length Gauge Formalism
+% =====================================================
 
-\label{theory}
-
-In this section we present the scheme used to calculate the surface 
-second-order nonlinear response using the length gauge formalism and
-the electron 
-density operator. 
-Some of the results presented  in Subsections \ref{pa} and \ref{lgf}
-have already been discussed in earlier
-studies and can be applied to bulk and surface studies.\cite{sipePRB00,cabellosPRB09}
-We present them in order to have a self-contained derivation.
-The terms presented in these Subsections are used in Subsection \ref{cd}, 
-where we derive the expression 
-for the surface $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$.
-
-\subsection{Perturbation approach}\label{pa}
-
-We assume the IPA, a classical electromagnetic field, and quantum mechanical 
-matter. We neglect local field and excitonic effects. We can describe the 
-system using 
-the one electron density operator ${\rho}$, with which we can calculate the 
-expectation value of a single-particle observable $\mathcal{O}$ as 
-$\langle{\cal O}\rangle = \mbox{Tr}({\rho}{\cal O}),\label{calo}$
-with $\mathcal{O}$ the associated quantum mechanical operator and Tr the 
-trace. The density operator satisfies
-$i\hbar(d{\rho}/dt) = [H(t), {\rho}], \label{rho}$
-with $H(t)$ as the total single electron Hamiltonian, written as 
-\begin{equation*}
-H(t) = H_{0} + H_{I}(t),  
-\label{ache}
-\end{equation*}
-where $H_{0}$ is the unperturbed time-independent Hamiltonian, and $H_{I}(t)$
-is the time-dependent potential energy due to the interaction of the electron 
-with the electromagnetic field.
-To proceed with the solution of $\rho$ it is convenient to use the interaction 
-picture, where a unitary operator $U = \exp({iH_{0}t/\hbar})$ transforms any 
-operator $\mathcal{O}$ into $\tilde{\mathcal{O}} = U{\cal O}U^{\dagger}$. 
-Even if $\mathcal{O}$ is time-independent, $\tilde{\mathcal{O}}$ is 
-time-dependent through the explicit time dependence of $U$. The dynamical 
-equation for $\tilde{\rho}$ is given by
-\begin{equation*}
-i\hbar \frac{d\tilde{\rho}}{dt} = [\tilde{H}_{I}(t), \tilde{\rho}],
-\label{rho1}
-\end{equation*}
-with solution 
-\begin{equation}
-i\hbar \tilde{\rho}(t) = 
-    i\hbar \tilde{\rho}_{0} + 
-    \int_{-\infty}^{t}dt^{\prime}
-    [\tilde{H}_{I}(t^{\prime}),\tilde{\rho}(t^{\prime})],  
-\label{trans}
-\end{equation}
-where $\tilde{\rho}_{0} = \tilde{\rho}(t = -\infty)$ is the unperturbed 
-density matrix. We look for the standard perturbation series solution, 
-$\tilde{\rho}(t) = \tilde{\rho}^{(0)} + \tilde{\rho}^{(1)} + 
-\tilde{\rho}^{(2)} + \cdots$, where the superscript denotes the order (power) 
-with which each term depends on the perturbation $H_{I}(t)$. From 
-Eq.~\eqref{trans} the $N$-th order term is 
-\begin{equation}
-\tilde{\rho}^{(N)}(t) = 
-    \frac{1}{i\hbar}\int_{-\infty }^{t}dt^{\prime}
-    [\tilde{H}_{I}(t^{\prime }), \tilde{\rho}^{(N - 1)}(t^{\prime})].
-\label{rhop}
-\end{equation}
-The series is generated by the unperturbed density operator $\tilde{\rho}^{(0)}
-\equiv \tilde{\rho}_{0}$, assumed to be the diagonal Fermi-Dirac distribution, 
-$\langle n\mathbf{k}|\tilde{\rho}_{0}|n\mathbf{k}\rangle = 
-f(\hbar\omega_{n}(\mathbf{k}))\equiv f_{n}$. For a clean, cold semiconductor 
-$f_{n}=1$ when $n$ is a valence ($v$) or occupied band, and zero when $n$ is 
-a conduction ($c$) or empty band. 
-We assume this for the remainder of the  
-article.  
-As we neglect spin-orbit coupling, the
-final expression for
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$   
-has to be multiplied by a factor of 2
-to account for spin-degeneracy.
-The expectation values must satisfy
-$\langle{\cal O}\rangle = \mbox{Tr}({\rho}{\cal O}) = 
-\mbox{Tr}(\tilde{\rho}\tilde{\cal O})$.
-
-We will look for the expectation value of the microscopic current density, 
-$\mathbf{J}$, given by 
-\begin{equation*}
-\mathbf{J} = \langle{\mathbf{J}}\rangle 
-           = \frac{e}{A}\mbox{Tr}({\rho}\dot{\mathbf{r}}),
-\end{equation*}
-where $\dot{\mathbf{r}}$ is the time derivative of the position operator of the
-electron with charge $e$, defined as
-\begin{equation}
-\mathbf{v}\equiv \dot{\mathbf{r}}=\frac{1}{i\hbar }[\mathbf{r},H_0],  
-\label{mv}
-\end{equation}
-with $\mathbf{v}$ the velocity operator of the electron, and $A$ the
-area of the unit cell. We calculate the polarization density 
-$\mathbf{P}$, related to $\mathbf{J}$ by
-$\mathbf{J}=d\mathbf{P}/dt$. For a 
-perturbing electromagnetic field, $\mathbf{E}(t)=
-\mathbf{E}(\omega )e^{-i\tilde{\omega} t}+c.c.$,
-where $\tilde\omega=\omega+i\eta $,
-and $\eta >0$ adiabatically switches on the interaction,
-we write the second-order nonlinear
-polarization as, 
-\begin{equation}
-{\mathbf{\cal P}}^{\mathrm{a}}(2\omega)=
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)
-E^{\mathrm{b}}(\omega)E^{\mathrm{c}}(\omega),  
-\label{pshg}
-\end{equation}
-where $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega ;\omega ,\omega )$ is the nonlinear
-susceptibility responsible for surface second-harmonic generation
-(SSHG). 
-The 
-superscripts in Eq.~\eqref{pshg} denote Cartesian components, and if
-repeated are to be summed over. Without loss of generality we will
-define $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$
- to satisfy intrinsic permutation
-symmetry, 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega ;\omega ,\omega )
-=\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega ;\omega,\omega )$.
-
-The unperturbed Hamiltonian 
-is used to solve the Kohn-Sham equations\cite{kohnPR65} of Density  
-Functional Theory (DFT). It is convenient to work within the Local 
-Density Approximation (LDA), so we label the hamiltonian with the corresponding  
-LDA superscript. Any other approximation can be used (like GGA) and our 
-derivation remains the same. Then,
-\begin{equation*}
-H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})
-=\frac{p^{2}}{2m_e}+V(\mathbf{r},\mathbf{p})
-\end{equation*}
-with $m_e$ the mass of the electron, $\mathbf{p}$ its canonical momentum, and 
-$V$
- the periodic crystal potential, where we neglect spin-orbit terms.
-To be more general in our derivation of
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$, we asume
-contribution as is customary for most
-pseudopotentials, and then we replace $V$ with
-\begin{align*}
-V^{\mathrm{ps}}(\mathbf{r},\mathbf{p})=V(\mathbf{r})+V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})
-,
-\end{align*}
-where 
-$V(\mathbf{r})$ and 
-$V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})$  
-are the local and nonlocal parts, 
- respectively.
-The argument $(\mathbf{r},\mathbf{p})$
-  is equivalente to the
-explicit $(\mathbf{r},\mathbf{r}')$ nonlocal notation.\cite{ismailPRL01}
-In case of
-a local potential, i.e. $V= V(\mathbf{r})$, 
-like that of all-electron schemes, 
-we simply omit the contribution of $V^{\mathrm{nl}}(\mathbf{r},\mathbf{p})$
-from the results that we have derived.
-
-It is well known that the use of the LDA
-leads to an underestimation of the band gap. A standard
-procedure to correct for this is to
-use the ``scissors approximation'', where the 
-conduction bands are rigidly shifted in energy so that the band gap 
-corresponds to the accepted 
-experimental electronic band gap.\cite{levinePRL89,levinePRL91,delsolePRB93} 
-This is often in fairly good agreement with the GW
-band gap based on a more sophisticated calculation.\cite{hybertsenPRB86}
-The LDA wave functions are used since they produce band
-structures with dispersion relations similar to those predicted by the GW.
-Mathematically, the scissors (non-local) operator 
-$S$ is added to the unperturbed or unscissored Hamiltonian $H^{\mathrm{LDA}}_{0}$ ,
-\begin{equation*}
-H^\Sigma_{0}(\mathbf{r},\mathbf{p})=H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})+S(\mathbf{r},\mathbf{p})
-\end{equation*}
-where 
-\begin{equation}
-S(\mathbf{r},\mathbf{p})=\hbar \Delta\sum_{n}\int d^{3}k(1-f_{n})
-|n\mathbf{k}\rangle\langle n\mathbf{k}|,
-\label{hats}
-\end{equation}
-with $\hbar \Delta$  the rigid ($\mathbf{k}$-independent) energy correction to be
-applied. 
-The unscissored and scissored Hamiltonians satisfy 
-\begin{align*}
-H^{\mathrm{LDA}}_{0}(\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=
-\hbar \omega^{\mathrm{LDA}}_{n}(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),\label{hamils} \\
-H_{0}^\Sigma (\mathbf{r},\mathbf{p})\psi _{n\mathbf{k}}(\mathbf{r}) &=
-\hbar \omega_{n}^\Sigma(\mathbf{k})\psi _{n\mathbf{k}}(\mathbf{r}),
-\end{align*}
-where the scissor-shifted energies, 
-$\omega_{n}^\Sigma(\mathbf{k})$ , are given by
-\begin{equation*}
-\omega_{n}^\Sigma(\mathbf{k})=\omega^\mathrm{LDA}_{n}(\mathbf{k})+(1-f_{n})\Delta.
-\end{equation*}
-We emphasize that the scissored and unscissored Hamiltonian 
-have the same eigenfunctions, where
-$\psi_{n\mathbf{k}}(\mathbf{r})=\langle\mathbf{r}|n\mathbf{k}\rangle=e^{i\mathbf{k}\cdot\mathbf{r}}u_{n\mathbf{k}}(\mathbf{r})$,
-are the real space representations of the Bloch states $| n\mathbf{k}\rangle$ labelled 
-by the band index $n$ and the crystal momentum $\mathbf{k}$, and $u_{n\mathbf{k}}(\mathbf{r})$
-are cell periodic.
-
-\subsection{Length Gauge Formalism}\label{lgf}
-
-To calculate the optical response 
-within the Length Gauge formalism, we take the 
-interaction Hamiltonian as
+According to ~\onlinecite{ismailPRL01}, we first start with the interaction 
+Hamiltonian expressed in the velocity gauge, containing the nonlocal parts
+$V^{nl}$ and $S$. Within the dipole approximation and using a gauge 
+transformation, it can be transformed into an effective Hamiltonian 
+\cite{Veniard, unpublished}
 \begin{equation}
@@
 \end{equation}
-The 
-treatment of the position operator
-$\mathbf{r}$ for extended Bloch states, $\langle\mathbf{r}|n\mathbf{k}\rangle$ has 
-been described in detail.\cite{adamsJCP53,blountSSP62} 
-In particular, Ref.~\onlinecite{aversaPRB95} develops
-the treatment of $\mathbf{r}$ for
-the calculation of the nonlinear optical response.
+The treatment of the position operator $\mathbf{r}$ for extended Bloch states 
+is problematic and has been discussed in Refs.\cite{adamsJCP53,blountSSP62}
+%%%% Valerie and Nicolas remove these lines
+% In particular, Ref.~\onlinecite{aversaPRB95} develops the treatment of 
+% $\mathbf{r}$ for the calculation of the nonlinear optical response.
+%%%%
+Following Ref.~\onlinecite{sipePRB00} we take the
+matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of 
+Eq.~\eqref{rde}, we obtain 
+$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\Sigma_{nm}-\tilde\omega)t}$,
+with 
+\begin{equation}
+B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\Sigma_{nm}-\tilde\omega},
+\label{j.1}
+\end{equation}
+and 
+\begin{equation}
+(\tilde{\rho}^{(2)}(t))_{nm} = \frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
+i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
+\Big)  
+-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\Sigma_{nm}-2\tilde\omega)t}
+.
+\label{j.2}
+\end{equation}
 $\mathbf{r}$ is split into the {\it intraband} 
@@
 with $\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ the canonical momentum operator, 
-and using $[r^a,p^b]=i\hbar\delta_{ab}$, where $\delta_{ab}$ is the Kronecker delta.
+and using $[r^a,p^b]=i\hbar\delta_{ab}$, where $\delta_{ab}$ is the 
+Kronecker delta.
 Using Eq.~\eqref{hats} we obtain 
@@
 where we see that $\mathbf{v}^S_{nn}=0$. From Eqs.~\eqref{pmnrmn} and
-\eqref{vop2}
- it follows that
+\eqref{vop2} it follows that
 \begin{equation}
@@
 
-Following Ref.~\onlinecite{sipePRB00} we take the
-matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of 
-Eq.~\eqref{rde}, we obtain 
-$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\Sigma_{nm}-\tilde\omega)t}$,
-with 
-\begin{equation}
-B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\Sigma_{nm}-\tilde\omega},
-\label{j.1}
-\end{equation}
-and 
-\begin{equation}
-(\tilde{\rho}^{(2)}(t))_{nm} = \frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
-i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
-\Big)  
--(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\Sigma_{nm}-2\tilde\omega)t}
-.
-\label{j.2}
-\end{equation}
 We have used the fact that for a cold semiconductor $\partial
@@
 
-\subsection{Layered Current Density}\label{cd}
+% =====================================================
 
-The approach we use to study the surface of a semi-infinite
-semiconductor crystal is as follows. Instead of using a
-semi-infinite system, we replace it by a super-cell that 
-consists of a finite
-slab of atomic layers and a vacuum region (see Fig.~\ref{fslab}). This
-super-cell
-is repeated to form a full three dimensional crystalline structure.
-The slab itself consists of front, back, and 
-sub-surface regions, and in between these
-a region that is equivalent to the
-bulk of the system. 
-In general the surface of a crystal reconstructs or relaxes as the atoms
-move to find equilibrium positions. This is due to the fact that
-the otherwise
-balanced forces are disrupted when the surface atoms do not find their 
-partner atoms that are now absent at the slab surface.
-To take the reconstruction or relaxation into account, 
-we take ``surface'' to mean
-the true surface of the first layer of atoms and
-some of the atomic sub-layers adjacent to it.
-Since the front and the back
-surfaces of the slab are usually identical the total slab is
-centrosymmetric. This would imply that 
-$\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}=0$ so we must
-find a scheme 
-in order to have a finite $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$ representative of the
-surface. Even if the front and back surfaces of the slab 
-are different, breaking the centrosymmetry and therefore giving an
-overall $\chi^{\mathrm{slab},\mathrm{a}\mathrm{b}\mathrm{c}}\ne 0$; we still
-need a procedure to extract the front surface $\chi^{f,\mathrm{a}\mathrm{b}\mathrm{c}}$
-and the back surface $\chi^{b,\mathrm{a}\mathrm{b}\mathrm{c}}$ from the slab
-susceptibility. 
-We have omitted the frequency dependence of $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$ for 
-convenience of notation.
-
-A convenient way to accomplish the separation of the SH signal of
-either surface is to introduce a ``cut function'', ${\mathbf{\cal C}}(z)$, which is 
-usually taken to be unity over one half of the slab and zero over 
-the other half.\cite{reiningPRB94}
-In this case ${\mathbf{\cal C}}(z)$ will give the contribution of the 
-side of the slab for which ${\mathbf{\cal C}}(z)=1$. 
-As was done for the linear response,\cite{mendozaPRB06}
-we can generalize this 
-simple choice for ${\mathbf{\cal C}}(z)$ by a top-hat cut function
-${\mathbf{\cal C}}^\ell(z)$ that selects a given layer,
-\begin{equation}
-{\mathbf{\cal C}}^\ell(z)=\Theta(z-z_\ell+\Delta_\ell^{b})  
-            \Theta(z_\ell-z+\Delta_\ell^f),
-\label{sz}
-\end{equation} 
-where $\Theta$ is the Heaviside function. Here, $\Delta_\ell^{f/b}$
-is the distance that the $\ell$-th layer extends towards the front
-$(f)$ or back $(b)$ from its $z_\ell$ position. 
-We take $z_\ell$ to be at the center of an atom that 
-belongs to layer $\ell$, so the previous
-equation would give the $\ell$-th atomic-layer 
-contribution to the nonlinear optical response.
-$\Delta_\ell^f+\Delta_\ell^b$ is the thickness of layer $\ell$ 
-(see Fig.~\ref{fslab}).
-\begin{figure}
-\centering
-\includegraphics[scale=.7]{images/slab}
-\caption{(color on line) A sketch of the super-cell. 
-The atomic slab corresponds to the
-circles representing the atoms of the system.
-\label{fslab}} 
-\end{figure}
-
-To introduce the
-cut function ${\mathbf{\cal C}}(z)$ in
-the calculation of $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}$, we start from 
-the operator for the electron current,
-$\mathbf{j}(\mathbf{r})=\frac{e}{2}\left(\mathbf{v}^\Sigma | \mathbf{r}\rangle\langle\mathbf{r} |
-+ | \mathbf{r}\rangle\langle\mathbf{r} | \mathbf{v}^\Sigma\right)$, that leads to
-\begin{equation}
-\mathbf{j}^{(N)}(\mathbf{r},t)=\mathrm{Tr}(\mathbf{j}(\mathbf{r})\rho^{(N)}(t))
-=
-\int \frac{dk^3}{8\pi^3}
-\sum_{nm}
-\rho^{(N)}_{nm}(\mathbf{k};t)\mathbf{j}_{mn}(\mathbf{k};\mathbf{r})
-,
-\label{jmic}
-\end{equation}
-where 
-\begin{equation}
-\mathbf{j}_{mn}(\mathbf{k};\mathbf{r})=
-\frac{e}{2}
-\left(
-\langle m\mathbf{k} | \mathbf{v}^\Sigma | \mathbf{r}\rangle \langle\mathbf{r}|n\mathbf{k}\rangle
-+
-\langle m\mathbf{k}|\mathbf{r}\rangle\langle\mathbf{r} | \mathbf{v}^\Sigma | n\mathbf{k}\rangle
-\right).
-\label{jmic3}
-\end{equation}
-Integrating the microscopic current $\mathbf{j}(\mathbf{r},t)$ over
-the entire slab gives the averaged microscopic current density, $\mathbf{J}(t)$. 
-If we want the contribution from only one region of the unit cell 
-towards the total current, we can integrate $\mathbf{j}({\mathbf r},t)$ 
-over the desired region. Then the contribution 
-to the current density from the
-chosen region of the slab is given by
-\begin{equation*}
-\frac{1}{A}\int d^3r\, {\mathbf{\cal C}}(z)\, 
-\mathbf{J}^{(N)} (\mathbf{r},t)
- \equiv \boldsymbol{\mathcal{\cal J}}^{(N)}(t),
-\end{equation*}
-where $\boldsymbol{\mathcal{\cal J}}^{(N)}(t)$ is the $N$-th order current induced in the
-region specified by ${\mathbf{\cal C}}(z)$.
-Therefore we define
-\begin{equation}
-e{\boldsymbol{\mathcal{V}}}^{\Sigma}_{mn}(\mathbf{k})
-\equiv
-\int d^3r\, {\mathbf{\cal C}}(z)\,\mathbf{j}_{mn}({\mathbf{k}};\mathbf{r}),
-\label{vcal}
-\end{equation}
-to write the Fourier transform of Eq.~\eqref{jmic} as
-\begin{equation}
-\boldsymbol{\mathcal{\cal J}}^{(N)}(2\omega)=\frac{e}{A}
-\int \frac{dk^3}{8\pi^3}
-\sum_{mn}
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{mn}(\mathbf{k}) 
-\rho^{(N)}_{nm}(\mathbf{k};2\omega) 
-, 
-\label{jmac2}
-\end{equation}
-that gives the induced microscopic current of the chosen region, to order $N$ 
-in the external perturbation. 
-From
-Eqs.~\eqref{vcal} and \eqref{jmic3} we obtain
-\begin{align*}
-{\boldsymbol{\mathcal{V}}}^{\Sigma}_{mn}({\mathbf k})
-&=
-\frac{1}{2}
-\int \mathrm{d}^3 r\,
- {\mathbf{\cal C}}(z)
-\bigg[
-\langle m\mathbf{k}|\mathbf{v}^\Sigma | \mathbf{r}\rangle
-\langle \mathbf{r} | n \mathbf k \rangle +
-\langle m\mathbf{k} | \mathbf{r}\rangle
-\langle \mathbf{r} | \mathbf{v}^\Sigma | n \mathbf k \rangle\bigg]
-\nonumber\\
-&=
-\frac{1}{2}
-\int \mathrm{d}^3 r\,
- {\mathbf{\cal C}}(z)
- \bigg[
-\psi_{n\mathbf{k}}(\mathbf{r})
-\mathbf{v}^{\Sigma *}\psi^*_{m\mathbf{k}}(\mathbf{r})
-+ 
-\psi^*_{m\mathbf{k}}(\mathbf{r})\mathbf{v}^\Sigma
-\psi_{n\mathbf{k}}(\mathbf{r})
-\bigg]
-\nonumber\\
-&=
-\int \mathrm{d}^3 r\,
-\psi^*_{m\mathbf{k}}(\mathbf{r})
-\left[\frac{{\mathbf{\cal C}}(z) \mathbf{v}^\Sigma +
-\mathbf{v}^\Sigma {\mathbf{\cal C}}(z)}{2}\right]
-\psi_{n\mathbf{k}}(\mathbf{r})
-\nonumber\\
-&=
-\int \mathrm{d}^3 r\,
-\psi^*_{m\mathbf{k}}(\mathbf{r})
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}
-\psi_{n\mathbf{k}}(\mathbf{r})
-,
-\end{align*}
-where, we used the hermitian property of $\mathbf{v}^\Sigma$ and defined
-\begin{equation*}
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}
-=
-\frac{{\mathbf{\cal C}}(z) \mathbf{v}^\Sigma +
-\mathbf{v}^\Sigma {\mathbf{\cal C}}(z)}{2}
-.
-\end{equation*} 
-We see that the replacement
-\begin{equation}
-\mathbf{V} \to \boldsymbol{\mathcal{\cal V}}=\frac{{\mathbf{\cal C}}(z) \mathbf{V} +
-\mathbf{V} {\mathbf{\cal C}}(z)}{2}
-,
-\label{vcali}
-\end{equation} 
-is all that is needed to change any of the
-electron velocity operators $\mathbf{V}$ to the new velocity
-operator $\boldsymbol{\mathcal{\cal V}}$ that implicitly takes into account the
-contribution of the region of the slab given by ${\mathbf{\cal C}}(z)$.
-We note that this modified operator is hermitian as it should.\cite{note2}
-The operator $\mathbf{V}$ could be any of those given by Eq.~\eqref{vop2},
-thus
-\begin{align*}
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}
-&=
-\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}
-+
-\boldsymbol{\mathcal{\cal V}}^{S}
-\nonumber\\
-\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}
-&=
-\boldsymbol{\mathcal{\cal V}}
-+
-\boldsymbol{\mathcal{\cal V}}^{\mathrm{nl}}
-.
-\end{align*}
-To calculate
-$\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{nm}(\mathbf{k})$ 
-we calculate the matrix elements of 
-$\boldsymbol{\mathcal{\cal V}}^{\mathrm{LDA}}$ and $\boldsymbol{\mathcal{\cal V}}^{S}$
- (separately)
-according to the expressions of
-Appendices \ref{calpcalc} and \ref{calvs}.
-If not stated differently, calligraphic letters correspond to layer quantities.
-
-To limit the SHG response to one surface, Eq.~\eqref{vcali} 
-for $\boldsymbol{\mathcal{\cal V}}$ was proposed in 
-Ref.~\onlinecite{reiningPRB94} and later used in Refs.
-\onlinecite{mendozaPRL98},
-\onlinecite{mendozaPRB01},
-\onlinecite{sanoPRB02},
- and \onlinecite{mejiaRMF04}. 
-In this article, for the first time, we formally introduce the cut function ${\mathbf{\cal C}}(z)$ 
-for the second-harmonic optical response of semiconductor surfaces,
-from an average of the second order polarization over the region of interest.
-
-Using
-$\boldsymbol{\mathcal{\cal J}}=d\boldsymbol{\mathcal{\cal P}}/dt$ 
-and Eq.~\eqref{jmac2} 
-we obtain the SH polarization of a given region as
-\begin{equation}
-\boldsymbol{\mathcal{\cal P}}^{(2)}(2\omega)
-=\frac{ie}{2A\tilde\omega}
-\int \frac{dk^3}{8\pi^3}
-\sum_{mn}
-\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{mn}(\mathbf{k})
-\rho^{(2)}_{nm}(\mathbf{k};2\omega)
-,
-\label{Pjikn}
-\end{equation}
-and using Eqs.~\eqref{pshg} and \eqref{j.2} 
-leads to
-\begin{align}
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega) 
-&=
-\frac{e^2}{2A\hbar\tilde\omega}
-\int \frac{dk^3}{8\pi^3}
-\sum_{mn}
-\frac{{\mathbf{\cal V}}^{\Sigma,\mathrm{a}}_{mn}(\mathbf{k})}
-{\omega^\Sigma_{nm\mathbf{k}}-2\tilde\omega}
-\bigg[
--(B_{nm}^{\mathrm{c}}(\mathbf{k},\omega))_{;k^{\mathrm{b}}}
-\nonumber \\
-&
-+i\sum_q\left(r_{nq}^{\mathrm{b}}B_{qm}^{\mathrm{c}}(\mathbf{k},\omega) -
-  B_{nq}^{\mathrm{c}}(\mathbf{k},\omega) 
-  r_{qm}^{\mathrm{b}}\right)
-\bigg]
-,
-\label{Pjikn2}
-\end{align}
-which gives the susceptibility 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ 
-of the layers of the slab specified by ${\mathbf{\cal C}}(z)$. 
-We mention that the units of 
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$
-are m$^2$/V, as they should be for a surface SH susceptibility.
-Using Eq.~\eqref{j.1} we
-split this equation into
-two contributions from the first and second terms on the right hand side
-of Eq.~\eqref{Pjikn2}:
-\begin{equation}
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_i (-2\omega;\omega,\omega)
-=-\frac{e^3}{A\hbar^22\tilde\omega}
-\int \frac{dk^3}{8\pi^3}
-\sum_{mn}
-\frac{{\mathbf{\cal V}}_{mn}^{\Sigma,\mathrm{a}}}{\omega^\Sigma_{nm}-2\tilde\omega}
-\left(\frac{f_{mn}r_{nm}^{\mathrm{b}}}{\omega^\Sigma_{nm}-\tilde\omega}\right)_{;k^{\mathrm{c}}}
-,
-\label{chii}
-\end{equation} 
-related to intraband transitions, and 
-\begin{equation}
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_e (-2\omega;\omega,\omega)
-=\frac{ie^3}{A\hbar^22\tilde\omega}
-\int \frac{dk^3}{8\pi^3}
-\sum_{qmn}
-\frac{{\mathbf{\cal V}}_{mn}^{\Sigma,\mathrm{a}}}{\omega^\Sigma_{nm}-2\tilde\omega}
-\left(
-\frac{r_{nq}^{\mathrm{c}} r_{qm}^{\mathrm{b}} 
-f_{mq}}{\omega^\Sigma_{qm}-\tilde\omega}
--\frac{r_{nq}^{\mathrm{b}} r_{qm}^{\mathrm{c}} 
-f_{qn}}{\omega^\Sigma_{nq}-\tilde\omega}
-\right),
-\label{chie}
-\end{equation} 
-related to interband transitions.
-The generalized derivative in Eq.~\eqref{chii} is dealt with by the chain rule 
-\begin{equation}
-\left(\frac{f_{mn}r_{nm}^{\mathrm{b}}}{\omega^\Sigma_{nm}-\tilde\omega}\right)_{;k^{\mathrm{c}}}=
-\frac{f_{mn}}{\omega^\Sigma_{nm}-\tilde\omega}\left(r_{nm}^\mathrm{b}\right)_{;k^{\mathrm{c}}}
--\frac{f_{mn}r_{nm}^{\mathrm{b}}\Delta_{nm}^\mathrm{c}}{(\omega^\Sigma_{nm}-\tilde\omega)^2}
-,
-\label{gene2}
-\end{equation}
-where substituting $H^\Sigma_0$ 
-into Eq.~\eqref{conmri3n} and then
-Eq.~\eqref{chon.10}
-we obtain
-\begin{equation*}
-\left(\omega^\Sigma_{nm}\right)_{;k^{\mathrm{a}}}
-=
-\left(\omega^\mathrm{LDA}_{nm}\right)_{;k^{\mathrm{a}}}
-= 
-v_{nn}^{\mathrm{LDA},\mathrm{a}}-v_{mm}^{\mathrm{LDA},\mathrm{a}}\equiv\Delta_{nm}^{\mathrm{a}}
-.
-\end{equation*} 
-The apparent divergence as $\tilde\omega\to 0$
-in Eqs. \eqref{chii} and \eqref{chie},  
-is removed  by
- a partial fraction expansion over $\tilde\omega$. 
-Using time-reversal symmetry, an integration by parts to 
-remove the square in the denominator of the second term of Eq.~\eqref{gene2}, 
-and taking the limit of $\eta\to 0$, 
-we obtain the following expressions for the imaginary parts of 
-Eqs. \eqref{chii} and \eqref{chie}, 
-\begin{subequations}\label{chis}
-\begin{equation}
-\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}]= 
-\frac{\pi |e|^3}{2\hbar^2}
-\int \frac{dk^3}{8\pi^3}
-\sum_{vc}\sum_{q\neq(v,c)}\frac{1}{\omega^\Sigma_{cv}}
-\left[
-\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{qc}\{r^{\mathrm{b}}_{cv}r^{\mathrm{c}}_{vq}\}]}
-{(2\omega^\Sigma_{cv}-\omega^\Sigma_{cq})} 
--\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vq}\{r^{\mathrm{c}}_{qc}r^{\mathrm{b}}_{cv}\}]}
-{(2\omega^\Sigma_{cv}-\omega^\Sigma_{qv})}
-\right]\delta(\omega^\Sigma_{cv}-\omega),
-\end{equation}  
-\begin{equation}
-\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}]= 
-\frac{\pi\vert e\vert^3}{2\hbar^2}
-\int \frac{dk^3}{8\pi^3}
-\sum_{cv}\frac{1}{(\omega^\Sigma_{cv})^{2}}
-\left[
-\mathrm{Re}\left[\left\{r^{\mathrm{b}}_{cv}\left(\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\right)_{;k^{\mathrm{c}}}\right\}\right]
-+\frac{\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
-\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\Sigma_{cv}} 
-\right]\delta(\omega^\Sigma_{cv}-\omega),
-\end{equation}
-\begin{equation}
-\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}]= 
--\frac{\pi |e|^3}{2\hbar^2}
-\int \frac{dk^3}{8\pi^3}
-\sum_{vc}\frac{4}{\omega^\Sigma_{cv}}
-\left[
-\sum_{v'\ne
-  v}\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\{r^{\mathrm{b}}_{cv'}r^{\mathrm{c}}_{v'v}\}]}
-{2\omega^\Sigma_{cv'}-\omega^\Sigma_{cv}}
-- \sum_{c'\ne
-  c}\frac{\mathrm{Im}[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\{r^{\mathrm{c}}_{cc'}r^{\mathrm{b}}_{c'v}\}]}
-{2\omega^\Sigma_{c'v}-\omega^\Sigma_{cv}}
-\right]\delta(\omega^\Sigma_{cv}-2\omega),
-\end{equation}
-\begin{equation}
-\mathrm{Im}[\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}]= 
- \frac{\pi \vert
-   e\vert^{3}}{2\hbar^2}
-\int \frac{dk^3}{8\pi^3}
-\sum_{vc}\frac{4}{(\omega^\Sigma_{cv})^{2}}
-\left[\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{\left(r^{\mathrm{b}}_{cv}\right)_{;k^{\mathrm{c}}}
-\right\}\right] -
-\frac{2\mathrm{Re}\left[\mathcal{V}^{\Sigma,\mathrm{a}}_{vc}\left\{r^{\mathrm{b}}_{cv}
-\Delta^{\mathrm{c}}_{cv}\right\}\right]}{\omega^\Sigma_{cv}}\right]\delta(\omega^\Sigma_{cv}-2\omega)
-,
-\end{equation}
-\end{subequations}
-where we have split the interband and intraband $1\omega$ and $2\omega$
-contributions and supressed the $\omega$ arguments for 
-convenience of notation.
-The factor of 2 for spin degeneracy is not included in Eq.~\eqref{chis}. 
-The real part of each contribution can be obtained through
-a Kramers-Kronig transformation\cite{nicolasPRB14} and
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}=
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega} 
-+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}
-+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega} 
-+\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}
-$.
-To fulfill the required intrinsic permutation symmetry, 
-the $\{\}$ notation symmetrizes the $\mathrm{b}\mathrm{c}$ Cartesian indices, i.e. 
-$\{u^{\mathrm{b}}s^{\mathrm{c}}\}=(u^{\mathrm{b}}s^{\mathrm{c}}+u^{\mathrm{c}}s^{\mathrm{b}})/2$,
-and thus
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}=\chi^{\mathrm{a}\mathrm{c}\mathrm{b}}$.
-The full expressions for
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$, along 
-with the various quantities involved in Eq.~\eqref{chis} are given in
-the Appendix \ref{appe}. 
-We mention that if we take ${\mathbf{\cal C}}(z)=1$ through out, the layered
-matrix elements $\boldsymbol{\mathcal{\cal V}}^{\Sigma}_{nm}$ become standard bulk-like
-$\mathbf{v}^{\Sigma}_{nm}$ matrix elements. We mention that in this
-case, Eq.~\eqref{chis} is equivalent to the expressions of
-Ref.~\onlinecite{cabellosPRB09}, valid for bulk semiconductors.
-
-Finally, we could also calculate the nonlinear surface susceptibility as 
-\begin{equation}
-\boldsymbol{\chi}(-2\omega;\omega,\omega)
-= \sum_{\{\ell\}}\boldsymbol{\chi}^\ell(-2\omega;\omega,\omega),
-\label{chiijksur}
-\end{equation} 
-where $\ell$ would denote a particular layer chosen through
-${\mathbf{\cal C}}^\ell(z)$ of Eq.~\eqref{sz} and
-$\{\ell\}$
-is meant to be a chosen set of layers. For instance, 
-one can take a single layer 
-encompassing half of the slab, or take each 
-atomic layer individually to the middle
-of the slab. For the first case there is 
-a single summand
-in Eq.~\eqref{chiijksur}. For the second case
-there is a sum from $\ell=1$, denoting the first layer 
-right at the surface, to $\ell=N$, denoting the layer at the middle of the slab 
-that behaves like a bulk layer.
-We remark that the value of 
-$N$ is not universal and
-the slab needs to have enough atomic layers 
-in order to give converged results for 
-$\boldsymbol{\chi} (-2\omega;\omega,\omega)$. 
-We can use Eq.~\eqref{chiijksur} for 
-either the front or the back surface.
-
-\section{Results}\label{results}
-
-In this section we present a 
-relevant test case to check the
-consistency of our approach. We have selected
-a clean Si(001) surface with a $2\times 1$ surface reconstruction.
-The slab for such a surface could be chosen to be centrosymmetric 
-by creating the front and back
-surfaces with the same $2\times 1$
-reconstruction. However, we choose to terminate one 
-of the surfaces with hydrogen producing an ideal  
-terminated bulk Si surface. The H atoms simply saturate the 
-dangling bonds of the bulk-like Si atoms at the surface, as seen in 
-Fig.~\ref{si2x1}.
-We take the $z$ coordinate pointing out of the surface and the $x$
-coordinate along the crystallographic [011] direction is parallel to the dimmers. 
-\begin{figure}
-\centering 
-\includegraphics[scale=.8]{images/si2x1-crop}
-\caption{(color on line) The slab shows a 
-    clean Si(001)$2\times 1$ front surface 
-    with an ideal terminated Si bulk back surface. The
-    dangling bonds are H (small balls) saturated. 
-    This image depicts 12 Si atomic layers
-    with one H atomic layer. 
-\label{si2x1}} 
-\end{figure} 
-The idea behind this slab configuration is that the 
-cristalline symmetry of the H terminated surface imposes that 
-$\chi_{\mathrm{H}}^{xxx}=0$.
- The $2\times 1$ surface has no 
-such restrictions, so $\chi_{2\times 1}^{xxx}\ne 0$.
-This is due to the fact that along the $y$ direction there is a mirror plane for the
-H-saturated surface, whereas 
-for the $2\times 1$ surface this mirror is lost as the dimers are
-asymmetric along $x$. 
-Thus, calculating $\chi^{xxx}$ for the full-slab, or the 
-half-slab containing the $2\times 1$ surface\cite{note1}
-should yield the same result since the contribution from the H
-saturated surface is zero regardless. 
-We must check that the following 
-relationship is satisfied for this particular slab
-\begin{equation*}
-\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega) 
-=
-\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega) 
-,
-\end{equation*}
-where
-$\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega)$ is calculated using
-${\mathbf{\cal C}}(z)=1$ from the upper half containing the $2\times 1$ 
-surface reconstruction, as seen in Fig.~\ref{si2x1},
-and $\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega)$ is calculated using
-${\mathbf{\cal C}}(z)=1$ through the full slab.
-We show the results for this comparison in the remainder  
-of this section.  
-Also, we checked that 
-for the
-dihydride surface
-$\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega)=0$.
-
-The self-consistent ground state and the Kohn-Sham states were
-calculated in the DFT-LDA framework using the plane-wave 
-ABINIT code.\cite{abinit}
-We used Troullier-Martins pseudopotentials\cite{troullierPRB91} that are 
-fully separable nonlocal pseudopotentials in the Kleinman-Bylander 
-form.\cite{kleinmanPRL82}
-The contribution of $\mathbf{v}^\mathrm{nl}$ 
-and 
-$\boldsymbol{\mathcal{\cal V}}^\mathrm{nl}$ 
-to Eq.~\eqref{chis} is carried out 
-using the DP code.\cite{olevanoDP}
-The surfaces have been studied with the experimental lattice constant of 5.43 \AA. 
-Structural optimizations were performed 
-with the ABINIT code.\cite{abinit}  
-The geometry optimization has been carried out in slabs of 12
-atomic layers where the central four layers where fixed at the bulk
-positions.  
-The structures were relaxed until the Cartesian force components were less than 5 meV/\AA. 
-The geometry optimization for the clean surface gives
-a dimer buckling of 0.721 \AA, and a dimer length of 2.301 \AA.  
-For the  Si(001)$1\times 1$:2H dihydride surface, we have obtained a Si-H bond distance of 1.48 \AA. 
-This results are in good agreement with previous 
-theoretical studies.\cite{caramellaPRB09,mendozaPRB06}
-The vacuum size is equivalent to one quarter the
-size of the slab, avoiding the effects produced by possible 
-wave-function tunneling from the contiguous surfaces of the full
-crystal formed by the repeated super-cell scheme.\cite{mendozaPRB06}
-
-Spin-orbit, local field, and electron-hole 
-attraction\cite{beyond}
-effects on the SHG process are all neglected.
-Although these are important factors in the optical response of a semiconductor,
-their efficient calculation is still theoretically and  
-numerically challenging and  
-under debate. This merits further study but is beyond the scope of this paper.
-For a given slab size, we find the converged spectra 
-to obtain the relevant parameters. The most important of 
-these are: an energy cut-off of 10 Ha, an equal number of conduction and 
-valence bands, and a set of 244 $\mathbf{k}$-points.
-The $\mathbf{k}$-points are used for the linear analytic 
-tetrahedron method for evaluating the
-Brillouin zone integrals 
-where special care was taken to examine the double resonances
-of Eq.~\eqref{chis}. \cite{nastosPRB05}
-All spectra were calculated with a Gaussian smearing of 0.15 eV.
-\begin{figure}
-\centering 
-\includegraphics[scale=.8]{plots/fig1}
-\caption{(color on line) 
-$|\chi_{\mathrm{half-slab}}^{xxx}|$ vs $\hbar\omega$ 
-for the slab
-with 16, 24, 32, and 40 atomic Si layers. The front surface 
-is in a clean $2\times 1$ reconstruction and the back 
-surface is an ideal terminated bulk H-saturated dangling bonds (see Fig.~\ref{si2x1}).
-\label{fig1}} 
-\end{figure}
-\begin{figure}
-\centering 
-\includegraphics[scale=.8]{plots/fig2}
-\caption{(color on line) 
-$\chi^{xxx}_{\mathrm{half-slab}}$
-and 
-$\chi^{xxx}_{\mathrm{full-slab}}$
-vs $\hbar\omega$ for a slab with 32 
-atomic Si layers plus one H layer. 
-\label{fig2}} 
-\end{figure}
-
-We must evaluate 
-$T^{\mathrm{a}\mathrm{b}}_{nm}=(i/\hbar)[r^\mathrm{b},v^{\mathrm{nl},\mathrm{a}}]_{nm}$
-in order to obtain 
-Eqs.~\eqref{tau.1} and \eqref{tau.1n} that are required for 
-Eq.~\eqref{chis}.
-Computing second-order 
-derivatives is required thus making the numerical procedure very 
-time consuming. This adds significantly to the already lengthy time needed 
-for the calculation of the $\mathbf{v}^\mathrm{nl}$ contribution that is
-proportional only to 
-the first order derivatives. 
-Memory requirements are also increased for both $\mathbf{v}^\mathrm{nl}$ and 
-$[\mathbf{r},\mathbf{v}^\mathrm{nl}]$. However, the contribution from $[\mathbf{r},\mathbf{v}^\mathrm{nl}]$ 
-is very small\cite{valerie} and therefore we neglect it in this work.
-
-\subsection{Full-slab results}\label{fsresults}
-
-In Fig.~\ref{fig1} we show $|\chi_{\mathrm{full-slab}}^{xxx}|$
-for the slab with
-16, 24, 32, and 40 Si atomic layers, without the contribution of
-$\mathbf{v}^{\mathrm{nl}}$ and with no scissors correction.
-Since the clean Si(001) surface is $2\times 1$,
-there are two atoms per atomic layer, thus the total number of atoms per slab is 
-twice the number of atomic layers of the slab. 
-In making the slabs larger, we add steps of 8 layers of bulk-like atomic positions. 
-We note that the response differs substantially 
-for 16 and 24 layers but is quite similar for 32 and 40 layers.
-As explained above,
-the calculation of the $\mathbf{v}^\mathrm{nl}$ contribution is 
-computationally expensive.
-A good compromise between the accuracy in the convergence of
-$\chi^{xxx}_{\mathrm{full-slab}}$ as a function of the number
-of layers in the slab, and the computational 
-expense is to consider
-the slab with 32 Si atomic layers as 
-an accurate representation of our 
-system.
-
-\subsection{Half-slab vs. full-slab}
-
-In Fig.~\ref{fig2}
-we compare 
-$\chi^{xxx}_{\mathrm{half-slab}}$  
-vs. 
-$\chi^{xxx}_{\mathrm{full-slab}}$ 
-for the four different possibilities 
-between including or not including the
-effects of $\mathbf{v}^\mathrm{nl}$ or the scissors correction
-$\hbar\Delta$.   
-For these results we chose
-$\hbar\Delta=0.5$ eV, that is the GW gap reported in
-Refs.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
-We see that for all four instances the 
-difference between responses is quite small.
-Indeed, when the value $|\chi^{xxx}|$ 
-is large the difference between the two is very small; 
-when the value is small the difference increases only slightly, 
-but the spectra is so close to zero that it is negligible. 
-These differences would decrease as the number of atomic layers 
-increases. We remark that 32 layers in the slab is more than enough 
-to confirm that the extraction of the surface second-harmonic 
-susceptibility from the $2\times 1$ surface is readily possible 
-using the formalism contained in Eq.~\eqref{chis}.
-We have confirmed that for the dihydride surface
-$|\chi^{xxx}_{\mathrm{half-slab}}|\approx 0$ (not shown).
-This confirms the validity of our theory and is the main result of
-this article; through the proposed layer formalism we can calculate the surface SH
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$    
-including
-the
-contribution of the nonlocal part of the pseudopotentials 
-and the part of the many-body effects through the scissors correction.
-Our scheme should work for any slab.
-
-\subsection{\texorpdfstring{Results for $\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)$}
-{Results for Xxxx(2x1)(-2w;w,w)}}
-
-We proceed to explain 
-some of the features seen in $|\chi^{xxx}_{2\times 1}|$ that, as
-explained above, are obtained 
-by calculating $|\chi^{xxx}_{\mathrm{half-slab}}|$.
-First, from Fig.~\ref{fig2} we note a series of resonances 
-that derive from 1$\omega$ and 2$\omega$ terms in
-Eq.~\eqref{chis}. 
-Notice that the 
-2$\omega$ resonances start below $E_g/2$ where $E_g$ is the band gap
-(0.53 eV  for LDA and
-1.03 eV if the scissor is used with
-$\hbar\Delta=0.5$ eV).
-These resonances
-come from the electronic states of the 
-$2\times 1$ surface, that lie inside the bulk band gap of Si and are the 
-well known electronic surface states.\cite{rohlfingPRB95} 
-In Fig.~\ref{fig3}
-we see that the effect of $\mathbf{v}^\mathrm{nl}$
- reduces the value of   
-$|\chi^{xxx}_{2\times 1}|$  
-by 15-20\% 
-showing the importance of this contribution for a correct calculation  
-of SSHG, in agreement with the analysis for bulk semiconductors.\cite{luppiPRB08}
-However, the inclusion of $\mathbf{v}^\mathrm{nl}$ does not
-changes the 
-spectral shape of $|\chi^{xxx}_{2\times 1}|$;
- this also can be confirmed from
-the cases of zero scissors correction from Fig.~\ref{fig2}.
-\begin{figure}
-\centering 
-\includegraphics[scale=.8]{plots/fig3}
-\caption{(color on line) 
-$\chi^{xxx}_{2\times 1}$
-vs $\hbar\omega$ for a slab with 32 
-atomic Si layers plus one H layer, 
-with and without the contribution from $\mathbf{v}^\mathrm{nl}$.
-\label{fig3}} 
-\end{figure}
+% @1270: Results section C
+% =====================================================
 
@@
 $1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
-for the non-rigid shift. 
+for the non-rigid shift. The reduction of the spectral strength is in 
+agreement with previous calculations for bulk systems.
+\cite{nastosPRB05, luppiPRB10, leitsmannPRB05}
 When we compare 
@@
 \end{figure}
-
-Although local fields are neglected they should, in principle, 
-be small parallel to the interface as the electric field is continuous.
-So, we would expect that the $xxx$ component of 
-$\boldsymbol{\chi}(-2\omega;\omega,\omega)$ 
-would have a small influence from the local fields.
-Also, the excitonic effects ought to be explored, but
-their efficient calculation is theoretically and 
-numerically challenging\cite{beyond}   and beyond the scope 
-of this article.
-Unfortunately the experimental measurement of the $xxx$
-component of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$ is not possible as the SH 
-radiated intensity would be proportional not only to this component 
-but also to the other components of $\boldsymbol{\chi}(-2\omega;\omega,\omega)$.  
-However,
-in a forthcoming publication 
-we will present a study of SSHG from several 
-Si surfaces 
-with comparison to experimental 
-results.
-
-\section{Conclusions}\label{conc}
-
-We have presented a formulation to calculate the surface second-harmonic
-(SSH) susceptibility tensor $\boldsymbol{\chi}(-2\omega;\omega,\omega)$, using the length gauge
-formalism and within the independent particle approximation (IPA).
-It includes on equal footing:
-(i) the scissors correction, 
-(ii) the  
-contribution of the non-local part of the pseudopotentials, and  
-(iii) the cut function.
-We have used a Si(001)$2\times 1$ surface to confirm that our scheme
-correctly obtains the surface response as we confirm
-that 
-$\chi_{\mathrm{half-slab}}^{xxx}(-2\omega;\omega,\omega) 
-\approx
-\chi_{\mathrm{full-slab}}^{xxx}(-2\omega;\omega,\omega) 
-. 
-$
-Although one can in principle increase the number of atomic layers,
-$\mathbf{k}$-points, etc. to
-improve even further on the similarity of the half-slab and full-slab results, we
-have chosen a good compromise between accuracy and the burden and time
-of the computations. 
-We describe the effect of the independent inclusion of the three
-effects mentioned above in the calculation of  
-$\boldsymbol{\chi}(-2\omega;\omega,\omega)$. 
-The scissors correction shifts the spectrum to  
-higher energies though the shifting is not rigid and mixes the 
-$1\omega$ and $2\omega$ resonances,  
-and has a strong influence in the line-shape,
-as for the case of bulk semiconductors.\cite{luppiJCP10,luppiPRB10,leitsmannPRB05}
- The cut function allows us  
-to extract unequivocally $\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)$.
- The effects of the nonlocal part of the pseudopotentials  
-keeps the same line-shape of $|\chi^{xxx}_{2\times
-  1}(-2\omega;\omega,\omega)|$,
-but  
-reduces the value of   
-by 15-20\%. 
-The $xxx$ component of  
-$\boldsymbol{\chi}_{2\times 1}(-2\omega;\omega,\omega)$, 
-can not be
-experimentally isolated,
-however in a forthcoming publication 
-we will compare our formulation 
-against
-experimental 
-results. 
-We have neglected
-local field 
-and excitonic effects.
-Although these are important factors in the optical response of a semiconductor, 
-their efficient calculation is theoretically and  
-numerically challenging and still  
-under debate.\cite{beyond}   
-This merits further study but is beyond the scope of this paper. 
-Nevertheless, the inclusion of aforementioned contributions 
-in
-our scheme opens the unprecedented possibility to study 
-surface SHG with more versatility
-and 
-more accurate results.
-
-\section{Acknowledgments}
-
-B.S.M. acknowledges the Laboratoire des Solides Irradi\'es (Ecole
-Polytechnique, Palaiseau, France) for the support and hospitality
-during a sabbatical year. B.S.M. acknowledges partial support from
-CONACYT-M\'exico Grant 153930.
-
-%%%%%%%%%%%%%
-\appendix 
-\section{}\label{appe}
-We give explicit expressions for the quantities used in the evaluation 
-of Eq.~\eqref{chis}; when appropriate, some 
-intermediate steps are given for their derivation. 
-\subsection{ Expressions for 
-\texorpdfstring{$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$}{que} 
-}\label{fullchis}
-Omitting the frequency dependence for convenience of notation,
-$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}
-=
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}
-+
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}
-+
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}
-+
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}
-$, where 
-\begin{align*}
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,\omega}
-&= \frac{|e|^3}{2\hbar^2}
-\int \frac{dk^3}{8\pi^3}
-\sum_{l\ne mn}\frac{f_{mn}}{\omega^\Sigma_{nm}}
-\left[
-  \frac{\mathrm{Im}\left[\mathcal{V}^{\Sigma,\text{a}}_{ml}
-        \{r^{\text{c}}_{ln}
-        r^{\text{b}}_{nm}\}\right]}
-        {2\omega^\Sigma_{nm}-\omega^\Sigma_{lm}}
- -\frac{\mathrm{Im}\left[\mathcal{V}^{\Sigma,\text{a}}_{ln}
-        \{r^{\text{b}}_{nm}
-        r^{\text{c}}_{ml}\}\right]}
-        {2\omega^\Sigma_{nm}-\omega^\Sigma_{nl}}
-\right]\frac{1}{\omega^\Sigma_{nm}-\tilde\omega}
-,
-\nonumber\\
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,\omega}
- &=   \frac{|e|^3}{2\hbar^2}\int \frac{dk^3}{8\pi^3}
-\sum_{mn}
-    \frac{f_{mn}}{(\omega^{\Sigma}_{nm})^{2}}
-\Biggl[
-    \mathrm{Re}\left[r^{\text{b}}_{nm}
-    \left(\mathcal{V}^{\Sigma,\text{a}}_{mn}\right)_{;k^{\text{c}}}\right]
-+   \frac{\mathrm{Re}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
-    r^{\text{b}}_{nm}\right]\Delta^{\text{c}}_{nm}}{\omega^{\Sigma}_{nm}} 
-\Biggr]\frac{1}{\omega^{\Sigma}_{nm}-\tilde\omega}
-,
-\nonumber\\
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{e,2\omega}
-&= \frac{|e|^3}{2\hbar^2}\int \frac{dk^3}{8\pi^3}
-\sum_{l\ne mn}
-4\left[
-  \frac{f_{ln}\,\mathrm{Im}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
-        \{r^{\text{b}}_{nl}
-        r^{\text{c}}_{lm}\}\right]}
-        {2\omega^\Sigma_{nl}-\omega^\Sigma_{nm}}
- -\frac{f_{ml}\,\mathrm{Im}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
-        \{r^{\text{c}}_{nl}
-        r^{\text{b}}_{lm}\}\right]}
-        {2\omega^\Sigma_{lm}-\omega^\Sigma_{nm}}
-\right]\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}
-,
-\nonumber\\
-\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}_{i,2\omega}
-&=  \frac{|e|^3}{2\hbar^2}\int \frac{dk^3}{8\pi^3}
-\sum_{mn}
-    \frac{4f_{mn}}{(\omega^{\Sigma}_{nm})^{2}}
-    \left[\mathrm{Re}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
-    \left(r^{\text{b}}_{nm}\right)_{;k^{\text{c}}}\right] 
--   \frac{2\,\mathrm{Re}\left[\mathcal{V}^{\Sigma,\text{a}}_{mn}
-    r^{\text{b}}_{nm}\right]\Delta^{\text{c}}_{nm}}{\omega^{\Sigma}_{nm}}\right]
-    \frac{1}{\omega^{\Sigma}_{nm}-2\tilde\omega}
-. 
-\end{align*}
-
-%%%%
-\subsection{ Expressions for 
-\texorpdfstring{${\boldsymbol{\mathcal{\cal V}}}_{nm}(\mathbf{k})$}{que} 
-and 
-\texorpdfstring{${\cal C}_{nm}(\mathbf{k})$}{que}
-}\label{calpcalc}
-
-Expanding the wave function in plane waves we obtain
-\begin{equation*}
-\psi_{n\mathbf{k}}(\mathbf{r})=\sum_\mathbf{G} A_{n\mathbf{k}}(\mathbf{G})e^{i(\mathbf{k}+\mathbf{G})\cdot\mathbf{r}}
-,
-\end{equation*}
-where $\{\mathbf{G}\}$ are the reciprocal basis vectors satisfying
-$e^{i\mathbf{R}\cdot\mathbf{G}}=1$, with $\{\mathbf{R}\}$ the translation vectors in real
-space, and $A_{n\mathbf{k}}(\mathbf{G})$ the expansion coefficients. Using
-$m_e\mathbf{v}=\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ into 
-Eq.~\eqref{vcali}
-we obtain,\cite{mendozaPRB06}
-\begin{equation}
-\boldsymbol{\mathcal{\cal V}}_{nm}(\mathbf{k})=
-\frac{\hbar}{2m_e}
-\sum_{\mathbf{G},\mathbf{G}'} A^*_{n\mathbf{k}}(\mathbf{G}')  A_{m\mathbf{k}}(\mathbf{G})
-(2\mathbf{k}+\mathbf{G}+\mathbf{G}')
-\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel}  
-f(G_\perp-G'_\perp)
-,
-\label{eni.2}
-\end{equation}   
-where
-\begin{equation*}
-f(G_\perp)=\frac{1}{L}\int{\mathbf{\cal C}}(z) e^{iG_\perp z}dz,
-\end{equation*}
-with $f^*(G_\perp)=f(-G_\perp)$,
-and $L$
-is the length of the super-cell. 
-The reciprocal lattice vectors $\mathbf{G}$ are 
-decomposed into components
-parallel ($\mathbf{G}_\parallel$), and perpendicular ($G_\perp \hat z$)
-to the surface, so
-that $\mathbf{G} = \mathbf{G}_\parallel + G_\perp\hat z$.
-The double summation over the $\mathbf{G}$ vectors can be 
-calculated efficiently by  
-creating a pointer array to identify all the plane-wave coefficients  
-associated with the same $G_\parallel$.
-
-Likewise we obtain that
-\begin{equation}
-{\mathbf{\cal C}}_{nm}(\mathbf{k})=
-\sum_{\mathbf{G},\mathbf{G}'} A^*_{n\mathbf{k}}(\mathbf{G}')  A_{m\mathbf{k}}(\mathbf{G})
-\delta_{\mathbf{G}_\parallel \mathbf{G}'_\parallel} 
-f(G_\perp-G'_\perp)
-.
-\label{eni.4}
-\end{equation}  
-If ${\mathbf{\cal C}}(z)=1$, then $f(G_\perp)=\delta_{G_\perp 0}$ and we 
-obtain the full-slab/bulk values, 
-$\mathbf{v}_{nm}(\mathbf{k})$ and ${\mathbf{\cal C}}_{nm}(\mathbf{k})=\delta_{nm}$,
-from Eqs.~\eqref{eni.2} and \eqref{eni.4}.
-
-%%%%
-\subsection{Expressions for 
-\texorpdfstring{$({\mathbf{\cal V}}^{\mathrm{LDA},a}_{nm})_{;k^b}$}{Vnonlocal},
-\texorpdfstring{$(r^a_{nm})_{;k^b}$}{Vnonlocal}
-and \texorpdfstring{$({\mathbf{\cal C}}_{nm})_{;\mathbf{k}}$}{Vnonlocal}
-}\label{appvnl}
-
-Using Eqs.~\eqref{rnminn}, \eqref{conmri3n}, \eqref{gendevnn}, and
-defining 
-$
-{\cal T}^{\mathrm{a}\mathrm{b}}\equiv[r^{\mathrm{b}},{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}]
-\equiv
-[r^{\mathrm{b}},{\mathbf{\cal V}}^\mathrm{a}]
-$
-one can show that
-\begin{equation}
-({\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}=
-{\cal T}_{nm}^{\mathrm{a}\mathrm{b}}
-+i
-\sum_{q}
-\bigg(
-r^{\mathrm{b}}_{nq}  
-{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{q m}
--
-{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nq}   
-r^{\mathrm{b}}_{q m}
-\bigg)  
-+i  
-r^{\mathrm{b}}_{nm}
-\Delta^{\mathrm{a}}_{mn}
-,
-\label{nmesn}
-\end{equation}
-where
-\begin{equation*}
-\Delta^{\mathrm{a}}_{mn}
-=
-{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{nn}  
--
-{\mathbf{\cal V}}^{\mathrm{LDA},\mathrm{a}}_{mm}  
-,
-\end{equation*} 
-\begin{equation}
-{\cal T}_{nm}^{\mathrm{a}\mathrm{b}}
-=
-\frac{\hbar}{m_e}\delta_{\mathrm{a}\mathrm{b}} 
-{\cal C}_{nm} 
--\hbar 
-\sum_q  
-L^{\mathrm{a}\mathrm{b}}_{nq} 
-{\cal C}_{qm} 
-,
-\label{tau.1}
-\end{equation}   
-and
-\begin{equation}
-L_{nm}^{\mathrm{a}\mathrm{b}}
-=\frac{i}{\hbar}[r^{\mathrm{b}},v^{\mathrm{nl},\mathrm{a}}]_{nm}
-.
-\label{lab}
-\end{equation}  
-The matrix elements $L^{\mathrm{a}\mathrm{b}}_{nm}$
-are small as 
-compared to the other terms, thus we neglect it throwout this work.\cite{valerie} 
-Notice that
-$(v^{\mathrm{LDA},\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}$ is obtained 
-from Eq.~\eqref{nmesn} by 
-taking 
-${\mathbf{\cal C}}(z)=1$ or ${\cal C}_{nm}=\delta_{nm}$.
-
-To obtain $(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}$ we use Eq.~\eqref{chon.10} to
-write
-$(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}
-=(v^{\mathrm{LDA},\mathrm{a}}_{nm}/i\omega^\mathrm{LDA}_{nm})_{;k^{\mathrm{b}}}
-$ and simply apply the chain rule,
-\begin{equation}
-(r^{\mathrm{a}}_{nm})_{;k^{\mathrm{b}}}
-=
-T^{\mathrm{a}\mathrm{b}}_{nm}
-+
-\frac{ 
-r^{\mathrm{b}}_{nm}
-\Delta^{\mathrm{a}}_{mn}
-+r^{\mathrm{a}}_{nm}
-\Delta^{\mathrm{b}}_{mn}
-}
-{\omega^\mathrm{LDA}_{nm}}
-+
-\frac{i}{\omega^\mathrm{LDA}_{nm}}
-\sum_{q}
-\bigg(
-\omega^\mathrm{LDA}_{q m} 
-r^{\mathrm{b}}_{nq} 
-r^{\mathrm{a}}_{q m}
--
-\omega^\mathrm{LDA}_{nq} 
-r^{\mathrm{a}}_{nq} 
-r^{\mathrm{b}}_{q m}
-\bigg)
-,
-\label{rka}
-\end{equation} 
-where 
-\begin{equation*}
-\Delta^{\mathrm{a}}_{mn}
-=
-v^{\mathrm{LDA},\mathrm{a}}_{nn}  
--
-v^{\mathrm{LDA},\mathrm{a}}_{mm}  
-,
-\end{equation*}
-and
-\begin{equation}
-T_{nm}^{\mathrm{a}\mathrm{b}}
-=\frac{\hbar}{m_e}\delta_{ab}\delta_{nm} 
--\hbar L^{\mathrm{a}\mathrm{b}}_{nm} 
-.
-\label{tau.1n} 
-\end{equation}
-Eq.~\eqref{rka} generalizes the usual expresion of
-$(r^a_{nm})_{;k^b}$ for a local 
-Hamiltonian
-\cite{aversaPRB95,nastosPRB05,cabellosPRB09,rashkeevPRB98}
-to
-the case of a
-nonlocal Hamiltonian.
-Note that the layered term
-${\cal T}^{\mathrm{a}\mathrm{b}}_{nm}$ reduces to $T^{\mathrm{a}\mathrm{b}}_{nm}$
-for the full-slab/bulk case.
-
-Again, we use Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn},
-along with $[\mathbf{r},F(\mathbf{r})]=0$, valid for 
-and any function $F(\mathbf{r})$, 
-to obtain 
-\begin{equation*} 
-({\mathbf{\cal C}}_{nm})_{;\mathbf{k}}
-=
-i 
-\sum_{q} 
- \left(\mathbf{r}_{nq}
-{\mathbf{\cal C}}_{qm}
--
-{\mathbf{\cal C}}_{nq}
-\mathbf{r}_{qm}
-\right) 
-+i\mathbf{r}_{nm}({\mathbf{\cal C}}_{mm}-{\mathbf{\cal C}}_{nn}) 
-,
-\end{equation*} 
-where we remind the reader that $\mathbf{r}_{nm}$ 
-is calculated through 
-Eq.~\eqref{chon.10} for LDA.
-
-\subsection{Expressions for  \texorpdfstring{$\boldsymbol{\mathcal{\cal V}}^{S}_{nm}$}{Vnm}
-and
-\texorpdfstring{$\Big({\boldsymbol{\mathcal{\cal V}}}^{S}_{nm}\Big)_{;\mathbf{k}}$}{(Vnm);kb}
-}\label{calvs}
-
-From Eq.~\eqref{vcali}
-\begin{equation}
-\boldsymbol{\mathcal{\cal V}}^{S}_{nm}
-=
-\frac{1}{2}\sum_q\left(   
-\mathbf{v}^S_{nq}{\mathbf{\cal C}}_{qm}+{\mathbf{\cal C}}_{nq}\mathbf{v}^S_{qm}
-\right)  
-,
-\label{a.3b}
-\end{equation}    
-where $\sum_q | q\mathbf{k}\rangle\langle q\mathbf{k} | =1$ was used
-and $\mathbf{v}^S_{nm}$ is given in Eq.~\eqref{chon.2}.
-Taking the generalized derivative of Eq.~\eqref{a.3b}
-and applying
-the chain rule, we obtain
-\begin{equation*}
-\left(\boldsymbol{\mathcal{\cal V}}^{S}_{nm}\right)_{;\mathbf{k}}
-=
-\frac{1}{2}\sum_q\left(
-(\mathbf{v}^S_{nq})_{;\mathbf{k}}{\mathbf{\cal C}}_{qm}
-+    
-\mathbf{v}^S_{nq}({\mathbf{\cal C}}_{qm})_{;\mathbf{k}}
-+
-({\mathbf{\cal C}}_{nq})_{;\mathbf{k}} \mathbf{v}^S_{qm}
-+
-{\mathbf{\cal C}}_{nq} (\mathbf{v}^S_{qm})_{;\mathbf{k}}
-\right)  
-.
-\end{equation*}    
-Again, from
-Eq.~\eqref{chon.2}, 
-\begin{equation*}
-(\mathbf{v}^S_{nm})_{;\mathbf{k}}=i\Delta f_{mn}
-(\mathbf{r}_{nm})_{;\mathbf{k}}
-,
-\end{equation*}
-a result that is in agreement with Eq. A(6) of Ref.~\onlinecite{cabellosPRB09}.
-
-%%%%%%%%%%%%%
-\bibliography{ref}
-\end{document}
-%%%%%%%%%%%%%
-
-Figures log:
-fig1 -> images/slab
-fig2 -> images/si2x1-crop
-fig3 -> plots/fig1 
-fig4 -> plots/fig2
-fig5 -> plots/fig3
-fig6 -> plots/fig4

### 2014-12-24 corrections.tex
a more BMS friendly corrections.tex

--- 2014-12-03 manuscript.tex
@@
 \usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
+\def\chon{red}
 %\usepackage{showlabels}
@@
 \cite{levinePRL89,levinePRL91,delsolePRB93}  
-This allows us to ``open'' the DFT-LDA gap to 
-its correct experimental or GW value even when experiments are not available. 
+{\color{\chon}This allows us to ``open'' the DFT-LDA gap to 
+its correct experimental or GW value for most bulk semiconductors.
+%%%% Valerie and Nicolas remove these lines
+% even when experiments are not available. 
+%%%%
+This approximation has already been used in linear optical calculations for 
+surfaces, see \cite{kippPRL96}, thus improving the agreement with 
+experimental results.}
 In this context, to correct for the underestimation of the energy band gap of 
@@
 
-To calculate the optical response 
-within the Length Gauge formalism, we take the 
-interaction Hamiltonian as
+{\color{\chon}According to ~\onlinecite{ismailPRL01}, we first start with the interaction 
+Hamiltonian expressed in the velocity gauge, containing the nonlocal parts
+$V^{nl}$ and $S$. Within the dipole approximation and using a gauge 
+transformation, it can be transformed into an effective Hamiltonian 
+\cite{Veniard, unpublished}
 \begin{equation}
@@
 \end{equation}
-The 
-treatment of the position operator
-$\mathbf{r}$ for extended Bloch states, $\langle\mathbf{r}|n\mathbf{k}\rangle$ has 
-been described in detail.\cite{adamsJCP53,blountSSP62} 
-In particular, Ref.~\onlinecite{aversaPRB95} develops
-the treatment of $\mathbf{r}$ for
-the calculation of the nonlinear optical response.
+The treatment of the position operator $\mathbf{r}$ for extended Bloch states 
+is problematic and has been discussed in Refs.\cite{adamsJCP53,blountSSP62}
+%%%% Valerie and Nicolas remove these lines
+% In particular, Ref.~\onlinecite{aversaPRB95} develops the treatment of 
+% $\mathbf{r}$ for the calculation of the nonlinear optical response.
+%%%%
+Following Ref.~\onlinecite{sipePRB00} we take the
+matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of 
+Eq.~\eqref{rde}, we obtain 
+$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\Sigma_{nm}-\tilde\omega)t}$,
+with 
+\begin{equation}
+B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\Sigma_{nm}-\tilde\omega},
+\label{j.1}
+\end{equation}
+and 
+\begin{equation}
+(\tilde{\rho}^{(2)}(t))_{nm} = \frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
+i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
+\Big)  
+-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\Sigma_{nm}-2\tilde\omega)t}
+.
+\label{j.2}
+\end{equation}}
 $\mathbf{r}$ is split into the {\it intraband} 
@@
 to obtain the scissored results for any 
-susceptibility expression, whether linear or nonlinear.\cite{nastosPRB05}
-
-Following Ref.~\onlinecite{sipePRB00} we take the
-matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of 
-Eq.~\eqref{rde}, we obtain 
-$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\Sigma_{nm}-\tilde\omega)t}$,
-with 
-\begin{equation}
-B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\Sigma_{nm}-\tilde\omega},
-\label{j.1}
-\end{equation}
-and 
-\begin{equation}
-(\tilde{\rho}^{(2)}(t))_{nm} = \frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
-i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
-\Big)  
--(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\Sigma_{nm}-2\tilde\omega)t}
-.
-\label{j.2}
-\end{equation}
-We have used the fact that for a cold semiconductor $\partial
+{\color{\chon}susceptibility expression, whether linear or nonlinear.\cite{nastosPRB05}
+
+We have used the fact that for a cold semiconductor} $\partial
 f_{n}/\partial \mathbf{k}=0$ and thus the intraband contribution to the linear
@@
 Refs.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
+{\color{\chon}This is justified by the fact that the surface states of the clean 
+Si(001) surface investigated are found, from GW calculations 
+\cite{rohlfingPRB95}, to shift rigidly and to keep their dispersion with 
+respect to LDA.}
 We see that for all four instances the 
@@
 $1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
-for the non-rigid shift. 
+for the non-rigid shift. {\color{\chon}The reduction of the spectral strength is in 
+agreement with previous calculations for bulk systems.
+\cite{nastosPRB05, luppiPRB10, leitsmannPRB05}}
 When we compare 

### 2015-01-08 referees.tex
for Valerie and Nicolas

From prb@aps.org Wed Dec 17 09:19:00 2014
Date: Tue, 16 Dec 2014 21:23:48 -0500
From: prb@aps.org
To: bms@cio.mx
Subject: {Spam?} Your_manuscript BZ12227 Anderson

    [ The following text is in the "UTF-8" character set. ]
    [ Your display is set for the "ISO-8859-1" character set.  ]
    [ Some characters may be displayed incorrectly. ]

Re: BZ12227
    Theory of surface second-harmonic generation for  semiconductors
    including effects of nonlocal operators
    by Sean M. Anderson, Nicolas Tancogne-Dejean, Bernardo S. Mendoza, et
    al.

Dear Dr. Mendoza,

The above manuscript has been reviewed by two of our referees. Comments
from the reports appear below.

These comments suggest that considerable revision of your manuscript is
in order. When you resubmit your manuscript, please include a summary
of any changes made and a succinct response to all recommendations or
criticisms contained in the reports.

Yours sincerely,

Athanasios Chantis
Assistant Editor
Physical Review B
Email: prb@aps.org
http://journals.aps.org/prb/

PRB Rapid Communications: Quality and Speed
http://journals.aps.org/prb/rapids

All PRB Editors' Suggestions feature prominently at
http://journals.aps.org/prb/

----------------------------------------------------------------------
Report of the First Referee -- BZ12227/Anderson
----------------------------------------------------------------------

The authors present a new scheme to calculate surface second-harmonic 
generation including the effects of non-local operators like the 
non-local part of the pseudo-potentials and the scissor-operator. 
Their approach reduces to the bulk one when the cut-function goes to 
one. They validated their new formulation on a clear example the 
Si(001)2x1 surface. I think their approach is correct and worth to be 
published in Physical Review B, I have only few comments and remarks.

Comments are remarks:

1) The authors claim that in this new formulation the scissor operator 
is correctly implemented respect to previous attempts to include it in 
the surface SHG response. Since some of the authors of the present 
paper are also author of the previous approaches (PRB 63, 2054406; PRB 
74, 075318) that “... incorrectly implemented” scissor operator, may 
they explain more in details what was wrong in the old papers and why 
now it is correct.

2) The idea behind the scissor operator is to shift all the conduction 
bands of the same amount. While I can believe that this approximation 
is justified in a bulk material, why it should be in case for a 
surface? The shape and localization of surface states is different 
from the one of the bulk. There are cases where this approximation has 
been tested or compared with a full GW calculation for a surface?

3) At the top of page 6 the authors write that X2 satisfies intrinsic 
permutation symmetry and then they put an equation with the same X2 in 
both sides. They should correct this mistake.

----------------------------------------------------------------------
Report of the Second Referee -- BZ12227/Anderson
----------------------------------------------------------------------

The authors present a theory of the second-harmonic generation at 
surfaces. The underlying electronic structures are calculated within 
the density functional theory. They discuss

(i) the effect of quasiparticle corrections in the framework of the 
scissors-operator approach,

(ii) the correct description of the optical-transition operator, and

(iii) the reduction to the surface response by a cut function.

The approach is timely and reasonable. The discussion of the three 
effects (i), (ii), and (iii) in combination is new. An application of 
the developed SHG approach to the Si(111)2x1 surface is demonstrated 
as a model case. Therefore, I claim that the publication of the 
manuscript in Physical Review B as a regular article can be suggested.

However, prior to publication the article needs a substantial 
revision. It concerns the presentation of the theory, the removal of 
some misleading comments, and the discussion of the three effects. 
More in detail:

- In one of their own papers by one of the authors, C. Castillo et 
al., PRB 68, 041310(R) (2003), the importance of the elimination of 
the bottom layers of the slab in the case of linear optics has been 
pointed out. Can this argument simply be applied to non-linear optics, 
where at least three light beams have to be discussed?

- The reduction of the spectral strength in Fig. 4 in the two lower 
panels is a consequence of the inclusion of the scissors operator in 
the optical matrix elements. Am I right? Is this effect really real? 
Other treatments, e.g. in Ref. [36] and [54], claim that the scissors 
operator should not be taken into the reformulation of the optical 
matrix elements.

- Abstract / Introduction: cancel “for the first time”

- Abstract: cancel: “Our scheme … is exact”

- Formulation in introduction “Radiation can occur …” is wrong

- Sec.II.B: The description of light-matter interaction by the 
Hamiltonian (6) is invalid, because Bloch matrix elements of the space 
operator cannot be computed. However, another formulation as given in 
M. Gajdos et al., PRB 73, 045112 (2006) is possible. This 
representation is more convenient because no non-local contributions 
to the optical matrix element occur. A comparison as in B. Adolph et 
al., PRB 63, 125108 (2001), done for linear optics could also be 
helpful in the case of SHG.

- The k-integration in 3D as indicated in many formulas from (18) and 
below does not fit to the slab approximation. Usually, only a 2D 
integration in k-space has to be performed.

### 2015-01-08 reply/Scissors.tex
for Valerie and Nicolas

Introduction :

However this can be a very expensive calculation and thus one uses the much simpler scissors operator scheme. 34–36
It allows to “open” the DFT-LDA gap to its correct experimental or GW value for most bulk semiconductors. 
This approximation has already been used in linear optical calculations for surfaces, see [PRL 76 2810 (1996)], thus improving the agreement with experimental results.

Results B After Refs. 64,65
This is justified by the fact that the surface states of the clean Si(001) surface investigated are found, 
from GW calculations[PRB 52 1905 (1995)], to shift rigidly and to keep their dispersion with respect to LDA.

### 2015-01-08 reply/Sect_II_B.tex
for Valerie and Nicolas

\subsection{Length Gauge Formalism}
According to ~\onlinecite{ismailPRL01}, we first start with the interaction Hamiltonian expressed in the velocity gauge, containing the nonlocal parts
$V^{nl}$ and $S$. Within the dipole approximation and using a gauge transformation, it can be transformed into an effective Hamiltonian \cite{Veniard, unpublished}
\begin{equation}
\tilde H_{I}(t)=-e\mathbf{r}\cdot \mathbf{E}(t).  \label{rde}
\end{equation}
The treatment of the position operator $\bfr$ for extended Bloch states is problematic and has been discussed in Refs. \cite{adamsJCP53,blountSSP62} 
{\color{\chon}. Following Ref.~\onlinecite{sipePRB00}, we take} the matrix elements of Eq.~\eqref{rhop} with the Hamiltonian $\tilde H_{I}(t)$ of
Eq.~\eqref{rde}, we obtain 
$(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\sigma_{nm}-\tilde\omega)t}$,
with 
\begin{equation}\label{j.1}
B_{nm}^{b}=\frac{e}{\hbar }\frac{f_{mn}r_{nm}^{b}}{\omega^\Sigma_{nm}-\tilde\omega},
\end{equation}
and 
\begin{eqnarray}\label{j.2}
(\tilde{\rho}^{(2)}(t))_{nm} &=&\frac{e}{i\hbar }\frac{1}{\omega^\Sigma_{nm}-2\tilde\omega}\bigg[
i\sum_{q }\Big(r_{nq }^{b}B_{q m}^{c}-B_{nq}^{c}r_{q m}^{b}
\Big)  
-(B_{nm}^{c})_{;k^{b}}\bigg]E^{b}(\omega)E^{c}(\omega)e^{i(\omega^\Sigma_{nm}-2\tilde\omega)t}
.
\end{eqnarray}
$\bfr$ is split into the {\it intraband} 
($\bfr_i$) and {\it interband} ($\bfr_e$) parts, where 
$\bfr=\bfr_i+\bfr_e$. 
For $\bfr_e$ one uses
\begin{align}\label{rnminn}
\bra{n\bfk}\bfr_e\ket{m\bfk'} &=
(1-\gd_{nm})\gd(\bfk-\bfk')\bfgxi_{nm}(\bfk) 
,
\end{align}
such that $\bfr_{e,nm}=0$ for $n=m${\color{\chon} .
>From Eq.~\eqref{mv}} with $H_0\to
H^\Sigma_0$, we obtain
\begin{align}\label{pmnrmn}
\bfr_{e,nm}(\bfk) =
\bfgxi_{nm}(\bfk)\equiv 
\bfr_{nm}(\bfk) 
&=
\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
\quad\quad n\notin D_m 
,
\end{align}  
where we defined
$\go^\gs_{nm}(\bfk)\equiv\go^\gs_n(\bfk)-\go^\gs_m(\bfk)$, and
$D_m$ are all the possible degenerate $m$-states. 
%We recognize that for each instance in which 
%$n\notin D_m$ or $n\ne m$, $\bfr_{nm}(\bfk)$ are interband matrix
%elements.
For the intraband part, {\color{\chon} $\bfr_i$ only appears in
commutators during the derivation of
the optical response. We use},\cite{aversaPRB95}
\begin{equation}\label{conmri3n}
\bra{n\bfk}[\bfr_i,\calo]\ket{m\bfk'}
=i\gd(\bfk-\bfk')(\calo_{nm})_{;\bfk}
,
\end{equation}  
where
\begin{equation}\label{gendevnn}
(\calo_{nm})_{;\bfk}=
\nabla_\bfk 
\calo_{n m}(\bfk) 
- 
i 
\calo_{nm}(\bfk) 
\left(
\bfgxi_{nn}(\bfk) 
-
\bfgxi_{mm}(\bfk) 
\right) 
,
\end{equation} 
is
the generalized derivative 
of the operator $\calo$. 
The vectors $\bfgxi_{nn}(\bfk)$ are defined in 
Ref.~\onlinecite{aversaPRB95} {\color{\chon} though} they do not need to be 
calculated explicitly in what follows.

Before {\color{\chon} continuing,} 
we derive a key result for the length gauge formulation. 
{\color{\chon} Again, using} $H^\Sigma_0$ in
Eq.~\eqref{mv} we obtain
\begin{align}\label{vop2}
\bfv^\gs&=
\bfv 
+
\bfv^\nl 
+\bfv^S
=
\bfv^\lda 
+\bfv^S 
,
\end{align}
where we have defined 
\begin{subequations}
\begin{align}\label{ve}
\bfv 
&=\frac{\bfp}{m_e},
\\\label{vnl}
\bfv^\nl 
&=
\frac{1}{i\hbar}[\bfr,V^\nl],
\\\label{vs}
\bfv^S
&=
\frac{1}{i\hbar}[\bfr, S],
\\\label{vlda}
\bfv^\lda 
&=
\bfv 
+\bfv^\nl 
,
\end{align}  
\end{subequations}
 with $[r^a,p^b]=i\hbar\gd_{ab}$, where $\gd_{ab}$ is the Kronecker delta.
Using Eq.~\eqref{hats} we obtain 
\begin{equation}\label{chon.2} 
\bfv^S_{nm}=i\Delta f_{mn}\bfr_{nm},
\end{equation}
with $f_{nm}\equiv f_n-f_m$,
where we see that $\bfv^S_{nn}=0$. From Eq.~\eqref{pmnrmn} and Eq.~\eqref{vop2} {\color{\chon} it} follows that
\begin{align}\label{chon.10}
\bfr_{nm}(\bfk) 
=
\frac{\bfv^\gs_{nm}(\bfk)}{i\go^\gs_{nm}(\bfk)}
=
\frac{\bfv^\lda_{nm}(\bfk)}{i\go^\lda_{nm}(\bfk)}
\quad\quad n\notin D_m 
. 
\end{align}
The matrix elements 
of $\bfr_{nm}(\bfk)$ are {\color{\chon} identical using either} 
the LDA or {\color{\chon} scissored} 
Hamiltonian, {\color{\chon} thus negating the} need to label them.
Of course, {\color{\chon} it} is more convenient to calculate them
through $\bfv^\lda_{nm}(\bfk)$ {\color{\chon} which 
includes} only the contribution of 
$\bfv^\nl_{nm}(\bfk)$. These {\color{\chon} can} be readily
calculated
for 
fully separable nonlocal pseudopotentials in the 
Kleinman-Bylander 
form,\cite{francesco,mottaCMS10,kleinmanPRL82,adolphPRB96}{\color{\chon}.}
{\color{\chon} The} advantage of using the electron density operator along with the length gauge formalism for 
{\color{\chon} calculating linear and nonlinear} optical responses, for the scissored Hamiltonian, 
resides in the ease with {\color{\chon} which the scissors operator
can be introduced into the calculation by simply using the unscissored LDA Hamiltonian, $H_0^{\mathrm{LDA}}$,
for the unperturbed system 
with $-e\mathbf{r}\cdot \mathbf{E}(t)$ as the interaction{\color{\chon}. We stress that within the length gauge, 
we need only replace $\omega^{\mathrm{LDA}}_{n}$ with 
$\omega_{n}^{\Sigma}$ at the end of the derivation}
to obtain the scissored results for any 
susceptibility expression, whether linear or nonlinear \cite{PRBNastos}.

We have used the fact that for a cold semiconductor $\partial
f_{n}/\partial \mathbf{k}=0$ and thus the intraband contribution to the linear
term vanishes identically. 
%In Appendix \ref{gender} we show how the generalized derivative of 
%$(B^a_{nm})_{;k^b}$ can be obtained.
{\color{\chon} Note that the indices in Eq.~\eqref{j.2} are all} different from each
other. {\color{\chon} This is due to the $f_{nm}$ factor in Eq.~\eqref{j.1}, 
and therefore $B^a_{nn}=0$}. The dependence {\color{\chon} on} $\bfk$ 
of all quantities is implicitly understood from 
{\color{\chon} this point forward.}

### 2015-01-08 reply/referees.tex
for Valerie and Nicolas

--- 2015-01-08 referees.tex
@@
-From prb@aps.org Wed Dec 17 09:19:00 2014
+From prb@aps.org Mon Dec 22 13:48:36 2014
 Date: Tue, 16 Dec 2014 21:23:48 -0500

### 2015-01-08 sent-1/manuscript.tex
for Valerie and Nicolas

--- 2014-12-04 sent/manuscript.tex
@@
 $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega ;\omega ,\omega )
-=\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega ;\omega,\omega )$.
+=\chi^{\mathrm{a}\mathrm{c}\mathrm{b}}(-2\omega ;\omega,\omega )$.
 

### 2015-01-08 sent-2/manuscript.tex
for Valerie and Nicolas

--- 2014-12-24 corrections.tex
@@
 $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$    
-is calculated, including for the first time in one unique formulation
+is calculated, including in one unique formulation
 (i) the scissors correction, needed to have the correct value of the
@@
 $\chi^{xxx}(-2\omega;\omega,\omega)$ for the clean Si(001)$2\times 1$
-reconstructed surface and find that our scheme to extract the surface
-through the cut function
-readily works and is exact. The effects of the scissors correction and
+reconstructed surface.
+The effects of the scissors correction and
 of the nonlocal part of the pseudopotentials are discussed 
@@
 the bulk SHG signal of centrosymmetric materials is identically zero.
-Radiation can occur only at the surface where the inversion symmetry is 
+However, the SHG process  
+can occur only   
+at the surface where the inversion symmetry is   
 broken.   
@@
 This approximation has already been used in linear optical calculations for 
-surfaces, see \cite{kippPRL96}, thus improving the agreement with 
+surfaces,\cite{kippPRL96} thus improving the agreement with 
 experimental results.}
@@
 is required since one usually uses a slab approach when treating semi-infinite 
-surface systems.\cite{reiningPRB94} If the slab is centrosymmetric the susceptibility tensor will 
+surface systems.\cite{reiningPRB94} If the slab is centrosymmetric the 
+susceptibility tensor will 
 be identically zero. The cut function is such that it separates the nonlinear 
@@
 implemented. 
+{\color{\chon}In the works that include this operator,  the velocity gauge was
+  used to derive the expressions for $\chi^{\mathrm{abc}}(-2\omega;\omega,\omega)$.
+Nevertheless, a
+  term in the time-dependent perturbation scheme
+necessary to satisfy the gauge invariance 
+of $\chi^{\mathrm{abc}}(-2\omega;\omega,\omega)$  was omitted. 
+This was demonstrated
+in
+  Ref.~\onlinecite{cabellosPRB09} where a comparison between the
+  velocity and the length gauge  
+was carried out.
+}
 
@@
 
-{\color{\chon}According to ~\onlinecite{ismailPRL01}, we first start with the interaction 
+{\color{\chon}According to Ref. ~\onlinecite{ismailPRL01}, we first
+  start with the interaction  
 Hamiltonian expressed in the velocity gauge, containing the nonlocal parts
-$V^{nl}$ and $S$. Within the dipole approximation and using a gauge 
+$V^{nl}(\mathbf{r},\mathbf{p})$ and 
+$S (\mathbf{r},\mathbf{p})$. Within the dipole approximation and using a gauge 
 transformation, it can be transformed into an effective Hamiltonian 
-\cite{Veniard, unpublished}
+\cite{valerie}
 \begin{equation}
@@
 The treatment of the position operator $\mathbf{r}$ for extended Bloch states 
-is problematic and has been discussed in Refs.\cite{adamsJCP53,blountSSP62}
+is problematic and has been discussed in 
+Refs. \onlinecite{adamsJCP53,blountSSP62}.
 %%%% Valerie and Nicolas remove these lines
@@
 matrix elements of Eq.~\eqref{rhop} with the $H_{I}(t)$ of 
-Eq.~\eqref{rde}, we obtain 
+Eq.~\eqref{rde}, and we obtain 
 $(\tilde{\rho}^{(1)}(t))_{nm}=B_{nm}^{b}E^{b}(\omega)e^{i(\omega^\Sigma_{nm}-\tilde\omega)t}$,
@@
 \end{equation}}
+The position operator
 $\mathbf{r}$ is split into the {\it intraband} 
@@
 to obtain the scissored results for any 
-{\color{\chon}susceptibility expression, whether linear or nonlinear.\cite{nastosPRB05}
-
-We have used the fact that for a cold semiconductor} $\partial
+susceptibility expression, whether linear or nonlinear.\cite{nastosPRB05}
+
+We have used the fact that for a cold semiconductor $\partial
 f_{n}/\partial \mathbf{k}=0$ and thus the intraband contribution to the linear
@@
  and \onlinecite{mejiaRMF04}. 
-In this article, for the first time, we formally introduce the cut function ${\mathbf{\cal C}}(z)$ 
+In this article we formally introduce the cut function ${\mathbf{\cal C}}(z)$ 
 for the second-harmonic optical response of semiconductor surfaces,
@@
 valence bands, and a set of 244 $\mathbf{k}$-points.
-The $\mathbf{k}$-points are used for the linear analytic 
+The $\mathbf{k}$-points are used for 
+the linear analytic  
 tetrahedron method for evaluating the  
-Brillouin zone integrals 
+{\color{\chon}3D} Brillouin Zone (BZ) integrals  
 where special care was taken to examine the double resonances
 of Eq.~\eqref{chis}. \cite{nastosPRB05}
+{\color{\chon}Since we are dealing with a surface
+only one $\mathbf{k}$-point
+along $z$ is required for the 3D BZ integrals, and in this case
+the electronic bands 
+show no dispersion along $k_z$. 
+}
 All spectra were calculated with a Gaussian smearing of 0.15 eV.
@@
 {\color{\chon}This is justified by the fact that the surface states of the clean 
-Si(001) surface investigated are found, from GW calculations 
-\cite{rohlfingPRB95}, to shift rigidly and to keep their dispersion with 
-respect to LDA.}
+Si(001) surface are 
+rigidly shifted and maintain their
+dispersion relation with 
+respect to LDA according to
+ the GW calculations of Ref.
+\onlinecite{rohlfingPRB95}. 
+}
 We see that for all four instances the 

### 2015-01-08 cpfigures.sh

@@ -1,7 +1,7 @@
 pfix=pdf
-cp images/slab.$pfix sent/fig1.$pfix 
-cp images/si2x1-crop.$pfix sent/fig2.$pfix 
-cp plots/fig1.$pfix sent/fig3.$pfix 
-cp plots/fig2.$pfix sent/fig4.$pfix 
-cp plots/fig3.$pfix sent/fig5.$pfix 
-cp plots/fig4.$pfix sent/fig6.$pfix 
+cp images/slab.$pfix sent-2/fig1.$pfix 
+cp images/si2x1-crop.$pfix sent-2/fig2.$pfix 
+cp plots/fig1.$pfix sent-2/fig3.$pfix 
+cp plots/fig2.$pfix sent-2/fig4.$pfix 
+cp plots/fig3.$pfix sent-2/fig5.$pfix 
+cp plots/fig4.$pfix sent-2/fig6.$pfix 

### 2015-01-20 sent-2/manuscript.tex
sent back 1st round

@@
 \usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
-\def\chon{red}
-%\usepackage{showlabels}
 
@@
 These require ultrahigh vacuum (UHV) environments and provide no access to 
-buried interfaces. However, SSHG is compatibile with non-UHV conditions and 
+buried interfaces. However, SSHG is compatible with non-UHV conditions and 
 has access to interfaces buried beneath transparent overlayers. Even when 
@@
 \cite{levinePRL89,levinePRL91,delsolePRB93}  
-{\color{\chon}This allows us to ``open'' the DFT-LDA gap to 
+This allows us to ``open'' the DFT-LDA gap to 
 its correct experimental or GW value for most bulk semiconductors.
-%%%% Valerie and Nicolas remove these lines
-% even when experiments are not available. 
-%%%%
 This approximation has already been used in linear optical calculations for 
 surfaces,\cite{kippPRL96} thus improving the agreement with 
-experimental results.}
+experimental results.
 In this context, to correct for the underestimation of the energy band gap of 
 semiconductors Nastos et al.\cite{nastosPRB05} used the ``length gauge'' or 
-``$\mathbf{r}\cdot\mathbf{E}$ gauge'' to show how to correctly include the 
+``$\mathbf{r}\cdot\mathbf{E}$'' gauge to show how to correctly include the 
 many-body corrections through the scissors operator in the SH susceptibility.
@@
 implemented. 
-{\color{\chon}In the works that include this operator,  the velocity gauge was
+In the works that include this operator,  the velocity gauge was
 used to derive the expressions for $\chi^{\mathrm{abc}}(-2\omega;\omega,\omega)$.
@@
 was carried out.
-}
 
@@
 Density Approximation (LDA), so we label the hamiltonian with the corresponding  
-LDA superscript. Any other approximation can be used (like GGA) and our 
+LDA superscript. Any other approximation can be used (like the
+generalized gradient approximation) and our 
 derivation remains the same. Then,
@@
 \end{equation*}
-with $m_e$ the mass of the electron, $\mathbf{p}$ its canonical momentum, and 
+with $m_e$ the mass of the electron, 
+$\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ 
+its canonical momentum, and 
 $V$
@@
 
-{\color{\chon}According to Ref. ~\onlinecite{ismailPRL01}, we first
+According to Ref. ~\onlinecite{ismailPRL01}, we first
   start with the interaction  
@@
 Refs. \onlinecite{adamsJCP53,blountSSP62}.
-%%%% Valerie and Nicolas remove these lines
-% In particular, Ref.~\onlinecite{aversaPRB95} develops the treatment of 
-% $\mathbf{r}$ for the calculation of the nonlinear optical response.
-%%%%
 Following Ref.~\onlinecite{sipePRB00} we take the
@@
 \label{j.2}
-\end{equation}}
+\end{equation}
 The position operator
@@
 \end{align}  
-with $\mathbf{p}=-i\hbar\boldsymbol{\nabla}$ the canonical momentum operator, 
 and using $[r^a,p^b]=i\hbar\delta_{ab}$, where $\delta_{ab}$ is the 
@@
 tetrahedron method for evaluating the  
-{\color{\chon}3D} Brillouin Zone (BZ) integrals  
+3D Brillouin Zone (BZ) integrals  
 where special care was taken to examine the double resonances
 of Eq.~\eqref{chis}. \cite{nastosPRB05}
-{\color{\chon}Since we are dealing with a surface
-only one $\mathbf{k}$-point
-along $z$ is required for the 3D BZ integrals, and in this case
-the electronic bands 
-show no dispersion along $k_z$. 
-}
+Note that the Brillouin zone for the slab geometry collapses to a               
+2D-zone, with only one $\mathbf{k}$-point along the $z$-axis.
 All spectra were calculated with a Gaussian smearing of 0.15 eV.
@@
 Refs.~\onlinecite{rohlfingPRB95,garciaCPC01}. 
-{\color{\chon}This is justified by the fact that the surface states of the clean 
+This is justified by the fact that the surface states of the clean 
 Si(001) surface are 
@@
 \onlinecite{rohlfingPRB95}. 
-}
 We see that for all four instances the 
@@
 we note that the scissors correction 
-shifts the spectra from it LDA value to higher energies as expected.
+shifts the spectra from its LDA value to higher energies as expected.
  However, contrary 
@@
 $1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
-for the non-rigid shift. {\color{\chon}The reduction of the spectral strength is in 
+for the non-rigid shift. 
+The reduction of the spectral strength is in 
 agreement with previous calculations for bulk systems.
-\cite{nastosPRB05, luppiPRB10, leitsmannPRB05}}
+\cite{nastosPRB05, luppiPRB10, leitsmannPRB05}
 When we compare 
